### 2017-07-07 writeup/conStruct.tex
ms draft

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{float}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[authoryear]{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{bigints}
\usepackage{array}
\usepackage{tikz}
\usepackage{longtable}
\usepackage{geometry}

\bibliographystyle{plainnat}

\setlength{\parskip}{\baselineskip}
\setlength{\parindent}{0pt}

\newcommand{\e}[1]{{\mathbb E}\left[ #1 \right]}
\newcommand{\given}{\mid}
\newcommand{\secref}[1]{``\nameref{#1}''}

\newcommand{\gb}[1]{{\it\color{magenta}{(#1)}}}
\newcommand{\plr}[1]{{\it\color{purple}{(#1)}}}
\newcommand{\gc}[1]{{\it\color{blue}{(#1)}}}

\geometry{a4paper}

\title{Inferring Continuous and Discrete Population Genetic Structure Across Space}
\date{\vspace{-5ex}}
\author{
Gideon S. Bradburd$^{1,a}$, 
Peter L. Ralph$^{2,b}$, 
Graham M. Coop$^{3,c}$}

\begin{document}

\maketitle

\textsuperscript{1}
Department of Integrative Biology, 
Ecology, Evolutionary Biology, and Behavior Graduate Group,
Michigan State University, MI 48824

\textsuperscript{2}
Department of Molecular and Computational Biology, 
University of Southern California, Los Angeles, CA 90089

\textsuperscript{3}
Center for Population Biology,
Department of Evolution and Ecology, 
University of California, Davis, CA 95616

\textsuperscript{a}bradburd@msu.edu; 
\textsuperscript{b}plr@uoregon.edu;
\textsuperscript{c}gmcoop@ucdavis.edu\\\\\

\newpage

\begin{abstract}
One of the classic problems in population genetics is the characterization 
of discrete population structure when the genotyped samples also show 
continuous patterns of genetic differentiation.
Especially when sampling is discontinuous, 
clustering or assignment methods may incorrectly ascribe differentiation 
due to continuous processes (e.g., isolation by distance) 
to discrete processes, such as geographic, ecological, or reproductive barriers 
between populations.
This is partly a result of the difficulty of sampling uniformly and continuously 
across the range of a population or species, 
but more, it reflects a shortcoming of current methods for inferring and 
visualizing population structure from genetic data in the face of data 
that are characterized by both continuous and discrete population structure.
Here, we present a novel statistical framework for the simultaneous inference 
of continuous and discrete patterns of population structure.
The method estimates ancestry proportions for each 
sample from a set of discrete population clusters, 
and, within each cluster, estimates a rate at which relatedness decays with distance.
This model explicitly addresses the ``clines vs. clusters" problem in 
modeling population genetic variation by jointly accommodating both 
continuous and discrete patterns of differentiation. 
We demonstrate the utility of this approach using both simulations and an empirical application.
\end{abstract}

%evolutionary history of the genotyped samples has also been shaped by geographically continuous processes, 
%such as migration.

\newpage
\section*{Introduction}
The fundamental goal of evolutionary genetics is to link patterns of genetic variation 
to the processes that have generated and maintained them.
Often, the first step in the analysis of genetic data is to define evolutionary units of analysis
by delineating discrete populations from which individuals or alleles are sampled \ref{Wright 1949}.
These units are useful for defining a scope of study in research on selection or local adaptation \ref{},
or on identifying either locally adapted alleles (human skin color, ralph \& coop, gartersnakes) or alleles involved in disease (cystic fibrosis).
Delineating populations can also be useful for informing conservation priorities \ref{}.

There have been many methods proposed to characterize populations,
including generating population phylogenies \ref{cavalli-sforza, treemix},
$k$-means clustering analyses performed on dimensionality-reduction approaches \ref{}, 
such as principal components analysis \ref{Cavalli-sforza,Novembre and Stephens, Patterson2006},
and model-based clustering approaches \ref{Pritchard2000, all the other structure papers, admixture, structurama, BAPS, }, among others.
These methods each perform best under different specific demographic histories of samples being analyzed,
but all can give misleading results when applied to data that show a continuous pattern of isolation by distance \ref{Wright 1938, 1940, 1943}.

Isolation by distance (IBD) can

Patterns of population genetic differentiation are characterized by both discrete and 
continuous structure.  Identification of genetically discrete populations is important 
not just for understanding the distribution of genetic variation on a landscape, but also 
for understanding the demographic history of a species, or, from a conservation perspective, 
for characterizing discrete management units.  However, it is frequently difficult to disentangle 
continuous and discrete patterns of genetic differentiation.  
  By introducing a statistical framework that simultaneously 
models discrete and continuous patterns of population structure, we can do a more accurate 
job describing and quantifying both.

\section*{Methods}
\paragraph{Data}
The statistical framework of our approach is conceptually similar to \cite{Wasser2004} and \cite{spacemix}.
The genetic data modeled consist of allele frequencies $F$ at $L$ unlinked, bi-allelic single nucleotide polymorphisms (SNPs) genotyped across $N$ samples.
The sample frequency at locus $\ell$ in sample $n$, $f_{n,\ell}$, is calculated by first arbitrarily choosing an allele segregating 
at locus $\ell$ to count, 
then dividing the total number of observations of that counted allele by the total number of chromosomes genotyped at that locus
in sample $n$.
We then calculate the sample covariance in allele frequencies, $\widehat{\Omega}$, as

\begin{equation}
\widehat{\Omega} = \frac{1}{L}  FF^T
\label{sample_covariance}
\end{equation}

\paragraph{Continuous and discrete differentiation}
We wish to describe the observed patterns of genetic variation as a combination of 
discrete population clusters,
within which genetic variation is continuously distributed, 
and between which sampled individuals or populations can be admixed.
Each of these discrete clusters is modeled as a spatial process
in which migration between neighboring demes acts 
to homogenize allele frequency changes that arise locally due to drift;
this interplay between drift and migration 
results in a continuous pattern of isolation by distance within each cluster.
Multiple population cluster processes can co-occur in space, 
leading to discrete jumps in allele frequencies across many loci over small geographic distances,
and the genotyped samples can be admixed between different clusters.

The continuous decay of allele frequency covariance with geographic distance 
within a cluster is described using a simple and flexible powered exponential function;
the within-cluster covariance between samples $i$ and $j$ is given by:

\begin{equation}
G^{(k)}_{i,j} = \alpha^{(k)}_0 \left( \text{exp} \left( -(\alpha^{(k)}_D D_{i,j}) ^ {\alpha^{(k)}_2}	\right) \right) + \mu^{(k)}
\label{within_cluster_covariance}
\end{equation}

where $G_{i,j}$ is the covariance function between samples $i$ and $j$ within a cluster,
and the superscript $(k)$ denotes that this function is specific to the $k$th cluster.
The quantity $D_{i,j}$ is the observed geographic distance between samples $i$ and $j$,
the $\alpha$ parameters control the shape of the decay of covariance with distance,
and $\mu$ is a parameter that describes the amount of shared drift within a cluster.
The shared drift parameter $\mu^{(k)}$ can be interpreted as the branch length connecting 
the $k$th population to the population ancestral to all modeled clusters.

Genotyped samples can be admixed between these different clusters. 
The admixture proportion of the $i$th sample in the $k$th cluster, $w^{(k)}_i$,
gives the probability that an allele in sample $i$ was derived from cluster $k$;
each sample's admixture proportions across all clusters must sum to one.

We can then describe the covariance between samples $i$ and $j$ across all clusters, $\Omega_{i,j}$,
by summing their within-cluster spatial covariances ($G_{i,j}^{(k)}$ in cluster $k$) across all $K$ clusters
and weighting by the product of those samples' admixture proportions in each cluster.

\begin{equation}
\Omega_{i,j} = \gamma + \sum\limits_K w^{(k)}_i w^{(k)}_j G^{(k)}_{i,j} + \delta_{i,j}\eta_i
\label{cross_cluster_covariance}
\end{equation}

In addition to the admixture-weighted sum of the within-cluster spatial covariances,
this function contains two terms, $\gamma$ and $\delta_{i,j}\eta_i$.
The first, $\gamma$, describes the global covariance between all samples, 
and arises because all samples share an ancestral mean allele frequency at each locus.
In the second term, $\delta_{i,j}$ is an indicator variable that takes a value of 1 when $i$ equals $j$ and 0 otherwise. 
and $\eta_i$ gives the variance specific to sample $i$, 
This term on the diagonal of the parametric covariance matrix 
is meant to capture processes shaping variance within the sampled deme, 
such as inbreeding.

\paragraph{Likelihood and inference}
If we assume that the  allele frequencies $\widehat{F}$ are independent and multivariate normally distributed, 
their sample covariance $\widehat{\Omega}$ 
will be Wishart distributed with degrees of freedom equal to $L$, 
the number of loci genotyped.
Linkage disequilibrium (LD) between loci will decrease the effective number of degrees of freedom, 
which we discuss further below.
The likelihood of the sample standardized allele frequency covariance is therefore given by
\begin{equation}
P(\widehat{\Omega} \; | \; \Omega) = \mathcal{W} \left( L\widehat{\Omega} \; | \; \Omega,L\right)
\end{equation}

We estimate the values of the parameters of the model using a Bayesian approach.
Acknowledging the dependence of the parametric covariance matrix $\Omega$ on its constituent parameters
$w,\alpha,\mu,\eta,\gamma$ and on observed quantiy $D$ with the notation $\Omega(w,\alpha,\mu,\eta,\gamma,D)$,
we denote the posterior probability of the parameters as:

\begin{equation}
P\left( w,\alpha,\mu,\eta,\gamma \;	| \; \widehat{\Omega}, L \right) \propto
P\left(\widehat{\Omega} \; | \; \Omega(w,\alpha,\mu,\eta,\gamma,D) \right)
P(w)P(\alpha)P(\mu)P(\eta)P(\gamma)
\end{equation}

The priors, $P(w),P(\alpha),P(\mu),P(\eta),P(\gamma)$, are detailed in the Appendix,
and the constant of proportionality is the normalization constant.
We use a Hamiltonian Monte Carlo sampling algorithm implemented in the statistical language STAN
\citep{stan, NUTS, stan_lib, rstan} to estimate the posterior distribution on the parameters.
We also present an R package \citep{R} called \texttt{conStruct} that functions as a wrapper 
around this inference machinery.

\section*{Results}

\subsection*{Simulations}

\subsection*{Empirical Applications}

\section*{Discussion}

\section*{Acknowledgements}

This work was supported in part by 
the National Science Foundation under award number NSF \#1262645 (DBI) to PR and GC, 
the National Institute of General Medical Sciences of the National Institutes of Health under award numbers NIH RO1GM83098 and RO1GM107374 to GC,
and the National Science Foundation under award numbers NSF \# 1148897 and \# 1402725 to GB.

\newpage
\section*{Appendix}
\renewcommand{\theequation}{A\arabic{equation}}
\setcounter{equation}{0}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{table}{0}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

\section{Model rationale}
\subsection{Drift, admixture, and allele frequencies}
\paragraph{Drift} 
We imagine that the allele frequencies at each locus in each sample are the sum of 3 components: 
the ancestral allele frequency $\epsilon$ shared by all samples,
the deviation from that ancestral mean in the $k$th population, $\Delta^{(k)}$,
which is shared by all samples with 100\% ancestry in that population,
the deviation specific to the $i$th sample, $\Delta^{(i)}$,
which captures drift not shared by all samples at the population level
(i.e., subpopulation-specific drift due to, e.g., inbreeding).

If all samples drew all of their ancestry from any one of the $K$ clusters,
the  allele frequency in the $i$th sample at the $\ell$th locus would be given by:
\begin{equation}
F_{i,\ell} = \epsilon_{\ell} + \Delta^{(k)}_{\ell} + \Delta^{(i)}_{\ell}
\label{drift_terms_no_admix}
\end{equation}

\paragraph{Admixture} 
The model above describes the simple case 
in which samples draw 100\% of their ancestry from only a single cluster each. 
To accommodate admixture between clusters, 
we can model allele frequencies within samples as linear combinations of the population-specific 
drift terms across the clusters from which they draw ancestry.
To do so, we introduce a term $w^{(k)}_{i}$ that describes the admixture proportion of sample $i$ in cluster $k$,
which can be interpreted as the probability that an allele sampled in the $i$th sample came from the $k$th cluster.
The allele frequency in the $i$th sample at the $\ell$th locus can therefore be written as:
%
\begin{equation}
F_{i,\ell} = \epsilon_{\ell} + \sum\limits_{K} \left( w^{(k)}_{i} \Delta^{(k)}_{\ell} \right) + \Delta^{(i)}_{\ell}	
\label{drift_terms_admix}
\end{equation}
%
where
%
\begin{equation}
\sum\limits_{K}w^{(k)}_{i} = 1
\end{equation}

\subsection{The Normal approximation to drift}
Genetic drift is serial binomial sampling, 
but over short timescales, and for intermediate ancestral allele frequencies, 
we can reasonably model the allele frequencies at each locus across populations as multivariate normal.
This will be a good approximation if the amount of elapsed drift in each sample is relatively small,
as would be expected if time since divergence from a common ancestral population is short,
or effective population sizes are large.

Beginning with the simplest case of two clusters ($K=2$),
with all samples drawing 100\% of their ancestry from either one or the other cluster, 
we can model allele frequencies at locus $\ell$ as:
%
\begin{equation}
F_{\ell} \sim MVN\left(\mu = \epsilon_{\ell} + \iota^{(k)}\Delta_{\ell}^{(k)}, \Sigma = \Omega	\right)
\label{MVN_mean}
\end{equation}
%
where $\iota^{(k)}$ is an indicator variable that equals 1 for samples with 100\% membership in cluster $k$
and 0 otherwise.
In Eqn \eqref{MVN_mean}, all entries of the covariance $\Omega$ have value 0 except the diagonals,
which describe sample-specific drift, and have the following value:
%
\begin{equation}
\Omega_{i,i} = \text{Var}\left( \Delta^{(i)} \right)
\label{MVN_mean_diag}
\end{equation}

%

\subsection{Describing relatedness as a covariance}
However, our data can consist of many loci, 
so it may be computationally impractical to model 
the ancestral, cluster, and sample components of each sampled allele frequency.
Instead, we can describe the covariance induced between samples due to the sharing of the components.
All samples share the same ancestral mean at all loci, 
which leads to a global covariance equal to the variance of the ancestral frequencies across loci.
Likewise, samples that are in the same cluster will have a covariance equal to the variance 
of the cluster-specific deviates from the ancestral frequencies across loci.

In Eqn \ref{MVN_mean} above, we described the relatedness between samples
due to the shared ancestral allele frequency and shared population-level drift
via the mean of the multivariate normal.
We can describe this relatedness as a covariance as follows:

\begin{equation}
	\Omega_{i,j} = 
	\begin{cases}
	      \text{Var}(\epsilon) +  \text{Var}\left( \Delta^{(k)} \right) + \text{Var}\left( \Delta^{(i)} \right), & \text{if}\ i=j\ \\
      	      \text{Var}(\epsilon) +  \text{Var}\left( \Delta^{(k)} \right), & \text{if}\ \iota^{(k)}_i = \iota^{(k)}_j\ \text{and}\ i \neq j \\
	      \text{Var}(\epsilon), & \text{otherwise}
	\end{cases}
\label{MVN_covar}
\end{equation}

Generalizing this result to any number of clusters and incorporating admixture as in Eqn \ref{drift_terms_admix}, 
we can write the form of the covariance of samples that are admixed between $K$ clusters as:

\begin{equation}
\Omega_{i,j} = \text{Var}(\epsilon) + \sum\limits_K \left(	w^{(k)}_iw^{(k)}_j \text{Var}(\Delta^{(k)}) 	\right) + 
\delta_{i=j} \text{Var}(\Delta^{(i)})
\label{admixed_discrete_covariance}
\end{equation}

where $\delta_{i=j}$ is an indicator variable that equals 1 when $i$ is equal to $j$ and 0 otherwise,
as in Eqn \ref{cross_cluster_covariance}.

\subsection{Incorporating spatial differentiation}
Equation \ref{admixed_discrete_covariance} describes a model in which 
samples can be continuously admixed between a set of $K$ discrete clusters.
In this model, any pair of samples with 100\% ancestry in a cluster 
have exactly the same covariance with each other (namely $\text{Var}(\epsilon) + \text{Var}(\Delta^{(k)})$).
However, we also wish to model continuous decay of covariance with spatial separation between samples.
That is, we expect (and want our model to reflect) that samples within the same cluster 
will have higher covariance if they are sampled closer together than if they are sampled farther apart.
To describe this spatial pattern, we build in a spatial component to our covariance model.

Specifically, we write that the covariance between a pair of samples $i$ and $j$
that draw all their ancestry from a single cluster $k$ decays exponentially as a function 
of the distance between their sampling locations as follows:
\begin{equation}
G^{(k)}_{i,j} = \alpha^{(k)}_0 \times \left(\text{exp} \left(  -(\alpha^{(k)}_D D_{i,j})^{\alpha^{(k)}_2}\right) \right) + \mu^{(k)}	
\label{spatial_cov}
\end{equation}
where the $\alpha$ parameters control: 
the sill of the covariance ($\alpha_0$), 
the rate at which covariance decays with distance $D$ ($\alpha_D$),
and the shape of that decay ($\alpha_2$)
in the $k$th cluster.
The parameter $\mu^{(k)}$ describes the covariance shared by all samples in the $k$th cluster;
it and the spatial covariance function $G$ together
 describe the quantity $\text{Var}(\Delta^{(k)})$ from Eqn \eqref{admixed_discrete_covariance}.

Tying these ideas together, we can then construct a covariance that describes
\begin{enumerate}
\item continuous decay of covariance within a cluster
\item a discrete amount of covariance shared within a cluster
\item continuous admixture between clusters.
\end{enumerate}

\begin{equation}
\Omega_{i,j} = \text{Var}(\epsilon) + \sum\limits_K \left(	w^{(k)}_iw^{(k)}_j G^{(k)}_{i,j}(\theta) 	\right) +
\delta_{i=j} \text{Var}(\Delta_i)
\label{admixed_spatial_cov}
\end{equation}
%
where $G^{(k)}_{i,j}(\theta)$ denotes the dependence of the spatial covariance between samples $i$ and $j$ 
within a cluster on other quantities $\theta$: 
specifically, the parameters $\vec{\alpha}^{(k)}$ and $\mu^{(k)}$, which are specific to the $k$th cluster, 
as well as on the observed quantity $D_{i,j}$, the pairwise distance between samples $i$ and $j$.

\section{Likelihood}
If the allele frequency data are well approximated by a Gaussian, 
their sample covariance is a sufficient statistic,
so that calculating the likelihood of their sample covariance is the same as 
calculating the probability of the frequency data up to a constant. 
We can therefore model the covariance of the sample allele frequencies, $\widehat{\Omega}$, 
as a draw from a Wishart distribution with degrees of freedom equal to 
the number of loci $L$ across which the sample covariance is calculated:

\begin{align}
\widehat{\Omega} &= \frac{1}{L}FF^T \notag \\
\widehat{\Omega} &\sim \mathcal{W}\left( L\Omega, L	\right)
\label{wishart}
\end{align}

A benefit of directly modeling the sample allele frequency covariance is that, 
after the initial calculation of the sample covariance matrix,
the computation time of the likelihood is not a function of the number of loci,
so inference can theoretically be done using whole genome data.

\section{Models, parameters, and priors}
\paragraph{Spatial vs. nonspatial}
In this paper, we discuss two types of models, spatial and nonspatial.
The spatial model is parameterized as in Eqn \ref{admixed_spatial_cov},
and the nonspatial model is described in Eqn \ref{admixed_discrete_covariance}.
The nonspatial model therefore has $3K$ fewer parameters than the spatial model,
as there are three $\alpha$ parameters that describe the continuous differentiation effect of distance per cluster,
and they are not included in the nonspatial model.

\paragraph{Single cluster}
Each of these models can be run with a single cluster ($K=1$), 
which involves a slight re-parameterization, detailed below.

For the spatial model, the single-cluster parametric covariance is:
\begin{equation}
\Omega_{i,j} = \text{Var}(\epsilon) + 
\alpha^{(k)}_0 \times \left(\text{exp} \left(  -(\alpha^{(k)}_D D_{i,j})^{\alpha^{(k)}_2}\right) \right)	 +
\delta_{i=j} \text{Var}(\Delta_i)
\label{admixed_spatial_cov}
\end{equation}

For the nonspatial model, the single-cluster parametric covariance is:
\begin{equation}
\Omega_{i,j} = \text{Var}(\epsilon) + \delta_{i=j} \text{Var}(\Delta^{(i)})
\label{admixed_discrete_covariance}
\end{equation}

\paragraph{Priors}
We use a Bayesian approach to parameter inference,

A table of all parameters, their descriptions, and their priors is given in Table \ref{tab:param_prior_tab} below.

\begin{centering}
\begin{table}
\begin{tabular}{| >{\centering\arraybackslash}m{2.1cm} | m{5.2cm} | >{\centering\arraybackslash}m{5.1cm} |}
	\hline
	\textbf{Parameter} & \centering{\textbf{Description}} & \textbf{Prior}\\ \hline
	$\boldsymbol{\gamma}$ & 
		global covariance due to shared ancestral frequency & 
		$\gamma \sim \mathcal{N}(\mu = \text{Var}(\bar{f}), \sigma = 0.5)$\\ \hline
	$\boldsymbol{\alpha^{(k)}_0}$ & 
		controls the sill of the covariance matrix in cluster $k$& 
		$\alpha^{(k)}_0 \sim \mathcal{N}(\mu = 0, \sigma = 1)$\\ \hline
	$\boldsymbol{\alpha^{(k)}_D}$ & 
		controls the rate of the decay of covariance with distance in cluster $k$& 
		$\alpha^{(k)}_D \sim \mathcal{N}(\mu = 0, \sigma = 1)$\\ \hline
	$\boldsymbol{\alpha^{(k)}_2}$ & 
		controls the shape of the decay of covariance with distance in cluster $k$ & 
		$\alpha^{(k)}_2 \sim U(0,2)$\\ \hline
	$\boldsymbol{\eta_i}$ & 
		the nugget in population $i$ (population specific drift parameter)  & 
		$\eta_i \sim \mathcal{N}(\mu = 0, \sigma = 1)$\\ \hline
	$\boldsymbol{\mu^{(k)}}$ & 
		cluster-specific shared drift in cluster $k$ &
		 $\mu^{(k)} \sim \mathcal{N}(\mu = 0, \sigma = 1)$\\ \hline
	$\boldsymbol{w_i}$ &
		admixture proportions sample $i$ draws across $K$ clusters &
		$w_i \sim \text{Dir}(\alpha_{1} ... \alpha_{K}=0.1)$  \\ \hline
	\hline
\end{tabular}
\caption{
List of parameters used in the \texttt{conStruct} model, along with their descriptions and priors.
The mean of the Normal prior on $\gamma$, $\text{Var}(\bar{f})$, is the variance of the sample mean allele frequencies across loci.
}\label{tab:param_prior_tab}
\end{table}
\end{centering}

\newpage
\section{Cross validation}
To perform model comparison, we employ a Monte Carlo cross-validation approach,
also known as repeated random sub-sampling validation.
Briefly, we follow the following procedure:
\begin{enumerate}
\item For each of $X$ replicates:
	\begin{enumerate}
		\item partition the allele frequency data into a 90\% ``training'' partition ($F^x_1$) and a 10 \%``testing" partition ($F^x_2$) \label{partition}
		\item run our inference procedure on the training partition to estimate model parameters $\theta_{mk}$ for: \label{inference}
			\begin{enumerate}
				\item $m$: the spatial and the nonspatial model
				\item $k$: the number of clusters 1 through $K$
			\end{enumerate}	
	\item calculate the mean log likelihood of the testing data partition, 
	over the posterior distribution of training-estimated parameters for each model 
	($\bar{\mathcal{L}}(F^x_2 \mid \theta_{mk})$, henceforth $\bar{\mathcal{L}}_{xmk}$) \label{lnL}
	\item generate standardized mean log likelihoods, $\mathcal{Z}_{xmk}$, across models: \label{standardize}
		\begin{enumerate}
			\item identify the highest mean log likelihood, $\bar{\mathcal{L}}^\text{max}_{xmk}$
			\item subtract $\bar{\mathcal{L}}^\text{max}_{xmk}$ from $\bar{\mathcal{L}}_{xmk}$ across all models,
				such that the standardized log likelihood, $\mathcal{Z}_{xmk}$, of the best model is 0,
				and less than 0 for all inferior models. 
		\end{enumerate}
	\end{enumerate}
\item For each model (i.e., each combination of $m$ and $k$) calculate 
	the mean standardized log likelihood of the testing data partition across $X$ replicates, 
	as well its standard error and 95\% confidence interval:
	\begin{enumerate}
	\item mean
		\begin{equation}
			\bar{\mathcal{Z}}_{mk} = \frac{1}{X}\sum\limits_{x=1}^{X}\mathcal{Z}_{xmk}
		\end{equation}
	\item standard error
		\begin{equation}
			SE_{\bar{\mathcal{Z}}_{mk}} = \frac{1}{\sqrt{X}}\sum\limits_{x=1}^{X} \left[ {\mathcal{Z}_{xmk} - \bar{\mathcal{Z}}_{mk}} \right]
		\end{equation}
	\item 95\% confidence interval
		\begin{equation}
			95\% \text{CI} = \bar{\mathcal{Z}}_{mk} \pm 1.96 \times SE_{\bar{\mathcal{Z}}_{mk}}
		\end{equation}
	\end{enumerate}
\end{enumerate}

This cross-validation procedure generates a mean predictive accuracy for each model and each value of $K$, 
as well as a confidence interval around that mean,
which can then be used for model comparison or selection.

\newpage

\bibliography{reference.library.tex}

\newpage

\pagenumbering{gobble}

\end{document}

### 2017-07-07 writeup/conStruct.tex
draft of intro

@@
 Often, the first step in the analysis of genetic data is to define evolutionary units of analysis
-by delineating discrete populations from which individuals or alleles are sampled \ref{Wright 1949}.
-These units are useful for defining a scope of study in research on selection or local adaptation \ref{},
+by delineating discrete populations from which individuals or alleles are sampled \citep{wright1949genetical}.
+These units are useful for defining a scope of study in research on selection or local adaptation \citep{},
 or on identifying either locally adapted alleles (human skin color, ralph \& coop, gartersnakes) or alleles involved in disease (cystic fibrosis).
-Delineating populations can also be useful for informing conservation priorities \ref{}.
+Delineating populations can also be useful for informing conservation priorities \citep{}.
 
 There have been many methods proposed to characterize populations,
-including generating population phylogenies \ref{cavalli-sforza, treemix},
-$k$-means clustering analyses performed on dimensionality-reduction approaches \ref{}, 
-such as principal components analysis \ref{Cavalli-sforza,Novembre and Stephens, Patterson2006},
-and model-based clustering approaches \ref{Pritchard2000, all the other structure papers, admixture, structurama, BAPS, }, among others.
+including generating population phylogenies \citep{CavalliSforza1975, treemix},
+$k$-means clustering analyses performed on dimensionality-reduction approaches \citep{}, 
+such as principal components analysis \citep{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
+and model-based clustering approaches \citep[e.g.][]{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, FINESTRUCTURE, huelsenbeck2007inference, Corander2003}, among others.
+Here, we will focus on model-based clustering, the most widely used class of approaches.
 These methods each perform best under different specific demographic histories of samples being analyzed,
-but all can give misleading results when applied to data that show a continuous pattern of isolation by distance \ref{Wright 1938, 1940, 1943}.
+but all can give misleading results when applied to data that show a continuous pattern of isolation by distance \citep{Wright 1938, 1940, Wright1943}.
 
-Isolation by distance (IBD) can
+A pattern of isolation by distance (IBD) arises when migration in a population is spatially limited \citep{Slatkin1985},
+and has been modeled in both spatially continuous frameworks \citep{Wright1943,Malecot1948}, 
+in which an individual's parents are found within some radius of that individual's location,
+and discrete, ``stepping-stone'' frameworks \citep{Kimura1953},
+in which migration occurs only between neighboring demes.
+Given the generality of the circumstances that generate a pattern of isolation by distance, 
+it is unsurprising that IBD is ubiquitous in nature \citep{meirmans2012,Sexton_etal_2014}.
 
-Patterns of population genetic differentiation are characterized by both discrete and 
-continuous structure.  Identification of genetically discrete populations is important 
-not just for understanding the distribution of genetic variation on a landscape, but also 
-for understanding the demographic history of a species, or, from a conservation perspective, 
-for characterizing discrete management units.  However, it is frequently difficult to disentangle 
-continuous and discrete patterns of genetic differentiation.  
-  By introducing a statistical framework that simultaneously 
-models discrete and continuous patterns of population structure, we can do a more accurate 
-job describing and quantifying both.
+The ubiquity of isolation by distance presents a challenge for models of discrete population structure,
+as it is frequently difficult to determine whether observed patterns of genetic variation are 
+continuously distributed across a landscape, or instead are partitioned in discrete clusters.
+This problem can be compounded if sampling is done unevenly or discretely across a population or species' range,
+and has given rise to a debate in the population genetic literature about whether 
+individuals are best described as being drawn from continuous clines or discrete clusters 
+\citep[e.g.][]{SerrePaabo2004,rosenberg2005clines}.
+
+Existing model-based clustering approaches can only describe continuous patterns of variation using
+discrete clusters, 
+and so tend to erroneously describe continuously distributed variation with multiple clusters that 
+show spatially autocorrelated cluster membership \citep{Frantz2009,Meirmans2012}.
+In analyses of empirical datasets, which often are characterized by IBD,
+model-based clustering approaches will therefore tend to overestimate the number of discrete clusters present,
+with downstream implications for interpretation of results.
+For example, conservation resources might be spread too thinly over areas 
+that are spuriously assumed to be discrete management units.
+In studies of locally adaptive loci, or of the genetics of disease,
+discrete populations may be modeled in separately,
+reducing the global power to detect the genomic basis of complex phenotypes.
+
+What is desired, then, is a model-based clustering method that incorporates a spatial element.
+and specifically one that uses a model of isolation by distance as a biologically sensible null hypothesis 
+for the process structuring observed genetic variation.
+With an explicit spatial null, discrete population structure need only be invoked when genetic differentiation 
+in the data is above and beyond that expected given some estimated background rate at which divergence 
+increases with geographic separation.
+
+In this paper, we present a statistical method for simultaneously 
+modeling continuous and discrete patterns of population structure.
+We model genetic variation in genotyped individuals as 
+partitioned within or admixed across a user-specified number of discrete clusters,
+within each of which relatedness decays as a parametric function of the distance between samples.
+We also implement a cross-validation approach for comparing and selecting models across different numbers of clusters,
+and we demonstrate the utility of our approach using both simulated and empirical data.
+The implementation of this method, \texttt{conStruct}, is available as an R package at \href{https://github.com/gbradburd/conStruct}{https://github.com/gbradburd/conStruct}.
 
@@
 
-\bibliography{reference.library.tex}
+\bibliography{reference.library.tex.bib}
 

### 2017-07-08 writeup/conStruct.tex
updated Peter's affiliation

@@
 \textsuperscript{2}
-Department of Molecular and Computational Biology, 
-University of Southern California, Los Angeles, CA 90089
+Institute of Ecology and Evolution,
+Department of Mathematics,
+University of Oregon, Eugene, OR 97403
 
@@
 
+We thank Marjorie Weber, William Wetzel, Mariah Meek, Yaniv Brandvain, and Matthew Stephens for invaluable comments on the method and manuscript, as well as attendees at the 2017 SSE Meeting in Portland, OR, whose votes determined the name of the method.
 This work was supported in part by 

### 2017-07-10 sims/sim_code/exe.sim.in.ms.R
added sim code

@@ -1,0 +1,43 @@
+source("sim.in.ms.R")
+
+K <- 2
+if(K == 1){
+	ms.command.line <- generate.ms.command.line.values(diploid.population.size = 5e2,
+														locus.size = 1e6, 
+														per.bp.mu = 1e-8, 
+														migration.fraction = 5e-5,
+														deep.split = 1e6,
+														shallow.split = 1e6)
+} else {
+	ms.command.line <- generate.ms.command.line.values(diploid.population.size = 1e3,
+														locus.size = 1e6, 
+														per.bp.mu = 1e-8, 
+														migration.fraction = 1e-4,
+														deep.split = 1e4,
+														shallow.split = 1e3)
+}
+
+coords <- as.matrix(expand.grid(1:10,1:10))
+N <- nrow(coords)
+if(K > 1){
+	w <- sim.admix.props(N,K,coords)
+	admix.list <- make.admix.list(N,K,coords,w,0.0001)
+}
+
+sim.dataset <- 	generate.conStruct.dataset(n.loci = 5e3,
+										K = K,
+										coords = coords,
+										n.chromo = 20,
+										ms.params = ms.command.line,
+										admix.list = NULL,
+										subsample = which(coords[,1] > 2 & coords[,1] < 9 &
+														   coords[,2] > 2 & coords[,2] < 9))
+
+save(sim.dataset,file="sim.dataset.Robj")
+
+
+freq.data <- conStruct::process.freq.data(sim.dataset$freq.data$freqs)
+D <- fields::rdist(sim.dataset$coords)
+pdf(file="ibd.pdf")
+	plot(D,freq.data$obsCov)
+dev.off()

### 2017-07-10 sims/sim_code/sim.in.ms.R

@@ -1,0 +1,380 @@
+##################################################################
+##################################################################
+##	run ms on a lattice
+##################################################################
+##################################################################
+
+require(conStruct)
+
+within.cluster.m <- function(sample.inds,migration.index,pairwise.migration.matrix){
+	#recover()
+	migration.rate.vector <- c()
+	for(i in 1:nrow(migration.index)){
+		migration.rate.vector <- c(migration.rate.vector,
+									sprintf("-m %s %s %s",
+										sample.inds[migration.index[i,1]],
+										sample.inds[migration.index[i,2]],
+										pairwise.migration.matrix[migration.index[i,1],migration.index[i,2]]))
+	}
+	return(migration.rate.vector)
+}
+
+within.cluster.merge <- function(sample.inds,shallow.split){
+	merge.vector <- unlist(lapply(sample.inds[2]:sample.inds[length(sample.inds)],
+						function(i){
+							sprintf("-ej %s %s %s",
+									shallow.split,
+									i,
+									sample.inds[1])}))
+	return(merge.vector)
+}
+
+rewrite.admix.props <- function(w){
+	K <- length(w)
+	new.w <- w
+	for(k in 2:(K-1)){
+		adjustment <- (1-sum(w[1:(k-1)]))
+		if(adjustment == 0){
+			new.w[k] <- w[k]
+		} else {
+			new.w[k] <- w[k]/(1-sum(w[1:(k-1)]))
+		}
+	}
+	new.w[new.w > 1] <- 1
+	return(new.w[1:(K-1)])
+}
+
+write.admixture.event.calls <- function(admix.list,n.pops){
+	#recover()
+	N <- nrow(admix.list$w)
+	K <- ncol(admix.list$w)
+	admixture.call <- c()
+	source.ticker <- 1
+	n.time.ticker <- 1e-5/K
+	k.time.ticker <- 1e-5/(10*K)
+	for(n in 1:N){
+		if(K > 2){
+			new.w <- rewrite.admix.props(admix.list$w[n,])
+		} else if (K == 2){
+			new.w <- admix.list$w[n,1]
+		}
+		for(i in 1:length(new.w)){
+			admixture.call <- c(admixture.call,
+								sprintf("-es %s %s %s",
[…]

### 2017-07-10 code/conStruct/R/conStruct.R
reminder about STRUCTURE comparison

@@ -1020,7 +1020,7 @@ x.validation.rep <- function(rep.no,test.pct,K,freqs,geoDist,coords,prefix,n.ite
 											 freqs = train.data,
 											 geoDist = geoDist,
 											 coords = coords,
-											 prefix = paste0(prefix,"_sp_","rep",rep.no,"K=",k),
+											 prefix = paste0(prefix,"_sp_","rep",rep.no,"K",k),
 											 n.iter = n.iter,
 											 burnin = burnin)
 						})
@@ -1030,7 +1030,7 @@ x.validation.rep <- function(rep.no,test.pct,K,freqs,geoDist,coords,prefix,n.ite
 											 freqs = train.data,
 											 geoDist = geoDist,
 											 coords = coords,
-											 prefix = paste0(prefix,"_nsp_","rep",rep.no,"K=",k),
+											 prefix = paste0(prefix,"_nsp_","rep",rep.no,"K",k),
 											 n.iter = n.iter,
 											 burnin = burnin)
 						})

### 2017-07-17 writeup/conStruct.tex
reminder about STRUCTURE comparison
edit on x-validation eqn

@@
 \section{Cross validation}
-To perform model comparison, we employ a Monte Carlo cross-validation approach,
+To perform model comparison, we employ a Monte Carlo cross-validation approach \citep{picard1984},
 also known as repeated random sub-sampling validation.
@@
 		\begin{equation}
-			SE_{\bar{\mathcal{Z}}_{mk}} = \frac{1}{\sqrt{X}}\sum\limits_{x=1}^{X} \left[ {\mathcal{Z}_{xmk} - \bar{\mathcal{Z}}_{mk}} \right]
+			SE_{\bar{\mathcal{Z}}_{mk}} = \sqrt{\frac{\sum\limits_{x=1}^{X} \left[ {\mathcal{Z}_{xmk} - \bar{\mathcal{Z}}_{mk}} \right]}{X}}
 		\end{equation}
@@
 
+\section{Our Model vs. STRUCTURE}
+We should probably demonstrate that the nonspatial version of \texttt{conStruct} is basically the same as STRUCTURE.
+Maybe check out Engelhardt \& Stephens?
+
 \newpage

### 2017-07-21 writeup/conStruct.tex
added figs for sim K1

@@
 \subsection*{Simulations}
+We first test the method with scenarios generated using the coalescent simulator \texttt{ms} \citep{Hudson2002}.
+To generate discrete population structure, we simulate $K$ discrete clusters,
+each of which split from a common ancestor at some point in the past, $\tau_{\text{s}}$,
+then diverge for some time without migration between them.
+To generate continuous differentiation within each cluster,
+at time $\tau_{\text{e}}$ we have each of these discrete clusters instantaneously colonize a lattice of demes,
+for which we use a stepping stone model with symmetric migration 
+to nearest neighbors (eight neighbors, including diagonals).
+Finally, at time $\tau_{\text{a}}$ we generate a single dataset that captures both continuous and discrete patterns of population structure 
+by collapsing those $K$ discrete lattices into a single grid of demes sampled from the different clusters
+or admixed between them (see Fig \ref{sim_setup}).
+We simulated datasets under $K =$ 1 through 4, 
+sampling 10,000 unlinked loci from each of 20 haploid individuals from every deme
+(for details on all simulations, see Methods and Supplementary Materials).
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=4in,height=2.5in]{figs/sims/sim_setup.png}}
+		\caption{Schematic of how we simulate datasets with continuous and discrete differentiation, using $K=2$ as an example.  
+			    Going forward in time, the $K$ clusters split from a common ancestor at time $\tau_{\text{s}}$,
+			    then each expand to colonize a lattice of demes with nearest-neighbor symmetric migration at time $\tau_{\text{e}}$,
+			    then finally at time $\tau_{\text{a}}$ collapse into a single lattice consisting of demes 
+			    with ancestry entirely in one or the other of the clusters,
+			    or admixed between them.
+			    }\label{sim_setup}
+\end{figure}
+
+We ran \texttt{conStruct} analyses on each of these simulated datasets, 
+specifying both a spatial and nonspatial model for $K = $ 1 through 7, 
+then used cross-validation to compare the predictive performance of the models.
 
@@
 
-We thank Marjorie Weber, William Wetzel, Mariah Meek, Yaniv Brandvain, and Matthew Stephens for invaluable comments on the method and manuscript, as well as attendees at the 2017 SSE Meeting in Portland, OR, whose votes determined the name of the method.
+We thank Marjorie Weber, William Wetzel, Mariah Meek, Yaniv Brandvain, and Matthew Stephens 
+for invaluable comments on the method and manuscript, 
+as well as Quentin Cronk, who helped interpret the \textit{Populus} analyses, 
+and the attendees at the 2017 SSE Meeting in Portland, OR, whose votes determined the name of the method.
 This work was supported in part by 
@@
 \setcounter{table}{0}
-\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thefigure}{A\arabic{figure}}
 \setcounter{figure}{0}
@@
 \newpage
+\section*{Supplementary Materials}
+\renewcommand{\theequation}{S\arabic{equation}}
+\setcounter{equation}{0}
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{table}{0}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{figure}{0}
 
+\newpage
+\begin{figure}
+	\centering
+		\subcaptionbox{$K=2$\label{K2}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K2.pdf}}
+		\subcaptionbox{$K=3$\label{K3}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K3.pdf}}
+		\subcaptionbox{$K=4$\label{K4}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K4.pdf}}
+		\subcaptionbox{$K=5$\label{K5}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K5.pdf}}
+		\subcaptionbox{$K=6$\label{K6}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K6.pdf}}
+		\subcaptionbox{$K=7$\label{K7}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K7.pdf}}
+	\caption{
+	Map of admixture proportions estimated using a nonspatial model for $K=2$ through 7.
+	The data were simulated using a single cluster with nearest-neighbor symmetric migration between demes.
+    }\label{sfig:simK1_nsp_pies}
+\end{figure}
+
+\newpage
+\newpage
 \bibliography{reference.library.tex.bib}

### 2017-07-21 writeup/figs/sim_figs.R

@@ -1,0 +1,78 @@
+################################################################
+################################################################
+#	Sim Figures for conStruct paper
+################################################################
+################################################################
+
+source("~/Dropbox/conStruct/sims/cross_validation/summarize.xvals.R")
+
+#K1
+setwd("~/Dropbox/conStruct/sims/cross_validation/K_1/x_validation")
+n.reps <- 10
+K <- 7
+for(n in 1:n.reps){
+	load(sprintf("simK1_rep%s_test.lnl.Robj",n))
+	assign(paste0("tl",n),test.lnl)
+}
+x.vals <- lapply(1:n.reps,function(n){get(sprintf("tl%s",n))})
+x.vals.std <- lapply(x.vals,standardize.xvals)
+xval.CIs <- get.xval.CIs(x.vals.std,K)
+pdf(file="~/Dropbox/conStruct/writeup/figs/sims/simK1_std.xval.pdf",width=6,height=6,pointsize=14)
+	plot.xval.CIs(xval.CIs,K,simK=" (K = 1)")
+	legend(x="bottomright",pch=19,col=c("blue","green"),legend=c("spatial","nonspatial"))
+dev.off()
+
+pdf(file="~/Dropbox/conStruct/writeup/figs/sims/simK1_std.xval_sp.pdf",width=6,height=6,pointsize=14)
+	plot.xval.CIs(xval.CIs,K,simK=" (K = 1)",ylim=c(-10,0))
+	legend(x="topright",pch=c(19,NA),lty=c(NA,1),lwd=c(NA,2),col=c(1,adjustcolor(1,0.8)),legend=c("mean","95% CI"))
+dev.off()
+
+library(conStruct)
+for(k in 2:7){
+	load(sprintf("simK1__nsp_rep1K%s_data.block.Robj",k))
+	load(sprintf("simK1__nsp_rep1K%s_conStruct.results.Robj",k))
+	csr <- conStruct.results[[1]]
+	clst.match <- NULL
+	if(k < 4){
+		csr1.order <- NULL
+	}
+	clst.match$cols <- c(4,2)
+	if(k > 2){
+		load(sprintf("simK1__nsp_rep1K%s_conStruct.results.Robj",k-1))
+		tmp.csr <- conStruct.results[[1]]
[…]

### 2017-07-27 writeup/conStruct.tex
G's edits

@@
 \end{equation}
+\gc{SO IS THIS ACTUALLY the calc given the new diagonal?}
 
@@
 and between which sampled individuals or populations can be admixed.
-Each of these discrete clusters is modeled as a spatial process
-in which migration between neighboring demes acts 
-to homogenize allele frequency changes that arise locally due to drift;
-this interplay between drift and migration 
-results in a continuous pattern of isolation by distance within each cluster.
+Each of these discrete clusters is modeled as a spatial process where
+\gc{allele frequencies covary between geographically close locations,
+  but  %%changed as we dont model migration or drif
+this covariance is allowed to decay with geographic distance. This 
+pattern of spatial decay reflects} how migration between neighboring demes acts 
+to homogenize allele frequency changes that arise locally due to
+drift, \gc{but that migration is less effective at homogenizing
+  between geographic distance demes};
+%with this interplay between drift and migration 
+resulting in a continuous pattern of isolation by distance within each cluster.
 Multiple population cluster processes can co-occur in space, 
@@
 
-The continuous decay of allele frequency covariance with geographic distance 
-within a cluster is described using a simple and flexible powered exponential function;
-the within-cluster covariance between samples $i$ and $j$ is given by:
+\gc{To model the} continuous decay of allele frequency covariance with geographic distance 
+within a cluster \gc{we use a flexible powered-exponential function;
+  where the} within-cluster covariance between samples $i$ and $j$ is given by:
 
@@
 The quantity $D_{i,j}$ is the observed geographic distance between samples $i$ and $j$,
-the $\alpha$ parameters control the shape of the decay of covariance with distance,
-and $\mu$ is a parameter that describes the amount of shared drift within a cluster.
+the \gc{$\alpha^{(k)}$ parameters control the shape of the decay of
+  covariance with distance in the cluster},
+and $\mu$ is a parameter that describes the \gc{background allele
+  frequency   %%amount of shared drift
+  covariance within the cluster. If two samples draw 100\% of their
+  ancestry from cluster $k$ then, if they are geographically very
+  close ($D_{i,j}=0$) they will have allele frequency covariance
+  $\alpha^{(k)}_0 +  \mu^{(k)}$ with their covariance decaying to the
+  background level $\mu^{(k)}$ within the cluster for large enough geographic distances.}
 The shared drift parameter $\mu^{(k)}$ can be interpreted as the branch length connecting 
-the $k$th population to the population ancestral to all modeled clusters.
+the $k$th population to the population ancestral to all modeled
+clusters \gc{(see XXX)}.
 
-Genotyped samples can be admixed between these different clusters. 
+\gc{Genotyped samples can draw their ancestry from multiple different clusters,
+representing these samples being admixed between the clusters.} 
 The admixture proportion of the $i$th sample in the $k$th cluster, $w^{(k)}_i$,
-gives the probability that an allele in sample $i$ was derived from cluster $k$;
-each sample's admixture proportions across all clusters must sum to one.
+gives the probability that an allele in sample $i$ was derived from
+cluster $k$ \gc{($\sum_k w^{(k)}_i =1$).}
 
@@
 by summing their within-cluster spatial covariances ($G_{i,j}^{(k)}$ in cluster $k$) across all $K$ clusters
-and weighting by the product of those samples' admixture proportions in each cluster.
-
+and weighting by the product of those samples' admixture proportions in each cluster
 \begin{equation}
-\Omega_{i,j} = \gamma + \sum\limits_K w^{(k)}_i w^{(k)}_j G^{(k)}_{i,j} + \delta_{i,j}\eta_i
+\Omega_{i,j} = \gamma + \sum\limits_K w^{(k)}_i w^{(k)}_j
+G^{(k)}_{i,j} + \delta_{i,j}\eta_i .
 \label{cross_cluster_covariance}
@@
 The first, $\gamma$, describes the global covariance between all samples, 
-and arises because all samples share an ancestral mean allele frequency at each locus.
+and arises because all samples \gc{have a base-line covariance as they
+  share} an ancestral mean allele frequency at each locus.
 In the second term, $\delta_{i,j}$ is an indicator variable that takes a value of 1 when $i$ equals $j$ and 0 otherwise. 
 and $\eta_i$ gives the variance specific to sample $i$, 
-This term on the diagonal of the parametric covariance matrix 
-is meant to capture processes shaping variance within the sampled deme, 
-such as inbreeding.
+This term on the diagonal of the parametric covariance matrix \gc{captures} processes shaping variance within the sampled deme, 
+such as inbreeding \gc{and the sampling process}.
 
 \paragraph{Likelihood and inference}
-If we assume that the  allele frequencies $\widehat{F}$ are independent and multivariate normally distributed, 
+\gc{We} assume that the allele frequencies \gc{at each locus
+  ($\widehat{F}_{\ell}$) are independent among loci and multivariate normally
+  distributed across populations}, 
 their sample covariance $\widehat{\Omega}$ 
@@
 around this inference machinery.
+
+\gc{
+\paragraph{Relationship of this model to non-spatial structure models}
+
+}
+
+\gc{ 
+\paragraph{Choice of cluster number and cross-validation}
+
+}
 

### 2017-07-27 writeup/conStruct.tex
section on sim results

--- 2017-07-21 writeup/conStruct.tex
@@
 such as principal components analysis \citep{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
-and model-based clustering approaches \citep[e.g.][]{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, FINESTRUCTURE, huelsenbeck2007inference, Corander2003}, among others.
+and model-based clustering approaches \citep[e.g.][]{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, FINESTRUCTURE, huelsenbeck2007inference, Corander2003,TESS}, among others.
 Here, we will focus on model-based clustering, the most widely used class of approaches.
@@
 \paragraph{Data}
-The statistical framework of our approach is conceptually similar to \cite{Wasser2004} and \cite{spacemix}.
+The statistical framework of our approach is conceptually similar to \cite{Wasser2004} and \cite{BEDASSLE}, and \cite{spacemix}.
 The genetic data modeled consist of allele frequencies $F$ at $L$ unlinked, bi-allelic single nucleotide polymorphisms (SNPs) genotyped across $N$ samples.
@@
 or admixed between them (see Fig \ref{sim_setup}).
-We simulated datasets under $K =$ 1 through 4, 
+We simulated datasets under $K =$ 1 through 3, 
 sampling 10,000 unlinked loci from each of 20 haploid individuals from every deme
@@
 then used cross-validation to compare the predictive performance of the models.
+In Fig \ref{simK1_nsp_pies}, we show the results of the nonspatial model 
+run with between 1 and 7 clusters, applied to the dataset simulated under $K=1$,
+and in Fig \ref{xvals} we show the results of the cross-validation analysis for each of the three simulated datasets.
+The admixture pie plots for the spatial and nonspatial models run on the datasets simulated under $K=2$ and $K=3$
+are shown in Figs \ref{simK1_sp_pies}-\ref{simK3_sp_pies}.
+
+A number of results are immediately clear from these tests with simulated data.
+First, across all values of K and for all simulations, 
+the spatial model is preferred over the nonspatial model
+(this is not surprising, as the data were simulated under a spatial model).
+Second, the predictive accuracy of the nonspatial model increases with the number of clusters in the model, 
+as it describes continuous, spatial patterns of relatedness between demes within the same cluster 
+with gradients of admixture across more and more discrete clusters.
+This behavior of the nonspatial model is highlighted in the analyses of data simulated under $K=1$ (Fig \ref{simK1_nsp_pies}):
+as $K$ in the model increases, 
+the sampled area is increasingly divvied up between clusters 
+with ``centers of mass" in the $K$ geographical extremes of the simulated space.
+At $K=2$, the clusters are centered in opposite corners on the map; 
+at $K=3$, in the lower two corners and the top half; 
+at $K=4$, all four corners, and so on.
+
+In contrast, the predictive accuracy of the spatial model of \texttt{conStruct} 
+either plateaus at or declines from the true value of $K$ used to simulate the data (Fig \ref{xvals}, second column).
+And, as $K$ in the spatial model increases, 
+deme membership in clusters beyond the number used to simulate the data does not appreciably increase 
+(Fig \ref{simK1_sp_pies}, \ref{simK2_sp_pies}, \ref{simK3_sp_pies}).
+That is, even when the spatial model is run with $K=7$, 
+the inferred admixture proportions are nearly identical to 
+those estimated under the true value of $K$ for each simulation.
 
 \subsection*{Empirical Applications}
+
+\begin{figure}
+	\centering
+		\subcaptionbox{$K=2$\label{simK1_nsp_pies:K2}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K2.pdf}}
+		\subcaptionbox{$K=3$\label{simK1_nsp_pies:K3}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K3.pdf}}
+		\subcaptionbox{$K=4$\label{simK1_nsp_pies:K4}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K4.pdf}}
+		\subcaptionbox{$K=5$\label{simK1_nsp_pies:K5}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K5.pdf}}
+		\subcaptionbox{$K=6$\label{simK1_nsp_pies:K6}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K6.pdf}}
+		\subcaptionbox{$K=7$\label{simK1_nsp_pies:K7}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K7.pdf}}
+	\caption{
+	Map of admixture proportions estimated using a nonspatial model for $K=2$ through 7.
+	The data were simulated using a single cluster with nearest-neighbor symmetric migration between demes.
+    }\label{simK1_nsp_pies}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=8.5in]{figs/sims/xvals.pdf}}
+	\caption{
+	Cross-validation results for data simulated under $K=1$, $K=2$, and $K=3$, 
+	comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
+	The first panel in each row shows all results; 
+	the second panel shows results for just the spatial model.
+	The inset plots zoom in on cross-validation results outlined in the dotted boxes.
+    }\label{xvals}
+\end{figure}
 
@@
 	\centering
-		\subcaptionbox{$K=2$\label{K2}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K2.pdf}}
-		\subcaptionbox{$K=3$\label{K3}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K3.pdf}}
-		\subcaptionbox{$K=4$\label{K4}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K4.pdf}}
-		\subcaptionbox{$K=5$\label{K5}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K5.pdf}}
-		\subcaptionbox{$K=6$\label{K6}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K6.pdf}}
-		\subcaptionbox{$K=7$\label{K7}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K7.pdf}}
+		\subcaptionbox{$K=2$\label{simK1_sp_pies:K2}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K2.pdf}}
+		\subcaptionbox{$K=3$\label{simK1_sp_pies:K3}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K3.pdf}}
+		\subcaptionbox{$K=4$\label{simK1_sp_pies:K4}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K4.pdf}}
+		\subcaptionbox{$K=5$\label{simK1_sp_pies:K5}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K5.pdf}}
+		\subcaptionbox{$K=6$\label{simK1_sp_pies:K6}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K6.pdf}}
+		\subcaptionbox{$K=7$\label{simK1_sp_pies:K7}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K7.pdf}}
 	\caption{
-	Map of admixture proportions estimated using a nonspatial model for $K=2$ through 7.
-	The data were simulated using a single cluster with nearest-neighbor symmetric migration between demes.
-    }\label{sfig:simK1_nsp_pies}
+	Map of admixture proportions estimated using a spatial model for $K=2$ through 7.
+	The data were simulated using 1 cluster with nearest-neighbor symmetric migration between demes.
+    }\label{simK1_sp_pies}
 \end{figure}
@@
 \newpage
+\begin{figure}
+	\centering
+		\subcaptionbox{$K=2$\label{simK2_nsp_pies:K2}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_nsp_pies_K2.pdf}}
+		\subcaptionbox{$K=3$\label{simK2_nsp_pies:K3}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_nsp_pies_K3.pdf}}
+		\subcaptionbox{$K=4$\label{simK2_nsp_pies:K4}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_nsp_pies_K4.pdf}}
+		\subcaptionbox{$K=5$\label{simK2_nsp_pies:K5}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_nsp_pies_K5.pdf}}
+		\subcaptionbox{$K=6$\label{simK2_nsp_pies:K6}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_nsp_pies_K6.pdf}}
+		\subcaptionbox{$K=7$\label{simK2_nsp_pies:K7}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_nsp_pies_K7.pdf}}
+	\caption{
+	Map of admixture proportions estimated using a nonspatial model for $K=2$ through 7.
+	The data were simulated using two clusters with nearest-neighbor symmetric migration between demes.
+    }\label{simK2_nsp_pies}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\subcaptionbox{$K=2$\label{simK2_sp_pies:K2}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_sp_pies_K2.pdf}}
+		\subcaptionbox{$K=3$\label{simK2_sp_pies:K3}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_sp_pies_K3.pdf}}
+		\subcaptionbox{$K=4$\label{simK2_sp_pies:K4}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_sp_pies_K4.pdf}}
+		\subcaptionbox{$K=5$\label{simK2_sp_pies:K5}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_sp_pies_K5.pdf}}
+		\subcaptionbox{$K=6$\label{simK2_sp_pies:K6}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_sp_pies_K6.pdf}}
+		\subcaptionbox{$K=7$\label{simK2_sp_pies:K7}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_sp_pies_K7.pdf}}
+	\caption{
+	Map of admixture proportions estimated using a spatial model for $K=2$ through 7.
+	The data were simulated using two clusters with nearest-neighbor symmetric migration between demes.
+    }\label{simK2_sp_pies}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\subcaptionbox{$K=2$\label{simK3_nsp_pies:K2}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_nsp_pies_K2.pdf}}
+		\subcaptionbox{$K=3$\label{simK3_nsp_pies:K3}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_nsp_pies_K3.pdf}}
+		\subcaptionbox{$K=4$\label{simK3_nsp_pies:K4}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_nsp_pies_K4.pdf}}
+		\subcaptionbox{$K=5$\label{simK3_nsp_pies:K5}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_nsp_pies_K5.pdf}}
+		\subcaptionbox{$K=6$\label{simK3_nsp_pies:K6}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_nsp_pies_K6.pdf}}
+		\subcaptionbox{$K=7$\label{simK3_nsp_pies:K7}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_nsp_pies_K7.pdf}}
+	\caption{
+	Map of admixture proportions estimated using a nonspatial model for $K=2$ through 7.
+	The data were simulated using three clusters with nearest-neighbor symmetric migration between demes.
+    }\label{simK3_nsp_pies}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\subcaptionbox{$K=2$\label{simK3_sp_pies:K2}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K2.pdf}}
+		\subcaptionbox{$K=3$\label{simK3_sp_pies:K3}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K3.pdf}}
+		\subcaptionbox{$K=4$\label{simK3_sp_pies:K4}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K4.pdf}}
+		\subcaptionbox{$K=5$\label{simK3_sp_pies:K5}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K5.pdf}}
+		\subcaptionbox{$K=6$\label{simK3_sp_pies:K6}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K6.pdf}}
+		\subcaptionbox{$K=7$\label{simK3_sp_pies:K7}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K7.pdf}}
+	\caption{
+	Map of admixture proportions estimated using a spatial model for $K=2$ through 7.
+	The data were simulated using three clusters with nearest-neighbor symmetric migration between demes.
+    }\label{simK3_sp_pies}
+\end{figure}
+
 \newpage
+\clearpage
 \bibliography{reference.library.tex.bib}

### 2017-07-27 sims/cross_validation/summarize.xvals.R

@@ -60,7 +60,7 @@ get.xval.CIs <- function(x.vals.std,K){
 				"nsp.CIs" = nsp.CIs))
 }
 
-plot.xval.CIs <- function(xval.CIs,K,k.range=c(1:K),ylim=NULL,simK=NULL){
+plot.xval.CIs <- function(xval.CIs,K,k.range=c(1:K),ylim=NULL,simK=NULL,cex=1.5,...){
 	#recover()
 	if(is.null(ylim)){
 		ylim <- range(c(unlist(lapply(k.range,function(k){xval.CIs$sp.CIs[[k]]})),
@@ -69,9 +69,9 @@ plot.xval.CIs <- function(xval.CIs,K,k.range=c(1:K),ylim=NULL,simK=NULL){
 	plot(xval.CIs$sp.means,
 			ylim=ylim,
 			xlim=range(k.range),
-			main= paste0("Cross-validation results",simK),
-			ylab="mean log p(testing data partition | training params)",
-			xlab="number of clusters",type='n')
+			main= "",
+			ylab="",
+			xlab="",type='n',...)
 		lapply(1:K,function(k){
 				segments(k,xval.CIs$sp.CIs[[k]][1],
 						 k,xval.CIs$sp.CIs[[k]][2],
@@ -80,7 +80,58 @@ plot.xval.CIs <- function(xval.CIs,K,k.range=c(1:K),ylim=NULL,simK=NULL){
 				segments(k,xval.CIs$nsp.CIs[[k]][1],
 						 k,xval.CIs$nsp.CIs[[k]][2],
 						 col=adjustcolor("green",0.5),lwd=3)})
-		points(xval.CIs$sp.means,pch=19,col=4,cex=1.5,)			
-		points(xval.CIs$nsp.means,col="green",pch=19,cex=1.5)
+		points(xval.CIs$sp.means,pch=19,col=4,cex=cex)			
+		points(xval.CIs$nsp.means,col="green",pch=19,cex=cex)
 	return(invisible("plotted"))
+}
+
+make.admix.pie.plot.tmp <- function(coords, K, N, conStruct.results, cluster.colors, stat, radii = 2.7, add = FALSE, title = NULL, x.lim = NULL, y.lim = NULL) 
+{
+    if (is.null(coords)) {
+        message("\nuser has not specified sampling coordinates in the data block\n")
+    }
+    else {
+        require(caroline)
+        cluster.names <- paste0("cluster_", 1:K)
[…]

### 2017-07-27 writeup/figs/sim_figs.R

@@ -17,15 +17,18 @@ for(n in 1:n.reps){
 x.vals <- lapply(1:n.reps,function(n){get(sprintf("tl%s",n))})
 x.vals.std <- lapply(x.vals,standardize.xvals)
 xval.CIs <- get.xval.CIs(x.vals.std,K)
-pdf(file="~/Dropbox/conStruct/writeup/figs/sims/simK1_std.xval.pdf",width=6,height=6,pointsize=14)
+pdf(file="~/Dropbox/conStruct/writeup/figs/sims/simK1_std_xval.pdf",width=10,height=5,pointsize=14)
+#	quartz(width=10,height=5)
+	par(mfrow=c(1,2),mar=c(4,5,4,2))
 	plot.xval.CIs(xval.CIs,K,simK=" (K = 1)")
-	legend(x="bottomright",pch=19,col=c("blue","green"),legend=c("spatial","nonspatial"))
+		legend(x="bottomright",pch=19,col=c("blue","green"),legend=c("spatial","nonspatial"))
+	mtext("Predictive accuracy",side=2,padj=-5)
+	plot.xval.CIs(xval.CIs,K,simK=" (K = 1)",ylim=c(-10,0))
+		legend(x="bottomleft",pch=c(19,NA),lty=c(NA,1),lwd=c(NA,2),col=c(1,adjustcolor(1,0.8)),legend=c("mean","95% CI"))
+	mtext("number of clusters",side=1,adj=-0.95,padj=4)
+	mtext("Cross-validation results (K=1)",side=3,adj=70,padj=-2.5,font=2,cex=1.2)
 dev.off()
 
-pdf(file="~/Dropbox/conStruct/writeup/figs/sims/simK1_std.xval_sp.pdf",width=6,height=6,pointsize=14)
-	plot.xval.CIs(xval.CIs,K,simK=" (K = 1)",ylim=c(-10,0))
-	legend(x="topright",pch=c(19,NA),lty=c(NA,1),lwd=c(NA,2),col=c(1,adjustcolor(1,0.8)),legend=c("mean","95% CI"))
-dev.off()
 
 library(conStruct)
 for(k in 2:7){
@@ -51,8 +54,35 @@ for(k in 2:7){
 	dev.off()
 }
 
+library(conStruct)
+for(k in 2:7){
+	load(sprintf("simK1__sp_rep1K%s_data.block.Robj",k))
[…]

### 2017-07-27 writeup/conStruct.tex
merged graham's edits

--- 2017-07-27 writeup/conStruct.tex
@@
 Gideon S. Bradburd$^{1,a}$, 
-Peter L. Ralph$^{2,b}$, 
-Graham M. Coop$^{3,c}$}
+Graham M. Coop$^{2,b}$,
+Peter L. Ralph$^{3,c}$}
 
@@
 \textsuperscript{2}
+Center for Population Biology,
+Department of Evolution and Ecology, 
+University of California, Davis, CA 95616
+
+\textsuperscript{3}
 Institute of Ecology and Evolution,
@@
 
-\textsuperscript{3}
-Center for Population Biology,
-Department of Evolution and Ecology, 
-University of California, Davis, CA 95616
-
 \textsuperscript{a}bradburd@msu.edu; 
-\textsuperscript{b}plr@uoregon.edu;
-\textsuperscript{c}gmcoop@ucdavis.edu\\\\\
+\textsuperscript{c}gmcoop@ucdavis.edu;
+\textsuperscript{b}plr@uoregon.edu\\\\\
 
@@
 and model-based clustering approaches \citep[e.g.][]{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, FINESTRUCTURE, huelsenbeck2007inference, Corander2003,TESS}, among others.
-Here, we will focus on model-based clustering, the most widely used class of approaches.
+Here, we will focus on model-based clustering, the most widely used class of approaches for population delineation.
 These methods each perform best under different specific demographic histories of samples being analyzed,
@@
 Given the generality of the circumstances that generate a pattern of isolation by distance, 
-it is unsurprising that IBD is ubiquitous in nature \citep{meirmans2012,Sexton_etal_2014}.
+it is unsurprising that IBD is found everywhere in nature \citep{meirmans2012,Sexton_etal_2014}.
 
@@
 that are spuriously assumed to be discrete management units.
-In studies of locally adaptive loci, or of the genetics of disease,
+Or, in studies of locally adaptive loci, or of the genetics of disease,
 discrete populations may be modeled in separately,
@@
 
-What is desired, then, is a model-based clustering method that incorporates a spatial element.
+What is desired, then, is a model-based clustering method that incorporates a spatial element,
 and specifically one that uses a model of isolation by distance as a biologically sensible null hypothesis 
@@
 We model genetic variation in genotyped individuals as 
-partitioned within or admixed across a user-specified number of discrete clusters,
+partitioned within or admixed across a specified number of discrete clusters,
 within each of which relatedness decays as a parametric function of the distance between samples.
@@
 and we demonstrate the utility of our approach using both simulated and empirical data.
-The implementation of this method, \texttt{conStruct}, is available as an R package at \href{https://github.com/gbradburd/conStruct}{https://github.com/gbradburd/conStruct}.
+The implementation of this method, \texttt{conStruct}, is available as an R package at 
+\href{https://github.com/gbradburd/conStruct}{https://github.com/gbradburd/conStruct}.
 
@@
 in sample $n$.
-We then calculate the sample covariance in allele frequencies, $\widehat{\Omega}$, as
+We then calculate the sample covariance in allele frequencies, $\widehat{\Omega}_{i,j}$ between samples $i$ and $j$.
+However, for the diagonal of this sample covariance (the variances within each sample),
+we wish to describe the mean of the variance in allele frequencies between haplotypes in a sample, 
+rather than the variance of the mean frequencies across haplotypes in a sample.
+Therefore, we calculate our sample covariance as follows:
 
 \begin{equation}
-\widehat{\Omega} = \frac{1}{L}  FF^T
+\widehat{\Omega}_{i,j} = 
+	\begin{cases}
+		\frac{1}{L}
+		\sum\limits_{\ell=1}^{L}{\left[ 	\left(f_{i,\ell} - \bar{f}_{\ell} \right)
+		 	\left(f_{j,\ell} - \bar{f}_{\ell} \right)\right]
+			} & \text{for } i \neq j \\
+		\text{Var}_L\left( \text{B}(1,f_{i,\ell})	\right) & \text{for } i = j \\
+	\end{cases}
 \label{sample_covariance}
 \end{equation}
+\gb{ideas for how to describe what I'm doing on the diagonal in actual math notation?}
 
@@
 and between which sampled individuals or populations can be admixed.
-Each of these discrete clusters is modeled as a spatial process
-in which migration between neighboring demes acts 
-to homogenize allele frequency changes that arise locally due to drift;
-this interplay between drift and migration 
-results in a continuous pattern of isolation by distance within each cluster.
+Each of these discrete clusters is modeled as a spatial process in which
+allele frequencies covary between geographically close locations,
+but this covariance is allowed to decay with geographic distance.
+This pattern of spatial decay reflects how migration between neighboring demes  
+homogenizes allele frequency changes that arise locally due to drift, 
+but is less effective at homogenizing between geographic distant demes,
+resulting in a continuous pattern of isolation by distance within each cluster.
 Multiple population cluster processes can co-occur in space, 
@@
 
-The continuous decay of allele frequency covariance with geographic distance 
-within a cluster is described using a simple and flexible powered exponential function;
+To model the continuous decay of allele frequency covariance with geographic distance within a cluster, 
+we use a flexible powered-exponential function;
 the within-cluster covariance between samples $i$ and $j$ is given by:
@@
 The quantity $D_{i,j}$ is the observed geographic distance between samples $i$ and $j$,
-the $\alpha$ parameters control the shape of the decay of covariance with distance,
-and $\mu$ is a parameter that describes the amount of shared drift within a cluster.
-The shared drift parameter $\mu^{(k)}$ can be interpreted as the branch length connecting 
-the $k$th population to the population ancestral to all modeled clusters.
+the $\alpha^{(k)}$ parameters control the shape of the decay of covariance with distance in the cluster,
+and $\mu$ is a parameter that describes the background allele frequency covariance within the cluster. 
+If two samples draw 100\% of their ancestry from cluster $k$ then, 
+if they are geographically very close ($D_{i,j}=0$) 
+they will have allele frequency covariance $\alpha^{(k)}_0 +  \mu^{(k)}$.
+If the geographic distance between them is very large, 
+their covariance will decay to the background level $\mu^{(k)}$ within the cluster.
+The shared drift parameter $\mu^{(k)}$ can be interpreted as the branch length 
+connecting the $k$th population to the population ancestral to all modeled
+clusters \gc{(see XXX)}. \gb{is ``XXX" TreeMix? or one of the $f$stat papers?  or is it a figure?}
 
-Genotyped samples can be admixed between these different clusters. 
+Genotyped samples can draw their ancestry from multiple different clusters,
+representing these samples being admixed between the clusters.
 The admixture proportion of the $i$th sample in the $k$th cluster, $w^{(k)}_i$,
-gives the probability that an allele in sample $i$ was derived from cluster $k$;
-each sample's admixture proportions across all clusters must sum to one.
+gives the probability that an allele in sample $i$ was derived from
+cluster $k$ ($\sum_K w^{(k)}_i =1$).
 
@@
 The first, $\gamma$, describes the global covariance between all samples, 
-and arises because all samples share an ancestral mean allele frequency at each locus.
+and arises because all samples share an ancestral mean allele frequency at each locus,
+which generates a base-line covariance.
 In the second term, $\delta_{i,j}$ is an indicator variable that takes a value of 1 when $i$ equals $j$ and 0 otherwise. 
 and $\eta_i$ gives the variance specific to sample $i$, 
-This term on the diagonal of the parametric covariance matrix 
-is meant to capture processes shaping variance within the sampled deme, 
-such as inbreeding.
+This term on the diagonal of the parametric covariance matrix captures processes shaping variance within the sampled deme, 
+such as inbreeding and the sampling process.
 
 \paragraph{Likelihood and inference}
-If we assume that the  allele frequencies $\widehat{F}$ are independent and multivariate normally distributed, 
-their sample covariance $\widehat{\Omega}$ 
-will be Wishart distributed with degrees of freedom equal to $L$, 
+We assume that the allele frequencies at each locus ($\widehat{F}_{\ell}$) 
+are independent among loci and multivariate normally distributed across populations, 
+so that their sample covariance $\widehat{\Omega}$ 
+is  Wishart distributed with degrees of freedom equal to $L$, 
 the number of loci genotyped.
@@
 Acknowledging the dependence of the parametric covariance matrix $\Omega$ on its constituent parameters
-$w,\alpha,\mu,\eta,\gamma$ and on observed quantiy $D$ with the notation $\Omega(w,\alpha,\mu,\eta,\gamma,D)$,
+$w,\alpha,\mu,\eta,\gamma$ and on observed quantity $D$ 
+with the notation $\Omega(w,\alpha,\mu,\eta,\gamma,D)$,
 we denote the posterior probability of the parameters as:
@@
 
-The priors, $P(w),P(\alpha),P(\mu),P(\eta),P(\gamma)$, are detailed in the Appendix,
+The priors, $P(w),P(\alpha),P(\mu),P(\eta),P(\gamma)$, are detailed in Table \ref{tab:param_prior_tab},
 and the constant of proportionality is the normalization constant.
@@
 then used cross-validation to compare the predictive performance of the models.
-In Fig \ref{simK1_nsp_pies}, we show the results of the nonspatial model 
+In Fig \ref{simK1_pies}, we show the results of the nonspatial model 
 run with between 1 and 7 clusters, applied to the dataset simulated under $K=1$,
@@
 The admixture pie plots for the spatial and nonspatial models run on the datasets simulated under $K=2$ and $K=3$
-are shown in Figs \ref{simK1_sp_pies}-\ref{simK3_sp_pies}.
+are shown in Figs \ref{simK1_nsp_pies}-\ref{simK3_sp_pies}.
 
@@
 with gradients of admixture across more and more discrete clusters.
-This behavior of the nonspatial model is highlighted in the analyses of data simulated under $K=1$ (Fig \ref{simK1_nsp_pies}):
+This behavior of the nonspatial model is highlighted in the analyses of data simulated under $K=1$ (Fig \ref{simK1_pies} and \ref{simK1_nsp_pies}):
 as $K$ in the model increases, 
@@
 
-\subsection*{Empirical Applications}
-
 \begin{figure}
@@
 			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K4.pdf}}
-		\subcaptionbox{$K=5$\label{simK1_nsp_pies:K5}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K5.pdf}}
-		\subcaptionbox{$K=6$\label{simK1_nsp_pies:K6}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K6.pdf}}
-		\subcaptionbox{$K=7$\label{simK1_nsp_pies:K7}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K7.pdf}}
+		\subcaptionbox{$K=2$\label{simK1_sp_pies:K2}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K5.pdf}}
+		\subcaptionbox{$K=3$\label{simK1_sp_pies:K3}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K6.pdf}}
+		\subcaptionbox{$K=4$\label{simK1_sp_pies:K4}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K7.pdf}}
 	\caption{
-	Map of admixture proportions estimated using a nonspatial model for $K=2$ through 7.
+	Map of admixture proportions estimated using the nonspatial model for $K=2$ through 4 (top row)
+	and also the spatial model for $K=2$ through 4 (bottom row).
 	The data were simulated using a single cluster with nearest-neighbor symmetric migration between demes.
-    }\label{simK1_nsp_pies}
+    }\label{simK1_pies}
 \end{figure}
@@
 \end{figure}
+
+\subsection*{Empirical Applications}
 
@@
 	\centering
+		\subcaptionbox{$K=2$\label{simK1_nsp_pies:K2}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K2.pdf}}
+		\subcaptionbox{$K=3$\label{simK1_nsp_pies:K3}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K3.pdf}}
+		\subcaptionbox{$K=4$\label{simK1_nsp_pies:K4}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K4.pdf}}
+		\subcaptionbox{$K=5$\label{simK1_nsp_pies:K5}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K5.pdf}}
+		\subcaptionbox{$K=6$\label{simK1_nsp_pies:K6}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K6.pdf}}
+		\subcaptionbox{$K=7$\label{simK1_nsp_pies:K7}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K7.pdf}}
+	\caption{
+	Map of admixture proportions estimated using a nonspatial model for $K=2$ through 7.
+	The data were simulated using a single cluster with nearest-neighbor symmetric migration between demes.
+    }\label{simK1_nsp_pies}
+\end{figure}
+
+\begin{figure}
+	\centering
 		\subcaptionbox{$K=2$\label{simK1_sp_pies:K2}}

### 2017-07-28 writeup/conStruct.tex
added xvalidation section

@@
 
-\gc{ 
 \paragraph{Choice of cluster number and cross-validation}
+There are a number of reasons, both statistical and biological, 
+why choosing a single, ``best" number of clusters to describe the data may not be desirable.
+%, including ambiguity in what exactly it means for any given value of $K$ to be ``best."
+We discuss these reasons further in the Discussion below, 
+but we also recognize that in many circumstances it is either useful or convenient 
+to present only a single set of results from a clustering analysis with a particular number of clusters.
+ideally a number large enough to describe meaningful partitions in the data, 
+but not so large as to be difficult to parse visually.
+Toward that end, we have implemented a method for 
+statistically comparing \texttt{conStruct} runs across different values of $K$.
 
-}
+Several approaches have been proposed to determine the number of clusters in the data, 
+including: 
+the use of information theoretic criteria \citep[e.g.][]{ADMIXTURE}, such as AIC or BIC;
+comparisons of the likelihood of the data across different values of $K$, 
+with various criteria on how to choose a single value (e.g., Evanno 2005);
+or comparisons of the marginal likelihood, 
+generated either via various approximations \citep[e.g.][]{STRUCTURE}
+or via a more rigorous and computationally intensive approach such as thermodynamic integration \citep{verity_nichols2016}
+or inference using a Dirichlet process prior \citep{huelsenbeck2007inference}
+A comprehensive review of these approaches and their various merits is beyond the scope of this paper,
+but see \cite{verity_nichols2016} for an excellent discussion of and comparison between several.
+
+We choose to take a cross-validation approach, 
+estimating the parameters of the model using a ``training" partition of the data,
+then calculating the likelihood of a separate, ``testing" partition of the data given those parameters,
+repeating this procedure many times, 
+and selecting the model with the significantly best predictive accuracy
+(see Section \ref{Xvalidation} for a thorough discussion of our cross-validation procedure).
+This approach rewards models with larger values of $K$, which increase the likelihood of the training partition,
+up to the point that cluster membership is invoked to model noise specific to the training data, 
+at which point predictive accuracy decreases, 
+and models with additional clusters are penalized.
 
@@
 those estimated under the true value of $K$ for each simulation.
+We parse \texttt{conStruct}'s performance on these simulation tests further in the Discussion.
 
@@
 \subsection*{Empirical Applications}
+To demonstrate the utility of this method, we analyzed empirical population genomic data from two systems:
+a contact zone between two poplar species in northwestern North America,
+and a large North American sample of black bears.
+
+\subsubsection*{Poplars}
+
+\subsubsection*{Black bears}
 
 \section*{Discussion}
+
+\paragraph{Performance on simulated data}
+\paragraph{Empirical results}
+\paragraph{Choosing the ``best" number of clusters}
+\paragraph{Implications for management and conservation}
+\paragraph{Extensions and future directions}
 
@@
 \setcounter{figure}{0}
+\renewcommand{\thesection}{A\arabic{section}}
+\setcounter{section}{0}
 
@@
 \newpage
-\section{Cross validation}
+\section{Cross validation}\label{Xvalidation}
 To perform model comparison, we employ a Monte Carlo cross-validation approach \citep{picard1984},

### 2017-07-28 writeup/conStruct.tex
G's structure comparison

--- 2017-07-27 writeup/conStruct.tex
@@
 \paragraph{Relationship of this model to non-spatial structure models}
+
+A nice feature of our approach is that our model contains a non-spatial assignment model as a special case. 
+Setting $\alpha_0^{(k)}$ to zero for all $k$, we obtain a non-spatial model where
+each cluster has its own allele frequency at each SNP, and individuals
+draw a proportion of their ancestry from each cluster. This model is
+strongly conceptually similar to that of STRUCTURE (and related
+approaches ADMIXTURE), the only difference being that our likelihood
+assumes that allele frequencies are normally distributed around their
+expectations, while the standard assignment methods assume that the
+error is binomial distributed (see XXXX). Therefore, to compare the fit
+of
 

### 2017-07-30 writeup/conStruct.tex
added to x-validation

@@
 \paragraph{Choice of cluster number and cross-validation}
-There are a number of reasons, both statistical and biological, 
-why choosing a single, ``best" number of clusters to describe the data may not be desirable.
+There are a number of reasons, both statistical and biological, why
+there is no true (or right) number of clusters for real datasets. 
+%There are a number of reasons, both statistical and biological, 
+%why choosing a single, ``best" number of clusters to describe the data may not be desirable.
 %, including ambiguity in what exactly it means for any given value of $K$ to be ``best."
-We discuss these reasons further in the Discussion below, 
-but we also recognize that in many circumstances it is either useful or convenient 
-to present only a single set of results from a clustering analysis with a particular number of clusters.
-ideally a number large enough to describe meaningful partitions in the data, 
-but not so large as to be difficult to parse visually.
+We discuss these reasons further in the Discussion below. In practice
+it is useful to explore the implications of clustering at a variety of
+different levels (i.e. across multiple $K$) for biological
+intepretation and downstream analyses. 
+However, we also recognize that in many circumstances it is 
+useful to know the number of clusters with highest statistical
+support in the (current) dataset.  %, for example to focus su 
+%However, we also recognize that in many circumstances 
+i%t is either useful or convenient 
+%to present only a single set of results from a clustering analysis with a particular number of clusters.
+%ideally a number large enough to describe meaningful partitions in the data, 
+%but not so large as to be difficult to parse visually.
 Toward that end, we have implemented a method for 
-statistically comparing \texttt{conStruct} runs across different values of $K$.
+statistically comparing \texttt{conStruct} results across different values of $K$.
 
-Several approaches have been proposed to determine the number of clusters in the data, 
-including: 
+Several approaches have been used as model choice criteria for the number of
+clusters in \gc{population genetic} data, including: 
 the use of information theoretic criteria \citep[e.g.][]{ADMIXTURE}, such as AIC or BIC;
 comparisons of the likelihood of the data across different values of $K$, 
-with various criteria on how to choose a single value (e.g., Evanno 2005);
+with various criteria on how to choose a single value (e.g., \citep{Evanno2005});
 or comparisons of the marginal likelihood, 
@@
 or inference using a Dirichlet process prior \citep{huelsenbeck2007inference}
-A comprehensive review of these approaches and their various merits is beyond the scope of this paper,
-but see \cite{verity_nichols2016} for an excellent discussion of and comparison between several.
+See \cite{verity_nichols2016} for a discussion of these approaches and comparison
+between several methods. %A comprehensive review of these approaches and their various merits is beyond the scope of this paper,
+%but
 
-We choose to take a cross-validation approach, 
-estimating the parameters of the model using a ``training" partition of the data,
-then calculating the likelihood of a separate, ``testing" partition of the data given those parameters,
-repeating this procedure many times, 
+\gc{IS PREDICTION accuracy the right phrase?}
+We choose to take a cross-validation approach.  We estimate the
+posterior on the parameters of the model using a ``training" partition
+of the data, in practice the training set is a random subset of $90\%$
+of the loci. We then calculate the likelihood of the
+``testing" subset, the remaining $10\%$ of the loci, averaged over the
+posterior of the parameters. We repeat this procedure a number of
+times, take an average of the relative log-likelihood over the runs at
+sumarize each $K$. 
 and selecting the model with the significantly best predictive accuracy
-(see Section \ref{Xvalidation} for a thorough discussion of our cross-validation procedure).
-This approach rewards models with larger values of $K$, which increase the likelihood of the training partition,
-up to the point that cluster membership is invoked to model noise specific to the training data, 
-at which point predictive accuracy decreases, 
-and models with additional clusters are penalized.
+(see Section \ref{Xvalidation} for more on our cross-validation procedure).
+This approach rewards models with larger values of $K$, which increase
+the likelihood of the training partition, as they fit the data better,
+but the improvement in the likelihood will plateau (or peak) as above
+a certain $K$ we are fitting noise specific to the training data and
+so not fitting the testing set any better, 
+%up to the point that cluster membership is invoked to model noise specific to the training data, 
+%at which point predictive accuracy decreases, 
+%and models with additional clusters are penalized.
 

### 2017-08-01 writeup/conStruct.tex
added poplar figs

@@
 and so tend to erroneously describe continuously distributed variation with multiple clusters that 
-show spatially autocorrelated cluster membership \citep{Frantz2009,Meirmans2012}.
+show spatially autocorrelated cluster membership \citep{Frantz2009,meirmans2012}.
 In analyses of empirical datasets, which often are characterized by IBD,
@@
 run with between 1 and 7 clusters, applied to the dataset simulated under $K=1$,
-and in Fig \ref{xvals} we show the results of the cross-validation analysis for each of the three simulated datasets.
+and in Fig \ref{sim_xvals} we show the results of the cross-validation analysis for each of the three simulated datasets.
 The admixture pie plots for the spatial and nonspatial models run on the datasets simulated under $K=2$ and $K=3$
@@
 In contrast, the predictive accuracy of the spatial model of \texttt{conStruct} 
-either plateaus at or declines from the true value of $K$ used to simulate the data (Fig \ref{xvals}, second column).
+either plateaus at or declines from the true value of $K$ used to simulate the data (Fig \ref{sim_xvals}, second column).
 And, as $K$ in the spatial model increases, 
@@
 	The inset plots zoom in on cross-validation results outlined in the dotted boxes.
-    }\label{xvals}
+    }\label{sim_xvals}
 \end{figure}
@@
 \subsubsection*{Poplars}
+
+Trees in the genus \textit{Populus} (poplars, aspens, and cottonwoods) 
+are distributed throughout the Northern Hemisphere, 
+and include species of ecological and economic importance.
+Species in the genus regularly co-occur and, 
+where they do, they frequently hybridize (ref).
+In addition, many \textit{Populus} species display patterns of isolation by distance, 
+making the system well-suited to an analysis designed to tease apart 
+genetic structure partitioned discretely across population (or species) clusters,
+and continuously with geographic distance.
+
+Here, we focus on two \textit{Populus} species in the Pacific Northwest: 
+\textit{Populus trichocarpa}, the black cottonwood,
+and \textit{Populus trichocarpa}, the balsam poplar.
+These species have a broad zone of overlap and are known to hybridize \citep{geraldes_etal_2014}.
+
+We address this question using a dataset generated and originally published by \cite{geraldes_etal_2014}.
+The published data consist of 434 individuals sampled from 35 drainages 
+and genotyped at just over 33,000 loci (map of the sampling shown in Fig \ref{populus_map}).
+The data were generated using RAD-seq, 
+and showed a strong pattern of bias in allelic dropout 
+(the majority of missing data were from drainages with only \textit{Populus balsamifera}).
+To ameliorate some of the problems that arise when there is a strong bias in which data are missing, 
+we dropped loci for which any data were missing, 
+resulting in just over 20,200 loci retained for analysis.  
+We then analyzed these data using both the spatial and nonspatial \texttt{conStruct} models, 
+running on $K = 1$ through 7,
+and compared these models using the cross-validation procedure described above. 
+The results of these analyses are shown in 
+Figs \ref{populus_xvals}-\ref{populus_nsp_clst_covs}. %,populus_nsp_pies,populus_sp_clst_covs,populus_nsp_clst_covs
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=5in]{figs/populus/populus_sampling_map.pdf}}
+	\caption{
+	Map of the sampled \textit{Populus} populations included in the analysis.
+	The color of the sampling location denotes the putative species.
+    }\label{populus_map}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=3in]{figs/populus/populus_std_xval.pdf}}
+	\caption{
+	Cross-validation results for \textit{Populus} dataset 
+	comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
+	The first panel in each row shows all results; 
+	the second panel zooms in on the results from analyses run with $K = 2$ through 7.
+    }\label{populus_xvals}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\subcaptionbox{$K=2$\label{populus_sp_pies:K2}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp2.pdf}}
+		\subcaptionbox{$K=3$\label{populus_sp_pies:K3}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp3.pdf}}
+		\subcaptionbox{$K=4$\label{populus_sp_pies:K4}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp4.pdf}}
+		\subcaptionbox{$K=5$\label{populus_sp_pies:K5}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp5.pdf}}
+		\subcaptionbox{$K=6$\label{populus_sp_pies:K6}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp6.pdf}}
+		\subcaptionbox{$K=7$\label{populus_sp_pies:K7}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp7.pdf}}
+	\caption{
+	Map of admixture proportions estimated for the \textit{Populus} dataset 
+	using the spatial model for $K=2$ through 7.
+    }\label{populus_sp_pies}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\subcaptionbox{$K=2$\label{populus_nsp_pies:K2}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp2.pdf}}
+		\subcaptionbox{$K=3$\label{populus_nsp_pies:K3}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp3.pdf}}
+		\subcaptionbox{$K=4$\label{populus_nsp_pies:K4}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp4.pdf}}
+		\subcaptionbox{$K=5$\label{populus_nsp_pies:K5}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp5.pdf}}
+		\subcaptionbox{$K=6$\label{populus_nsp_pies:K6}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp6.pdf}}
+		\subcaptionbox{$K=7$\label{populus_nsp_pies:K7}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp7.pdf}}
+	\caption{
+	Map of admixture proportions estimated for the \textit{Populus} dataset 
+	using the nonspatial model for $K=2$ through 7.
+    }\label{populus_nsp_pies}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=4in]{figs/populus/populus_sp_clst_covs.pdf}}
+	\caption{
+	Plots showing the cluster-specific parametric covariance IBD-curves 
+	estimated for the \textit{Populus} data using 
+	the spatial \texttt{conStruct} model run with $K=2$ through 7.
+	Colors are consistent with Fig \ref{populus_sp_pies}.
+    }\label{populus_sp_clst_covs}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=4in]{figs/populus/populus_nsp_clst_covs.pdf}}
+	\caption{
+	Plots showing the cluster-specific parametric covariance IBD-curves 
+	estimated for the \textit{Populus} data using 
+	the nonspatial \texttt{conStruct} model run with $K=2$ through 7.
+	Colors are consistent with Fig \ref{populus_nsp_pies}.
+    }\label{populus_nsp_clst_covs}
+\end{figure}
 

### 2017-08-01 writeup/figs/populus_figs.R

@@ -1,0 +1,197 @@
+source("~/Dropbox/conStruct/sims/cross_validation/summarize.xvals.R")
+
+plot.cluster.curves <- function(data.block, conStruct.results, cluster.cols=NULL){
+	if(is.null(cluster.cols)){
+		cluster.cols <- c("blue", "red", "green", "yellow", "purple","orange", "lightblue", "darkgreen", "lightblue", "gray")
+	}
+    order.mat <- order(data.block$geoDist)
+    y.range <- range(c(
+    			unlist(lapply(1:data.block$K,
+    							function(k){
+							        conStruct.results$MAP$cluster.params[[k]]$cluster.cov
+				})) + conStruct.results$MAP$gamma, 
+				data.block$obsCov))
+    plot(data.block$geoDist[upper.tri(data.block$obsCov, diag = TRUE)], 
+        data.block$obsCov[upper.tri(data.block$obsCov, diag = TRUE)], 
+        xlim = range(data.block$geoDist), ylim = y.range,
+        xlab="",ylab="",pch=19,col=adjustcolor(1,0.7))
+    lapply(1:data.block$K, function(k) {
+             lines(data.block$geoDist[order.mat],
+             		conStruct.results$MAP$gamma + 
+             		conStruct.results$MAP$cluster.params[[k]]$cluster.cov[order.mat],
+                  col = 1,lwd=4.5,lty=1) ; 
+             lines(data.block$geoDist[order.mat],
+             		conStruct.results$MAP$gamma + 
+             		conStruct.results$MAP$cluster.params[[k]]$cluster.cov[order.mat],
+                  col = cluster.cols[k],lwd=4,lty=1)
+    })
+    return(invisible("cluster covs"))
+}
+
+
+setwd("~/Dropbox/gid_runs/populus/all/runs")
+n.reps <- 10
+K <- 7
+for(n in 1:n.reps){
+	load(sprintf("poplar_rep%srep%s_test.lnl.Robj",n,n))
+	assign(paste0("tl",n),test.lnl)
+}
+x.vals <- lapply(1:n.reps,function(n){get(sprintf("tl%s",n))})
+x.vals.std <- lapply(x.vals,standardize.xvals)
[…]

### 2017-08-03 writeup/conStruct.tex
new poplar figs, fixed eqn.1

@@
 	\begin{cases}
-		\frac{1}{L}
-		\sum\limits_{\ell=1}^{L}{\left[ 	\left(f_{i,\ell} - \bar{f}_{\ell} \right)
-		 	\left(f_{j,\ell} - \bar{f}_{\ell} \right)\right]
+		\frac{1}{L-1}
+		\sum\limits_{\ell=1}^{L}{\left[ 	\left(f_{i,\ell} - \bar{f}_{i} \right)
+		 	\left(f_{j,\ell} - \bar{f}_{j} \right)\right]
 			} & \text{for } i \neq j \\
-		\text{Var}_L\left( \text{B}(1,f_{i,\ell})	\right) & \text{for } i = j \\
+		\bar{f}_i (1-\bar{f}_i) & \text{for } i = j \\
 	\end{cases}
@@
 \end{equation}
-\gb{ideas for how to describe what I'm doing on the diagonal in actual math notation?}
 
@@
 } 
+\gb{support for the spatial model over the nonspatial model at a given value of $K$ 
+indicates that isolation-by-distance is likely a feature of the data at that level.}
 
@@
 	the spatial \texttt{conStruct} model run with $K=2$ through 7.
-	Colors are consistent with Fig \ref{populus_sp_pies}.
+	Line colors are consistent with Fig \ref{populus_sp_pies}.
+	Points are colored by the species they are a covariance between:
+	black on black points are sample covariances between populations of \textit{Populus balsamifera};
+	green on black points are sample covariances between \textit{P. balsamifera} and \textit{P. trichocarpa};
+	green on green points are sample covariances between \textit{P. trichocarpa} and \textit{P. trichocarpa}.
     }\label{populus_sp_clst_covs}
@@
 	the nonspatial \texttt{conStruct} model run with $K=2$ through 7.
-	Colors are consistent with Fig \ref{populus_nsp_pies}.
+	Line colors are consistent with Fig \ref{populus_nsp_pies}.
+	Points are colored by the species they are a covariance between:
+	black on black points are sample covariances between populations of \textit{Populus balsamifera};
+	green on black points are sample covariances between \textit{P. balsamifera} and \textit{P. trichocarpa};
+	green on green points are sample covariances between \textit{P. trichocarpa} and \textit{P. trichocarpa}.
     }\label{populus_nsp_clst_covs}

### 2017-08-03 sims/cross_validation/summarize.xvals.R

@@ -60,7 +60,7 @@ get.xval.CIs <- function(x.vals.std,K){
 				"nsp.CIs" = nsp.CIs))
 }
 
-plot.xval.CIs <- function(xval.CIs,K,k.range=c(1:K),ylim=NULL,simK=NULL,cex=1.5,...){
+plot.xval.CIs <- function(xval.CIs,K,k.range=c(1:K),ylim=NULL,cex=1.5,jitter=0,...){
 	#recover()
 	if(is.null(ylim)){
 		ylim <- range(c(unlist(lapply(k.range,function(k){xval.CIs$sp.CIs[[k]]})),
@@ -68,7 +68,6 @@ plot.xval.CIs <- function(xval.CIs,K,k.range=c(1:K),ylim=NULL,simK=NULL,cex=1.5,
 	}
 	plot(xval.CIs$sp.means,
 			ylim=ylim,
-			xlim=range(k.range),
 			main= "",
 			ylab="",
 			xlab="",type='n',...)
@@ -77,11 +76,11 @@ plot.xval.CIs <- function(xval.CIs,K,k.range=c(1:K),ylim=NULL,simK=NULL,cex=1.5,
 						 k,xval.CIs$sp.CIs[[k]][2],
 						 col=adjustcolor(4,0.5),lwd=3)})
 		lapply(1:K,function(k){
-				segments(k,xval.CIs$nsp.CIs[[k]][1],
-						 k,xval.CIs$nsp.CIs[[k]][2],
+				segments(k+jitter,xval.CIs$nsp.CIs[[k]][1],
+						 k+jitter,xval.CIs$nsp.CIs[[k]][2],
 						 col=adjustcolor("green",0.5),lwd=3)})
 		points(xval.CIs$sp.means,pch=19,col=4,cex=cex)			
-		points(xval.CIs$nsp.means,col="green",pch=19,cex=cex)
+		points(1:K + jitter, xval.CIs$nsp.means,col="green",pch=19,cex=cex)
 	return(invisible("plotted"))
 }
 

### 2017-08-03 writeup/figs/populus_figs.R

@@ -1,20 +1,22 @@
 source("~/Dropbox/conStruct/sims/cross_validation/summarize.xvals.R")
 
-plot.cluster.curves <- function(data.block, conStruct.results, cluster.cols=NULL){
+plot.cluster.curves <- function(data.block, conStruct.results, cluster.cols=NULL,sample.cols=NULL,add=FALSE){
 	if(is.null(cluster.cols)){
 		cluster.cols <- c("blue", "red", "green", "yellow", "purple","orange", "lightblue", "darkgreen", "lightblue", "gray")
 	}
-    order.mat <- order(data.block$geoDist)
-    y.range <- range(c(
-    			unlist(lapply(1:data.block$K,
-    							function(k){
-							        conStruct.results$MAP$cluster.params[[k]]$cluster.cov
-				})) + conStruct.results$MAP$gamma, 
-				data.block$obsCov))
-    plot(data.block$geoDist[upper.tri(data.block$obsCov, diag = TRUE)], 
-        data.block$obsCov[upper.tri(data.block$obsCov, diag = TRUE)], 
-        xlim = range(data.block$geoDist), ylim = y.range,
-        xlab="",ylab="",pch=19,col=adjustcolor(1,0.7))
+	if(add==FALSE){
+		order.mat <- order(data.block$geoDist)
+	    y.range <- range(c(
+	    			unlist(lapply(1:data.block$K,
+	    							function(k){
+								        conStruct.results$MAP$cluster.params[[k]]$cluster.cov
+					})) + conStruct.results$MAP$gamma, 
+					data.block$obsCov))
+	    plot(data.block$geoDist[upper.tri(data.block$obsCov, diag = TRUE)], 
+	        data.block$obsCov[upper.tri(data.block$obsCov, diag = TRUE)], 
+	        xlim = range(data.block$geoDist), ylim = y.range,
+	        xlab="",ylab="",pch=19,col=adjustcolor(1,0.7))
+    }
     lapply(1:data.block$K, function(k) {
              lines(data.block$geoDist[order.mat],
              		conStruct.results$MAP$gamma + 
@@ -42,10 +44,10 @@ xval.CIs <- get.xval.CIs(x.vals.std,K)
[…]

### 2017-08-06 writeup/conStruct.tex
G's appendix operation

@@
 \paragraph{Drift} 
-We imagine that the allele frequencies at each locus in each sample are the sum of 3 components: 
-the ancestral allele frequency $\epsilon$ shared by all samples,
-the deviation from that ancestral mean in the $k$th population, $\Delta^{(k)}$,
-which is shared by all samples with 100\% ancestry in that population,
-the deviation specific to the $i$th sample, $\Delta^{(i)}$,
+We first provide a simple model of allele frequencies within a
+cluster. Imagine a sample $i$ that draws all of its ancestry from cluster $k$.
+We imagine that the allele frequency ($F$) at locus $\ell$ can be
+written as a sum
+\begin{equation}
+F_{i,\ell} = \epsilon_{\ell} + \Delta^{(k)}_{\ell} +
+\Delta^{(k,i)}_{\ell} + \Delta^{(i)}_{\ell}
+\label{drift_terms_no_admix}
+\end{equation}
+the first two terms are the ancestral allele frequency
+$\epsilon_\ell$ shared by all samples and the deviation from that
+ancestral mean due to drift in the ancestral population of $k$th cluster, $\Delta^{(k)}$,
+which is shared by all samples within the cluster. The third term is
+the deviation of sample $i$ away from the cluster mean due to the
+spatial process of drift and migration within the cluster,
+$\Delta^{(k,i)}$, the final term is the deviation specific to the $i$th sample, $\Delta^{(i)}$,
 which captures drift not shared by all samples at the population level
-(i.e., subpopulation-specific drift due to, e.g., inbreeding).
-
-If all samples drew all of their ancestry from any one of the $K$ clusters,
-the  allele frequency in the $i$th sample at the $\ell$th locus would be given by:
-\begin{equation}
-F_{i,\ell} = \epsilon_{\ell} + \Delta^{(k)}_{\ell} + \Delta^{(i)}_{\ell}
-\label{drift_terms_no_admix}
-\end{equation}
+(i.e., subpopulation-specific drift due to, e.g., inbreeding). We will
+assume that these four deviations are all uncorrelated with each other.
+
+If we have two populations $i$ and $j$ drawn from cluster $k$ their
+covariance across loci will be 
+\begin{equation}
+\text{Var}(\epsilon) +  \text{Var}\left( \Delta^{(k)} \right) +
+\text{Covar}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell}) + \delta_{i,j} \text{Var}\left( \Delta^{(i)} \right)\ \\
+\end{equation}
+
+%We rewrite this as 
+%\begin{equation}
+%\gamma +\mu^{(k)} + \delta_{i,j} \eta_{i}
+%\end{equation}
+%We set $\gamma = \text{Var}(\epsilon)$, the variance of allele
+%frequencies due to drift
+%in the ancestor of cluster $k$  $\mu^{(k)}=\text{Var}\left(
+ % \Delta^{(k)} \right)$. We parameterize the spatial covariance of
+%population $i$ and $j$ as
+%\begin{equation}
+%\gamma+  \text{Var}\left( \Delta^{(k)} \right) +
+%\text{Covar}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell}) + \delta_{i,j} \text{Var}\left( \Delta^{(i)} \right)\ \\
+%\end{equation}
+
+%Under our non-spatial model we set
+%$\text{Covar}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell}) =0$, such
+%that
+
+%If all samples drew all of their ancestry from any one of the $K$ clusters,
+%the  allele frequency in the $i$th sample at the $\ell$th locus would be given by:
 
@@
 To accommodate admixture between clusters, 
-we can model allele frequencies within samples as linear combinations of the population-specific 
+we model allele frequencies within samples as linear combinations of the population-specific 
 drift terms across the clusters from which they draw ancestry.
-To do so, we introduce a term $w^{(k)}_{i}$ that describes the admixture proportion of sample $i$ in cluster $k$,
-which can be interpreted as the probability that an allele sampled in the $i$th sample came from the $k$th cluster.
+To do so, we introduce a term $w^{(k)}_{i}$ that describes the
+admixture proportion of sample $i$ in cluster $k$ ($\sum\limits_{K}w^{(k)}_{i} = 1$),
+which can be interpreted as the proportion of the genome in the $i$th
+sample came from the $k$th cluster (or the probability that an allele
+at a locus is drawn from $k$).
 The allele frequency in the $i$th sample at the $\ell$th locus can therefore be written as:
-%
-\begin{equation}
-F_{i,\ell} = \epsilon_{\ell} + \sum\limits_{K} \left( w^{(k)}_{i} \Delta^{(k)}_{\ell} \right) + \Delta^{(i)}_{\ell}	
+\begin{equation}
+F_{i,\ell} = \epsilon_{\ell} + \sum\limits_{K} w^{(k)}_{i}\left( 
+  \Delta^{(k)}_{\ell} + \Delta^{(k,i)}_{\ell}\right) + \Delta^{(i)}_{\ell}	
 \label{drift_terms_admix}
 \end{equation}
-%
-where
-%
-\begin{equation}
-\sum\limits_{K}w^{(k)}_{i} = 1
-\end{equation}
-
-\subsection{The Normal approximation to drift}
-Genetic drift is serial binomial sampling, 
-but over short timescales, and for intermediate ancestral allele frequencies, 
-we can reasonably model the allele frequencies at each locus across populations as multivariate normal.
-This will be a good approximation if the amount of elapsed drift in each sample is relatively small,
-as would be expected if time since divergence from a common ancestral population is short,
-or effective population sizes are large.
-
-Beginning with the simplest case of two clusters ($K=2$),
-with all samples drawing 100\% of their ancestry from either one or the other cluster, 
-we can model allele frequencies at locus $\ell$ as:
-%
-\begin{equation}
-F_{\ell} \sim MVN\left(\mu = \epsilon_{\ell} + \iota^{(k)}\Delta_{\ell}^{(k)}, \Sigma = \Omega	\right)
-\label{MVN_mean}
-\end{equation}
-%
-where $\iota^{(k)}$ is an indicator variable that equals 1 for samples with 100\% membership in cluster $k$
-and 0 otherwise.
-In Eqn \eqref{MVN_mean}, all entries of the covariance $\Omega$ have value 0 except the diagonals,
-which describe sample-specific drift, and have the following value:
-%
-\begin{equation}
-\Omega_{i,i} = \text{Var}\left( \Delta^{(i)} \right)
-\label{MVN_mean_diag}
-\end{equation}
-
-%
-
-\subsection{Describing relatedness as a covariance}
-However, our data can consist of many loci, 
-so it may be computationally impractical to model 
-the ancestral, cluster, and sample components of each sampled allele frequency.
-Instead, we can describe the covariance induced between samples due to the sharing of the components.
-All samples share the same ancestral mean at all loci, 
-which leads to a global covariance equal to the variance of the ancestral frequencies across loci.
-Likewise, samples that are in the same cluster will have a covariance equal to the variance 
-of the cluster-specific deviates from the ancestral frequencies across loci.
-
-In Eqn \ref{MVN_mean} above, we described the relatedness between samples
-due to the shared ancestral allele frequency and shared population-level drift
-via the mean of the multivariate normal.
-We can describe this relatedness as a covariance as follows:
-
-\begin{equation}
-	\Omega_{i,j} = 
-	\begin{cases}
-	      \text{Var}(\epsilon) +  \text{Var}\left( \Delta^{(k)} \right) + \text{Var}\left( \Delta^{(i)} \right), & \text{if}\ i=j\ \\
-      	      \text{Var}(\epsilon) +  \text{Var}\left( \Delta^{(k)} \right), & \text{if}\ \iota^{(k)}_i = \iota^{(k)}_j\ \text{and}\ i \neq j \\
-	      \text{Var}(\epsilon), & \text{otherwise}
-	\end{cases}
-\label{MVN_covar}
-\end{equation}
-
-Generalizing this result to any number of clusters and incorporating admixture as in Eqn \ref{drift_terms_admix}, 
-we can write the form of the covariance of samples that are admixed between $K$ clusters as:
-
-\begin{equation}
-\Omega_{i,j} = \text{Var}(\epsilon) + \sum\limits_K \left(	w^{(k)}_iw^{(k)}_j \text{Var}(\Delta^{(k)}) 	\right) + 
-\delta_{i=j} \text{Var}(\Delta^{(i)})
-\label{admixed_discrete_covariance}
-\end{equation}
-
-where $\delta_{i=j}$ is an indicator variable that equals 1 when $i$ is equal to $j$ and 0 otherwise,
-as in Eqn \ref{cross_cluster_covariance}.
-
-\subsection{Incorporating spatial differentiation}
-Equation \ref{admixed_discrete_covariance} describes a model in which 
-samples can be continuously admixed between a set of $K$ discrete clusters.
-In this model, any pair of samples with 100\% ancestry in a cluster 
-have exactly the same covariance with each other (namely $\text{Var}(\epsilon) + \text{Var}(\Delta^{(k)})$).
-However, we also wish to model continuous decay of covariance with spatial separation between samples.
-That is, we expect (and want our model to reflect) that samples within the same cluster 
-will have higher covariance if they are sampled closer together than if they are sampled farther apart.
-To describe this spatial pattern, we build in a spatial component to our covariance model.
-
-Specifically, we write that the covariance between a pair of samples $i$ and $j$
-that draw all their ancestry from a single cluster $k$ decays exponentially as a function 
-of the distance between their sampling locations as follows:
-\begin{equation}
-G^{(k)}_{i,j} = \alpha^{(k)}_0 \times \left(\text{exp} \left(  -(\alpha^{(k)}_D D_{i,j})^{\alpha^{(k)}_2}\right) \right) + \mu^{(k)}	
-\label{spatial_cov}
-\end{equation}
-where the $\alpha$ parameters control: 
-the sill of the covariance ($\alpha_0$), 
-the rate at which covariance decays with distance $D$ ($\alpha_D$),
-and the shape of that decay ($\alpha_2$)
-in the $k$th cluster.
-The parameter $\mu^{(k)}$ describes the covariance shared by all samples in the $k$th cluster;
-it and the spatial covariance function $G$ together
- describe the quantity $\text{Var}(\Delta^{(k)})$ from Eqn \eqref{admixed_discrete_covariance}.
-
-Tying these ideas together, we can then construct a covariance that describes
-\begin{enumerate}
-\item continuous decay of covariance within a cluster
-\item a discrete amount of covariance shared within a cluster
-\item continuous admixture between clusters.
-\end{enumerate}
-
-\begin{equation}
-\Omega_{i,j} = \text{Var}(\epsilon) + \sum\limits_K \left(	w^{(k)}_iw^{(k)}_j G^{(k)}_{i,j}(\theta) 	\right) +
+Then the covariance, across loci, between $i$ and $j$ is
+\begin{equation}
+\Omega_{i,j} = \text{Var}(\epsilon) + \sum\limits_K w^{(k)}_iw^{(k)}_j
+\left(
+  \text{Var}\left( \Delta^{(k)} \right) +
+\text{Covar}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell}) 	\right) +
 \delta_{i=j} \text{Var}(\Delta_i)
@@
 \end{equation}
+
+Under our non-spatial model we assume that
+$\text{Covar}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell})=0$ so that
+the only additional convariance between $i$ and $j$ is due to the
+drift in the ancestral population of our cluster (the variance of
+which is $\mu_k$). Under our spatial
+model we assume that the spatial covariance in allele frequencies
+between $i$ and $j$ decays as a function of the geographic distance
+between the pair $D_{i,j}$
+\begin{equation}
+\text{Covar}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell}) = \alpha^{(k)}_0 \times \left(\text{exp} \left(  -(\alpha^{(k)}_D D_{i,j})^{\alpha^{(k)}_2}\right) \right)	
+\end{equation}
+We note that this form is chosen for its flexibility and not because
+it matches any explicit population-genetic model of isolation by
+distance. 
 %
-where $G^{(k)}_{i,j}(\theta)$ denotes the dependence of the spatial covariance between samples $i$ and $j$ 
-within a cluster on other quantities $\theta$: 
-specifically, the parameters $\vec{\alpha}^{(k)}$ and $\mu^{(k)}$, which are specific to the $k$th cluster, 
-as well as on the observed quantity $D_{i,j}$, the pairwise distance between samples $i$ and $j$.
-
-\section{Likelihood}
-If the allele frequency data are well approximated by a Gaussian, 
+%where
+%
+%\begin{equation}
+
+%\end{equation}
+
+% \subsection{The Normal approximation to drift}
+% Genetic drift is serial binomial sampling, 
+% but over short timescales, and for intermediate ancestral allele frequencies, 
+% we can reasonably model the allele frequencies at each locus across populations as multivariate normal.
+% This will be a good approximation if the amount of elapsed drift in each sample is relatively small,
+% as would be expected if time since divergence from a common ancestral population is short,
+% or effective population sizes are large.
+
+% Beginning with the simplest case of two clusters ($K=2$),
+% with all samples drawing 100\% of their ancestry from either one or the other cluster, 
+% we can model allele frequencies at locus $\ell$ as:
+% %
+% \begin{equation}
+% F_{\ell} \sim MVN\left(\mu = \epsilon_{\ell} + \iota^{(k)}\Delta_{\ell}^{(k)}, \Sigma = \Omega	\right)
+% \label{MVN_mean}
+% \end{equation}
+% %
+% where $\iota^{(k)}$ is an indicator variable that equals 1 for samples with 100\% membership in cluster $k$
+% and 0 otherwise.
+% In Eqn \eqref{MVN_mean}, all entries of the covariance $\Omega$ have value 0 except the diagonals,
+% which describe sample-specific drift, and have the following value:
+% %
+% \begin{equation}
+% \Omega_{i,i} = \text{Var}\left( \Delta^{(i)} \right)
+% \label{MVN_mean_diag}
+% \end{equation}
+
+% %
+
+% \subsection{Describing relatedness as a covariance}
+% However, our data can consist of many loci, 
+% so it may be computationally impractical to model 
+% the ancestral, cluster, and sample components of each sampled allele frequency.
+% Instead, we can describe the covariance induced between samples due to the sharing of the components.
+% All samples share the same ancestral mean at all loci, 
+% which leads to a global covariance equal to the variance of the ancestral frequencies across loci.
+% Likewise, samples that are in the same cluster will have a covariance equal to the variance 
+% of the cluster-specific deviates from the ancestral frequencies across loci.
+
+% In Eqn \ref{MVN_mean} above, we described the relatedness between samples
+% due to the shared ancestral allele frequency and shared population-level drift
+% via the mean of the multivariate normal.
+% We can describe this relatedness as a covariance as follows:
+
+% \begin{equation}
+% 	\Omega_{i,j} = 
+% 	\begin{cases}
+% 	      \text{Var}(\epsilon) +  \text{Var}\left( \Delta^{(k)} \right) + \text{Var}\left( \Delta^{(i)} \right), & \text{if}\ i=j\ \\
+%       	      \text{Var}(\epsilon) +  \text{Var}\left( \Delta^{(k)} \right), & \text{if}\ \iota^{(k)}_i = \iota^{(k)}_j\ \text{and}\ i \neq j \\
+% 	      \text{Var}(\epsilon), & \text{otherwise}
+% 	\end{cases}
+% \label{MVN_covar}
+% \end{equation}
+
+% Generalizing this result to any number of clusters and incorporating admixture as in Eqn \ref{drift_terms_admix}, 
+% we can write the form of the covariance of samples that are admixed between $K$ clusters as:
+
+% \begin{equation}
+% \Omega_{i,j} = \text{Var}(\epsilon) + \sum\limits_K \left(	w^{(k)}_iw^{(k)}_j \text{Var}(\Delta^{(k)}) 	\right) + 
+% \delta_{i=j} \text{Var}(\Delta^{(i)})
+% \label{admixed_discrete_covariance}
+% \end{equation}
+
+% where $\delta_{i=j}$ is an indicator variable that equals 1 when $i$ is equal to $j$ and 0 otherwise,
+% as in Eqn \ref{cross_cluster_covariance}.
+
+% \subsection{Incorporating spatial differentiation}
+% Equation \ref{admixed_discrete_covariance} describes a model in which 
+% samples can be continuously admixed between a set of $K$ discrete clusters.
+% In this model, any pair of samples with 100\% ancestry in a cluster 
+% have exactly the same covariance with each other (namely $\text{Var}(\epsilon) + \text{Var}(\Delta^{(k)})$).
+% However, we also wish to model continuous decay of covariance with spatial separation between samples.
+% That is, we expect (and want our model to reflect) that samples within the same cluster 
+% will have higher covariance if they are sampled closer together than if they are sampled farther apart.
+% To describe this spatial pattern, we build in a spatial component to our covariance model.
+
+% Specifically, we write that the covariance between a pair of samples $i$ and $j$
+% that draw all their ancestry from a single cluster $k$ decays exponentially as a function 
+% of the distance between their sampling locations as follows:
+% \begin{equation}
+% G^{(k)}_{i,j} = \alpha^{(k)}_0 \times \left(\text{exp} \left(  -(\alpha^{(k)}_D D_{i,j})^{\alpha^{(k)}_2}\right) \right) + \mu^{(k)}	
+% \label{spatial_cov}
+% \end{equation}
+% where the $\alpha$ parameters control: 
+% the sill of the covariance ($\alpha_0$), 
+% the rate at which covariance decays with distance $D$ ($\alpha_D$),
+% and the shape of that decay ($\alpha_2$)
+% in the $k$th cluster.
+% The parameter $\mu^{(k)}$ describes the covariance shared by all samples in the $k$th cluster;
+% it and the spatial covariance function $G$ together
+%  describe the quantity $\text{Var}(\Delta^{(k)})$ from Eqn \eqref{admixed_discrete_covariance}.
+
+% Tying these ideas together, we can then construct a covariance that describes
+% \begin{enumerate}
+% \item continuous decay of covariance within a cluster
+% \item a discrete amount of covariance shared within a cluster
+% \item continuous admixture between clusters.
+% \end{enumerate}
+
+% \begin{equation}
+% \Omega_{i,j} = \text{Var}(\epsilon) + \sum\limits_K \left(	w^{(k)}_iw^{(k)}_j G^{(k)}_{i,j}(\theta) 	\right) +
+% \delta_{i=j} \text{Var}(\Delta_i)
+% \label{admixed_spatial_cov}
+% \end{equation}
+% %
+% where $G^{(k)}_{i,j}(\theta)$ denotes the dependence of the spatial covariance between samples $i$ and $j$ 
+% within a cluster on other quantities $\theta$: 
+% specifically, the parameters $\vec{\alpha}^{(k)}$ and $\mu^{(k)}$, which are specific to the $k$th cluster, 
+% as well as on the observed quantity $D_{i,j}$, the pairwise distance between samples $i$ and $j$.
+
+\paragraph{Likelihood}
+If our allele frequency deviations are well approximated by a Gaussian, 
 their sample covariance is a sufficient statistic,

### 2017-08-07 writeup/conStruct.tex
minor edits

--- 2017-08-03 writeup/conStruct.tex
@@
 
-\gc{
 \paragraph{Relationship of this model to non-spatial structure models}
 
-A nice feature of our approach is that our model contains a non-spatial assignment model as a special case. 
-Setting $\alpha_0^{(k)}$ to zero for all $k$, we obtain a non-spatial model where
-each cluster has its own allele frequency at each SNP, and individuals
-draw a proportion of their ancestry from each cluster. This model is
-strongly conceptually similar to that of STRUCTURE (and related
-approaches ADMIXTURE), the only difference being that our likelihood
-assumes that allele frequencies are normally distributed around their
-expectations, while the standard assignment methods assume that the
-error is binomial distributed (see XXXX). Therefore, to compare the fit
-of
-
-} 
-\gb{support for the spatial model over the nonspatial model at a given value of $K$ 
-indicates that isolation-by-distance is likely a feature of the data at that level.}
+A nice feature of our approach is that the model described in Eqn \refeq{cross_cluster_covariance} 
+contains a non-spatial assignment model as a special case 
+(see \ref{model_app} for a more in-depth discussion). 
+Setting $\alpha_0^{(k)}$ to zero for all $k$, 
+we obtain a non-spatial model in which each cluster has its own allele frequency at each SNP, 
+and individuals draw a proportion of their ancestry from each cluster. 
+This model is strongly conceptually similar to that of STRUCTURE \citep{STRUCTURE} 
+and related models \citep[e.g][]{ADMIXTURE}; 
+the only difference is that our likelihood assumes that allele frequencies are normally distributed 
+around their expectations, 
+while the standard assignment methods assume that the error is binomial distributed (see XXXX).  
+Therefore, we can evaluate the fit of the different models 
+--- spatial vs. nonspatial, across different values of $K$ --- 
+by comparing their performance in a common framework.
 
 \paragraph{Choice of cluster number and cross-validation}
-There are a number of reasons, both statistical and biological, why
-there is no true (or right) number of clusters for real datasets. 
-%There are a number of reasons, both statistical and biological, 
-%why choosing a single, ``best" number of clusters to describe the data may not be desirable.
-%, including ambiguity in what exactly it means for any given value of $K$ to be ``best."
-We discuss these reasons further in the Discussion below. In practice
-it is useful to explore the implications of clustering at a variety of
-different levels (i.e. across multiple $K$) for biological
-intepretation and downstream analyses. 
+There are a number of reasons, both statistical and biological, 
+why there is no true (or right) number of clusters for real datasets. 
+We discuss these reasons further in the Discussion below. 
+In practice, it is useful to explore the implications of clustering at a variety of different levels (
+i.e. across multiple $K$) for biological interpretation and downstream analyses. 
 However, we also recognize that in many circumstances it is 
 useful to know the number of clusters with highest statistical
-support in the (current) dataset.  %, for example to focus su 
-%However, we also recognize that in many circumstances 
-i%t is either useful or convenient 
-%to present only a single set of results from a clustering analysis with a particular number of clusters.
-%ideally a number large enough to describe meaningful partitions in the data, 
-%but not so large as to be difficult to parse visually.
+support in the (current) dataset.
 Toward that end, we have implemented a method for 
-statistically comparing \texttt{conStruct} results across different values of $K$.
-
-Several approaches have been used as model choice criteria for the number of
-clusters in \gc{population genetic} data, including: 
+statistically comparing \texttt{conStruct} results across different values of $K$ 
+and between the spatial and non-spatial models.
+
+Several approaches have been used as model choice criteria 
+for the number of clusters in population genetic data, including: 
 the use of information theoretic criteria \citep[e.g.][]{ADMIXTURE}, such as AIC or BIC;
@@
 See \cite{verity_nichols2016} for a discussion of these approaches and comparison
-between several methods. %A comprehensive review of these approaches and their various merits is beyond the scope of this paper,
-%but
+between several methods.
 
 \gc{IS PREDICTION accuracy the right phrase?}
-We choose to take a cross-validation approach.  We estimate the
-posterior on the parameters of the model using a ``training" partition
-of the data, in practice the training set is a random subset of $90\%$
-of the loci. We then calculate the likelihood of the
-``testing" subset, the remaining $10\%$ of the loci, averaged over the
-posterior of the parameters. We repeat this procedure a number of
-times, take an average of the relative log-likelihood over the runs at
-sumarize each $K$. 
-and selecting the model with the significantly best predictive accuracy
+We choose to take a cross-validation approach.
+We estimate the posterior on the parameters of the model using a ``training" partition of the data 
+(in practice the training set is a random subset of $90\%$ of the loci). 
+We then calculate the likelihood of the ``testing" subset (the remaining $10\%$ of the loci), 
+averaged over the posterior of the parameters. 
+We repeat this procedure a number of times, 
+take an average of the relative log-likelihoods over the runs at each $K$, 
+and select the model with the significantly best predictive accuracy
 (see Section \ref{Xvalidation} for more on our cross-validation procedure).
-This approach rewards models with larger values of $K$, which increase
-the likelihood of the training partition, as they fit the data better,
-but the improvement in the likelihood will plateau (or peak) as above
-a certain $K$ we are fitting noise specific to the training data and
-so not fitting the testing set any better, 
-%up to the point that cluster membership is invoked to model noise specific to the training data, 
-%at which point predictive accuracy decreases, 
-%and models with additional clusters are penalized.
+This approach rewards models with larger values of $K$, 
+which increase the likelihood of the training partition, 
+as they fit the data better,
+but the improvement in the likelihood will plateau (or peak),
+ as above a certain $K$ the model is fitting noise specific to the training data 
+ and therefore not fitting the testing set any better.
 
@@
 where they do, they frequently hybridize (ref).
-In addition, many \textit{Populus} species display patterns of isolation by distance, 
-making the system well-suited to an analysis designed to tease apart 
-genetic structure partitioned discretely across population (or species) clusters,
-and continuously with geographic distance.
 
@@
 and \textit{Populus trichocarpa}, the balsam poplar.
-These species have a broad zone of overlap and are known to hybridize \citep{geraldes_etal_2014}.
+These species have a broad zone of overlap and are hypothesized to hybridize \citep{geraldes_etal_2014}.
+Both species are sampled over a large geographic region, 
+making the system well-suited to an analysis designed to tease apart 
+proportion of genetic structure that is partitioned discretely across population (or species) clusters,
+as well as continuously with geographic distance.
 
@@
 and showed a strong pattern of bias in allelic dropout 
-(the majority of missing data were from drainages with only \textit{Populus balsamifera}).
+(the majority of missing data were from drainages with only \textit{Populus balsamifera} individuals).
 To ameliorate some of the problems that arise when there is a strong bias in which data are missing, 
@@
 for invaluable comments on the method and manuscript, 
-as well as Quentin Cronk, who helped interpret the \textit{Populus} analyses, 
-and the attendees at the 2017 SSE Meeting in Portland, OR, whose votes determined the name of the method.
+as well as Quentin Cronk, who provided input on the \textit{Populus} analyses, 
+and the attendees at the 2017 SSE Meeting in Portland, OR, 
+whose votes determined the name of the method.
 This work was supported in part by 
@@
 
-\section{Models, parameters, and priors}
+\section{Models, parameters, and priors}\label{model_app}
 \paragraph{Spatial vs. nonspatial}
-In this paper, we discuss two types of models, spatial and nonspatial.
+In this paper, we discuss two types of models, spatial and nonspatial, 
+each of which can be implemented with different numbers of clusters.
 The spatial model is parameterized as in Eqn \ref{admixed_spatial_cov},
 and the nonspatial model is described in Eqn \ref{admixed_discrete_covariance}.
+The nonspatial model is a special case of the spatial model, 
+in which the value of all $\alpha$ parameters is set to 0.
 The nonspatial model therefore has $3K$ fewer parameters than the spatial model,
 as there are three $\alpha$ parameters that describe the continuous differentiation effect of distance per cluster,
-and they are not included in the nonspatial model.
+and they are fixed in the nonspatial model.
 
@@
 Each of these models can be run with a single cluster ($K=1$), 
-which involves a slight re-parameterization, detailed below.
+in which case the cluster-specific covariance parameter $\mu^{(k})$ 
+and the global covariance parameter $\gamma$ become redundant.
+The single-cluster model is therefore a special case of the multi-cluster model, 
+in which we set $\mu$ to zero, as described below:
 

### 2017-08-07 writeup/conStruct.tex
merged graham's appendectomy

--- 2017-08-07 writeup/conStruct.tex
@@
 \textsuperscript{a}bradburd@msu.edu; 
-\textsuperscript{c}gmcoop@ucdavis.edu;
-\textsuperscript{b}plr@uoregon.edu\\\\\
+\textsuperscript{b}gmcoop@ucdavis.edu;
+\textsuperscript{c}plr@uoregon.edu\\\\\
 
@@
 
-\paragraph{Relationship of this model to non-spatial structure models}
+\paragraph{Relationship of this model to nonspatial structure models}
 
 A nice feature of our approach is that the model described in Eqn \refeq{cross_cluster_covariance} 
-contains a non-spatial assignment model as a special case 
+contains a nonspatial assignment model as a special case 
 (see \ref{model_app} for a more in-depth discussion). 
 Setting $\alpha_0^{(k)}$ to zero for all $k$, 
-we obtain a non-spatial model in which each cluster has its own allele frequency at each SNP, 
+we obtain a nonspatial model in which each cluster has its own allele frequency at each SNP, 
 and individuals draw a proportion of their ancestry from each cluster. 
@@
 statistically comparing \texttt{conStruct} results across different values of $K$ 
-and between the spatial and non-spatial models.
+and between the spatial and nonspatial models.
 
@@
 \section{Model rationale}
-\subsection{Drift, admixture, and allele frequencies}
+\subsection{Drift, admixture, and space}
 \paragraph{Drift} 
-We imagine that the allele frequencies at each locus in each sample are the sum of 3 components: 
-the ancestral allele frequency $\epsilon$ shared by all samples,
-the deviation from that ancestral mean in the $k$th population, $\Delta^{(k)}$,
-which is shared by all samples with 100\% ancestry in that population,
-the deviation specific to the $i$th sample, $\Delta^{(i)}$,
+We first provide a simple model of allele frequencies within a cluster. 
+Imagine a sample $i$ that draws all of its ancestry from cluster $k$.
+The allele frequency in sample $i$ at locus $\ell$, $F_{i,\ell}$ can be
+written as the sum
+\begin{equation}
+F_{i,\ell} = \epsilon_{\ell} + \Delta^{(k)}_{\ell} +
+\Delta^{(k,i)}_{\ell} + \Delta^{(i)}_{\ell}
+\label{drift_terms_no_admix}
+\end{equation}
+The first term is the ancestral allele frequency $\epsilon_\ell$ shared by all samples; 
+the second is the deviation from that ancestral frequency 
+due to drift in the ancestral population of the $k$th cluster,
+which is shared by all samples within the cluster. 
+The third term is the deviation of the $i$th sample away from the $k$th cluster mean 
+due to the spatial process of drift and migration within the cluster.
+The final term is the deviation specific to the $i$th sample,
 which captures drift not shared by all samples at the population level
 (i.e., subpopulation-specific drift due to, e.g., inbreeding). 
+We will assume that these four deviations are all uncorrelated with each other.
 
-If all samples drew all of their ancestry from any one of the $K$ clusters,
-the  allele frequency in the $i$th sample at the $\ell$th locus would be given by:
+If we have two samples $i$ and $j$ drawn from cluster $k$, 
+their covariance across loci will be 
 \begin{equation}
-F_{i,\ell} = \epsilon_{\ell} + \Delta^{(k)}_{\ell} + \Delta^{(i)}_{\ell}
-\label{drift_terms_no_admix}
+\text{Var}(\epsilon) +  \text{Var}\left( \Delta^{(k)} \right) +
+\text{Cov}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell}) + \delta_{i=j} \text{Var}\left( \Delta^{(i)} \right)\ \\
 \end{equation}
+where the quantity $\delta_{i=j}$ is an indicator variable that equals 1 when $i$ is equal to $j$ and 0 otherwise, 
+as in Eqn. \eqref{cross_cluster_covariance}.
 
@@
 To accommodate admixture between clusters, 
-we can model allele frequencies within samples as linear combinations of the population-specific 
+we model allele frequencies within samples as linear combinations of the population-specific 
 drift terms across the clusters from which they draw ancestry.
-To do so, we introduce a term $w^{(k)}_{i}$ that describes the admixture proportion of sample $i$ in cluster $k$,
-which can be interpreted as the probability that an allele sampled in the $i$th sample came from the $k$th cluster.
+To do so, we introduce a term $w^{(k)}_{i}$ that describes the
+admixture proportion of sample $i$ in cluster $k$ ($\sum_{K}w^{(k)}_{i} = 1$),
+which can be interpreted as the proportion of the genome in the $i$th
+sample that came from the $k$th cluster 
+(or the probability that an allele at a locus is drawn from $k$).
 The allele frequency in the $i$th sample at the $\ell$th locus can therefore be written as:
-%
 \begin{equation}
-F_{i,\ell} = \epsilon_{\ell} + \sum\limits_{K} \left( w^{(k)}_{i} \Delta^{(k)}_{\ell} \right) + \Delta^{(i)}_{\ell}	
+F_{i,\ell} = \epsilon_{\ell} + \sum\limits_{K} w^{(k)}_{i}\left( 
+  \Delta^{(k)}_{\ell} + \Delta^{(k,i)}_{\ell}\right) + \Delta^{(i)}_{\ell}	
 \label{drift_terms_admix}
 \end{equation}
-%
-where
-%
+and the covariance between $i$ and $j$ across loci is
 \begin{equation}
-\sum\limits_{K}w^{(k)}_{i} = 1
-\end{equation}
-
-\subsection{The Normal approximation to drift}
-Genetic drift is serial binomial sampling, 
-but over short timescales, and for intermediate ancestral allele frequencies, 
-we can reasonably model the allele frequencies at each locus across populations as multivariate normal.
-This will be a good approximation if the amount of elapsed drift in each sample is relatively small,
-as would be expected if time since divergence from a common ancestral population is short,
-or effective population sizes are large.
-
-Beginning with the simplest case of two clusters ($K=2$),
-with all samples drawing 100\% of their ancestry from either one or the other cluster, 
-we can model allele frequencies at locus $\ell$ as:
-%
-\begin{equation}
-F_{\ell} \sim MVN\left(\mu = \epsilon_{\ell} + \iota^{(k)}\Delta_{\ell}^{(k)}, \Sigma = \Omega	\right)
-\label{MVN_mean}
-\end{equation}
-%
-where $\iota^{(k)}$ is an indicator variable that equals 1 for samples with 100\% membership in cluster $k$
-and 0 otherwise.
-In Eqn \eqref{MVN_mean}, all entries of the covariance $\Omega$ have value 0 except the diagonals,
-which describe sample-specific drift, and have the following value:
-%
-\begin{equation}
-\Omega_{i,i} = \text{Var}\left( \Delta^{(i)} \right)
-\label{MVN_mean_diag}
-\end{equation}
-
-%
-
-\subsection{Describing relatedness as a covariance}
-However, our data can consist of many loci, 
-so it may be computationally impractical to model 
-the ancestral, cluster, and sample components of each sampled allele frequency.
-Instead, we can describe the covariance induced between samples due to the sharing of the components.
-All samples share the same ancestral mean at all loci, 
-which leads to a global covariance equal to the variance of the ancestral frequencies across loci.
-Likewise, samples that are in the same cluster will have a covariance equal to the variance 
-of the cluster-specific deviates from the ancestral frequencies across loci.
-
-In Eqn \ref{MVN_mean} above, we described the relatedness between samples
-due to the shared ancestral allele frequency and shared population-level drift
-via the mean of the multivariate normal.
-We can describe this relatedness as a covariance as follows:
-
-\begin{equation}
-	\Omega_{i,j} = 
-	\begin{cases}
-	      \text{Var}(\epsilon) +  \text{Var}\left( \Delta^{(k)} \right) + \text{Var}\left( \Delta^{(i)} \right), & \text{if}\ i=j\ \\
-      	      \text{Var}(\epsilon) +  \text{Var}\left( \Delta^{(k)} \right), & \text{if}\ \iota^{(k)}_i = \iota^{(k)}_j\ \text{and}\ i \neq j \\
-	      \text{Var}(\epsilon), & \text{otherwise}
-	\end{cases}
-\label{MVN_covar}
-\end{equation}
-
-Generalizing this result to any number of clusters and incorporating admixture as in Eqn \ref{drift_terms_admix}, 
-we can write the form of the covariance of samples that are admixed between $K$ clusters as:
-
-\begin{equation}
-\Omega_{i,j} = \text{Var}(\epsilon) + \sum\limits_K \left(	w^{(k)}_iw^{(k)}_j \text{Var}(\Delta^{(k)}) 	\right) + 
-\delta_{i=j} \text{Var}(\Delta^{(i)})
-\label{admixed_discrete_covariance}
-\end{equation}
-
-where $\delta_{i=j}$ is an indicator variable that equals 1 when $i$ is equal to $j$ and 0 otherwise,
-as in Eqn \ref{cross_cluster_covariance}.
-
-\subsection{Incorporating spatial differentiation}
-Equation \ref{admixed_discrete_covariance} describes a model in which 
-samples can be continuously admixed between a set of $K$ discrete clusters.
-In this model, any pair of samples with 100\% ancestry in a cluster 
-have exactly the same covariance with each other (namely $\text{Var}(\epsilon) + \text{Var}(\Delta^{(k)})$).
-However, we also wish to model continuous decay of covariance with spatial separation between samples.
-That is, we expect (and want our model to reflect) that samples within the same cluster 
-will have higher covariance if they are sampled closer together than if they are sampled farther apart.
-To describe this spatial pattern, we build in a spatial component to our covariance model.
-
-Specifically, we write that the covariance between a pair of samples $i$ and $j$
-that draw all their ancestry from a single cluster $k$ decays exponentially as a function 
-of the distance between their sampling locations as follows:
-\begin{equation}
-G^{(k)}_{i,j} = \alpha^{(k)}_0 \times \left(\text{exp} \left(  -(\alpha^{(k)}_D D_{i,j})^{\alpha^{(k)}_2}\right) \right) + \mu^{(k)}	
-\label{spatial_cov}
-\end{equation}
-where the $\alpha$ parameters control: 
-the sill of the covariance ($\alpha_0$), 
-the rate at which covariance decays with distance $D$ ($\alpha_D$),
-and the shape of that decay ($\alpha_2$)
-in the $k$th cluster.
-The parameter $\mu^{(k)}$ describes the covariance shared by all samples in the $k$th cluster;
-it and the spatial covariance function $G$ together
- describe the quantity $\text{Var}(\Delta^{(k)})$ from Eqn \eqref{admixed_discrete_covariance}.
-
-Tying these ideas together, we can then construct a covariance that describes
-\begin{enumerate}
-\item continuous decay of covariance within a cluster
-\item a discrete amount of covariance shared within a cluster
-\item continuous admixture between clusters.
-\end{enumerate}
-
-\begin{equation}
-\Omega_{i,j} = \text{Var}(\epsilon) + \sum\limits_K \left(	w^{(k)}_iw^{(k)}_j G^{(k)}_{i,j}(\theta) 	\right) +
+\Omega_{i,j} = \text{Var}(\epsilon) + \sum\limits_K w^{(k)}_iw^{(k)}_j
+\left(
+  \text{Var}\left( \Delta^{(k)} \right) +
+\text{Cov}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell}) 	\right) +
 \delta_{i=j} \text{Var}(\Delta_i)
@@
 \end{equation}
-%
-where $G^{(k)}_{i,j}(\theta)$ denotes the dependence of the spatial covariance between samples $i$ and $j$ 
-within a cluster on other quantities $\theta$: 
-specifically, the parameters $\vec{\alpha}^{(k)}$ and $\mu^{(k)}$, which are specific to the $k$th cluster, 
-as well as on the observed quantity $D_{i,j}$, the pairwise distance between samples $i$ and $j$.
 
-\section{Likelihood}
-If the allele frequency data are well approximated by a Gaussian, 
+\paragraph{Space}
+Under our nonspatial model, 
+we assume that $\text{Cov}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell})=0$,
+so that the only additional covariance between $i$ and $j$ 
+(above that induced by a shared ancestral frequency at each locus)
+is due to the drift in the ancestral population of their cluster 
+(the variance of which is $\mu_k$). 
+Under our spatial model we assume that some of the covariance in allele frequencies
+between $i$ and $j$ decays as a function of the geographic distance
+between the pair, $D_{i,j}$,
+so that 
+\begin{equation}
+\text{Cov}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell}) = \alpha^{(k)}_0 \times \left(\text{exp} \left(  -(\alpha^{(k)}_D D_{i,j})^{\alpha^{(k)}_2}\right) \right)
+\end{equation}
+We note that this form is chosen for its flexibility, 
+and not because it matches any explicit population genetic model of isolation by distance.
+
+\paragraph{Likelihood}
+If our allele frequency deviations are well approximated by a Gaussian, 
 their sample covariance is a sufficient statistic,
@@
 \begin{align}
-\widehat{\Omega} &= \frac{1}{L}FF^T \notag \\
-\widehat{\Omega} &\sim \mathcal{W}\left( L\Omega, L	\right)
+&\widehat{\Omega}_{i,j} = 
+	\begin{cases}
+		\frac{1}{L-1}
+		\sum\limits_{\ell=1}^{L}{\left[ 	\left(f_{i,\ell} - \bar{f}_{i} \right)
+		 	\left(f_{j,\ell} - \bar{f}_{j} \right)\right] \notag
+			} & \text{for } i \neq j \\
+		\bar{f}_i (1-\bar{f}_i) & \text{for } i = j \\
+	\end{cases}\\
+&\widehat{\Omega} \sim \mathcal{W}\left( L\Omega, L	\right)
 \label{wishart}
@@
 Each of these models can be run with a single cluster ($K=1$), 
-in which case the cluster-specific covariance parameter $\mu^{(k})$ 
+in which case the cluster-specific covariance parameter $\mu^{(k)}$ 
 and the global covariance parameter $\gamma$ become redundant.

### 2017-08-10 writeup/conStruct.tex
added to discussion

@@
 \newcommand{\secref}[1]{``\nameref{#1}''}
+\newcommand{\tri}{\textit{trichocarpa}}
+\newcommand{\bals}{\textit{balsamifera}}
 
@@
 and therefore not fitting the testing set any better.
+At any value of $K$, support for the spatial model over the nonspatial model 
+means that isolation by distance is likely a feature of the data.
+\gb{something about how we choose the best model?}
 
@@
 A number of results are immediately clear from these tests with simulated data.
-First, across all values of K and for all simulations, 
-the spatial model is preferred over the nonspatial model
-(this is not surprising, as the data were simulated under a spatial model).
+First, \texttt{conStruct} performs well on all simulated scenarios.
+Across all values of K and for all simulations, 
+the spatial model is correctly preferred over the nonspatial model, 
+and it infers the true admixture proportions with high accuracy, tight precision, and good coverage (Fig ).
 Second, the predictive accuracy of the nonspatial model increases with the number of clusters in the model, 
@@
 with gradients of admixture across more and more discrete clusters.
-This behavior of the nonspatial model is highlighted in the analyses of data simulated under $K=1$ (Fig \ref{simK1_pies} and \ref{simK1_nsp_pies}):
+This behavior of the nonspatial model is highlighted in 
+the analyses of data simulated under $K=1$ (Fig \ref{simK1_pies} and \ref{simK1_nsp_pies}):
 as $K$ in the model increases, 
@@
 those estimated under the true value of $K$ for each simulation.
-We parse \texttt{conStruct}'s performance on these simulation tests further in the Discussion.
+
+\gb{could talk about proportion of relatedness explained by each layer here.
+e.g., Another way to visualize the difference in the behavior of the spatial and nonspatial models 
+is to summarize the proportion of relatedness between samples explained by each cluster...}
 
@@
 Both species are sampled over a large geographic region, 
-making the system well-suited to an analysis designed to tease apart 
-proportion of genetic structure that is partitioned discretely across population (or species) clusters,
+making the system well-suited to an analysis designed to determine  
+the proportion of genetic structure that is partitioned discretely across population (or species) clusters,
 as well as continuously with geographic distance.
-
-We address this question using a dataset generated and originally published by \cite{geraldes_etal_2014}.
+We had the following questions for the system: 
+\begin{enumerate}
+\item Given the known history of hybridization between this pair of poplar species, 
+is there support for two discrete clusters, 
+rather than a single cline of ancestry across their overlap zone?
+\item If there is evidence for discrete population structuring, 
+does it fall solely along species boundaries, 
+or is there sub-structuring between different populations of the same species?
+\item Is there support for a pattern of isolation by distance within
+the inferred clusters, and, if so, 
+does it differ between different clusters?
+\end{enumerate}
+
+We address these questions using a dataset generated and originally published by \cite{geraldes_etal_2014}.
 The published data consist of 434 individuals sampled from 35 drainages 
@@
 	the spatial \texttt{conStruct} model run with $K=2$ through 7.
-	Line colors are consistent with Fig \ref{populus_sp_pies}.
+	Line colors are consistent with cluster colors in Fig \ref{populus_sp_pies}.
 	Points are colored by the species they are a covariance between:
 	black on black points are sample covariances between populations of \textit{Populus balsamifera};
-	green on black points are sample covariances between \textit{P. balsamifera} and \textit{P. trichocarpa};
-	green on green points are sample covariances between \textit{P. trichocarpa} and \textit{P. trichocarpa}.
+	green on black points are sample covariances between \bals{} and \tri{};
+	green on green points are sample covariances between \tri{} and \tri{}.
     }\label{populus_sp_clst_covs}
@@
 	the nonspatial \texttt{conStruct} model run with $K=2$ through 7.
-	Line colors are consistent with Fig \ref{populus_nsp_pies}.
+	Line colors are consistent with cluster colors in Fig \ref{populus_nsp_pies}.
 	Points are colored by the species they are a covariance between:
 	black on black points are sample covariances between populations of \textit{Populus balsamifera};
-	green on black points are sample covariances between \textit{P. balsamifera} and \textit{P. trichocarpa};
-	green on green points are sample covariances between \textit{P. trichocarpa} and \textit{P. trichocarpa}.
+	green on black points are sample covariances between \bals{} and \tri{};
+	green on green points are sample covariances between \tri{} and \tri{}.
     }\label{populus_nsp_clst_covs}
@@
 
+Although admixture was inferred between \textit{P. balsamifera} and \tri{}
+across all values of $K>1$, and in both the spatial and nonspatial models,
+there is nonetheless strong evidence for discrete population structuring, 
+rather than a single, continuous cline of ancestry.
+At all values of $K>1$, discrete population structure was mostly partitioned along species lines; 
+at values of $K$ above 2, further discrete substructure was inferred within the \textit{P. trichocarpa} samples, 
+while \bals{} samples displayed no substructure up to $K=7$.
+At all values of $K$, the spatial model was preferred over the nonspatial model, 
+indicating that the data exhibit isolation by distance.
+However, most of this signal of IBD seems to be coming from the \textit{P. trichocarpa} samples;
+in the spatial model, the cluster-specific covariance curves for clusters that are predominantly \tri{}
+have sharper slopes than that of the \bals{} cluster (Fig \ref{populus_sp_clst_covs}).
+The most strongly supported model, given the current dataset, was XXXX.
+
 \subsubsection*{Black bears}
@@
 \section*{Discussion}
-
-\paragraph{Performance on simulated data}
-\paragraph{Empirical results}
+Isolation by distance is a pattern that is ubiquitous in nature:
+it can arise as long as mating opportunities are distance-limited, 
+so we might reasonably expect to observe it in more species than not.
+This common pattern can flummox model-based clustering methods, 
+which attempt to ascribe the observed, continuous population differentiation 
+to unnecessary additional discrete clusters.
+The inference of these spurious clusters can have implications for 
+downstream analysis and interpretation of genetic data, 
+including conservation management decisions and 
+identification of the genetic basis of adaptation or disease traits.
+
+In this paper, we have presented a statistical framework, \texttt{conStruct}, for simultaneously 
+modeling continuous and discrete patterns of population structure.
+By employing the sensible ``null hypothesis" 
+that relatedness ought to decay with distance, even within a population, 
+we avoid erroneously ascribing population differentiation to discrete population clusters.
+We also present a cross-validation approach for comparing between models, 
+although we stress the limits of this or any model selection procedure for genetic clustering.
+
+Our tests on simulated data show that the method performs well; 
+we are able to accurately infer both the admixture proportions used to simulate the data 
+as well as accurately pick the simulating model as the best model using our cross-validation procedure.
+Empirical applications of \texttt{conStruct} to a sample of North American poplars 
+and a sample of North American black bears yield reasonable results, 
+largely in agreement with previous studies in both systems (Geraldes, Puckett).
+
+The proposed method combines the utility of model-based clustering algorithms 
+with a biologically realistic model of isolation by distance.
+We anticipate that \texttt{conStruct} will be useful for identifying populations 
+and determining samples' ancestry in and across them, 
+especially when the populations exhibit spatial patterns of relatedness.
+
+\paragraph{Comparison to nonspatial model-based clustering}
+
 \paragraph{Choosing the ``best" number of clusters}
+Although we recognize the utility of choosing a single, ``best'' value of $K$, 
+and using only that analysis to communicate results, 
+we want to emphasize that there are both statistical and biological reasons 
+why there is no true best value of $K$.
+From a statistical perspective, the choice of best $K$ is always relative to the data in hand.
+This means both that sampling different individuals or loci 
+from the same populations may yield a different ``best" $K$, 
+but also that the support for larger values of $K$ is likely to increase with increasing amounts of data.
+In the limit of infinite data, the best value of $K$ 
+will be the number of individuals included in the dataset.
+From a biological perspective, 
+it is important to stress that patterns of relatedness between individuals and populations 
+are shaped by complex spatial and hierarchical processes.
+All individuals within a species (and indeed, all individuals across all species), 
+are related to one another in some way, 
+and summarizing that relatedness with a single value of $K$ may be reductive or misleading.
+We therefore encourage users to perform analyses across different values of $K$ and 
+observe which clusters split out at what levels 
+(effectively taking successively shallower cross-sections of the population phylogeny),
+and also to take the results of the proposed cross-validation procedure with a large grain of salt.
+
 \paragraph{Implications for management and conservation}

### 2017-08-11 writeup/conStruct.tex
edits up to Relationship

@@
 \newcommand{\bals}{\textit{balsamifera}}
+
+\include{macros}
 
@@
 These units are useful for defining a scope of study in research on selection or local adaptation \citep{},
-or on identifying either locally adapted alleles (human skin color, ralph \& coop, gartersnakes) or alleles involved in disease (cystic fibrosis).
+or for identifying either locally adapted alleles (human skin color, ralph \& coop, gartersnakes) or alleles involved in disease (cystic fibrosis).
 Delineating populations can also be useful for informing conservation priorities \citep{}.
+However, if underlying population structure is continuous, this may entail a loss of power.
 
@@
 Here, we will focus on model-based clustering, the most widely used class of approaches for population delineation.
-These methods each perform best under different specific demographic histories of samples being analyzed,
-but all can give misleading results when applied to data that show a continuous pattern of isolation by distance \citep{Wright 1938, 1940, Wright1943}.
+Like the method we present here, these are based on a statistical model,
+but do not perform inference directly with a demographic model.
+Each of these methods perform best in particular situations,
+but all can give misleading results when applied to data that show a continuous pattern of isolation by distance \citep{Wright 1938, 1940, Wright1943, novembre_interpreting_2008}.
 
@@
 \citep[e.g.][]{SerrePaabo2004,rosenberg2005clines}.
+As a further complication, if local genetic drift is strong,
+continous models are expected to give rise to spatially patchy patterns
+\citep{SLFV,durrett_crabgrass}.
 
@@
 
-What is desired, then, is a model-based clustering method that incorporates a spatial element,
-and specifically one that uses a model of isolation by distance as a biologically sensible null hypothesis 
-for the process structuring observed genetic variation.
-With an explicit spatial null, discrete population structure need only be invoked when genetic differentiation 
-in the data is above and beyond that expected given some estimated background rate at which divergence 
-increases with geographic separation.
+To address this, we set out to develop
+a model-based clustering method that when possible uses isolation by distance 
+to explain observed genetic variation.
+With an explicit spatial component, discrete population structure need only be invoked when genetic differentiation 
+in the data is deviates significantly from that expected given geographic separation.
 
@@
 
+%%%%%%%%%%%%%%%%%
 \section*{Methods}
+
 \paragraph{Data}
-The statistical framework of our approach is conceptually similar to \cite{Wasser2004} and \cite{BEDASSLE}, and \cite{spacemix}.
-The genetic data modeled consist of allele frequencies $F$ at $L$ unlinked, bi-allelic single nucleotide polymorphisms (SNPs) genotyped across $N$ samples.
-The sample frequency at locus $\ell$ in sample $n$, $f_{n,\ell}$, is calculated by first arbitrarily choosing an allele segregating 
-at locus $\ell$ to count, 
-then dividing the total number of observations of that counted allele by the total number of chromosomes genotyped at that locus
+The statistical framework of our approach is conceptually similar to that in \cite{Wasser2004}, \cite{BEDASSLE}, and \cite{spacemix},
+although we use a somewhat different summary statistic than in this previous work.
+The model works with allele frequencies at $L$ unlinked, bi-allelic single nucleotide polymorphisms (SNPs) genotyped across $N$ samples.
+Each ``sample'' may be a collection of diploids, or frequencies estimated from pooled sequencing.
+We write $N_i$ for the number of chromosomes sequenced in the $i^\text{th}$ sample.
+The \emph{sample frequency} at locus $\ell$ in sample $n$, denoted $f_{n,\ell}$, 
+is calculated by first arbitrarily choosing one of the observed alleles at locus $\ell$ to count, 
+then dividing the number of observations of that counted allele by the total number of chromosomes genotyped at that locus
 in sample $n$.
-We then calculate the sample covariance in allele frequencies, $\widehat{\Omega}_{i,j}$ between samples $i$ and $j$.
-However, for the diagonal of this sample covariance (the variances within each sample),
-we wish to describe the mean of the variance in allele frequencies between haplotypes in a sample, 
-rather than the variance of the mean frequencies across haplotypes in a sample.
-Therefore, we calculate our sample covariance as follows:
-
+The choice of allele does not affect subsequent calculations, and so may be arbitrary.
+\plr{How about calling what we calculate the ``allelic covariance''?  ``individual covariance''?}
+We then calculate the \emph{sample covariance} between samples $i$ and $j$, denoted $\widehat{\Omega}_{i,j}$,
+as the expected covariance of distinct individual alleles chosen from each of the two samples at a random locus,
+choosing a random allele to code as `1' (and the other as `0').
+Concretely, this is
 \begin{equation}
 \widehat{\Omega}_{i,j} = 
-	\begin{cases}
-		\frac{1}{L-1}
-		\sum\limits_{\ell=1}^{L}{\left[ 	\left(f_{i,\ell} - \bar{f}_{i} \right)
-		 	\left(f_{j,\ell} - \bar{f}_{j} \right)\right]
-			} & \text{for } i \neq j \\
-		\bar{f}_i (1-\bar{f}_i) & \text{for } i = j \\
-	\end{cases}
+    \frac{1}{L} \sum_{\ell=1}^L \left( (f_{i,\ell}-1/2) (f_{j,\ell}-1/2) - \delta_{i,j} f_{i,\ell}^2 / (N_i-1) \right),
 \label{sample_covariance}
 \end{equation}
+where $\delta_{i,j}=1$ if $i=j$ and is 0 otherwise.
+Although we describe this as a covariance between individually drawn alleles,
+$\widehat{\Omega}_{i,j}$ is in fact also the covariance between the allele frequencies
+of a randomly chosen allele in samples $i$ and $j$, as long as $i \neq j$.
+The diagonal (where $i=j$) does not have this interpretation.
+
+In addition to these differences on the diagonal,
+this differs from the usual ``genetic covariance'' \cite{mcvean_covariance}
+in that (a) we do not subtract locus means,
+and so to make the statistic invariant under choices of reference allele, 
+(b) we randomly choose alleles to count.
+To see how to arrive at equation \eqref{sample_covariance},
+pick a random locus and
+let $A$ and $B$ be randomly drawn alleles at that locus from populations $i$ and $j$ respectively.
+If $i=j$ then these are chosen \emph{without} replacement.
+Suppose these are each coded as `0' or `1` (where `0' denotes a reference allele),
+but we randomly ``flip'' this coding, so that we let $X=A$ and $Y=B$ with probability 1/2,
+but otherwise we let $X=1-A$ and $Y=1-B$.
+We then let $\widehat{\Omega}_{i,j} = \cov[X,Y]$ --
+sampling without replacement makes the statistic insensitive to sample size,
+and random allele flipping makes it independent of the choice of reference allele.
+By conditioning on the flip,
+and using the fact that $\E[X] = \E[Y] = 1/2$,
+equation \eqref{sample_covariance} comes from the calculation that
+$\cov[X,Y] = \E[XY] - 1/4 = (1/2)\E[AB + (1-A)(1-B)] - 1/4 = \E[(A-1/2)(B-1/2)]$.
+The factor of $\delta_{i,j}/(N_i-1)$ comes from the fact that if $i=j$,
+then $\E[AB]$ is the probability that both alleles drawn without replacement are reference,
+which is $f_i^2 n/(n-1)$.
+As noted in \cite{desi_thesis},
+this covariance has a close relationship to the pairwise genetic distance:
+if $D_{i,j}$ is the mean density of sites at which random samples from $i$ and $j$ differ at a randomly chosen locus,
+then $\Sigma_{i,j} = (1 - 2 D_{i,j})/4$.
+Therefore, unlike genetic covariance,
+this covariance is affected in shape by singleton sites,
+so it may be advisable to filter these prior to analysis
+if these are likely to contain a large percentage of errors.
 
 \paragraph{Continuous and discrete differentiation}
-We wish to describe the observed patterns of genetic variation as a combination of 
-discrete population clusters,
-within which genetic variation is continuously distributed, 
-and between which sampled individuals or populations can be admixed.
-Each of these discrete clusters is modeled as a spatial process in which
-allele frequencies covary between geographically close locations,
-but this covariance is allowed to decay with geographic distance.
+Clustering approaches to describing genetic variation
+are useful because population history can often be meaningfully described on a coarse scale
+by interactions between discrete ``populations''
+whose relationships are delimited by patterns of glaciation,
+large-scale migration,
+mountain ranges, and the like.
+Here we add a spatial component within each such discrete historical component,
+which we refer to as a set of ``layers'' which overlay the modern map.
+We imagine each layer as a geographically distributed population,
+and that each sample is composed of a mixture of contributions from each of these layers.
+These layers thus take the place of ``clusters'' or ``populations'' in clustering methods,
+but we do not adopt these terms as they connote a lack of geographical separation within each.
+
+Within each of these layers,
+allele frequencies have positive covariance at geographically close locations,
+but this covariance is allowed to decay towards zero as geographic distance increases.
 This pattern of spatial decay reflects how migration between neighboring demes  
 homogenizes allele frequency changes that arise locally due to drift, 
-but is less effective at homogenizing between geographic distant demes,
+but this mixing less effectively homogenizes geographic distant demes,
 resulting in a continuous pattern of isolation by distance within each cluster.
-Multiple population cluster processes can co-occur in space, 
-leading to discrete jumps in allele frequencies across many loci over small geographic distances,
-and the genotyped samples can be admixed between different clusters.
-
-To model the continuous decay of allele frequency covariance with geographic distance within a cluster, 
-we use a flexible powered-exponential function;
-the within-cluster covariance between samples $i$ and $j$ is given by:
-
-\begin{equation}
-G^{(k)}_{i,j} = \alpha^{(k)}_0 \left( \text{exp} \left( -(\alpha^{(k)}_D D_{i,j}) ^ {\alpha^{(k)}_2}	\right) \right) + \mu^{(k)}
+There is a fixed amount of covariance between layers, irrespective of spatial location.
+Within each layer allele frequencies are expected to change gradually with distance,
+but observed frequencies can change abruptly at many loci 
+if the proportions by which individuals' ancestry derive from each layer 
+(the ``admixture'' coefficients) 
+do so as well.
+
+To allow flexibility in the form of the decay of allele frequency covariance with geographic distance within each cluster, 
+we define the covariance within cluster $k$ between samples $i$ and $j$ to be:
+\begin{equation}
+G^{(k)}_{i,j} 
+    = 
+    \alpha^{(k)}_0 \left( \text{exp} \left( -(\alpha^{(k)}_D D_{i,j}) ^ {\alpha^{(k)}_2}	\right) \right) + \mu^{(k)}
 \label{within_cluster_covariance}
 \end{equation}
-
-where $G_{i,j}$ is the covariance function between samples $i$ and $j$ within a cluster,
-and the superscript $(k)$ denotes that this function is specific to the $k$th cluster.
+where the superscript $(k)$ denotes parameters specific to the $k$th cluster.
 The quantity $D_{i,j}$ is the observed geographic distance between samples $i$ and $j$,
 the $\alpha^{(k)}$ parameters control the shape of the decay of covariance with distance in the cluster,
-and $\mu$ is a parameter that describes the background allele frequency covariance within the cluster. 
-If two samples draw 100\% of their ancestry from cluster $k$ then, 
-if they are geographically very close ($D_{i,j}=0$) 
+and $\mu^{(k)}$ is a parameter that describes the background allele frequency covariance within the cluster. 
+If two samples draw 100\% of their ancestry from cluster $k$, then their covariance under the model is $G^{(k)}_{i,j}$;
+if they are furthermore geographically very close ($D_{i,j}=0$) 
 they will have allele frequency covariance $\alpha^{(k)}_0 +  \mu^{(k)}$.
 If the geographic distance between them is very large, 
-their covariance will decay to the background level $\mu^{(k)}$ within the cluster.
-The shared drift parameter $\mu^{(k)}$ can be interpreted as the branch length 
+their covariance will be equal to to the background level $\mu^{(k)}$ within the cluster.
+The ``shared drift'' parameter $\mu^{(k)}$ can be interpreted as the branch length 
 connecting the $k$th population to the population ancestral to all modeled
 clusters \gc{(see XXX)}. \gb{is ``XXX" TreeMix? or one of the $f$stat papers?  or is it a figure?} 
-
-Genotyped samples can draw their ancestry from multiple different clusters,
-representing these samples being admixed between the clusters.
-The admixture proportion of the $i$th sample in the $k$th cluster, $w^{(k)}_i$,
-gives the probability that an allele in sample $i$ was derived from
-cluster $k$ ($\sum_K w^{(k)}_i =1$).
+\plr{an F-stat paper; also reference Ben Peters review}
+
+We then allow samples to draw their ancestry from more than one cluster.
+The ``admixture'' proportion of the $i$th sample in the $k$th cluster, denoted $w^{(k)}_i$,
+gives the proportion of alleles from sample $i$ that derive from
+cluster $k$ (and so $\sum_K w^{(k)}_i =1$).
 
 We can then describe the covariance between samples $i$ and $j$ across all clusters, $\Omega_{i,j}$,
-by summing their within-cluster spatial covariances ($G_{i,j}^{(k)}$ in cluster $k$) across all $K$ clusters
-and weighting by the product of those samples' admixture proportions in each cluster
+by summing their within-cluster spatial covariances ($G_{i,j}^{(k)}$ in cluster $k$) across all $K$ clusters,
+weighted by the relevant admixture proportions.
 \begin{equation}
@@
 \end{equation}
+In this equation, $w^{(k)}_i w^{(k)}_j$ is the proportion of alleles that \emph{both}
+sample $i$ and sample $j$ have inherited from cluster $k$.
 
@@
 which generates a base-line covariance.
-In the second term, $\delta_{i,j}$ is an indicator variable that takes a value of 1 when $i$ equals $j$ and 0 otherwise. 
-and $\eta_i$ gives the variance specific to sample $i$, 
+In the final term, $\delta_{i,j}$ is an indicator variable that takes a value of 1 when $i$ equals $j$ and 0 otherwise,
+and $\eta_i$ adds variance specific to sample $i$.
 This term on the diagonal of the parametric covariance matrix captures processes shaping variance within the sampled deme, 
@@
 \paragraph{Likelihood and inference}
-We assume that the allele frequencies at each locus ($\widehat{F}_{\ell}$) 
-are independent among loci and multivariate normally distributed across populations, 
-so that their sample covariance $\widehat{\Omega}$ 
-is  Wishart distributed with degrees of freedom equal to $L$, 
+If allele frequencies at each locus
+were independent between loci and multivariate normally distributed across populations, 
+then their sample covariance $\widehat{\Omega}$ 
+would be Wishart distributed with degrees of freedom equal to $L$, 
 the number of loci genotyped.
+We use this as a convenient approximation to the true distribution described above,
+and so define the likelihood of the sample standardized allele frequency covariance to be
+\begin{equation}
+P(\widehat{\Omega} \; | \; \Omega) = \mathcal{W} \left( L\widehat{\Omega} \; | \; \Omega,L\right) .
+\end{equation}
 Linkage disequilibrium (LD) between loci will decrease the effective number of degrees of freedom, 
-which we discuss further below.
-The likelihood of the sample standardized allele frequency covariance is therefore given by
-\begin{equation}
-P(\widehat{\Omega} \; | \; \Omega) = \mathcal{W} \left( L\widehat{\Omega} \; | \; \Omega,L\right)
-\end{equation}
+a complication we discuss further below.
 
@@
 Acknowledging the dependence of the parametric covariance matrix $\Omega$ on its constituent parameters
-$w,\alpha,\mu,\eta,\gamma$ and on observed quantity $D$ 
+$w,\alpha,\mu,\eta,\gamma$ and on the (observed) geographic distances $D$ 
 with the notation $\Omega(w,\alpha,\mu,\eta,\gamma,D)$,
-we denote the posterior probability of the parameters as:
-
+we denote the posterior probability density of the parameters as:
 \begin{equation}
@@
 P\left(\widehat{\Omega} \; | \; \Omega(w,\alpha,\mu,\eta,\gamma,D) \right)
-P(w)P(\alpha)P(\mu)P(\eta)P(\gamma)
-\end{equation}
-
-The priors, $P(w),P(\alpha),P(\mu),P(\eta),P(\gamma)$, are detailed in Table \ref{tab:param_prior_tab},
-and the constant of proportionality is the normalization constant.
-We use a Hamiltonian Monte Carlo sampling algorithm implemented in the statistical language STAN
-\citep{stan, NUTS, stan_lib, rstan} to estimate the posterior distribution on the parameters.
-We also present an R package \citep{R} called \texttt{conStruct} that functions as a wrapper 
+P(w)P(\alpha)P(\mu)P(\eta)P(\gamma) ,
+\end{equation}
+where $P(w)$, $P(\alpha)$, $P(\mu)$, $P(\eta)$, and $P(\gamma)$, are prior distributions.
+All parameters are given (half-)Gaussian priors except for $\alpha_2$, which is uniform on $(0,2)$ and $w$, which is Dirichlet
+(see Table \ref{tab:param_prior_tab} for specifics).
+Parameters are independent between clusters.
+We use Hamiltonian Monte Carlo as implemented in STAN \citep{stan, NUTS, stan_lib, rstan} 
+to estimate the posterior distribution on the parameters.
+We also provide an R package \citep{R} called \texttt{conStruct} that functions as a wrapper 
 around this inference machinery.

### 2017-08-11 writeup/conStruct.tex
added to discussion

@@
 Species in the genus regularly co-occur and, 
-where they do, they frequently hybridize (ref).
+where they do, they frequently hybridize \citep{eckenwalder_1984, cronk_2005}.
 
@@
 and \textit{Populus trichocarpa}, the balsam poplar.
-These species have a broad zone of overlap and are hypothesized to hybridize \citep{geraldes_etal_2014}.
+These species have a broad zone of overlap and are hypothesized to hybridize \citep{geraldes_etal_2014,suarez-gonzalez_etal_2016}.
 Both species are sampled over a large geographic region, 
+and show spatial patterns of genetic and phenotypic variation \citep{Slavov_etal_2012, McKown_etal_2014},
 making the system well-suited to an analysis designed to determine  
@@
 resulting in just over 20,200 loci retained for analysis.  
-We then analyzed these data using both the spatial and nonspatial \texttt{conStruct} models, 
+We then analyzed these data, grouped by drainage, using both the spatial and nonspatial \texttt{conStruct} models, 
 running on $K = 1$ through 7,
@@
 while \bals{} samples displayed no substructure up to $K=7$.
+This lack of substructure within the sampled \bals{} populations is in agreement with results from 
+\citet{Keller_etal_2010}, who found that \bals{} in the region is characterized by relatively low diversity, 
+and probably represents a recent, postglacial expansion.
+
+A consistent split between clusters within \tri{} fell along the ``no-cottonwood belt," 
+a region along the central coast of British Columbia in which black cottonwood is absent. 
+The no-cottonwood belt is hypothesized to divide the species' distribution 
+into northern and southern groups, which, in a provenance test, 
+were experimentally shown to display different phenotypes 
+\citep[pathogen resitsance][]{Xie_etal_2009,Xie_etal_2012}.  
+At higher values of $K$, drainages at the southern tip of the \tri{} sampling 
+begin to split out into their own clusters.  
+This may potentially be the result of introgression from Southern \textit{Populus} species, 
+such as \textit{P. angustifolia} or \textit{fremontii} \citep{Zhou2012,Geraldes_etal_2014}
+
 At all values of $K$, the spatial model was preferred over the nonspatial model, 
@@
 \paragraph{Implications for management and conservation}
+If isolation by distance is common, 
+then a likely result of applying \texttt{conStruct} to existing data is that 
+populations previously identified using nonspatial clustering methods 
+will be collapsed into each other.  
+This ``lumping" might better reflect biological reality, 
+but may also have implications for management decisions and conservation policy, 
+both of which are often predicated on the identification of discrete ``management units" (MUs) 
+identified using genetic data.
+
+It is therefore important to emphasize that individuals sampled from the same \texttt{conStruct} cluster 
+may be quite genetically diverged from one another, 
+and that a \texttt{conStruct} cluster may still contain multiple distinct MUs.  
+Alternatively, the collapse of multiple, previously identified MUs into a single \texttt{conStruct} cluster 
+may reflect a biological reality that these populations are currently 
+(or were recently) exchanging migrants, 
+and thus might emphasize the importance of maintaining habitat corridors (ref) 
+between demes, or of implementing an integrated conservation plan across multiple demes within a cluster.
+
+\paragraph{Caveats and considerations}
+There are a few important caveats to consider in the interpretation of \texttt{conStruct} results. 
+First, we have modeled allele frequency covariance within a cluster as a spatial process.
+Although there is flexibility built into the model about the shape of that covariance, 
+inference may be misleading if the sampling geography departs radically from the way 
+the sampled organisms experience (or have experienced) their landscape.
+For example, if we were to run a \texttt{conStruct} analysis using geographic distances between 
+sampled individuals of greenish warblers (Irwin ref) or \textit{Ensatina} salamanders (Wake ref)
+--- two canonical examples of rings species --- 
+we might get misleading results.
+This is because distance between locations on either side of the species' distributions
+(across the Tibetan plateau and the Central Valley, respectively) 
+is not representative of the path traversed in the coalescent of a pair of alleles sampled at those locations.
+
+A second caveat is that, in some instances, 
+membership in the same cluster may not mean that samples are particularly related.
+If covariance within a cluster decays sharply with distance, 
+and the cluster-specific relatedness parameter $\mu^{(k)}$ is low, 
+individuals at a far spatial remove may be in the same cluster but have very low pairwise relatedness.
+It is possible that this is happening in Fig \ref{populus_sp_pies}\subref{populus_sp_pies:K3}. 
+At $K=3$, the southernmost populations of \textit{P. trichocarpa} are in the green cluster, 
+whose other neighbors are to the north, with an intervening group of populations in the red cluster, 
+and at $K=4$, those southernmost samples split out and become their own yellow cluster.
+Again, we encourage users to run analyses across multiple values of $K$, 
+and to examine the spatial covariance functions within clusters when interpreting results.
+\gb{missing data?}
+
 \paragraph{Extensions and future directions}
+There are several ways in which the model described in this paper might be extended or improved.  
+For example, we currently assume that all layers within a model are equally unrelated 
+(a star population phylogeny, although the branches can have different lengths thanks to the $\mu^{(k)}$ parameter), 
+similar to the F-model in (REF the right STRUCTURE paper).
+However, we could extend the existing model by implementing 
+a relatedness structure between the clusters by, for example, 
+estimating a population phylogeny between them \citep[e.g.][]{treemix}.
+
+A second exciting direction is to model relatedness within a cluster as a spatiotemporal process, 
+in which covariance decays both with distance in space and in time.  
+As the number of genotyped historical or ancient samples increases, 
+it is becoming possible to ask whether there is genetic continuity at a point in space across time, 
+or whether populations are being replaced (yamnaya paper, slatkin and racimo, nielsen 2016, schraiber 2017).
+However, we expect allele frequencies to change through time in a population, even without replacement, 
+simply due to drift.
+Therefore, a natural way to test for population replacement is to estimate the rates 
+at which relatedness within a cluster decays with time in the same way we do in the current model with space, 
+in which case discrete population structure in space is comparable to population replacement in time.
 

### 2017-08-13 writeup/conStruct.tex
added missing refs

@@
 by delineating discrete populations from which individuals or alleles are sampled \citep{wright1949genetical}.
-These units are useful for defining a scope of study in research on selection or local adaptation \citep{},
+These units are useful for defining a scope of study in research on selection or local adaptation (ref),
 or for identifying either locally adapted alleles (human skin color, ralph \& coop, gartersnakes) or alleles involved in disease (cystic fibrosis).
-Delineating populations can also be useful for informing conservation priorities \citep{}.
+Delineating populations can also be useful for informing conservation priorities (ref).
 However, if underlying population structure is continuous, this may entail a loss of power.
@@
 including generating population phylogenies \citep{CavalliSforza1975, treemix},
-$k$-means clustering analyses performed on dimensionality-reduction approaches \citep{}, 
+$k$-means clustering analyses performed on dimensionality-reduction approaches (ref), 
 such as principal components analysis \citep{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
@@
 Each of these methods perform best in particular situations,
-but all can give misleading results when applied to data that show a continuous pattern of isolation by distance \citep{Wright 1938, 1940, Wright1943, novembre_interpreting_2008}.
+but all can give misleading results when applied to data that show a continuous pattern of isolation by distance 
+\citep{Wright1943, novembre_interpreting_2008, Frantz2009}.
 
 A pattern of isolation by distance (IBD) arises when migration in a population is spatially limited \citep{Slatkin1985},
-and has been modeled in both spatially continuous frameworks \citep{Wright1943,Malecot1948}, 
+and has been modeled in both spatially continuous frameworks \citep{Wright1943,malecot1969mathematics}, 
 in which an individual's parents are found within some radius of that individual's location,
-and discrete, ``stepping-stone'' frameworks \citep{Kimura1953},
+and discrete, ``stepping-stone'' frameworks \citep{kimura1964stepping},
 in which migration occurs only between neighboring demes.
@@
 continous models are expected to give rise to spatially patchy patterns
-\citep{SLFV,durrett_crabgrass}.
+\citep{SLFV,bramson1989crabgrass}.
 
@@
 In addition to these differences on the diagonal,
-this differs from the usual ``genetic covariance'' \cite{mcvean_covariance}
+this differs from the usual ``genetic covariance'' \cite{mcvean_genealogical_2009}
 in that (a) we do not subtract locus means,
@@
 which is $f_i^2 n/(n-1)$.
-As noted in \cite{desi_thesis},
+As noted in \cite{EEMS},
 this covariance has a close relationship to the pairwise genetic distance:
@@
 Species in the genus regularly co-occur and, 
-where they do, they frequently hybridize \citep{eckenwalder_1984, cronk_2005}.
+where they do, they frequently hybridize \citep{eckenwalder1984, Cronk2005}.
 
@@
 and \textit{Populus trichocarpa}, the balsam poplar.
-These species have a broad zone of overlap and are hypothesized to hybridize \citep{geraldes_etal_2014,suarez-gonzalez_etal_2016}.
+These species have a broad zone of overlap and are hypothesized to hybridize \citep{geraldes_etal_2014, suarezgonzalez_etal_2016}.
 Both species are sampled over a large geographic region, 
-and show spatial patterns of genetic and phenotypic variation \citep{Slavov_etal_2012, McKown_etal_2014},
+and show spatial patterns of genetic and phenotypic variation \citep{slavov_etal_2012, mckown_etal_2013},
 making the system well-suited to an analysis designed to determine  
@@
 This lack of substructure within the sampled \bals{} populations is in agreement with results from 
-\citet{Keller_etal_2010}, who found that \bals{} in the region is characterized by relatively low diversity, 
+\citet{keller_etal_2010}, who found that \bals{} in the region is characterized by relatively low diversity, 
 and probably represents a recent, postglacial expansion.
@@
 were experimentally shown to display different phenotypes 
-\citep[pathogen resitsance][]{Xie_etal_2009,Xie_etal_2012}.  
+\citep[pathogen resitsance][]{xie2009,xie2012}.  
 At higher values of $K$, drainages at the southern tip of the \tri{} sampling 
@@
 This may potentially be the result of introgression from Southern \textit{Populus} species, 
-such as \textit{P. angustifolia} or \textit{fremontii} \citep{Zhou2012,Geraldes_etal_2014}
+such as \textit{P. angustifolia} or \textit{fremontii} \citep{Zhou2012,geraldes_etal_2014}
 

### 2017-08-15 writeup/conStruct.tex
more edits

@@
 (see \ref{model_app} for a more in-depth discussion). 
-Setting $\alpha_0^{(k)}$ to zero for all $k$, 
+By setting $\alpha_0^{(k)}$ to zero for all $k$, 
 we obtain a nonspatial model in which each cluster has its own allele frequency at each SNP, 
@@
 This model is strongly conceptually similar to that of STRUCTURE \citep{STRUCTURE} 
-and related models \citep[e.g][]{ADMIXTURE}; 
+and related models \citep[e.g.][]{ADMIXTURE}; 
 the only difference is that our likelihood assumes that allele frequencies are normally distributed 
@@
 while the standard assignment methods assume that the error is binomial distributed (see XXXX).  
-Therefore, we can evaluate the fit of the different models 
+Therefore, we can compare the fit of the different models 
 --- spatial vs. nonspatial, across different values of $K$ --- 
@@
 There are a number of reasons, both statistical and biological, 
-why there is no true (or right) number of clusters for real datasets. 
-We discuss these reasons further in the Discussion below. 
-In practice, it is useful to explore the implications of clustering at a variety of different levels (
-i.e. across multiple $K$) for biological interpretation and downstream analyses. 
-However, we also recognize that in many circumstances it is 
-useful to know the number of clusters with highest statistical
-support in the (current) dataset.
+why there is no true (or right) number of clusters for real datasets,
+discussed further in the Discussion.
+However, it is still important to assess whether additional spatial layers (larger $K$)
+meaningfully model pattern in the data
+or merely explain spurious variation introduced by noise
+-- in other words, whether additional model complexity
+provides significant explanatory power.
 Toward that end, we have implemented a method for 
@@
 between several methods.
-
-\gc{IS PREDICTION accuracy the right phrase?}
-We choose to take a cross-validation approach.
-We estimate the posterior on the parameters of the model using a ``training" partition of the data 
-(in practice the training set is a random subset of $90\%$ of the loci). 
-We then calculate the likelihood of the ``testing" subset (the remaining $10\%$ of the loci), 
-averaged over the posterior of the parameters. 
-We repeat this procedure a number of times, 
-take an average of the relative log-likelihoods over the runs at each $K$, 
-and select the model with the significantly best predictive accuracy
-(see Section \ref{Xvalidation} for more on our cross-validation procedure).
-This approach rewards models with larger values of $K$, 
-which increase the likelihood of the training partition, 
-as they fit the data better,
-but the improvement in the likelihood will plateau (or peak), 
-as above a certain $K$ the model is fitting noise specific to the training data 
-and therefore not fitting the testing set any better.
+\plr{I wonder if we can just cite \cite{verity_nichols2016} rather than this laundry list here?}
+
+We use cross-validation to attack this problem.
+To do this,
+we use a ``training" partition of the data (in practice, a random 90\% subset of the loci)
+to estimate the posterior distribution of the parameters,
+and then calculate the log-likelihood of the remaining ``testing'' loci,
+averaged over the posterior.
+Prediction accuracy of a particular model (e.g., value of $K$)
+is then measured using this log-likelihood,
+averaged over a number of independent data partitions.
+The best model is judged to be the simplest one with significantly better predictive accuracy
+than others (see Section \ref{Xvalidation} for more on our cross-validation procedure).
+In general, larger values of $K$ allows the model more flexibility,
+and thus increases the likelihood of the training partition, 
+but this improvement in the likelihood will plateau (or even peak), 
+since above a certain $K$ the model only fits noise specific to the training data 
+rather than generalizable patterns.
 At any value of $K$, support for the spatial model over the nonspatial model 
 means that isolation by distance is likely a feature of the data.
+
 \gb{something about how we choose the best model?}
+\plr{how's this?}
+Cross-validation provides a valuable summary of how much explanatory power
+is added by spatial structure within each layer, and each additional layer.
+A simple $t$-test can be used to assess whether additional layers
+show statistically significantly greater predictive accuracy.
+However, we remind users that ``statistical significance does not imply real-world significance'',
+and so small but statistically significant differences between models
+should probably not be relied on too strongly.
 
@@
 We first test the method with scenarios generated using the coalescent simulator \texttt{ms} \citep{Hudson2002}.
-To generate discrete population structure, we simulate $K$ discrete clusters,
-each of which split from a common ancestor at some point in the past, $\tau_{\text{s}}$,
-then diverge for some time without migration between them.
+To generate discrete population structure, we simulate $K$ distinct populations,
+each of which split from a common ancestor $\tau_{\text{s}}$ units of coalescent time in the past,
+without subsequent migration between them.
 To generate continuous differentiation within each cluster,
-at time $\tau_{\text{e}}$ we have each of these discrete clusters instantaneously colonize a lattice of demes,
+at time $\tau_{\text{e}}$ in the past,
+each of these discrete clusters instantaneously colonizes a lattice of demes,
 for which we use a stepping stone model with symmetric migration 
 to nearest neighbors (eight neighbors, including diagonals).
-Finally, at time $\tau_{\text{a}}$ we generate a single dataset that captures both continuous and discrete patterns of population structure 
-by collapsing those $K$ discrete lattices into a single grid of demes sampled from the different clusters
-or admixed between them (see Fig \ref{sim_setup}).
-We simulated datasets under $K =$ 1 through 3, 
-sampling 10,000 unlinked loci from each of 20 haploid individuals from every deme
+Finally, at time $\tau_{\text{a}}$ in the past 
+we generate a single dataset % that captures both continuous and discrete patterns of population structure 
+by collapsing those $K$ discrete lattices into a single grid of demes that are admixed to various degrees from these different layers
+(see Fig \ref{sim_setup}).
+We simulated datasets using $K=1$, 2, and 3 layers,
+in each sampling 10,000 unlinked loci from each of 20 haploid individuals from every deme
 (for details on all simulations, see Methods and Supplementary Materials).
@@
 the spatial model is correctly preferred over the nonspatial model, 
-and it infers the true admixture proportions with high accuracy, tight precision, and good coverage (Fig ).
-Second, the predictive accuracy of the nonspatial model increases with the number of clusters in the model, 
+and it infers the true admixture proportions with high accuracy, tight precision, and good coverage 
+(Fig \ref{sim_xvals}).
+Second, the predictive accuracy of the nonspatial model increases 
+with the number of clusters in the model, 
 as it describes continuous, spatial patterns of relatedness between demes within the same cluster 
-with gradients of admixture across more and more discrete clusters.
+using gradients of admixture across more and more discrete clusters.
 This behavior of the nonspatial model is highlighted in 
@@
 at $K=4$, all four corners, and so on.
-
-In contrast, the predictive accuracy of the spatial model of \texttt{conStruct} 
+\plr{Need a concluding sentence, maybe like this?}
+While it is true that each such cluster is more similar within itself
+than it is to other clusters,
+we know that these boundaries are arbitray,
+since the data were simulated with no such boundaries.
+
+In contrast, the predictive accuracy of the spatial \texttt{conStruct} model
 either plateaus at or declines from the true value of $K$ used to simulate the data (Fig \ref{sim_xvals}, second column).
@@
 \textit{Populus trichocarpa}, the black cottonwood,
-and \textit{Populus trichocarpa}, the balsam poplar.
+and \textit{Populus balsamifera}, the balsam poplar.
 These species have a broad zone of overlap and are hypothesized to hybridize \citep{geraldes_etal_2014, suarezgonzalez_etal_2016}.
@@
 and show spatial patterns of genetic and phenotypic variation \citep{slavov_etal_2012, mckown_etal_2013},
-making the system well-suited to an analysis designed to determine  
-the proportion of genetic structure that is partitioned discretely across population (or species) clusters,
-as well as continuously with geographic distance.
+making the system well-suited to an analysis designed to describe
+how genetic structure is partitioned both discretely across population (or species) clusters
+and continuously with geographic distance.
 We had the following questions for the system: 
@@
 We then analyzed these data, grouped by drainage, using both the spatial and nonspatial \texttt{conStruct} models, 
-running on $K = 1$ through 7,
-and compared these models using the cross-validation procedure described above. 
+using $K = 1$ through 7,
+and compared these models using cross-validation.
 The results of these analyses are shown in 
-Figs \ref{populus_xvals}-\ref{populus_nsp_clst_covs}. %,populus_nsp_pies,populus_sp_clst_covs,populus_nsp_clst_covs
+Figs \ref{populus_xvals}--\ref{populus_nsp_clst_covs}. %,populus_nsp_pies,populus_sp_clst_covs,populus_nsp_clst_covs
 
@@
 
-Although admixture was inferred between \textit{P. balsamifera} and \tri{}
-across all values of $K>1$, and in both the spatial and nonspatial models,
-there is nonetheless strong evidence for discrete population structuring, 
+All models with $K>1$ assigned the majority of each species
+to distinct layers, with some populations drawing ancestry from both layers.
+This provides strong support for discrete population structure between the two species,
+with some admixture,
+% Although admixture was inferred between \textit{P. balsamifera} and \tri{}
+% across all values of $K>1$, and in both the spatial and nonspatial models,
+% there is nonetheless strong evidence for discrete population structuring, 
 rather than a single, continuous cline of ancestry.
@@
 were experimentally shown to display different phenotypes 
-\citep[pathogen resitsance][]{xie2009,xie2012}.  
-At higher values of $K$, drainages at the southern tip of the \tri{} sampling 
+\citep[pathogen resistance][]{xie2009,xie2012}.  
+At higher values of $K$, drainages at the southern tip of \tri{} sampling 
 begin to split out into their own clusters.  
+\plr{cut this if we pick lower $K$? Or add a disclaimer?}
 This may potentially be the result of introgression from Southern \textit{Populus} species, 
@@
 
-At all values of $K$, the spatial model was preferred over the nonspatial model, 
-indicating that the data exhibit isolation by distance.
-However, most of this signal of IBD seems to be coming from the \textit{P. trichocarpa} samples;
-in the spatial model, the cluster-specific covariance curves for clusters that are predominantly \tri{}
-have sharper slopes than that of the \bals{} cluster (Fig \ref{populus_sp_clst_covs}).
+At all values of $K$, the spatial model was preferred over the nonspatial model.
+However, most of this signal seems to derive from the \textit{P. trichocarpa} samples:
+covariance within predominantly \tri{} layers decays more quickly with distance
+than within the majority \bals{} layer, as seen in Fig \ref{populus_sp_clst_covs}.
 The most strongly supported model, given the current dataset, was XXXX.
@@
 
+%%%%%%%%%
 \section*{Discussion}
-Isolation by distance is a pattern that is ubiquitous in nature:
-it can arise as long as mating opportunities are distance-limited, 
-so we might reasonably expect to observe it in more species than not.
-This common pattern can flummox model-based clustering methods, 
-which attempt to ascribe the observed, continuous population differentiation 
-to unnecessary additional discrete clusters.
-The inference of these spurious clusters can have implications for 
+
+The pattern of isolation by distance is ubiquitous in nature,
+unsuprisingly, since mating opportunities are distance-limited in most species.
+However,
+this common pattern can flummox commonly used clustering methods, 
+which must use a possibly large number of discrete clusters
+to describe such continuous population differentiation .
+Inference of these spurious clusters can have implications for 
 downstream analysis and interpretation of genetic data, 
@@
 modeling continuous and discrete patterns of population structure.
-By employing the sensible ``null hypothesis" 
+By employing the sensible default assumption
 that relatedness ought to decay with distance, even within a population, 
@@
 
-Our tests on simulated data show that the method performs well; 
-we are able to accurately infer both the admixture proportions used to simulate the data 
-as well as accurately pick the simulating model as the best model using our cross-validation procedure.
+The method performs well on simulated data:
+we accurately infer the admixture proportions used to simulate the data 
+and accurately pick the simulating model as the best model using our cross-validation procedure.
 Empirical applications of \texttt{conStruct} to a sample of North American poplars 
@@
 \paragraph{Comparison to nonspatial model-based clustering}
+Above, we \plr{hopefully?} showed that 
+(a) the nonspatial conStruct model recapitulates results of 
+other, commonly-used nonspatial clustering methods,
+and 
+(b) conStruct can concisely capture spatial structure, which is nearly always expected within clusters.
+Given this, when should methods without spatial capability be used?
+One advantage these have over conStruct is speed \plr{details?}.
+The other main difference is the model of noise,
+but it is unclear what effects this may have,
+or indeed, which is closer to the truth.
 
@@
 from the same populations may yield a different ``best" $K$, 
-but also that the support for larger values of $K$ is likely to increase with increasing amounts of data.
+but also 
+-- unless the data were generated under the model itself --
+that the support for larger values of $K$ is likely to increase with increasing amounts of data
+(although perhaps only logarithmically).
 In the limit of infinite data, the best value of $K$ 
-will be the number of individuals included in the dataset.
+may be the number of individuals included in the dataset.
 From a biological perspective, 

### 2017-08-15 writeup/conStruct.tex
added more refs

--- 2017-08-13 writeup/conStruct.tex
@@
 such as principal components analysis \citep{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
-and model-based clustering approaches \citep[e.g.][]{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, FINESTRUCTURE, huelsenbeck2007inference, Corander2003,TESS}, among others.
+and model-based clustering approaches \citep[e.g.][]{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, FINESTRUCTURE, huelsenbeck2007inference, Corander2003,TESS,geneland}, among others.
 Here, we will focus on model-based clustering, the most widely used class of approaches for population delineation.
@@
 \gc{IS PREDICTION accuracy the right phrase?}
-We choose to take a cross-validation approach.
+We choose to take a cross-validation approach \citep[similar in spirit to][]{ADMIXTURE_xval}.
 We estimate the posterior on the parameters of the model using a ``training" partition of the data 

### 2017-08-15 writeup/conStruct.tex

--- 2017-08-15 writeup/conStruct.tex
@@
 between several methods.
+
+\gc{IS PREDICTION accuracy the right phrase?}
+We choose to take a cross-validation approach \citep[similar in spirit to][]{ADMIXTURE_xval}.
+We estimate the posterior on the parameters of the model using a ``training" partition of the data 
+(in practice the training set is a random subset of $90\%$ of the loci). 
+We then calculate the likelihood of the ``testing" subset (the remaining $10\%$ of the loci), 
+averaged over the posterior of the parameters. 
+We repeat this procedure a number of times, 
+take an average of the relative log-likelihoods over the runs at each $K$, 
+and select the model with the significantly best predictive accuracy
+(see Section \ref{Xvalidation} for more on our cross-validation procedure).
+This approach rewards models with larger values of $K$, 
+which increase the likelihood of the training partition, 
+as they fit the data better,
+but the improvement in the likelihood will plateau (or peak), 
+as above a certain $K$ the model is fitting noise specific to the training data 
+and therefore not fitting the testing set any better.
 \plr{I wonder if we can just cite \cite{verity_nichols2016} rather than this laundry list here?}

### 2017-08-15 writeup/conStruct.tex
changed mu to phi

@@
     = 
-    \alpha^{(k)}_0 \left( \text{exp} \left( -(\alpha^{(k)}_D D_{i,j}) ^ {\alpha^{(k)}_2}	\right) \right) + \mu^{(k)}
+    \alpha^{(k)}_0 \left( \text{exp} \left( -(\alpha^{(k)}_D D_{i,j}) ^ {\alpha^{(k)}_2}	\right) \right) + \phi^{(k)}
 \label{within_cluster_covariance}
@@
 the $\alpha^{(k)}$ parameters control the shape of the decay of covariance with distance in the cluster,
-and $\mu^{(k)}$ is a parameter that describes the background allele frequency covariance within the cluster. 
+and $\phi^{(k)}$ is a parameter that describes the background allele frequency covariance within the cluster. 
 If two samples draw 100\% of their ancestry from cluster $k$, then their covariance under the model is $G^{(k)}_{i,j}$;
 if they are furthermore geographically very close ($D_{i,j}=0$)
-they will have allele frequency covariance $\alpha^{(k)}_0 +  \mu^{(k)}$.
+they will have allele frequency covariance $\alpha^{(k)}_0 +  \phi^{(k)}$.
 If the geographic distance between them is very large, 
-their covariance will be equal to to the background level $\mu^{(k)}$ within the cluster.
-The ``shared drift'' parameter $\mu^{(k)}$ can be interpreted as the branch length 
+their covariance will be equal to to the background level $\phi^{(k)}$ within the cluster.
+The ``shared drift'' parameter $\phi^{(k)}$ can be interpreted as the branch length 
 connecting the $k$th population to the population ancestral to all modeled
@@
 Acknowledging the dependence of the parametric covariance matrix $\Omega$ on its constituent parameters
-$w,\alpha,\mu,\eta,\gamma$ and on the (observed) geographic distances $D$ 
-with the notation $\Omega(w,\alpha,\mu,\eta,\gamma,D)$,
+$w,\alpha,\phi,\eta,\gamma$ and on the (observed) geographic distances $D$ 
+with the notation $\Omega(w,\alpha,\phi,\eta,\gamma,D)$,
 we denote the posterior probability density of the parameters as:
 \begin{equation}
-P\left( w,\alpha,\mu,\eta,\gamma \;	| \; \widehat{\Omega}, L \right) \propto
-P\left(\widehat{\Omega} \; | \; \Omega(w,\alpha,\mu,\eta,\gamma,D) \right)
-P(w)P(\alpha)P(\mu)P(\eta)P(\gamma) ,
-\end{equation}
-where $P(w)$, $P(\alpha)$, $P(\mu)$, $P(\eta)$, and $P(\gamma)$, are prior distributions.
+P\left( w,\alpha,\phi,\eta,\gamma \;	| \; \widehat{\Omega}, L \right) \propto
+P\left(\widehat{\Omega} \; | \; \Omega(w,\alpha,\phi,\eta,\gamma,D) \right)
+P(w)P(\alpha)P(\phi)P(\eta)P(\gamma) ,
+\end{equation}
+where $P(w)$, $P(\alpha)$, $P(\phi)$, $P(\eta)$, and $P(\gamma)$, are prior distributions.
 All parameters are given (half-)Gaussian priors except for $\alpha_2$, which is uniform on $(0,2)$ and $w$, which is Dirichlet
@@
 If covariance within a cluster decays sharply with distance, 
-and the cluster-specific relatedness parameter $\mu^{(k)}$ is low, 
+and the cluster-specific relatedness parameter $\phi^{(k)}$ is low, 
 individuals at a far spatial remove may be in the same cluster but have very low pairwise relatedness.
@@
 For example, we currently assume that all layers within a model are equally unrelated 
-(a star population phylogeny, although the branches can have different lengths thanks to the $\mu^{(k)}$ parameter), 
+(a star population phylogeny, although the branches can have different lengths thanks to the $\phi^{(k)}$ parameter), 
 similar to the F-model in (REF the right STRUCTURE paper).
@@
 is due to the drift in the ancestral population of their cluster 
-(the variance of which is $\mu_k$). 
+(the variance of which is $\phi_k$). 
 Under our spatial model we assume that some of the covariance in allele frequencies
@@
 Each of these models can be run with a single cluster ($K=1$), 
-in which case the cluster-specific covariance parameter $\mu^{(k)}$ 
+in which case the cluster-specific covariance parameter $\phi^{(k)}$ 
 and the global covariance parameter $\gamma$ become redundant.
 The single-cluster model is therefore a special case of the multi-cluster model, 
-in which we set $\mu$ to zero, as described below:
+in which we set $\phi$ to zero, as described below:
 
@@
 		$\eta_i \sim \mathcal{N}(\mu = 0, \sigma = 1)$\\ \hline
-	$\boldsymbol{\mu^{(k)}}$ & 
+	$\boldsymbol{\phi^{(k)}}$ & 
 		cluster-specific shared drift in cluster $k$ &
-		 $\mu^{(k)} \sim \mathcal{N}(\mu = 0, \sigma = 1)$\\ \hline
+		 $\phi^{(k)} \sim \mathcal{N}(\mu = 0, \sigma = 1)$\\ \hline
 	$\boldsymbol{w_i}$ &

### 2017-08-15 writeup/conStruct.tex
added method figure

@@
 The method estimates ancestry proportions for each 
-sample from a set of discrete population clusters, 
-and, within each cluster, estimates a rate at which relatedness decays with distance.
+sample from a set of discrete population layers, 
+and, within each layer, estimates a rate at which relatedness decays with distance.
 This model explicitly addresses the ``clines vs. clusters" problem in 
@@
 \end{abstract}
-
-%evolutionary history of the genotyped samples has also been shaped by geographically continuous processes, 
-%such as migration.
 
@@
 or for identifying either locally adapted alleles (human skin color, ralph \& coop, gartersnakes) or alleles involved in disease (cystic fibrosis).
-Delineating populations can also be useful for informing conservation priorities (ref).
+Delineating populations can also be useful for informing conservation priorities \citep{Moritz1994,Waples_1998,Moritz_etal_2002}.
 However, if underlying population structure is continuous, this may entail a loss of power.
@@
 
-A pattern of isolation by distance (IBD) arises when migration in a population is spatially limited \citep{Slatkin1985},
-and has been modeled in both spatially continuous frameworks \citep{Wright1943,malecot1969mathematics}, 
+A pattern of isolation by distance (IBD) arises when migration in a population is spatially limited \citep{Wright1943},
+and has been modeled in both spatially continuous frameworks \citep{malecot1969mathematics,Slatkin1985}, 
 in which an individual's parents are found within some radius of that individual's location,
@@
 We model genetic variation in genotyped individuals as 
-partitioned within or admixed across a specified number of discrete clusters,
+partitioned within or admixed across a specified number of discrete layers,
 within each of which relatedness decays as a parametric function of the distance between samples.
-We also implement a cross-validation approach for comparing and selecting models across different numbers of clusters,
+We also implement a cross-validation approach for comparing and selecting models across different numbers of layers,
 and we demonstrate the utility of our approach using both simulated and empirical data.
@@
 The choice of allele does not affect subsequent calculations, and so may be arbitrary.
-\plr{How about calling what we calculate the ``allelic covariance''?  ``individual covariance''?}
-We then calculate the \emph{sample covariance} between samples $i$ and $j$, denoted $\widehat{\Omega}_{i,j}$,
+We then calculate the \emph{allelic covariance} between samples $i$ and $j$, denoted $\widehat{\Omega}_{i,j}$,
 as the expected covariance of distinct individual alleles chosen from each of the two samples at a random locus,
@@
     \frac{1}{L} \sum_{\ell=1}^L \left( (f_{i,\ell}-1/2) (f_{j,\ell}-1/2) - \delta_{i,j} f_{i,\ell}^2 / (N_i-1) \right),
-\label{sample_covariance}
+\label{allelic_covariance}
 \end{equation}
@@
 In addition to these differences on the diagonal,
-this differs from the usual ``genetic covariance'' \cite{mcvean_genealogical_2009}
+this differs from the usual ``genetic covariance'' \citep{mcvean_genealogical_2009}
 in that (a) we do not subtract locus means,
@@
 (b) we randomly choose alleles to count.
-To see how to arrive at equation \eqref{sample_covariance},
+To see how to arrive at equation \eqref{allelic_covariance},
 pick a random locus and
@@
 and using the fact that $\E[X] = \E[Y] = 1/2$,
-equation \eqref{sample_covariance} comes from the calculation that
+equation \eqref{allelic_covariance} comes from the calculation that
 $\cov[X,Y] = \E[XY] - 1/4 = (1/2)\E[AB + (1-A)(1-B)] - 1/4 = \E[(A-1/2)(B-1/2)]$.
@@
 Therefore, unlike genetic covariance,
-this covariance is affected in shape by singleton sites,
+this allelic covariance is affected in shape by singleton sites,
 so it may be advisable to filter these prior to analysis
@@
 
-To allow flexibility in the form of the decay of allele frequency covariance with geographic distance within each cluster, 
+To allow flexibility in the form of the decay of allelic covariance with geographic distance within each cluster, 
 we define the covariance within cluster $k$ between samples $i$ and $j$ to be:
@@
 the $\alpha^{(k)}$ parameters control the shape of the decay of covariance with distance in the cluster,
-and $\phi^{(k)}$ is a parameter that describes the background allele frequency covariance within the cluster. 
+and $\phi^{(k)}$ is a parameter that describes the background covariance within the cluster. 
 If two samples draw 100\% of their ancestry from cluster $k$, then their covariance under the model is $G^{(k)}_{i,j}$;
 if they are furthermore geographically very close ($D_{i,j}=0$)
-they will have allele frequency covariance $\alpha^{(k)}_0 +  \phi^{(k)}$.
+they will have covariance $\alpha^{(k)}_0 +  \phi^{(k)}$.
 If the geographic distance between them is very large, 
@@
 connecting the $k$th population to the population ancestral to all modeled
-clusters \gc{(see XXX)}. \gb{is ``XXX" TreeMix? or one of the $f$stat papers?  or is it a figure?} 
-\plr{an F-stat paper; also reference Ben Peters review}
+layers \citep{patterson_ancient_2012, peter_fstats}.
 
@@
 cluster $k$ (and so $\sum_K w^{(k)}_i =1$).
-
-We can then describe the covariance between samples $i$ and $j$ across all clusters, $\Omega_{i,j}$,
-by summing their within-cluster spatial covariances ($G_{i,j}^{(k)}$ in cluster $k$) across all $K$ clusters,
+A visual representation of the method mechanics is shown in Fig \ref{schematic}.
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=4in,height=2.5in]{figs/schematic/method_schematic.png}}
+		\caption{Schematic of our method, using $K=3$ as an example.
+			     The pattern of isolation by distance within each layer is denoted by the color gradient in each of the layer ellipses, 
+			     and $\phi_{(k)}$ denotes the covariance shared by samples with ancestry entirely in the $k$th cluster.
+			     Sampled populations on the landscape are inferred to be admixed between these layers; 
+			     the $i$th sample draws proportion $w^{(k)}_i$ of its ancestry from cluster $k$.
+			    }\label{sim_setup}
+\end{figure}
+
+We can then describe the covariance between samples $i$ and $j$ across all layers, $\Omega_{i,j}$,
+by summing their within-cluster spatial covariances ($G_{i,j}^{(k)}$ in cluster $k$) across all $K$ layers,
 weighted by the relevant admixture proportions.
@@
 this function contains two terms, $\gamma$ and $\delta_{i,j}\eta_i$.
-The first, $\gamma$, describes the global covariance between all samples, 
+The first, $\gamma$, describes the global allelic covariance between all samples, 
 and arises because all samples share an ancestral mean allele frequency at each locus,
@@
 were independent between loci and multivariate normally distributed across populations, 
-then their sample covariance $\widehat{\Omega}$ 
+then their allelic covariance $\widehat{\Omega}$ 
 would be Wishart distributed with degrees of freedom equal to $L$, 
@@
 We use this as a convenient approximation to the true distribution described above,
-and so define the likelihood of the sample standardized allele frequency covariance to be
+and so define the likelihood of the allelic covariance to be
 \begin{equation}
@@
 (see Table \ref{tab:param_prior_tab} for specifics).
-Parameters are independent between clusters.
+Parameters are independent between layers.
 We use Hamiltonian Monte Carlo as implemented in STAN \citep{stan, NUTS, stan_lib, rstan} 
@@
 There are a number of reasons, both statistical and biological, 
-why there is no true (or right) number of clusters for real datasets,
+why there is no true (or right) number of layers for real datasets,
 discussed further in the Discussion.
@@
 we generate a single dataset % that captures both continuous and discrete patterns of population structure 
-by collapsing those $K$ discrete lattices into a single grid of demes that are admixed to various degrees from these different layers
-(see Fig \ref{sim_setup}).
+by collapsing those $K$ discrete lattices into a single grid of demes that are admixed to various degrees from these different layers.
+%(see Fig \ref{sim_setup}).
 We simulated datasets using $K=1$, 2, and 3 layers,
-in each sampling 10,000 unlinked loci from each of 20 haploid individuals from every deme
-(for details on all simulations, see Methods and Supplementary Materials).
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=4in,height=2.5in]{figs/sims/sim_setup.png}}
-		\caption{Schematic of how we simulate datasets with continuous and discrete differentiation, using $K=2$ as an example.  
-			    Going forward in time, the $K$ clusters split from a common ancestor at time $\tau_{\text{s}}$,
-			    then each expand to colonize a lattice of demes with nearest-neighbor symmetric migration at time $\tau_{\text{e}}$,
-			    then finally at time $\tau_{\text{a}}$ collapse into a single lattice consisting of demes 
-			    with ancestry entirely in one or the other of the clusters,
-			    or admixed between them.
-			    }\label{sim_setup}
-\end{figure}
+in each sampling 10,000 unlinked loci from each of 20 haploid individuals from every deme.
+
+%\begin{figure}
+%	\centering
+%		{\includegraphics[width=4in,height=2.5in]{figs/sims/sim_setup.png}}
+%		\caption{Schematic of how we simulate datasets with continuous and discrete differentiation, using $K=2$ as an example.  
+%			    Going forward in time, the $K$ clusters split from a common ancestor at time $\tau_{\text{s}}$,
+%			    then each expand to colonize a lattice of demes with nearest-neighbor symmetric migration at time $\tau_{\text{e}}$,
+%			    then finally at time $\tau_{\text{a}}$ collapse into a single lattice consisting of demes 
+%			    with ancestry entirely in one or the other of the clusters,
+%			    or admixed between them.
+%			    }\label{sim_setup}
+%\end{figure}
 
@@
 at $K=4$, all four corners, and so on.
-\plr{Need a concluding sentence, maybe like this?}
 While it is true that each such cluster is more similar within itself
@@
 And, as $K$ in the spatial model increases, 
-deme membership in clusters beyond the number used to simulate the data does not appreciably increase 
+deme membership in layers beyond the number used to simulate the data does not appreciably increase 
 (Fig \ref{simK1_sp_pies}, \ref{simK2_sp_pies}, \ref{simK3_sp_pies}).
@@
 the inferred clusters, and, if so, 
-does it differ between different clusters?
+does it differ between different layers?
 \end{enumerate}
@@
 
-A consistent split between clusters within \tri{} fell along the ``no-cottonwood belt," 
+A consistent split between layers within \tri{} fell along the ``no-cottonwood belt," 
 a region along the central coast of British Columbia in which black cottonwood is absent. 
@@
 At higher values of $K$, drainages at the southern tip of \tri{} sampling 
-begin to split out into their own clusters.  
+begin to split out into their own layers.  
 \plr{cut this if we pick lower $K$? Or add a disclaimer?}
@@
 We therefore encourage users to perform analyses across different values of $K$ and 
-observe which clusters split out at what levels 
+observe which layers split out at what levels 
 (effectively taking successively shallower cross-sections of the population phylogeny),
@@
 both of which are often predicated on the identification of discrete ``management units" (MUs) 
-identified using genetic data.
+identified using genetic data \citep{Moritz1994,Waples_1998,Moritz_etal_2002}.
 
@@
 There are a few important caveats to consider in the interpretation of \texttt{conStruct} results. 
-First, we have modeled allele frequency covariance within a cluster as a spatial process.
+First, we have modeled allelic covariance within a cluster as a spatial process.
 Although there is flexibility built into the model about the shape of that covariance, 
@@
 Again, we encourage users to run analyses across multiple values of $K$, 
-and to examine the spatial covariance functions within clusters when interpreting results.
+and to examine the spatial covariance functions within layers when interpreting results.
 \gb{missing data?}
@@
 However, we could extend the existing model by implementing 
-a relatedness structure between the clusters by, for example, 
+a relatedness structure between the layers by, for example, 
 estimating a population phylogeny between them \citep[e.g.][]{treemix}.
@@
 If our allele frequency deviations are well approximated by a Gaussian, 
-their sample covariance is a sufficient statistic,
-so that calculating the likelihood of their sample covariance is the same as 
+their sample allelic covariance is a sufficient statistic,
+so that calculating the likelihood of their sample allelic covariance is the same as 
 calculating the probability of the frequency data up to a constant. 
@@
 as a draw from a Wishart distribution with degrees of freedom equal to 
-the number of loci $L$ across which the sample covariance is calculated:
+the number of loci $L$ across which the sample allelic covariance is calculated:
 
@@
 
-A benefit of directly modeling the sample allele frequency covariance is that, 
+A benefit of directly modeling the sample allelic covariance is that, 
 after the initial calculation of the sample covariance matrix,

### 2017-08-17 writeup/conStruct.tex
added fastStructure results

@@
 such as principal components analysis \citep{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
-and model-based clustering approaches \citep[e.g.][]{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, FINESTRUCTURE, huelsenbeck2007inference, Corander2003,TESS,geneland}, among others.
+and model-based clustering approaches \citep[e.g.][]{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, FINESTRUCTURE, fastStructure, huelsenbeck2007inference, Corander2003,TESS,geneland}, among others.
 Here, we will focus on model-based clustering, the most widely used class of approaches for population delineation.
@@
 are shown in Figs \ref{simK1_nsp_pies}-\ref{simK3_sp_pies}.
+To confirm that results from our nonspatial model were qualitatively similar to those of 
+other nonspatial model-based clustering algorithms, 
+we also analyzed each simulated dataset using fastSTRUCTURE \citep{fastStructure}, 
+inferring admixture under models with $K=2$ through 4 (Fig \ref{fastStr_simK1}-\ref{fastStr_simK3}).
 
@@
 \end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=2.25in]{figs/fastStr/fastStr_simK1_pies.pdf}}
+	\caption{
+	Map of admixture proportions estimated using fastSTRUCTURE \citep{fastStructure} for $K=2$ through 4.
+	The data were simulated using one cluster with nearest-neighbor symmetric migration between demes.
+    }\label{fastStr_simK1}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=2.25in]{figs/fastStr/fastStr_simK2_pies.pdf}}
+	\caption{
+	Map of admixture proportions estimated using fastSTRUCTURE \citep{fastStructure} for $K=2$ through 4.
+	The data were simulated using two clusters with nearest-neighbor symmetric migration between demes.
+    }\label{fastStr_simK2}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=2.25in]{figs/fastStr/fastStr_simK3_pies.pdf}}
+	\caption{
+	Map of admixture proportions estimated using fastSTRUCTURE \citep{fastStructure} for $K=2$ through 4.
+	The data were simulated using three clusters with nearest-neighbor symmetric migration between demes.
+    }\label{fastStr_simK3}
+\end{figure}
 

### 2017-08-17 sims/structure/datasets/simK1/exe.fastStr.R

@@ -1,0 +1,19 @@
+setwd("/Applications/STRUCTURE/fastStructure")
+
+K <- 2
+infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK1/simK1"
+outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK1/simK1_estK",K)
+call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
+system(call)
+
+K <- 3
+infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK1/simK1"
+outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK1/simK1_estK",K)
+call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
+system(call)
+
+K <- 4
+infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK1/simK1"
+outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK1/simK1_estK",K)
+call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
+system(call)

### 2017-08-17 sims/structure/datasets/simK2/exe.fastStr.R

@@ -1,0 +1,19 @@
+setwd("/Applications/STRUCTURE/fastStructure")
+
+K <- 2
+infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK2/simK2"
+outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK2/simK2_est",K)
+call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
+system(call)
+
+K <- 3
+infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK2/simK2"
+outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK2/simK2_estK",K)
+call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
+system(call)
+
+K <- 4
+infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK2/simK2"
+outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK2/simK2_estK",K)
+call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
+system(call)

### 2017-08-17 sims/structure/datasets/simK3/exe.fastStr.R

@@ -1,0 +1,19 @@
+setwd("/Applications/STRUCTURE/fastStructure")
+
+K <- 2
+infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK3/simK3"
+outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK3/simK3_estK",K)
+call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
+system(call)
+
+K <- 3
+infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK3/simK3"
+outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK3/simK3_estK",K)
+call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
+system(call)
+
+K <- 4
+infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK3/simK3"
+outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK3/simK3_estK",K)
+call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
+system(call)

### 2017-08-17 sims/structure/format.data.R

@@ -1,0 +1,49 @@
+make.ind.genos <- function(ones,N){
+	geno <- numeric(N)
+	geno[sample(1:N,ones)] <- 1
+	return(geno)
+}
+
+split.pop.into.inds <- function(pop.counts,pop.nChromo){
+	ind.geno <- apply(pop.counts,2,function(x){make.ind.genos(x,pop.nChromo)})
+	return(ind.geno)
+}
+
+make.structure.data.file <- function(counts,sample.sizes){
+	geno.matrix <- matrix(NA,sum(sample.sizes),ncol(freqs))
+	index <- 0
+	for(n in 1:nrow(freqs)){
+		geno.matrix[index+c(1:sample.sizes[n]),] <- split.pop.into.inds(counts[n,,drop=FALSE],sample.sizes[n])
+		index <- index + sample.sizes[n]
+	}
+	PopData <- unlist(lapply(1:length(sample.sizes),
+					function(n){
+						lapply(1:sample.sizes[n],
+							function(i){
+								sprintf("ind_%s.%s",n,i)
+							})
+					}))
+	LocData <- unlist(lapply(1:length(sample.sizes),function(n){rep(n,sample.sizes[n])}))
+	return(cbind(PopData,LocData,geno.matrix))
+}
+
+load("~/Dropbox/conStruct/sims/cross_validation/K_1/sim.dataset.Robj")
+freqs <- sim.dataset$freq.data$freqs
+sample.sizes <- sim.dataset$freq.data$sample.sizes
+counts <- freqs*sample.sizes
+str.data <- make.structure.data.file(counts,sample.sizes[,1])
+write.table(str.data,file="datasets/simK1/simK1.str",quote=FALSE,row.names=FALSE,col.names=FALSE)
+
+load("~/Dropbox/conStruct/sims/cross_validation/K_2/sim.dataset.Robj")
+freqs <- sim.dataset$freq.data$freqs
+sample.sizes <- sim.dataset$freq.data$sample.sizes
+counts <- freqs*sample.sizes
+str.data <- make.structure.data.file(counts,sample.sizes[,1])
+write.table(str.data,file="datasets/simK2/simK2.str",quote=FALSE,row.names=FALSE,col.names=FALSE)
+
+load("~/Dropbox/conStruct/sims/cross_validation/K_3/sim.dataset.Robj")
+freqs <- sim.dataset$freq.data$freqs
+sample.sizes <- sim.dataset$freq.data$sample.sizes
[…]

### 2017-08-17 writeup/figs/fastStr_figs.R

@@ -1,0 +1,95 @@
+collapse.ind.Q <- function(Q,pop.vec){
+	n.pops <- length(unique(pop.vec))
+	admix.props <- matrix(NA,n.pops,ncol(Q))
+	for(i in 1:n.pops){
+		admix.props[i,] <- colMeans(Q[which(pop.vec==i),])
+	}
+	return(admix.props)
+}
+
+pie.plot <- function(Q.file,coords,pop.vec,cluster.colors){
+	Q <- read.table(Q.file,stringsAsFactors=FALSE)
+	admix.props <- collapse.ind.Q(Q,pop.vec)
+	K <- ncol(admix.props)
+	N <- length(unique(pop.vec))
+	cluster.names <- paste0("cluster_",1:K)
+	sample.names <- paste0("sample_", 1:N)
+	color.tab <- caroline::nv(c(cluster.colors[1:K]), cluster.names)
+	pie.list <- lapply(1:N, function(i) {
+	    caroline::nv(admix.props[i, ], cluster.names)
+	})
+	names(pie.list) <- sample.names
+	caroline::pies(pie.list, x0 = coords[, 1], 
+	    y0 = coords[, 2], color.table = color.tab, 
+	    border = "black", radii = 3, xlab = "", ylab = "", 
+	    lty = 1, density = NULL,
+	    xlim = range(coords[,1]) + c(-0.5,0.5),
+	    ylim = range(coords[,2]) + c(-0.5,0.5))
+	box(lwd = 2)
+}
+
+load("~/Dropbox/conStruct/sims/cross_validation/K_1/sim.dataset.Robj")
+
+pdf(file="~/Dropbox/conStruct/writeup/figs/fastStr/fastStr_simK1_pies.pdf",width=12,height=4.5,pointsize=13)
+	par(mfrow=c(1,3))
+		pie.plot(Q.file = "~/Dropbox/conStruct/sims/structure/datasets/simK1/simK1_estK2.2.meanQ",
+				 coords = sim.dataset$coords,
+				 pop.vec = unlist(lapply(1:sim.dataset$N,function(n){rep(n,10)})),
+				 cluster.colors = c(4,2))
+			mtext(side=3,font=2,text="K=2")
+		pie.plot(Q.file = "~/Dropbox/conStruct/sims/structure/datasets/simK1/simK1_estK3.3.meanQ",
+				 coords = sim.dataset$coords,
[…]

### 2017-08-17 writeup/conStruct.tex
moved allelic cov to Appendix

@@
 
-%%%%%%%%%%%%%%%%%
 \section*{Methods}
@@
 and so to make the statistic invariant under choices of reference allele, 
-(b) we randomly choose alleles to count.
-To see how to arrive at equation \eqref{allelic_covariance},
-pick a random locus and
-let $A$ and $B$ be randomly drawn alleles at that locus from populations $i$ and $j$ respectively.
-If $i=j$ then these are chosen \emph{without} replacement.
-Suppose these are each coded as `0' or `1` (where `0' denotes a reference allele),
-but we randomly ``flip'' this coding, so that we let $X=A$ and $Y=B$ with probability 1/2,
-but otherwise we let $X=1-A$ and $Y=1-B$.
-We then let $\widehat{\Omega}_{i,j} = \cov[X,Y]$ --
-sampling without replacement makes the statistic insensitive to sample size,
-and random allele flipping makes it independent of the choice of reference allele.
-By conditioning on the flip,
-and using the fact that $\E[X] = \E[Y] = 1/2$,
-equation \eqref{allelic_covariance} comes from the calculation that
-$\cov[X,Y] = \E[XY] - 1/4 = (1/2)\E[AB + (1-A)(1-B)] - 1/4 = \E[(A-1/2)(B-1/2)]$.
-The factor of $\delta_{i,j}/(N_i-1)$ comes from the fact that if $i=j$,
-then $\E[AB]$ is the probability that both alleles drawn without replacement are reference,
-which is $f_i^2 n/(n-1)$.
+(b) we randomly choose an allele to count at each locus.
+We discuss the derivation of Eqn. \eqref{allelic_covariance} further in the Appendix (\secref{allelic_cov}).
 As noted in \cite{EEMS},
@@
 			     the $i$th sample draws proportion $w^{(k)}_i$ of its ancestry from cluster $k$.
-			    }\label{sim_setup}
+			    }\label{schematic}
 \end{figure}
@@
 
-A nice feature of our approach is that the model described in Eqn \refeq{cross_cluster_covariance} 
+A nice feature of our approach is that the model described in Eqn \eqref{cross_cluster_covariance} 
 contains a nonspatial assignment model as a special case 
-(see \ref{model_app} for a more in-depth discussion). 
+(see \secref{model_app} for a more in-depth discussion). 
 By setting $\alpha_0^{(k)}$ to zero for all $k$, 
@@
 
-\gc{IS PREDICTION accuracy the right phrase?}
 We choose to take a cross-validation approach \citep[similar in spirit to][]{ADMIXTURE_xval}.
@@
 and select the model with the significantly best predictive accuracy
-(see Section \ref{Xvalidation} for more on our cross-validation procedure).
+(see \secref{Xvalidation} for more on our cross-validation procedure).
 This approach rewards models with larger values of $K$, 
@@
 The best model is judged to be the simplest one with significantly better predictive accuracy
-than others (see Section \ref{Xvalidation} for more on our cross-validation procedure).
+than others (see \secref{Xvalidation} for more on our cross-validation procedure).
 In general, larger values of $K$ allows the model more flexibility,
@@
 Finally, at time $\tau_{\text{a}}$ in the past 
-we generate a single dataset % that captures both continuous and discrete patterns of population structure 
+we generate a single dataset 
 by collapsing those $K$ discrete lattices into a single grid of demes that are admixed to various degrees from these different layers.
-%(see Fig \ref{sim_setup}).
+(see Fig \ref{sim_setup}).
 We simulated datasets using $K=1$, 2, and 3 layers,
 in each sampling 10,000 unlinked loci from each of 20 haploid individuals from every deme.
-
-%\begin{figure}
-%	\centering
-%		{\includegraphics[width=4in,height=2.5in]{figs/sims/sim_setup.png}}
-%		\caption{Schematic of how we simulate datasets with continuous and discrete differentiation, using $K=2$ as an example.  
-%			    Going forward in time, the $K$ clusters split from a common ancestor at time $\tau_{\text{s}}$,
-%			    then each expand to colonize a lattice of demes with nearest-neighbor symmetric migration at time $\tau_{\text{e}}$,
-%			    then finally at time $\tau_{\text{a}}$ collapse into a single lattice consisting of demes 
-%			    with ancestry entirely in one or the other of the clusters,
-%			    or admixed between them.
-%			    }\label{sim_setup}
-%\end{figure}
 
@@
 we also analyzed each simulated dataset using fastSTRUCTURE \citep{fastStructure}, 
-inferring admixture under models with $K=2$ through 4 (Fig \ref{fastStr_simK1}-\ref{fastStr_simK3}).
+inferring admixture under models with $K=2$ through 4.
 
@@
 at $K=4$, all four corners, and so on.
+These results are qualitatively similar to those generated using fastSTRUCTURE \citep{fastStructure}, 
+which also describes continuous differentiation using admixture between clusters
+ (Fig \ref{fastStr_simK1}-\ref{fastStr_simK3}).
 While it is true that each such cluster is more similar within itself
 than it is to other clusters,
-we know that these boundaries are arbitray,
+we know that these boundaries are arbitrary,
 since the data were simulated with no such boundaries.
@@
 Trees in the genus \textit{Populus} (poplars, aspens, and cottonwoods) 
-are distributed throughout the Northern Hemisphere, 
-and include species of ecological and economic importance.
-Species in the genus regularly co-occur and, 
+are distributed throughout the Northern Hemisphere;
+species in the genus regularly co-occur and, 
 where they do, they frequently hybridize \citep{eckenwalder1984, Cronk2005}.
@@
 and show spatial patterns of genetic and phenotypic variation \citep{slavov_etal_2012, mckown_etal_2013},
-making the system well-suited to an analysis designed to describe
+making the system well-suited to test an analysis designed to describe
 how genetic structure is partitioned both discretely across population (or species) clusters
 and continuously with geographic distance.
-We had the following questions for the system: 
+We organize the results of our analyses around the following questions: 
 \begin{enumerate}
@@
 with some admixture,
-% Although admixture was inferred between \textit{P. balsamifera} and \tri{}
-% across all values of $K>1$, and in both the spatial and nonspatial models,
-% there is nonetheless strong evidence for discrete population structuring, 
 rather than a single, continuous cline of ancestry.
@@
 
+\section{Allelic Covariance}\label{allelic_cov}
+To see how to arrive at Eqn. \eqref{allelic_covariance},
+pick a random locus and
+let $A$ and $B$ be randomly drawn alleles at that locus from populations $i$ and $j$ respectively.
+If $i=j$, then these are chosen \emph{without} replacement.
+Suppose these are each coded as `0' or `1` (where `0' denotes a reference allele),
+but we randomly ``flip'' this coding, so that we let $X=A$ and $Y=B$ with probability 1/2,
+but otherwise we let $X=1-A$ and $Y=1-B$.
+We then let $\widehat{\Omega}_{i,j} = \cov[X,Y]$ ; 
+sampling without replacement makes the statistic insensitive to sample size,
+and random allele flipping makes it independent of the choice of reference allele.
+By conditioning on the flip,
+and using the fact that $\E[X] = \E[Y] = 1/2$,
+Eqn. \eqref{allelic_covariance} comes from the calculation that
+\begin{align}
+\cov[X,Y] &= \E[XY] - 1/4 = (1/2)\E[AB + (1-A)(1-B)] - 1/4\\\notag
+&= \E[(A-1/2)(B-1/2)]
+\end{align}
+The factor of $\delta_{i,j}/(N_i-1)$ comes from the fact that if $i=j$,
+then $\E[AB]$ is the probability that both alleles drawn without replacement are reference,
+which is $f_i^2 n/(n-1)$.
 \paragraph{Likelihood}
@@
 the number of loci $L$ across which the sample allelic covariance is calculated:
-
+\gb{should we take this equation out?}
 \begin{align}
@@
 each of which can be implemented with different numbers of clusters.
-The spatial model is parameterized as in Eqn \ref{admixed_spatial_cov},
-and the nonspatial model is described in Eqn \ref{admixed_discrete_covariance}.
+The spatial model is parameterized as in Eqn \eqref{admixed_spatial_cov},
+and the nonspatial model is described in Eqn \eqref{admixed_discrete_covariance}.
 The nonspatial model is a special case of the spatial model, 
@@
 
-\section{Our Model vs. STRUCTURE}
-We should probably demonstrate that the nonspatial version of \texttt{conStruct} is basically the same as STRUCTURE.
-Maybe check out Engelhardt \& Stephens?
-
 \newpage
@@
 \setcounter{figure}{0}
+
+\clearpage
+\begin{figure}
+	\centering
+		{\includegraphics[width=4in,height=2.5in]{figs/sims/sim_setup.png}}
+		\caption{Schematic of how we simulate datasets with continuous and discrete differentiation, using $K=2$ as an example.  
+			    Going forward in time, the $K$ clusters split from a common ancestor at time $\tau_{\text{s}}$,
+			    then each expand to colonize a lattice of demes with nearest-neighbor symmetric migration at time $\tau_{\text{e}}$,
+			    then finally at time $\tau_{\text{a}}$ collapse into a single lattice consisting of demes 
+			    with ancestry entirely in one or the other of the clusters,
+			    or admixed between them.
+			    }\label{sim_setup}
+\end{figure}
 

### 2017-08-18 writeup/conStruct.tex
intro changes

@@
 \section*{Introduction}
-The fundamental goal of evolutionary genetics is to link patterns of genetic variation 
-to the processes that have generated and maintained them.
-Often, the first step in the analysis of genetic data is to define evolutionary units of analysis
-by delineating discrete populations from which individuals or alleles are sampled \citep{wright1949genetical}.
-These units are useful for defining a scope of study in research on selection or local adaptation (ref),
-or for identifying either locally adapted alleles (human skin color, ralph \& coop, gartersnakes) or alleles involved in disease (cystic fibrosis).
-Delineating populations can also be useful for informing conservation priorities \citep{Moritz1994,Waples_1998,Moritz_etal_2002}.
-However, if underlying population structure is continuous, this may entail a loss of power.
-
-There have been many methods proposed to characterize populations,
+A fundemental quandary in the description of biological diversity, is
+the fact that diversity shows both discrete and continuous patterns. 
+For example, there is disagreement on species concepts because the
+process of speciation is usually gradual, and so there is no set point
+in the continuous divergence of populations when they become discrete species.
+This issue extends below the species level as patterns of phenotypic
+and genetic diversity within and among populations are shaped
+by continuous migration and drift, as well as more discrete events and
+processes such as geographic barriers, rapid expansions, bottlenecks, and rare
+long distance migration.
+
+From a practical standpoint, even while acknowledging continuous
+processes, we often need to identify somewhat separable populations 
+from which individuals are sampled \citep{wright1949genetical}.
+Delineating populations is useful for systematics and for
+informing conservation priorities \citep{Moritz1994,Waples_1998,Moritz_etal_2002}.
+Furthermore, we often need to identify sets of indivduals resulting from reasonably coherent
+evolutionary histories for downstream analyses to learn about population history and adaptation.
+
+%In studying populations we need group individuals 
+%Many downstream
+
+%Biology diversity
+
+%The fundamental goal of evolutionary genetics is to link patterns of genetic variation 
+%to the processes that have generated and maintained them.
+%Often, the first step in the analysis of genetic data is to define evolutionary units of analysis
+%by delineating discrete populations from which individuals or alleles are sampled \citep{wright1949genetical}.
+%These units are useful for defining a scope of study in research on selection or local adaptation (ref),
+%or for identifying either locally adapted alleles (human skin color, ralph \& coop, gartersnakes) or alleles involved in disease (cystic fibrosis).
+%Delineating populations can also be useful for informing conservation priorities \citep{Moritz1994,Waples_1998,Moritz_etal_2002}.
+%However, if underlying population structure is continuous, this may entail a loss of power.
+
+There have been many methods proposed to characterize population
+genetic structure,
 including generating population phylogenies \citep{CavalliSforza1975, treemix},

### 2017-08-18 writeup/conStruct.tex
intro changes

@@
 Given the generality of the circumstances that generate a pattern of isolation by distance, 
-it is unsurprising that IBD is found everywhere in nature \citep{meirmans2012,Sexton_etal_2014}.
+it is unsurprising that IBD is very widespread in nature \citep{meirmans2012,Sexton_etal_2014}.
 
@@
 \citep[e.g.][]{SerrePaabo2004,rosenberg2005clines}.
-As a further complication, if local genetic drift is strong,
-continous models are expected to give rise to spatially patchy patterns
-\citep{SLFV,bramson1989crabgrass}.
+%As a further complication, if local genetic drift is strong,
+%continous models are expected to give rise to spatially patchy patterns
+%\citep{SLFV,bramson1989crabgrass}.
 
 Existing model-based clustering approaches can only describe continuous patterns of variation using
-discrete clusters, 
-and so tend to erroneously describe continuously distributed variation with multiple clusters that 
+discrete clusters, and so tend to erroneously describe continuously distributed variation with multiple clusters that 
 show spatially autocorrelated cluster membership \citep{Frantz2009,meirmans2012}.
 In analyses of empirical datasets, which often are characterized by IBD,
-model-based clustering approaches will therefore tend to overestimate the number of discrete clusters present,
-with downstream implications for interpretation of results.
-For example, conservation resources might be spread too thinly over areas 
-that are spuriously assumed to be discrete management units.
-Or, in studies of locally adaptive loci, or of the genetics of disease,
-discrete populations may be modeled in separately,
-reducing the global power to detect the genomic basis of complex phenotypes.
+model-based clustering approaches will therefore tend to overestimate
+the number of discrete clusters present.
+
+%with downstream implications for interpretation of results.
+%For example, conservation resources might be spread too thinly over areas 
+%that are spuriously assumed to be discrete management units.
+%Or, in studies of locally adaptive loci, or of the genetics of disease,
+%discrete populations may be modeled in separately,
+%reducing the global power to detect the genomic basis of complex phenotypes.
 

### 2017-08-23 data/bears/exe.xval.R
added bear results

@@ -1,0 +1,38 @@
+library(conStruct)
+library(doParallel)
+
+x.validation <- function(test.pct,n.reps,K,freqs,geoDist,coords,prefix,n.iter){
+	x.val <- mclapply(1:n.reps,
+					function(i){
+						x.validation.rep(rep.no = i,
+										 test.pct,
+										 K,
+										 freqs,
+										 geoDist,
+										 coords,
+										 prefix,
+										 n.iter)
+					},mc.cores=6,mc.allow.recursive=FALSE,mc.preschedule=FALSE)
+	names(x.val) <- paste0("rep_",1:n.reps)
+	return(x.val)
+}
+
+
+load("bear.dataset.Robj")
+
+missing.data <- apply(bear.dataset$sample.freqs,1,function(x){length(which(is.na(x)))})
+too.much.missing.data <- which(missing.data/ncol(bear.dataset$sample.freqs) > 0.04)
+
+bear.dataset <- list("sample.freqs" = bear.dataset$sample.freqs[-too.much.missing.data,],
+					 "sample.sizes" = bear.dataset$sample.sizes[-too.much.missing.data],
+					 "sample.coords" = bear.dataset$sample.coords[-too.much.missing.data,])
+
+save(bear.dataset,file="bear.dataset.redux.Robj")
+x.validation(test.pct = 0.10,
+			 n.reps = 10,
+			 K = 1:7,
+			 freqs = bear.dataset$sample.freqs,
+			 geoDist = fields::rdist.earth(bear.dataset$sample.coords),
+			 coords = bear.dataset$sample.coords,
+			 prefix = "bear_",
+			 n.iter = 5e3)

### 2017-08-23 writeup/figs/bear_figs.R

@@ -1,0 +1,112 @@
+source("~/Dropbox/conStruct/writeup/figs/fig_funcs.R")
+
+n.reps <- 6
+K <- 7
+for(n in 1:n.reps){
+	load(sprintf("bear_rep%s_test.lnl.Robj",n,n))
+	assign(paste0("tl",n),test.lnl)
+}
+x.vals <- lapply(1:n.reps,function(n){get(sprintf("tl%s",n))})
+x.vals.std <- lapply(x.vals,standardize.xvals)
+xval.CIs <- get.xval.CIs(x.vals.std,K)
+pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bear_std_xval.pdf",width=10,height=5,pointsize=14)
+#	quartz(width=10,height=5)
+	par(mfrow=c(1,2),mar=c(4,5,4,2))
+	plot.xval.CIs(xval.CIs,K,jitter=0.1,xlim=c(0.75,7.25))
+		legend(x="bottomright",pch=19,col=c("blue","green"),legend=c("spatial","nonspatial"))
+	mtext("Predictive accuracy",side=2,padj=-5,font=2)
+	plot.xval.CIs(xval.CIs,K,ylim=c(-200,0),jitter=0.1,xlim=c(3.75,7.25))
+		legend(x="bottomright",pch=c(19,NA),lty=c(NA,1),lwd=c(NA,2),col=c(1,adjustcolor(1,0.8)),legend=c("mean","95% CI"))
+	mtext("number of clusters",side=1,adj=-1.1,padj=4,font=2)
+	mtext("Cross-validation results (Bears)",side=3,adj=5,padj=-2.5,font=2,cex=1.2)
+dev.off()
+
+
+library(conStruct)
+library(maps)
+load("~/Dropbox/conStruct/data/bears/bear.dataset.redux.Robj")
+pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bear_sampling_map.pdf",width=6,height=4,pointsize=13)
+	#quartz(width=6,height=4,pointsize=13)
+	par(mar=c(1,1.5,0,0),oma=c(0,0,0,0))
+	map(xlim = range(bear.dataset$sample.coords[,1]) + c(-5,5), ylim = range(bear.dataset$sample.coords[,2])+c(-2,2), col="gray")
+	points(bear.dataset$sample.coords,pch=19,cex=1.2)
+	map.axes()
+dev.off()
+
+
+freq.data <- conStruct:::process.freq.data(bear.dataset$sample.freqs)
[…]

### 2017-08-23 data/poplars/exe.xval.R
added poplars

@@ -1,0 +1,31 @@
+library(conStruct)
+library(doParallel)
+
+x.validation <- function(test.pct,n.reps,K,freqs,geoDist,coords,prefix,n.iter){
+	x.val <- mclapply(1:n.reps,
+					function(i){
+						x.validation.rep(rep.no = i,
+										 test.pct,
+										 K,
+										 freqs,
+										 geoDist,
+										 coords,
+										 prefix,
+										 n.iter)
+					},mc.cores=7,mc.allow.recursive=FALSE,mc.preschedule=FALSE)
+	names(x.val) <- paste0("rep_",1:n.reps)
+	return(x.val)
+}
+
+
+load("poplar.spStr.dataset.Robj")
+freqs <- poplar.data$freqs[,!apply(poplar.data$freqs,2,function(x){any(is.na(x))})]
+
+x.validation(test.pct = 0.10,
+			 n.reps = 10,
+			 K = 1:7,
+			 freqs = freqs,
+			 geoDist = fields::rdist.earth(poplar.data$coords),
+			 coords = poplar.data$coords,
+			 prefix = "poplar_",
+			 n.iter = 2e3)

### 2017-08-23 writeup/conStruct.tex
minor tinkering

--- 2017-08-17 writeup/conStruct.tex
@@
 The results of these analyses are shown in 
-Figs \ref{populus_xvals}--\ref{populus_nsp_clst_covs}. %,populus_nsp_pies,populus_sp_clst_covs,populus_nsp_clst_covs
+Figs \ref{populus_xvals}--\ref{populus_sp_clst_covs} and \ref{populus_nsp_clst_covs}. %,populus_nsp_pies,populus_sp_clst_covs,populus_nsp_clst_covs
 
@@
     }\label{populus_sp_clst_covs}
-\end{figure}
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=6in,height=4in]{figs/populus/populus_nsp_clst_covs.pdf}}
-	\caption{
-	Plots showing the cluster-specific parametric covariance IBD-curves 
-	estimated for the \textit{Populus} data using 
-	the nonspatial \texttt{conStruct} model run with $K=2$ through 7.
-	Line colors are consistent with cluster colors in Fig \ref{populus_nsp_pies}.
-	Points are colored by the species they are a covariance between:
-	black on black points are sample covariances between populations of \textit{Populus balsamifera};
-	green on black points are sample covariances between \bals{} and \tri{};
-	green on green points are sample covariances between \tri{} and \tri{}.
-    }\label{populus_nsp_clst_covs}
 \end{figure}
@@
 
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=4in]{figs/populus/populus_nsp_clst_covs.pdf}}
+	\caption{
+	Plots showing the cluster-specific parametric covariances 
+	estimated for the \textit{Populus} data using 
+	the nonspatial \texttt{conStruct} model run with $K=2$ through 7.
+	Line colors are consistent with cluster colors in Fig \ref{populus_nsp_pies}.
+	Points are colored by the species they are a covariance between:
+	black on black points are sample covariances between populations of \textit{Populus balsamifera};
+	green on black points are sample covariances between \bals{} and \tri{};
+	green on green points are sample covariances between \tri{} and \tri{}.
+    }\label{populus_nsp_clst_covs}
+\end{figure}
+
 \newpage

### 2017-08-23 writeup/conStruct.tex
getting tex to compile

@@
 \section*{Introduction}
-A fundemental quandary in the description of biological diversity, is
+A fundamental quandary in the description of biological diversity is
 the fact that diversity shows both discrete and continuous patterns. 
@@
 process of speciation is usually gradual, and so there is no set point
-in the continuous divergence of populations when they become discrete species.
-This issue extends below the species level as patterns of phenotypic
+in the continuous divergence of populations when they unambiguously become discrete species.
+This issue extends below the species level, as patterns of phenotypic
 and genetic diversity within and among populations are shaped
-by continuous migration and drift, as well as more discrete events and
-processes such as geographic barriers, rapid expansions, bottlenecks, and rare
+by continuous migration and drift, as well as by more discrete events, 
+such as geographic barriers, rapid expansions, bottlenecks, and rare
 long distance migration.

### 2017-08-30 writeup/figs/sim_figs.R
sim results

@@ -13,6 +13,7 @@ pdf(file="~/Dropbox/conStruct/writeup/figs/sims/simK1_std_xval.pdf",width=10,hei
 	plot.sim.xvals(dir="~/Dropbox/conStruct/sims/cross_validation/K_1/x_validation",n.reps=10,K=7,simK=1,y.lim=c(-10,0))
 dev.off()
 
+setwd("~/Dropbox/conStruct/sims/cross_validation/K_1")
 library(conStruct)
 load("sim.dataset.Robj")
 freq.data <- conStruct:::process.freq.data(sim.dataset$freq.data$freqs)
@@ -21,17 +22,87 @@ data.block <- conStruct:::make.data.block(K = 1,
 										  coords = sim.dataset$coords,
 										  spatial = TRUE,
 										  geoDist = fields::rdist(sim.dataset$coords))
-load("training.runs.nsp.Robj")
+
+output.list.sp <- vector("list",7)
+for(k in 1:7){
+	load(sprintf("~/Dropbox/conStruct/sims/cross_validation/K_1/runs/simK1_K%s_sp_conStruct.results.Robj",k))
+	output.list.sp[[k]] <- conStruct.results
+}
+
+output.list.nsp <- vector("list",7)
+for(k in 1:7){
+	load(sprintf("~/Dropbox/conStruct/sims/cross_validation/K_1/runs/simK1_K%s_nsp_conStruct.results.Robj",k))
+	output.list.nsp[[k]] <- conStruct.results
+}
+
+
 plot.sim.pies(data.block = data.block,
 			  K = 7,
-			  training.runs = training.runs.nsp,
+			  output.list = output.list.nsp,
 			  file.name = "~/Dropbox/conStruct/writeup/figs/sims/simK1_nsp_pies_K")
 
-load("training.runs.sp.Robj")
 plot.sim.pies(data.block = data.block,
 			  K = 7,
-			  training.runs = training.runs.sp,
+			  output.list = output.list.sp,
 			  file.name = "~/Dropbox/conStruct/writeup/figs/sims/simK1_sp_pies_K")
+
+K <- 7
+laycon.sp <- matrix(0,nrow=7,ncol=7)
+colnames(laycon.sp) <- paste(1:7)
+for(k in 1:K){
+	csr <- output.list.sp[[k]][[1]]
+	clst.match <- NULL
[…]

### 2017-08-30 writeup/figs/fig_funcs.R
updated fig making functions

@@ -83,7 +83,6 @@ plot.xval.CIs <- function(xval.CIs,K,k.range=c(1:K),ylim=NULL,cex=1.5,jitter=0,.
 }
 
 calculate.qij <- function(cluster.params,data.block,i,j){
-	cluster.params <- fix.pars(cluster.params)
 	q_ij <- 2 * cluster.params$alpha0 * 
 				exp(-(cluster.params$alphaD * 
 						data.block$geoDist[i,j])^cluster.params$alpha2) + 
@@ -124,40 +123,43 @@ get.admix.CIs <- function(conStruct.results){
 	CIs <- apply(conStruct.results$posterior$admix.proportions,
 					c(2,3),
 					function(x){
-						quantile(x,c(0.025,0.975))
+						quantile(x,c(0.005,0.995))
 					}
 			)
 	return(CIs)
 }
 
 viz.admix.results <- function(sim.admix.props,conStruct.results,clst.order=NULL){
-	layout(matrix(c(1,1,2),1,3))
+	#layout(matrix(c(1,1,2),1,3))
 	k.cols <- c("blue", "red", "green", "yellow", "purple", "orange", "lightblue", "darkgreen", "lightblue", "gray")	
 	K <- ncol(sim.admix.props)
 	if(is.null(clst.order)){
 		clst.order <- 1:K
 	}
 	N <- nrow(sim.admix.props)
-	plot(0,ylim=c(0,1),xlim=c(1,N) + c(-0.5,0.5),type='n',ylab="admixture proportion",xlab="samples")
+	#plot(0,ylim=c(0,1),xlim=c(1,N) + c(-0.5,0.5),type='n',ylab="admixture proportion",xlab="samples")
 	admix.prop.CIs <- get.admix.CIs(conStruct.results)
-	for(k in 1:K){
-		segments(1:N,admix.prop.CIs[1,,k],1:N,admix.prop.CIs[2,,k],col=adjustcolor(k.cols[k],0.7),lwd=2.5)
-		points(sim.admix.props[,clst.order[k]],col="gray",bg=k.cols[k],pch=23,cex=1.2)
-	}
+	# for(k in 1:K){
+		# segments(1:N,admix.prop.CIs[1,,k],1:N,admix.prop.CIs[2,,k],col=adjustcolor(k.cols[k],0.7),lwd=2.5)
+		# points(sim.admix.props[,clst.order[k]],col="gray",bg=k.cols[k],pch=23,cex=1.2)
+	# }
[…]

### 2017-08-30 writeup/conStruct.tex
complete draft

@@
 continuous and discrete patterns of differentiation. 
-We demonstrate the utility of this approach using both simulations and an empirical application.
+We demonstrate the utility of this approach using both simulations and empirical applications.
 \end{abstract}
@@
 
-%In studying populations we need group individuals 
-%Many downstream
-
-%Biology diversity
-
-%The fundamental goal of evolutionary genetics is to link patterns of genetic variation 
-%to the processes that have generated and maintained them.
-%Often, the first step in the analysis of genetic data is to define evolutionary units of analysis
-%by delineating discrete populations from which individuals or alleles are sampled \citep{wright1949genetical}.
-%These units are useful for defining a scope of study in research on selection or local adaptation (ref),
-%or for identifying either locally adapted alleles (human skin color, ralph \& coop, gartersnakes) or alleles involved in disease (cystic fibrosis).
-%Delineating populations can also be useful for informing conservation priorities \citep{Moritz1994,Waples_1998,Moritz_etal_2002}.
-%However, if underlying population structure is continuous, this may entail a loss of power.
-
 There have been many methods proposed to characterize population
@@
 including generating population phylogenies \citep{CavalliSforza1975, treemix},
-$k$-means clustering analyses performed on dimensionality-reduction approaches (ref), 
+$k$-means clustering analyses performed on dimensionality-reduction approaches \citep{meirmans2009genodive}, 
 such as principal components analysis \citep{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
@@
 To address this, we set out to develop
-a model-based clustering method that when possible uses isolation by distance 
+a model-based clustering method that, when possible, uses isolation by distance 
 to explain observed genetic variation.
 With an explicit spatial component, discrete population structure need only be invoked when genetic differentiation 
-in the data is deviates significantly from that expected given geographic separation.
+in the data deviates significantly from that expected given geographic separation.
 
@@
 if $D_{i,j}$ is the mean density of sites at which random samples from $i$ and $j$ differ at a randomly chosen locus,
-then $\Sigma_{i,j} = (1 - 2 D_{i,j})/4$.
+then $\Omega_{i,j} = (1 - 2 D_{i,j})/4$.
 Therefore, unlike genetic covariance,
@@
 Here we add a spatial component within each such discrete historical component,
-which we refer to as a set of ``layers'' which overlay the modern map.
+which we refer to as a set of ``layers'' that overlay the modern map.
 We imagine each layer as a geographically distributed population,
@@
 allele frequencies have positive covariance at geographically close locations,
-but this covariance is allowed to decay towards zero as geographic distance increases.
+but this covariance is allowed to decay as geographic distance increases.
 This pattern of spatial decay reflects how migration between neighboring demes  
 homogenizes allele frequency changes that arise locally due to drift, 
-but this mixing less effectively homogenizes geographic distant demes,
+but less effectively homogenizes geographic distant demes,
 resulting in a continuous pattern of isolation by distance within each cluster.
@@
 around their expectations, 
-while the standard assignment methods assume that the error is binomial distributed (see XXXX).  
+while the standard assignment methods assume that the error is binomially distributed \citep{Engelhardt2012}.  
 Therefore, we can compare the fit of the different models 
@@
 However, it is still important to assess whether additional spatial layers (larger $K$)
-meaningfully model pattern in the data
+meaningfully model patterns in the data
 or merely explain spurious variation introduced by noise
@@
 for the number of clusters in population genetic data, including: 
-the use of information theoretic criteria \citep[e.g.][]{ADMIXTURE}, such as AIC or BIC;
+information theoretic criteria \citep[e.g.][]{ADMIXTURE}, such as AIC or BIC;
 comparisons of the likelihood of the data across different values of $K$, 
@@
 or via a more rigorous and computationally intensive approach such as thermodynamic integration \citep{verity_nichols2016}
-or inference using a Dirichlet process prior \citep{huelsenbeck2007inference}
+or inference using a Dirichlet process prior \citep{huelsenbeck2007inference}.
 See \cite{verity_nichols2016} for a discussion of these approaches and comparison
@@
 
-We choose to take a cross-validation approach \citep[similar in spirit to][]{ADMIXTURE_xval}.
-We estimate the posterior on the parameters of the model using a ``training" partition of the data 
-(in practice the training set is a random subset of $90\%$ of the loci). 
-We then calculate the likelihood of the ``testing" subset (the remaining $10\%$ of the loci), 
-averaged over the posterior of the parameters. 
-We repeat this procedure a number of times, 
-take an average of the relative log-likelihoods over the runs at each $K$, 
-and select the model with the significantly best predictive accuracy
-(see \secref{Xvalidation} for more on our cross-validation procedure).
-This approach rewards models with larger values of $K$, 
-which increase the likelihood of the training partition, 
-as they fit the data better,
-but the improvement in the likelihood will plateau (or peak), 
-as above a certain $K$ the model is fitting noise specific to the training data 
-and therefore not fitting the testing set any better.
-\plr{I wonder if we can just cite \cite{verity_nichols2016} rather than this laundry list here?}
-
-We use cross-validation to attack this problem.
+We use cross-validation \citep[similar in spirit to][]{ADMIXTURE_xval} to attack this problem.
 To do this,
@@
 
-\gb{something about how we choose the best model?}
-\plr{how's this?}
 Cross-validation provides a valuable summary of how much explanatory power
@@
 should probably not be relied on too strongly.
+
+An additional heuristic to evaluate the inclusion of additional layers is to calculate 
+each layer's relative contribution to total covariance as their importance, 
+and to choose a value of $K$ where all layers have an importance above some cutoff (e.g. 0.1\%).
+Because the prior on admixture proportions is a Dirichlet distribution 
+that is quite harsh against intermediate admixture values (see Table \ref{tab:param_prior_tab}),  
+the model will tend to ``not use" unnecessary layers, 
+and they will have a low contribution to overall covariance.
+The contribution of the $k$th layer can be calculated as
+its contribution to the parametric covariance between samples 
+summed over the upper triangle (excluding the diagonal) of the covariance matrix.
+The importance of the $k$th layer is its contribution relative to those of all other layers, 
+i.e., its contribution divided by the sum of the contributions of all $K$ layers
+(for details on calculating this importance, see Appendix \secref{layer_importance}). 
+This layer importance is similar to the ``ancestry contribution'' proposed by \cite{fastStructure}. 
+However, each of our layers can induce a different amount of covariance between samples embedded in them, 
+so we take that into account when calculating each layer's contribution to the whole.
 
@@
 then used cross-validation to compare the predictive performance of the models.
-In Fig \ref{simK1_pies}, we show the results of the nonspatial model 
-run with between 1 and 7 clusters, applied to the dataset simulated under $K=1$,
-and in Fig \ref{sim_xvals} we show the results of the cross-validation analysis for each of the three simulated datasets.
-The admixture pie plots for the spatial and nonspatial models run on the datasets simulated under $K=2$ and $K=3$
-are shown in Figs \ref{simK1_nsp_pies}-\ref{simK3_sp_pies}.
+In Fig \ref{simK1_pies}, we show the results of the analyses of the dataset simulated under $K=1$, 
+contrasting the results of the nonspatial model runs with 2 - 4 clusters 
+with those of spatial model with 2 - 4 layers.
+In Fig \ref{sim_xvals} we show the results of the cross-validation analysis for each of the three simulated datasets.
+The results of all analyses run on datasets simulated using $K=1$ through 3 are shown in 
+Figs \ref{simK1_nsp_pies}-\ref{simK3_adprop_fit}.
 To confirm that results from our nonspatial model were qualitatively similar to those of 
@@
 we also analyzed each simulated dataset using fastSTRUCTURE \citep{fastStructure}, 
-inferring admixture under models with $K=2$ through 4.
+inferring admixture under models with $K=2$ through 4 (Figs \ref{fastStr_simK1}-\ref{fastStr_simK3}).
 
@@
 Across all values of K and for all simulations, 
-the spatial model is correctly preferred over the nonspatial model, 
+the spatial model is correctly preferred over the nonspatial model (Fig \ref{sim_xvals})., 
 and it infers the true admixture proportions with high accuracy, tight precision, and good coverage 
-(Fig \ref{sim_xvals}).
+(Figs \ref{simK2_adprop_fit} and \ref{simK3_adprop_fit}).
 Second, the predictive accuracy of the nonspatial model increases 
@@
  (Fig \ref{fastStr_simK1}-\ref{fastStr_simK3}).
-While it is true that each such cluster is more similar within itself
+While it is true that each cluster is more similar within itself
 than it is to other clusters,
@@
 In contrast, the predictive accuracy of the spatial \texttt{conStruct} model
-either plateaus at or declines from the true value of $K$ used to simulate the data (Fig \ref{sim_xvals}, second column).
+either plateaus at or declines from the true value of $K$ used to simulate the data 
+(Figs \ref{sim_xvals}, \ref{simK1_xval}, \ref{simK2_xval}, and \ref{simK3_xval}).
 And, as $K$ in the spatial model increases, 
@@
 those estimated under the true value of $K$ for each simulation.
-
-\gb{could talk about proportion of relatedness explained by each layer here.
-e.g., Another way to visualize the difference in the behavior of the spatial and nonspatial models 
-is to summarize the proportion of relatedness between samples explained by each cluster...}
-
-\begin{figure}
-	\centering
-		\subcaptionbox{$K=2$\label{simK1_nsp_pies:K2}}
+These results are highlighted in the calculation of layer importance for 
+analyses run in each simulation (Fig \ref{simK1_laycon}, \ref{simK2_laycon}, and \ref{simK3_laycon}).
+In the spatial analyses run using the true value of $K$,
+the importance of the $K$ layers remains constant even with additional layers;
+in contrast, additional layers in the nonspatial analyses continue to be important for describing variation.
+
+\begin{figure}
+	\centering
+		\subcaptionbox{$K=2$\label{simK1_pies:nsp_K2}}
 			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K2.pdf}}
-		\subcaptionbox{$K=3$\label{simK1_nsp_pies:K3}}
+		\subcaptionbox{$K=3$\label{simK1_pies:nsp_K3}}
 			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K3.pdf}}
-		\subcaptionbox{$K=4$\label{simK1_nsp_pies:K4}}
+		\subcaptionbox{$K=4$\label{simK1_pies:nsp_K4}}
 			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K4.pdf}}
-		\subcaptionbox{$K=2$\label{simK1_sp_pies:K2}}
+		\subcaptionbox{$K=2$\label{simK1_pies:sp_K2}}
 			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K5.pdf}}
-		\subcaptionbox{$K=3$\label{simK1_sp_pies:K3}}
+		\subcaptionbox{$K=3$\label{simK1_pies:sp_K3}}
 			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K6.pdf}}
-		\subcaptionbox{$K=4$\label{simK1_sp_pies:K4}}
+		\subcaptionbox{$K=4$\label{simK1_pies:sp_K4}}
 			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K7.pdf}}
@@
 	\centering
-		{\includegraphics[width=6in,height=8.5in]{figs/sims/xvals.pdf}}
+		{\includegraphics[width=6in,height=2in]{figs/sims/sim_xvals.pdf}}
 	\caption{
@@
 	comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
-	The first panel in each row shows all results; 
-	the second panel shows results for just the spatial model.
 	The inset plots zoom in on cross-validation results outlined in the dotted boxes.
@@
 
-We address these questions using a dataset generated and originally published by \cite{geraldes_etal_2014}.
-The published data consist of 434 individuals sampled from 35 drainages 
+Our data, originally published by \cite{geraldes_etal_2014}, 
+consist of 434 individuals sampled from 35 drainages 
 and genotyped at just over 33,000 loci (map of the sampling shown in Fig \ref{populus_map}).
@@
 The results of these analyses are shown in 
-Figs \ref{populus_xvals}--\ref{populus_sp_clst_covs} and \ref{populus_nsp_clst_covs}. %,populus_nsp_pies,populus_sp_clst_covs,populus_nsp_clst_covs
+Figs \ref{populus_xvals} and \ref{populus_pies}, 
+as well as in Figs \ref{populus_sp_pies} - \ref{populus_laycon} in the Supplementary Materials.
 
@@
 	\centering
-		\subcaptionbox{$K=2$\label{populus_sp_pies:K2}}
+		\subcaptionbox{$K=2$\label{populus_pies:K2}}
 			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp2.pdf}}
-		\subcaptionbox{$K=3$\label{populus_sp_pies:K3}}
+		\subcaptionbox{$K=3$\label{populus_pies:K3}}
 			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp3.pdf}}
-		\subcaptionbox{$K=4$\label{populus_sp_pies:K4}}
+		\subcaptionbox{$K=4$\label{populus_pies:K4}}
 			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp4.pdf}}
-		\subcaptionbox{$K=5$\label{populus_sp_pies:K5}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp5.pdf}}
-		\subcaptionbox{$K=6$\label{populus_sp_pies:K6}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp6.pdf}}
-		\subcaptionbox{$K=7$\label{populus_sp_pies:K7}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp7.pdf}}
 	\caption{
 	Map of admixture proportions estimated for the \textit{Populus} dataset 
-	using the spatial model for $K=2$ through 7.
-    }\label{populus_sp_pies}
-\end{figure}
-
-\begin{figure}
-	\centering
-		\subcaptionbox{$K=2$\label{populus_nsp_pies:K2}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp2.pdf}}
-		\subcaptionbox{$K=3$\label{populus_nsp_pies:K3}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp3.pdf}}
-		\subcaptionbox{$K=4$\label{populus_nsp_pies:K4}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp4.pdf}}
-		\subcaptionbox{$K=5$\label{populus_nsp_pies:K5}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp5.pdf}}
-		\subcaptionbox{$K=6$\label{populus_nsp_pies:K6}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp6.pdf}}
-		\subcaptionbox{$K=7$\label{populus_nsp_pies:K7}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp7.pdf}}
-	\caption{
-	Map of admixture proportions estimated for the \textit{Populus} dataset 
-	using the nonspatial model for $K=2$ through 7.
-    }\label{populus_nsp_pies}
-\end{figure}
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=6in,height=4in]{figs/populus/populus_sp_clst_covs.pdf}}
-	\caption{
-	Plots showing the cluster-specific parametric covariance IBD-curves 
-	estimated for the \textit{Populus} data using 
-	the spatial \texttt{conStruct} model run with $K=2$ through 7.
-	Line colors are consistent with cluster colors in Fig \ref{populus_sp_pies}.
-	Points are colored by the species they are a covariance between:
-	black on black points are sample covariances between populations of \textit{Populus balsamifera};
-	green on black points are sample covariances between \bals{} and \tri{};
-	green on green points are sample covariances between \tri{} and \tri{}.
-    }\label{populus_sp_clst_covs}
+	using the spatial model for $K=2$ through 4.
+    }\label{populus_pies}
 \end{figure}
@@
 were experimentally shown to display different phenotypes 
-\citep[pathogen resistance][]{xie2009,xie2012}.  
+\citep[e.g., pathogen resistance,][]{xie2009,xie2012}.  
 At higher values of $K$, drainages at the southern tip of \tri{} sampling 
 begin to split out into their own layers.  
-\plr{cut this if we pick lower $K$? Or add a disclaimer?}
-This may potentially be the result of introgression from Southern \textit{Populus} species, 
+This may potentially be the result of introgression from southern \textit{Populus} species, 
 such as \textit{P. angustifolia} or \textit{fremontii} \citep{Zhou2012,geraldes_etal_2014}
 
-At all values of $K$, the spatial model was preferred over the nonspatial model.
-However, most of this signal seems to derive from the \textit{P. trichocarpa} samples:
+At all values of $K$, the spatial model was preferred over the nonspatial model (Fig \ref{populus_xvals}).
+Most of this signal seems to derive from the \textit{P. trichocarpa} samples:
 covariance within predominantly \tri{} layers decays more quickly with distance
 than within the majority \bals{} layer, as seen in Fig \ref{populus_sp_clst_covs}.  
-The most strongly supported model, given the current dataset, was XXXX.
+This indicates that, although isolation by distance is a feature of the dataset as a whole, 
+it's magnitude varies considerably across biological units - 
+an aspect of the data \texttt{conStruct} is able capture well.
 
 \subsubsection*{Black bears}
+The American black bear, \textit{Ursus americanus}, is endemic to North America
+and has a broad distribution across the continent.
+During the last glacial maximum, 
+black bears were confined to isolated glacial refugia, 
+from which they subsequently expanded to occupy their current range
+\citep{WoodingWard1997,Byun1997,Stone2000,Puckett2015}.
+The retreat of the glaciers was relatively recent in units of 
+black bear generation times \citep{Onorato2004}, 
+so we might reasonably expect the divergence accrued between refugia 
+to be reflected in discrete genetic structure between extant populations.
+Black bears are strong dispersers \citep{Moore2014}, 
+but they are distributed across most of the continent, 
+so we may therefore expect them to experience distance-limited 
+mating opportunities and exhibit a pattern of IBD.
+As with the poplars described above, 
+this system is likely to present patterns of genetic structure 
+that are both continuous and discrete, 
+making it well-suited as a test for our method.
+We organize the results of our analyses of 
+the complicated landscape of genetic structure in black bears 
+around the following questions:
+\begin{enumerate}
+\item How many historical clusters are reflected in modern patterns of genetic variation?
+\item Do we see support for a pattern of isolation by distance within each of the inferred discrete groups?
+\end{enumerate}
+
+Our data, originally published by \cite{Puckett2015}, consist of 95 individuals 
+sampled across the United States and on the West coast of Canada,
+genotyped at just under 22,000 bi-allelic loci (map of the sampling shown in Fig \ref{bear_map}).
+The distribution of missing data across these individuals was uneven, 
+with a few individuals representing most of the missing data, 
+so we removed individuals with greater than 4\% missing data, 
+resulting in a final dataset of 78 individuals.
+We then analyzed these data, treating individuals as the unit of analysis, 
+using both the spatial and nonspatial \texttt{conStruct} models with a $K$ of between 1 and 7, 
+and compared these models using cross-validation.
+The results of these analyses are shown in Figs \ref{bear_xvals} - \ref{bear_laycon}, 
+as well as in Figs \ref{bear_sp_pies} - \ref{bear_nsp_clst_covs} in the Supplementary Materials.
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=4in]{figs/bears/bear_sampling_map.pdf}}
+	\caption{
+	Map of the sampled black bear individuals included in the analysis.
+    }\label{bear_map}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=3in]{figs/bears/bear_std_xval.pdf}}
+	\caption{
+	Cross-validation results for black bear dataset 
+	comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
+	The first panel in each row shows all results; 
+	the second panel zooms in on the results from the spatial analyses run with $K = 4$ through 7.
+    }\label{bear_xvals}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\subcaptionbox{$K=3$ (spatial)\label{bear_K3:sp}}
+			{\includegraphics[width=2.5in,height=2.5in]{figs/bears/bear_sp3.pdf}}
+		\subcaptionbox{$K=3$ (nonspatial)\label{bear_K3:nsp}}
+			{\includegraphics[width=2.5in,height=2.5in]{figs/bears/bear_nsp3.pdf}}
+	\caption{
+	Map sof admixture proportions estimated for the black bear dataset 
+	using the spatial model (left) and the nonspatial model (right) for $K=3$.
+	Pies show mean admixture results across individuals within their diameter, 
+	and the admixture results for all individuals included within each group are 
+	shown in the plot above.
+    }\label{bear_K3}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=3in]{figs/bears/bears_laycon_barplots.pdf}}
+	\caption{
+	Layer ``contributions" (i.e., how much total covariance is contributed by each layer), 
+	for all layers estimated in runs using $K = 1$ through 7 
+	for the spatial model (left) and the nonspatial model (right).
+	For each value of $K$ along the x-axis, there are an equal number of layer contributions plotted.
+	Colors are consistent with Fig \ref{bear_K3}.
+    }\label{bear_laycon}
+\end{figure}
+
+Bears in the sample are partitioned into two main groups:
+one to the east of the Rocky Mountains, 
+which also occurs in Alaska, and, presumably, 
+in the intervening (unsampled) northern Great Plains and Canadian Shield;
+the other primarily west of the Rockies, 
+but presumably stretching up into British Columbia.
+Additional layers in the spatial model have strong statistical support 
+up until around $K=5$ or 6 (Fig \ref{bear_xvals}).
+However, upon examining layer importance (Fig \ref{bear_laycon}),
+it is clear that the additional layers beyond $K=2$ are contributing only a very small amount toward 
+total covariance between individuals.
+The locations of the individuals inferred to be admixed between the two layers 
+provide information about a possible zone of contact and migration between these groups 
+along the northwestern coast of North America and near the Cascade Range in the northern U.S.
+These findings are consistent with a scenario of postglacial expansion from two refugia, 
+one in the American Southwest and one in the American Southeast, 
+but lack of any samples from Canada and Mexico, 
+and lack of denser sampling across northern North America, 
+make more detailed interpretations untrustworthy.
+In addition, our model is not an explicit test of any demographic scenario, 
+and also uses a dataset that consists solely of nuclear (rather than mitochondrial) DNA.
+
+Results from the nonspatial model clearly exhibit the tendency 
+of nonspatial clustering algorithms to describe continuous spatial patterns of divergence 
+using gradients of admixture between clusters.
+For example, in Fig \ref{bear_K3:nsp}, 
+the third cluster (in green) exhibits a clear East-West gradient that 
+is overlain across the discrete structure between the Southwest cluster and the Southeast
+
+Across all values of $K$ for which we ran \texttt{conStruct},
+we see strong support for the spatial model over the nonspatial model (Fig \ref{bear_xvals}),
+indicating the isolation by distance seems to be a prominent feature of the data.
+This pattern may resolve a discrepancy between our results and 
+previous analyses that split Alaskan and British Columbian bears out 
+into their own cluster with an inferred Beringian glacial refugium
+\citep{Byun1997,Stone2000,Puckett2015}.
+Our model, which explicitly incorporates a spatial decay of relatedness, 
+allows individuals sampled at a large remove to belong to the same layer
+instead of splitting these individuals out into successive clusters.
 
@@
 
-The pattern of isolation by distance is ubiquitous in nature,
-unsuprisingly, since mating opportunities are distance-limited in most species.
-However,
-this common pattern can flummox commonly used clustering methods, 
+The pattern of isolation by distance is ubiquitous in nature;
+this is unsurprising, as mating opportunities are distance-limited in most species.
+This common pattern can flummox commonly used clustering methods, 
 which must use a possibly large number of discrete clusters
-to describe such continuous population differentiation .
+to describe such continuous population differentiation.
 Inference of these spurious clusters can have implications for 
@@
 Empirical applications of \texttt{conStruct} to a sample of North American poplars 
-and a sample of North American black bears yield reasonable results, 
-largely in agreement with previous studies in both systems (Geraldes, Puckett).
+and a sample of North American black bears yield reasonable results.
 
@@
 \paragraph{Comparison to nonspatial model-based clustering}
-Above, we \plr{hopefully?} showed that 
-(a) the nonspatial conStruct model recapitulates results of 
+Above, we showed that 
+(a) the nonspatial \texttt{conStruct} model recapitulates results of 
 other, commonly-used nonspatial clustering methods,
 and 
-(b) conStruct can concisely capture spatial structure, which is nearly always expected within clusters.
+(b) \texttt{conStruct} can concisely capture spatial structure, which is nearly always expected within clusters.
 Given this, when should methods without spatial capability be used?
-One advantage these have over conStruct is speed \plr{details?}.
+One advantage these have over \texttt{conStruct} is speed.
+Although \texttt{conStruct}'s computation time is invariant 
+with the number of loci included in the dataset 
+(after the initial calculation of the allelic covariance), 
+it scales poorly with number of samples.
+The computationally limiting step is the inversion of the parametric covariance, 
+which is an operation of order $O(n^3)$, 
+whereas computation time for, e.g., STRUCTURE, 
+scales linearly with number of samples. \gb{is this true?}
 The other main difference is the model of noise,
@@
 and also to take the results of the proposed cross-validation procedure with a large grain of salt.
+Calculating layer importance may also be a useful heuristic, 
+as it can reveal layers with strong statistical support but small biological import.
 
@@
 
-It is therefore important to emphasize that individuals sampled from the same \texttt{conStruct} cluster 
+It is therefore important to emphasize that individuals sampled from the same \texttt{conStruct} layer 
 may be quite genetically diverged from one another, 
-and that a \texttt{conStruct} cluster may still contain multiple distinct MUs.  
-Alternatively, the collapse of multiple, previously identified MUs into a single \texttt{conStruct} cluster 
+and that a \texttt{conStruct} layer may still contain multiple distinct MUs.  
+Alternatively, the collapse of multiple, previously identified MUs into a single \texttt{conStruct} layer 
 may reflect a biological reality that these populations are currently 
 (or were recently) exchanging migrants, 
-and thus might emphasize the importance of maintaining habitat corridors (ref) 
-between demes, or of implementing an integrated conservation plan across multiple demes within a cluster.
+and thus might emphasize the importance of maintaining habitat corridors between demes, 
+or of implementing an integrated conservation plan across multiple demes within a layer.
 
@@
 There are a few important caveats to consider in the interpretation of \texttt{conStruct} results. 
-First, we have modeled allelic covariance within a cluster as a spatial process.
+First, we have modeled allelic covariance within a layer as a spatial process.
 Although there is flexibility built into the model about the shape of that covariance, 
@@
 For example, if we were to run a \texttt{conStruct} analysis using geographic distances between 
-sampled individuals of greenish warblers (Irwin ref) or \textit{Ensatina} salamanders (Wake ref)
+sampled individuals of greenish warblers \citep{Irwin2001} or \textit{Ensatina} salamanders \citep{wake_schneider1998} 
 --- two canonical examples of rings species --- 
@@
 and to examine the spatial covariance functions within layers when interpreting results.
-\gb{missing data?}
 
@@
 (a star population phylogeny, although the branches can have different lengths thanks to the $\phi^{(k)}$ parameter), 
-similar to the F-model in (REF the right STRUCTURE paper).
+similar to the F-model in \citep{falush2003}.
 However, we could extend the existing model by implementing 
@@
 
-A second exciting direction is to model relatedness within a cluster as a spatiotemporal process, 
+Similarly, we currently assume that all loci are independent, 
+such that the number of loci in the dataset gives the degrees of 
+freedom of the Wishart distribution in our likelihood calculations.  
+Linkage disequilibrium (LD) between loci will decrease 
+the effective number of degrees of freedom, 
+artificially increasing users' confidence in their results.  
+One possible solution is to estimate an \emph{effective} number of loci 
+by introducing a parameter to modify the given degrees of freedom 
+and thereby model linkage between loci \citep[e.g.][]{EEMS}.
+
+A final exciting direction is to model relatedness within a cluster as a spatiotemporal process, 
 in which covariance decays both with distance in space and in time.  
@@
 it is becoming possible to ask whether there is genetic continuity at a point in space across time, 
-or whether populations are being replaced (yamnaya paper, slatkin and racimo, nielsen 2016, schraiber 2017).
+or whether populations are being replaced \citep{lazaridis_ancient_2014, Haak2015, slatkin_racimo2016, Nielsen2017, Schraiber2017}.
 However, we expect allele frequencies to change through time in a population, even without replacement, 
@@
 \section*{Appendix}
+\pagenumbering{gobble}
 \renewcommand{\theequation}{A\arabic{equation}}
@@
 is due to the drift in the ancestral population of their cluster 
-(the variance of which is $\phi_k$). 
+(the variance of which is $\phi^{(k)}$). 
 Under our spatial model we assume that some of the covariance in allele frequencies
@@
 the number of loci $L$ across which the sample allelic covariance is calculated:
-\gb{should we take this equation out?}
-\begin{align}
-&\widehat{\Omega}_{i,j} = 
-	\begin{cases}
-		\frac{1}{L-1}
-		\sum\limits_{\ell=1}^{L}{\left[ 	\left(f_{i,\ell} - \bar{f}_{i} \right)
-		 	\left(f_{j,\ell} - \bar{f}_{j} \right)\right] \notag
-			} & \text{for } i \neq j \\
-		\bar{f}_i (1-\bar{f}_i) & \text{for } i = j \\
-	\end{cases}\\
-&\widehat{\Omega} \sim \mathcal{W}\left( L\Omega, L	\right)
+\begin{equation}
+\widehat{\Omega} \sim \mathcal{W}\left( L\Omega, L	\right)
 \label{wishart}
-\end{align}
+\end{equation}
 
@@
 \delta_{i=j} \text{Var}(\Delta_i)
-\label{admixed_spatial_cov}
+\label{admixed_continuous_cov}
 \end{equation}
@@
 \section{Cross validation}\label{Xvalidation}
-To perform model comparison, we employ a Monte Carlo cross-validation approach \citep{picard1984},
-also known as repeated random sub-sampling validation.
+To perform model comparison, we employ a Monte Carlo cross-validation approach \citep{picard1984}.
+This cross-validation procedure generates a mean predictive accuracy for each model and each value of $K$, 
+as well as a confidence interval around that mean,
+which can then be used for model comparison or selection.
 Briefly, we follow the following procedure:
@@
 
-This cross-validation procedure generates a mean predictive accuracy for each model and each value of $K$, 
-as well as a confidence interval around that mean,
-which can then be used for model comparison or selection.
+\newpage
+\section{Calculating layer importance}\label{layer_importance}
+First, we describe the allelic covariance between two samples, $i$, and $j$.
+Note that this is the allelic covariance described in Eqn \eqref{allelic_covariance},
+in which we randomly polarize alleles,
+and not the genetic covariance, which subtracts mean allele frequencies.
+Let $A$ and $B$ be a randomly chosen alleles from the two samples, respectively,
+at a randomly chosen locus 
+(if the two populations are the same, then choose without replacement).
+Then, if we let $X=2(A-1/2)$ and $Y=2(B-1/2)$,
+since $X$ and $Y$ take the values $\pm 1$,
+and thanks to the random polarization, $\E[A] = \E[A] = 1/2$,
+$$\begin{aligned}
+\E[XY]
+        &= \P\{ X=Y \} - \P\{ X \neq Y \} \\
+        &= 2 \P\{ X=Y \} - 1 \\
+        &= 1 - 2 \P\{ X \neq Y \} .
+\end{aligned}$$
+To translate, $\P\{ X=Y \}$ is the probability that the alleles from our two focal samples agree with each other,
+while $\P\{ X \neq Y \}$ is the probability that they disagree.
+For the first thing to happen, $i$ and $j$ must coalesce before coalescing with the reference samples,
+so for $X$ and $Y$ to agree they must be more closely related than typical in the population.
+The last part, $\P\{ XY \neq 0\}$, is the probability that both samples disagree from the references.
+(This implies that $\E[XY] = 1 - 2 D_{ij}$, 
+where $D_{ij}$ is the probability that two randomly chosen alleles differ, 
+which is the genetic divergence.)
+
+Now, here is a generative model that gives us the form of the covariance we have postulated.
+To decide whether or not $X$ and $Y$ will agree,
+first each sample randomly chooses a layer: call these $I$ and $J$.
+The probability that $X$ chooses layer $k$ is $\P{I=k}$, which is $w_i^{(k)}$, 
+the $i$th sample's admixture proportion in the $k$th layer.
+The same holds true for $Y$.
+If they do not choose the same layer, the probability that they agree is $p_\gamma$.
+If they do choose the same layer, 
+then they agree with a probability $p_\gamma + q^{(k)}_{ij}$ that depends on their distance apart.
+Here
+
+$$\begin{aligned}
+    p_\gamma &= (2 (\gamma + \delta_{ij} \eta_i) + 1/2) \\
+    q^{(k)}_{ij} &= \left(2 \alpha_0^{(k)} \exp\left( - \left(\alpha_D^{(k)} D_{ij}\right)^{\alpha_2^{(k)}} \right) + 2 \mu^{(k)}  + 1/2 \right) .
+\end{aligned}$$
+
+One way to summarize the contribution of each layer is to partition the probability of agreement
+into contributions due to agreement ``in" each layer.
+So, the contribution from layer $k$ to agreement between $i$ and $j$ is 
+$w_i^{(k)} w_j^{(k)} q^{(k)}_{ij} / (p_\gamma + \sum_K w_i^{(k)} w_j^{(k)} q^{(k)}_{ij})$,
+which is the probability, given that they agree, that they agree thanks to layer $k$.
+Similarly, the ``background" contribution is $p_\gamma / (p_\gamma + \sum_K w_i^{(k)} w_j^{(k)} q^{(k)}_{ij})$.
+Because our signal comes from \emph{variation} in covariance, we omit the $p_\gamma$ terms.
+
+This suggests defining the overall contribution of layer $k$ to agreement, $\xi^{(k)}$, 
+to be the average of that quantity over $i$ and $j$.
+
+%
+\begin{equation}
+\xi^{(k)} = 
+	\sum\limits_{i=1}^{N-1}
+		\sum\limits_{j=i+1}^N
+			w^{(k)}_iw^{(k)}_j 
+				\left(	
+					2\alpha^{(k)}_0 
+						\left( \text{exp} 
+							\left( -(\alpha^{(k)}_D D_{i,j}) ^ {\alpha^{(k)}_2}\right) 
+						\right) + 2\phi^{(k)} + \frac{1}{2}
+				\right)
+\label{layer_contribution}
+\end{equation}
+%
+which is that layer's contribution to agreement between samples 
+summed over the upper triangle (excluding the diagonal) of the covariance matrix.
+We define the importance of the $k$th layer, $\Xi^{(k)}$, 
+as the relative contribution of the $k$th layer to total agreement:
+%
+\begin{equation}
+\Xi^{(k)} = 
+\frac{\xi^{(k)}}
+	{\sum\limits_{k=1}^{K}
+	    	\xi^{(k)}}
+\label{rel_layer_contribution}
+\end{equation}
+%
 
@@
 
+\begin{figure}
+	\centering
+		{\includegraphics[width=5in,height=2.5in]{figs/sims/simK1_laycon_barplots.pdf}}
+		\caption{
+			Layer ``importance" (i.e., how much total covariance is contributed by each layer), 
+			for all layers estimated in runs using $K = 1$ through 7 
+			for the spatial model (left) and the nonspatial model (right).
+			Data were simulated using $K=1$.
+			For each value of $K$ along the x-axis, there are an equal number of layer contributions plotted.
+		}\label{simK1_laycon}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=5in,height=2.5in]{figs/sims/simK1_std_xval.pdf}}
+		\caption{
+			Cross-validation results for data simulated under $K=1$,
+			comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
+			The right panel zooms in on just the spatial cross-validation results.
+		}\label{simK1_xval}
+\end{figure}
+
 \newpage
@@
 	\centering
+		{\includegraphics[width=5in,height=2.5in]{figs/sims/simK2_laycon_barplots.pdf}}
+		\caption{
+			Layer ``importance" (i.e., how much total covariance is contributed by each layer), 
+			for all layers estimated in runs using $K = 1$ through 7 
+			for the spatial model (left) and the nonspatial model (right).
+			Data were simulated using $K=2$.
+			For each value of $K$ along the x-axis, there are an equal number of layer contributions plotted.
+		}\label{simK2_laycon}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=5in,height=2.5in]{figs/sims/simK2_std_xval.pdf}}
+		\caption{
+			Cross-validation results for data simulated under $K=2$,
+			comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
+			The right panel zooms in on just the spatial cross-validation results.
+		}\label{simK2_xval}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=4in,height=4in]{figs/sims/simK2_adprop_fit.pdf}}
+		\caption{
+			Plot of \texttt{conStruct} ability to correctly estimate admixture proportions on simulated data.
+			Results are from an analysis with a spatial model using $K=2$.
+		}\label{simK2_adprop_fit}
+\end{figure}
+
+\begin{figure}
+	\centering
 		\subcaptionbox{$K=2$\label{simK3_nsp_pies:K2}}
@@
 	\centering
+		{\includegraphics[width=5in,height=2.5in]{figs/sims/simK3_laycon_barplots.pdf}}
+		\caption{
+			Layer ``importance" (i.e., how much total covariance is contributed by each layer), 
+			for all layers estimated in runs using $K = 1$ through 7 
+			for the spatial model (left) and the nonspatial model (right).
+			Data were simulated using $K=3$.
+			For each value of $K$ along the x-axis, there are an equal number of layer contributions plotted.
+		}\label{simK3_laycon}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=5in,height=2.5in]{figs/sims/simK3_std_xval.pdf}}
+		\caption{
+			Cross-validation results for data simulated under $K=3$,
+			comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
+			The right panel zooms in on just the spatial cross-validation results.
+		}\label{simK3_xval}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=4in,height=4in]{figs/sims/simK3_adprop_fit.pdf}}
+		\caption{
+			Plot of \texttt{conStruct} ability to correctly estimate admixture proportions on simulated data.
+			Results are from an analysis with a spatial model using $K=3$.
+		}\label{simK3_adprop_fit}
+\end{figure}
+
+\begin{figure}
+	\centering
 		{\includegraphics[width=6in,height=2.25in]{figs/fastStr/fastStr_simK1_pies.pdf}}
@@
 	\centering
+		\subcaptionbox{$K=2$\label{populus_sp_pies:K2}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp2.pdf}}
+		\subcaptionbox{$K=3$\label{populus_sp_pies:K3}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp3.pdf}}
+		\subcaptionbox{$K=4$\label{populus_sp_pies:K4}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp4.pdf}}
+		\subcaptionbox{$K=5$\label{populus_sp_pies:K5}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp5.pdf}}
+		\subcaptionbox{$K=6$\label{populus_sp_pies:K6}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp6.pdf}}
+		\subcaptionbox{$K=7$\label{populus_sp_pies:K7}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp7.pdf}}
+	\caption{
+	Map of admixture proportions estimated for the \textit{Populus} dataset 
+	using the spatial model for $K=2$ through 7.
+    }\label{populus_sp_pies}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\subcaptionbox{$K=2$\label{populus_nsp_pies:K2}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp2.pdf}}
+		\subcaptionbox{$K=3$\label{populus_nsp_pies:K3}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp3.pdf}}
+		\subcaptionbox{$K=4$\label{populus_nsp_pies:K4}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp4.pdf}}
+		\subcaptionbox{$K=5$\label{populus_nsp_pies:K5}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp5.pdf}}
+		\subcaptionbox{$K=6$\label{populus_nsp_pies:K6}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp6.pdf}}
+		\subcaptionbox{$K=7$\label{populus_nsp_pies:K7}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp7.pdf}}
+	\caption{
+	Map of admixture proportions estimated for the \textit{Populus} dataset 
+	using the nonspatial model for $K=2$ through 7.
+    }\label{populus_nsp_pies}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=4in]{figs/populus/populus_sp_clst_covs.pdf}}
+	\caption{
+	Plots showing the cluster-specific parametric covariance IBD-curves 
+	estimated for the \textit{Populus} data using 
+	the spatial \texttt{conStruct} model run with $K=1$ through 6.
+	Line colors are consistent with cluster colors in Fig \ref{populus_sp_pies}.
+	Points are colored by the species they are a covariance between:
+	black on black points are sample covariances between populations of \textit{Populus balsamifera};
+	green on black points are sample covariances between \bals{} and \tri{};
+	green on green points are sample covariances between \tri{} and \tri{}.
+    }\label{populus_sp_clst_covs}
+\end{figure}
+
+\begin{figure}
+	\centering
 		{\includegraphics[width=6in,height=4in]{figs/populus/populus_nsp_clst_covs.pdf}}
@@
 	estimated for the \textit{Populus} data using 
-	the nonspatial \texttt{conStruct} model run with $K=2$ through 7.
+	the nonspatial \texttt{conStruct} model run with $K=1$ through 6.
 	Line colors are consistent with cluster colors in Fig \ref{populus_nsp_pies}.
@@
 
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=3in]{figs/populus/populus_laycon_barplots.pdf}}
+	\caption{
+	Layer ``contributions" (i.e., how much total covariance is contributed by each layer), 
+	for all layers estimated in runs using $K = 1$ through 7 
+	for the spatial model (left) and the nonspatial model (right).
+	For each value of $K$ along the x-axis, there are an equal number of layer contributions plotted.
+	Colors are consistent with Figs \ref{populus_sp_pies}, \ref{populus_sp_clst_covs}, \ref{populus_nsp_pies}, and \ref{populus_nsp_clst_covs}.
+    }\label{populus_laycon}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\subcaptionbox{$K=2$\label{bear_sp_pies:K2}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_sp2.pdf}}
+		\subcaptionbox{$K=3$\label{bear_sp_pies:K3}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_sp3.pdf}}
+		\subcaptionbox{$K=4$\label{bear_sp_pies:K4}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_sp4.pdf}}
+		\subcaptionbox{$K=5$\label{bear_sp_pies:K5}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_sp5.pdf}}
+		\subcaptionbox{$K=6$\label{bear_sp_pies:K6}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_sp6.pdf}}
+		\subcaptionbox{$K=7$\label{bear_sp_pies:K7}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_sp7.pdf}}
+	\caption{
+	Map of admixture proportions estimated for the \textit{bear} dataset 
+	using the spatial model for $K=2$ through 7.
+    }\label{bear_sp_pies}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\subcaptionbox{$K=2$\label{bear_nsp_pies:K2}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_nsp2.pdf}}
+		\subcaptionbox{$K=3$\label{bear_nsp_pies:K3}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_nsp3.pdf}}
+		\subcaptionbox{$K=4$\label{bear_nsp_pies:K4}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_nsp4.pdf}}
+		\subcaptionbox{$K=5$\label{bear_nsp_pies:K5}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_nsp5.pdf}}
+		\subcaptionbox{$K=6$\label{bear_nsp_pies:K6}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_nsp6.pdf}}
+		\subcaptionbox{$K=7$\label{bear_nsp_pies:K7}}
+			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_nsp7.pdf}}
+	\caption{
+	Map of admixture proportions estimated for the \textit{bear} dataset 
+	using the nonspatial model for $K=2$ through 7.
+    }\label{bear_nsp_pies}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=4in]{figs/bears/bear_sp_clst_covs.pdf}}
+	\caption{
+	Plots showing the cluster-specific parametric covariance IBD-curves 
+	estimated for the black bear data using 
+	the spatial \texttt{conStruct} model run with $K=1$ through 6.
+	 }\label{bear_sp_clst_covs}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=4in]{figs/bears/bear_nsp_clst_covs.pdf}}
+	\caption{
+	Plots showing the cluster-specific parametric covariances 
+	estimated for the black bear data using 
+	the nonspatial \texttt{conStruct} model run with $K=1$ through 6.
+    }\label{bear_nsp_clst_covs}
+\end{figure}
+
 \newpage
@@
 
-\pagenumbering{gobble}
-
 \end{document}

### 2017-08-30 writeup/conStruct.tex
dropped sampling maps

@@
 	\centering
-		{\includegraphics[width=6in,height=5in]{figs/populus/populus_sampling_map.pdf}}
-	\caption{
-	Map of the sampled \textit{Populus} populations included in the analysis.
-	The color of the sampling location denotes the putative species.
-    }\label{populus_map}
-\end{figure}
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=6in,height=3in]{figs/populus/populus_std_xval.pdf}}
-	\caption{
-	Cross-validation results for \textit{Populus} dataset 
-	comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
-	The first panel in each row shows all results; 
-	the second panel zooms in on the results from analyses run with $K = 2$ through 7.
-    }\label{populus_xvals}
-\end{figure}
-
-\begin{figure}
-	\centering
 		\subcaptionbox{$K=2$\label{populus_pies:K2}}
@@
     }\label{populus_pies}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=3in]{figs/populus/populus_std_xval.pdf}}
+	\caption{
+	Cross-validation results for \textit{Populus} dataset 
+	comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
+	The first panel in each row shows all results; 
+	the second panel zooms in on the results from analyses run with $K = 2$ through 7.
+    }\label{populus_xvals}
 \end{figure}
@@
 sampled across the United States and on the West coast of Canada,
-genotyped at just under 22,000 bi-allelic loci (map of the sampling shown in Fig \ref{bear_map}).
+genotyped at just under 22,000 bi-allelic loci.
 The distribution of missing data across these individuals was uneven, 
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/bears/bear_sampling_map.pdf}}
-	\caption{
-	Map of the sampled black bear individuals included in the analysis.
-    }\label{bear_map}
-\end{figure}
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=6in,height=3in]{figs/bears/bear_std_xval.pdf}}
-	\caption{
-	Cross-validation results for black bear dataset 
-	comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
-	The first panel in each row shows all results; 
-	the second panel zooms in on the results from the spatial analyses run with $K = 4$ through 7.
-    }\label{bear_xvals}
-\end{figure}
-
-\begin{figure}
-	\centering
 		\subcaptionbox{$K=3$ (spatial)\label{bear_K3:sp}}
@@
     }\label{bear_K3}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=3in]{figs/bears/bear_std_xval.pdf}}
+	\caption{
+	Cross-validation results for black bear dataset 
+	comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
+	The first panel in each row shows all results; 
+	the second panel zooms in on the results from the spatial analyses run with $K = 4$ through 7.
+    }\label{bear_xvals}
 \end{figure}
@@
 	\centering
+		{\includegraphics[width=6in,height=5in]{figs/populus/populus_sampling_map.pdf}}
+	\caption{
+	Map of the sampled \textit{Populus} populations included in the analysis.
+	The color of the sampling location denotes the putative species.
+    }\label{populus_map}
+\end{figure}
+
+\begin{figure}
+	\centering
 		\subcaptionbox{$K=2$\label{populus_sp_pies:K2}}

### 2017-09-05 writeup/conStruct.tex
G's comments

--- 2017-08-30 writeup/conStruct.tex
@@
 \begin{abstract}
-One of the classic problems in population genetics is the characterization 
-of discrete population structure when the genotyped samples also show 
+\gc{A classic problem in population genetics is the characterization 
+of discrete population structure in the presence of} 
 continuous patterns of genetic differentiation.
@@
 the fact that diversity shows both discrete and continuous patterns. 
-For example, there is disagreement on species concepts because the
+For example, \gc{reasonable people can disagree whether two
+  populations are separate species, because the} %there is disagreement on species concepts because the
 process of speciation is usually gradual, and so there is no set point
 in the continuous divergence of populations when they unambiguously become discrete species.
-This issue extends below the species level, as patterns of phenotypic
+\gc{The issue of identifying meaningful biological subunits} extends below the species level, as patterns of phenotypic
 and genetic diversity within and among populations are shaped
@@
 informing conservation priorities \citep{Moritz1994,Waples_1998,Moritz_etal_2002}.
-Furthermore, we often need to identify sets of indivduals resulting from reasonably coherent
+Furthermore, we often need to identify \gc{subsets} of indivduals resulting from reasonably coherent
 evolutionary histories for downstream analyses to learn about population history and adaptation.
@@
 including generating population phylogenies \citep{CavalliSforza1975, treemix},
-$k$-means clustering analyses performed on dimensionality-reduction approaches \citep{meirmans2009genodive}, 
+\gc{dimensionality-reduction} approaches \citep{meirmans2009genodive},   %$k$-means clustering analyses performed
 such as principal components analysis \citep{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
@@
 Here, we will focus on model-based clustering, the most widely used class of approaches for population delineation.
+\gc{We need more on CLUSTERING approaches here. SOMETHING LIKE:
+These clustering methods model each individual's genotypes as random
+draws from a set of underlying, unobserved
+population clusters, each with a characteristic set of allele
+frequencies. 
+These underlying frequencies are identical for all individuals
+assigned to a cluster, regardless of their spatial location. 
+Spatial information has been incorporated in some of these methods, for
+example in placing priors on cluster memberships \citep{geneland}, but
+this does not address the underlying issue that allele frequencies are
+assumed constant in a cluster across the species range.  
+}
 Like the method we present here, these are based on a statistical model,
@@
 
-A pattern of isolation by distance (IBD) arises when migration in a population is spatially limited \citep{Wright1943},
-and has been modeled in both spatially continuous frameworks \citep{malecot1969mathematics,Slatkin1985}, 
+A pattern of isolation by distance (IBD) arises when migration in a
+population is spatially limited, \gc{and slow enough betweeen distant
+locations that genetic drift can build up differentiation}
+\citep{Wright1943}. IBD has been modeled in both spatially continuous frameworks \citep{malecot1969mathematics,Slatkin1985}, 
 in which an individual's parents are found within some radius of that individual's location,
@@
 This problem can be compounded if sampling is done unevenly or discretely across a population or species' range,
-and has given rise to a debate in the population genetic literature about whether 
-individuals are best described as being drawn from continuous clines or discrete clusters 
+and has given rise to a debate in the population genetic literature
+about \gc{how best to describe sets of individuals using continuous clines and} discrete clusters 
 \citep[e.g.][]{SerrePaabo2004,rosenberg2005clines}.
@@
 The model works with allele frequencies at $L$ unlinked, bi-allelic single nucleotide polymorphisms (SNPs) genotyped across $N$ samples.
-Each ``sample'' may be a collection of diploids, or frequencies estimated from pooled sequencing.
+Each ``sample'' may \gc{be a single individual, a collection of
+  individuals from a location}, or frequencies estimated from pooled sequencing.
 We write $N_i$ for the number of chromosomes sequenced in the $i^\text{th}$ sample.
@@
 of a randomly chosen allele in samples $i$ and $j$, as long as $i \neq j$.
-The diagonal (where $i=j$) does not have this interpretation.
-
-In addition to these differences on the diagonal,
-this differs from the usual ``genetic covariance'' \citep{mcvean_genealogical_2009}
+The diagonal (where $i=j$) does not have this interpretation, \gc{and
+  reflects the covariance between indivdual alleles drawn from within
+  the population}.
+
+\gc{This definition of convariance} differs from the usual ``genetic covariance'' \citep{mcvean_genealogical_2009}
 in that (a) we do not subtract locus means,
@@
 Therefore, unlike genetic covariance,
-this allelic covariance is affected in shape by singleton sites,
+this allelic covariance is \gc{more} affected in shape by singleton sites,
 so it may be advisable to filter these prior to analysis
@@
 We imagine each layer as a geographically distributed population,
-and that each sample is composed of a mixture of contributions from each of these layers.
+\gc{each layer extends over the entire sampled range of the
+  populations}. 
+Each sample is composed of a mixture of contributions from each of
+these layers. \gc{See Figure \ref{schematic} for a cartoon representation
+of our model.} 
 These layers thus take the place of ``clusters'' or ``populations'' in clustering methods,
-but we do not adopt these terms as they connote a lack of geographical separation within each.
+but we do not adopt these terms as \gc{terms like ``spatial'' cluster
+  suggest a clustering in space while our layers (and samples from
+  them) may be present at all sampled locations.}
+
+%they connote a lack of geographical separation within each.
 
@@
 The quantity $D_{i,j}$ is the observed geographic distance between samples $i$ and $j$,
-the $\alpha^{(k)}$ parameters control the shape of the decay of covariance with distance in the cluster,
-and $\phi^{(k)}$ is a parameter that describes the background covariance within the cluster. 
+the $\alpha^{(k)}$ parameters control the shape of the decay of
+covariance with distance in the cluster.
+\gc{Our choice of a powered-exponential decay, as parameterized by the
+  $\alpha$s, is a flexible, and standard choice in spatial statistics
+  REF, and is not chosen to match  a particular population genetics model. 
+The} $\phi^{(k)}$ is a parameter that describes the background covariance within the cluster. 
 If two samples draw 100\% of their ancestry from cluster $k$, then their covariance under the model is $G^{(k)}_{i,j}$;
@@
 connecting the $k$th population to the population ancestral to all modeled
-layers \citep{patterson_ancient_2012, peter_fstats}.
+layers \citep[see for example ][]{patterson_ancient_2012, peter_fstats}.
 
@@
 The ``admixture'' proportion of the $i$th sample in the $k$th cluster, denoted $w^{(k)}_i$,
-gives the proportion of alleles from sample $i$ that derive from
+gives the \gc{genome-wide} proportion of alleles from sample $i$ that derive from
 cluster $k$ (and so $\sum_K w^{(k)}_i =1$).

### 2017-09-06 code/conStruct/R/model.comparison.R
documented pkg

@@ -1,24 +1,205 @@
+#' Run a conStruct cross-validation analysis
+#' 
+#' \code{x.validation} runs a conStruct cross-validation analysis
+#' 
+#' This function initiates a cross-validation analysis that 
+#' uses Monte Carlo cross-validation to determine the statistical 
+#' support for models with different numbers of clusters or 
+#' with and without a spatial component.
+#' 
+#' @param train.prop A numeric value between 0 and 1 that gives 
+#'		the proportions of the data to be used in the 
+#'		training partition of the analysis. Default is 0.9.
+#' @param n.reps An \code{integer} giving the number of cross-
+#'		validation replicates to be run.
+#' @param K A numeric \code{vector} giving the numbers of clusters 
+#'		to be tested in each cross-validation replicate.
+#'		E.g., \code{K=1:7}.
+#' @param freqs A \code{matrix} of allele frequencies with one column per 
+#'		locus and one row per sample.
+#' 		Missing data should be indicated with \code{NA}.
+#' @param geoDist A \code{matrix} of geographic distance between samples. 
+#'		If \code{NULL}, user can only run the nonspatial model.
+#' @param coords A \code{matrix} giving the longitude and latitude 
+#'		(or X and Y coordinates) of the samples.
+#' @param prefix A character \code{vector} giving the prefix to be attached 
+#'		to all output files.
+#' @param n.iter An \code{integer} giving the number of iterations each MCMC 
+#'		chain is run. Default is 1e3.  If the number of iterations 
+#'		is greater than 500, the MCMC is thinned so that the number 
+#'		of retained iterations is 500 (before burn-in).
+#' @param make.figs A \code{logical} value indicating whether to automatically 
+#'		make figures during the course of the cross-validation analysis. 
+#'		Default is \code{FALSE}.
+#' @param save.files A \code{logical} value indicating whether to automatically 
+#'		save output and intermediate files once the analysis is
+#'		complete. Default is \code{FALSE}.
+#'
+#' @return This function returns a \code{list} containing the 
+#'		standardized results of the cross-validation analysis
[…]

### 2017-09-06 code/conStruct/R/plot.output.R

@@ -1,86 +1,235 @@
+#' Make output plots
+#'
+#' \code{make.all.the.plots} makes figures from the output from a 
+#' 	conStruct analysis.
+#'
+#' This function takes the output from a conStruct analysis and 
+#' generates a number of plots for visualizing results and 
+#' diagnosing MCMC performance.
+#'
+#' @param conStruct.results The list output by a 
+#'			\code{conStruct} run.
+#' @param data.block A \code{data.block} list saved during a 
+#'			\code{conStruct} run.
+#' @param prefix A character vector to be prepended to all figures.
+#' @param cluster.colors A \code{vector} of colors to be used in 
+#'			plotting results for different clusters. Users must 
+#'			specify one color per cluster.  If \code{NULL}, plots 
+#'			will use a pre-specified vector of colors.
+#' @return This function has only invisible return values.
+#'
+#'	@details This function produces a variety of plots that can be 
+#'	useful for visualizing results or diagnosing MCMC performance. 
+#'  The plots made are by no means an exhaustive, and users are 
+#' 	encouraged to make further plots, or customize these plots as they 
+#'	see fit.  For each plot, one file is generated for each MCMC chain 
+#'	(specified withe the \code{n.chains} argument in the function 
+#'	\code{conStruct}. The plots generated (as .pdf files) are:
+#'	\itemize{
+#'		\item Structure plot - STRUCTURE-style plot, where each sample 
+#'			is represented as a stacked bar plot, and the length of the 
+#'			bar plot segments of each color represent that sample's 
+#'			admixture proportion in that cluster. Described further 
+#'			in the help page for \code{make.structure.plot}.
+#'		\item Admixture pie plot - A map of samples in which each sample's 
+#'				location is denoted with a pie chart, and the proportion 
+#'				of a pie chart of each color represents that sample's 
+#'				admixture in each cluster. Described further in the help 
+#'				page for \code{make.admix.pie.plot}
+#'		\item model.fit.CIs - A plot of the sample allelic covariance 
+#'			shown with the 95\% credible interval of the parametric 
+#'			covariance for each entry in the matrix.
+#'		\item cluster.covariances - A plot of the cluster-specific 
[…]

### 2017-09-06 code/conStruct/R/stan.model.blocks.R

@@ -92,13 +92,13 @@ model {
 
 multiK.stan.block <- "
 functions {
-	matrix admixed_covariance(int N, int K, matrix w_mat, vector nugget, vector mu, real gamma) {
+	matrix admixed_covariance(int N, int K, matrix w_mat, vector nugget, vector phi, real gamma) {
 		matrix[N,N] parCov;
 		matrix[N,N] Nug_mat;
 		parCov = rep_matrix(0,N,N);
 		Nug_mat = diag_matrix(nugget);
 		for(k in 1:K){
-			parCov = parCov + tcrossprod(to_matrix(w_mat[,k])) * mu[k];
+			parCov = parCov + tcrossprod(to_matrix(w_mat[,k])) * phi[k];
 		}
 		parCov = gamma + parCov + Nug_mat;
 		return parCov;	
@@ -125,7 +125,7 @@ data {
 	real<lower=0,upper=1> temp;			// temperature parameter for estimating marginal likelihood	
 }
 parameters {
-	positive_ordered[K] mu;				// shared drift effect in cluster k
+	positive_ordered[K] phi;				// shared drift effect in cluster k
 	real<lower=0> gamma;				// covariance between all pairs of clusters
   	vector<lower=0>[N] nugget; 			// sample-specific variance (allele sampling error + sample-specific drift)
 	simplex[K]    w[N];    				// every sample (N in total) has a K simplex (i.e. K clusters)
@@ -136,11 +136,11 @@ transformed parameters {
 	vector[K] dirConPar;
 	dirConPar = rep_vector(0.1,K);	
 	w_mat = make_w_matrix(N,K,w);
-	parCov = admixed_covariance(N, K, w_mat, nugget, mu, gamma);
+	parCov = admixed_covariance(N, K, w_mat, nugget, phi, gamma);
 }
 model {
 	nugget ~ normal(0,1);										// prior on nugget
-	mu ~ normal(0,1);
+	phi ~ normal(0,1);
 	gamma ~ normal(varMeanFreqs,0.5);
 	for(i in 1:N) w[i] ~ dirichlet(dirConPar);				    // prior on admixture proportions
 	target += dWish(temp, L, obsCov, parCov);					// likelihood function
@@ -149,23 +149,23 @@ model {
 
 space.multiK.stan.block <- "
 functions {
-	matrix spCov(int N, real a0, real aD, real a2, matrix D, real mu){
[…]

### 2017-09-06 code/conStruct/R/run.conStruct.R
documented pkg

@@ -1,18 +1,118 @@
-#'@export
-conStruct <- function(spatial=TRUE,K,freqs,geoDist=NULL,temp=NULL,coords,prefix="",n.chains=1,n.iter=1e3,make.figs=TRUE,save.files=TRUE){
+#' Run a conStruct analysis.
+#'
+#' \code{conStruct} runs a conStruct analysis of genetic data.
+#'
+#' This function initiates an analysis that uses  
+#' geographic and genetic relationships between samples 
+#' to estimate sample membership (admixture proportions) across 
+#' a user-specified number of layers.
+#'
+#' @param spatial A logical indicating whether to perform a spatial analysis. 
+#' 				  Default is \code{TRUE}. 
+#' @param K An \code{integer} that indicates the number of layers to be 
+#' 				  included in the analysis.
+#' @param freqs A \code{matrix} of allele frequencies with one column per 
+#'				locus and one row per sample.
+#' 				Missing data should be indicated with \code{NA}.
+#' @param geoDist A \code{matrix} of geographic distance between samples. 
+#'					If \code{NULL}, user can only run the nonspatial model.
+#' @param coords A \code{matrix} giving the longitude and latitude 
+#'					(or X and Y coordinates) of the samples.
+#' @param prefix A character \code{vector} giving the prefix to be attached 
+#'					 to all output files.
+#' @param n.chains An integer indicating the number of MCMC chains to be run 
+#'					in the analysis. Default is 1.
+#' @param n.iter An \code{integer} giving the number of iterations each MCMC 
+#'				 chain is run. Default is 1e3.  If the number of iterations 
+#'				 is greater than 500, the MCMC is thinned so that the number 
+#'				 of retained iterations is 500 (before burn-in).
+#' @param make.figs A \code{logical} value indicating whether to automatically 
+#'					make figures once the analysis is complete. Default is 
+#'					\code{TRUE}.
+#' @param save.files A \code{logical} value indicating whether to automatically 
+#'						save output and intermediate files once the analysis is
+#'						 complete. Default is \code{TRUE}.
+#'
+#' @return This function returns a list with one entry for each chain run 
+#'			(specified with \code{n.chains}). The entry for each chain is named 
+#'			"chain_X" for the Xth chain.  The components of the entries for each 
+#'			are detailed below: 
[…]

### 2017-09-06 code/conStruct/R/process.model.fit.R
documented pkg

@@ -1,15 +1,15 @@
 get.conStruct.results <- function(data.block,model.fit,n.chains){
-	conStruct.results <- setNames(
-						lapply(1:n.chains,
-							function(i){
-								get.conStruct.chain.results(data.block,model.fit,i)
-							}),
-					  paste0("chain_",1:n.chains))
+	conStruct.results <- stats::setNames(
+							lapply(1:n.chains,
+								function(i){
+									get.conStruct.chain.results(data.block,model.fit,i)
+								}),
+						  paste0("chain_",1:n.chains))
 	return(conStruct.results)
 }
 
 get.MAP.iter <- function(model.fit,chain.no){
-	lpd <- get_logposterior(model.fit)
+	lpd <- rstan::get_logposterior(model.fit)
 	MAP.iter <- lapply(lpd,which.max)[[chain.no]]
 	return(MAP.iter)
 }
@@ -19,9 +19,9 @@ get.admix.props <- function(model.fit,chain.no,N,n.clusters){
 	admix.props <- array(1,dim=c(model.fit@sim$n_save[chain.no],N,n.clusters))
 	if(any(grepl("w",model.fit@model_pars))){
 		for(k in 1:n.clusters){
-			admix.props[,,k] <- extract(model.fit,
-										pars=unlist(lapply(1:N,function(j){sprintf("w[%s,%s]",j,k)})),
-										permuted=FALSE,inc_warmup=TRUE)[,chain.no,]
+			admix.props[,,k] <- rstan::extract(model.fit,
+											pars=unlist(lapply(1:N,function(j){sprintf("w[%s,%s]",j,k)})),
+											permuted=FALSE,inc_warmup=TRUE)[,chain.no,]
 		}
 	}
 	return(admix.props)
@@ -32,19 +32,19 @@ get.par.cov <- function(model.fit,chain.no,N){
 	for(i in 1:N){
 		for(j in 1:N){
 			my.par <- sprintf("parCov[%s,%s]",i,j)
-			par.cov[,i,j] <- extract(model.fit,pars=my.par,inc_warmup=TRUE,permuted=FALSE)[,chain.no,]
+			par.cov[,i,j] <- rstan::extract(model.fit,pars=my.par,inc_warmup=TRUE,permuted=FALSE)[,chain.no,]
 		}
 	}
 	return(par.cov)
 }
 
 get.nuggets <- function(model.fit,chain.no,N){
[…]

### 2017-09-06 writeup/figs/fig_funcs.R
added fig functions

@@ -418,3 +418,36 @@ make.bear.redux.result.plot <- function(csr,coords,lump.dist,cluster.colors,clus
 			 lwd=0.5,
 			 col = adjustcolor(1,0.6))
 }
+
+match.clusters.x.runs <- function(admix.mat1,admix.mat2,admix.mat1.order=NULL){
+	cluster.colors <- c("blue","red","green","yellow","purple","orange","lightblue","darkgreen","lightblue","gray")
+	K1 <- ncol(run1$MAP$admix.proportions)
+		if(!is.null(run1.order)){
+			run1$MAP$admix.proportions <- run1$MAP$admix.proportions[,run1.order]
+		}
+		K1.cols <- cluster.colors[1:K1]
+	K2 <- ncol(run2$MAP$admix.proportions)
+		K2.cols <- numeric(K2)
+	k.combn <- expand.grid(1:K1,1:K2)
+	clst.sims <- unlist(lapply(1:nrow(k.combn),
+						function(n){
+							measure.frob.similarity(run1$MAP$admix.proportions[,k.combn[n,1],drop=FALSE],
+													run2$MAP$admix.proportions[,k.combn[n,2],drop=FALSE],
+													K=1)
+							}))
+	while(length(which(K2.cols == 0)) > (K2-K1)){
+		tmp.max <- which.max(rank(clst.sims,na.last=FALSE))
+		run2.match <- k.combn[tmp.max,2]
+		run1.match <- k.combn[tmp.max,1]
+		K2.cols[run2.match] <- K1.cols[run1.match]
+		clst.sims[which(k.combn[,1]==run1.match)] <- NA
+		clst.sims[which(k.combn[,2]==run2.match)] <- NA
+	}
+	if(K2 > K1){
+		K2.cols[which(K2.cols==0)] <- cluster.colors[(K1+1):K2]
+	}
+	K2.clst.order <- unique(match(cluster.colors,K2.cols)[which(!is.na(match(cluster.colors,K2.cols)))])
+	clst2.info <- list("cols" = K2.cols,
+					   "clst.order" = K2.clst.order)
+	return(clst2.info)
+}

### 2017-09-06 writeup/conStruct.tex

@@
 connecting the $k$th population to the population ancestral to all modeled
-layers \citep[see for example ][]{patterson_ancient_2012, peter_fstats}.
+layers \citep[see for example ][]{patterson_ancient_2012,
+  peter_fstats}. In the methods section \ref{rationale} we lay out a
+simple model of allele frequencies underlying this covariance model.
 
@@
 \paragraph{Likelihood and inference}
-If allele frequencies at each locus
+\gc{Assuming that the} allele frequencies at each locus
 were independent between loci and multivariate normally distributed across populations, 
@@
 \end{equation}
-Linkage disequilibrium (LD) between loci will decrease the effective number of degrees of freedom, 
-a complication we discuss further below.
+Linkage disequilibrium (LD) between loci will decrease the effective
+number of degrees of freedom.  \gc{One possible solution, that we have
+  not implemented yet, is to estimate an \emph{effective} number of loci 
+by introducing a parameter to modify the given degrees of freedom 
+and thereby informally model linkage between loci \citep[e.g.][]{EEMS}.}
 
@@
 where $P(w)$, $P(\alpha)$, $P(\phi)$, $P(\eta)$, and $P(\gamma)$, are prior distributions.
-All parameters are given (half-)Gaussian priors except for $\alpha_2$, which is uniform on $(0,2)$ and $w$, which is Dirichlet
+All parameters are given (half-)Gaussian priors except for $\alpha_2$,
+which is uniform on $(0,2)$ and $w$, which is an \gc{independent Dirichlet, of
+  dimension $K$, for each individual}
 (see Table \ref{tab:param_prior_tab} for specifics).
@@
 --- spatial vs. nonspatial, across different values of $K$ --- 
-by comparing their performance in a common framework.
+by comparing their performance in a common framework. \gc{We also
+  validate our claim that the results of our non-spatial method are a
+  close match to those of STRUCTURE-based approaches.}
 
 \paragraph{Choice of cluster number and cross-validation}
-There are a number of reasons, both statistical and biological, 
-why there is no true (or right) number of layers for real datasets,
+There are a number of \gc{reasons 
+why there is no true (or right)} number of layers for real datasets,
 discussed further in the Discussion.
@@
 A simple $t$-test can be used to assess whether additional layers
-show statistically significantly greater predictive accuracy.
+show statistically significantly greater predictive accuracy. \gc{DO
+  WE DO t-test? IF NOT DROP THIS'}
 However, we remind users that ``statistical significance does not imply real-world significance'',
@@
 a relatedness structure between the layers by, for example, 
-estimating a population phylogeny between them \citep[e.g.][]{treemix}.
-
-Similarly, we currently assume that all loci are independent, 
-such that the number of loci in the dataset gives the degrees of 
-freedom of the Wishart distribution in our likelihood calculations.  
-Linkage disequilibrium (LD) between loci will decrease 
-the effective number of degrees of freedom, 
-artificially increasing users' confidence in their results.  
-One possible solution is to estimate an \emph{effective} number of loci 
-by introducing a parameter to modify the given degrees of freedom 
-and thereby model linkage between loci \citep[e.g.][]{EEMS}.
+estimating a population phylogeny between them
+\citep[e.g.][]{treemix}.
+
+%%Similarly, we currently assume that all loci are independent, 
+%such that the number of loci in the dataset gives the degrees of 
+%freedom of the Wishart distribution in our likelihood calculations.  
+%Linkage disequilibrium (LD) between loci will decrease 
+%the effective number of degrees of freedom, 
+%artificially increasing users' confidence in their results.
 
@@
 \section{Model rationale}
-\subsection{Drift, admixture, and space}
+\subsection{Drift, admixture, and space} \label{rationale}
 \paragraph{Drift} 

### 2017-09-06 writeup/conStruct.tex
edits on graham's edits

@@
 \begin{abstract}
-\gc{A classic problem in population genetics is the characterization 
-of discrete population structure in the presence of} 
+A classic problem in population genetics is the characterization 
+of discrete population structure in the presence of 
 continuous patterns of genetic differentiation.
@@
 the fact that diversity shows both discrete and continuous patterns. 
-For example, \gc{reasonable people can disagree whether two
-  populations are separate species, because the} %there is disagreement on species concepts because the
+For example, reasonable people can disagree about whether two 
+populations are separate species, because the 
 process of speciation is usually gradual, and so there is no set point
 in the continuous divergence of populations when they unambiguously become discrete species.
-\gc{The issue of identifying meaningful biological subunits} extends below the species level, as patterns of phenotypic
+The issue of identifying meaningful biological subunits 
+extends below the species level, as patterns of phenotypic
 and genetic diversity within and among populations are shaped
@@
 informing conservation priorities \citep{Moritz1994,Waples_1998,Moritz_etal_2002}.
-Furthermore, we often need to identify \gc{subsets} of indivduals resulting from reasonably coherent
+Furthermore, we often need to identify subsets of indivduals resulting from reasonably coherent
 evolutionary histories for downstream analyses to learn about population history and adaptation.
@@
 including generating population phylogenies \citep{CavalliSforza1975, treemix},
-\gc{dimensionality-reduction} approaches \citep{meirmans2009genodive},   %$k$-means clustering analyses performed
-such as principal components analysis \citep{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
-and model-based clustering approaches \citep[e.g.][]{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, FINESTRUCTURE, fastStructure, huelsenbeck2007inference, Corander2003,TESS,geneland}, among others.
+dimensionality-reduction approaches \citep{meirmans2009genodive}, 
+such as principal components analysis 
+\citep{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
+and model-based clustering approaches 
+\citep[e.g.][]{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, 
+FINESTRUCTURE, fastStructure, huelsenbeck2007inference, 
+Corander2003,TESS,geneland}, among others.
 Here, we will focus on model-based clustering, the most widely used class of approaches for population delineation.
-\gc{We need more on CLUSTERING approaches here. SOMETHING LIKE:
-These clustering methods model each individual's genotypes as random
-draws from a set of underlying, unobserved
-population clusters, each with a characteristic set of allele
-frequencies. 
-These underlying frequencies are identical for all individuals
-assigned to a cluster, regardless of their spatial location. 
-Spatial information has been incorporated in some of these methods, for
-example in placing priors on cluster memberships \citep{geneland}, but
-this does not address the underlying issue that allele frequencies are
-assumed constant in a cluster across the species range.  
-}
+Existing model-based clustering methods model each individual's genotypes 
+as random draws from a set of underlying, unobserved population clusters, 
+each with a characteristic set of allele frequencies, which are estimated. 
+These underlying frequencies are identical 
+for all individuals assigned to a cluster, 
+\emph{regardless} of their spatial location. 
+Spatial information has been incorporated into some of these methods, 
+by, for example, placing spatial priors on cluster memberships \citep{geneland}, 
+but this does not address the underlying issue that 
+these methods assume that allele frequencies 
+are constant in a cluster across the species' range.  
 Like the method we present here, these are based on a statistical model,
@@
 A pattern of isolation by distance (IBD) arises when migration in a
-population is spatially limited, \gc{and slow enough betweeen distant
-locations that genetic drift can build up differentiation}
-\citep{Wright1943}. IBD has been modeled in both spatially continuous frameworks \citep{malecot1969mathematics,Slatkin1985}, 
+population is spatially limited, and slow enough between distant
+locations that genetic drift can build up differentiation between them
+\citep{Wright1943}. 
+IBD has been modeled in both spatially continuous frameworks 
+\citep{malecot1969mathematics,Slatkin1985}, 
 in which an individual's parents are found within some radius of that individual's location,
@@
 and has given rise to a debate in the population genetic literature
-about \gc{how best to describe sets of individuals using continuous clines and} discrete clusters 
+about how best to describe sets of individuals using continuous clines and discrete clusters 
 \citep[e.g.][]{SerrePaabo2004,rosenberg2005clines}.
-%As a further complication, if local genetic drift is strong,
-%continous models are expected to give rise to spatially patchy patterns
-%\citep{SLFV,bramson1989crabgrass}.
 
@@
 The model works with allele frequencies at $L$ unlinked, bi-allelic single nucleotide polymorphisms (SNPs) genotyped across $N$ samples.
-Each ``sample'' may \gc{be a single individual, a collection of
-  individuals from a location}, or frequencies estimated from pooled sequencing.
+Each ``sample'' may be a single individual,  
+a collection of individuals from a location, 
+or frequencies estimated from pooled sequencing.
 We write $N_i$ for the number of chromosomes sequenced in the $i^\text{th}$ sample.
@@
 of a randomly chosen allele in samples $i$ and $j$, as long as $i \neq j$.
-The diagonal (where $i=j$) does not have this interpretation, \gc{and
-  reflects the covariance between indivdual alleles drawn from within
-  the population}.
-
-\gc{This definition of convariance} differs from the usual ``genetic covariance'' \citep{mcvean_genealogical_2009}
+The diagonal (where $i=j$) does not have this interpretation, 
+and reflects the covariance between individual alleles drawn from within the population.
+
+This definition of covariance differs from the usual ``genetic covariance'' \citep{mcvean_genealogical_2009}
 in that (a) we do not subtract locus means,
@@
 then $\Omega_{i,j} = (1 - 2 D_{i,j})/4$.
-Therefore, unlike genetic covariance,
-this allelic covariance is \gc{more} affected in shape by singleton sites,
+Therefore, 
+this allelic covariance is more affected in shape by singleton sites
+than the standard genetic covariance,
 so it may be advisable to filter these prior to analysis
@@
 which we refer to as a set of ``layers'' that overlay the modern map.
-We imagine each layer as a geographically distributed population,
-\gc{each layer extends over the entire sampled range of the
-  populations}. 
+We imagine each layer as a geographically distributed population 
+that extends over the entire sampled range of the populations. 
 Each sample is composed of a mixture of contributions from each of
-these layers. \gc{See Figure \ref{schematic} for a cartoon representation
-of our model.} 
-These layers thus take the place of ``clusters'' or ``populations'' in clustering methods,
-but we do not adopt these terms as \gc{terms like ``spatial'' cluster
-  suggest a clustering in space while our layers (and samples from
-  them) may be present at all sampled locations.}
+these layers. 
+See Figure \ref{schematic} for a cartoon representation of our model. 
+These layers thus take the place of ``clusters'' in clustering methods,
+but we do not adopt this term, 
+as ``spatial cluster'' suggests a clustering in space, 
+while our layers may be present at all sampled locations.
 
@@
 covariance with distance in the cluster.
-\gc{Our choice of a powered-exponential decay, as parameterized by the
-  $\alpha$s, is a flexible, and standard choice in spatial statistics
-  REF, and is not chosen to match  a particular population genetics model. 
-The} $\phi^{(k)}$ is a parameter that describes the background covariance within the cluster. 
+Our choice of a powered-exponential decay, 
+as parameterized by the $\alpha$s, 
+is a flexible and standard choice in spatial statistics \citep{Diggle1998}, 
+and is not chosen to match a particular population genetics model. 
+The $\phi^{(k)}$ is a parameter that describes the background covariance within the cluster. 
 If two samples draw 100\% of their ancestry from cluster $k$, then their covariance under the model is $G^{(k)}_{i,j}$;
@@
 The ``admixture'' proportion of the $i$th sample in the $k$th cluster, denoted $w^{(k)}_i$,
-gives the \gc{genome-wide} proportion of alleles from sample $i$ that derive from
+gives the genome-wide proportion of alleles from sample $i$ that derive from
 cluster $k$ (and so $\sum_K w^{(k)}_i =1$).
@@
 		\caption{Schematic of our method, using $K=3$ as an example.
-			     The pattern of isolation by distance within each layer is denoted by the color gradient in each of the layer ellipses, 
-			     and $\phi_{(k)}$ denotes the covariance shared by samples with ancestry entirely in the $k$th cluster.
+			     The pattern of isolation by distance within each layer is denoted by their color gradients, 
+			     and $\phi_{(k)}$ denotes the covariance shared by samples with ancestry entirely in the $k$th layer.
 			     Sampled populations on the landscape are inferred to be admixed between these layers; 
-			     the $i$th sample draws proportion $w^{(k)}_i$ of its ancestry from cluster $k$.
+			     the $i$th sample draws proportion $w^{(k)}_i$ of its ancestry from layer $k$.
+			     For convenience, each layer is depicted as a small square, but in fact, 
+			     each layer exists everywhere in the sampled area,
+			     so the small dashed circles on each layer show where 
+			     the location of the highlighted, admixed sample intersects each layer.
 			    }\label{schematic}
@@
 \paragraph{Likelihood and inference}
-\gc{Assuming that the} allele frequencies at each locus
-were independent between loci and multivariate normally distributed across populations, 
-then their allelic covariance $\widehat{\Omega}$ 
+Assuming that the allele frequencies at each locus
+are independent between loci and multivariate normally distributed across populations, 
+their allelic covariance $\widehat{\Omega}$ 
 would be Wishart distributed with degrees of freedom equal to $L$, 
@@
 Linkage disequilibrium (LD) between loci will decrease the effective
-number of degrees of freedom.  \gc{One possible solution, that we have
-  not implemented yet, is to estimate an \emph{effective} number of loci 
+number of degrees of freedom.  
+One possible solution, which we have not yet implemented, 
+is to estimate an \emph{effective} number of loci 
 by introducing a parameter to modify the given degrees of freedom 
-and thereby informally model linkage between loci \citep[e.g.][]{EEMS}.}
+and thereby informally model linkage between loci \citep[e.g.][]{EEMS}.
 
@@
 where $P(w)$, $P(\alpha)$, $P(\phi)$, $P(\eta)$, and $P(\gamma)$, are prior distributions.
-All parameters are given (half-)Gaussian priors except for $\alpha_2$,
-which is uniform on $(0,2)$ and $w$, which is an \gc{independent Dirichlet, of
-  dimension $K$, for each individual}
+All parameters are given (half-)Gaussian priors 
+except for $\alpha_2$, which is uniform on $(0,2)$, 
+and $w$, for which we use an independent Dirichlet of dimension $K$ for each individual
 (see Table \ref{tab:param_prior_tab} for specifics).
@@
 --- spatial vs. nonspatial, across different values of $K$ --- 
-by comparing their performance in a common framework. \gc{We also
-  validate our claim that the results of our non-spatial method are a
-  close match to those of STRUCTURE-based approaches.}
+by comparing their performance in a common framework. 
+We also validate our claim that the results of our non-spatial method 
+are a close match to those of STRUCTURE-based approaches.
 
 \paragraph{Choice of cluster number and cross-validation}
-There are a number of \gc{reasons 
-why there is no true (or right)} number of layers for real datasets,
+There are a number of reasons 
+why there is no true (or right) number of layers for real datasets,
 discussed further in the Discussion.
@@
 is added by spatial structure within each layer, and each additional layer.
-A simple $t$-test can be used to assess whether additional layers
-show statistically significantly greater predictive accuracy. \gc{DO
-  WE DO t-test? IF NOT DROP THIS'}
 However, we remind users that ``statistical significance does not imply real-world significance'',
@@
 	\caption{
-	Map sof admixture proportions estimated for the black bear dataset 
+	Maps of admixture proportions estimated for the black bear dataset 
 	using the spatial model (left) and the nonspatial model (right) for $K=3$.
@@
 
-We thank Marjorie Weber, William Wetzel, Mariah Meek, Yaniv Brandvain, and Matthew Stephens 
+We thank Marjorie Weber, Yaniv Brandvain, William Wetzel, 
+Mariah Meek, Doc Edge, and Matthew Stephens 
 for invaluable comments on the method and manuscript, 
 as well as Quentin Cronk, who provided input on the \textit{Populus} analyses, 
-and the attendees at the 2017 SSE Meeting in Portland, OR, 
+and Emily Puckett, who provided input on the black bear analyses.
+We also thank the attendees at the 2017 SSE Meeting in Portland, OR, 
 whose votes determined the name of the method.

### 2017-09-06 writeup/conStruct.tex
fixed wording for new cols

@@
 It is possible that this is happening in Fig \ref{populus_sp_pies}\subref{populus_sp_pies:K3}. 
-At $K=3$, the southernmost populations of \textit{P. trichocarpa} are in the green cluster, 
+At $K=3$, the southernmost populations of \textit{P. trichocarpa} are in the yellow cluster, 
 whose other neighbors are to the north, with an intervening group of populations in the red cluster, 
-and at $K=4$, those southernmost samples split out and become their own yellow cluster.
+and at $K=5$, those southernmost samples split out and become their own cluster.
 Again, we encourage users to run analyses across multiple values of $K$, 

### 2017-09-07 writeup/conStruct.tex
minor edits for clarity

@@
 to estimate the posterior distribution on the parameters.
-We also provide an R package \citep{R} called \texttt{conStruct} that functions as a wrapper 
-around this inference machinery.
+We also provide an R package \citep{R} called \texttt{conStruct} 
+(for ``continuous structure") that functions as a wrapper around this inference machinery.
 
@@
 	Cross-validation results for data simulated under $K=1$, $K=2$, and $K=3$, 
-	comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
+	comparing the spatial and nonspatial \texttt{conStruct} models 
+	(in blue and green, respectively) run with $K=1$ through 7.  
 	The inset plots zoom in on cross-validation results outlined in the dotted boxes.
@@
 The results of these analyses are shown in 
-Figs \ref{populus_xvals} and \ref{populus_pies}, 
+Figs \ref{populus_pies} and \ref{populus_xvals}, 
 as well as in Figs \ref{populus_sp_pies} - \ref{populus_laycon} in the Supplementary Materials.
@@
 All models with $K>1$ assigned the majority of each species
-to distinct layers, with some populations drawing ancestry from both layers.
+to distinct layers, with some populations drawing ancestry from multiple layers.
 This provides strong support for discrete population structure between the two species,
@@
 while \bals{} samples displayed no substructure up to $K=7$.
-This lack of substructure within the sampled \bals{} populations is in agreement with results from 
+This lack of substructure within the sampled \bals{} populations is largely in agreement with results from 
 \citet{keller_etal_2010}, who found that \bals{} in the region is characterized by relatively low diversity, 
@@
 This indicates that, although isolation by distance is a feature of the dataset as a whole, 
-it's magnitude varies considerably across biological units - 
+its magnitude varies considerably across biological units - 
 an aspect of the data \texttt{conStruct} is able capture well.
@@
 and compared these models using cross-validation.
-The results of these analyses are shown in Figs \ref{bear_xvals} - \ref{bear_laycon}, 
+The results of these analyses are shown in Figs \ref{bear_K3} - \ref{bear_laycon}, 
 as well as in Figs \ref{bear_sp_pies} - \ref{bear_nsp_clst_covs} in the Supplementary Materials.
@@
 Our model, which explicitly incorporates a spatial decay of relatedness, 
-allows individuals sampled at a large remove to belong to the same layer
-instead of splitting these individuals out into successive clusters.
+allows somewhat genetically differentiated individuals 
+that are sampled far from one another to belong to the same layer,
+instead of splitting these individuals out into successive clusters 
+(e.g., Fig \ref{bear_sp_pies:K5} vs \ref{bear_nsp_pies:K5}).
 
@@
 	\caption{
-	Map of admixture proportions estimated for the \textit{bear} dataset 
+	Map of admixture proportions estimated for the bear dataset 
 	using the spatial model for $K=2$ through 7.
@@
 	\caption{
-	Map of admixture proportions estimated for the \textit{bear} dataset 
+	Map of admixture proportions estimated for the bear dataset 
 	using the nonspatial model for $K=2$ through 7.

### 2017-09-08 writeup/conStruct.tex
convert multipanel figs for submission

@@
 \usepackage{float}
-\usepackage{subcaption}
 \usepackage{hyperref}
@@
 	\centering
-		\subcaptionbox{$K=2$\label{simK1_pies:nsp_K2}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K2.pdf}}
-		\subcaptionbox{$K=3$\label{simK1_pies:nsp_K3}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K3.pdf}}
-		\subcaptionbox{$K=4$\label{simK1_pies:nsp_K4}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K4.pdf}}
-		\subcaptionbox{$K=2$\label{simK1_pies:sp_K2}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K5.pdf}}
-		\subcaptionbox{$K=3$\label{simK1_pies:sp_K3}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K6.pdf}}
-		\subcaptionbox{$K=4$\label{simK1_pies:sp_K4}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K7.pdf}}
+		{\includegraphics[width=6in,height=4in]{figs/sims/Fig2_simK1_sp_vs_nsp.pdf}}
 	\caption{
@@
 	\centering
-		\subcaptionbox{$K=2$\label{populus_pies:K2}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp2.pdf}}
-		\subcaptionbox{$K=3$\label{populus_pies:K3}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp3.pdf}}
-		\subcaptionbox{$K=4$\label{populus_pies:K4}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp4.pdf}}
+		{\includegraphics[width=6in,height=2in]{figs/populus/Fig4_pop_sp_results.pdf}}
 	\caption{
@@
 	\centering
-		\subcaptionbox{$K=3$ (spatial)\label{bear_K3:sp}}
-			{\includegraphics[width=2.5in,height=2.5in]{figs/bears/bear_sp3.pdf}}
-		\subcaptionbox{$K=3$ (nonspatial)\label{bear_K3:nsp}}
-			{\includegraphics[width=2.5in,height=2.5in]{figs/bears/bear_nsp3.pdf}}
+		{\includegraphics[width=6in,height=3in]{figs/bears/Fig6_sp_vs_nsp.pdf}}
 	\caption{
@@
 using gradients of admixture between clusters.
-For example, in Fig \ref{bear_K3:nsp}, 
+For example, in Fig \ref{bear_K3}b, 
 the third cluster (in green) exhibits a clear East-West gradient that 
@@
 instead of splitting these individuals out into successive clusters 
-(e.g., Fig \ref{bear_sp_pies:K5} vs \ref{bear_nsp_pies:K5}).
+(e.g., Fig \ref{bear_sp_pies}d vs \ref{bear_nsp_pies}d).
 
@@
 individuals at a far spatial remove may be in the same cluster but have very low pairwise relatedness.
-It is possible that this is happening in Fig \ref{populus_sp_pies}\subref{populus_sp_pies:K3}. 
-At $K=3$, the southernmost populations of \textit{P. trichocarpa} are in the yellow cluster, 
+It is possible that this is happening in Fig \ref{populus_sp_pies}. 
+At $K=3$, the southernmost populations of \textit{P. trichocarpa} are in the gold cluster, 
 whose other neighbors are to the north, with an intervening group of populations in the red cluster, 
@@
 	\centering
-		\subcaptionbox{$K=2$\label{simK1_nsp_pies:K2}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K2.pdf}}
-		\subcaptionbox{$K=3$\label{simK1_nsp_pies:K3}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K3.pdf}}
-		\subcaptionbox{$K=4$\label{simK1_nsp_pies:K4}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K4.pdf}}
-		\subcaptionbox{$K=5$\label{simK1_nsp_pies:K5}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K5.pdf}}
-		\subcaptionbox{$K=6$\label{simK1_nsp_pies:K6}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K6.pdf}}
-		\subcaptionbox{$K=7$\label{simK1_nsp_pies:K7}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_nsp_pies_K7.pdf}}
+		{\includegraphics[width=6in,height=4in]{figs/sims/simK1_nsp_pies.pdf}}
 	\caption{
@@
 	\centering
-		\subcaptionbox{$K=2$\label{simK1_sp_pies:K2}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K2.pdf}}
-		\subcaptionbox{$K=3$\label{simK1_sp_pies:K3}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K3.pdf}}
-		\subcaptionbox{$K=4$\label{simK1_sp_pies:K4}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K4.pdf}}
-		\subcaptionbox{$K=5$\label{simK1_sp_pies:K5}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K5.pdf}}
-		\subcaptionbox{$K=6$\label{simK1_sp_pies:K6}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K6.pdf}}
-		\subcaptionbox{$K=7$\label{simK1_sp_pies:K7}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK1_sp_pies_K7.pdf}}
+		{\includegraphics[width=6in,height=4in]{figs/sims/simK1_sp_pies.pdf}}
 	\caption{
@@
 	\centering
-		\subcaptionbox{$K=2$\label{simK2_nsp_pies:K2}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_nsp_pies_K2.pdf}}
-		\subcaptionbox{$K=3$\label{simK2_nsp_pies:K3}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_nsp_pies_K3.pdf}}
-		\subcaptionbox{$K=4$\label{simK2_nsp_pies:K4}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_nsp_pies_K4.pdf}}
-		\subcaptionbox{$K=5$\label{simK2_nsp_pies:K5}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_nsp_pies_K5.pdf}}
-		\subcaptionbox{$K=6$\label{simK2_nsp_pies:K6}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_nsp_pies_K6.pdf}}
-		\subcaptionbox{$K=7$\label{simK2_nsp_pies:K7}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_nsp_pies_K7.pdf}}
+		{\includegraphics[width=6in,height=4in]{figs/sims/simK2_nsp_pies.pdf}}
 	\caption{
@@
 	\centering
-		\subcaptionbox{$K=2$\label{simK2_sp_pies:K2}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_sp_pies_K2.pdf}}
-		\subcaptionbox{$K=3$\label{simK2_sp_pies:K3}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_sp_pies_K3.pdf}}
-		\subcaptionbox{$K=4$\label{simK2_sp_pies:K4}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_sp_pies_K4.pdf}}
-		\subcaptionbox{$K=5$\label{simK2_sp_pies:K5}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_sp_pies_K5.pdf}}
-		\subcaptionbox{$K=6$\label{simK2_sp_pies:K6}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_sp_pies_K6.pdf}}
-		\subcaptionbox{$K=7$\label{simK2_sp_pies:K7}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK2_sp_pies_K7.pdf}}
+		{\includegraphics[width=6in,height=4in]{figs/sims/simK2_sp_pies.pdf}}
 	\caption{
@@
 	\centering
-		\subcaptionbox{$K=2$\label{simK3_nsp_pies:K2}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_nsp_pies_K2.pdf}}
-		\subcaptionbox{$K=3$\label{simK3_nsp_pies:K3}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_nsp_pies_K3.pdf}}
-		\subcaptionbox{$K=4$\label{simK3_nsp_pies:K4}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_nsp_pies_K4.pdf}}
-		\subcaptionbox{$K=5$\label{simK3_nsp_pies:K5}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_nsp_pies_K5.pdf}}
-		\subcaptionbox{$K=6$\label{simK3_nsp_pies:K6}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_nsp_pies_K6.pdf}}
-		\subcaptionbox{$K=7$\label{simK3_nsp_pies:K7}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_nsp_pies_K7.pdf}}
+		{\includegraphics[width=6in,height=4in]{figs/sims/simK3_nsp_pies.pdf}}
 	\caption{
@@
 	\centering
-		\subcaptionbox{$K=2$\label{simK3_sp_pies:K2}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K2.pdf}}
-		\subcaptionbox{$K=3$\label{simK3_sp_pies:K3}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K3.pdf}}
-		\subcaptionbox{$K=4$\label{simK3_sp_pies:K4}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K4.pdf}}
-		\subcaptionbox{$K=5$\label{simK3_sp_pies:K5}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K5.pdf}}
-		\subcaptionbox{$K=6$\label{simK3_sp_pies:K6}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K6.pdf}}
-		\subcaptionbox{$K=7$\label{simK3_sp_pies:K7}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K7.pdf}}
+		{\includegraphics[width=6in,height=4in]{figs/sims/simK3_sp_pies.pdf}}
 	\caption{
@@
 	\centering
-		\subcaptionbox{$K=2$\label{populus_sp_pies:K2}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp2.pdf}}
-		\subcaptionbox{$K=3$\label{populus_sp_pies:K3}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp3.pdf}}
-		\subcaptionbox{$K=4$\label{populus_sp_pies:K4}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp4.pdf}}
-		\subcaptionbox{$K=5$\label{populus_sp_pies:K5}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp5.pdf}}
-		\subcaptionbox{$K=6$\label{populus_sp_pies:K6}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp6.pdf}}
-		\subcaptionbox{$K=7$\label{populus_sp_pies:K7}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_sp7.pdf}}
+		{\includegraphics[width=6in,height=4in]{figs/populus/populus_sp_pies.pdf}}
 	\caption{
@@
 	\centering
-		\subcaptionbox{$K=2$\label{populus_nsp_pies:K2}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp2.pdf}}
-		\subcaptionbox{$K=3$\label{populus_nsp_pies:K3}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp3.pdf}}
-		\subcaptionbox{$K=4$\label{populus_nsp_pies:K4}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp4.pdf}}
-		\subcaptionbox{$K=5$\label{populus_nsp_pies:K5}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp5.pdf}}
-		\subcaptionbox{$K=6$\label{populus_nsp_pies:K6}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp6.pdf}}
-		\subcaptionbox{$K=7$\label{populus_nsp_pies:K7}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/populus/populus_nsp7.pdf}}
+		{\includegraphics[width=6in,height=4in]{figs/populus/populus_nsp_pies.pdf}}
 	\caption{
@@
 	\centering
-		\subcaptionbox{$K=2$\label{bear_sp_pies:K2}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_sp2.pdf}}
-		\subcaptionbox{$K=3$\label{bear_sp_pies:K3}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_sp3.pdf}}
-		\subcaptionbox{$K=4$\label{bear_sp_pies:K4}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_sp4.pdf}}
-		\subcaptionbox{$K=5$\label{bear_sp_pies:K5}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_sp5.pdf}}
-		\subcaptionbox{$K=6$\label{bear_sp_pies:K6}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_sp6.pdf}}
-		\subcaptionbox{$K=7$\label{bear_sp_pies:K7}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_sp7.pdf}}
+		{\includegraphics[width=6in,height=4in]{figs/bears/bear_sp_results.pdf}}
 	\caption{
@@
 	\centering
-		\subcaptionbox{$K=2$\label{bear_nsp_pies:K2}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_nsp2.pdf}}
-		\subcaptionbox{$K=3$\label{bear_nsp_pies:K3}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_nsp3.pdf}}
-		\subcaptionbox{$K=4$\label{bear_nsp_pies:K4}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_nsp4.pdf}}
-		\subcaptionbox{$K=5$\label{bear_nsp_pies:K5}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_nsp5.pdf}}
-		\subcaptionbox{$K=6$\label{bear_nsp_pies:K6}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_nsp6.pdf}}
-		\subcaptionbox{$K=7$\label{bear_nsp_pies:K7}}
-			{\includegraphics[width=1.8in,height=1.8in]{figs/bears/bear_nsp7.pdf}}
+		{\includegraphics[width=6in,height=4in]{figs/bears/bear_nsp_results.pdf}}
 	\caption{

### 2017-09-08 writeup/figs/fig_funcs.R

@@ -180,60 +180,45 @@ plot.sim.pies <- function(data.block,K,output.list,file.name){
 	}
 }
 
-# admix.pie.plot <- function (coords, w, cluster.colors, radii = 2.7, add = FALSE,box=TRUE) {
-	# K <- ncol(w)
-	# N <- nrow(w)
-   	# cluster.names <- paste0("cluster_", 1:K)
-   	# sample.names <- paste0("sample_", 1:N)
-   	# color.tab <- caroline::nv(c(cluster.colors[1:K]),cluster.names)
-   	# pie.list <- lapply(1:N, function(i) {
-    		   # caroline::nv(w[i, ], cluster.names)
-   	# })
-   	# names(pie.list) <- sample.names
-   	# if (add) {
-    	   # par(new = TRUE)
-   	# }
-   	# else {
-    	   # par(mar = c(2, 2, 2, 2))
-   	# }
-   	# caroline::pies(pie.list, x0 = coords[, 1], 
-       # y0 = coords[, 2], color.table = color.tab, 
-       # border = "black", radii = radii, xlab = "", ylab = "", 
-       # main = "", lty = 1, density = NULL)
-	# if(box){
-	   	# box(lwd = 2)
-	# }
-    # return(invisible(0))
-# }
+plot.sim.pies.multipanel <- function(data.block,K,output.list,radii,mar=NULL){
+	#recover()
+	cluster.colors <- c("blue","red","goldenrod1","forestgreen","darkorchid1","deepskyblue","darkorange1","seagreen2","yellow1","black")
+	csr1.order <- NULL
+	for(k in 2:K){
+		data.block$K <- k
+		csr <- output.list[[k]][[1]]
+		if(k > 2){
+			tmp.csr <- output.list[[k-1]][[1]]
+			csr1.order <- conStruct:::match.clusters.x.runs(tmp.csr$MAP$admix.proportions,csr$MAP$admix.proportions,csr1.order)
+		}
+		if(is.null(csr1.order)){
+			csr1.order <- 1:k
+		}
+		make.admix.pie.plot(csr$MAP$admix.proportions[,csr1.order],data.block$coords,cluster.colors=cluster.colors,radii=radii,x.lim=c(2.5,8.5),y.lim=c(2.5,8.5),mar=mar)
+			mtext(side=1,text=bquote(paste("(",.(letters[k-1]),") ",italic("K")," = ",.(k))),padj=2.7,adj=0.4)
+	}
+}
 
[…]

### 2017-09-08 writeup/figs/sim_figs.R

@@ -51,6 +51,49 @@ plot.sim.pies(data.block = data.block,
 			  output.list = output.list.sp,
 			  file.name = "~/Dropbox/conStruct/writeup/figs/sims/simK1_sp_pies_K")
 
+pdf(file="~/Dropbox/conStruct/writeup/figs/sims/simK1_nsp_pies.pdf",width=8,height=5.34,pointsize=14)
+	par(mfrow=c(2,3))
+	plot.sim.pies.multipanel(data.block = data.block,
+							 K = 7,
+							 output.list = output.list.nsp,
+							 radii = 1.7,
+							 mar = c(5,3,1,3))
+dev.off()
+
+pdf(file="~/Dropbox/conStruct/writeup/figs/sims/simK1_sp_pies.pdf",width=8,height=5.34,pointsize=14)
+	par(mfrow=c(2,3))
+	plot.sim.pies.multipanel(data.block = data.block,
+							 K = 7,
+							 output.list = output.list.sp,
+							 radii = 1.7,
+							 mar = c(5,3,1,3))
+dev.off()
+
+pdf(file="~/Dropbox/conStruct/writeup/figs/sims/Fig2_simK1_sp_vs_nsp.pdf",width=8,height=5.34,pointsize=14)
+	#quartz(width=6,height=4,pointsize=14)
+	radii <- 1.7
+	mar <- c(5,3,1,3)
+	par(mfrow=c(2,3))
+		make.admix.pie.plot(output.list.nsp[[2]][[1]]$MAP$admix.proportions,
+								data.block$coords,radii=radii,x.lim=c(2.5,8.5),y.lim=c(2.5,8.5),mar= mar)
+			mtext(side=1,text=expression(paste("(a) ",italic("K")," = 2")),padj=2.7,adj=0.4)
+		make.admix.pie.plot(output.list.nsp[[3]][[1]]$MAP$admix.proportions[,c(2,3,1)],
+								data.block$coords,radii=radii,x.lim=c(2.5,8.5),y.lim=c(2.5,8.5),mar= mar)
+			mtext(side=1,text=expression(paste("(b) ",italic("K")," = 3")),padj=2.7,adj=0.4)
+		make.admix.pie.plot(output.list.nsp[[4]][[1]]$MAP$admix.proportions[,c(4,2,3,1)],
+								data.block$coords,radii=radii,x.lim=c(2.5,8.5),y.lim=c(2.5,8.5),mar= mar)
[…]

### 2017-09-09 writeup/conStruct.tex
G'c changes

--- 2017-09-07 writeup/conStruct.tex
@@
 such as geographic barriers, rapid expansions, bottlenecks, and rare
-long distance migration.
+long-distance migration.
 
@@
 Corander2003,TESS,geneland}, among others.
+\gc{Each of these methods perform best in particular situations,
+but many can give misleading results when applied to data that show a continuous pattern of isolation by distance 
+\citep{Wright1943, novembre_interpreting_2008, Frantz2009}.}
 Here, we will focus on model-based clustering, the most widely used class of approaches for population delineation.
@@
 Spatial information has been incorporated into some of these methods, 
-by, for example, placing spatial priors on cluster memberships \citep{geneland}, 
+by, for example, placing spatial priors on cluster memberships \citep{geneland,TESS}, 
 but this does not address the underlying issue that 
@@
 are constant in a cluster across the species' range.  
-Like the method we present here, these are based on a statistical model,
-but do not perform inference directly with a demographic model.
-Each of these methods perform best in particular situations,
-but all can give misleading results when applied to data that show a continuous pattern of isolation by distance 
-\citep{Wright1943, novembre_interpreting_2008, Frantz2009}.
+%Like the method we present here, these are based on a statistical model,
+%but do not perform inference directly with a demographic model.
 
@@
 at time $\tau_{\text{e}}$ in the past,
-each of these discrete clusters instantaneously colonizes a lattice of demes,
+each of these discrete clusters instantaneously colonizes an
+\gc{independent} lattice of demes,
 for which we use a stepping stone model with symmetric migration 
@@
 by collapsing those $K$ discrete lattices into a single grid of demes that are admixed to various degrees from these different layers.
-(see Fig \ref{sim_setup}).
-We simulated datasets using $K=1$, 2, and 3 layers,
+(see Fig \ref{sim_setup}). \gc{We collapse the layers together using
+random admixture proportions that are spatially autocorrelated  (see
+{\bf XXXX} for details), so
+that both the frequencies in clusters and the cluster memberships are
+spatially autocorrelated.} We simulated datasets using $K=1$, 2, and 3 layers,
 in each sampling 10,000 unlinked loci from each of 20 haploid individuals from every deme.
@@
 Across all values of K and for all simulations, 
-the spatial model is correctly preferred over the nonspatial model (Fig \ref{sim_xvals})., 
-and it infers the true admixture proportions with high accuracy, tight precision, and good coverage 
-(Figs \ref{simK2_adprop_fit} and \ref{simK3_adprop_fit}).
-Second, the predictive accuracy of the nonspatial model increases 
+the spatial model is correctly preferred over the nonspatial model
+(Fig \ref{sim_xvals}). Second, the predictive accuracy of the nonspatial model increases 
 with the number of clusters in the model, 
@@
  (Fig \ref{fastStr_simK1}-\ref{fastStr_simK3}).
-While it is true that each cluster is more similar within itself
+While it is true that each non-spatial cluster is \gc{genetically} more similar within itself
 than it is to other clusters,
@@
 (Fig \ref{simK1_sp_pies}, \ref{simK2_sp_pies}, \ref{simK3_sp_pies}).
-That is, even when the spatial model is run with $K=7$, 
+\gc{Even} when the spatial model is run with $K=7$, 
 the inferred admixture proportions are nearly identical to 
 those estimated under the true value of $K$ for each simulation.
-These results are highlighted in the calculation of layer importance for 
+\gc{Moreover, the method infers the true admixture proportions with high accuracy, tight precision, and good coverage 
+(Figs \ref{simK2_adprop_fit} and \ref{simK3_adprop_fit}).}
+
+\gc{Finally, we calculated} of layer importance for 
 analyses run in each simulation (Fig \ref{simK1_laycon}, \ref{simK2_laycon}, and \ref{simK3_laycon}).
-In the spatial analyses run using the true value of $K$,
-the importance of the $K$ layers remains constant even with additional layers;
-in contrast, additional layers in the nonspatial analyses continue to be important for describing variation.
+In the spatial analyses, \gc{once we pass the true $K$, subsequent
+layers add little in terms of (co)variance explained}  %$K$ layers remains constant even with additional layers;
+in contrast, additional layers in the nonspatial analyses continue to
+be important for describing variation, \gc{as they are forced to partition
+up the spatial patterns within clusters}.
 
@@
 
-Our data, originally published by \cite{geraldes_etal_2014}, 
+Our data, \gc{from} \cite{geraldes_etal_2014}, 
 consist of 434 individuals sampled from 35 drainages 
@@
 
-All models with $K>1$ assigned the majority of each species
+All models with $K>1$ assigned the majority of each \gc{of the two} species
 to distinct layers, with some populations drawing ancestry from multiple layers.
@@
 
-A consistent split between layers within \tri{} fell along the ``no-cottonwood belt," 
-a region along the central coast of British Columbia in which black cottonwood is absent. 
-The no-cottonwood belt is hypothesized to divide the species' distribution 
-into northern and southern groups, which, in a provenance test, 
-were experimentally shown to display different phenotypes 
-\citep[e.g., pathogen resistance,][]{xie2009,xie2012}.  
-At higher values of $K$, drainages at the southern tip of \tri{} sampling 
-begin to split out into their own layers.  
-This may potentially be the result of introgression from southern \textit{Populus} species, 
-such as \textit{P. angustifolia} or \textit{fremontii} \citep{Zhou2012,geraldes_etal_2014}
-
-At all values of $K$, the spatial model was preferred over the nonspatial model (Fig \ref{populus_xvals}).
+\gc{For values of $K<4$,} the spatial model was strongly preferred over the nonspatial model (Fig \ref{populus_xvals}).
 Most of this signal seems to derive from the \textit{P. trichocarpa} samples:
@@
 an aspect of the data \texttt{conStruct} is able capture well.
+\gc{Based on our cross-validation results we view $K=3$ spatial-model
+  as sufficient description of the data.}
+
+A consistent split between layers within \tri{} fell along the ``no-cottonwood belt," 
+a region along the central coast of British Columbia in which black
+cottonwood is absent \gc{(the break between yellow and red, for $K \geq 3$)} . 
+The no-cottonwood belt is hypothesized to divide the species' distribution 
+into northern and southern groups, which, in a provenance test, 
+were experimentally shown to display \gc{differences in ecologically relevant} phenotypes 
+\citep[e.g., pathogen resistance,][]{xie2009,xie2012}.  
+At higher values of $K$, drainages at the southern tip of \tri{} sampling 
+begin to split out into their own layers.  
+This may potentially be the result of introgression from southern \textit{Populus} species, 
+such as \textit{P. angustifolia} or \textit{fremontii} \citep{Zhou2012,geraldes_etal_2014}
 
@@
 black bears were confined to isolated glacial refugia, 
-from which they subsequently expanded to occupy their current range
+from which they subsequently expanded to occupy their current \gc{range
 \citep{WoodingWard1997,Byun1997,Stone2000,Puckett2015}.
-The retreat of the glaciers was relatively recent in units of 
-black bear generation times \citep{Onorato2004}, 
-so we might reasonably expect the divergence accrued between refugia 
-to be reflected in discrete genetic structure between extant populations.
-Black bears are strong dispersers \citep{Moore2014}, 
-but they are distributed across most of the continent, 
-so we may therefore expect them to experience distance-limited 
-mating opportunities and exhibit a pattern of IBD.
-As with the poplars described above, 
+%The retreat of the glaciers was relatively recent in units of 
+%black bear generation times \citep{Onorato2004}, 
+%so we might reasonably expect the divergence accrued between refugia 
+%to be reflected in discrete genetic structure between extant populations.
+%Black bears are strong dispersers \citep{Moore2014}, 
+%but they are distributed across most of the continent, 
+%so we may therefore expect them to experience distance-limited 
+%mating opportunities and exhibit a pattern of IBD.
+As} with the poplars described above, 
 this system is likely to present patterns of genetic structure 
@@
 
-Bears in the sample are partitioned into two main groups:
-one to the east of the Rocky Mountains, 
-which also occurs in Alaska, and, presumably, 
-in the intervening (unsampled) northern Great Plains and Canadian Shield;
-the other primarily west of the Rockies, 
-but presumably stretching up into British Columbia.
+Bears in the sample are partitioned into two main groups \gc{($K \geq 2$)}:
+one \gc{(red)} to the east of the Rocky Mountains, 
+which also occurs in Alaska,
+the other primarily west of the Rockies \gc{(blue)}.
+\gc{The disjointed range of the blue layer, likely reflects the fact
+that Canada was not sampled, and so the blue layer may extend
+through the intervening (unsampled) northern Great Plains and Canadian
+Shield, with the red layer presumably then stretching up into British Columbia.}
+
 Additional layers in the spatial model have strong statistical support 
@@
 However, upon examining layer importance (Fig \ref{bear_laycon}),
-it is clear that the additional layers beyond $K=2$ are contributing only a very small amount toward 
+it is clear that the additional \gc{spatial} layers beyond $K=2$ are contributing only a very small amount toward 
 total covariance between individuals.
@@
 and lack of denser sampling across northern North America, 
-make more detailed interpretations untrustworthy.
-In addition, our model is not an explicit test of any demographic scenario, 
-and also uses a dataset that consists solely of nuclear (rather than mitochondrial) DNA.
+\gc{make more detailed interpretations untrustworthy.}
+%In addition, our model is not an explicit test of any demographic scenario, 
+%and also uses a dataset that consists solely of nuclear (rather than
+%mitochondrial) DNA. WHY DO WE CARE about BEAR MTDNS
 
@@
 downstream analysis and interpretation of genetic data, 
-including conservation management decisions and 
-identification of the genetic basis of adaptation or disease traits.
+including conservation management decisions.
+
+%and 
+%identification of the genetic basis of adaptation or disease traits.
 
@@
 By employing the sensible default assumption
-that relatedness ought to decay with distance, even within a population, 
+that relatedness ought to decay with \gc{geographic} distance, even within a population, 
 we avoid erroneously ascribing population differentiation to discrete population clusters.
@@
 Empirical applications of \texttt{conStruct} to a sample of North American poplars 
-and a sample of North American black bears yield reasonable results.
+and a sample of North American black bears yield reasonable results,
+\gc{and demonstrate that real datasets can better described using fewer
+  layers, by acknowledging isolation by distance. }
 
@@
 and 
-(b) \texttt{conStruct} can concisely capture spatial structure, which is nearly always expected within clusters.
+(b) \texttt{conStruct} can concisely capture spatial structure, which
+\gc{will be common within} clusters.
 Given this, when should methods without spatial capability be used?

### 2017-09-09 writeup/conStruct.tex
appendix on sim details

@@
 Corander2003,TESS,geneland}, among others.
-\gc{Each of these methods perform best in particular situations,
-but many can give misleading results when applied to data that show a continuous pattern of isolation by distance 
-\citep{Wright1943, novembre_interpreting_2008, Frantz2009}.}
+Each of these methods perform best in particular situations,
+but many can give misleading results when applied to data 
+that show a continuous pattern of isolation by distance 
+\citep{Wright1943, novembre_interpreting_2008, Frantz2009}.
 Here, we will focus on model-based clustering, the most widely used class of approaches for population delineation.
@@
 are constant in a cluster across the species' range.
-%Like the method we present here, these are based on a statistical model,
-%but do not perform inference directly with a demographic model.
 
@@
 the number of discrete clusters present.
-
-%with downstream implications for interpretation of results.
-%For example, conservation resources might be spread too thinly over areas 
-%that are spuriously assumed to be discrete management units.
-%Or, in studies of locally adaptive loci, or of the genetics of disease,
-%discrete populations may be modeled in separately,
-%reducing the global power to detect the genomic basis of complex phenotypes.
 
@@
 without subsequent migration between them.
-To generate continuous differentiation within each cluster,
+To generate continuous differentiation within each layer,
 at time $\tau_{\text{e}}$ in the past,
-each of these discrete clusters instantaneously colonizes an
-\gc{independent} lattice of demes,
+each of these discrete populations instantaneously colonizes an
+independent lattice of demes,
 for which we use a stepping stone model with symmetric migration 
@@
 we generate a single dataset 
-by collapsing those $K$ discrete lattices into a single grid of demes that are admixed to various degrees from these different layers.
-(see Fig \ref{sim_setup}). \gc{We collapse the layers together using
-random admixture proportions that are spatially autocorrelated  (see
-{\bf XXXX} for details), so
-that both the frequencies in clusters and the cluster memberships are
-spatially autocorrelated.} We simulated datasets using $K=1$, 2, and 3 layers,
+by collapsing those $K$ discrete layers into a single grid of demes 
+that have various amounts of admixture from these different layers
+(see Fig \ref{sim_setup}). 
+We collapse the layers together using random admixture proportions 
+that are spatially autocorrelated (see Appendix \secref{sim_details} for details), 
+so that both the frequencies in layers and the cluster memberships are
+spatially autocorrelated. 
+We simulated datasets using $K=1$, 2, and 3 layers,
 in each sampling 10,000 unlinked loci from each of 20 haploid individuals from every deme.
@@
  (Fig \ref{fastStr_simK1}-\ref{fastStr_simK3}).
-While it is true that each non-spatial cluster is \gc{genetically} more similar within itself
+While it is true that each non-spatial cluster is genetically more similar within itself
 than it is to other clusters,
@@
 (Fig \ref{simK1_sp_pies}, \ref{simK2_sp_pies}, \ref{simK3_sp_pies}).
-\gc{Even} when the spatial model is run with $K=7$, 
+Even when the spatial model is run with $K=7$, 
 the inferred admixture proportions are nearly identical to 
 those estimated under the true value of $K$ for each simulation.
-\gc{Moreover, the method infers the true admixture proportions with high accuracy, tight precision, and good coverage 
-(Figs \ref{simK2_adprop_fit} and \ref{simK3_adprop_fit}).}
-
-\gc{Finally, we calculated} of layer importance for 
-analyses run in each simulation (Fig \ref{simK1_laycon}, \ref{simK2_laycon}, and \ref{simK3_laycon}).
-In the spatial analyses, \gc{once we pass the true $K$, subsequent
-layers add little in terms of (co)variance explained}  %$K$ layers remains constant even with additional layers;
+Moreover, the method infers the true admixture proportions with high accuracy, tight precision, and good coverage 
+(Figs \ref{simK2_adprop_fit} and \ref{simK3_adprop_fit}).
+
+Finally, we calculated layer importance for 
+analyses run in each simulation 
+(Fig \ref{simK1_laycon}, \ref{simK2_laycon}, and \ref{simK3_laycon}).
+In the spatial analyses, once we pass the true $K$, 
+subsequent layers add little in terms of (co)variance explained; 
 in contrast, additional layers in the nonspatial analyses continue to
-be important for describing variation, \gc{as they are forced to partition
-up the spatial patterns within clusters}.
+be important for describing variation, 
+as they are forced to partition up the spatial patterns within clusters.
 
@@
     }\label{sim_xvals}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=5in,height=2.5in]{figs/sims/simK1_laycon_barplots.pdf}}
+		\caption{
+			Layer ``importance" (i.e., how much total covariance is contributed by each layer), 
+			for all layers estimated in runs using $K = 1$ through 7 
+			for the spatial model (left) and the nonspatial model (right).
+			Data were simulated using $K=1$.
+			For each value of $K$ along the x-axis, there are an equal number of layer contributions plotted.
+		}\label{simK1_laycon}
 \end{figure}
@@
 
-Our data, \gc{from} \cite{geraldes_etal_2014}, 
+Our data, from \cite{geraldes_etal_2014}, 
 consist of 434 individuals sampled from 35 drainages 
@@
 
-All models with $K>1$ assigned the majority of each \gc{of the two} species
+All models with $K>1$ assigned the majority of each of the two species
 to distinct layers, with some populations drawing ancestry from multiple layers.
@@
 
-\gc{For values of $K<4$,} the spatial model was strongly preferred over the nonspatial model (Fig \ref{populus_xvals}).
+For values of $K<4$, the spatial model was strongly preferred over the nonspatial model (Fig \ref{populus_xvals}).
 Most of this signal seems to derive from the \textit{P. trichocarpa} samples:
@@
 an aspect of the data \texttt{conStruct} is able capture well.
-\gc{Based on our cross-validation results we view $K=3$ spatial-model
-  as sufficient description of the data.}
+Based on our cross-validation results we view the $K=3$ spatial model 
+as a sufficient description of the data.
 
@@
 a region along the central coast of British Columbia in which black
-cottonwood is absent \gc{(the break between yellow and red, for $K \geq 3$)} . 
+cottonwood is absent (the break between yellow and red, for $K \geq 3$). 
 The no-cottonwood belt is hypothesized to divide the species' distribution 
 into northern and southern groups, which, in a provenance test, 
-were experimentally shown to display \gc{differences in ecologically relevant} phenotypes 
+were experimentally shown to display differences in ecologically relevant phenotypes 
 \citep[e.g., pathogen resistance,][]{xie2009,xie2012}.  
@@
 black bears were confined to isolated glacial refugia, 
-from which they subsequently expanded to occupy their current \gc{range
+from which they subsequently expanded to occupy their current range
 \citep{WoodingWard1997,Byun1997,Stone2000,Puckett2015}.
@@
 %mating opportunities and exhibit a pattern of IBD.
-As} with the poplars described above, 
+As with the poplars described above, 
 this system is likely to present patterns of genetic structure 
@@
 
-Our data, originally published by \cite{Puckett2015}, consist of 95 individuals 
+Our data, from \cite{Puckett2015}, consist of 95 individuals 
 sampled across the United States and on the West coast of Canada,
@@
 
-Bears in the sample are partitioned into two main groups \gc{($K \geq 2$)}:
-one \gc{(red)} to the east of the Rocky Mountains, 
+Bears in the sample are partitioned into two main groups ($K \geq 2$):
+one (red) to the east of the Rocky Mountains, 
 which also occurs in Alaska,
-the other primarily west of the Rockies \gc{(blue)}.
-\gc{The disjointed range of the blue layer, likely reflects the fact
-that Canada was not sampled, and so the blue layer may extend
-through the intervening (unsampled) northern Great Plains and Canadian
-Shield, with the red layer presumably then stretching up into British Columbia.}
+the other primarily west of the Rockies (blue).
+The disjointed range of the blue layer likely reflects the fact
+that Canada was not sampled, 
+and so the blue layer may extend
+through the intervening (unsampled) northern Great Plains and Canadian Shield, 
+with the red layer presumably then stretching up into British Columbia.
 
@@
 However, upon examining layer importance (Fig \ref{bear_laycon}),
-it is clear that the additional \gc{spatial} layers beyond $K=2$ are contributing only a very small amount toward 
+it is clear that the additional spatial layers beyond $K=2$ are contributing only a very small amount toward 
 total covariance between individuals.
@@
 and lack of denser sampling across northern North America, 
-\gc{make more detailed interpretations untrustworthy.}
+make more detailed interpretations untrustworthy.
 %In addition, our model is not an explicit test of any demographic scenario, 
@@
 By employing the sensible default assumption
-that relatedness ought to decay with \gc{geographic} distance, even within a population, 
+that relatedness ought to decay with geographic distance, even within a population, 
 we avoid erroneously ascribing population differentiation to discrete population clusters.
@@
 and a sample of North American black bears yield reasonable results,
-\gc{and demonstrate that real datasets can better described using fewer
-  layers, by acknowledging isolation by distance. }
+and demonstrate that,
+by acknowledging isolation by distance,
+real datasets can be better described using fewer layers.
 
@@
 (b) \texttt{conStruct} can concisely capture spatial structure, which
-\gc{will be common within} clusters.
+will be common within populations.
 Given this, when should methods without spatial capability be used?
@@
 \newpage
+\section{Simulation details}\label{sim_details}
+We wished to simulate data under a model that had some biological realism, 
+but at the same time had unambiguous true admixture proportions 
+(so as to test the behavior of the method).
+This second requirement precluded scenarios of, e.g, 
+recent secondary contact between populations expanding out of different refugia, 
+which would have more biological realism, 
+but no unambiguous ancestry proportions for admixed populations.
+Here, we describe in more detail the procedure we use to simulate our test dataset, 
+using a cartoon schematic with $K=2$ as an example (Fig \ref{sim_setup}).
+
+Using the program \texttt{ms} \citep{Hudson2002}, 
+we generated discrete population structure by simulating $K$ distinct populations,
+each of which split from a common ancestor $\tau_{\text{s}}$ units of coalescent time in the past,
+without subsequent migration between them.
+Then, to generate continuous differentiation within each cluster,
+at time $\tau_{\text{e}}$ in the past,
+each of these discrete clusters instantaneously colonizes an
+independent lattice of demes,
+for which we use a stepping stone model with symmetric migration 
+to nearest neighbors (eight neighbors, including diagonals).
+
+Finally, at time $\tau_{\text{a}}$ in the past 
+we generate a single dataset 
+by collapsing those $K$ discrete lattices into a single grid of demes 
+that are admixed to various degrees from these different layers. 
+We wish to simulate realistic patterns of admixture 
+(and thereby set a more difficult test for the method), 
+by generating spatially autocorrelated admixture proportions 
+in each diverged population.
+To do so, we first place $K$ equidistant points on the circle centered on our lattice.  
+These points serve as ``foci" of ancestry in each of the $K$ layers.  
+We then calculate the distance from each deme in the sampled lattice to each of these $K$ foci, 
+and draw admixture proportions for each deme 
+from a Dirichlet distribution for which the concentration parameter 
+for deme $i$ in layer $k$ is inversely proportional to the distance between deme $i$ and focus $k$. 
+This creates a pattern in which the admixture proportions in a given layer decreases 
+with the distance from that layer's focus, 
+as might be expected if a spatial process were mediating admixture between diverged populations.
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=4in,height=2.5in]{figs/sims/sim_setup.png}}
+		\caption{Schematic of how we simulate datasets with continuous and discrete differentiation, using $K=2$ as an example.  
+			    Going forward in time, the $K$ clusters split from a common ancestor at time $\tau_{\text{s}}$,
+			    then expand to each  colonize a lattice of demes with nearest-neighbor symmetric migration at time $\tau_{\text{e}}$,
+			    then finally at time $\tau_{\text{a}}$ collapse into a single lattice consisting of demes 
+			    with ancestry entirely in one or the other of the clusters,
+			    or admixed between them.
+			    }\label{sim_setup}
+\end{figure}
+
+\newpage
 \section*{Supplementary Materials}
@@
 
-\clearpage
-\begin{figure}
-	\centering
-		{\includegraphics[width=4in,height=2.5in]{figs/sims/sim_setup.png}}
-		\caption{Schematic of how we simulate datasets with continuous and discrete differentiation, using $K=2$ as an example.  
-			    Going forward in time, the $K$ clusters split from a common ancestor at time $\tau_{\text{s}}$,
-			    then each expand to colonize a lattice of demes with nearest-neighbor symmetric migration at time $\tau_{\text{e}}$,
-			    then finally at time $\tau_{\text{a}}$ collapse into a single lattice consisting of demes 
-			    with ancestry entirely in one or the other of the clusters,
-			    or admixed between them.
-			    }\label{sim_setup}
-\end{figure}
-
 \newpage
@@
     }\label{simK1_sp_pies}
-\end{figure}
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=5in,height=2.5in]{figs/sims/simK1_laycon_barplots.pdf}}
-		\caption{
-			Layer ``importance" (i.e., how much total covariance is contributed by each layer), 
-			for all layers estimated in runs using $K = 1$ through 7 
-			for the spatial model (left) and the nonspatial model (right).
-			Data were simulated using $K=1$.
-			For each value of $K$ along the x-axis, there are an equal number of layer contributions plotted.
-		}\label{simK1_laycon}
 \end{figure}

### 2017-09-10 writeup/conStruct.tex
poplar fastStr figs

@@
 and we demonstrate the utility of our approach using both simulated and empirical data.
-The implementation of this method, \texttt{conStruct}, is available as an R package at 
+The implementation of this method, \texttt{conStruct} (for ``\emph{con}tinuous \emph{struct}ure"), 
+is available as an R package at 
 \href{https://github.com/gbradburd/conStruct}{https://github.com/gbradburd/conStruct}.
@@
 and compared these models using cross-validation.
-The results of these analyses are shown in 
+For comparison, we also ran fastSTRUCTURE \citep{fastStructure} 
+using $K = 1$ through 7.
+The results of all these analyses are shown in 
 Figs \ref{populus_pies} and \ref{populus_xvals}, 
-as well as in Figs \ref{populus_sp_pies} - \ref{populus_laycon} in the Supplementary Materials.
+as well as in Figs \ref{populus_sp_pies} - \ref{populus_fastStr} in the Supplementary Materials.
 
@@
 
-For values of $K<4$, the spatial model was strongly preferred over the nonspatial model (Fig \ref{populus_xvals}).
+For values of $K<4$, the spatial model was strongly preferred over the nonspatial model (Fig \ref{populus_xvals}),
+the results of which were broadly concordant with those of the fastSTRUCTURE runs (Fig \ref{populus_fastStr}).
 Most of this signal seems to derive from the \textit{P. trichocarpa} samples:
@@
 	\caption{
-	Map of admixture proportions estimated for the \textit{Populus} dataset 
+	Maps of admixture proportions estimated for the \textit{Populus} dataset 
 	using the spatial model for $K=2$ through 7.
@@
 	\caption{
-	Map of admixture proportions estimated for the \textit{Populus} dataset 
+	Maps of admixture proportions estimated for the \textit{Populus} dataset 
 	using the nonspatial model for $K=2$ through 7.
@@
 	\centering
+		{\includegraphics[width=6in,height=4in]{figs/populus/poplar_fastStr_results.pdf}}
+	\caption{
+	Maps of admixture proportions estimated for the \textit{Populus} dataset 
+	using fastSTRUCTURE for $K=2$ through 7.
+    }\label{populus_fastStr}
+\end{figure}
+
+\begin{figure}
+	\centering
 		{\includegraphics[width=6in,height=4in]{figs/bears/bear_sp_results.pdf}}

### 2017-09-10 data/poplars/fastStructure/exe.poplar.fastStr.R

@@ -1,0 +1,10 @@
+setwd("/Applications/STRUCTURE/fastStructure")
+
+
+for(k in 2:7){
+	infile <- "/Users/bradburd/Dropbox/conStruct/data/poplars/fastStructure/poplars"
+	outfile <- paste0("/Users/bradburd/Dropbox/conStruct/data/poplars/fastStructure/poplars_K",k)
+	call <- paste0("python structure.py -K ",k," --input=",infile," --output=",outfile," --prior=logistic --format=str")
+	system(call)
+}
+

### 2017-09-10 data/poplars/fastStructure/format.poplar.fastStr.R

@@ -1,0 +1,40 @@
+make.ind.genos <- function(x,N){
+		geno <- numeric(N)
+	if(is.na(x)){
+		geno <- rep(-9,N)
+	} else {
+		geno[sample(1:N,x)] <- 1	
+	}
+	return(geno)
+}
+
+split.pop.into.inds <- function(pop.counts,pop.nChromo){
+	ind.geno <- apply(pop.counts,2,function(x){make.ind.genos(x,pop.nChromo)})
+	return(ind.geno)
+}
+
+make.structure.data.file <- function(counts,sample.sizes){
+	geno.matrix <- matrix(NA,sum(sample.sizes),ncol(freqs))
+	index <- 0
+	for(n in 1:nrow(freqs)){
+		geno.matrix[index+c(1:sample.sizes[n]),] <- split.pop.into.inds(counts[n,,drop=FALSE],sample.sizes[n])
+		index <- index + sample.sizes[n]
+	}
+	PopData <- unlist(lapply(1:length(sample.sizes),
+					function(n){
+						lapply(1:sample.sizes[n],
+							function(i){
+								sprintf("ind_%s.%s",n,i)
+							})
+					}))
+	LocData <- unlist(lapply(1:length(sample.sizes),function(n){rep(n,sample.sizes[n])}))
+	return(cbind(PopData,LocData,geno.matrix))
+}
+
+load("~/Dropbox/conStruct/data/poplars/poplar.data.redux.Robj")
+load("~/Dropbox/conStruct/data/poplars/fastStructure/poplar.sample.sizes.Robj")
+freqs <- poplar.data$freqs
+sample.sizes <- poplar.sample.sizes*2
+counts <- freqs*matrix(sample.sizes,nrow(freqs),ncol(freqs))
+str.data <- make.structure.data.file(counts,poplar.sample.sizes*2)
+write.table(str.data,file="~/Dropbox/conStruct/data/poplars/fastStructure/poplars.str",quote=FALSE,row.names=FALSE,col.names=FALSE)

### 2017-09-10 writeup/figs/populus_figs.R

@@ -24,7 +24,6 @@ dev.off()
 
 
 library(maps)
-my.rep <- 1
 load("~/Dropbox/conStruct/data/poplars/poplar.data.redux.Robj")
 col.mat1 <- matrix(ifelse(poplar.data$sp.ID=="Populus trichocarpa","forestgreen","black"),byrow=TRUE,length(poplar.data$sp.ID),length(poplar.data$sp.ID))
 col.mat2 <- matrix(ifelse(poplar.data$sp.ID=="Populus trichocarpa","forestgreen","black"),byrow=FALSE,length(poplar.data$sp.ID),length(poplar.data$sp.ID))
@@ -230,4 +229,40 @@ pdf(file="~/Dropbox/conStruct/writeup/figs/populus/populus_laycon_barplots.pdf",
 			xlab="",ylab="")
 			mtext(side=1,text="number of layers",padj=4.25,adj=-1)
 			mtext(side=3,text="Layer contributions (Poplars)",padj=-2.25,adj=7,font=2,cex=1.2)
+dev.off()
+
+
+collapse.ind.Q <- function(Q,pop.vec){
+	n.pops <- length(unique(pop.vec))
+	admix.props <- matrix(NA,n.pops,ncol(Q))
+	for(i in 1:n.pops){
+		admix.props[i,] <- colMeans(Q[which(pop.vec==i),,drop=FALSE])
+	}
+	return(admix.props)
+}
+load("~/Dropbox/conStruct/data/poplars/fastStructure/poplar.sample.sizes.Robj")
+poplar.pop.vec <- unlist(lapply(1:length(poplar.sample.sizes),function(n){rep(n,poplar.sample.sizes[n])}))
+
+pdf(file="~/Dropbox/conStruct/writeup/figs/populus/poplar_fastStr_results.pdf",width=12,height=8,pointsize=14)
+	par(mfrow=c(2,3),mar=c(5,3,0,1.5))
+	for(k in 2:7){
+		w <- collapse.ind.Q(Q = as.matrix(read.table(sprintf("~/Dropbox/conStruct/data/poplars/fastStructure/poplars_K%s.%s.meanQ",k,k),stringsAsFactors=FALSE)),
+							pop.vec = poplar.pop.vec)	
+		if(k <= 2){
+			csr1.order <- NULL
+		}
+		if(k > 2){
[…]

### 2017-09-10 writeup/cover_letter.tex
added cover letter

\documentclass[11pt]{letter}
\usepackage[hidelinks]{hyperref}
\usepackage[green]{msuletterhead}
%\topmarginskip{-0.35in}
%\bottommarginskip{-1.5in}
%\leftmarginsize{1in}
%\rightmarginsize{1.25in}
%\sigskipbefore{0.4in}
%\sigskipafter{0in}
%\noLines
%\nolines
%\noHeadline
%\noheadline
%\signature{Gideon Bradburd, Graham Coop, and Peter Ralph}
%
%\namefrom{}
%\addrfrom{Gideon Bradburd \\ Dept. of Integrative Biology \\ Michigan State University \\ East Lansing, MI 48823 USA \\ 
%\href{mailto:bradburd@msu.edu}{bradburd@msu.edu}}
%\emailfrom{bradburd@msu.edu}
%
%\greetto{To the Editor(s) --}
%\closeline{Sincerely,}

\begin{document}
%\begin{newlfm}

Submission Cover Letter\\
\emph{PLoS Biology}
\\\\\\
To the Editor(s),

Please find enclosed our manuscript,  
``Inferring Continuous and Discrete Population Genetic Structure Across Space,"
for consideration for publication in PLoS Biology.

In the manuscript, we present a novel statistical method 
for the simultaneous inference of continuous and discrete patterns of population structure.
The method estimates ancestry proportions for each 
sample from a set of discrete population layers, 
and, within each layer, estimates a rate at which relatedness decays with distance.
This model explicitly addresses the ``clines vs. clusters" problem in 
modeling population genetic variation by jointly accommodating 
continuous and discrete patterns of differentiation. 
We demonstrate the utility of this approach using both simulations 
and empirical applications, with datasets from poplars and black bears. 
We also release an associated R package, called \texttt{conStruct}, 
for ``\emph{con}tinuous \emph{struct}ure."

This work is motivated by two observations:
first, that patterns of isolation by distance (IBD) are ubiquitous in nature; 
and second, that existing model-based clustering methods, 
such as \texttt{STRUCTURE} (Pritchard \emph{et al.} 2000) 
or \texttt{ADMIXTURE} (Alexander \emph{et al.} 2009), 
can only describe continuous patterns of variation using discrete clusters, 
and so must use a possibly large number of clusters
to describe continuous population differentiation.
Inference of these spurious clusters can throw a 
wrench in the gears of efforts to describe and categorize natural variation, 
one of the first steps in any analysis of genetic data, 
and a principal mission of biology.

In this manuscript,
we use simulations and empirical data to demonstrate 
the ways in which nonspatial clustering approaches 
fail in the presence of IBD; 
we also present two different metrics for choosing an appropriate 
number of layers to describe the data, 
and, using both, 
are able to establish that our method performs quite well.

As of this submission, \texttt{STRUCTURE} alone has been cited more than 20,000 times, 
and the falling costs of sequencing are facilitating an explosion of next-generation sequence datasets.  
The first step in the analysis of these data will, for most researchers, 
involve a model-based clustering analysis.
We therefore believe that our work represents an important and timely contribution to the field.

John Novembre (jnovembre@uchicago.edu), \\
Desislava Petkova (desislavka@gmail.com), \\
Jonathan Pritchard (pritch@stanford.edu),  \\
Sohini Ramachandran (sramachandran@brown.edu), \\
and Matthew Stephens (mstephens@uchicago.edu) \\
would be suitable reviewers.

Sincerely,
Gideon Bradburd, 
Graham Coop, and 
Peter Ralph
%\end{newlfm}
\end{document}

%What is the scientific question you are addressing?
%What is the key finding that answers this question?
%What is the nature of the evidence you provide in support of your conclusion?
%What are the three most recently published articles that are relevant to this question?
%What significance do your results have for the field?
%What significance do your results have for the broader community (of biologists and/or the public)?
%What other novel findings do you present?
%Is there additional information that we should take into account?

### 2017-09-12 writeup/conStruct.tex
g's changes

@@
 (after the initial calculation of the allelic covariance), 
-it scales poorly with number of samples.
+it \gc{currently scales} poorly with number of samples.
 The computationally limiting step is the inversion of the parametric covariance, 
@@
 scales linearly with number of samples. \gb{is this true?}
-The other main difference is the model of noise,
-but it is unclear what effects this may have,
-or indeed, which is closer to the truth.
+\gc{Our speed, on large sample sizes, could be improved by adopting rank one updates to the
+ inverse of the covariance matrix for updating an individual's
+ admixture proportions, which alters only a single row/column of the
+ covariance matrix. However, we have not implemented this yet as it
+ would likely mean losing our Hamilton MCMC for our parameters.}
+
+%The other main difference is the model of noise,
+%but it is unclear what effects this may have,
+
+\gc{SO I THINK IF YOU SAY THIS you need to give some run time specs,
+  mainly to show that it's not insanely slow.}
+
+%Binomially clearly is, genotypes are discrete:  "or indeed, which is closer to the truth."
 
@@
 and using only that analysis to communicate results, 
-we want to emphasize that there are both statistical and biological reasons 
-why there is no true best value of $K$.
-From a statistical perspective, the choice of best $K$ is always relative to the data in hand.
-This means both that sampling different individuals or loci 
-from the same populations may yield a different ``best" $K$, 
-but also 
--- unless the data were generated under the model itself --
+\gc{we emphasize that there} is no true best value of $K$.
+\gc{The choice} of best $K$ is always relative to the data in hand.
+%This means both that sampling different individuals or loci 
+%from the same populations may yield a different ``best" $K$, 
+From the statistical perspective, Unless the data were generated under the model itself 
 that the support for larger values of $K$ is likely to increase with increasing amounts of data
-(although perhaps only logarithmically).
+%(although perhaps only logarithmically).
 In the limit of infinite data, the best value of $K$ 
-may be the number of individuals included in the dataset.
+may be the number of samples included in the dataset \citep{Patterson2006}.
 From a biological perspective, 
@@
 observe which layers split out at what levels 
-(effectively taking successively shallower cross-sections of the population phylogeny),
+(\gc{conceptually similar to} taking successively shallower
+cross-sections of the population phylogeny),  %%note that the
+                                %%equivalent would only be ~true if we
+                                %%had same sample size from each
+                                %%proportion.
 and also to take the results of the proposed cross-validation procedure with a large grain of salt.
 Calculating layer importance may also be a useful heuristic, 
-as it can reveal layers with strong statistical support but small biological import.
+as it can reveal layers with statistical support but small biological import.
+
+\gc{In adding to spatial realism to the groups used by clustering methods,
+it is important to re-emphasize that that clusters/layers detected by
+these methods do not necessarily accord to distinct, ancestral
+populations. Nor does a non-zero admixture proportion indicate that admixture (i.e. gene
+flow) must have occurred. Both groupings and admixture proportions
+should be viewed as hypotheses that should be subject to futher
+testing, see \citet{Falush:16} for futher discussion of these points. }
 
@@
 will be collapsed into each other.  
-This ``lumping" might better reflect biological reality, 
+This ``lumping" might better reflect biological \gc{``reality'',  
 but may also have implications for management decisions and conservation policy, 
@@
 It is therefore important to emphasize that individuals sampled from the same \texttt{conStruct} layer 
-may be quite genetically diverged from one another, 
+may be quite genetically diverged from one another, \gc{perhaps especially at
+  loci underlying adaptive traits}, 
 and that a \texttt{conStruct} layer may still contain multiple distinct MUs.  
@@
 inference may be misleading if the sampling geography departs radically from the way 
-the sampled organisms experience (or have experienced) their landscape.
+the sampled organisms experience (or have experienced) \gc{dispersal
+  on } their landscape.
 For example, if we were to run a \texttt{conStruct} analysis using geographic distances between 
@@
 \citep[e.g.][]{treemix}.
+
+%\gc{We could imagine modeling }
 

### 2017-09-12 writeup/conStruct.tex
incorporating graham's edits

@@
 (after the initial calculation of the allelic covariance), 
-it \gc{currently scales} poorly with number of samples.
+it currently scales poorly with number of samples.
 The computationally limiting step is the inversion of the parametric covariance, 
@@
 whereas computation time for, e.g., STRUCTURE, 
-scales linearly with number of samples. \gb{is this true?}
-\gc{Our speed, on large sample sizes, could be improved by adopting rank one updates to the
- inverse of the covariance matrix for updating an individual's
- admixture proportions, which alters only a single row/column of the
- covariance matrix. However, we have not implemented this yet as it
- would likely mean losing our Hamilton MCMC for our parameters.}
-
-%The other main difference is the model of noise,
-%but it is unclear what effects this may have,
+scales linearly with number of samples.
+Our speed, on datasets with large sample sizes, 
+could be improved by adopting rank-one updates 
+to the inverse of the covariance matrix 
+\citep[e.g.][]{woodbury1950,sherman_morrison1950}
+when updating a sample's admixture proportions, 
+which alters only a single row/column of the covariance matrix. 
+However, we have not implemented this yet, 
+as it would likely mean losing the ability to do efficient,  
+Hamiltonian Monte Carlo sampling of our parameters.
 
@@
   mainly to show that it's not insanely slow.}
-
-%Binomially clearly is, genotypes are discrete:  "or indeed, which is closer to the truth."
 
@@
 and using only that analysis to communicate results, 
-\gc{we emphasize that there} is no true best value of $K$.
-\gc{The choice} of best $K$ is always relative to the data in hand.
+we emphasize that there is no true best value of $K$.
+The choice of best $K$ is always relative to the data in hand.
 %This means both that sampling different individuals or loci 
@@
 may be the number of samples included in the dataset \citep{Patterson2006}.
+
 From a biological perspective, 
@@
 observe which layers split out at what levels 
-(\gc{conceptually similar to} taking successively shallower
-cross-sections of the population phylogeny),  %%note that the
-                                %%equivalent would only be ~true if we
-                                %%had same sample size from each
-                                %%proportion.
+(this is conceptually similar to taking successively shallower
+cross-sections of the population phylogeny), 
 and also to take the results of the proposed cross-validation procedure with a large grain of salt.
@@
 
-\gc{In adding to spatial realism to the groups used by clustering methods,
-it is important to re-emphasize that that clusters/layers detected by
-these methods do not necessarily accord to distinct, ancestral
-populations. Nor does a non-zero admixture proportion indicate that admixture (i.e. gene
-flow) must have occurred. Both groupings and admixture proportions
-should be viewed as hypotheses that should be subject to futher
-testing, see \citet{Falush:16} for futher discussion of these points. }
+Although we believe our model adds spatial realism to the groups used by clustering methods,
+it is important to note that the clusters/layers detected by our method 
+do not necessarily correspond  to distinct, ancestral populations; 
+nor does a non-zero admixture proportion indicate that admixture 
+(i.e. gene flow) must have occurred. 
+Both groupings and admixture proportions
+should be viewed as hypotheses that should be subject to further testing
+\citep[see][for an in-depth discussion of these points]{Falush:16}.
 
@@
 will be collapsed into each other.  
-This ``lumping" might better reflect biological \gc{``reality'',  
+This ``lumping" might better reflect biological ``reality'',  
 but may also have implications for management decisions and conservation policy, 
@@
 
-It is therefore important to emphasize that individuals sampled from the same \texttt{conStruct} layer 
-may be quite genetically diverged from one another, \gc{perhaps especially at
-  loci underlying adaptive traits}, 
+It is therefore important to stress that individuals sampled from the same \texttt{conStruct} layer 
+may be quite genetically diverged from one another, 
+perhaps especially at loci underlying adaptive traits, 
 and that a \texttt{conStruct} layer may still contain multiple distinct MUs.  
@@
 inference may be misleading if the sampling geography departs radically from the way 
-the sampled organisms experience (or have experienced) \gc{dispersal
-  on } their landscape.
+the sampled organisms disperse (or have dispersed) on their landscape.
 For example, if we were to run a \texttt{conStruct} analysis using geographic distances between 

### 2017-09-12 writeup/conStruct.tex
changes to disco

@@
 
-%\gc{We could imagine modeling }
-
+\gc{Here we have assumed that samples have known geographic
+  co-ordinates, and draw ancestry from the layers only at those sampled locations. A
+  natural extension would be to attempt to ``geo-locate'' the ancestry
+of samples without geographic co-ordinates \citep{Elephants}. We could
+also imagine letting samples drawn ancestry from other geographic
+co-ordinates, as we have done in a previous approach
+\citep{Spaceymix}, to model long distance dispersal. Indeed we could
+allow entire layers to draw their mean frequency from a particular
+location on another layer. This would enable the modeling of scenerios
+such as range expansion, or domestication, where
+a set of individuals are thought to have ancestry that originated from
+a particular geographic location embedded in a larger pattern of IBD. }
 %%Similarly, we currently assume that all loci are independent, 
@@
 
-A final exciting direction is to model relatedness within a cluster as a spatiotemporal process, 
+A final direction is to model relatedness within a \gc{layer} as a spatiotemporal process, 
 in which covariance decays both with distance in space and in time.  
@@
 Therefore, a natural way to test for population replacement is to estimate the rates 
-at which relatedness within a cluster decays with time in the same way we do in the current model with space, 
-in which case discrete population structure in space is comparable to population replacement in time.
+at which relatedness within a \gc{layer} decays with time in the same way we do in the current model with space, 
+in which case \gc{changes in discrete population structure across space is comparable to population replacement across time.}
 

### 2017-09-12 writeup/conStruct.tex
changes to appendix

@@
 the National Science Foundation under award number NSF \#1262645 (DBI) to PR and GC, 
-the National Institute of General Medical Sciences of the National Institutes of Health under award numbers NIH RO1GM83098 and RO1GM107374 to GC,
+the National Institute of General Medical Sciences of the National
+Institutes of Health under award numbers NIH R01-GM108779 to GC,
 and the National Science Foundation under award numbers NSF \# 1148897 and \# 1402725 to GB.
@@
 \subsection{Drift, admixture, and space} \label{rationale}
+\gc{Here we sketch a simple model of allele frequencies and their
+covariances, to justify the form given in the main text.}
 \paragraph{Drift} 
-We first provide a simple model of allele frequencies within a cluster. 
+We first provide a simple model of allele frequencies within a \gc{layer}. 
 Imagine a sample $i$ that draws all of its ancestry from cluster $k$.
@@
 the second is the deviation from that ancestral frequency 
-due to drift in the ancestral population of the $k$th cluster,
+due to drift in the ancestral population of the $k$th \gc{layer},
 which is shared by all samples within the cluster. 
-The third term is the deviation of the $i$th sample away from the $k$th cluster mean 
-due to the spatial process of drift and migration within the cluster.
+The third term is the deviation of the $i$th sample away from the
+$k$th \gc{layer} mean 
+due to the spatial process of drift and migration within the \gc{layer}.
 The final term is the deviation specific to the $i$th sample,
@@
 
-If we have two samples $i$ and $j$ drawn from cluster $k$, 
+If we have two samples $i$ and $j$ drawn from \gc{layer} $k$, 
 their covariance across loci will be 
@@
 The model above describes the simple case 
-in which samples draw 100\% of their ancestry from only a single cluster each. 
-To accommodate admixture between clusters, 
+in which samples draw 100\% of their ancestry from only a single \gc{layer} each. 
+To accommodate admixture between \gc{layers}, 
 we model allele frequencies within samples as linear combinations of the population-specific 
@@
 To do so, we introduce a term $w^{(k)}_{i}$ that describes the
-admixture proportion of sample $i$ in cluster $k$ ($\sum_{K}w^{(k)}_{i} = 1$),
+admixture proportion of sample $i$ in \gc{layer} $k$ ($\sum_{K}w^{(k)}_{i} = 1$),
 which can be interpreted as the proportion of the genome in the $i$th
-sample that came from the $k$th cluster 
+sample that came from the $k$th \gc{layer} 
 (or the probability that an allele at a locus is drawn from $k$).
@@
 \section{Allelic Covariance}\label{allelic_cov}
-To see how to arrive at Eqn. \eqref{allelic_covariance},
+To see how to arrive at \gc{our empirical variance-covariance expression,} Eqn. \eqref{allelic_covariance},
 pick a random locus and
@@
 \item For each model (i.e., each combination of $m$ and $k$) calculate 
-	the mean standardized log likelihood of the testing data partition across $X$ replicates, 
-	as well its standard error and 95\% confidence interval:
-	\begin{enumerate}
-	\item mean
-		\begin{equation}
-			\bar{\mathcal{Z}}_{mk} = \frac{1}{X}\sum\limits_{x=1}^{X}\mathcal{Z}_{xmk}
-		\end{equation}
-	\item standard error
-		\begin{equation}
-			SE_{\bar{\mathcal{Z}}_{mk}} = \sqrt{\frac{\sum\limits_{x=1}^{X} \left[ {\mathcal{Z}_{xmk} - \bar{\mathcal{Z}}_{mk}} \right]}{X}}
-		\end{equation}
-	\item 95\% confidence interval
-		\begin{equation}
-			95\% \text{CI} = \bar{\mathcal{Z}}_{mk} \pm 1.96 \times SE_{\bar{\mathcal{Z}}_{mk}}
-		\end{equation}
-	\end{enumerate}
-\end{enumerate}
+	the mean ($	\bar{\mathcal{Z}}_{mk} $) standardized log likelihood of the testing data partition across $X$ replicates, 
+	as well its standard error ($SE_{\bar{\mathcal{Z}}_{mk}}$) and
+        95\% confidence interval ($\bar{\mathcal{Z}}_{mk} \pm 1.96 \times SE_{\bar{\mathcal{Z}}_{mk}}$).
+% 	\begin{enumerate}
+% 	\item mean
+% 		\begin{equation}
+% 			\bar{\mathcal{Z}}_{mk} = \frac{1}{X}\sum\limits_{x=1}^{X}\mathcal{Z}_{xmk}
+% 		\end{equation}
+% 	\item standard error
+% 		\begin{equation}
+% 			SE_{\bar{\mathcal{Z}}_{mk}} = \sqrt{\frac{\sum\limits_{x=1}^{X} \left[ {\mathcal{Z}_{xmk} - \bar{\mathcal{Z}}_{mk}} \right]}{X}}
+% 		\end{equation}
+% 	\item 95\% confidence interval
+% 		\begin{equation}
+% 			95\% \text{CI} = \bar{\mathcal{Z}}_{mk} \pm 1.96 \times SE_{\bar{\mathcal{Z}}_{mk}}
+% 		\end{equation}
+% 	\end{enumerate}
+% \end{enumerate}
 
 \newpage
-\section{Calculating layer importance}\label{layer_importance}
+\section{Calculating layer \gc{contributions}}\label{layer_importance}
 First, we describe the allelic covariance between two samples, $i$, and $j$.
@@
 first each sample randomly chooses a layer: call these $I$ and $J$.
-The probability that $X$ chooses layer $k$ is $\P{I=k}$, which is $w_i^{(k)}$, 
+The probability that $X$ chooses layer $k$ is \gc{$\P({I=k})$}, which is $w_i^{(k)}$, 
 the $i$th sample's admixture proportion in the $k$th layer.
@@
 Similarly, the ``background" contribution is $p_\gamma / (p_\gamma + \sum_K w_i^{(k)} w_j^{(k)} q^{(k)}_{ij})$.
-Because our signal comes from \emph{variation} in covariance, we omit the $p_\gamma$ terms.
+Because our signal comes from \emph{variation} in covariance, we omit
+the $p_\gamma$ terms. \gc{Stated in this way this quantity is the
+relative contribution of the $k$th layer to the (model-based) kinship
+coefficient between $i$ and $j$.}
 
@@
 summed over the upper triangle (excluding the diagonal) of the covariance matrix.
-We define the importance of the $k$th layer, $\Xi^{(k)}$, 
+We define the \gc{contribution} of the $k$th layer, $\Xi^{(k)}$, 
 as the relative contribution of the $k$th layer to total agreement:

### 2017-09-13 writeup/conStruct.tex
minor edits

--- 2017-09-12 writeup/conStruct.tex
@@
 An additional heuristic to evaluate the inclusion of additional layers is to calculate 
-each layer's relative contribution to total covariance as their importance, 
-and to choose a value of $K$ where all layers have an importance above some cutoff (e.g. 0.1\%).
+each layer's relative contribution to total covariance, 
+and to choose a value of $K$ where all layers have a contribution above some cutoff (e.g. 0.1\%).
 Because the prior on admixture proportions is a Dirichlet distribution 
@@
 summed over the upper triangle (excluding the diagonal) of the covariance matrix.
-The importance of the $k$th layer is its contribution relative to those of all other layers, 
+The contribution of the $k$th layer is its absolute contribution relative to those of all other layers, 
 i.e., its contribution divided by the sum of the contributions of all $K$ layers
-(for details on calculating this importance, see Appendix \secref{layer_importance}). 
-This layer importance is similar to the ``ancestry contribution'' proposed by \cite{fastStructure}. 
+(for details on calculating this contribution, see Appendix \secref{layer_contribution}). 
+This layer contribution is similar to the ``ancestry contribution'' proposed by \cite{fastStructure}. 
 However, each of our layers can induce a different amount of covariance between samples embedded in them, 
@@
 
-Finally, we calculated layer importance for 
+Finally, we calculated layer contribution for 
 analyses run in each simulation 
@@
 		\caption{
-			Layer ``importance" (i.e., how much total covariance is contributed by each layer), 
+			Layer contributions (i.e., how much total covariance is contributed by each layer), 
 			for all layers estimated in runs using $K = 1$ through 7 
@@
 	\caption{
-	Layer ``contributions" (i.e., how much total covariance is contributed by each layer), 
+	Layer contributions (i.e., how much total covariance is contributed by each layer), 
 	for all layers estimated in runs using $K = 1$ through 7 
@@
 that Canada was not sampled, 
-and so the blue layer may extend
+and so the red layer may extend
 through the intervening (unsampled) northern Great Plains and Canadian Shield, 
-with the red layer presumably then stretching up into British Columbia.
+with the blue layer presumably then stretching up into British Columbia.
 
@@
 up until around $K=5$ or 6 (Fig \ref{bear_xvals}).
-However, upon examining layer importance (Fig \ref{bear_laycon}),
+However, upon examining layer contributions (Fig \ref{bear_laycon}),
 it is clear that the additional spatial layers beyond $K=2$ are contributing only a very small amount toward 
@@
 Hamiltonian Monte Carlo sampling of our parameters.
-
-\gc{SO I THINK IF YOU SAY THIS you need to give some run time specs,
-  mainly to show that it's not insanely slow.}
+On a desktop machine with a 4.2 GHz Intel Core i7 processor, 
+an analysis of the the poplar dataset (35 samples, 22,500 loci), 
+running the spatial model with 7 layers for 5,000 MCMC iterations 
+took 29.8 minutes.
+For the black bear dataset (84 samples, 21,000 loci) 
+running the spatial model with 7 layers for 5,000 MCMC iterations, 
+which was more than sufficient for convergence,
+took 6.9 hours.
 
@@
 and also to take the results of the proposed cross-validation procedure with a large grain of salt.
-Calculating layer importance may also be a useful heuristic, 
+Calculating layer contributions may also be a useful heuristic, 
 as it can reveal layers with statistical support but small biological import.
@@
 will be collapsed into each other.  
-This ``lumping" might better reflect biological ``reality'',  
+This ``lumping" might better reflect biological reality, 
 but may also have implications for management decisions and conservation policy, 
@@
 perhaps especially at loci underlying adaptive traits, 
-and that a \texttt{conStruct} layer may still contain multiple distinct MUs.  
-Alternatively, the collapse of multiple, previously identified MUs into a single \texttt{conStruct} layer 
-may reflect a biological reality that these populations are currently 
+and that a \texttt{conStruct} layer may still contain multiple distinct MUs worthy of independent protections.  
+Alternatively, the inclusion of multiple MUs into a single \texttt{conStruct} layer 
+may occur if these populations are currently 
 (or were recently) exchanging migrants, 
@@
 \newpage
-\section{Calculating layer importance}\label{layer_importance}
+\section{Calculating layer contributions}\label{layer_contribution}
 First, we describe the allelic covariance between two samples, $i$, and $j$.
@@
 summed over the upper triangle (excluding the diagonal) of the covariance matrix.
-We define the importance of the $k$th layer, $\Xi^{(k)}$, 
+We define the contribution of the $k$th layer, $\Xi^{(k)}$, 
 as the relative contribution of the $k$th layer to total agreement:

### 2017-09-13 writeup/conStruct.tex

@@
 
-%\gc{We could imagine modeling }
-
-%%Similarly, we currently assume that all loci are independent, 
-%such that the number of loci in the dataset gives the degrees of 
-%freedom of the Wishart distribution in our likelihood calculations.  
-%Linkage disequilibrium (LD) between loci will decrease 
-%the effective number of degrees of freedom, 
-%artificially increasing users' confidence in their results.
-
-A final exciting direction is to model relatedness within a cluster as a spatiotemporal process, 
+In addition, here we have assumed that samples have known geographic coordinates, 
+and that they draw ancestry from layers only at those sampled locations. 
+A natural extension would be to attempt to ``geo-locate'' 
+the ancestry of samples without geographic coordinates \citep{Wasser2004}. 
+We could also imagine letting samples draw ancestry from other geographic coordinates, 
+as we have done in a previous approach \citep{spacemix} to model long distance dispersal. 
+Indeed, we could allow entire layers to bud off of a particular location on another layer. 
+This would enable us to model scenarios of range expansion or domestication, 
+in which a set of individuals are thought to have ancestry that originated from
+a particular geographic location embedded in a larger pattern of IBD.
+
+A final direction is to model relatedness within a layer as a spatiotemporal process, 
 in which covariance decays both with distance in space and in time.  
@@
 or whether populations are being replaced \citep{lazaridis_ancient_2014, Haak2015, slatkin_racimo2016, Nielsen2017, Schraiber2017}.
-However, we expect allele frequencies to change through time in a population, even without replacement, 
-simply due to drift.
+However, we expect allele frequencies to change through time in a population, 
+even without replacement, simply due to drift.
 Therefore, a natural way to test for population replacement is to estimate the rates 
-at which relatedness within a cluster decays with time in the same way we do in the current model with space, 
-in which case discrete population structure in space is comparable to population replacement in time.
+at which relatedness within a layer decays with time in the same way we do in the current model with space, 
+in which case a change in discrete population structure across space is comparable to population replacement across time.
 
@@
 \subsection{Drift, admixture, and space} \label{rationale}
+Here we sketch a simple model of allele frequencies and their
+covariances, to justify the form given in the main text.
 \paragraph{Drift} 
-We first provide a simple model of allele frequencies within a cluster. 
+We first provide a simple model of allele frequencies within a layer. 
 Imagine a sample $i$ that draws all of its ancestry from cluster $k$.
@@
 the second is the deviation from that ancestral frequency 
-due to drift in the ancestral population of the $k$th cluster,
+due to drift in the ancestral population of the $k$th layer,
 which is shared by all samples within the cluster. 
-The third term is the deviation of the $i$th sample away from the $k$th cluster mean 
-due to the spatial process of drift and migration within the cluster.
+The third term is the deviation of the $i$th sample away from the $k$th layer mean 
+due to the spatial process of drift and migration within the layer.
 The final term is the deviation specific to the $i$th sample,
@@
 
-If we have two samples $i$ and $j$ drawn from cluster $k$, 
+If we have two samples $i$ and $j$ drawn from layer $k$, 
 their covariance across loci will be 
@@
 The model above describes the simple case 
-in which samples draw 100\% of their ancestry from only a single cluster each. 
-To accommodate admixture between clusters, 
+in which samples draw 100\% of their ancestry from only a single layer each. 
+To accommodate admixture between layers, 
 we model allele frequencies within samples as linear combinations of the population-specific 
@@
 To do so, we introduce a term $w^{(k)}_{i}$ that describes the
-admixture proportion of sample $i$ in cluster $k$ ($\sum_{K}w^{(k)}_{i} = 1$),
+admixture proportion of sample $i$ in layer $k$ ($\sum_{K}w^{(k)}_{i} = 1$),
 which can be interpreted as the proportion of the genome in the $i$th
-sample that came from the $k$th cluster 
+sample that came from the $k$th layer 
 (or the probability that an allele at a locus is drawn from $k$).
@@
 \section{Allelic Covariance}\label{allelic_cov}
-To see how to arrive at Eqn. \eqref{allelic_covariance},
+To see how to arrive at our empirical variance-covariance expression, Eqn. \eqref{allelic_covariance},
 pick a random locus and
@@
 \item For each model (i.e., each combination of $m$ and $k$) calculate 
-	the mean standardized log likelihood of the testing data partition across $X$ replicates, 
-	as well its standard error and 95\% confidence interval:
-	\begin{enumerate}
-	\item mean
-		\begin{equation}
-			\bar{\mathcal{Z}}_{mk} = \frac{1}{X}\sum\limits_{x=1}^{X}\mathcal{Z}_{xmk}
-		\end{equation}
-	\item standard error
-		\begin{equation}
-			SE_{\bar{\mathcal{Z}}_{mk}} = \sqrt{\frac{\sum\limits_{x=1}^{X} \left[ {\mathcal{Z}_{xmk} - \bar{\mathcal{Z}}_{mk}} \right]}{X}}
-		\end{equation}
-	\item 95\% confidence interval
-		\begin{equation}
-			95\% \text{CI} = \bar{\mathcal{Z}}_{mk} \pm 1.96 \times SE_{\bar{\mathcal{Z}}_{mk}}
-		\end{equation}
-	\end{enumerate}
-\end{enumerate}
+	the mean ($	\bar{\mathcal{Z}}_{mk} $) standardized log likelihood of the testing data partition across $X$ replicates, 
+	as well its standard error ($SE_{\bar{\mathcal{Z}}_{mk}}$) and
+        95\% confidence interval ($\bar{\mathcal{Z}}_{mk} \pm 1.96 \times SE_{\bar{\mathcal{Z}}_{mk}}$).
+% 	\begin{enumerate}
+% 	\item mean
+% 		\begin{equation}
+% 			\bar{\mathcal{Z}}_{mk} = \frac{1}{X}\sum\limits_{x=1}^{X}\mathcal{Z}_{xmk}
+% 		\end{equation}
+% 	\item standard error
+% 		\begin{equation}
+% 			SE_{\bar{\mathcal{Z}}_{mk}} = \sqrt{\frac{\sum\limits_{x=1}^{X} \left[ {\mathcal{Z}_{xmk} - \bar{\mathcal{Z}}_{mk}} \right]}{X}}
+% 		\end{equation}
+% 	\item 95\% confidence interval
+% 		\begin{equation}
+% 			95\% \text{CI} = \bar{\mathcal{Z}}_{mk} \pm 1.96 \times SE_{\bar{\mathcal{Z}}_{mk}}
+% 		\end{equation}
+% 	\end{enumerate}
+% \end{enumerate}
 
@@
 first each sample randomly chooses a layer: call these $I$ and $J$.
-The probability that $X$ chooses layer $k$ is $\P{I=k}$, which is $w_i^{(k)}$, 
+The probability that $X$ chooses layer $k$ is $\P({I=k})$, which is $w_i^{(k)}$, 
 the $i$th sample's admixture proportion in the $k$th layer.
@@
 Similarly, the ``background" contribution is $p_\gamma / (p_\gamma + \sum_K w_i^{(k)} w_j^{(k)} q^{(k)}_{ij})$.
-Because our signal comes from \emph{variation} in covariance, we omit the $p_\gamma$ terms.
+Because our signal comes from \emph{variation} in covariance, we omit
+the $p_\gamma$ terms. 
+Stated in this way, this quantity is the
+relative contribution of the $k$th layer to the (model-based) kinship
+coefficient between $i$ and $j$.
 

### 2017-09-13 writeup/cover_letter.tex
tweaked cover letter

@@
 
-Submission Cover Letter\\
+Submission Cover Letter \hfill	\today\\
 \emph{PLoS Biology}
@@
 As of this submission, \texttt{STRUCTURE} alone has been cited more than 20,000 times, 
-and the falling costs of sequencing are facilitating an explosion of next-generation sequence datasets.  
+and the ever-falling costs of sequencing are facilitating an explosion of next-generation sequence datasets.  
 The first step in the analysis of these data will, for most researchers, 

### 2017-09-13 writeup/conStruct.tex
added makefile; ended enumerate
added bear fastStr figs

@@
 and compared these models using cross-validation.
+We also ran fastSTRUCTURE \citep{fastStructure} on the same dataset, 
+using $K = 2$ through 4.
 The results of these analyses are shown in Figs \ref{bear_K3} - \ref{bear_laycon}, 
-as well as in Figs \ref{bear_sp_pies} - \ref{bear_nsp_clst_covs} in the Supplementary Materials.
+as well as in Figs \ref{bear_sp_pies} - \ref{bear_fastStr} in the Supplementary Materials.
 
@@
 the third cluster (in green) exhibits a clear East-West gradient that 
-is overlain across the discrete structure between the Southwest cluster and the Southeast
+is overlain across the discrete structure between the Southwest cluster and the Southeast.
+The results from fastSTRUCTURE are are not identical to 
+those obtained using the nonspatial model, 
+but they do showcase the same tendency: 
+e.g., at $K=3$, fastSTRUCTURE splits the westernmost Alaskan samples 
+out of the cluster with the eastern samples, 
+and at $K=4$, it splits them into their own cluster entirely \ref{bear_fastStr}.
 
@@
 
-A table of all parameters, their descriptions, and their priors is given in Table \ref{tab:param_prior_tab} below.
+A table of all parameters, their descriptions, and their priors is given in Table \ref{tab:param_prior_tab}.
 
@@
         95\% confidence interval ($\bar{\mathcal{Z}}_{mk} \pm 1.96 \times SE_{\bar{\mathcal{Z}}_{mk}}$).
-% 	\begin{enumerate}
-% 	\item mean
-% 		\begin{equation}
-% 			\bar{\mathcal{Z}}_{mk} = \frac{1}{X}\sum\limits_{x=1}^{X}\mathcal{Z}_{xmk}
-% 		\end{equation}
-% 	\item standard error
-% 		\begin{equation}
-% 			SE_{\bar{\mathcal{Z}}_{mk}} = \sqrt{\frac{\sum\limits_{x=1}^{X} \left[ {\mathcal{Z}_{xmk} - \bar{\mathcal{Z}}_{mk}} \right]}{X}}
-% 		\end{equation}
-% 	\item 95\% confidence interval
-% 		\begin{equation}
-% 			95\% \text{CI} = \bar{\mathcal{Z}}_{mk} \pm 1.96 \times SE_{\bar{\mathcal{Z}}_{mk}}
-% 		\end{equation}
-% 	\end{enumerate}
-% \end{enumerate}
+\end{enumerate}
 
@@
 \xi^{(k)} = 
-	\sum\limits_{i=1}^{N-1}
-		\sum\limits_{j=i+1}^N
+	\sum\limits_{i=1}^{N}
+		\sum\limits_{j=i}^N
 			w^{(k)}_iw^{(k)}_j 
@@
 	Maps of admixture proportions estimated for the \textit{Populus} dataset 
-	using fastSTRUCTURE for $K=2$ through 7.
+	using fastSTRUCTURE \citep{fastStructure} for $K=2$ through 7.
     }\label{populus_fastStr}
@@
 
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=2in]{figs/bears/bear_fastStr_results.pdf}}
+	\caption{
+	Maps of admixture proportions estimated for the black bear dataset 
+	using fastSTRUCTURE \citep{fastStructure} for $K=2$ through 4.
+    }\label{bear_fastStr}
+\end{figure}
+
 \newpage

### 2017-09-13 data/bears/fastStructure/exe.bear.fastStr.R

@@ -1,0 +1,10 @@
+setwd("/Applications/STRUCTURE/fastStructure")
+
+
+for(k in 2:7){
+	infile <- "/Users/bradburd/Dropbox/conStruct/data/bears/fastStructure/bears"
+	outfile <- paste0("/Users/bradburd/Dropbox/conStruct/data/bears/fastStructure/bears_K",k)
+	call <- paste0("python structure.py -K ",k," --input=",infile," --output=",outfile," --prior=logistic --format=str")
+	system(call)
+}
+

### 2017-09-13 data/bears/fastStructure/format.bear.fastStr.R

@@ -1,0 +1,40 @@
+make.ind.genos <- function(x,N){
+		geno <- numeric(N)
+	if(is.na(x)){
+		geno <- rep(-9,N)
+	} else {
+		geno[sample(1:N,x)] <- 1	
+	}
+	return(geno)
+}
+
+split.pop.into.inds <- function(pop.counts,pop.nChromo){
+	ind.geno <- apply(pop.counts,2,function(x){make.ind.genos(x,pop.nChromo)})
+	return(ind.geno)
+}
+
+make.structure.data.file <- function(counts,sample.sizes){
+	geno.matrix <- matrix(NA,sum(sample.sizes),ncol(freqs))
+	index <- 0
+	for(n in 1:nrow(freqs)){
+		geno.matrix[index+c(1:sample.sizes[n]),] <- split.pop.into.inds(counts[n,,drop=FALSE],sample.sizes[n])
+		index <- index + sample.sizes[n]
+	}
+	PopData <- unlist(lapply(1:length(sample.sizes),
+					function(n){
+						lapply(1:sample.sizes[n],
+							function(i){
+								sprintf("ind_%s.%s",n,i)
+							})
+					}))
+	LocData <- unlist(lapply(1:length(sample.sizes),function(n){rep(n,sample.sizes[n])}))
+	return(cbind(PopData,LocData,geno.matrix))
+}
+
+load("~/Dropbox/conStruct/data/bears/bear.dataset.redux.Robj")
+freqs <- bear.dataset$sample.freqs
+sample.sizes <- matrix(2,nrow(freqs),ncol(freqs))
+sample.sizes[which(is.na(freqs),arr.ind=TRUE)] <- 0
+counts <- freqs*sample.sizes
+str.data <- make.structure.data.file(counts,rep(2,nrow(freqs)))
+write.table(str.data,file="~/Dropbox/conStruct/data/bears/fastStructure/bears.str",quote=FALSE,row.names=FALSE,col.names=FALSE)

### 2017-09-13 writeup/figs/bear_figs.R

@@ -229,16 +229,10 @@ pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bears_laycon_barplots.pdf",widt
 dev.off()
 
 
-pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bear_fastStr_results.pdf",width=15,height=10,pointsize=14)
-	layout(
-		rbind(
-			Reduce("cbind",lapply(1:3,function(x){
-				matrix(((x-1)*2)+c(rep(1,10),rep(2,15)),
-					nrow=5,ncol=5,byrow=TRUE)})),
-			Reduce("cbind",lapply(4:6,function(x){
-				matrix(((x-1)*2)+c(rep(1,10),rep(2,15)),
-					nrow=5,ncol=5,byrow=TRUE)}))
-			))
+pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bear_fastStr_results.pdf",width=15,height=5,pointsize=14)
+	layout(Reduce("cbind",lapply(1:3,function(x){
+			matrix(((x-1)*2)+c(rep(1,10),rep(2,15)),
+					nrow=5,ncol=5,byrow=TRUE)})))
 	for(k in 2:4){
 		w <- as.matrix(read.table(sprintf("~/Dropbox/conStruct/data/bears/fastStructure/bears_K%s.%s.meanQ",k,k),stringsAsFactors=FALSE))
 		if(k <= 2){
@@ -251,11 +245,11 @@ pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bear_fastStr_results.pdf",width
 		if(is.null(csr1.order)){
 			csr1.order <- 1:k
 		}
-		make.bear.redux.result.plot.multipanel1(admix.proportions = w,
+		make.bear.redux.result.plot.multipanel1(admix.proportions = w[,csr1.order],
 												coords = bear.dataset$sample.coords,
 												lump.dist = 200,
 												cluster.colors = cluster.colors,
 												cluster.order=NULL)
-		mtext(side=1,text=bquote(paste("(",.(letters[k-1]),") ",italic("K")," = ",.(k))),padj=2.7,adj=0.4)
+		mtext(side=1,text=bquote(paste("(",.(letters[k-1]),") ",italic("K")," = ",.(k))),padj=0.8,adj=0.4)
 	}
 dev.off()

### 2017-09-13 writeup/figs/fastStr_figs.R

@@ -93,3 +93,4 @@ pdf(file="~/Dropbox/conStruct/writeup/figs/fastStr/fastStr_simK3_pies.pdf",width
 				 cluster.colors = c(2,4,"goldenrod1","forestgreen"))
 			mtext(side=3,font=2,text="K=4")
 dev.off()
+

### 2017-09-13 writeup/conStruct.tex
fastStr time comparison

@@
 Given this, when should methods without spatial capability be used?
-One advantage these have over \texttt{conStruct} is speed.
+One advantage these have over \texttt{conStruct} is speed when the number of samples is large.
 Although \texttt{conStruct}'s computation time is invariant 
@@
 which alters only a single row/column of the covariance matrix. 
-However, we have not implemented this yet, 
+We have not implemented this yet, 
 as it would likely mean losing the ability to do efficient,  
 Hamiltonian Monte Carlo sampling of our parameters.
+
+For a relatively small number of samples, 
+\texttt{conStruct} can be much faster than other nonspatial clustering methods.
 On a desktop machine with a 4.2 GHz Intel Core i7 processor, 
-an analysis of the the poplar dataset (35 samples, 22,500 loci), 
-running the spatial model with 7 layers for 5,000 MCMC iterations 
-took 29.8 minutes.
-For the black bear dataset (84 samples, 21,000 loci) 
-running the spatial model with 7 layers for 5,000 MCMC iterations, 
+an analysis of the black bear dataset (84 samples, 21,000 loci) 
+running \texttt{conStruct}'s spatial model with 4 layers for 5,000 MCMC iterations, 
 which was more than sufficient for convergence,
-took 6.9 hours.
+took XXX hours.
+On the same machine and dataset, 
+a fastSTRUCTURE run with $K=4$ took 3.76 days.
+However, for a large number of samples, 
+fastSTRUCTURE (or ADMIXTURE) would be much faster.
 

### 2017-09-13 writeup/conStruct.tex
changed cluster to layer in figs and text

@@
 but less effectively homogenizes geographic distant demes,
-resulting in a continuous pattern of isolation by distance within each cluster.
+resulting in a continuous pattern of isolation by distance within each layer.
 There is a fixed amount of covariance between layers, irrespective of spatial location.
@@
 
-To allow flexibility in the form of the decay of allelic covariance with geographic distance within each cluster, 
-we define the covariance within cluster $k$ between samples $i$ and $j$ to be:
+To allow flexibility in the form of the decay of allelic covariance with geographic distance within each layer, 
+we define the covariance within layer $k$ between samples $i$ and $j$ to be:
 \begin{equation}
@@
     \alpha^{(k)}_0 \left( \text{exp} \left( -(\alpha^{(k)}_D D_{i,j}) ^ {\alpha^{(k)}_2}	\right) \right) + \phi^{(k)}
-\label{within_cluster_covariance}
+\label{within_layer_covariance}
 \end{equation}
-where the superscript $(k)$ denotes parameters specific to the $k$th cluster.
+where the superscript $(k)$ denotes parameters specific to the $k$th layer.
 The quantity $D_{i,j}$ is the observed geographic distance between samples $i$ and $j$,
 the $\alpha^{(k)}$ parameters control the shape of the decay of
-covariance with distance in the cluster.
+covariance with distance in the layer.
 Our choice of a powered-exponential decay, 
@@
 and is not chosen to match a particular population genetics model. 
-The $\phi^{(k)}$ is a parameter that describes the background covariance within the cluster. 
-If two samples draw 100\% of their ancestry from cluster $k$, then their covariance under the model is $G^{(k)}_{i,j}$;
+The $\phi^{(k)}$ is a parameter that describes the background covariance within the layer. 
+If two samples draw 100\% of their ancestry from layer $k$, then their covariance under the model is $G^{(k)}_{i,j}$;
 if they are furthermore geographically very close ($D_{i,j}=0$)
@@
 If the geographic distance between them is very large, 
-their covariance will be equal to to the background level $\phi^{(k)}$ within the cluster.
+their covariance will be equal to to the background level $\phi^{(k)}$ within the layer.
 The ``shared drift'' parameter $\phi^{(k)}$ can be interpreted as the branch length 
@@
 
-We then allow samples to draw their ancestry from more than one cluster.
-The ``admixture'' proportion of the $i$th sample in the $k$th cluster, denoted $w^{(k)}_i$,
+We then allow samples to draw their ancestry from more than one layer.
+The ``admixture'' proportion of the $i$th sample in the $k$th layer, denoted $w^{(k)}_i$,
 gives the genome-wide proportion of alleles from sample $i$ that derive from
-cluster $k$ (and so $\sum_K w^{(k)}_i =1$).
+layer $k$ (and so $\sum_K w^{(k)}_i =1$).
 A visual representation of the method mechanics is shown in Fig \ref{schematic}.
@@
 We can then describe the covariance between samples $i$ and $j$ across all layers, $\Omega_{i,j}$,
-by summing their within-cluster spatial covariances ($G_{i,j}^{(k)}$ in cluster $k$) across all $K$ layers,
+by summing their within-layer spatial covariances ($G_{i,j}^{(k)}$ in layer $k$) across all $K$ layers,
 weighted by the relevant admixture proportions.
@@
 G^{(k)}_{i,j} + \delta_{i,j}\eta_i .
-\label{cross_cluster_covariance}
+\label{cross_layer_covariance}
 \end{equation}
 In this equation, $w^{(k)}_i w^{(k)}_j$ is the proportion of alleles that \emph{both}
-sample $i$ and sample $j$ have inherited from cluster $k$.
-
-In addition to the admixture-weighted sum of the within-cluster spatial covariances,
+sample $i$ and sample $j$ have inherited from layer $k$.
+
+In addition to the admixture-weighted sum of the within-layer spatial covariances,
 this function contains two terms, $\gamma$ and $\delta_{i,j}\eta_i$.
@@
 
-A nice feature of our approach is that the model described in Eqn \eqref{cross_cluster_covariance} 
+A nice feature of our approach is that the model described in Eqn \eqref{cross_layer_covariance} 
 contains a nonspatial assignment model as a special case 
@@
 
-\paragraph{Choice of cluster number and cross-validation}
+\paragraph{Choice of layer number and cross-validation}
 There are a number of reasons 
@@
 Several approaches have been used as model choice criteria 
-for the number of clusters in population genetic data, including: 
+for the number of discrete clusters in population genetic data, including: 
 information theoretic criteria \citep[e.g.][]{ADMIXTURE}, such as AIC or BIC;
@@
 that are spatially autocorrelated (see Appendix \secref{sim_details} for details), 
-so that both the frequencies in layers and the cluster memberships are
+so that both the frequencies in layers and the layer memberships are
 spatially autocorrelated. 
@@
 In Fig \ref{simK1_pies}, we show the results of the analyses of the dataset simulated under $K=1$, 
-contrasting the results of the nonspatial model runs with 2 - 4 clusters 
+contrasting the results of the nonspatial model runs with 2 - 4 layers 
 with those of spatial model with 2 - 4 layers.
@@
 (Fig \ref{sim_xvals}). Second, the predictive accuracy of the nonspatial model increases 
-with the number of clusters in the model, 
+with the number of layers in the model, 
 as it describes continuous, spatial patterns of relatedness between demes within the same cluster 
@@
 subsequent layers add little in terms of (co)variance explained; 
-in contrast, additional layers in the nonspatial analyses continue to
+in contrast, additional clusters in the nonspatial analyses continue to
 be important for describing variation, 
@@
 	and also the spatial model for $K=2$ through 4 (bottom row).
-	The data were simulated using a single cluster with nearest-neighbor symmetric migration between demes.
+	The data were simulated using a single layer with nearest-neighbor symmetric migration between demes.
     }\label{simK1_pies}
@@
 making the system well-suited to test an analysis designed to describe
-how genetic structure is partitioned both discretely across population (or species) clusters
+how genetic structure is partitioned both discretely across populations (or species)
 and continuously with geographic distance.
@@
 \item Given the known history of hybridization between this pair of poplar species, 
-is there support for two discrete clusters, 
+is there support for two discrete populations, 
 rather than a single cline of ancestry across their overlap zone?
@@
 \item Is there support for a pattern of isolation by distance within
-the inferred clusters, and, if so, 
+the inferred layers, and, if so, 
 does it differ between different layers?
@@
 covariance within predominantly \tri{} layers decays more quickly with distance
-than within the majority \bals{} layer, as seen in Fig \ref{populus_sp_clst_covs}.  
+than within the majority \bals{} layer, as seen in Fig \ref{populus_sp_layer_covs}.  
 This indicates that, although isolation by distance is a feature of the dataset as a whole, 
@@
 \begin{enumerate}
-\item How many historical clusters are reflected in modern patterns of genetic variation?
+\item How many historical populations are reflected in modern patterns of genetic variation?
 \item Do we see support for a pattern of isolation by distance within each of the inferred discrete groups?
@@
 
-\paragraph{Choosing the ``best" number of clusters}
+\paragraph{Choosing the ``best" number of layers}
 Although we recognize the utility of choosing a single, ``best'' value of $K$, 
@@
 A second caveat is that, in some instances, 
-membership in the same cluster may not mean that samples are particularly related.
-If covariance within a cluster decays sharply with distance, 
-and the cluster-specific relatedness parameter $\phi^{(k)}$ is low, 
-individuals at a far spatial remove may be in the same cluster but have very low pairwise relatedness.
+membership in the same layer may not mean that samples are particularly related.
+If covariance within a layer decays sharply with distance, 
+and the layer-specific relatedness parameter $\phi^{(k)}$ is low, 
+individuals at a far spatial remove may be in the same layer but have very low pairwise relatedness.
 It is possible that this is happening in Fig \ref{populus_sp_pies}. 
-At $K=3$, the southernmost populations of \textit{P. trichocarpa} are in the gold cluster, 
-whose other neighbors are to the north, with an intervening group of populations in the red cluster, 
-and at $K=5$, those southernmost samples split out and become their own cluster.
+At $K=3$, the southernmost populations of \textit{P. trichocarpa} are in the gold layer, 
+whose other neighbors are to the north, with an intervening group of populations in the red layer, 
+and at $K=5$, those southernmost samples split out and become their own layer.
 Again, we encourage users to run analyses across multiple values of $K$, 
@@
 We first provide a simple model of allele frequencies within a layer. 
-Imagine a sample $i$ that draws all of its ancestry from cluster $k$.
+Imagine a sample $i$ that draws all of its ancestry from layer $k$.
 The allele frequency in sample $i$ at locus $\ell$, $F_{i,\ell}$ can be
@@
 due to drift in the ancestral population of the $k$th layer,
-which is shared by all samples within the cluster. 
+which is shared by all samples within the layer. 
 The third term is the deviation of the $i$th sample away from the $k$th layer mean 
@@
 where the quantity $\delta_{i=j}$ is an indicator variable that equals 1 when $i$ is equal to $j$ and 0 otherwise, 
-as in Eqn. \eqref{cross_cluster_covariance}.
+as in Eqn. \eqref{cross_layer_covariance}.
 
@@
 we model allele frequencies within samples as linear combinations of the population-specific 
-drift terms across the clusters from which they draw ancestry.
+drift terms across the layers from which they draw ancestry.
 To do so, we introduce a term $w^{(k)}_{i}$ that describes the
@@
 (above that induced by a shared ancestral frequency at each locus)
-is due to the drift in the ancestral population of their cluster 
+is due to the drift in the ancestral population of their layer 
 (the variance of which is $\phi^{(k)}$). 
@@
 In this paper, we discuss two types of models, spatial and nonspatial, 
-each of which can be implemented with different numbers of clusters.
+each of which can be implemented with different numbers of layers/clusters.
 The spatial model is parameterized as in Eqn \eqref{admixed_spatial_cov},
@@
 The nonspatial model therefore has $3K$ fewer parameters than the spatial model,
-as there are three $\alpha$ parameters that describe the continuous differentiation effect of distance per cluster,
+as there are three $\alpha$ parameters that describe the continuous differentiation effect of distance per layer,
 and they are fixed in the nonspatial model.
 
-\paragraph{Single cluster}
-Each of these models can be run with a single cluster ($K=1$), 
-in which case the cluster-specific covariance parameter $\phi^{(k)}$ 
+\paragraph{Single layer}
+Each of these models can be run with a single layer ($K=1$), 
+in which case the layer-specific covariance parameter $\phi^{(k)}$ 
 and the global covariance parameter $\gamma$ become redundant.
-The single-cluster model is therefore a special case of the multi-cluster model, 
+The single-layer model is therefore a special case of the multi-layer model, 
 in which we set $\phi$ to zero, as described below:
 
-For the spatial model, the single-cluster parametric covariance is:
+For the spatial model, the single-layer parametric covariance is:
 \begin{equation}
@@
 	$\boldsymbol{\alpha^{(k)}_0}$ & 
-		controls the sill of the covariance matrix in cluster $k$& 
+		controls the sill of the covariance matrix in layer $k$& 
 		$\alpha^{(k)}_0 \sim \mathcal{N}(\mu = 0, \sigma = 1)$\\ \hline
 	$\boldsymbol{\alpha^{(k)}_D}$ & 
-		controls the rate of the decay of covariance with distance in cluster $k$& 
+		controls the rate of the decay of covariance with distance in layer $k$& 
 		$\alpha^{(k)}_D \sim \mathcal{N}(\mu = 0, \sigma = 1)$\\ \hline
 	$\boldsymbol{\alpha^{(k)}_2}$ & 
-		controls the shape of the decay of covariance with distance in cluster $k$ & 
+		controls the shape of the decay of covariance with distance in layer $k$ & 
 		$\alpha^{(k)}_2 \sim U(0,2)$\\ \hline
@@
 	$\boldsymbol{\phi^{(k)}}$ & 
-		cluster-specific shared drift in cluster $k$ &
+		layer-specific shared drift in layer $k$ &
 		 $\phi^{(k)} \sim \mathcal{N}(\mu = 0, \sigma = 1)$\\ \hline
 	$\boldsymbol{w_i}$ &
-		admixture proportions sample $i$ draws across $K$ clusters &
+		admixture proportions sample $i$ draws across $K$ layers &
 		$w_i \sim \text{Dir}(\alpha_{1} ... \alpha_{K}=0.1)$  \\ \hline
@@
 				\item $m$: the spatial and the nonspatial model
-				\item $k$: the number of clusters 1 through $K$
+				\item $k$: the number of layers/clusters 1 through $K$
 			\end{enumerate}	
@@
 without subsequent migration between them.
-Then, to generate continuous differentiation within each cluster,
+Then, to generate continuous differentiation within each population,
 at time $\tau_{\text{e}}$ in the past,
-each of these discrete clusters instantaneously colonizes an
+each of these discrete populations instantaneously colonizes an
 independent lattice of demes,
@@
 		\caption{Schematic of how we simulate datasets with continuous and discrete differentiation, using $K=2$ as an example.  
-			    Going forward in time, the $K$ clusters split from a common ancestor at time $\tau_{\text{s}}$,
+			    Going forward in time, the $K$ populations split from a common ancestor at time $\tau_{\text{s}}$,
 			    then expand to each  colonize a lattice of demes with nearest-neighbor symmetric migration at time $\tau_{\text{e}}$,
 			    then finally at time $\tau_{\text{a}}$ collapse into a single lattice consisting of demes 
-			    with ancestry entirely in one or the other of the clusters,
+			    with ancestry entirely in one or the other of the populations,
 			    or admixed between them.
@@
 	Map of admixture proportions estimated using a nonspatial model for $K=2$ through 7.
-	The data were simulated using a single cluster with nearest-neighbor symmetric migration between demes.
+	The data were simulated using one layer with nearest-neighbor symmetric migration between demes.
     }\label{simK1_nsp_pies}
@@
 	Map of admixture proportions estimated using a spatial model for $K=2$ through 7.
-	The data were simulated using 1 cluster with nearest-neighbor symmetric migration between demes.
+	The data were simulated using one layer with nearest-neighbor symmetric migration between demes.
     }\label{simK1_sp_pies}
@@
 	Map of admixture proportions estimated using a nonspatial model for $K=2$ through 7.
-	The data were simulated using two clusters with nearest-neighbor symmetric migration between demes.
+	The data were simulated using two layers with nearest-neighbor symmetric migration between demes.
     }\label{simK2_nsp_pies}
@@
 	Map of admixture proportions estimated using a spatial model for $K=2$ through 7.
-	The data were simulated using two clusters with nearest-neighbor symmetric migration between demes.
+	The data were simulated using two layers with nearest-neighbor symmetric migration between demes.
     }\label{simK2_sp_pies}
@@
 	Map of admixture proportions estimated using a nonspatial model for $K=2$ through 7.
-	The data were simulated using three clusters with nearest-neighbor symmetric migration between demes.
+	The data were simulated using three layers with nearest-neighbor symmetric migration between demes.
     }\label{simK3_nsp_pies}
@@
 	Map of admixture proportions estimated using a spatial model for $K=2$ through 7.
-	The data were simulated using three clusters with nearest-neighbor symmetric migration between demes.
+	The data were simulated using three layers with nearest-neighbor symmetric migration between demes.
     }\label{simK3_sp_pies}
@@
 	Map of admixture proportions estimated using fastSTRUCTURE \citep{fastStructure} for $K=2$ through 4.
-	The data were simulated using one cluster with nearest-neighbor symmetric migration between demes.
+	The data were simulated using one layer with nearest-neighbor symmetric migration between demes.
     }\label{fastStr_simK1}
@@
 	Map of admixture proportions estimated using fastSTRUCTURE \citep{fastStructure} for $K=2$ through 4.
-	The data were simulated using two clusters with nearest-neighbor symmetric migration between demes.
+	The data were simulated using two layers with nearest-neighbor symmetric migration between demes.
     }\label{fastStr_simK2}
@@
 	Map of admixture proportions estimated using fastSTRUCTURE \citep{fastStructure} for $K=2$ through 4.
-	The data were simulated using three clusters with nearest-neighbor symmetric migration between demes.
+	The data were simulated using three layers with nearest-neighbor symmetric migration between demes.
     }\label{fastStr_simK3}
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/populus/populus_sp_clst_covs.pdf}}
-	\caption{
-	Plots showing the cluster-specific parametric covariance IBD-curves 
+		{\includegraphics[width=6in,height=4in]{figs/populus/populus_sp_layer_covs.pdf}}
+	\caption{
+	Plots showing the layer-specific parametric covariance IBD-curves 
 	estimated for the \textit{Populus} data using 
 	the spatial \texttt{conStruct} model run with $K=1$ through 6.
-	Line colors are consistent with cluster colors in Fig \ref{populus_sp_pies}.
+	Line colors are consistent with layer colors in Fig \ref{populus_sp_pies}.
 	Points are colored by the species they are a covariance between:
@@
 	green on green points are sample covariances between \tri{} and \tri{}.
-    }\label{populus_sp_clst_covs}
-\end{figure}
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=6in,height=4in]{figs/populus/populus_nsp_clst_covs.pdf}}
-	\caption{
-	Plots showing the cluster-specific parametric covariances 
+    }\label{populus_sp_layer_covs}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=4in]{figs/populus/populus_nsp_layer_covs.pdf}}
+	\caption{
+	Plots showing the layer-specific parametric covariances 
 	estimated for the \textit{Populus} data using 
 	the nonspatial \texttt{conStruct} model run with $K=1$ through 6.
-	Line colors are consistent with cluster colors in Fig \ref{populus_nsp_pies}.
+	Line colors are consistent with layer colors in Fig \ref{populus_nsp_pies}.
 	Points are colored by the species they are a covariance between:
@@
 	green on green points are sample covariances between \tri{} and \tri{}.
-    }\label{populus_nsp_clst_covs}
+    }\label{populus_nsp_layer_covs}
 \end{figure}
@@
 	For each value of $K$ along the x-axis, there are an equal number of layer contributions plotted.
-	Colors are consistent with Figs \ref{populus_sp_pies}, \ref{populus_sp_clst_covs}, \ref{populus_nsp_pies}, and \ref{populus_nsp_clst_covs}.
+	Colors are consistent with Figs \ref{populus_sp_pies}, \ref{populus_sp_layer_covs}, \ref{populus_nsp_pies}, and \ref{populus_nsp_layer_covs}.
     }\label{populus_laycon}
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/bears/bear_sp_clst_covs.pdf}}
-	\caption{
-	Plots showing the cluster-specific parametric covariance IBD-curves 
+		{\includegraphics[width=6in,height=4in]{figs/bears/bear_sp_layer_covs.pdf}}
+	\caption{
+	Plots showing the layer-specific parametric covariance IBD-curves 
 	estimated for the black bear data using 
 	the spatial \texttt{conStruct} model run with $K=1$ through 6.
-	 }\label{bear_sp_clst_covs}
-\end{figure}
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=6in,height=4in]{figs/bears/bear_nsp_clst_covs.pdf}}
-	\caption{
-	Plots showing the cluster-specific parametric covariances 
+	 }\label{bear_sp_layer_covs}
+\end{figure}
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=6in,height=4in]{figs/bears/bear_nsp_layer_covs.pdf}}
+	\caption{
+	Plots showing the layer-specific parametric covariances 
 	estimated for the black bear data using 
 	the nonspatial \texttt{conStruct} model run with $K=1$ through 6.
-    }\label{bear_nsp_clst_covs}
+    }\label{bear_nsp_layer_covs}
 \end{figure}

### 2017-09-13 writeup/figs/bear_figs.R

@@ -18,7 +18,7 @@ pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bear_std_xval.pdf",width=10,hei
 	plot.xval.CIs(xval.CIs,K,ylim=c(-200,0),jitter=0.1,xlim=c(3.75,7.25),xaxt='n')
 		axis(1,at=c(4:7))
 		legend(x="bottomright",pch=c(19,NA),lty=c(NA,1),lwd=c(NA,2),col=c(1,adjustcolor(1,0.8)),legend=c("mean","95% CI"))
-	mtext("number of clusters",side=1,adj=-1.1,padj=4,font=2)
+	mtext("number of layers",side=1,adj=-1.1,padj=4,font=2)
 	mtext("Cross-validation results (Bears)",side=3,adj=10.5,padj=-2.5,font=2,cex=1.2)
 dev.off()
 
@@ -41,8 +41,9 @@ data.block <- conStruct:::make.data.block(K = 1,
 output.list.sp <- vector("list",7)
 for(k in 1:7){
 	load(sprintf("~/Dropbox/conStruct/data/bears/runs/bearsK%s_sp_conStruct.results.Robj",k))
+	conStruct.results <- cluster.2.layer(conStruct.results)
 	for(j in 1:k){
-		names(conStruct.results[[1]]$MAP$cluster.params[[j]])[4] <- "phi"
+		names(conStruct.results[[1]]$MAP$layer.params[[j]])[4] <- "phi"
 	}
 	output.list.sp[[k]] <- conStruct.results
 }
@@ -50,19 +51,21 @@ for(k in 1:7){
 output.list.nsp <- vector("list",7)
 for(k in 1:7){
 	load(sprintf("~/Dropbox/conStruct/data/bears/runs/bearsK%s_nsp_conStruct.results.Robj",k))
+	conStruct.results <- cluster.2.layer(conStruct.results)
 	for(j in 1:k){
-		names(conStruct.results[[1]]$MAP$cluster.params[[j]])[4] <- "phi"
+		names(conStruct.results[[1]]$MAP$layer.params[[j]])[4] <- "phi"
 	}
 	output.list.nsp[[k]] <- conStruct.results
 }
 
+if(FALSE){
 csr1.order <- NULL
 for(k in 2:7){
 	data.block$K <- k
 	csr <- output.list.sp[[k]][[1]]
 	if(k > 2){
 		tmp.csr <- output.list.sp[[k-1]][[1]]
[…]

### 2017-09-13 writeup/figs/fastStr_figs.R

@@ -7,16 +7,16 @@ collapse.ind.Q <- function(Q,pop.vec){
 	return(admix.props)
 }
 
-pie.plot <- function(Q.file,coords,pop.vec,cluster.colors){
+pie.plot <- function(Q.file,coords,pop.vec,layer.colors){
 	Q <- read.table(Q.file,stringsAsFactors=FALSE)
 	admix.props <- collapse.ind.Q(Q,pop.vec)
 	K <- ncol(admix.props)
 	N <- length(unique(pop.vec))
-	cluster.names <- paste0("cluster_",1:K)
+	layer.names <- paste0("layer_",1:K)
 	sample.names <- paste0("sample_", 1:N)
-	color.tab <- caroline::nv(c(cluster.colors[1:K]), cluster.names)
+	color.tab <- caroline::nv(c(layer.colors[1:K]), layer.names)
 	pie.list <- lapply(1:N, function(i) {
-	    caroline::nv(admix.props[i, ], cluster.names)
+	    caroline::nv(admix.props[i, ], layer.names)
 	})
 	names(pie.list) <- sample.names
 	caroline::pies(pie.list, x0 = coords[, 1], 
@@ -35,18 +35,18 @@ pdf(file="~/Dropbox/conStruct/writeup/figs/fastStr/fastStr_simK1_pies.pdf",width
 		pie.plot(Q.file = "~/Dropbox/conStruct/sims/structure/datasets/simK1/simK1_estK2.2.meanQ",
 				 coords = sim.dataset$coords,
 				 pop.vec = unlist(lapply(1:sim.dataset$N,function(n){rep(n,10)})),
-				 cluster.colors = c(4,2))
+				 layer.colors = c(4,2))
 			mtext(side=3,font=2,text="K=2")
 		pie.plot(Q.file = "~/Dropbox/conStruct/sims/structure/datasets/simK1/simK1_estK3.3.meanQ",
 				 coords = sim.dataset$coords,
 				 pop.vec = unlist(lapply(1:sim.dataset$N,function(n){rep(n,10)})),
-				 cluster.colors = c(2,"goldenrod1",4))
+				 layer.colors = c(2,"goldenrod1",4))
 			mtext(side=3,font=2,text="K=3")
 			mtext(side=1,font=2,text="true K = 1",cex=1.2,padj=2.7)
[…]

### 2017-09-13 writeup/figs/fig_funcs.R

@@ -1,5 +1,5 @@
 require(conStruct)
-cluster.colors <- c("blue","red","goldenrod1","forestgreen","darkorchid1","deepskyblue","darkorange1","seagreen2","yellow1","black")
+layer.colors <- c("blue","red","goldenrod1","forestgreen","darkorchid1","deepskyblue","darkorange1","seagreen2","yellow1","black")
 
 plot.xval.CIs <- function(xval.CIs,K,k.range=c(1:K),ylim=NULL,cex=1.5,jitter=0,...){
 	#recover()
@@ -35,49 +35,49 @@ get.admix.CIs <- function(conStruct.results){
 	return(CIs)
 }
 
-viz.admix.results <- function(sim.admix.props,conStruct.results,clst.order=NULL){
+viz.admix.results <- function(sim.admix.props,conStruct.results,layer.order=NULL){
 	#layout(matrix(c(1,1,2),1,3))
 	k.cols <- c("blue","red","goldenrod1","forestgreen","darkorchid1","deepskyblue","darkorange1","seagreen2","yellow1","black")
 	K <- ncol(sim.admix.props)
-	if(is.null(clst.order)){
-		clst.order <- 1:K
+	if(is.null(layer.order)){
+		layer.order <- 1:K
 	}
 	N <- nrow(sim.admix.props)
 	#plot(0,ylim=c(0,1),xlim=c(1,N) + c(-0.5,0.5),type='n',ylab="admixture proportion",xlab="samples")
 	admix.prop.CIs <- get.admix.CIs(conStruct.results)
 	# for(k in 1:K){
 		# segments(1:N,admix.prop.CIs[1,,k],1:N,admix.prop.CIs[2,,k],col=adjustcolor(k.cols[k],0.7),lwd=2.5)
-		# points(sim.admix.props[,clst.order[k]],col="gray",bg=k.cols[k],pch=23,cex=1.2)
+		# points(sim.admix.props[,layer.order[k]],col="gray",bg=k.cols[k],pch=23,cex=1.2)
 	# }
-	plot(sim.admix.props,conStruct.results$MAP$admix.proportions[,clst.order],type='n',
+	plot(sim.admix.props,conStruct.results$MAP$admix.proportions[,layer.order],type='n',
 			xlab="true admixture proportions",
 			ylab="estimated admixture proportions",
 			main=sprintf("Fitting admixture parameters (K = %s)",K))
 		abline(0,1,col=1,lty=2)
 		for(k in 1:K){
[…]

### 2017-09-13 writeup/figs/populus_figs.R

@@ -18,7 +18,7 @@ pdf(file="~/Dropbox/conStruct/writeup/figs/populus/populus_std_xval.pdf",width=1
 	mtext("Predictive accuracy",side=2,padj=-5,font=2)
 	plot.xval.CIs(xval.CIs,K,ylim=c(-40,0),jitter=0.1,xlim=c(1.75,7.25))
 		legend(x="bottomright",pch=c(19,NA),lty=c(NA,1),lwd=c(NA,2),col=c(1,adjustcolor(1,0.8)),legend=c("mean","95% CI"))
-	mtext("number of clusters",side=1,adj=-1.1,padj=4,font=2)
+	mtext("number of layers",side=1,adj=-1.1,padj=4,font=2)
 	mtext("Cross-validation results (Populus)",side=3,adj=5,padj=-2.5,font=2,cex=1.2)
 dev.off()
 
@@ -46,8 +46,9 @@ data.block <- conStruct:::make.data.block(K = 1,
 output.list.sp <- vector("list",7)
 for(k in 1:7){
 	load(sprintf("~/Dropbox/conStruct/data/poplars/runs/poplarsK%s_sp_conStruct.results.Robj",k))
+	conStruct.results <- cluster.2.layer(conStruct.results)
 	for(j in 1:k){
-		names(conStruct.results[[1]]$MAP$cluster.params[[j]])[4] <- "phi"
+		names(conStruct.results[[1]]$MAP$layer.params[[j]])[4] <- "phi"
 	}
 	output.list.sp[[k]] <- conStruct.results
 }
@@ -55,19 +56,21 @@ for(k in 1:7){
 output.list.nsp <- vector("list",7)
 for(k in 1:7){
 	load(sprintf("~/Dropbox/conStruct/data/poplars/runs/poplarsK%s_nsp_conStruct.results.Robj",k))
+	conStruct.results <- cluster.2.layer(conStruct.results)
 	for(j in 1:k){
-		names(conStruct.results[[1]]$MAP$cluster.params[[j]])[4] <- "phi"
+		names(conStruct.results[[1]]$MAP$layer.params[[j]])[4] <- "phi"
 	}
 	output.list.nsp[[k]] <- conStruct.results
 }
 
+if(FALSE){
 csr1.order <- NULL
 for(k in 2:7){
 	data.block$K <- k
 	csr <- output.list.sp[[k]][[1]]
 	if(k > 2){
[…]

### 2017-09-13 writeup/figs/sim_figs.R

@@ -25,8 +25,9 @@ data.block <- conStruct:::make.data.block(K = 1,
 output.list.sp <- vector("list",7)
 for(k in 1:7){
 	load(sprintf("~/Dropbox/conStruct/sims/cross_validation/K_1/runs/simK1_K%s_sp_conStruct.results.Robj",k))
+	conStruct.results <- cluster.2.layer(conStruct.results)
 	for(j in 1:k){
-		names(conStruct.results[[1]]$MAP$cluster.params[[j]])[4] <- "phi"
+		names(conStruct.results[[1]]$MAP$layer.params[[j]])[4] <- "phi"
 	}
 	output.list.sp[[k]] <- conStruct.results
 }
@@ -34,13 +35,14 @@ for(k in 1:7){
 output.list.nsp <- vector("list",7)
 for(k in 1:7){
 	load(sprintf("~/Dropbox/conStruct/sims/cross_validation/K_1/runs/simK1_K%s_nsp_conStruct.results.Robj",k))
+	conStruct.results <- cluster.2.layer(conStruct.results)
 	for(j in 1:k){
-		names(conStruct.results[[1]]$MAP$cluster.params[[j]])[4] <- "phi"
+		names(conStruct.results[[1]]$MAP$layer.params[[j]])[4] <- "phi"
 	}
 	output.list.nsp[[k]] <- conStruct.results
 }
 
-
+if(FALSE){
 plot.sim.pies(data.block = data.block,
 			  K = 7,
 			  output.list = output.list.nsp,
@@ -50,6 +52,7 @@ plot.sim.pies(data.block = data.block,
 			  K = 7,
 			  output.list = output.list.sp,
 			  file.name = "~/Dropbox/conStruct/writeup/figs/sims/simK1_sp_pies_K")
+}
 
 pdf(file="~/Dropbox/conStruct/writeup/figs/sims/simK1_nsp_pies.pdf",width=8,height=5.34,pointsize=14)
 	par(mfrow=c(2,3))
@@ -104,7 +107,7 @@ for(k in 1:K){
 	}
 	if(k > 2){
 		tmp.csr <- output.list.sp[[k-1]][[1]]
-		csr1.order <- match.clusters.x.runs(tmp.csr$MAP$admix.proportions,csr$MAP$admix.proportions,csr1.order)
[…]

### 2017-09-13 writeup/conStruct.tex
added fastStr time comparison

@@
 the results of which were broadly concordant with those of the fastSTRUCTURE runs (Fig \ref{populus_fastStr}).
+\gb{come back to this}.
 Most of this signal seems to derive from the \textit{P. trichocarpa} samples:
@@
 which was more than sufficient for convergence,
-took XXX hours.
+took 2.8 hours.
 On the same machine and dataset, 
-a fastSTRUCTURE run with $K=4$ took 3.76 days.
+a fastSTRUCTURE run with $K=4$ took 90.2 hours.
 However, for a large number of samples, 

### 2017-09-14 writeup/conStruct.tex
.

@@
 Especially when sampling is discontinuous, 
-clustering or assignment methods may incorrectly ascribe differentiation 
-due to continuous processes (e.g., isolation by distance) 
+the use of clustering or assignment methods may incorrectly ascribe
+differentiation due to continuous processes (e.g., geographic isolation by distance)
 to discrete processes, such as geographic, ecological, or reproductive barriers 
 between populations.
-This is partly a result of the difficulty of sampling uniformly and continuously 
-across the range of a population or species, 
-but more, it reflects a shortcoming of current methods for inferring and 
-visualizing population structure from genetic data in the face of data 
-that are characterized by both continuous and discrete population structure.
+% This is partly a result of the difficulty of sampling uniformly and continuously 
+% across the range of a population or species, but more, it 
+This reflects a shortcoming of current methods for inferring and 
+visualizing population structure when applied to genetic data
+deriving from geographically distributed populations.
 Here, we present a novel statistical framework for the simultaneous inference 
@@
 The method estimates ancestry proportions for each 
-sample from a set of discrete population layers, 
+sample from a set of two-dimensional population layers, 
 and, within each layer, estimates a rate at which relatedness decays with distance.
 This model explicitly addresses the ``clines vs. clusters" problem in 
-modeling population genetic variation by jointly accommodating both 
-continuous and discrete patterns of differentiation. 
-We demonstrate the utility of this approach using both simulations and empirical applications.
+modeling population genetic variation.
+% by jointly accommodating both continuous and discrete patterns of differentiation. 
+The method produces useful descriptions of structure in genetic relatedness
+in situations where separated, geographically distributed populations interact,
+as after a range expansion or secondary contact.
+We demonstrate the utility of this approach using simulations 
+and by applying it to empirical datasets of poplars and black bears in North America.
 \end{abstract}
+
+\plr{I felt like we said "continuous and discrete" too many times in the abstract.}
 
@@
 in the continuous divergence of populations when they unambiguously become discrete species.
+Descriptions of taxa using phylogenetic trees omit within-taxon variation,
+especially when these are intermingling subgroups.
 The issue of identifying meaningful biological subunits 
@@
 long-distance migration. 
+Both discrete and continuous components
+are required to accurately describe most species' patterns of genetic relatedness.
 
@@
 Furthermore, we often need to identify subsets of indivduals resulting from reasonably coherent
-evolutionary histories for downstream analyses to learn about population history and adaptation.
+evolutionary histories for downstream analyses to learn about population history and adaptation,
+\plr{as for example XXX?}
+Conversely,
+the substantial information available from continuous, geographic differentiation
+(e.g., adaptation along a climatic gradient)
+can be confounded by discrete historical processes (e.g., admixture),
+requiring methods that can disentangle the two.
 
@@
 but many can give misleading results when applied to data 
-that show a continuous pattern of isolation by distance 
+that show a continuous pattern of differentiation,
+as that produced by geographic isolation by distance 
 \citep{Wright1943, novembre_interpreting_2008, Frantz2009}.
@@
 
-A pattern of isolation by distance (IBD) arises when migration in a
-population is spatially limited, and slow enough between distant
-locations that genetic drift can build up differentiation between them
+\emph{Isolation by distance} refers to a pattern of increasing genetic differentiation
+with geographic separation,
+which occurs when geographically restricted dispersal allows
+genetic drift to build up differentiation between distant locations
 \citep{Wright1943}. 
-IBD has been modeled in both spatially continuous frameworks 
-\citep{malecot1969mathematics,Slatkin1985}, 
-in which an individual's parents are found within some radius of that individual's location,
-and discrete, ``stepping-stone'' frameworks \citep{kimura1964stepping},
-in which migration occurs only between neighboring demes.
+Theoretical work,
+mostly derived from ``stepping-stone'' models 
+\citep{kimura1964stepping,sawyer1976stepping,shiga1984stepping},
+gives us some analytical predictions for isolation by distance
+\citep{malecot1969mathematics,Slatkin1985,epperson2003geographical}, 
+but substantial work remains to be done \citep{barton2002neutral,barton2013modelling}.
+% IBD has been modeled in both spatially continuous frameworks 
+% in which an individual's parents are found within some radius of that individual's location,
+% and discrete, ``stepping-stone'' frameworks \citep{kimura1964stepping},
+% in which migration occurs only between neighboring demes.
 Given the generality of the circumstances that generate a pattern of isolation by distance, 
@@
 in the data deviates significantly from that expected given geographic separation.
-
-In this paper, we present a statistical method for simultaneously 
-modeling continuous and discrete patterns of population structure.
-We model genetic variation in genotyped individuals as 
+In this paper, 
+% we present a statistical method for simultaneously 
+% modeling continuous and discrete patterns of population structure.
+we model genetic variation in genotyped individuals as 
 partitioned within or admixed across a specified number of discrete layers,
@@
 We then calculate the \emph{allelic covariance} between samples $i$ and $j$, denoted $\widehat{\Omega}_{i,j}$,
-as the expected covariance of distinct individual alleles chosen from each of the two samples at a random locus,
-choosing a random allele to code as `1' (and the other as `0').
+as the expected covariance of distinct individual alleles 
+chosen from each of the two samples at a random locus,
+coding the counted allele as `1' (and the other as `0').
 Concretely, this is
@@
 that extends over the entire sampled range of the populations. 
-Each sample is composed of a mixture of contributions from each of
-these layers. 
-See Figure \ref{schematic} for a cartoon representation of our model. 
+As depicted in Figure \ref{schematic},
+each sample is composed of a mixture of contributions from each of these layers,
+with the relative contributions of each layer described by
+a set of ``admixture proportions'' (the $w_i^{(k)}$).
 These layers thus take the place of ``clusters'' in clustering methods,
@@
 as ``spatial cluster'' suggests a clustering in space, 
-while our layers may be present at all sampled locations.
+while our layers may contribute to genetic variation 
+across the entire geographic range.
 
@@
 if the proportions by which individuals' ancestry derive from each layer 
-(the ``admixture'' coefficients) 
+(the ``admixture'' proportions) 
 do so as well.
@@
 The quantity $D_{i,j}$ is the observed geographic distance between samples $i$ and $j$,
-the $\alpha^{(k)}$ parameters control the shape of the decay of
-covariance with distance in the layer.
+and the $\alpha^{(k)}$ parameters control the shape of the decay of
+covariance with distance in the cluster.
 Our choice of a powered-exponential decay, 
@@
 and is not chosen to match a particular population genetics model. 
-The $\phi^{(k)}$ is a parameter that describes the background covariance within the layer. 
-If two samples draw 100\% of their ancestry from layer $k$, then their covariance under the model is $G^{(k)}_{i,j}$;
+% (Analytical theory \citep{barton2002neutral} predicts the decay should be proportional to $\exp{-D}/\sqrt{D}$.)
+The $\phi^{(k)}$ is a parameter that describes the background covariance within the cluster. 
+If two samples draw 100\% of their ancestry from cluster $k$, then their covariance under the model is $G^{(k)}_{i,j}$;
 if they are furthermore geographically very close ($D_{i,j}=0$)
@@
 If the geographic distance between them is very large, 
-their covariance will be equal to to the background level $\phi^{(k)}$ within the layer.
-The ``shared drift'' parameter $\phi^{(k)}$ can be interpreted as the branch length 
-connecting the $k$th population to the population ancestral to all modeled
-layers \citep[see for example ][]{patterson_ancient_2012,
-  peter_fstats}. In the methods section \ref{rationale} we lay out a
+their covariance will be equal to to the background level $\phi^{(k)}$ within the cluster.
+%% plr: this is no longer true
+% The ``shared drift'' parameter $\phi^{(k)}$ can be interpreted as the branch length 
+% connecting the $k$th population to the population ancestral to all modeled
+% layers \citep[see for example ][]{patterson_ancient_2012,peter_fstats}. 
+In the methods section \ref{rationale} we lay out a
 simple model of allele frequencies underlying this covariance model.
@@
 gives the genome-wide proportion of alleles from sample $i$ that derive from
-layer $k$ (and so $\sum_K w^{(k)}_i =1$).
+layer 
+$k$ (and so $\sum_{k=1}^K w^{(k)}_i =1$).
 A visual representation of the method mechanics is shown in Fig \ref{schematic}.
@@
 			     The pattern of isolation by distance within each layer is denoted by their color gradients, 
-			     and $\phi_{(k)}$ denotes the covariance shared by samples with ancestry entirely in the $k$th layer.
+			     and $\phi^{(k)}$ denotes the covariance shared by samples with ancestry entirely in the $k$th layer.
 			     Sampled populations on the landscape are inferred to be admixed between these layers; 
@@
 \begin{equation}
-\Omega_{i,j} = \gamma + \sum\limits_K w^{(k)}_i w^{(k)}_j
+    \Omega_{i,j} = \gamma + \sum_{k=1}^K w^{(k)}_i w^{(k)}_j
 G^{(k)}_{i,j} + \delta_{i,j}\eta_i .
@@
 \paragraph{Likelihood and inference}
-Assuming that the allele frequencies at each locus
-are independent between loci and multivariate normally distributed across populations, 
+If the allele frequencies at each locus
+were independent between loci and multivariate normally distributed across populations, 
 their allelic covariance $\widehat{\Omega}$ 
@@
 \begin{equation}
-P(\widehat{\Omega} \; | \; \Omega) = \mathcal{W} \left( L\widehat{\Omega} \; | \; \Omega,L\right) .
+    P(\widehat{\Omega} \; | \; \Omega) 
+    = 
+    \mathcal{W} \left( L\widehat{\Omega} \; | \; \Omega,L\right) ,
 \end{equation}
-Linkage disequilibrium (LD) between loci will decrease the effective
+where $\mathcal{W}$ is the Wishart likelihood function.
+Statistical nonindependence between loci (linkage disequilibrium)
+will decrease the effective
 number of degrees of freedom.  
-One possible solution, which we have not yet implemented, 
-is to estimate an \emph{effective} number of loci 
+One possible solution, which we have not yet found necessary to implement,
+would be to estimate an \emph{effective} number of loci 
 by introducing a parameter to modify the given degrees of freedom 
@@
 \begin{equation}
-P\left( w,\alpha,\phi,\eta,\gamma \;	| \; \widehat{\Omega}, L \right) \propto
+P\left( w,\alpha,\phi,\eta,\gamma \;|\; \widehat{\Omega} \right) 
+    \propto
     P\left(\widehat{\Omega} \; | \; \Omega(w,\alpha,\phi,\eta,\gamma,D) \right)
@@
 contains a nonspatial assignment model as a special case 
-(see \secref{model_app} for a more in-depth discussion). 
+(see \secref{model_app} in the appendix for a more in-depth discussion). 
 By setting $\alpha_0^{(k)}$ to zero for all $k$, 
@@
 and individuals draw a proportion of their ancestry from each cluster. 
-This model is strongly conceptually similar to that of STRUCTURE \citep{STRUCTURE} 
+This model is very similar to that of STRUCTURE \citep{STRUCTURE} 
 and related models \citep[e.g.][]{ADMIXTURE}; 
@@
 while the standard assignment methods assume that the error is binomially distributed \citep{Engelhardt2012}.  
+(We make this approximation for the substantial advantages in computational speed.)
 Therefore, we can compare the fit of the different models 
@@
 The best model is judged to be the simplest one with significantly better predictive accuracy
-than others (see \secref{Xvalidation} for more on our cross-validation procedure).
+than others (see \secref{Xvalidation} of the appendix for more on our cross-validation procedure).
 In general, larger values of $K$ allows the model more flexibility,
@@
 
-An additional heuristic to evaluate the inclusion of additional layers is to calculate 
-each layer's relative contribution to total covariance, 
+Another way to describe how additional layers contribute to model fit
+is to calculate each layer's relative contribution to total covariance, 
 and to choose a value of $K$ where all layers have a contribution above some cutoff (e.g. 0.1\%).
-Because the prior on admixture proportions is a Dirichlet distribution 
-that is quite harsh against intermediate admixture values (see Table \ref{tab:param_prior_tab}),  
-the model will tend to ``not use" unnecessary layers, 
-and they will have a low contribution to overall covariance.
+The Dirichlet prior on admixture proportions
+is quite harsh against intermediate admixture values (see Table \ref{tab:param_prior_tab}),  
+encouraging the model to ``not use" unnecessary layers, 
+so that they will have a low contribution to overall covariance.
+We use the connection of allelic covariance to the probability of agreement (identity) between genotypes
+to calculate a measure of overall contribution of each layer to the model.
+XXX PLR XXX
+To calculate the overall contribution of the $k$th layer to the model,
+we use the connection between allelic covariance and
+
 The contribution of the $k$th layer can be calculated as
@@
 \begin{equation}
-\Omega_{i,j} = \text{Var}(\epsilon) + \sum\limits_K w^{(k)}_iw^{(k)}_j
+\Omega_{i,j} = \text{Var}(\epsilon) + \sum_{k=1}^K w^{(k)}_iw^{(k)}_j
 \left(
@@
 So, the contribution from layer $k$ to agreement between $i$ and $j$ is 
-$w_i^{(k)} w_j^{(k)} q^{(k)}_{ij} / (p_\gamma + \sum_K w_i^{(k)} w_j^{(k)} q^{(k)}_{ij})$,
+$w_i^{(k)} w_j^{(k)} q^{(k)}_{ij} / (p_\gamma + \sum_{k=1}^K w_i^{(k)} w_j^{(k)} q^{(k)}_{ij})$,
 which is the probability, given that they agree, that they agree thanks to layer $k$.
-Similarly, the ``background" contribution is $p_\gamma / (p_\gamma + \sum_K w_i^{(k)} w_j^{(k)} q^{(k)}_{ij})$.
+Similarly, the ``background" contribution is $p_\gamma / (p_\gamma + \sum_{k=1}^K w_i^{(k)} w_j^{(k)} q^{(k)}_{ij})$.
 Because our signal comes from \emph{variation} in covariance, we omit

### 2017-09-14 writeup/conStruct.tex
edited up to empirical apps

@@
 
-Another way to describe how additional layers contribute to model fit
+Another way to describe the practical significance of additional layers
 is to calculate each layer's relative contribution to total covariance, 
@@
 is quite harsh against intermediate admixture values (see Table \ref{tab:param_prior_tab}),  
-encouraging the model to ``not use" unnecessary layers, 
+encouraging the model to ``not use" unnecessary layers if they are present, 
 so that they will have a low contribution to overall covariance.
-We use the connection of allelic covariance to the probability of agreement (identity) between genotypes
-to calculate a measure of overall contribution of each layer to the model.
-XXX PLR XXX
-To calculate the overall contribution of the $k$th layer to the model,
-we use the connection between allelic covariance and
-
-The contribution of the $k$th layer can be calculated as
-its contribution to the parametric covariance between samples 
-summed over the upper triangle (excluding the diagonal) of the covariance matrix.
-The contribution of the $k$th layer is its absolute contribution relative to those of all other layers, 
-i.e., its contribution divided by the sum of the contributions of all $K$ layers
-(for details on calculating this contribution, see Appendix \secref{layer_contribution}). 
+To do this, we use the following alternative description of our covariance model:
+the genomes of any pair of individuals \emph{agree} with some background probability,
+but this probability of agreement is increased on any segment of genome 
+that both have inherited through the same layer.
+(The amount it increases depends on how far apart they are.)
+We use this characterization to quantify the relative contributions of each layer,
+by computing the average contribution to increased probability of agreement
+as described in Appendix \secref{layer_contribution}). 
 This layer contribution is similar to the ``ancestry contribution'' proposed by \cite{fastStructure}. 
@@
 \subsection*{Simulations}
-We first test the method with scenarios generated using the coalescent simulator \texttt{ms} \citep{Hudson2002}.
-To generate discrete population structure, we simulate $K$ distinct populations,
-each of which split from a common ancestor $\tau_{\text{s}}$ units of coalescent time in the past,
-without subsequent migration between them.
-To generate continuous differentiation within each layer,
-at time $\tau_{\text{e}}$ in the past,
-each of these discrete populations instantaneously colonizes an
-independent lattice of demes,
-for which we use a stepping stone model with symmetric migration 
-to nearest neighbors (eight neighbors, including diagonals).
+
+To test the method, we first generated data using the coalescent simulator \texttt{ms} \citep{Hudson2002}.
+In each simulation,
+we split a single ancestral population into $K$ subpopulations $\tau_{\text{s}}$ units of coalescent time in the past,
+and at time $\tau_{\text{e}}$ in the past,
+each of these discrete populations instantaneously colonized a
+separate $6 \times 6$ square lattice of demes.
+Migration on each lattice was to nearest neighbors (eight neighbors, including diagonals).
 Finally, at time $\tau_{\text{a}}$ in the past 
-we generate a single dataset 
-by collapsing those $K$ discrete layers into a single grid of demes 
-that have various amounts of admixture from these different layers
+we collapsed those $K$ discrete layers into a single grid of demes,
+choosing various amounts of admixture from these different layers
 (see Fig \ref{sim_setup}). 
-We collapse the layers together using random admixture proportions 
-that are spatially autocorrelated (see Appendix \secref{sim_details} for details), 
+We collapsed the layers together using random admixture, spatially autocorrelated proportions 
+(see Appendix \secref{sim_details} for details), 
 so that both the frequencies in layers and the layer memberships are
@@
 in each sampling 10,000 unlinked loci from each of 20 haploid individuals from every deme.
-
-We ran \texttt{conStruct} analyses on each of these simulated datasets, 
-specifying both a spatial and nonspatial model for $K = $ 1 through 7, 
-then used cross-validation to compare the predictive performance of the models.
-In Fig \ref{simK1_pies}, we show the results of the analyses of the dataset simulated under $K=1$, 
-contrasting the results of the nonspatial model runs with 2 - 4 layers 
-with those of spatial model with 2 - 4 layers.
-In Fig \ref{sim_xvals} we show the results of the cross-validation analysis for each of the three simulated datasets.
-The results of all analyses run on datasets simulated using $K=1$ through 3 are shown in 
-Figs \ref{simK1_nsp_pies}-\ref{simK3_adprop_fit}.
-To confirm that results from our nonspatial model were qualitatively similar to those of 
-other nonspatial model-based clustering algorithms, 
-we also analyzed each simulated dataset using fastSTRUCTURE \citep{fastStructure}, 
-inferring admixture under models with $K=2$ through 4 (Figs \ref{fastStr_simK1}-\ref{fastStr_simK3}).
-
-A number of results are immediately clear from these tests with simulated data.
-First, \texttt{conStruct} performs well on all simulated scenarios.
-Across all values of K and for all simulations, 
-the spatial model is correctly preferred over the nonspatial model
-(Fig \ref{sim_xvals}). Second, the predictive accuracy of the nonspatial model increases 
-with the number of layers in the model, 
-as it describes continuous, spatial patterns of relatedness between demes within the same cluster 
-using gradients of admixture across more and more discrete clusters.
-This behavior of the nonspatial model is highlighted in 
-the analyses of data simulated under $K=1$ (Fig \ref{simK1_pies} and \ref{simK1_nsp_pies}):
-as $K$ in the model increases, 
-the sampled area is increasingly divvied up between clusters 
-with ``centers of mass" in the $K$ geographical extremes of the simulated space.
-At $K=2$, the clusters are centered in opposite corners on the map; 
-at $K=3$, in the lower two corners and the top half; 
-at $K=4$, all four corners, and so on.
-These results are qualitatively similar to those generated using fastSTRUCTURE \citep{fastStructure}, 
-which also describes continuous differentiation using admixture between clusters
- (Fig \ref{fastStr_simK1}-\ref{fastStr_simK3}).
-While it is true that each non-spatial cluster is genetically more similar within itself
-than it is to other clusters,
-we know that these boundaries are arbitrary,
-since the data were simulated with no such boundaries.
-
-In contrast, the predictive accuracy of the spatial \texttt{conStruct} model
-either plateaus at or declines from the true value of $K$ used to simulate the data 
-(Figs \ref{sim_xvals}, \ref{simK1_xval}, \ref{simK2_xval}, and \ref{simK3_xval}).
-And, as $K$ in the spatial model increases, 
-deme membership in layers beyond the number used to simulate the data does not appreciably increase 
-(Fig \ref{simK1_sp_pies}, \ref{simK2_sp_pies}, \ref{simK3_sp_pies}).
+We then ran both spatial and nonspatial \texttt{conStruct} analyses on each simulated dataset with $K$ between 1 and 7,
+and compared predictive performance of the models using cross-validation to compare the predictive performance.
+% To confirm that results from our nonspatial model were qualitatively similar to those of 
+% other nonspatial model-based clustering algorithms, 
+For comparison,
+we also analyzed each simulated dataset using fastSTRUCTURE \citep{fastStructure}, using $K=2--4$.
+% inferring admixture under models with $K=2$ through 4 (Figs \ref{fastStr_simK1}--\ref{fastStr_simK3}).
+
+\plr{rearranged this to focus more on take-homes}
+
+% In Fig \ref{simK1_pies}, we show the results of the dataset simulated with $K=1$, 
+% contrasting the results of the nonspatial model runs with 2--4 layers 
+% with those of spatial model with 2--4 layers.
+% 
+% In Fig \ref{sim_xvals} we show the results of the cross-validation analysis for each of the three simulated datasets.
+% The results of all analyses run on datasets simulated using $K=1$ through 3 are shown in 
+% Figs \ref{simK1_nsp_pies}-\ref{simK3_adprop_fit}.
+% 
+% To confirm that results from our nonspatial model were qualitatively similar to those of 
+% other nonspatial model-based clustering algorithms, 
+% we also analyzed each simulated dataset using fastSTRUCTURE \citep{fastStructure}, 
+% inferring admixture under models with $K=2$ through 4 (Figs \ref{fastStr_simK1}--\ref{fastStr_simK3}).
+
+With these simulations,
+spatial \texttt{conStruct} does not create spurious clusters when there are none:
+Figures \ref{simK1_pies}, \ref{simK2_sp_pies}, and \ref{simK3_sp_pies} 
+show that subsequent layers beyond the number used for simulation are unused 
+when data is simulated with $K=1$ but analyzed with $K>1$ --
+the layers are present, but contribute very little to any population.
 Even when the spatial model is run with $K=7$, 
@@
 
-Finally, we calculated layer contribution for 
-analyses run in each simulation 
-(Fig \ref{simK1_laycon}, \ref{simK2_laycon}, and \ref{simK3_laycon}).
+In contrast, the nonspatial model describes geographic variation
+using gradients of admixture between more and more discrete clusters
+to better approximate the continuous, spatial patterns of relatedness
+(depicted in Figs \ref{simK1_pies} and \ref{simK1_nsp_pies}).
+% as $K$ in the model increases, 
+% the sampled area is increasingly divvied up between clusters 
+% with ``centers of mass" in the $K$ geographical extremes of the simulated space.
+% At $K=2$, the clusters are centered in opposite corners on the map; 
+% at $K=3$, in the lower two corners and the top half; 
+% at $K=4$, all four corners, and so on.
+The fastSTRUCTURE results are qualitatively similar, as shown in Figs \ref{fastStr_simK1}-\ref{fastStr_simK3}.
+Each non-spatial cluster is in truth genetically more similar within itself
+than it is to other clusters,
+but we know that these boundaries are arbitrary,
+since the data were simulated with no such boundaries.
+
+The spatial model's better fit is reflected by increased predictive accuracy:
+as shown in Fig \ref{sim_xvals},
+across all models and choices of $K$, the spatial model is correctly preferred over the nonspatial model.
+As desired, predictive accuracy of the spatial model increases until the true value of $K$,
+and then plateaus or declines
+(Figs \ref{sim_xvals}, \ref{simK1_xval}, \ref{simK2_xval}, and \ref{simK3_xval}),
+while the predictive accuracy of the nonspatial model increases as subsequent clusters are added.
+
+% In contrast, the predictive accuracy of the spatial \texttt{conStruct} model
+% either plateaus at or declines from the true value of $K$ used to simulate the data 
+% (Figs \ref{sim_xvals}, \ref{simK1_xval}, \ref{simK2_xval}, and \ref{simK3_xval}).
+
+The unimportance of spurious layers can be seen in plots of layer contributions,
+Figs \ref{simK1_laycon}, \ref{simK2_laycon}, and \ref{simK3_laycon}.
 In the spatial analyses, once we pass the true $K$, 
 subsequent layers add little in terms of (co)variance explained; 
-in contrast, additional clusters in the nonspatial analyses continue to
-be important for describing variation, 
-as they are forced to partition up the spatial patterns within clusters.
+in contrast, additional clusters in the nonspatial analyses continue to contribute substantially.
 

### 2017-09-14 writeup/conStruct.tex
edited populus

@@
 \author{
-Gideon S. Bradburd$^{1,a}$, 
-Graham M. Coop$^{2,b}$,
-Peter L. Ralph$^{3,c}$}
+Gideon S.\ Bradburd$^{1,a}$, 
+Graham M.\ Coop$^{2,b}$,
+Peter L.\ Ralph$^{3,c}$}
 
@@
 \subsection*{Empirical Applications}
+
 To demonstrate the utility of this method, we analyzed empirical population genomic data from two systems:
@@
 
-Here, we focus on two \textit{Populus} species in the Pacific Northwest: 
 \textit{Populus trichocarpa}, the black cottonwood,
-and \textit{Populus balsamifera}, the balsam poplar.
-These species have a broad zone of overlap and are hypothesized to hybridize \citep{geraldes_etal_2014, suarezgonzalez_etal_2016}.
+and \textit{Populus balsamifera}, the balsam poplar,
+have a broad zone of overlap in the Pacific Northwest,
+where they are hypothesized to hybridize \citep{geraldes_etal_2014, suarezgonzalez_etal_2016}.
 Both species are sampled over a large geographic region, 
 and show spatial patterns of genetic and phenotypic variation \citep{slavov_etal_2012, mckown_etal_2013},
-making the system well-suited to test an analysis designed to describe
-how genetic structure is partitioned both discretely across populations (or species)
-and continuously with geographic distance.
+making the system well-suited for application of our method.
+% to test an analysis designed to describe
+% how genetic structure is partitioned both discretely across populations (or species)
+% and continuously with geographic distance.
 We organize the results of our analyses around the following questions: 
 \begin{enumerate}
-\item Given the known history of hybridization between this pair of poplar species, 
-is there support for two discrete populations, 
-rather than a single cline of ancestry across their overlap zone?
-\item If there is evidence for discrete population structuring, 
-does it fall solely along species boundaries, 
-or is there sub-structuring between different populations of the same species?
-\item Is there support for a pattern of isolation by distance within
-the inferred layers, and, if so, 
-does it differ between different layers?
+
+    \item To what degree has hybridization blurred the boundaries between 
+        \textit{trichocarpa} and \textit{balsamifera}?
+        (As an extreme case, does genetic differentiation support these as separate species,
+        as opposed to a single cline of ancestry?)
+    % Given the known history of hybridization between this pair of poplar species, 
+    % is there support for two discrete populations, 
+    % rather than a single cline of ancestry across their overlap zone?
+
+    \item Does the only significant boundary of population structure
+        fall along the species boundary (if any),
+        % If there is evidence for discrete population structuring, 
+        % does it fall solely along species boundaries, 
+        or is there sub-structuring within species?
+
+    \item Does the strength of isolation by distance differ between inferred layers?
+        This may indicate, e.g., different speeds of postglacial expansion
+        or primary modes of dispersal.
+        % Is there support for a pattern of isolation by distance within
+        % the inferred layers, and, if so, 
+        % does it differ between different layers?
+
 \end{enumerate}
 
-Our data, from \cite{geraldes_etal_2014}, 
-consist of 434 individuals sampled from 35 drainages 
-and genotyped at just over 33,000 loci (map of the sampling shown in Fig \ref{populus_map}).
+We use data from \cite{geraldes_etal_2014}, 
+consisting of 434 individuals sampled from 35 drainages 
+genotyped at just over 33,000 loci (map of the sampling shown in Fig \ref{populus_map}).
+\plr{How do the indivs partition across the drainages?  Perhaps give range of number of indivs per drainage?}
 The data were generated using RAD-seq, 
@@
 	Map of admixture proportions estimated for the \textit{Populus} dataset 
-	using the spatial model for $K=2$ through 4.
+    using the spatial \texttt{conStruct} model for $K=2$ through 4.
     }\label{populus_pies}
@@
 to distinct layers, with some populations drawing ancestry from multiple layers.
+Based on cross-validation results, we view the $K=3$ spatial model 
+as a sufficient description of the data,
+with additional structure of uncertain significance.
 This provides strong support for discrete population structure between the two species,
@@
 at values of $K$ above 2, further discrete substructure was inferred within the \textit{P. trichocarpa} samples, 
-while \bals{} samples displayed no substructure up to $K=7$.
-This lack of substructure within the sampled \bals{} populations is largely in agreement with results from 
-\citet{keller_etal_2010}, who found that \bals{} in the region is characterized by relatively low diversity, 
-and probably represents a recent, postglacial expansion.
-
-For values of $K<4$, the spatial model was strongly preferred over the nonspatial model (Fig \ref{populus_xvals}),
-the results of which were broadly concordant with those of the fastSTRUCTURE runs (Fig \ref{populus_fastStr}).
+with no substructure within \bals{}.
+Most of this signal seems to derive from the \textit{P. trichocarpa} samples:
+% covariance within predominantly \tri{} layers decays more quickly with distance than within the majority \bals{} layer, 
+as seen in Fig \ref{populus_sp_layer_covs},
+there is almost no isolation by distance within the \bals{} layer ($\alpha_D \approx 0$).
+Both points are
+in agreement with \citet{keller_etal_2010}, 
+who found low diversity within the region's \bals{},
+probably as the result of a recent postglacial expansion.
+
+For all values of $K$, the spatial model was preferred over the nonspatial model (Fig \ref{populus_xvals}) (strongly for $K<4$).
+The nonspatial results were broadly concordant with those of the fastSTRUCTURE runs (Fig \ref{populus_fastStr}).
 \gb{come back to this}.
-Most of this signal seems to derive from the \textit{P. trichocarpa} samples:
-covariance within predominantly \tri{} layers decays more quickly with distance
-than within the majority \bals{} layer, as seen in Fig \ref{populus_sp_layer_covs}.  
-This indicates that, although isolation by distance is a feature of the dataset as a whole, 
-its magnitude varies considerably across biological units - 
-an aspect of the data \texttt{conStruct} is able capture well.
-Based on our cross-validation results we view the $K=3$ spatial model 
-as a sufficient description of the data.
 
@@
 At higher values of $K$, drainages at the southern tip of \tri{} sampling 
-begin to split out into their own layers.  
-This may potentially be the result of introgression from southern \textit{Populus} species, 
-such as \textit{P. angustifolia} or \textit{fremontii} \citep{Zhou2012,geraldes_etal_2014}
+begin to split out into their own layers,
+perhaps due to introgression from the southern neighbors
+\textit{P. angustifolia} or \textit{fremontii} \citep{Zhou2012,geraldes_etal_2014}
 

### 2017-09-14 writeup/conStruct.tex
edited bears

@@
 \subsubsection*{Black bears}
+
 The American black bear, \textit{Ursus americanus}, is endemic to North America
@@
 from which they subsequently expanded to occupy their current range
-\citep{WoodingWard1997,Byun1997,Stone2000,Puckett2015}.
+\citep{WoodingWard1997,Byun1997,Stone2000,Puckett2015},
 %The retreat of the glaciers was relatively recent in units of 
@@
 %mating opportunities and exhibit a pattern of IBD.
-As with the poplars described above, 
-this system is likely to present patterns of genetic structure 
-that are both continuous and discrete, 
-making it well-suited as a test for our method.
-We organize the results of our analyses of 
-the complicated landscape of genetic structure in black bears 
-around the following questions:
+likely leading to both continuous and discrete patterns of genetic structure.
+We organize our results around the following questions:
 \begin{enumerate}
-\item How many historical populations are reflected in modern patterns of genetic variation?
-\item Do we see support for a pattern of isolation by distance within each of the inferred discrete groups?
+    \item How many distinct populations (probably, distinct glacial refugia)
+        are reflected in modern patterns of genetic variation?
+    \item How strong is isolation by distance within each inferred group?
 \end{enumerate}
-
-Our data, from \cite{Puckett2015}, consist of 95 individuals 
+Differing strengths of isolation by distance might indicate 
+different levels of habitat connectivity,
+dispersal behavior,
+or different postglacial histories.
+
+We use data from \cite{Puckett2015}, consisting of 95 individuals 
 sampled across the United States and on the West coast of Canada,
@@
 
-Bears in the sample are partitioned into two main groups ($K \geq 2$):
+The results partition the sampled bears into two main groups:
 one (red) to the east of the Rocky Mountains, 
@@
 Additional layers in the spatial model have strong statistical support 
-up until around $K=5$ or 6 (Fig \ref{bear_xvals}).
-However, upon examining layer contributions (Fig \ref{bear_laycon}),
-it is clear that the additional spatial layers beyond $K=2$ are contributing only a very small amount toward 
-total covariance between individuals.
-The locations of the individuals inferred to be admixed between the two layers 
-provide information about a possible zone of contact and migration between these groups 
-along the northwestern coast of North America and near the Cascade Range in the northern U.S.
-These findings are consistent with a scenario of postglacial expansion from two refugia, 
+up until around $K=5$ or 6 (Fig \ref{bear_xvals}),
+but additional spatial layers beyond $K=2$ contribute little to total covariance (Fig \ref{bear_laycon}).
+% However, upon examining layer contributions (Fig \ref{bear_laycon}),
+% it is clear that the additional spatial layers beyond $K=2$ are contributing only a very small amount toward total covariance between individuals.
+The locations admixed individuals 
+% between the two layers provide information about a possible zone of contact and migration between these groups 
+% along the northwestern coast of North America and near the Cascade Range in the northern U.S.
+are consistent with a scenario of postglacial expansion from two refugia, 
 one in the American Southwest and one in the American Southeast, 
-but lack of any samples from Canada and Mexico, 
+meeting near the Northwest coast of North America and the Cascade Range.
+However, lack of any samples from Canada and Mexico, 
 and lack of denser sampling across northern North America, 
@@
 the third cluster (in green) exhibits a clear East-West gradient that 
-is overlain across the discrete structure between the Southwest cluster and the Southeast.
+overlays the discrete structure between the Southwest cluster and the Southeast.
 The results from fastSTRUCTURE are are not identical to 
@@
 Across all values of $K$ for which we ran \texttt{conStruct},
-we see strong support for the spatial model over the nonspatial model (Fig \ref{bear_xvals}),
-indicating the isolation by distance seems to be a prominent feature of the data.
+we see strong support for the spatial model over the nonspatial model (Fig \ref{bear_xvals}).
+% indicating the isolation by distance seems to be a prominent feature of the data.
 This pattern may resolve a discrepancy between our results and 

### 2017-09-14 writeup/conStruct.tex
edited through disco

@@
 
-The pattern of isolation by distance is ubiquitous in nature;
-this is unsurprising, as mating opportunities are distance-limited in most species.
-This common pattern can flummox commonly used clustering methods, 
-which must use a possibly large number of discrete clusters
-to describe such continuous population differentiation.
-Inference of these spurious clusters can have implications for 
+The difficulty of dispersing across geography makes
+the pattern of isolation by distance ubiquitous in nature.
+% this is unsurprising, as mating opportunities are distance-limited in most species.
+% This common pattern can flummox 
+Commonly used clustering methods
+must use a possibly large number of discrete clusters
+to describe this simple pattern
+such continuous population differentiation.
+The resulting spurious clusters can have implications for 
 downstream analysis and interpretation of genetic data, 
 including conservation management decisions. 
-
-%and 
-%identification of the genetic basis of adaptation or disease traits.
+However, discrete groupings are also often produced by important biological processes.
 
@@
 we avoid erroneously ascribing population differentiation to discrete population clusters.
-We also present a cross-validation approach for comparing between models, 
-although we stress the limits of this or any model selection procedure for genetic clustering.
+To aid comparison between models,
+we also present a cross-validation approach 
+and a way to describe the contribution of each spatial layer to the model
+(but caution against overly strict interpretation of these).
 
@@
 One advantage these have over \texttt{conStruct} is speed when the number of samples is large.
-Although \texttt{conStruct}'s computation time is invariant 
-with the number of loci included in the dataset 
+Although \texttt{conStruct}'s computation time is independent
+of the number of loci included in the dataset 
 (after the initial calculation of the allelic covariance), 
@@
 The computationally limiting step is the inversion of the parametric covariance, 
-which is an operation of order $O(n^3)$, 
+% actually it is like O(n^2.3) but I don't know which algorithm STAN uses
+% which is an operation of order $O(n^3)$, 
+which scales more than quadratically with the number of samples,
 whereas computation time for, e.g., STRUCTURE, 
@@
 and using only that analysis to communicate results, 
-we emphasize that there is no true best value of $K$.
-The choice of best $K$ is always relative to the data in hand.
+we emphasize that 
+% there is no true best value of $K$. 
+% HOW ABOUT: there is no truth.
+the choice of best $K$ is always relative to the data in hand
+and the questions to be answered.
 %This means both that sampling different individuals or loci 
 %from the same populations may yield a different ``best" $K$, 
-From the statistical perspective, Unless the data were generated under the model itself 
-that the support for larger values of $K$ is likely to increase with increasing amounts of data
+From a statistical perspective, unless the data were generated under the model itself,
+the support for larger values of $K$ is likely to increase with increasing amounts of data.
 %(although perhaps only logarithmically).
@@
 or of implementing an integrated conservation plan across multiple demes within a layer.
+Conversely, two populations that are easily distinguishable based on abundant genetic data
+need not be strongly differentiated --
+cluster assignment is not a replacement for absolute measures of genetic divergence.
 
@@
 and the layer-specific relatedness parameter $\phi^{(k)}$ is low, 
-individuals at a far spatial remove may be in the same layer but have very low pairwise relatedness.
+individuals at a large spatial distance may be in the same layer but have very low pairwise relatedness.
 It is possible that this is happening in Fig \ref{populus_sp_pies}. 
@@
 (a star population phylogeny, although the branches can have different lengths thanks to the $\phi^{(k)}$ parameter), 
-similar to the F-model in \citep{falush2003}.
+similar to the F-model in \citet{falush2003}.
 However, we could extend the existing model by implementing 
@@
 as we have done in a previous approach \citep{spacemix} to model long distance dispersal. 
-Indeed, we could allow entire layers to bud off of a particular location on another layer. 
-This would enable us to model scenarios of range expansion or domestication, 
+We could even allow entire layers to bud off of a particular location on another layer. 
+This would enable more explicit modeling of range expansion or domestication, 
 in which a set of individuals are thought to have ancestry that originated from
@@
 
-A final direction is to model relatedness within a layer as a spatiotemporal process, 
+A final direction would be to model relatedness within a layer as a spatiotemporal process, 
 in which covariance decays both with distance in space and in time.  
@@
 Institutes of Health under award numbers NIH R01-GM108779 to GC,
-and the National Science Foundation under award numbers NSF \# 1148897 and \# 1402725 to GB.
+and the National Science Foundation under award numbers NSF \#1148897 and \#1402725 to GB.
 

### 2017-09-14 writeup/conStruct.tex
finished edits

@@
 this covariance has a close relationship to the pairwise genetic distance:
-if $D_{i,j}$ is the mean density of sites at which random samples from $i$ and $j$ differ at a randomly chosen locus,
-then $\Omega_{i,j} = (1 - 2 D_{i,j})/4$.
+if $\pi_{i,j}$ is the mean density of sites at which random samples from $i$ and $j$ differ at a randomly chosen locus,
+then $\Omega_{i,j} = (1 - 2 \pi_{i,j})/4$.
 Therefore, 
@@
 To accommodate admixture between layers, 
-we model allele frequencies within samples as linear combinations of the population-specific 
-drift terms across the layers from which they draw ancestry.
-To do so, we introduce a term $w^{(k)}_{i}$ that describes the
-admixture proportion of sample $i$ in layer $k$ ($\sum_{K}w^{(k)}_{i} = 1$),
-which can be interpreted as the proportion of the genome in the $i$th
+we model sampled genomes as drawn from
+allele frequencies that are weighted averages of the local frequencies in each layer
+from which they draw ancestry.
+The weights, $w^{(k)}_{i}$, describe the
+``admixture proportion'' of sample $i$ in layer $k$
+(Note that $\sum_{k=1}^K w^{(k)}_{i} = 1$ for each $i$.)
+These can be interpreted as the proportion of the genome in the $i$th
 sample that came from the $k$th layer 
-(or the probability that an allele at a locus is drawn from $k$).
+(or the probability that an allele at a locus is drawn from layer $k$).
 The allele frequency in the $i$th sample at the $\ell$th locus can therefore be written as:
@@
 F_{i,\ell} = \epsilon_{\ell} + \sum\limits_{K} w^{(k)}_{i}\left( 
-  \Delta^{(k)}_{\ell} + \Delta^{(k,i)}_{\ell}\right) + \Delta^{(i)}_{\ell}	
+  \Delta^{(k)}_{\ell} + \Delta^{(k,i)}_{\ell}\right) + \Delta^{(i)}_{\ell}	 ,
 \label{drift_terms_admix}
 \end{equation}
-and the covariance between $i$ and $j$ across loci is
+and so the covariance between $i$ and $j$ across loci is
 \begin{equation}
@@
 \text{Cov}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell}) 	\right) +
-\delta_{i=j} \text{Var}(\Delta_i)
+\delta_{i=j} \text{Var}(\Delta_i) .
 \label{admixed_spatial_cov}
@@
 \begin{equation}
-\text{Cov}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell}) = \alpha^{(k)}_0 \times \left(\text{exp} \left(  -(\alpha^{(k)}_D D_{i,j})^{\alpha^{(k)}_2}\right) \right)
+\text{Cov}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell}) = \alpha^{(k)}_0 \times \left(\text{exp} \left(  -(\alpha^{(k)}_D D_{i,j})^{\alpha^{(k)}_2}\right) \right) .
 \end{equation}
@@
 which is $f_i^2 n/(n-1)$.
+
 \paragraph{Likelihood}
-If our allele frequency deviations are well approximated by a Gaussian, 
+If allele frequency deviations are well approximated by a Gaussian, 
 their sample allelic covariance is a sufficient statistic,
@@
 \end{equation}
+where $\mathcal{W}$ is the Wishart likelihood function.
 
@@
 \paragraph{Priors}
-We use a Bayesian approach to parameter inference,
-
+We use a Bayesian approach to parameter inference.
 A table of all parameters, their descriptions, and their priors is given in Table \ref{tab:param_prior_tab}.
@@
 \section{Cross validation}\label{Xvalidation}
-To perform model comparison, we employ a Monte Carlo cross-validation approach \citep{picard1984}.
-This cross-validation procedure generates a mean predictive accuracy for each model and each value of $K$, 
+
+We employ Monte Carlo cross-validation approach for model comparison \citep{picard1984}. 
+This procedure generates a mean predictive accuracy for each model and each value of $K$, 
 as well as a confidence interval around that mean,
@@
 \section{Calculating layer contributions}\label{layer_contribution}
-First, we describe the allelic covariance between two samples, $i$, and $j$.
-Note that this is the allelic covariance described in Eqn \eqref{allelic_covariance},
-in which we randomly polarize alleles,
-and not the genetic covariance, which subtracts mean allele frequencies.
-Let $A$ and $B$ be a randomly chosen alleles from the two samples, respectively,
+
+Let $A$ and $B$ be randomly chosen alleles from samples $i$ and $j$ respectively,
 at a randomly chosen locus 
@@
 since $X$ and $Y$ take the values $\pm 1$,
-and thanks to the random polarization, $\E[A] = \E[A] = 1/2$,
+and thanks to the random polarization, $\E[A] = \E[B] = 1/2$,
 $$\begin{aligned}
@@
 while $\P\{ X \neq Y \}$ is the probability that they disagree.
-For the first thing to happen, $i$ and $j$ must coalesce before coalescing with the reference samples,
-so for $X$ and $Y$ to agree they must be more closely related than typical in the population.
-The last part, $\P\{ XY \neq 0\}$, is the probability that both samples disagree from the references.
-(This implies that $\E[XY] = 1 - 2 D_{ij}$, 
-where $D_{ij}$ is the probability that two randomly chosen alleles differ, 
-which is the genetic divergence.)
+This implies that $\E[XY] = 1 - 2 \pi_{ij}$, 
+where $\pi_{ij}$ is the probability that two randomly chosen alleles differ, 
+which is the genetic divergence.
 
@@
 To decide whether or not $X$ and $Y$ will agree,
-first each sample randomly chooses a layer: call these $I$ and $J$.
-The probability that $X$ chooses layer $k$ is $\P({I=k})$, which is $w_i^{(k)}$, 
+first each sample randomly chooses a layer: call these layers $I$ and $J$.
+The probability that $X$ chooses layer $k$ is $\P({I=k})=w_i^{(k)}$, 
 the $i$th sample's admixture proportion in the $k$th layer.
@@
 then they agree with a probability $p_\gamma + q^{(k)}_{ij}$ that depends on their distance apart.
-Here
-
+By the above,
+the probability of agreement is $\P\{A=B\} = 2 \cov[A,B] + 1/2$,
+and so
 $$\begin{aligned}
     p_\gamma &= (2 (\gamma + \delta_{ij} \eta_i) + 1/2) \\
-    q^{(k)}_{ij} &= \left(2 \alpha_0^{(k)} \exp\left( - \left(\alpha_D^{(k)} D_{ij}\right)^{\alpha_2^{(k)}} \right) + 2 \mu^{(k)}  + 1/2 \right) .
+    q^{(k)}_{ij} &= \left(2 \alpha_0^{(k)} \exp\left( - \left(\alpha_D^{(k)} D_{ij}\right)^{\alpha_2^{(k)}} \right) + 2 \phi^{(k)}  + 1/2 \right) .
 \end{aligned}$$
@@
 This suggests defining the overall contribution of layer $k$ to agreement, $\xi^{(k)}$, 
-to be the average of that quantity over $i$ and $j$.
-
-%
+to be the average of that quantity over $i$ and $j$:
 \begin{equation}
@@
 						\right) + 2\phi^{(k)} + \frac{1}{2}
-				\right)
+				\right) ,
 \label{layer_contribution}
 \end{equation}
-%
 which is that layer's contribution to agreement between samples 
@@
 as the relative contribution of the $k$th layer to total agreement:
-%
 \begin{equation}
@@
 	{\sum\limits_{k=1}^{K}
-	    	\xi^{(k)}}
+	    	\xi^{(k)}} .
 \label{rel_layer_contribution}
 \end{equation}
-%
+This is the quantity that is plotted in, e.g., Figures \ref{simK1_laycon} and \ref{bear_laycon}.
 

### 2017-09-14 writeup/conStruct.tex
merged P's changes and added Fig4

--- 2017-09-14 writeup/conStruct.tex
@@
 \author{
-Gideon S.\ Bradburd$^{1,a}$, 
-Graham M.\ Coop$^{2,b}$,
-Peter L.\ Ralph$^{3,c}$}
+Gideon S. Bradburd$^{1,a}$, 
+Graham M. Coop$^{2,b}$,
+Peter L. Ralph$^{3,c}$}
 
@@
 
-\plr{I felt like we said "continuous and discrete" too many times in the abstract.}
-
 \newpage
@@
 in the continuous divergence of populations when they unambiguously become discrete species.
-Descriptions of taxa using phylogenetic trees omit within-taxon variation,
-especially when these are intermingling subgroups.
 The issue of identifying meaningful biological subunits 
@@
 Furthermore, we often need to identify subsets of indivduals resulting from reasonably coherent
-evolutionary histories for downstream analyses to learn about population history and adaptation,
-\plr{as for example XXX?}
+evolutionary histories for downstream analyses to learn about population history and adaptation.
 Conversely,
@@
 so it may be advisable to filter these prior to analysis
-if these are likely to contain a large percentage of errors.
+if they are likely to contain a large percentage of errors.
 
@@
 There is a fixed amount of covariance between layers, irrespective of spatial location.
-Within each layer allele frequencies are expected to change gradually with distance,
+Within each layer, allele frequencies are expected to change gradually with distance,
 but observed frequencies can change abruptly at many loci 
-if the proportions by which individuals' ancestry derive from each layer 
+if the proportions of ancestry individuals derive from each layer 
 (the ``admixture'' proportions) 
@@
 and the $\alpha^{(k)}$ parameters control the shape of the decay of
-covariance with distance in the cluster.
+covariance with distance in the layer.
 Our choice of a powered-exponential decay, 
@@
 % (Analytical theory \citep{barton2002neutral} predicts the decay should be proportional to $\exp{-D}/\sqrt{D}$.)
-The $\phi^{(k)}$ is a parameter that describes the background covariance within the cluster. 
-If two samples draw 100\% of their ancestry from cluster $k$, then their covariance under the model is $G^{(k)}_{i,j}$;
+The $\phi^{(k)}$ is a parameter that describes the background covariance within the layer. 
+If two samples draw 100\% of their ancestry from layer $k$, then their covariance under the model is $G^{(k)}_{i,j}$;
 if they are furthermore geographically very close ($D_{i,j}=0$)
@@
 If the geographic distance between them is very large, 
-their covariance will be equal to to the background level $\phi^{(k)}$ within the cluster.
-%% plr: this is no longer true
-% The ``shared drift'' parameter $\phi^{(k)}$ can be interpreted as the branch length 
-% connecting the $k$th population to the population ancestral to all modeled
-% layers \citep[see for example ][]{patterson_ancient_2012,peter_fstats}. 
-In the methods section \ref{rationale} we lay out a
+their covariance will be equal to to the background level $\phi^{(k)}$ within the layer.
+The ``shared drift'' parameter $\phi^{(k)}$ is analogous to 
+the branch length connecting the $k$th population to the population ancestral to all modeled
+layers \citep[see for example ][]{patterson_ancient_2012,peter_fstats},
+although they cannot be directly compared because 
+we are modeling the allelic, rather than genetic, covariance. 
+In the Appendix section \ref{rationale} we lay out a
 simple model of allele frequencies underlying this covariance model.
@@
 contains a nonspatial assignment model as a special case 
-(see \secref{model_app} in the appendix for a more in-depth discussion). 
+(see \secref{model_app} in the Appendix for a more in-depth discussion). 
 By setting $\alpha_0^{(k)}$ to zero for all $k$, 
@@
 The best model is judged to be the simplest one with significantly better predictive accuracy
-than others (see \secref{Xvalidation} of the appendix for more on our cross-validation procedure).
+than others (see \secref{Xvalidation} in the Appendix for more on our cross-validation procedure).
 In general, larger values of $K$ allows the model more flexibility,
@@
 
-Another way to describe the practical significance of additional layers
+Another way to describe how additional layers contribute to model fit
 is to calculate each layer's relative contribution to total covariance, 
@@
 is quite harsh against intermediate admixture values (see Table \ref{tab:param_prior_tab}),  
-encouraging the model to ``not use" unnecessary layers if they are present, 
+encouraging the model to ``not use" unnecessary layers, 
 so that they will have a low contribution to overall covariance.
-To do this, we use the following alternative description of our covariance model:
-the genomes of any pair of individuals \emph{agree} with some background probability,
-but this probability of agreement is increased on any segment of genome 
-that both have inherited through the same layer.
-(The amount it increases depends on how far apart they are.)
-We use this characterization to quantify the relative contributions of each layer,
-by computing the average contribution to increased probability of agreement
-as described in Appendix \secref{layer_contribution}). 
+We use the connection of allelic covariance to the probability of agreement (identity) between genotypes
+to calculate a measure of overall contribution of each layer to the model.
+
+The contribution of the $k$th layer can be calculated as
+its contribution to the parametric covariance between samples 
+summed over the upper triangle (excluding the diagonal) of the covariance matrix.
+The contribution of the $k$th layer is its absolute contribution relative to those of all other layers, 
+i.e., its contribution divided by the sum of the contributions of all $K$ layers
+(for details on calculating this contribution, see Appendix \secref{layer_contribution}). 
 This layer contribution is similar to the ``ancestry contribution'' proposed by \cite{fastStructure}. 
@@
 \subsection*{Simulations}
-
-To test the method, we first generated data using the coalescent simulator \texttt{ms} \citep{Hudson2002}.
-In each simulation,
-we split a single ancestral population into $K$ subpopulations $\tau_{\text{s}}$ units of coalescent time in the past,
-and at time $\tau_{\text{e}}$ in the past,
-each of these discrete populations instantaneously colonized a
-separate $6 \times 6$ square lattice of demes.
-Migration on each lattice was to nearest neighbors (eight neighbors, including diagonals).
+We first test the method with scenarios generated using the coalescent simulator \texttt{ms} \citep{Hudson2002}.
+To generate discrete population structure, we simulate $K$ distinct populations,
+each of which split from a common ancestor $\tau_{\text{s}}$ units of coalescent time in the past,
+without subsequent migration between them.
+To generate continuous differentiation within each layer,
+at time $\tau_{\text{e}}$ in the past,
+each of these discrete populations instantaneously colonizes an
+independent lattice of demes,
+for which we use a stepping stone model with symmetric migration 
+to nearest neighbors (eight neighbors, including diagonals).
 Finally, at time $\tau_{\text{a}}$ in the past 
-we collapsed those $K$ discrete layers into a single grid of demes,
-choosing various amounts of admixture from these different layers
+we generate a single dataset 
+by collapsing those $K$ discrete layers into a single grid of demes 
+that have various amounts of admixture from these different layers
 (see Fig \ref{sim_setup}). 
-We collapsed the layers together using random admixture, spatially autocorrelated proportions 
-(see Appendix \secref{sim_details} for details), 
+We collapse the layers together using random admixture proportions 
+that are spatially autocorrelated (see Appendix \secref{sim_details} for details), 
 so that both the frequencies in layers and the layer memberships are
@@
 in each sampling 10,000 unlinked loci from each of 20 haploid individuals from every deme.
-We then ran both spatial and nonspatial \texttt{conStruct} analyses on each simulated dataset with $K$ between 1 and 7,
-and compared predictive performance of the models using cross-validation to compare the predictive performance.
-% To confirm that results from our nonspatial model were qualitatively similar to those of 
-% other nonspatial model-based clustering algorithms, 
-For comparison,
-we also analyzed each simulated dataset using fastSTRUCTURE \citep{fastStructure}, using $K=2--4$.
-% inferring admixture under models with $K=2$ through 4 (Figs \ref{fastStr_simK1}--\ref{fastStr_simK3}).
-
-\plr{rearranged this to focus more on take-homes}
-
-% In Fig \ref{simK1_pies}, we show the results of the dataset simulated with $K=1$, 
-% contrasting the results of the nonspatial model runs with 2--4 layers 
-% with those of spatial model with 2--4 layers.
-% 
-% In Fig \ref{sim_xvals} we show the results of the cross-validation analysis for each of the three simulated datasets.
-% The results of all analyses run on datasets simulated using $K=1$ through 3 are shown in 
-% Figs \ref{simK1_nsp_pies}-\ref{simK3_adprop_fit}.
-% 
-% To confirm that results from our nonspatial model were qualitatively similar to those of 
-% other nonspatial model-based clustering algorithms, 
-% we also analyzed each simulated dataset using fastSTRUCTURE \citep{fastStructure}, 
-% inferring admixture under models with $K=2$ through 4 (Figs \ref{fastStr_simK1}--\ref{fastStr_simK3}).
-
-With these simulations,
-spatial \texttt{conStruct} does not create spurious clusters when there are none:
-Figures \ref{simK1_pies}, \ref{simK2_sp_pies}, and \ref{simK3_sp_pies} 
-show that subsequent layers beyond the number used for simulation are unused 
-when data is simulated with $K=1$ but analyzed with $K>1$ --
-the layers are present, but contribute very little to any population.
+
+We ran \texttt{conStruct} analyses on each of these simulated datasets, 
+specifying both a spatial and nonspatial model for $K = $ 1 through 7, 
+then used cross-validation to compare the predictive performance of the models.
+In Fig \ref{simK1_pies}, we show the results of the analyses of the dataset simulated under $K=1$, 
+contrasting the results of the nonspatial model runs with 2 - 4 layers 
+with those of spatial model with 2 - 4 layers.
+In Fig \ref{sim_xvals} we show the results of the cross-validation analysis for each of the three simulated datasets.
+The results of all analyses run on datasets simulated using $K=1$ through 3 are shown in 
+Figs \ref{simK1_nsp_pies}-\ref{simK3_adprop_fit}.
+To confirm that results from our nonspatial model were qualitatively similar to those of 
+other nonspatial model-based clustering algorithms, 
+we also analyzed each simulated dataset using fastSTRUCTURE \citep{fastStructure}, 
+inferring admixture under models with $K=2$ through 4 (Figs \ref{fastStr_simK1}-\ref{fastStr_simK3}).
+
+A number of results are immediately clear from these tests with simulated data.
+First, \texttt{conStruct} performs well on all simulated scenarios.
+Across all values of K and for all simulations, 
+the spatial model is correctly preferred over the nonspatial model
+(Fig \ref{sim_xvals}). Second, the predictive accuracy of the nonspatial model increases 
+with the number of layers in the model, 
+as it describes continuous, spatial patterns of relatedness between demes within the same cluster 
+using gradients of admixture across more and more discrete clusters.
+This behavior of the nonspatial model is highlighted in 
+the analyses of data simulated under $K=1$ (Fig \ref{simK1_pies} and \ref{simK1_nsp_pies}):
+as $K$ in the model increases, 
+the sampled area is increasingly divvied up between clusters 
+with ``centers of mass" in the $K$ geographical extremes of the simulated space.
+At $K=2$, the clusters are centered in opposite corners on the map; 
+at $K=3$, in the lower two corners and the top half; 
+at $K=4$, all four corners, and so on.
+These results are qualitatively similar to those generated using fastSTRUCTURE \citep{fastStructure}, 
+which also describes continuous differentiation using admixture between clusters
+ (Fig \ref{fastStr_simK1}-\ref{fastStr_simK3}).
+While it is true that each non-spatial cluster is genetically more similar within itself
+than it is to other clusters,
+we know that these boundaries are arbitrary,
+since the data were simulated with no such boundaries.
+
+In contrast, the predictive accuracy of the spatial \texttt{conStruct} model
+either plateaus at or declines from the true value of $K$ used to simulate the data 
+(Figs \ref{sim_xvals}, \ref{simK1_xval}, \ref{simK2_xval}, and \ref{simK3_xval}).
+And, as $K$ in the spatial model increases, 
+deme membership in layers beyond the number used to simulate the data does not appreciably increase 
+(Fig \ref{simK1_sp_pies}, \ref{simK2_sp_pies}, \ref{simK3_sp_pies}).
 Even when the spatial model is run with $K=7$, 
@@
 
-In contrast, the nonspatial model describes geographic variation
-using gradients of admixture between more and more discrete clusters
-to better approximate the continuous, spatial patterns of relatedness
-(depicted in Figs \ref{simK1_pies} and \ref{simK1_nsp_pies}).
-% as $K$ in the model increases, 
-% the sampled area is increasingly divvied up between clusters 
-% with ``centers of mass" in the $K$ geographical extremes of the simulated space.
-% At $K=2$, the clusters are centered in opposite corners on the map; 
-% at $K=3$, in the lower two corners and the top half; 
-% at $K=4$, all four corners, and so on.
-The fastSTRUCTURE results are qualitatively similar, as shown in Figs \ref{fastStr_simK1}-\ref{fastStr_simK3}.
-Each non-spatial cluster is in truth genetically more similar within itself
-than it is to other clusters,
-but we know that these boundaries are arbitrary,
-since the data were simulated with no such boundaries.
-
-The spatial model's better fit is reflected by increased predictive accuracy:
-as shown in Fig \ref{sim_xvals},
-across all models and choices of $K$, the spatial model is correctly preferred over the nonspatial model.
-As desired, predictive accuracy of the spatial model increases until the true value of $K$,
-and then plateaus or declines
-(Figs \ref{sim_xvals}, \ref{simK1_xval}, \ref{simK2_xval}, and \ref{simK3_xval}),
-while the predictive accuracy of the nonspatial model increases as subsequent clusters are added.
-
-% In contrast, the predictive accuracy of the spatial \texttt{conStruct} model
-% either plateaus at or declines from the true value of $K$ used to simulate the data 
-% (Figs \ref{sim_xvals}, \ref{simK1_xval}, \ref{simK2_xval}, and \ref{simK3_xval}).
-
-The unimportance of spurious layers can be seen in plots of layer contributions,
-Figs \ref{simK1_laycon}, \ref{simK2_laycon}, and \ref{simK3_laycon}.
+Finally, we calculated layer contribution for 
+analyses run in each simulation 
+(Fig \ref{simK1_laycon}, \ref{simK2_laycon}, and \ref{simK3_laycon}).
 In the spatial analyses, once we pass the true $K$, 
 subsequent layers add little in terms of (co)variance explained; 
-in contrast, additional clusters in the nonspatial analyses continue to contribute substantially.
+in contrast, additional clusters in the nonspatial analyses continue to
+be important for describing variation, 
+as they are forced to partition up the spatial patterns within clusters.
 
@@
 \subsection*{Empirical Applications}
-
 To demonstrate the utility of this method, we analyzed empirical population genomic data from two systems:
@@
 
+Here, we focus on two \textit{Populus} species in the Pacific Northwest: 
 \textit{Populus trichocarpa}, the black cottonwood,
-and \textit{Populus balsamifera}, the balsam poplar,
-have a broad zone of overlap in the Pacific Northwest,
-where they are hypothesized to hybridize \citep{geraldes_etal_2014, suarezgonzalez_etal_2016}.
+and \textit{Populus balsamifera}, the balsam poplar.
+These species have a broad zone of overlap and are hypothesized to hybridize \citep{geraldes_etal_2014, suarezgonzalez_etal_2016}.
 Both species are sampled over a large geographic region, 
 and show spatial patterns of genetic and phenotypic variation \citep{slavov_etal_2012, mckown_etal_2013},
-making the system well-suited for application of our method.
-% to test an analysis designed to describe
-% how genetic structure is partitioned both discretely across populations (or species)
-% and continuously with geographic distance.
+making the system well-suited to test an analysis designed to describe
+how genetic structure is partitioned both discretely across populations (or species)
+and continuously with geographic distance.
 We organize the results of our analyses around the following questions: 
 \begin{enumerate}
-
-    \item To what degree has hybridization blurred the boundaries between 
-        \textit{trichocarpa} and \textit{balsamifera}?
-        (As an extreme case, does genetic differentiation support these as separate species,
-        as opposed to a single cline of ancestry?)
-    % Given the known history of hybridization between this pair of poplar species, 
-    % is there support for two discrete populations, 
-    % rather than a single cline of ancestry across their overlap zone?
-
-    \item Does the only significant boundary of population structure
-        fall along the species boundary (if any),
-        % If there is evidence for discrete population structuring, 
-        % does it fall solely along species boundaries, 
-        or is there sub-structuring within species?
-
-    \item Does the strength of isolation by distance differ between inferred layers?
-        This may indicate, e.g., different speeds of postglacial expansion
-        or primary modes of dispersal.
-        % Is there support for a pattern of isolation by distance within
-        % the inferred layers, and, if so, 
-        % does it differ between different layers?
-
+\item Given the known history of hybridization between this pair of poplar species, 
+is there support for two discrete populations, 
+rather than a single cline of ancestry across their overlap zone?
+\item If there is evidence for discrete population structuring, 
+does it fall solely along species boundaries, 
+or is there sub-structuring between different populations of the same species?
+\item Is there support for a pattern of isolation by distance within
+the inferred layers, and, if so, 
+does it differ between different layers?
 \end{enumerate}
 
-We use data from \cite{geraldes_etal_2014}, 
-consisting of 434 individuals sampled from 35 drainages 
-genotyped at just over 33,000 loci (map of the sampling shown in Fig \ref{populus_map}).
-\plr{How do the indivs partition across the drainages?  Perhaps give range of number of indivs per drainage?}
+Our data, from \cite{geraldes_etal_2014}, 
+consist of 434 individuals sampled from 35 drainages 
+and genotyped at just over 33,000 loci (map of the sampling shown in Fig \ref{populus_map}).
 The data were generated using RAD-seq, 
@@
 	\centering
-		{\includegraphics[width=6in,height=2in]{figs/populus/Fig4_pop_sp_results.pdf}}
-	\caption{
-	Map of admixture proportions estimated for the \textit{Populus} dataset 
-    using the spatial \texttt{conStruct} model for $K=2$ through 4.
+		{\includegraphics[width=6in,height=4in]{figs/populus/Fig4_pop_sp_results.pdf}}
+	\caption{
+	Maps of admixture proportions estimated for the \textit{Populus} dataset 
+	using the spatial model for $K=2$ through 4 (a-c), 
+	as well as the corresponding layer-specific covariance curves estimated under each model (d-f).
     }\label{populus_pies}
@@
 to distinct layers, with some populations drawing ancestry from multiple layers.
-Based on cross-validation results, we view the $K=3$ spatial model 
-as a sufficient description of the data,
-with additional structure of uncertain significance.
 This provides strong support for discrete population structure between the two species,
@@
 at values of $K$ above 2, further discrete substructure was inferred within the \textit{P. trichocarpa} samples, 
-with no substructure within \bals{}.
+while \bals{} samples displayed no substructure up to $K=7$.
+This lack of substructure within the sampled \bals{} populations is largely in agreement with results from 
+\citet{keller_etal_2010}, who found that \bals{} in the region is characterized by relatively low diversity, 
+and probably represents a recent, postglacial expansion.
+
+For values of $K<4$, the spatial model was strongly preferred over the nonspatial model (Fig \ref{populus_xvals}).
 Most of this signal seems to derive from the \textit{P. trichocarpa} samples:
-% covariance within predominantly \tri{} layers decays more quickly with distance than within the majority \bals{} layer, 
-as seen in Fig \ref{populus_sp_layer_covs},
-there is almost no isolation by distance within the \bals{} layer ($\alpha_D \approx 0$).
-Both points are
-in agreement with \citet{keller_etal_2010}, 
-who found low diversity within the region's \bals{},
-probably as the result of a recent postglacial expansion.
-
-For all values of $K$, the spatial model was preferred over the nonspatial model (Fig \ref{populus_xvals}) (strongly for $K<4$).
-The nonspatial results were broadly concordant with those of the fastSTRUCTURE runs (Fig \ref{populus_fastStr}).
-\gb{come back to this}.
+covariance within predominantly \tri{} layers decays more quickly with distance
+than within the majority \bals{} layer, as seen in Fig \ref{populus_sp_layer_covs}.  
+This indicates that, although isolation by distance is a feature of the dataset as a whole, 
+its magnitude varies considerably across biological units - 
+an aspect of the data \texttt{conStruct} is able capture well.
+Based on our cross-validation results we view the $K=3$ spatial model 
+as a sufficient description of the data.
 
@@
 At higher values of $K$, drainages at the southern tip of \tri{} sampling 
-begin to split out into their own layers,
-perhaps due to introgression from the southern neighbors
-\textit{P. angustifolia} or \textit{fremontii} \citep{Zhou2012,geraldes_etal_2014}
+begin to split out into their own layers.  
+This may potentially be the result of introgression from southern \textit{Populus} species, 
+such as \textit{P. angustifolia} or \textit{fremontii} \citep{Zhou2012,geraldes_etal_2014}
+
+\gb{The results of the nonspatial analysis differed from those of the fastSTRUCTURE runs (Fig \ref{populus_fastStr}), 
+although both showcased the tendency of nonspatial models to describe continuous variation with multiple clusters.
+Replicate runs of the nonspatial \texttt{conStruct} model produced different results, 
+evidence of multimodality in the likelihood surface, 
+which we might expect given the poor correspondence of the model to the data.}
 
 \subsubsection*{Black bears}
-
 The American black bear, \textit{Ursus americanus}, is endemic to North America
@@
 from which they subsequently expanded to occupy their current range
-\citep{WoodingWard1997,Byun1997,Stone2000,Puckett2015},
+\citep{WoodingWard1997,Byun1997,Stone2000,Puckett2015}.
 %The retreat of the glaciers was relatively recent in units of 
@@
 %mating opportunities and exhibit a pattern of IBD.
-likely leading to both continuous and discrete patterns of genetic structure.
-We organize our results around the following questions:
+As with the poplars described above, 
+this system is likely to present patterns of genetic structure 
+that are both continuous and discrete, 
+making it well-suited as a test for our method.
+We organize the results of our analyses of 
+the complicated landscape of genetic structure in black bears 
+around the following questions:
 \begin{enumerate}
-    \item How many distinct populations (probably, distinct glacial refugia)
-        are reflected in modern patterns of genetic variation?
-    \item How strong is isolation by distance within each inferred group?
+\item How many historical populations are reflected in modern patterns of genetic variation?
+\item Do we see support for a pattern of isolation by distance within each of the inferred discrete groups?
 \end{enumerate}
-Differing strengths of isolation by distance might indicate 
-different levels of habitat connectivity,
-dispersal behavior,
-or different postglacial histories.
-
-We use data from \cite{Puckett2015}, consisting of 95 individuals 
+
+Our data, from \cite{Puckett2015}, consist of 95 individuals 
 sampled across the United States and on the West coast of Canada,
@@
 
-The results partition the sampled bears into two main groups:
+Bears in the sample are partitioned into two main groups ($K \geq 2$):
 one (red) to the east of the Rocky Mountains, 
@@
 Additional layers in the spatial model have strong statistical support 
-up until around $K=5$ or 6 (Fig \ref{bear_xvals}),
-but additional spatial layers beyond $K=2$ contribute little to total covariance (Fig \ref{bear_laycon}).
-% However, upon examining layer contributions (Fig \ref{bear_laycon}),
-% it is clear that the additional spatial layers beyond $K=2$ are contributing only a very small amount toward total covariance between individuals.
-The locations admixed individuals 
-% between the two layers provide information about a possible zone of contact and migration between these groups 
-% along the northwestern coast of North America and near the Cascade Range in the northern U.S.
-are consistent with a scenario of postglacial expansion from two refugia, 
+up until around $K=5$ or 6 (Fig \ref{bear_xvals}).
+However, upon examining layer contributions (Fig \ref{bear_laycon}),
+it is clear that the additional spatial layers beyond $K=2$ are contributing only a very small amount toward 
+total covariance between individuals.
+The locations of the individuals inferred to be admixed between the two layers 
+provide information about a possible zone of contact and migration between these groups 
+along the northwestern coast of North America and near the Cascade Range in the northern U.S.
+These findings are consistent with a scenario of postglacial expansion from two refugia, 
 one in the American Southwest and one in the American Southeast, 
-meeting near the Northwest coast of North America and the Cascade Range.
-However, lack of any samples from Canada and Mexico, 
+but lack of any samples from Canada and Mexico, 
 and lack of denser sampling across northern North America, 
@@
 the third cluster (in green) exhibits a clear East-West gradient that 
-overlays the discrete structure between the Southwest cluster and the Southeast.
+is overlain across the discrete structure between the Southwest cluster and the Southeast.
 The results from fastSTRUCTURE are are not identical to 
@@
 Across all values of $K$ for which we ran \texttt{conStruct},
-we see strong support for the spatial model over the nonspatial model (Fig \ref{bear_xvals}).
-% indicating the isolation by distance seems to be a prominent feature of the data.
+we see strong support for the spatial model over the nonspatial model (Fig \ref{bear_xvals}),
+indicating the isolation by distance seems to be a prominent feature of the data.
 This pattern may resolve a discrepancy between our results and 

### 2017-09-14 writeup/figs/bear_figs.R

@@ -138,7 +138,7 @@ pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bear_sp_results.pdf",width=15,h
 	make.bear.redux.result.plot.multipanel2(output.list = output.list.sp,
 											coords = data.block$coords,
 											lump.dist = 200,
-											layer.colors,csr1.order=NULL)
+											layer.colors,csr1.order=c(2,1))
 dev.off()
 
 pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bear_nsp_results.pdf",width=15,height=10)
@@ -154,7 +154,7 @@ pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bear_nsp_results.pdf",width=15,
 	make.bear.redux.result.plot.multipanel2(output.list = output.list.nsp,
 											coords = data.block$coords,
 											lump.dist = 200,
-											layer.colors,csr1.order=c(2,1))
+											layer.colors,csr1.order=NULL)
 dev.off()
 
 

### 2017-09-14 writeup/figs/fig_funcs.R

@@ -70,7 +70,7 @@ viz.admix.results <- function(sim.admix.props,conStruct.results,layer.order=NULL
 
 plot.layer.curves <- function(data.block, conStruct.results, layer.cols=NULL,sample.cols=NULL,add=FALSE){
 	if(is.null(layer.cols)){
-		layer.colors <- c("blue","red","goldenrod1","forestgreen","darkorchid1","deepskyblue","darkorange1","seagreen2","yellow1","black")
+		layer.cols <- c("blue","red","goldenrod1","forestgreen","darkorchid1","deepskyblue","darkorange1","seagreen2","yellow1","black")
 	}
 	order.mat <- order(data.block$geoDist)
 	if(add==FALSE){

### 2017-09-14 writeup/figs/populus_figs.R

@@ -102,9 +102,29 @@ pdf(file=paste0("~/Dropbox/conStruct/writeup/figs/populus/populus_sp_layer_covs.
 	mtext(text="allele frequency covariance",side=2,font=2,cex.axis=2,padj=-59.5,adj=20)
 dev.off()
 
-pdf(file="~/Dropbox/conStruct/writeup/figs/populus/Fig4_pop_sp_results.pdf",width=12,height=4,pointsize=14)
+pdf(file=paste0("~/Dropbox/conStruct/writeup/figs/populus/populus_sp_layer_covs.pdf"),width=12,height=8,pointsize=14)
+	#quartz(width=12,height=8)
+	layout(matrix(c(1:6),nrow=2,ncol=3,byrow=TRUE))
+	par(mar=c(4,5,3,2),oma=c(3,3,3,1))	
+	plot.K.layer.curves(K=1:6,data.block=data.block,output.list=output.list.sp,col.mat1=col.mat1,col.mat2=col.mat2)
+	par(xpd=NA)
+		legend(-0.5,0.06,pch=21,
+				pt.bg=c(1,"forestgreen","forestgreen"),
+				col=c(1,1,"forestgreen"),
+				legend=c("balsamifera - balsamifera",
+						 "balsamifera - trichocarpa",
+						 "trichocarpa - trichocarpa"),cex=0.9,pt.cex=1.5)
+	legend(-11,0.057,pch=c(19,NA),lty=c(NA,1),lwd=c(NA,4),legend=c("sample covariance","layer covariance"))
+	mtext(text="geographic distance",side=1,font=2,cex.axis=2,padj=4.5,adj=-3.55)
+	mtext(text="allele frequency covariance",side=2,font=2,cex.axis=2,padj=-59.5,adj=20)
+dev.off()
+
+pdf(file="~/Dropbox/conStruct/writeup/figs/populus/Fig4_pop_sp_results.pdf",width=12,height=8,pointsize=14)
+	#quartz(width=12,height=8,pointsize=14)
+	layout(matrix(c(1:6),nrow=2,ncol=3,byrow=TRUE))
 	csr1.order <- NULL
-	par(mfrow=c(1,3),mar=c(5,5,1,1))
+	mar <- c(0,3,0,0)
+	par(mar=mar,oma=c(4,4,1,1))
 		for(k in 2:4){
 			csr <- output.list.sp[[k]][[1]]
 			if(k > 2){
[…]

### 2017-09-14 writeup/conStruct.tex

--- 2017-09-14 writeup/conStruct.tex
@@
 by comparing their performance in a common framework. 
-We also validate our claim that the results of our non-spatial method 
+We also validate our claim that the results of our nonspatial method 
 are a close match to those of STRUCTURE-based approaches.
@@
 so that they will have a low contribution to overall covariance.
-To do this, we use the following alternative description of our covariance model:
+
+To calculate layer contributions, we use the following alternative description of our covariance model:
 the genomes of any pair of individuals \emph{agree} with some background probability,
 but this probability of agreement is increased on any segment of genome 
-that both have inherited through the same layer.
-(The amount it increases depends on how far apart they are.)
+that both have inherited through the same layer 
+(the amount it increases depends on how far apart they are.)
 We use this characterization to quantify the relative contributions of each layer,
@@
 Migration on each lattice was to nearest neighbors (eight neighbors, including diagonals).
-Finally, at time $\tau_{\text{a}}$ in the past 
+Finally, at time $\tau_{\text{a}}$ in the past, 
 we collapsed those $K$ discrete layers into a single grid of demes,
@@
 (see Fig \ref{sim_setup}). 
-We collapsed the layers together using random admixture, spatially autocorrelated proportions 
+We collapsed the layers together using random admixture with spatially autocorrelated proportions 
 (see Appendix \secref{sim_details} for details), 
@@
 spatially autocorrelated. 
-We simulated datasets using $K=1$, 2, and 3 layers,
-in each sampling 10,000 unlinked loci from each of 20 haploid individuals from every deme.
+We simulated datasets using $K=1$, 2, and 3 layers;
+in each simulation we sampled 10,000 unlinked loci from each of 20 haploid individuals from every deme.
 We then ran both spatial and nonspatial \texttt{conStruct} analyses on each simulated dataset with $K$ between 1 and 7,
-and compared predictive performance of the models using cross-validation to compare the predictive performance.
+and compared predictive performance of the models using cross-validation.
 % To confirm that results from our nonspatial model were qualitatively similar to those of 
@@
 For comparison,
-we also analyzed each simulated dataset using fastSTRUCTURE \citep{fastStructure}, using $K=2--4$.
+we also analyzed each simulated dataset using fastSTRUCTURE \citep{fastStructure}, using $K=2-4$.
 % inferring admixture under models with $K=2$ through 4 (Figs \ref{fastStr_simK1}--\ref{fastStr_simK3}).
-
-\plr{rearranged this to focus more on take-homes}
 
@@
 Figures \ref{simK1_pies}, \ref{simK2_sp_pies}, and \ref{simK3_sp_pies} 
-show that subsequent layers beyond the number used for simulation are unused 
-when data is simulated with $K=1$ but analyzed with $K>1$ --
+show that subsequent layers beyond the number used for simulation are unused.
+When data are simulated with $K=1$ but analyzed with $K>1$ 
 the layers are present, but contribute very little to any population.
@@
 The fastSTRUCTURE results are qualitatively similar, as shown in Figs \ref{fastStr_simK1}-\ref{fastStr_simK3}.
-Each non-spatial cluster is in truth genetically more similar within itself
+Each nonspatial cluster is in truth genetically more similar within itself
 than it is to other clusters,
 but we know that these boundaries are arbitrary,
-since the data were simulated with no such boundaries.
+because the data were simulated with no such boundaries.
 
@@
 
-The unimportance of spurious layers can be seen in plots of layer contributions,
-Figs \ref{simK1_laycon}, \ref{simK2_laycon}, and \ref{simK3_laycon}.
+The unimportance of spurious layers can be seen in plots of layer contributions
+(Figs \ref{simK1_laycon}, \ref{simK2_laycon}, and \ref{simK3_laycon}).
 In the spatial analyses, once we pass the true $K$, 
@@
 genotyped at just over 33,000 loci (map of the sampling shown in Fig \ref{populus_map}).
-\plr{How do the indivs partition across the drainages?  Perhaps give range of number of indivs per drainage?}
+The range in number of individuals per drainage was between 1 and 50, 
+with most sampling concentrated on \tri{} drainages.
 The data were generated using RAD-seq, 
@@
 with no substructure within \bals{}.
-Most of this signal seems to derive from the \textit{P. trichocarpa} samples:
+There was also strong support for isolation by distance in the dataset, 
+but most of this signal seems to derive from the \textit{P. trichocarpa} samples:
 % covariance within predominantly \tri{} layers decays more quickly with distance than within the majority \bals{} layer, 
-as seen in Fig \ref{populus_sp_layer_covs},
+as seen in Figs \ref{populus_pies}d-f and \ref{populus_sp_layer_covs},
 there is almost no isolation by distance within the \bals{} layer ($\alpha_D \approx 0$).
@@
 probably as the result of a recent postglacial expansion.
-
-For all values of $K$, the spatial model was preferred over the nonspatial model (Fig \ref{populus_xvals}) (strongly for $K<4$).
-The nonspatial results were broadly concordant with those of the fastSTRUCTURE runs (Fig \ref{populus_fastStr}).
-\gb{come back to this}.
 
@@
 \textit{P. angustifolia} or \textit{fremontii} \citep{Zhou2012,geraldes_etal_2014}
+
+Across multiple replicate runs, 
+both nonspatial \texttt{conStruct} and fastSTRUCTURE displayed 
+the successive partitioning of space and the clines of admixture seen in the simulation results.
+The details of each were somewhat different 
+(e.g. Fig \ref{populus_nsp_pies} vs. Fig \ref{populus_fastStr}), 
+and also differed across the replicate analyses.
 
@@
 % it is clear that the additional spatial layers beyond $K=2$ are contributing only a very small amount toward total covariance between individuals.
-The locations admixed individuals 
+The locations of admixed individuals 
 % between the two layers provide information about a possible zone of contact and migration between these groups 
@@
 For example, in Fig \ref{bear_K3}b, 
-the third cluster (in green) exhibits a clear East-West gradient that 
+the third cluster (in gold) exhibits a clear East-West gradient that 
 overlays the discrete structure between the Southwest cluster and the Southeast.
@@
 out of the cluster with the eastern samples, 
-and at $K=4$, it splits them into their own cluster entirely \ref{bear_fastStr}.
+and at $K=4$, it splits them into their own cluster entirely (Fig \ref{bear_fastStr}).
 
@@
 Commonly used clustering methods
-must use a possibly large number of discrete clusters
-to describe this simple pattern
-such continuous population differentiation.
+must employ a possibly large number of discrete clusters
+to describe this simple pattern.
 The resulting spurious clusters can have implications for 
@@
 including conservation management decisions.
-However, discrete groupings are also often produced by important biological processes.
 
@@
 we also present a cross-validation approach 
-and a way to describe the contribution of each spatial layer to the model
-(but caution against overly strict interpretation of these).
+as well as a way to describe the contribution of each spatial layer to the model
+(but caution against overly strict interpretation of either).
 
@@
 On a desktop machine with a 4.2 GHz Intel Core i7 processor, 
-an analysis of the black bear dataset (84 samples, 21,000 loci) 
-running \texttt{conStruct}'s spatial model with 4 layers for 5,000 MCMC iterations, 
-which was more than sufficient for convergence,
+an analysis of the black bear dataset (78 samples, 21,000 loci) 
+running \texttt{conStruct}'s spatial model with 4 layers for 5,000 MCMC iterations 
+(which was more than sufficient for convergence) 
 took 2.8 hours.
@@
 and the layer-specific relatedness parameter $\phi^{(k)}$ is low, 
-individuals at a large spatial distance may be in the same layer but have very low pairwise relatedness.
+individuals separated by a large spatial distance may be in the same layer but have very low pairwise relatedness.
 It is possible that this is happening in Fig \ref{populus_sp_pies}. 
@@
 
+\clearpage
 \newpage

### 2017-09-14 writeup/figs/fig_funcs.R
colored cov by sp comparison in Fig4

@@ -68,9 +68,15 @@ viz.admix.results <- function(sim.admix.props,conStruct.results,layer.order=NULL
 			# legend=c("true admixture proportion","99% credible interval"))
 }
 
-plot.layer.curves <- function(data.block, conStruct.results, layer.cols=NULL,sample.cols=NULL,add=FALSE){
+plot.layer.curves <- function(data.block, conStruct.results, layer.cols=NULL,add=FALSE,col.mat1=NULL,col.mat2=NULL){
 	if(is.null(layer.cols)){
 		layer.cols <- c("blue","red","goldenrod1","forestgreen","darkorchid1","deepskyblue","darkorange1","seagreen2","yellow1","black")
+	}	
+	if(is.null(col.mat1)){
+		col.mat1 <- matrix(adjustcolor(1,0.7),data.block$N,data.block$N)
+	}
+	if(is.null(col.mat2)){
+		col.mat2 <- matrix(adjustcolor(1,0.7),data.block$N,data.block$N)
 	}
 	order.mat <- order(data.block$geoDist)
 	if(add==FALSE){
@@ -79,11 +85,13 @@ plot.layer.curves <- function(data.block, conStruct.results, layer.cols=NULL,sam
 	    							function(k){
 								        conStruct.results$MAP$layer.params[[k]]$layer.cov
 					})) + conStruct.results$MAP$gamma, 
-					data.block$obsCov))
+					data.block$obsCov)) + c(0,0.02)
 	    plot(data.block$geoDist[upper.tri(data.block$obsCov, diag = TRUE)], 
 	        data.block$obsCov[upper.tri(data.block$obsCov, diag = TRUE)], 
 	        xlim = range(data.block$geoDist), ylim = y.range,
-	        xlab="",ylab="",pch=19,col=adjustcolor(1,0.7))
+	        xlab="",ylab="",pch=21,cex=1.3,
+			col=adjustcolor(col.mat1[upper.tri(data.block$obsCov, diag = TRUE)],0.7),
+			bg=adjustcolor(col.mat2[upper.tri(data.block$obsCov, diag = TRUE)],0.7))
     }
     lapply(1:data.block$K, function(k) {
              lines(data.block$geoDist[order.mat],

### 2017-09-14 writeup/figs/populus_figs.R

@@ -152,13 +152,22 @@ pdf(file="~/Dropbox/conStruct/writeup/figs/populus/Fig4_pop_sp_results.pdf",widt
 			csr1.order <- 1:k
 		}
 		data.block$K <- k
-		plot.layer.curves(data.block = data.block, conStruct.results = output.list.sp[[k]][[1]], layer.cols=layer.colors[order(csr1.order)],sample.cols=NULL,add=FALSE)
+		plot.layer.curves(data.block = data.block, conStruct.results = output.list.sp[[k]][[1]], layer.cols=layer.colors[order(csr1.order)],add=FALSE,col.mat1=col.mat1,col.mat2=col.mat2)
 		box(lwd=2)
 		if(k==2){
 			mtext(side=2,text="allelic covariance",padj=-3)
+			legend(x="topright",pch=c(19,NA),lty=c(NA,1),lwd=c(NA,4),legend=c("sample covariance","layer covariance"))	
 		}
 		if(k==3){
 			mtext(side=1,text="pairwise geographic distance",padj=3)
+		}
+		if(k==4){
+			legend(x="topright",pch=21,
+					pt.bg=c(1,"forestgreen","forestgreen"),
+					col=c(1,1,"forestgreen"),
+					legend=c("balsamifera - balsamifera",
+							 "balsamifera - trichocarpa",
+							 "trichocarpa - trichocarpa"),cex=0.9,pt.cex=1.5)
 		}
 		mtext(side=1,text=bquote(paste("(",.(letters[k+2]),") ",italic("K")," = ",.(k))),padj=4.6,adj=0.4)
 	}

### 2017-09-14 writeup/cover_letter.tex

@@
 continuous and discrete patterns of differentiation. 
-We demonstrate the utility of this approach using both simulations 
-and empirical applications, with datasets from poplars and black bears. 
+We demonstrate the utility of this approach using simulations 
+and two empirical applications. %, with datasets from poplars and black bears. 
 We also release an associated R package, called \texttt{conStruct}, 
@@
 
-This work is motivated by two observations:
-first, that patterns of isolation by distance (IBD) are ubiquitous in nature; 
-and second, that existing model-based clustering methods, 
+This work is motivated by the failure of existing model-based clustering methods, 
 such as \texttt{STRUCTURE} (Pritchard \emph{et al.} 2000)
 or \texttt{ADMIXTURE} (Alexander \emph{et al.} 2009), 
-can only describe continuous patterns of variation using discrete clusters, 
-and so must use a possibly large number of clusters
-to describe continuous population differentiation.
-Inference of these spurious clusters can throw a 
+to model continuous patterns of geographical variaion (i.e., isolation by distance, or IBD).
+Such geographical patterns are ubiquitous in nature,
+but are described by these clustering methods using a possibly large number of discrete clusters.
+These clusters are spurious, and can throw a 
 wrench in the gears of efforts to describe and categorize natural variation, 
@@
 
-In this manuscript,
+Besides presenting a solution to this problem,
 we use simulations and empirical data to demonstrate 
@@
 fail in the presence of IBD; 
-we also present two different metrics for choosing an appropriate 
-number of layers to describe the data, 
-and, using both, 
-are able to establish that our method performs quite well.
+and provide two different metrics for choosing an appropriate 
+number of layers to describe the data.
 
-As of this submission, \texttt{STRUCTURE} alone has been cited more than 20,000 times, 
-and the ever-falling costs of sequencing are facilitating an explosion of next-generation sequence datasets.  
-The first step in the analysis of these data will, for most researchers, 
-involve a model-based clustering analysis.
-We therefore believe that our work represents an important and timely contribution to the field.
+As of this submission, \texttt{STRUCTURE} alone has been cited more than 20,000 times.
+However, as the costs of sequencing continue to fall
+and good geospatial information becomes more readily available,
+the limitations we cite above will become more severe.
+Our method has the potential to fill this gap.
+% We therefore believe that our work represents an important and timely contribution to the field.
 

### 2017-09-14 writeup/conStruct.tex
G's F model change

@@
 and related models \citep[e.g.][]{ADMIXTURE}; 
-the only difference is that our likelihood assumes that allele frequencies are normally distributed 
+the \gc{main} difference is that our likelihood assumes that allele frequencies are normally distributed 
 around their expectations, 
 while the standard assignment methods assume that the error is binomially distributed \citep{Engelhardt2012}.  
-(We make this approximation for the substantial advantages in computational speed.)
-Therefore, we can compare the fit of the different models 
+(We make this approximation for the substantial advantages in
+computational speed.)
+\gc{The second small difference is that unlike the original STRUCTURE where
+  allele frequencies at a locus are independently drawn for each
+  cluster \citep{STRUCTURE}, our non-spatial model of allele
+  frequencies at locus assumes that each layer has an allele frequency
+  that deviates away ( due to drift) from some shared
+  global allele frequency. Our model of allele frequencies is therefore more
+  closely related to the F-model prior for allele frequencies of \citet{falush2003}.
+As these differences are relatively small,} we can compare the fit of the different models 
 --- spatial vs. nonspatial, across different values of $K$ --- 

### 2017-09-15 writeup/cover_letter.tex
tweaked cover letter

@@
 \usepackage[green]{msuletterhead}
-%\topmarginskip{-0.35in}
-%\bottommarginskip{-1.5in}
-%\leftmarginsize{1in}
-%\rightmarginsize{1.25in}
-%\sigskipbefore{0.4in}
-%\sigskipafter{0in}
-%\noLines
-%\nolines
-%\noHeadline
-%\noheadline
-%\signature{Gideon Bradburd, Graham Coop, and Peter Ralph}
-%
-%\namefrom{}
-%\addrfrom{Gideon Bradburd \\ Dept. of Integrative Biology \\ Michigan State University \\ East Lansing, MI 48823 USA \\ 
-%\href{mailto:bradburd@msu.edu}{bradburd@msu.edu}}
-%\emailfrom{bradburd@msu.edu}
-%
-%\greetto{To the Editor(s) --}
-%\closeline{Sincerely,}
-
 \begin{document}
-%\begin{newlfm}
 
@@
 We demonstrate the utility of this approach using simulations 
-and two empirical applications. %, with datasets from poplars and black bears. 
+and two empirical applications.
 We also release an associated R package, called \texttt{conStruct}, 
@@
 the limitations we cite above will become more severe.
-Our method has the potential to fill this gap.
-% We therefore believe that our work represents an important and timely contribution to the field.
+Our method provides an important extension to \texttt{STRUCTURE} and similar approaches, 
+one that solves a critical and widely known basic conceptual drawback of the model. 
+We therefore believe that our work represents an 
+important and timely contribution to the field.
 
@@
 Peter Ralph
-%\end{newlfm}
 \end{document}
-
-%What is the scientific question you are addressing?
-%What is the key finding that answers this question?
-%What is the nature of the evidence you provide in support of your conclusion?
-%What are the three most recently published articles that are relevant to this question?
-%What significance do your results have for the field?
-%What significance do your results have for the broader community (of biologists and/or the public)?
-%What other novel findings do you present?
-%Is there additional information that we should take into account?

### 2017-09-15 writeup/conStruct.tex
edited sentence

@@
 computational speed.)
-\gc{The second small difference is that unlike the original STRUCTURE where
-  allele frequencies at a locus are independently drawn for each
-  cluster \citep{STRUCTURE}, our non-spatial model of allele
-  frequencies at locus assumes that each layer has an allele frequency
-  that deviates away ( due to drift) from some shared
-  global allele frequency. Our model of allele frequencies is therefore more
-  closely related to the F-model prior for allele frequencies of \citet{falush2003}.
+\gc{The second small difference is that in the original STRUCTURE model,
+  allele frequencies at each locus are independently drawn for each
+  cluster \citep{STRUCTURE}, while in conStruct's non-spatial model,
+  it is more natural to envision each layer's allele frequency
+  as being drifted away from a single, global allele frequency. 
+  This makes our model more
+  closely related to the ``$F$-model'' prior for allele frequencies of \citet{falush2003}.
 As these differences are relatively small,} we can compare the fit of the different models 
@@
 
-\section{Allelic Covariance}\label{allelic_cov}
+\section{Allelic covariance}\label{allelic_cov}
 To see how to arrive at our empirical variance-covariance expression, Eqn. \eqref{allelic_covariance},

### 2017-09-15 writeup/conStruct.tex
final changes

--- 2017-09-14 writeup/conStruct.tex
@@
 is quite harsh against intermediate admixture values (see Table \ref{tab:param_prior_tab}),  
-encouraging the model to ``not use" unnecessary layers if they are present, 
+encouraging the model to ``not use" unnecessary layers if they are
+present \gc{in the model'}, 
 so that they will have a low contribution to overall covariance.
@@
 To calculate layer contributions, we use the following alternative description of our covariance model:
-the genomes of any pair of individuals \emph{agree} with some background probability,
+the genomes of any pair of individuals \emph{agree} with some
+background probability \gc{at a locus},
 but this probability of agreement is increased on any segment of genome 
-that both have inherited through the same layer 
-(the amount it increases depends on how far apart they are.)
+that both have inherited \gc{from} the same layer 
+(the amount it increases depends on how far apart they are
+\gc{geographically and the decay of IBD}.)
 We use this characterization to quantify the relative contributions of each layer,

### 2017-09-15 writeup/conStruct.tex
moved over to plos template

--- 2017-09-15 writeup/conStruct.tex
@@
-\documentclass[12pt]{article}
-\usepackage{graphicx}
-\usepackage{float}
-\usepackage{hyperref}
-\usepackage{mathtools}
-\usepackage[usenames,dvipsnames]{xcolor}
-\usepackage[authoryear]{natbib}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{bigints}
+% Template for PLoS
+% Version 3.4 January 2017
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% This template contains comments intended 
+% to minimize problems and delays during our production 
+% process. Please follow the template instructions
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Once your paper is accepted for publication, 
+% PLEASE REMOVE ALL TRACKED CHANGES in this file 
+% and leave only the final text of your manuscript. 
+% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
+% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
+%
+%
+% There are no restrictions on package use within the LaTeX files except that 
+% no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% For figure citations, please use "Fig" instead of "Figure".
+% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - spacing/line breaks within cells to alter layout or alignment
+% - do not nest tabular environments (no tabular environments within tabular environments)
+% - no graphics or colored text (cell background color/shading OK)
+% See http://journals.plos.org/plosone/s/tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
+%
+% For inline equations, please be sure to include all portions of an equation in the math environment.
+For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
+%
+% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
+%
+% Please add line breaks to long display equations when possible in order to fit size of the column. 
+%
+% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
+%
+% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
+%
+% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
+%
+% % % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
+\usepackage{amsmath,amssymb}
+
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
+
+% Use Unicode characters when possible
+\usepackage[utf8x]{inputenc}
+
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% color can be used to apply background shading to table cells only
+\usepackage[table]{xcolor}
+
+% array package and thick rules for tables
 \usepackage{array}
-\usepackage{tikz}
-\usepackage{longtable}
-\usepackage{geometry}
-
-\bibliographystyle{plainnat}
-
-\setlength{\parskip}{\baselineskip}
-\setlength{\parindent}{0pt}
+
+% create "+" rule type for thick vertical lines
+\newcolumntype{+}{!{\vrule width 2pt}}
+
+% create \thickcline for thick horizontal lines of variable length
+\newlength\savedwidth
+\newcommand\thickcline[1]{%
+  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+  \cline{#1}%
+  \noalign{\vskip\arrayrulewidth}%
+  \noalign{\global\arrayrulewidth\savedwidth}%
+}
+
+% \thickhline command for thick horizontal lines that span the table
+\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+\hline
+\noalign{\global\arrayrulewidth\savedwidth}}
+
+% Remove comment for double spacing
+%\usepackage{setspace} 
+%\doublespacing
+
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in 
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+\renewcommand{\figurename}{Fig}
+
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2015}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Leave date blank
+\date{}
+
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+\setlength{\headheight}{27.023pt}
+\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\sf PLOS}
+
+%% Include all macros below
 
@@
 
-\geometry{a4paper}
-
-\title{Inferring Continuous and Discrete Population Genetic Structure Across Space}
-\date{\vspace{-5ex}}
-\author{
-Gideon S.\ Bradburd$^{1,a}$, 
-Graham M.\ Coop$^{2,b}$,
-Peter L.\ Ralph$^{3,c}$}
+%% END MACROS SECTION
 
 \begin{document}
-
-\maketitle
-
-\textsuperscript{1}
+\vspace*{0.2in}
+
+% Title must be 250 characters or less.
+\begin{flushleft}
+{\Large
+\textbf\newline{Inferring continuous and discrete population genetic structure across space}
+}
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+Gideon S.\  Bradburd\textsuperscript{1,a},
+Graham M.\  Coop\textsuperscript{2,b,\Yinyang},
+Peter L.\  Ralph\textsuperscript{3,c,\Yinyang},
+\\
+\bigskip
+\textbf{1}
 Department of Integrative Biology, 
@@
 Michigan State University, MI 48824
-
-\textsuperscript{2}
+\\
+\textbf{2}
 Center for Population Biology,
@@
 University of California, Davis, CA 95616
-
-\textsuperscript{3}
+\\
+\textbf{3} 
 Institute of Ecology and Evolution,
-Department of Mathematics,
+Departments of Mathematics and Biology,
 University of Oregon, Eugene, OR 97403
-
-\textsuperscript{a}bradburd@msu.edu; 
-\textsuperscript{b}gmcoop@ucdavis.edu;
-\textsuperscript{c}plr@uoregon.edu\\\\\
-
-\newpage
-
-\begin{abstract}
+\\
+\textbf{a} bradburd@msu.edu; \\
+\textbf{b} gmcoop@ucdavis.edu;\\
+\textbf{c} plr@uoregon.edu\\
+
+\bigskip
+
+%
+% Primary Equal Contribution Note
+\Yinyang These authors contributed equally to this work.
+
+\end{flushleft}
+% Please keep the abstract below 300 words
+\section*{Abstract}
+
 A classic problem in population genetics is the characterization 
@@
 and by applying it to empirical datasets of poplars and black bears in North America.
-\end{abstract}
+
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step. 
+% Author Summary not valid for PLOS ONE submissions.   
+\section*{Author summary}
+\plr{WRITE AUTHOR SUMMARY}
+
+\linenumbers
 
@@
 
-\section*{Methods}
+% \section*{Methods}
+\section*{Results}
 
@@
 (b) we randomly choose an allele to count at each locus.
-We discuss the derivation of Eqn. \eqref{allelic_covariance} further in the Appendix (\secref{allelic_cov}).
+We discuss the derivation of Eq.\ \eqref{allelic_covariance} further in \secref{allelic_cov}.
 As noted in \cite{EEMS},
@@
 we are modeling the allelic, rather than genetic, covariance. 
-In the Appendix section \ref{rationale} we lay out a
+In \ref{rationale} we lay out a
 simple model of allele frequencies underlying this covariance model.
@@
 
-A nice feature of our approach is that the model described in Eqn \eqref{cross_layer_covariance} 
+A nice feature of our approach is that the model described in Eq.\ \eqref{cross_layer_covariance} 
 contains a nonspatial assignment model as a special case 
-(see \secref{model_app} in the Appendix for a more in-depth discussion). 
+(see \secref{model_app} for a more in-depth discussion). 
 By setting $\alpha_0^{(k)}$ to zero for all $k$, 
@@
 The best model is judged to be the simplest one with significantly better predictive accuracy
-than others (see \secref{Xvalidation} in the Appendix for more on our cross-validation procedure).
+than others (see \secref{Xvalidation} for more on our cross-validation procedure).
 In general, larger values of $K$ allows the model more flexibility,
@@
 by computing the average contribution to increased probability of agreement
-as described in Appendix \secref{layer_contribution}). 
+as described in \secref{layer_contribution}). 
 This layer contribution is similar to the ``ancestry contribution'' proposed by \cite{fastStructure}. 
@@
 
-\section*{Results}
+% \section*{Results}
 
@@
 We collapsed the layers together using random admixture with spatially autocorrelated proportions 
-(see Appendix \secref{sim_details} for details), 
+(see \secref{sim_details} for details), 
 so that both the frequencies in layers and the layer memberships are
@@
 
-\section*{Acknowledgements}
-
-We thank Marjorie Weber, Yaniv Brandvain, William Wetzel, 
-Mariah Meek, Doc Edge, and Matthew Stephens 
-for invaluable comments on the method and manuscript, 
-as well as Quentin Cronk, who provided input on the \textit{Populus} analyses, 
-and Emily Puckett, who provided input on the black bear analyses.
-We also thank the attendees at the 2017 SSE Meeting in Portland, OR, 
-whose votes determined the name of the method.
-This work was supported in part by 
-the National Science Foundation under award number NSF \#1262645 (DBI) to PR and GC, 
-the National Institute of General Medical Sciences of the National
-Institutes of Health under award numbers NIH R01-GM108779 to GC,
-and the National Science Foundation under award numbers NSF \#1148897 and \#1402725 to GB.
-
 \newpage
-\section*{Appendix}
-\pagenumbering{gobble}
-\renewcommand{\theequation}{A\arabic{equation}}
-\setcounter{equation}{0}
-\renewcommand{\thetable}{A\arabic{table}}
-\setcounter{table}{0}
-\renewcommand{\thefigure}{A\arabic{figure}}
-\setcounter{figure}{0}
-\renewcommand{\thesection}{A\arabic{section}}
-\setcounter{section}{0}
-
-\section{Model rationale}
-\subsection{Drift, admixture, and space} \label{rationale}
+\section*{Materials and Methods}
+
+% \section*{Appendix}
+% \pagenumbering{gobble}
+% \renewcommand{\theequation}{A\arabic{equation}}
+% \setcounter{equation}{0}
+% \renewcommand{\thetable}{A\arabic{table}}
+% \setcounter{table}{0}
+% \renewcommand{\thefigure}{A\arabic{figure}}
+% \setcounter{figure}{0}
+% \renewcommand{\thesection}{A\arabic{section}}
+% \setcounter{section}{0}
+
+\section*{Model rationale}
+\subsection*{Drift, admixture, and space} \label{rationale}
 Here we sketch a simple model of allele frequencies and their
@@
 where the quantity $\delta_{i=j}$ is an indicator variable that equals 1 when $i$ is equal to $j$ and 0 otherwise, 
-as in Eqn. \eqref{cross_layer_covariance}.
+as in Eq.\ \eqref{cross_layer_covariance}.
 
@@
 
-\section{Allelic covariance}\label{allelic_cov}
-To see how to arrive at our empirical variance-covariance expression, Eqn. \eqref{allelic_covariance},
+\section*{Allelic covariance}\label{allelic_cov}
+To see how to arrive at our empirical variance-covariance expression, Eq.\ \eqref{allelic_covariance},
 pick a random locus and
@@
 and using the fact that $\E[X] = \E[Y] = 1/2$,
-Eqn. \eqref{allelic_covariance} comes from the calculation that
+Eq.\ \eqref{allelic_covariance} comes from the calculation that
 \begin{align}
@@
 
-\section{Models, parameters, and priors}\label{model_app}
+\section*{Models, parameters, and priors}\label{model_app}
 \paragraph{Spatial vs. nonspatial}
@@
 each of which can be implemented with different numbers of layers/clusters.
-The spatial model is parameterized as in Eqn \eqref{admixed_spatial_cov},
-and the nonspatial model is described in Eqn \eqref{admixed_discrete_covariance}.
+The spatial model is parameterized as in Eq.\ \eqref{admixed_spatial_cov},
+and the nonspatial model is described in Eq.\ \eqref{admixed_discrete_covariance}.
 The nonspatial model is a special case of the spatial model, 
@@
 \newpage
-\section{Cross validation}\label{Xvalidation}
+\section*{Cross validation}\label{Xvalidation}
 
@@
 \newpage
-\section{Calculating layer contributions}\label{layer_contribution}
+\section*{Calculating layer contributions}\label{layer_contribution}
 
@@
 \newpage
-\section{Simulation details}\label{sim_details}
+\section*{Simulation details}\label{sim_details}
 We wished to simulate data under a model that had some biological realism, 
@@
 \newpage
-\section*{Supplementary Materials}
+\section*{Supplementary Information}
 \renewcommand{\theequation}{S\arabic{equation}}
@@
 
+\section*{Acknowledgements}
+
+We thank Marjorie Weber, Yaniv Brandvain, William Wetzel, 
+Mariah Meek, Doc Edge, and Matthew Stephens 
+for invaluable comments on the method and manuscript, 
+as well as Quentin Cronk, who provided input on the \textit{Populus} analyses, 
+and Emily Puckett, who provided input on the black bear analyses.
+We also thank the attendees at the 2017 SSE Meeting in Portland, OR, 
+whose votes determined the name of the method.
+This work was supported in part by 
+the National Science Foundation under award number NSF \#1262645 (DBI) to PR and GC, 
+the National Institute of General Medical Sciences of the National
+Institutes of Health under award numbers NIH R01-GM108779 to GC,
+and the National Science Foundation under award numbers NSF \#1148897 and \#1402725 to GB.
+
 \newpage
@@
 
-\newpage
-
 \end{document}

### 2017-09-15 writeup/conStruct.tex
penultimate edits

--- 2017-09-15 writeup/conStruct.tex
@@
 Gideon S.\ Bradburd$^{1,a}$, 
-Graham M.\ Coop$^{2,b}$,
-Peter L.\ Ralph$^{3,c}$}
+Graham M.\ Coop$^{2,b,}$*,
+Peter L.\ Ralph$^{3,c,}$*}
 
@@
 \textsuperscript{b}gmcoop@ucdavis.edu;
-\textsuperscript{c}plr@uoregon.edu\\\\\
-
+\textsuperscript{c}plr@uoregon.edu
+
+*these authors contributed equally to this project
 \newpage
@@
 within each of which relatedness decays as a parametric function of the distance between samples.
-We also implement a cross-validation approach for comparing and selecting models across different numbers of layers,
+We also implement a cross-validation approach for comparing and selecting models across different numbers of layers/clusters,
 and we demonstrate the utility of our approach using both simulated and empirical data.
@@
 			     so the small dashed circles on each layer show where 
-			     the location of the highlighted, admixed sample intersects each layer.
+			     the location of the highlighted admixed sample intersects each layer.
 			    }\label{schematic}
@@
 and related models \citep[e.g.][]{ADMIXTURE}; 
-the \gc{main} difference is that our likelihood assumes that allele frequencies are normally distributed 
+the main difference is that our likelihood assumes that allele frequencies are normally distributed 
 around their expectations, 
@@
 computational speed.)
-\gc{The second small difference is that unlike the original STRUCTURE where
-  allele frequencies at a locus are independently drawn for each
-  cluster \citep{STRUCTURE}, our non-spatial model of allele
-  frequencies at locus assumes that each layer has an allele frequency
-  that deviates away ( due to drift) from some shared
-  global allele frequency. Our model of allele frequencies is therefore more
-  closely related to the F-model prior for allele frequencies of \citet{falush2003}.
-As these differences are relatively small,} we can compare the fit of the different models 
+The second small difference is that, unlike the original STRUCTURE model,
+in which allele frequencies at a locus are independently drawn for each cluster \citep{STRUCTURE}, 
+our non-spatial model of allele frequencies assumes 
+that each cluster has an allele frequency that deviates 
+(due to drift) away from some shared global allele frequency. 
+This model of allele frequencies is therefore more 
+closely related to STRUCTURE's F-model prior for allele frequencies of \citet{falush2003}.
+As these differences are relatively small, 
+we can compare the fit of the different models 
 --- spatial vs. nonspatial, across different values of $K$ --- 
@@
 discussed further in the Discussion.
-However, it is still important to assess whether additional spatial layers (larger $K$)
+However, it is still important to assess whether additional layers (larger $K$)
 meaningfully model patterns in the data
@@
 is quite harsh against intermediate admixture values (see Table \ref{tab:param_prior_tab}),  
-encouraging the model to ``not use" unnecessary layers if they are
-present \gc{in the model'}, 
+encouraging the model to ``not use" unnecessary layers if they are present, 
 so that they will have a low contribution to overall covariance.
@@
 To calculate layer contributions, we use the following alternative description of our covariance model:
-the genomes of any pair of individuals \emph{agree} with some
-background probability \gc{at a locus},
+the genomes of any pair of individuals \emph{agree} with some background probability,
 but this probability of agreement is increased on any segment of genome 
-that both have inherited \gc{from} the same layer 
-(the amount it increases depends on how far apart they are
-\gc{geographically and the decay of IBD}.)
+that both have inherited through the same layer 
+(the amount it increases depends on how far apart they are.)
 We use this characterization to quantify the relative contributions of each layer,
@@
 With these simulations,
-spatial \texttt{conStruct} does not create spurious clusters when there are none:
+spatial \texttt{conStruct} does not create spurious discrete groupings when there are none:
 Figures \ref{simK1_pies}, \ref{simK2_sp_pies}, and \ref{simK3_sp_pies} 
@@
 		\caption{
-			Layer contributions (i.e., how much total covariance is contributed by each layer), 
+			Layer/cluster contributions (i.e., how much total covariance is contributed by each layer/cluster), 
 			for all layers estimated in runs using $K = 1$ through 7 
-			for the spatial model (left) and the nonspatial model (right).
+			for the spatial model (left),
+			and for all clusters using the nonspatial model (right).
 			Data were simulated using $K=1$.
-			For each value of $K$ along the x-axis, there are an equal number of layer contributions plotted.
+			For each value of $K$ along the x-axis, there are an equal number of contributions plotted.
 		}\label{simK1_laycon}
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/populus/Fig4_pop_sp_results.pdf}}
+		{\includegraphics[width=6in,height=4in]{figs/populus/pop_sp_results.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=3in]{figs/bears/Fig6_sp_vs_nsp.pdf}}
+		{\includegraphics[width=6in,height=3in]{figs/bears/bears_sp_vs_nsp.pdf}}
 	\caption{
@@
 	\caption{
-	Layer contributions (i.e., how much total covariance is contributed by each layer), 
+	Layer/cluster contributions (i.e., how much total covariance is contributed by each layer/cluster), 
 	for all layers estimated in runs using $K = 1$ through 7 
-	for the spatial model (left) and the nonspatial model (right).
-	For each value of $K$ along the x-axis, there are an equal number of layer contributions plotted.
+	for the spatial model (left),
+	and for all clusters using the nonspatial model (right).
+	For each value of $K$ along the x-axis, there are an equal number of contributions plotted.
 	Colors are consistent with Fig \ref{bear_K3}.
@@
 Although we believe our model adds spatial realism to the groups used by clustering methods,
-it is important to note that the clusters/layers detected by our method 
+it is important to note that the layers detected by our method 
 do not necessarily correspond  to distinct, ancestral populations; 
@@
 need not be strongly differentiated --
-cluster assignment is not a replacement for absolute measures of genetic divergence.
+pouplation assignment is not a replacement for absolute measures of genetic divergence.
 
@@
 				\right) ,
-\label{layer_contribution}
+\label{abs_layer_contribution}
 \end{equation}
@@
 		\caption{
-			Layer contributions (i.e., how much total covariance is contributed by each layer), 
+			Layer/cluster contributions (i.e., how much total covariance is contributed by each layer/cluster), 
 			for all layers estimated in runs using $K = 1$ through 7 
-			for the spatial model (left) and the nonspatial model (right).
+			for the spatial model (left), 
+			and for all clusters using the nonspatial model (right).
 			Data were simulated using $K=2$.
-			For each value of $K$ along the x-axis, there are an equal number of layer contributions plotted.
+			For each value of $K$ along the x-axis, there are an equal number of contributions plotted.
 		}\label{simK2_laycon}
@@
 		\caption{
-			Layer contributions (i.e., how much total covariance is contributed by each layer), 
+			Layer/cluster contributions (i.e., how much total covariance is contributed by each layer/cluster), 
 			for all layers estimated in runs using $K = 1$ through 7 
-			for the spatial model (left) and the nonspatial model (right).
+			for the spatial model (left), 
+			and for all clusters using the nonspatial model (right).
 			Data were simulated using $K=3$.
-			For each value of $K$ along the x-axis, there are an equal number of layer contributions plotted.
+			For each value of $K$ along the x-axis, there are an equal number of contributions plotted.
 		}\label{simK3_laycon}
@@
 	\caption{
-	Plots showing the layer-specific parametric covariances 
+	Plots showing the cluster-specific parametric covariances 
 	estimated for the \textit{Populus} data using 
 	the nonspatial \texttt{conStruct} model run with $K=1$ through 6.
-	Line colors are consistent with layer colors in Fig \ref{populus_nsp_pies}.
+	Line colors are consistent with cluster colors in Fig \ref{populus_nsp_pies}.
 	Points are colored by the species they are a covariance between:
@@
 	\caption{
-	Layer contributions (i.e., how much total covariance is contributed by each layer), 
+	Layer/cluster contributions (i.e., how much total covariance is contributed by each layer/cluster), 
 	for all layers estimated in runs using $K = 1$ through 7 
-	for the spatial model (left) and the nonspatial model (right).
-	For each value of $K$ along the x-axis, there are an equal number of layer contributions plotted.
+	for the spatial model (left), 
+	and for all clusters using the nonspatial model (right).
+	For each value of $K$ along the x-axis, there are an equal number of contributions plotted.
 	Colors are consistent with Figs \ref{populus_sp_pies}, \ref{populus_sp_layer_covs}, \ref{populus_nsp_pies}, and \ref{populus_nsp_layer_covs}.
@@
 	\caption{
-	Plots showing the layer-specific parametric covariances 
+	Plots showing the cluster-specific parametric covariances 
 	estimated for the black bear data using 

### 2017-09-15 writeup/figs/bear_figs.R

@@ -106,7 +106,7 @@ for(k in 2:7){
 }
 }
 
-pdf(file="~/Dropbox/conStruct/writeup/figs/bears/Fig6_sp_vs_nsp.pdf",width=15,height=7.5,pointsize=14)
+pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bears_sp_vs_nsp.pdf",width=15,height=7.5,pointsize=14)
 	layout(cbind(matrix(c(rep(1,10),rep(2,15)),nrow=5,ncol=5,byrow=TRUE),
 				 matrix(2+c(rep(1,10),rep(2,15)),nrow=5,ncol=5,byrow=TRUE)))
 	par(mar=c(5,5,1,1))

### 2017-09-15 writeup/figs/populus_figs.R

@@ -119,7 +119,7 @@ pdf(file=paste0("~/Dropbox/conStruct/writeup/figs/populus/populus_sp_layer_covs.
 	mtext(text="allele frequency covariance",side=2,font=2,cex.axis=2,padj=-59.5,adj=20)
 dev.off()
 
-pdf(file="~/Dropbox/conStruct/writeup/figs/populus/Fig4_pop_sp_results.pdf",width=12,height=8,pointsize=14)
+pdf(file="~/Dropbox/conStruct/writeup/figs/populus/pop_sp_results.pdf",width=12,height=8,pointsize=14)
 	#quartz(width=12,height=8,pointsize=14)
 	layout(matrix(c(1:6),nrow=2,ncol=3,byrow=TRUE))
 	csr1.order <- NULL

### 2017-09-15 writeup/conStruct.tex

@@
 computational speed.)
-The second small difference is that, unlike the original STRUCTURE model,
-in which allele frequencies at a locus are independently drawn for each cluster \citep{STRUCTURE}, 
-our non-spatial model of allele frequencies assumes 
-that each cluster has an allele frequency that deviates 
-(due to drift) away from some shared global allele frequency. 
-This model of allele frequencies is therefore more 
-closely related to STRUCTURE's F-model prior for allele frequencies of \citet{falush2003}.
-As these differences are relatively small, 
-we can compare the fit of the different models 
+The second small difference is that, in the original STRUCTURE model, 
+allele frequencies at each locus are independently drawn for each 
+cluster \citep{STRUCTURE}, while in conStruct's non-spatial model, 
+it is more natural to envision each cluster's allele frequency 
+as being drifted away from a single, global allele frequency. 
+This makes our model more
+closely related to the ``$F$-model'' prior for allele frequencies of \citet{falush2003}.
+As these differences are relatively small, we can compare the fit of the different models 
 --- spatial vs. nonspatial, across different values of $K$ --- 
@@
 is quite harsh against intermediate admixture values (see Table \ref{tab:param_prior_tab}),  
-encouraging the model to ``not use" unnecessary layers if they are present, 
+encouraging the model to ``not use" unnecessary layers if they are
+present in the model, 
 so that they will have a low contribution to overall covariance.
@@
 To calculate layer contributions, we use the following alternative description of our covariance model:
-the genomes of any pair of individuals \emph{agree} with some background probability,
+the genomes of any pair of individuals \emph{agree} with some
+background probability at a locus,
 but this probability of agreement is increased on any segment of genome 
-that both have inherited through the same layer 
-(the amount it increases depends on how far apart they are.)
+that both have inherited from the same layer 
+(the amount it increases depends on how far apart they are
+geographically and on the decay of IBD).
 We use this characterization to quantify the relative contributions of each layer,

### 2017-09-15 writeup/latex/plos_latex_template.tex

% Template for PLoS
% Version 3.4 January 2017
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{27.023pt}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Title of submission to PLOS journals} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Name1 Surname\textsuperscript{1,2\Yinyang},
Name2 Surname\textsuperscript{2\Yinyang},
Name3 Surname\textsuperscript{2,3\textcurrency},
Name4 Surname\textsuperscript{2},
Name5 Surname\textsuperscript{2\ddag},
Name6 Surname\textsuperscript{2\ddag},
Name7 Surname\textsuperscript{1,2,3*},
with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
\\
\bigskip
\textbf{1} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
\Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
\ddag These authors also contributed equally to this work.

% Current address notes
\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
% \textcurrency b Insert second current address 
% \textcurrency c Insert third current address

% Deceased author note
\dag Deceased

% Group/Consortium Author Note
\textpilcrow Membership list can be found in the Acknowledgments section.

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* correspondingauthor@institute.edu

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author summary}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}
Lorem ipsum dolor sit~\cite{bib1} amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam Eq~(\ref{eq:schemeP}) sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id.~\cite{bib2} Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

\begin{eqnarray}
\label{eq:schemeP}
	\mathrm{P_Y} = \underbrace{H(Y_n) - H(Y_n|\mathbf{V}^{Y}_{n})}_{S_Y} + \underbrace{H(Y_n|\mathbf{V}^{Y}_{n})- H(Y_n|\mathbf{V}^{X,Y}_{n})}_{T_{X\rightarrow Y}},
\end{eqnarray}

\section*{Materials and methods}
\subsection*{Etiam eget sapien nibh}

% For figure citations, please use "Fig" instead of "Figure".
Nulla mi mi, Fig~\ref{fig1} venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, \nameref{S1_Video} vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

% Place figure captions after the first paragraph in which they are cited.
\begin{figure}[!h]
\caption{{\bf Bold the figure title.}
Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
\label{fig1}
\end{figure}

% Results and Discussion can be combined.
\section*{Results}
Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

% Place tables after the first paragraph in which they are cited.
\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{
{\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
\begin{tabular}{|l+l|l|l|l|l|l|l|}
\hline
\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \thickhline
$cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
$cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
$cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
\end{tabular}
\begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
\end{flushleft}
\label{table1}
\end{adjustwidth}
\end{table}

%PLOS does not support heading levels beyond the 3rd (no 4th level headings).
\subsection*{\lorem\ and \ipsum\ nunc blandit a tortor}
\subsubsection*{3rd level heading} 
Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit. Sed non aliquet felis. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris commodo justo ac dui pretium imperdiet. Sed suscipit iaculis mi at feugiat.

\begin{enumerate}
	\item{react}
	\item{diffuse free particles}
	\item{increment time by dt and go to 1}
\end{enumerate}

\subsection*{Sed ac quam id nisi malesuada congue}

Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

\begin{itemize}
	\item First bulleted item.
	\item Second bulleted item.
	\item Third bulleted item.
\end{itemize}

\section*{Discussion}
Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero~\cite{bib3}.

\section*{Conclusion}

CO\textsubscript{2} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.

Sed non aliquet felis. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris commodo justo ac dui pretium imperdiet. Sed suscipit iaculis mi at feugiat. Ut neque ipsum, luctus id lacus ut, laoreet scelerisque urna. Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed. Nam condimentum sem eget mollis euismod. Nullam dui urna, gravida venenatis dui et, tincidunt sodales ex. Nunc est dui, sodales sed mauris nec, auctor sagittis leo. Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor. For more information, see \nameref{S1_Appendix}.

\section*{Supporting information}

% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
\paragraph*{S1 Fig.}
\label{S1_Fig}
{\bf Bold the title sentence.} Add descriptive text after the title of the item (optional).

\paragraph*{S2 Fig.}
\label{S2_Fig}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 File.}
\label{S1_File}
{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Video.}
\label{S1_Video}
{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Appendix.}
\label{S1_Appendix}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Table.}
\label{S1_Table}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\section*{Acknowledgments}
Cras egestas velit mauris, eu mollis turpis pellentesque sit amet. Interdum et malesuada fames ac ante ipsum primis in faucibus. Nam id pretium nisi. Sed ac quam id nisi malesuada congue. Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.

\nolinenumbers

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% or
%
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here. See http://journals.plos.org/plosone/s/latex for 
% step-by-step instructions.
% 
\begin{thebibliography}{10}

\bibitem{bib1}
Conant GC, Wolfe KH.
\newblock {{T}urning a hobby into a job: how duplicated genes find new
  functions}.
\newblock Nat Rev Genet. 2008 Dec;9(12):938--950.

\bibitem{bib2}
Ohno S.
\newblock Evolution by gene duplication.
\newblock London: George Alien \& Unwin Ltd. Berlin, Heidelberg and New York:
  Springer-Verlag.; 1970.

\bibitem{bib3}
Magwire MM, Bayer F, Webster CL, Cao C, Jiggins FM.
\newblock {{S}uccessive increases in the resistance of {D}rosophila to viral
  infection through a transposon insertion followed by a {D}uplication}.
\newblock PLoS Genet. 2011 Oct;7(10):e1002337.

\end{thebibliography}

\end{document}

### 2017-09-15 writeup/conStruct.tex
tidbits

--- 2017-09-15 writeup/conStruct.tex
@@
 % Use the PLoS provided BiBTeX style
-\bibliographystyle{plos2015}
+\bibliographystyle{latex/plos2015}
 
@@
 \setlength{\headheight}{27.023pt}
-\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\lhead{\includegraphics[width=2.0in]{latex/PLOS-submission.eps}}
 \rfoot{\thepage/\pageref{LastPage}}
@@
 
-\include{macros}
-
 \newcommand{\gb}[1]{{\it\color{magenta}{(#1)}}}
@@
 \newcommand{\gc}[1]{{\it\color{blue}{(#1)}}}
+
+\newcommand{\R}{\mathbb{R}}
+\renewcommand{\P}{\mathbb{P}}
+\newcommand{\E}{\mathbb{E}}
+\newcommand{\var}{\mathop{\mbox{Var}}}
+\newcommand{\cov}{\mathop{\mbox{cov}}}
+\newcommand{\supp}{\mathop{\mbox{supp}}}
+\newcommand{\sgn}{\mathop{\mbox{sgn}}}
+\newcommand{\bone}{\mathbf{1}}
 
@@
 processes, we often need to identify somewhat separable populations 
-from which individuals are sampled \citep{wright1949genetical}.
+from which individuals are sampled \cite{wright1949genetical}.
 Delineating populations is useful for systematics and for
-informing conservation priorities \citep{Moritz1994,Waples_1998,Moritz_etal_2002}.
+informing conservation priorities \cite{Moritz1994,Waples_1998,Moritz_etal_2002}.
 Furthermore, we often need to identify subsets of indivduals resulting from reasonably coherent
@@
 genetic structure,
-including generating population phylogenies \citep{CavalliSforza1975, treemix},
-dimensionality-reduction approaches \citep{meirmans2009genodive}, 
+including generating population phylogenies \cite{CavalliSforza1975, treemix},
+dimensionality-reduction approaches \cite{meirmans2009genodive}, 
 such as principal components analysis 
-\citep{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
+\cite{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
 and model-based clustering approaches 
-\citep[e.g.][]{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, 
+(e.g., \cite{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, 
 FINESTRUCTURE, fastStructure, huelsenbeck2007inference, 
-Corander2003,TESS,geneland}, among others.
+Corander2003,TESS,geneland}), among others.
 Each of these methods perform best in particular situations,
@@
 as that produced by geographic isolation by distance 
-\citep{Wright1943, novembre_interpreting_2008, Frantz2009}.
+\cite{Wright1943, novembre_interpreting_2008, Frantz2009}.
 Here, we will focus on model-based clustering, the most widely used class of approaches for population delineation.
@@
 Spatial information has been incorporated into some of these methods, 
-by, for example, placing spatial priors on cluster memberships \citep{geneland,TESS}, 
+by, for example, placing spatial priors on cluster memberships \cite{geneland,TESS}, 
 but this does not address the underlying issue that 
@@
 genetic drift to build up differentiation between distant locations
-\citep{Wright1943}. 
+\cite{Wright1943}. 
 Theoretical work,
 mostly derived from ``stepping-stone'' models 
-\citep{kimura1964stepping,sawyer1976stepping,shiga1984stepping},
+\cite{kimura1964stepping,sawyer1976stepping,shiga1984stepping},
 gives us some analytical predictions for isolation by distance
-\citep{malecot1969mathematics,Slatkin1985,epperson2003geographical}, 
-but substantial work remains to be done \citep{barton2002neutral,barton2013modelling}.
+\cite{malecot1969mathematics,Slatkin1985,epperson2003geographical}, 
+but substantial work remains to be done \cite{barton2002neutral,barton2013modelling}.
 % IBD has been modeled in both spatially continuous frameworks 
 % in which an individual's parents are found within some radius of that individual's location,
-% and discrete, ``stepping-stone'' frameworks \citep{kimura1964stepping},
+% and discrete, ``stepping-stone'' frameworks \cite{kimura1964stepping},
 % in which migration occurs only between neighboring demes.
 Given the generality of the circumstances that generate a pattern of isolation by distance, 
-it is unsurprising that IBD is very widespread in nature \citep{meirmans2012,Sexton_etal_2014}.
+it is unsurprising that IBD is very widespread in nature \cite{meirmans2012,Sexton_etal_2014}.
 
@@
 about how best to describe sets of individuals using continuous clines and discrete clusters 
-\citep[e.g.][]{SerrePaabo2004,rosenberg2005clines}.
+(e.g., \cite{SerrePaabo2004,rosenberg2005clines}).
 
@@
 discrete clusters, and so tend to erroneously describe continuously distributed variation with multiple clusters that 
-show spatially autocorrelated cluster membership \citep{Frantz2009,meirmans2012}.
+show spatially autocorrelated cluster membership \cite{Frantz2009,meirmans2012}.
 In analyses of empirical datasets, which often are characterized by IBD,
@@
 
-This definition of covariance differs from the usual ``genetic covariance'' \citep{mcvean_genealogical_2009}
+This definition of covariance differs from the usual ``genetic covariance'' \cite{mcvean_genealogical_2009}
 in that (a) we do not subtract locus means,
@@
 as parameterized by the $\alpha$s, 
-is a flexible and standard choice in spatial statistics \citep{Diggle1998}, 
+is a flexible and standard choice in spatial statistics \cite{Diggle1998}, 
 and is not chosen to match a particular population genetics model. 
-% (Analytical theory \citep{barton2002neutral} predicts the decay should be proportional to $\exp{-D}/\sqrt{D}$.)
+% (Analytical theory \cite{barton2002neutral} predicts the decay should be proportional to $\exp{-D}/\sqrt{D}$.)
 The $\phi^{(k)}$ is a parameter that describes the background covariance within the layer. 
@@
 the branch length connecting the $k$th population to the population ancestral to all modeled
-layers \citep[see for example ][]{patterson_ancient_2012,peter_fstats},
+layers (see for example \cite{patterson_ancient_2012,peter_fstats}),
 although they cannot be directly compared because 
@@
 by introducing a parameter to modify the given degrees of freedom 
-and thereby informally model linkage between loci \citep[e.g.][]{EEMS}.
+and thereby informally model linkage between loci (e.g., \cite{EEMS}).
 
@@
 Parameters are independent between layers.
-We use Hamiltonian Monte Carlo as implemented in STAN \citep{stan, NUTS, stan_lib, rstan} 
+We use Hamiltonian Monte Carlo as implemented in STAN \cite{stan, NUTS, stan_lib, rstan} 
 to estimate the posterior distribution on the parameters.
-We also provide an R package \citep{R} called \texttt{conStruct} 
+We also provide an R package \cite{R} called \texttt{conStruct} 
 (for ``continuous structure") that functions as a wrapper around this inference machinery.
@@
 and individuals draw a proportion of their ancestry from each cluster. 
-This model is very similar to that of STRUCTURE \citep{STRUCTURE} 
-and related models \citep[e.g.][]{ADMIXTURE}; 
+This model is very similar to that of STRUCTURE \cite{STRUCTURE} 
+and related models (e.g., \cite{ADMIXTURE}); 
 the \gc{main} difference is that our likelihood assumes that allele frequencies are normally distributed 
 around their expectations, 
-while the standard assignment methods assume that the error is binomially distributed \citep{Engelhardt2012}.  
+while the standard assignment methods assume that the error is binomially distributed \cite{Engelhardt2012}.  
 (We make this approximation for the substantial advantages in
@@
   allele frequencies at each locus are independently drawn for each
-  cluster \citep{STRUCTURE}, while in conStruct's non-spatial model,
+  cluster \cite{STRUCTURE}, while in conStruct's non-spatial model,
   it is more natural to envision each layer's allele frequency
@@
   This makes our model more
-  closely related to the ``$F$-model'' prior for allele frequencies of \citet{falush2003}.
+  closely related to the ``$F$-model'' prior for allele frequencies \cite{falush2003}.
 As these differences are relatively small,} we can compare the fit of the different models 
@@
 for the number of discrete clusters in population genetic data, including: 
-information theoretic criteria \citep[e.g.][]{ADMIXTURE}, such as AIC or BIC;
+information theoretic criteria (e.g., \cite{ADMIXTURE}), such as AIC or BIC;
 comparisons of the likelihood of the data across different values of $K$, 
-with various criteria on how to choose a single value (e.g., \citep{Evanno2005});
+with various criteria on how to choose a single value (e.g., \cite{Evanno2005});
 or comparisons of the marginal likelihood, 
-generated either via various approximations \citep[e.g.][]{STRUCTURE}
-or via a more rigorous and computationally intensive approach such as thermodynamic integration \citep{verity_nichols2016}
-or inference using a Dirichlet process prior \citep{huelsenbeck2007inference}.
+generated either via various approximations (e.g., \cite{STRUCTURE})
+or via a more rigorous and computationally intensive approach such as thermodynamic integration \cite{verity_nichols2016}
+or inference using a Dirichlet process prior \cite{huelsenbeck2007inference}.
 See \cite{verity_nichols2016} for a discussion of these approaches and comparison
@@
 
-We use cross-validation \citep[similar in spirit to][]{ADMIXTURE_xval} to attack this problem.
+We use cross-validation (similar in spirit to \cite{ADMIXTURE_xval}) to attack this problem.
 To do this,
@@
 
-To test the method, we first generated data using the coalescent simulator \texttt{ms} \citep{Hudson2002}.
+To test the method, we first generated data using the coalescent simulator \texttt{ms} \cite{Hudson2002}.
 In each simulation,
@@
 For comparison,
-we also analyzed each simulated dataset using fastSTRUCTURE \citep{fastStructure}, using $K=2-4$.
+we also analyzed each simulated dataset using fastSTRUCTURE \cite{fastStructure}, using $K=2-4$.
 % inferring admixture under models with $K=2$ through 4 (Figs \ref{fastStr_simK1}--\ref{fastStr_simK3}).
@@
 % other nonspatial model-based clustering algorithms, 
-% we also analyzed each simulated dataset using fastSTRUCTURE \citep{fastStructure}, 
+% we also analyzed each simulated dataset using fastSTRUCTURE \cite{fastStructure}, 
 % inferring admixture under models with $K=2$ through 4 (Figs \ref{fastStr_simK1}--\ref{fastStr_simK3}).
@@
 species in the genus regularly co-occur and, 
-where they do, they frequently hybridize \citep{eckenwalder1984, Cronk2005}.
+where they do, they frequently hybridize \cite{eckenwalder1984, Cronk2005}.
 
@@
 have a broad zone of overlap in the Pacific Northwest,
-where they are hypothesized to hybridize \citep{geraldes_etal_2014, suarezgonzalez_etal_2016}.
+where they are hypothesized to hybridize \cite{geraldes_etal_2014, suarezgonzalez_etal_2016}.
 Both species are sampled over a large geographic region, 
-and show spatial patterns of genetic and phenotypic variation \citep{slavov_etal_2012, mckown_etal_2013},
+and show spatial patterns of genetic and phenotypic variation \cite{slavov_etal_2012, mckown_etal_2013},
 making the system well-suited for application of our method.
@@
 and compared these models using cross-validation.
-For comparison, we also ran fastSTRUCTURE \citep{fastStructure} 
+For comparison, we also ran fastSTRUCTURE \cite{fastStructure} 
 using $K = 1$ through 7.
@@
 Both points are
-in agreement with \citet{keller_etal_2010}, 
+in agreement with \cite{keller_etal_2010}, 
 who found low diversity within the region's \bals{},
@@
 were experimentally shown to display differences in ecologically relevant phenotypes 
-\citep[e.g., pathogen resistance,][]{xie2009,xie2012}.  
+(e.g., pathogen resistance, \cite{xie2009,xie2012}).  
 At higher values of $K$, drainages at the southern tip of \tri{} sampling 
@@
 perhaps due to introgression from the southern neighbors
-\textit{P. angustifolia} or \textit{fremontii} \citep{Zhou2012,geraldes_etal_2014}
+\textit{P. angustifolia} or \textit{fremontii} \cite{Zhou2012,geraldes_etal_2014}
 
@@
 from which they subsequently expanded to occupy their current range
-\citep{WoodingWard1997,Byun1997,Stone2000,Puckett2015},
+\cite{WoodingWard1997,Byun1997,Stone2000,Puckett2015},
 %The retreat of the glaciers was relatively recent in units of 
-%black bear generation times \citep{Onorato2004}, 
+%black bear generation times \cite{Onorato2004}, 
 %so we might reasonably expect the divergence accrued between refugia 
 %to be reflected in discrete genetic structure between extant populations.
-%Black bears are strong dispersers \citep{Moore2014}, 
+%Black bears are strong dispersers \cite{Moore2014}, 
 %but they are distributed across most of the continent, 
@@
 and compared these models using cross-validation.
-We also ran fastSTRUCTURE \citep{fastStructure} on the same dataset, 
+We also ran fastSTRUCTURE \cite{fastStructure} on the same dataset, 
 using $K = 2$ through 4.
@@
 into their own cluster with an inferred Beringian glacial refugium
-\citep{Byun1997,Stone2000,Puckett2015}.
+\cite{Byun1997,Stone2000,Puckett2015}.
 Our model, which explicitly incorporates a spatial decay of relatedness, 
@@
 to the inverse of the covariance matrix 
-\citep[e.g.][]{woodbury1950,sherman_morrison1950}
+(e.g., \cite{woodbury1950,sherman_morrison1950})
 when updating a sample's admixture proportions, 
@@
 In the limit of infinite data, the best value of $K$ 
-may be the number of samples included in the dataset \citep{Patterson2006}.
+may be the number of samples included in the dataset \cite{Patterson2006}.
 
@@
 should be viewed as hypotheses that should be subject to further testing
-\citep[see][for an in-depth discussion of these points]{Falush:16}.
+(see \cite{Falush:16} for an in-depth discussion of these points).
 
@@
 both of which are often predicated on the identification of discrete ``management units" (MUs) 
-identified using genetic data \citep{Moritz1994,Waples_1998,Moritz_etal_2002}.
+identified using genetic data \cite{Moritz1994,Waples_1998,Moritz_etal_2002}.
 
@@
 For example, if we were to run a \texttt{conStruct} analysis using geographic distances between 
-sampled individuals of greenish warblers \citep{Irwin2001} or \textit{Ensatina} salamanders \citep{wake_schneider1998} 
+sampled individuals of greenish warblers \cite{Irwin2001} or \textit{Ensatina} salamanders \cite{wake_schneider1998} 
 --- two canonical examples of rings species --- 
@@
 (a star population phylogeny, although the branches can have different lengths thanks to the $\phi^{(k)}$ parameter), 
-similar to the F-model in \citet{falush2003}.
+similar to the F-model of \cite{falush2003}.
 However, we could extend the existing model by implementing 
@@
 estimating a population phylogeny between them
-\citep[e.g.][]{treemix}.
+(e.g, \cite{treemix}).
 
@@
 A natural extension would be to attempt to ``geo-locate'' 
-the ancestry of samples without geographic coordinates \citep{Wasser2004}. 
+the ancestry of samples without geographic coordinates \cite{Wasser2004}. 
 We could also imagine letting samples draw ancestry from other geographic coordinates, 
-as we have done in a previous approach \citep{spacemix} to model long distance dispersal. 
+as we have done in a previous approach \cite{spacemix} to model long distance dispersal. 
 We could even allow entire layers to bud off of a particular location on another layer. 
@@
 it is becoming possible to ask whether there is genetic continuity at a point in space across time, 
-or whether populations are being replaced \citep{lazaridis_ancient_2014, Haak2015, slatkin_racimo2016, Nielsen2017, Schraiber2017}.
+or whether populations are being replaced \cite{lazaridis_ancient_2014, Haak2015, slatkin_racimo2016, Nielsen2017, Schraiber2017}.
 However, we expect allele frequencies to change through time in a population, 
@@
 \text{Var}(\epsilon) +  \text{Var}\left( \Delta^{(k)} \right) +
-\text{Cov}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell}) + \delta_{i=j} \text{Var}\left( \Delta^{(i)} \right)\ \\
+\text{Cov}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell}) + \delta_{i=j} \text{Var}\left( \Delta^{(i)} \right),
 \end{equation}
@@
 
-We employ Monte Carlo cross-validation approach for model comparison \citep{picard1984}. 
+We employ Monte Carlo cross-validation approach for model comparison \cite{picard1984}. 
 This procedure generates a mean predictive accuracy for each model and each value of $K$, 
@@
 
-Using the program \texttt{ms} \citep{Hudson2002}, 
+Using the program \texttt{ms} \cite{Hudson2002}, 
 we generated discrete population structure by simulating $K$ distinct populations,
@@
 	\caption{
-	Map of admixture proportions estimated using fastSTRUCTURE \citep{fastStructure} for $K=2$ through 4.
+	Map of admixture proportions estimated using fastSTRUCTURE \cite{fastStructure} for $K=2$ through 4.
 	The data were simulated using one layer with nearest-neighbor symmetric migration between demes.
@@
 	\caption{
-	Map of admixture proportions estimated using fastSTRUCTURE \citep{fastStructure} for $K=2$ through 4.
+	Map of admixture proportions estimated using fastSTRUCTURE \cite{fastStructure} for $K=2$ through 4.
 	The data were simulated using two layers with nearest-neighbor symmetric migration between demes.
@@
 	\caption{
-	Map of admixture proportions estimated using fastSTRUCTURE \citep{fastStructure} for $K=2$ through 4.
+	Map of admixture proportions estimated using fastSTRUCTURE \cite{fastStructure} for $K=2$ through 4.
 	The data were simulated using three layers with nearest-neighbor symmetric migration between demes.
@@
 	Maps of admixture proportions estimated for the \textit{Populus} dataset 
-	using fastSTRUCTURE \citep{fastStructure} for $K=2$ through 7.
+	using fastSTRUCTURE \cite{fastStructure} for $K=2$ through 7.
     }\label{populus_fastStr}
@@
 	Maps of admixture proportions estimated for the black bear dataset 
-	using fastSTRUCTURE \citep{fastStructure} for $K=2$ through 4.
+	using fastSTRUCTURE \cite{fastStructure} for $K=2$ through 4.
     }\label{bear_fastStr}

### 2017-09-15 writeup/conStruct.tex
latexified

@@
 and related models (e.g., \cite{ADMIXTURE}); 
-the \gc{main} difference is that our likelihood assumes that allele frequencies are normally distributed 
+the main difference is that our likelihood assumes that allele frequencies are normally distributed 
 around their expectations, 
@@
 computational speed.)
-\gc{The second small difference is that in the original STRUCTURE model,
+The second small difference is that, in the original STRUCTURE model, 
 allele frequencies at each locus are independently drawn for each 
 cluster \cite{STRUCTURE}, while in conStruct's non-spatial model, 
-  it is more natural to envision each layer's allele frequency
+it is more natural to envision each cluster's allele frequency 
 as being drifted away from a single, global allele frequency. 
 This makes our model more
-  closely related to the ``$F$-model'' prior for allele frequencies \cite{falush2003}.
-As these differences are relatively small,} we can compare the fit of the different models 
+closely related to the ``$F$-model'' prior for allele frequencies of \cite{falush2003}.
+As these differences are relatively small, we can compare the fit of the different models 
 --- spatial vs. nonspatial, across different values of $K$ --- 
@@
 encouraging the model to ``not use" unnecessary layers if they are
-present \gc{in the model'}, 
+present in the model, 
 so that they will have a low contribution to overall covariance.
@@
 the genomes of any pair of individuals \emph{agree} with some
-background probability \gc{at a locus},
+background probability at a locus,
 but this probability of agreement is increased on any segment of genome 
-that both have inherited \gc{from} the same layer 
+that both have inherited from the same layer 
 (the amount it increases depends on how far apart they are
-\gc{geographically and the decay of IBD}.)
+geographically and on the decay of IBD).
 We use this characterization to quantify the relative contributions of each layer,

### 2017-09-15 writeup/conStruct.tex
author summary

--- 2017-09-15 writeup/conStruct.tex
@@
 \end{abstract}
+\newpage
+
+\section*{Author Summary}
+One of the first steps in the analysis of genetic data, 
+and a principal mission of biology, 
+is to describe and categorize natural variation. 
+A continuous pattern of differentiation, 
+such as when individuals found closer together in space 
+are, on average, more genetically similar than individuals sampled farther apart, 
+can confound attempts to categorize natural variation into discrete groups.  
+This is because current statistical methods for assigning individuals to discrete clusters 
+cannot accommodate continuous spatial patterns, 
+making it seem as though there are many more clusters than there actually are. 
+Such patterns are common in nature, 
+so this shortcoming of existing methods may have widespread repercussions.
+In this study, we introduce a new statistical method 
+for categorizing natural genetic variation - 
+one that describes variation as a combination of continuous and discrete patterns.
+Using simulated data, we show how current approaches can be misleading, 
+and also demonstrate that our new method works well.  
+We predict that our method will be an important tool for the analysis of genetic data.
 

### 2017-09-15 writeup/conStruct.tex

--- 2017-09-15 writeup/conStruct.tex
@@
 and by applying it to empirical datasets of poplars and black bears in North America.
+\end{abstract}
+\newpage
 
@@
 \section*{Author summary}
-\plr{WRITE AUTHOR SUMMARY}
+One of the first steps in the analysis of genetic data, 
+and a principal mission of biology, 
+is to describe and categorize natural variation. 
+A continuous pattern of differentiation, 
+such as when individuals found closer together in space 
+are, on average, more genetically similar than individuals sampled farther apart, 
+can confound attempts to categorize natural variation into discrete groups.  
+This is because current statistical methods for assigning individuals to discrete clusters 
+cannot accommodate continuous spatial patterns, 
+making it seem as though there are many more clusters than there actually are. 
+Such patterns are common in nature, 
+so this shortcoming of existing methods may have widespread repercussions.
+In this study, we introduce a new statistical method 
+for categorizing natural genetic variation - 
+one that describes variation as a combination of continuous and discrete patterns.
+Using simulated data, we show how current approaches can be misleading, 
+and also demonstrate that our new method works well.  
+We predict that our method will be an important tool for the analysis of genetic data.
 

### 2017-09-15 writeup/conStruct.tex
fixed typos
more latexing

--- 2017-09-15 writeup/conStruct.tex
@@
 
+%% added PLR
+\newif\ifsubmissionversion
+\submissionversiontrue
+\newif\ifincludesupplement
+\includesupplementtrue
+\newif\iffloatsatend
+\floatsatendfalse
+
+\iffloatsatend
+\usepackage{endfloat}
+\fi
+%% end added PLR
+
 % amsmath and amssymb packages, useful for mathematical formulas and symbols
@@
 % Use nameref to cite supporting information files (see Supporting Information section for more info)
-\usepackage{nameref,hyperref}
+\usepackage{nameref}
+\usepackage[hidelinks]{hyperref}
 
 % line numbers
+\ifsubmissionversion
 \usepackage[right]{lineno}
+\fi
 
@@
 \setlength{\headheight}{27.023pt}
-\lhead{\includegraphics[width=2.0in]{latex/PLOS-submission.eps}}
+\lhead{\includegraphics[width=\textwidth]{latex/PLOS-submission.eps}}
 \rfoot{\thepage/\pageref{LastPage}}
@@
 	\centering
-		{\includegraphics[width=4in,height=2.5in]{figs/schematic/method_schematic.png}}
+		{\includegraphics[width=\textwidth]{figs/schematic/method_schematic.png}}
 		\caption{Schematic of our method, using $K=3$ as an example.
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/sims/Fig2_simK1_sp_vs_nsp.pdf}}
+		{\includegraphics[width=\textwidth]{figs/sims/Fig2_simK1_sp_vs_nsp.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=2in]{figs/sims/sim_xvals.pdf}}
+		{\includegraphics[width=\textwidth]{figs/sims/sim_xvals.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=5in,height=2.5in]{figs/sims/simK1_laycon_barplots.pdf}}
+		{\includegraphics[width=\textwidth]{figs/sims/simK1_laycon_barplots.pdf}}
 		\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/populus/pop_sp_results.pdf}}
+		{\includegraphics[width=\textwidth]{figs/populus/pop_sp_results.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=3in]{figs/populus/populus_std_xval.pdf}}
+		{\includegraphics[width=\textwidth]{figs/populus/populus_std_xval.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=3in]{figs/bears/bears_sp_vs_nsp.pdf}}
+		{\includegraphics[width=\textwidth]{figs/bears/bears_sp_vs_nsp.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=3in]{figs/bears/bear_std_xval.pdf}}
+		{\includegraphics[width=\textwidth]{figs/bears/bear_std_xval.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=3in]{figs/bears/bears_laycon_barplots.pdf}}
+		{\includegraphics[width=\textwidth]{figs/bears/bears_laycon_barplots.pdf}}
 	\caption{
@@
 
-\newpage
 \section*{Materials and Methods}
@@
 
-\newpage
 \section*{Cross validation}\label{Xvalidation}
@@
 
-\newpage
 \section*{Calculating layer contributions}\label{layer_contribution}
@@
 
-\newpage
 \section*{Simulation details}\label{sim_details}
@@
 	\centering
-		{\includegraphics[width=4in,height=2.5in]{figs/sims/sim_setup.png}}
+		{\includegraphics[width=\textwidth]{figs/sims/sim_setup.png}}
 		\caption{Schematic of how we simulate datasets with continuous and discrete differentiation, using $K=2$ as an example.  
@@
 \end{figure}
+
+\section*{Acknowledgements}
+
+We thank Marjorie Weber, Yaniv Brandvain, William Wetzel, 
+Mariah Meek, Doc Edge, and Matthew Stephens 
+for invaluable comments on the method and manuscript, 
+as well as Quentin Cronk, who provided input on the \textit{Populus} analyses, 
+and Emily Puckett, who provided input on the black bear analyses.
+We also thank the attendees at the 2017 SSE Meeting in Portland, OR, 
+whose votes determined the name of the method.
+% This work was supported in part by 
+% the National Science Foundation under award number NSF \#1262645 (DBI) to PR and GC, 
+% the National Institute of General Medical Sciences of the National
+% Institutes of Health under award numbers NIH R01-GM108779 to GC,
+% and the National Science Foundation under award numbers NSF \#1148897 and \#1402725 to GB.
+
+\newpage
+\clearpage
+\bibliography{reference.library.tex.bib}
+
+%%% BEGIN IF INCLUDE SUPPLEMENT
+\ifincludesupplement
 
@@
 
+% \linespread{1}
+\iffloatsatend
+\processdelayedfloats
+\fi
+
 \newpage
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/sims/simK1_nsp_pies.pdf}}
+		{\includegraphics[width=\textwidth]{figs/sims/simK1_nsp_pies.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/sims/simK1_sp_pies.pdf}}
+		{\includegraphics[width=\textwidth]{figs/sims/simK1_sp_pies.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=5in,height=2.5in]{figs/sims/simK1_std_xval.pdf}}
+		{\includegraphics[width=\textwidth]{figs/sims/simK1_std_xval.pdf}}
 		\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/sims/simK2_nsp_pies.pdf}}
+		{\includegraphics[width=\textwidth]{figs/sims/simK2_nsp_pies.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/sims/simK2_sp_pies.pdf}}
+		{\includegraphics[width=\textwidth]{figs/sims/simK2_sp_pies.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=5in,height=2.5in]{figs/sims/simK2_laycon_barplots.pdf}}
+		{\includegraphics[width=\textwidth]{figs/sims/simK2_laycon_barplots.pdf}}
 		\caption{
@@
 	\centering
-		{\includegraphics[width=5in,height=2.5in]{figs/sims/simK2_std_xval.pdf}}
+		{\includegraphics[width=\textwidth]{figs/sims/simK2_std_xval.pdf}}
 		\caption{
@@
 	\centering
-		{\includegraphics[width=4in,height=4in]{figs/sims/simK2_adprop_fit.pdf}}
+		{\includegraphics[width=\textwidth]{figs/sims/simK2_adprop_fit.pdf}}
 		\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/sims/simK3_nsp_pies.pdf}}
+		{\includegraphics[width=\textwidth]{figs/sims/simK3_nsp_pies.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/sims/simK3_sp_pies.pdf}}
+		{\includegraphics[width=\textwidth]{figs/sims/simK3_sp_pies.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=5in,height=2.5in]{figs/sims/simK3_laycon_barplots.pdf}}
+		{\includegraphics[width=\textwidth]{figs/sims/simK3_laycon_barplots.pdf}}
 		\caption{
@@
 	\centering
-		{\includegraphics[width=5in,height=2.5in]{figs/sims/simK3_std_xval.pdf}}
+		{\includegraphics[width=\textwidth]{figs/sims/simK3_std_xval.pdf}}
 		\caption{
@@
 	\centering
-		{\includegraphics[width=4in,height=4in]{figs/sims/simK3_adprop_fit.pdf}}
+		{\includegraphics[width=\textwidth]{figs/sims/simK3_adprop_fit.pdf}}
 		\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=2.25in]{figs/fastStr/fastStr_simK1_pies.pdf}}
+		{\includegraphics[width=\textwidth]{figs/fastStr/fastStr_simK1_pies.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=2.25in]{figs/fastStr/fastStr_simK2_pies.pdf}}
+		{\includegraphics[width=\textwidth]{figs/fastStr/fastStr_simK2_pies.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=2.25in]{figs/fastStr/fastStr_simK3_pies.pdf}}
+		{\includegraphics[width=\textwidth]{figs/fastStr/fastStr_simK3_pies.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=5in]{figs/populus/populus_sampling_map.pdf}}
+		{\includegraphics[width=\textwidth]{figs/populus/populus_sampling_map.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/populus/populus_sp_pies.pdf}}
+		{\includegraphics[width=\textwidth]{figs/populus/populus_sp_pies.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/populus/populus_nsp_pies.pdf}}
+		{\includegraphics[width=\textwidth]{figs/populus/populus_nsp_pies.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/populus/populus_sp_layer_covs.pdf}}
+		{\includegraphics[width=\textwidth]{figs/populus/populus_sp_layer_covs.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/populus/populus_nsp_layer_covs.pdf}}
+		{\includegraphics[width=\textwidth]{figs/populus/populus_nsp_layer_covs.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=3in]{figs/populus/populus_laycon_barplots.pdf}}
+		{\includegraphics[width=\textwidth]{figs/populus/populus_laycon_barplots.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/populus/poplar_fastStr_results.pdf}}
+		{\includegraphics[width=\textwidth]{figs/populus/poplar_fastStr_results.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/bears/bear_sp_results.pdf}}
+		{\includegraphics[width=\textwidth]{figs/bears/bear_sp_results.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/bears/bear_nsp_results.pdf}}
+		{\includegraphics[width=\textwidth]{figs/bears/bear_nsp_results.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/bears/bear_sp_layer_covs.pdf}}
+		{\includegraphics[width=\textwidth]{figs/bears/bear_sp_layer_covs.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=4in]{figs/bears/bear_nsp_layer_covs.pdf}}
+		{\includegraphics[width=\textwidth]{figs/bears/bear_nsp_layer_covs.pdf}}
 	\caption{
@@
 	\centering
-		{\includegraphics[width=6in,height=2in]{figs/bears/bear_fastStr_results.pdf}}
+		{\includegraphics[width=\textwidth]{figs/bears/bear_fastStr_results.pdf}}
 	\caption{
@@
 
-\section*{Acknowledgements}
-
-We thank Marjorie Weber, Yaniv Brandvain, William Wetzel, 
-Mariah Meek, Doc Edge, and Matthew Stephens 
-for invaluable comments on the method and manuscript, 
-as well as Quentin Cronk, who provided input on the \textit{Populus} analyses, 
-and Emily Puckett, who provided input on the black bear analyses.
-We also thank the attendees at the 2017 SSE Meeting in Portland, OR, 
-whose votes determined the name of the method.
-This work was supported in part by 
-the National Science Foundation under award number NSF \#1262645 (DBI) to PR and GC, 
-the National Institute of General Medical Sciences of the National
-Institutes of Health under award numbers NIH R01-GM108779 to GC,
-and the National Science Foundation under award numbers NSF \#1148897 and \#1402725 to GB.
-
-\newpage
-\clearpage
-\bibliography{reference.library.tex.bib}
+%%% BEGIN IF INCLUDE SUPPLEMENT
+\fi
 

### 2017-09-15 writeup/conStruct.tex

@@
 
+\newpage
+
 % Please keep the Author Summary between 150 and 200 words

### 2017-09-15 writeup/conStruct.tex
toggle format
G's summary

--- 2017-09-15 writeup/conStruct.tex
@@
 \section*{Author summary}
-\plr{WRITE AUTHOR SUMMARY}
+One of the first steps in the analysis of genetic data, 
+and a principal mission of biology, 
+is to describe and categorize natural variation. 
+A continuous pattern of differentiation (isolation by distance), 
+where individuals found closer together in space 
+are, on average, more genetically similar than individuals sampled farther apart, 
+can confound attempts to categorize natural variation into groups.  
+This is because current statistical methods for assigning individuals to discrete clusters 
+cannot accommodate spatial patterns, 
+and so are forced to use clusters to describe what is in fact
+continuous variation. 
+As isolation by distance is common in nature, 
+this is a substantial shortcoming of existing methods.
+In this study, we introduce a new statistical method 
+for categorizing natural genetic variation - 
+one that describes variation as a combination of continuous and discrete patterns.
+We demonstrate that this method works well and can capture patterns in
+population genomic data without resorting to splitting populations
+where they can be described by continuous patterns of variation.
+
+%Using simulated data, we show how current approaches can be misleading, 
+%and also demonstrate that our new method works well.  
+%We predict that our method will be an important tool for the analysis of genetic data.
 

### 2017-09-15 writeup/conStruct.tex
.

@@
 \newif\ifsubmissionversion
+% \submissionversionfalse
 \submissionversiontrue
@@
 % line numbers
-\ifsubmissionversion
 \usepackage[right]{lineno}
-\fi
 
@@
 % Text layout
+\ifsubmissionversion
     \raggedright
@@
     \textwidth 5.25in 
-\textheight 8.75in
+    \textheight 8.25in
+\else
+    \usepackage{fullpage}
+\fi
 
@@
 % Header and Footer with logo
-\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{graphicx}
+\ifsubmissionversion
+    \usepackage{lastpage,fancyhdr}
     \usepackage{epstopdf}
@@
     \lfoot{\sf PLOS}
+\fi
 
@@
 \\
-\textbf{a} bradburd@msu.edu; \\
-\textbf{b} gmcoop@ucdavis.edu;\\
-\textbf{c} plr@uoregon.edu\\
+\textbf{a} bradburd@msu.edu; 
+\textbf{b} gmcoop@ucdavis.edu;
+\textbf{c} plr@uoregon.edu
 
@@
 deriving from geographically distributed populations.
-Here, we present a novel statistical framework for the simultaneous inference 
+Here, we present a statistical framework for the simultaneous inference 
 of continuous and discrete patterns of population structure.
@@
 and, within each layer, estimates a rate at which relatedness decays with distance.
-This model explicitly addresses the ``clines vs. clusters" problem in 
+This thereby explicitly addresses the ``clines versus clusters" problem in 
 modeling population genetic variation.
@@
 
+\ifsubmissionversion
 \newpage
+\fi
 
@@
 
+\ifsubmissionversion
 \linenumbers
+\fi
 
@@
 including generating population phylogenies \cite{CavalliSforza1975, treemix},
-dimensionality-reduction approaches \cite{meirmans2009genodive}, 
+dimensionality reduction approaches \cite{meirmans2009genodive}
 such as principal components analysis 
@@
 but substantial work remains to be done \cite{barton2002neutral,barton2013modelling}.
-% IBD has been modeled in both spatially continuous frameworks 
+% Isolation by distance has been modeled in both spatially continuous frameworks 
 % in which an individual's parents are found within some radius of that individual's location,
@@
 Given the generality of the circumstances that generate a pattern of isolation by distance, 
-it is unsurprising that IBD is very widespread in nature \cite{meirmans2012,Sexton_etal_2014}.
+it is unsurprising that isolation by distance is very widespread in nature \cite{meirmans2012,Sexton_etal_2014}.
 
@@
 show spatially autocorrelated cluster membership \cite{Frantz2009,meirmans2012}.
-In analyses of empirical datasets, which often are characterized by IBD,
+In analyses of empirical datasets, which often show strong isolation by distance,
 model-based clustering approaches will therefore tend to overestimate
@@
 in that (a) we do not subtract locus means,
-and so to make the statistic invariant under choices of reference allele, 
+and so to make the statistic invariant under choice of reference allele, 
 (b) we randomly choose an allele to count at each locus.
@@
 we are modeling the allelic, rather than genetic, covariance. 
-In \ref{rationale} we lay out a
+In \secref{rationale} we lay out a
 simple model of allele frequencies underlying this covariance model.
@@
 $k$ (and so $\sum_{k=1}^K w^{(k)}_i =1$).
-A visual representation of the method mechanics is shown in Fig \ref{schematic}.
+A visual representation of the method is shown in Fig \ref{schematic}.
 
@@
 		\caption{Schematic of our method, using $K=3$ as an example.
-			     The pattern of isolation by distance within each layer is denoted by their color gradients, 
+                 Spatial autocorrelation of allele frequencies within each layer is depicted by color gradients, 
 			     and $\phi^{(k)}$ denotes the covariance shared by samples with ancestry entirely in the $k$th layer.
@@
 
-We can then describe the covariance between samples $i$ and $j$ across all layers, $\Omega_{i,j}$,
-by summing their within-layer spatial covariances ($G_{i,j}^{(k)}$ in layer $k$) across all $K$ layers,
+We can then describe the covariance between samples $i$ and $j$ across all $K$ layers, $\Omega_{i,j}$,
+by summing their within-layer spatial covariances ($G_{i,j}^{(k)}$ in layer $k$),
 weighted by the relevant admixture proportions.
@@
 \paragraph{Likelihood and inference}
-If the allele frequencies at each locus
+If the allele frequency deviations at each locus
 were independent between loci and multivariate normally distributed across populations, 
@@
 to estimate the posterior distribution on the parameters.
-We also provide an R package \cite{R} called \texttt{conStruct} 
-(for ``continuous structure") that functions as a wrapper around this inference machinery.
+Our R package, \texttt{conStruct},
+functions as a wrapper around this inference machinery.
 
@@
 for the number of discrete clusters in population genetic data, including: 
-information theoretic criteria (e.g., \cite{ADMIXTURE}), such as AIC or BIC;
 comparisons of the likelihood of the data across different values of $K$,
-with various criteria on how to choose a single value (e.g., \cite{Evanno2005});
+with various criteria on how to choose a single value (e.g., \cite{Evanno2005}),
+or with information theoretic penalizations such as AIC or BIC (e.g., \cite{ADMIXTURE});
 or comparisons of the marginal likelihood, 
@@
 (the amount it increases depends on how far apart they are
-geographically and on the decay of IBD).
+geographically and on the decay of isolation by distance).
 We use this characterization to quantify the relative contributions of each layer,
@@
 (see Fig \ref{sim_setup}). 
-We collapsed the layers together using random admixture with spatially autocorrelated proportions 
-(see \secref{sim_details} for details), 
-so that both the frequencies in layers and the layer memberships are
-spatially autocorrelated. 
+We collapsed the layers together using random, spatially autocorrelated admixture proportions 
+(see \secref{sim_details}).
+% so that both the frequencies in layers and the layer memberships are spatially autocorrelated. 
 We simulated datasets using $K=1$, 2, and 3 layers;
@@
 When data are simulated with $K=1$ but analyzed with $K>1$ 
-the layers are present, but contribute very little to any population.
+additional layers are present, but contribute very little to any population.
 Even when the spatial model is run with $K=7$, 
@@
 (Figs \ref{sim_xvals}, \ref{simK1_xval}, \ref{simK2_xval}, and \ref{simK3_xval}),
-while the predictive accuracy of the nonspatial model increases as subsequent clusters are added.
+while predictive accuracy of the nonspatial model increases as subsequent clusters are added.
 
@@
 %so we may therefore expect them to experience distance-limited 
-%mating opportunities and exhibit a pattern of IBD.
+%mating opportunities and exhibit a pattern of isolation by distance.
 likely leading to both continuous and discrete patterns of genetic structure.
@@
 in which a set of individuals are thought to have ancestry that originated from
-a particular geographic location embedded in a larger pattern of IBD.
+a particular geographic location embedded in a larger pattern of isolation by distance.
 
@@
 
-\section*{Model rationale}
-\subsection*{Drift, admixture, and space} \label{rationale}
+\section*{Model rationale} \label{rationale}
+\subsection*{Drift, admixture, and space}
 Here we sketch a simple model of allele frequencies and their
@@
 
-\newpage
 \begin{figure}
@@
 	\caption{
-	Plots showing the layer-specific parametric covariance IBD-curves 
+	Plots showing the layer-specific parametric covariance curves
 	estimated for the \textit{Populus} data using 
@@
 	\caption{
-	Plots showing the layer-specific parametric covariance IBD-curves 
+	Plots showing the layer-specific parametric covariance curves 
 	estimated for the black bear data using 
@@
 
-%%% BEGIN IF INCLUDE SUPPLEMENT
+%%% END IF INCLUDE SUPPLEMENT
 \fi

### 2017-09-15 writeup/conStruct.tex
gid edits

--- 2017-09-15 writeup/conStruct.tex
@@
 \newif\ifsubmissionversion
+%\submissionversionfalse
 \submissionversiontrue
@@
     \textheight 8.75in
+\else
+    \usepackage{fullpage}
+\fi
 
@@
 For example, reasonable people can disagree about whether two 
-populations are separate species, because the 
+populations are separate species because the 
 process of speciation is usually gradual, and so there is no set point
@@
 including generating population phylogenies \cite{CavalliSforza1975, treemix},
-dimensionality-reduction approaches \cite{meirmans2009genodive}, 
+dimensionality-reduction approaches, 
 such as principal components analysis 
-\cite{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
+\cite{meirmans2009genodive,menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
 and model-based clustering approaches 
@@
 
-\emph{Isolation by distance} refers to a pattern of increasing genetic differentiation
+\emph{Isolation by distance} (IBD) refers to a pattern of increasing genetic differentiation
 with geographic separation,
@@
 within each of which relatedness decays as a parametric function of the distance between samples.
-We also implement a cross-validation approach for comparing and selecting models across different numbers of layers/clusters,
+We also implement a cross-validation approach for comparing and selecting models across different numbers of layers,
 and we demonstrate the utility of our approach using both simulated and empirical data.
@@
 so it may be advisable to filter these prior to analysis
-if they are likely to contain a large percentage of errors.
+if they are likely to contain a large percentage of errors \cite{linck_battey2017}.
 
@@
 If the geographic distance between them is very large, 
-their covariance will be equal to to the background level $\phi^{(k)}$ within the layer.
+their covariance will be equal to the background level $\phi^{(k)}$ within the layer.
 The ``shared drift'' parameter $\phi^{(k)}$ is analogous to 
@@
 except for $\alpha_2$, which is uniform on $(0,2)$, 
-and $w$, for which we use an independent Dirichlet of dimension $K$ for each individual
+and $w$, for which we use an independent Dirichlet of dimension $K$ for each sample
 (see Table \ref{tab:param_prior_tab} for specifics).
@@
 We also provide an R package \cite{R} called \texttt{conStruct} 
-(for ``continuous structure") that functions as a wrapper around this inference machinery.
+(for ``\emph{con}tinuous \emph{struct}ure") that functions as a wrapper around this inference machinery.
 
@@
 than others (see \secref{Xvalidation} for more on our cross-validation procedure).
-In general, larger values of $K$ allows the model more flexibility,
+In general, larger values of $K$ allow the model more flexibility,
 and thus increases the likelihood of the training partition, 
 but this improvement in the likelihood will plateau (or even peak), 
-since above a certain $K$ the model only fits noise specific to the training data 
+as above a certain $K$ the model only fits noise specific to the training data 
 rather than generalizable patterns.
@@
 is added by spatial structure within each layer, and each additional layer.
-However, we remind users that ``statistical significance does not imply real-world significance'',
+However, we remind users that ``statistical significance does not imply real-world significance','
 and so small but statistically significant differences between models
@@
 show that subsequent layers beyond the number used for simulation are unused.
-When data are simulated with $K=1$ but analyzed with $K>1$ 
+When data are simulated with $K=1$ but analyzed with $K>1$, 
 the layers are present, but contribute very little to any population.
@@
 but we know that these boundaries are arbitrary,
-because the data were simulated with no such boundaries.
+because the data were simulated without them.
 
@@
 
-To demonstrate the utility of this method, we analyzed empirical population genomic data from two systems:
+To further demonstrate the utility of this method, 
+we also analyzed empirical population genomic data from two systems:
 a contact zone between two poplar species in northwestern North America,
@@
 The details of each were somewhat different 
-(e.g. Fig \ref{populus_nsp_pies} vs. Fig \ref{populus_fastStr}), 
+(Fig \ref{populus_nsp_pies} vs. Fig \ref{populus_fastStr}), 
 and also differed across the replicate analyses.
@@
 which also occurs in Alaska,
-the other primarily west of the Rockies (blue).
+the other primarily west of the Rockies (blue)  (Fig \ref{bear_K3}a).
 The disjointed range of the blue layer likely reflects the fact
@@
 \section*{Allelic covariance}\label{allelic_cov}
-To see how to arrive at our empirical variance-covariance expression, Eq.\ \eqref{allelic_covariance},
+To see how to arrive at our empirical variance-covariance expression (Eq.\ \eqref{allelic_covariance})
 pick a random locus and
@@
 			\end{enumerate}	
-	\item calculate the mean log likelihood of the testing data partition, 
+	\item calculate the mean log likelihood of the testing data partition  
 	over the posterior distribution of training-estimated parameters for each model 
@@
 
-\newpage
-\begin{figure}
+\begin{figure}[h]
 	\centering

### 2017-09-15 writeup/conStruct.tex

@@
 is to describe and categorize natural variation. 
-A continuous pattern of differentiation, 
-such as when individuals found closer together in space 
+A continuous pattern of differentiation (isolation by distance), 
+where individuals found closer together in space 
 are, on average, more genetically similar than individuals sampled farther apart, 
-can confound attempts to categorize natural variation into discrete groups.  
+can confound attempts to categorize natural variation into groups.  
 This is because current statistical methods for assigning individuals to discrete clusters 
-cannot accommodate continuous spatial patterns, 
-making it seem as though there are many more clusters than there actually are. 
-Such patterns are common in nature, 
-so this shortcoming of existing methods may have widespread repercussions.
+cannot accommodate spatial patterns, 
+and so are forced to use clusters to describe what is in fact
+continuous variation. 
+As isolation by distance is common in nature, 
+this is a substantial shortcoming of existing methods.
 In this study, we introduce a new statistical method 
@@
 one that describes variation as a combination of continuous and discrete patterns.
-Using simulated data, we show how current approaches can be misleading, 
-and also demonstrate that our new method works well.  
-We predict that our method will be an important tool for the analysis of genetic data.
+We demonstrate that this method works well and can capture patterns in
+population genomic data without resorting to splitting populations
+where they can be described by continuous patterns of variation.
+
+%Using simulated data, we show how current approaches can be misleading, 
+%and also demonstrate that our new method works well.  
+%We predict that our method will be an important tool for the analysis of genetic data.
 

### 2017-09-15 writeup/conStruct.tex
minor tweaks

--- 2017-09-15 writeup/conStruct.tex
@@
 	\caption{
-	Map of admixture proportions estimated using the nonspatial model for $K=2$ through 4 (top row)
-	and also the spatial model for $K=2$ through 4 (bottom row).
-	The data were simulated using a single layer with nearest-neighbor symmetric migration between demes.
+        Results for data simulated using $K=1$,
+	showing maps of admixture proportions estimated using the nonspatial model for $K=2$ through 4 (top row)
+	and the spatial model for $K=2$ through 4 (bottom row).
+	% The data were simulated using a single layer with nearest-neighbor symmetric migration between demes.
+    Since there is only a single layer in the simulation, the spatial model accurately depicts the data
+    in all cases, while the nonspatial model creates spurious clusters.
     }\label{simK1_pies}
@@
 	The inset plots zoom in on cross-validation results outlined in the dotted boxes.
+    The spatial model shows better model fit at every value of $K$.
     }\label{sim_xvals}
@@
 		\caption{
-			Layer/cluster contributions (i.e., how much total covariance is contributed by each layer/cluster), 
-			for all layers estimated in runs using $K = 1$ through 7 
-			for the spatial model (left),
-			and for all clusters using the nonspatial model (right).
-			Data were simulated using $K=1$.
-			For each value of $K$ along the x-axis, there are an equal number of contributions plotted.
+            Results for data simulated using $K=1$, showing
+            layer/cluster contributions (i.e., how much each layer/cluster contributes to total covariance),
+            from \texttt{conStruct} runs using $K = 1$ through 7 
+			for the spatial model (left), and the nonspatial model (right).
+			In each run of the nonspatial model, a single layer explained nearly all the covariance
+            (additional bars are present but not visible).
 		}\label{simK1_laycon}
@@
 
-To demonstrate the utility of this method, we analyzed empirical population genomic data from two systems:
+To demonstrate the utility of this method, we applied \texttt{conStruct} to empirical population genomic data from two systems:
 a contact zone between two poplar species in northwestern North America,
@@
 genotyped at just over 33,000 loci (map of the sampling shown in Fig \ref{populus_map}).
-The range in number of individuals per drainage was between 1 and 50, 
+The number of individuals per drainage ranged between 1 and 50, 
 with most sampling concentrated on \tri{} drainages.
@@
 
-Across multiple replicate runs, 
-both nonspatial \texttt{conStruct} and fastSTRUCTURE displayed 
+Both nonspatial \texttt{conStruct} and fastSTRUCTURE displayed 
 the successive partitioning of space and the clines of admixture seen in the simulation results.
@@
 (e.g. Fig \ref{populus_nsp_pies} vs. Fig \ref{populus_fastStr}), 
-and also differed across the replicate analyses.
+and also differed across independent runs of each analysis.
+% suggesting a problem of multimodality.
 
@@
 \begin{enumerate}
-    \item How many distinct populations (probably, distinct glacial refugia)
+    \item How many distinct populations
         are reflected in modern patterns of genetic variation?
@@
 \end{enumerate}
-Differing strengths of isolation by distance might indicate 
+Distinct popualations likely represent different glacial refugia,
+and differing strengths of isolation by distance might indicate 
 different levels of habitat connectivity,
@@
 	\caption{
-	Cross-validation results for black bear dataset 
-	comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
+	Cross-validation results for the black bear dataset,
+	comparing spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
 	The first panel in each row shows all results; 
@@
 those obtained using the nonspatial model, 
-but they do showcase the same tendency: 
+but they do show the same tendency: 
 e.g., at $K=3$, fastSTRUCTURE splits the westernmost Alaskan samples 
@@
 
-The difficulty of dispersing across geography makes
-the pattern of isolation by distance ubiquitous in nature.
-% this is unsurprising, as mating opportunities are distance-limited in most species.
-% This common pattern can flummox 
-Commonly used clustering methods
-must employ a possibly large number of discrete clusters
-to describe this simple pattern.
-The resulting spurious clusters can have implications for 
-downstream analysis and interpretation of genetic data, 
-including conservation management decisions.
+% The difficulty of dispersing across geography makes
+% the pattern of isolation by distance ubiquitous in nature.
+% % this is unsurprising, as mating opportunities are distance-limited in most species.
+% % This common pattern can flummox 
+% Commonly used clustering methods
+% must employ a possibly large number of discrete clusters
+% to describe this simple pattern.
+% The resulting spurious clusters can have implications for 
+% downstream analysis and interpretation of genetic data, 
+% including conservation management decisions.
 
@@
 To aid comparison between models,
-we also present a cross-validation approach 
+we present a cross-validation approach 
 as well as a way to describe the contribution of each spatial layer to the model
@@
 and accurately pick the simulating model as the best model using our cross-validation procedure.
-Empirical applications of \texttt{conStruct} to a sample of North American poplars 
-and a sample of North American black bears yield reasonable results,
+Two empirical applications of \texttt{conStruct} to samples of North American poplars and black bears
+yield reasonable results,
 and demonstrate that,
@@
 \texttt{conStruct} can be much faster than other nonspatial clustering methods.
-On a desktop machine with a 4.2 GHz Intel Core i7 processor, 
+On a desktop machine, using a single 4.2 GHz Intel Core i7 processor, 
 an analysis of the black bear dataset (78 samples, 21,000 loci) 
@@
 \paragraph{Implications for management and conservation}
-If isolation by distance is common, 
-then a likely result of applying \texttt{conStruct} to existing data is that 
+Since isolation by distance is common, 
+a likely result of applying \texttt{conStruct} to existing data is that some
 populations previously identified using nonspatial clustering methods 
-will be collapsed into each other.  
-This ``lumping" might better reflect biological reality, 
+may be collapsed into each other.  
+This ``lumping'' might better reflect biological reality, 
 but may also have implications for management decisions and conservation policy, 
-both of which are often predicated on the identification of discrete ``management units" (MUs) 
+both of which are often predicated on the identification of discrete ``management units'' (MUs) 
 identified using genetic data \cite{Moritz1994,Waples_1998,Moritz_etal_2002}.
@@
 may be quite genetically diverged from one another, 
-perhaps especially at loci underlying adaptive traits, 
+perhaps especially at loci underlying adaptive traits, % \citep{toews2016plumage}, 
 and that a \texttt{conStruct} layer may still contain multiple distinct MUs worthy of independent protections.  
@@
 or of implementing an integrated conservation plan across multiple demes within a layer.
-Conversely, two populations that are easily distinguishable based on abundant genetic data
-need not be strongly differentiated --
-pouplation assignment is not a replacement for absolute measures of genetic divergence.
+% Conversely, two populations that are easily distinguishable based on abundant genetic data
+% need not be strongly differentiated --
+% population assignment is not a replacement for absolute measures of genetic divergence.
 
@@
 Imagine a sample $i$ that draws all of its ancestry from layer $k$.
-The allele frequency in sample $i$ at locus $\ell$, $F_{i,\ell}$ can be
+The allele frequency in sample $i$ at locus $\ell$, denoted $F_{i,\ell}$, can be
 written as the sum
@@
 F_{i,\ell} = \epsilon_{\ell} + \Delta^{(k)}_{\ell} +
-\Delta^{(k,i)}_{\ell} + \Delta^{(i)}_{\ell}
+\Delta^{(k,i)}_{\ell} + \Delta^{(i)}_{\ell} .
 \label{drift_terms_no_admix}
@@
 The weights, $w^{(k)}_{i}$, describe the
-``admixture proportion'' of sample $i$ in layer $k$
+``admixture proportion'' of sample $i$ in layer $k$.
 (Note that $\sum_{k=1}^K w^{(k)}_{i} = 1$ for each $i$.)
@@
 but otherwise we let $X=1-A$ and $Y=1-B$.
-We then let $\widehat{\Omega}_{i,j} = \cov[X,Y]$ ; 
-sampling without replacement makes the statistic insensitive to sample size,
+We then let $\widehat{\Omega}_{i,j} = \cov[X,Y]$. 
+Sampling without replacement makes the statistic insensitive to sample size,
 and random allele flipping makes it independent of the choice of reference allele.
@@
 and using the fact that $\E[X] = \E[Y] = 1/2$,
-Eq.\ \eqref{allelic_covariance} comes from the calculation that
-\begin{align}
-\cov[X,Y] &= \E[XY] - 1/4 = (1/2)\E[AB + (1-A)(1-B)] - 1/4\\\notag
-&= \E[(A-1/2)(B-1/2)]
+Eq.\ \eqref{allelic_covariance} comes from the observation that
+\begin{align} \label{eqn:cov_xy}
+\cov[X,Y] % &= \E[XY] - 1/4 = (1/2)\E[AB + (1-A)(1-B)] - 1/4\\\notag
+    &= \E[(A-1/2)(B-1/2)] .
 \end{align}
@@
 \section*{Models, parameters, and priors}\label{model_app}
-\paragraph{Spatial vs. nonspatial}
+\paragraph{Spatial versus nonspatial}
 In this paper, we discuss two types of models, spatial and nonspatial, 
@@
 The spatial model is parameterized as in Eq.\ \eqref{admixed_spatial_cov},
-and the nonspatial model is described in Eq.\ \eqref{admixed_discrete_covariance}.
-The nonspatial model is a special case of the spatial model, 
-in which the value of all $\alpha$ parameters is set to 0.
+and the nonspatial model
+is a special case of the spatial model with all $\alpha$ parameters is set to 0.
+% is described in Eq.\ \eqref{admixed_discrete_covariance}.  % that's not the right eq'n
 The nonspatial model therefore has $3K$ fewer parameters than the spatial model,
-as there are three $\alpha$ parameters that describe the continuous differentiation effect of distance per layer,
-and they are fixed in the nonspatial model.
+since three $\alpha$ parameters that describe the continuous differentiation effect of distance in each layer.
 
@@
 The single-layer model is therefore a special case of the multi-layer model, 
-in which we set $\phi$ to zero, as described below:
-
+in which we set $\phi$ to zero.
 For the spatial model, the single-layer parametric covariance is:
@@
 \alpha^{(k)}_0 \times \left(\text{exp} \left(  -(\alpha^{(k)}_D D_{i,j})^{\alpha^{(k)}_2}\right) \right)	 +
-\delta_{i=j} \text{Var}(\Delta_i)
+\delta_{i=j} \text{Var}(\Delta_i), 
 \label{admixed_continuous_cov}
 \end{equation}
-
-For the nonspatial model, the single-cluster parametric covariance is:
+and for the nonspatial model, it is:
 \begin{equation}
-\Omega_{i,j} = \text{Var}(\epsilon) + \delta_{i=j} \text{Var}(\Delta^{(i)})
+\Omega_{i,j} = \text{Var}(\epsilon) + \delta_{i=j} \text{Var}(\Delta^{(i)}) .
 \label{admixed_discrete_covariance}
@@
 (if the two populations are the same, then choose without replacement).
-Then, if we let $X=2(A-1/2)$ and $Y=2(B-1/2)$,
-since $X$ and $Y$ take the values $\pm 1$,
-and thanks to the random polarization, $\E[A] = \E[B] = 1/2$,
+Then, if we let $U=2(A-1/2)$ and $V=2(B-1/2)$,
+since $U$ and $V$ take the values $\pm 1$,
+so as in Eq.\ \eqref{eqn:cov_xy},
 $$\begin{aligned}
-\E[XY]
-        &= \P\{ X=Y \} - \P\{ X \neq Y \} \\
-        &= 2 \P\{ X=Y \} - 1 \\
-        &= 1 - 2 \P\{ X \neq Y \} .
+\E[UV]
+        &= \P\{ U=V \} - \P\{ U \neq V \} \\
+        &= 2 \P\{ U=V \} - 1 \\
+        % &= 1 - 2 \P\{ U \neq V \} .
+        &= 2 \P\{ A=B \} - 1 
 \end{aligned}$$
-To translate, $\P\{ X=Y \}$ is the probability that the alleles from our two focal samples agree with each other,
-while $\P\{ X \neq Y \}$ is the probability that they disagree.
-This implies that $\E[XY] = 1 - 2 \pi_{ij}$, 
+To translate, $\P\{ U=V \}$ is the probability that the alleles from our two focal samples agree with each other,
+while $\P\{ U \neq V \}$ is the probability that they disagree.
+This implies that $\E[UV] = 1 - 2 \pi_{ij}$, 
 where $\pi_{ij}$ is the probability that two randomly chosen alleles differ, 
@@
 Now, here is a generative model that gives us the form of the covariance we have postulated.
-To decide whether or not $X$ and $Y$ will agree,
+To decide whether or not $A$ and $B$ will agree,
 first each sample randomly chooses a layer: call these layers $I$ and $J$.
-The probability that $X$ chooses layer $k$ is $\P({I=k})=w_i^{(k)}$, 
+The probability that $A$ chooses layer $k$ is $\P({I=k})=w_i^{(k)}$, 
 the $i$th sample's admixture proportion in the $k$th layer.
-The same holds true for $Y$.
+The same holds true for $B$.
 If they do not choose the same layer, the probability that they agree is $p_\gamma$.
@@
 whose votes determined the name of the method.
-% This work was supported in part by 
-% the National Science Foundation under award number NSF \#1262645 (DBI) to PR and GC, 
-% the National Institute of General Medical Sciences of the National
-% Institutes of Health under award numbers NIH R01-GM108779 to GC,
-% and the National Science Foundation under award numbers NSF \#1148897 and \#1402725 to GB.
+\ifsubmissionversion
+\else
+This work was supported in part by 
+the National Science Foundation under award number NSF \#1262645 (DBI) to PR and GC, 
+the National Institute of General Medical Sciences of the National
+Institutes of Health under award numbers NIH R01-GM108779 to GC,
+and the National Science Foundation under award numbers NSF \#1148897 and \#1402725 to GB.
+\fi
 

### 2017-09-15 writeup/conStruct.tex

@@
 \newif\ifsubmissionversion
-% \submissionversionfalse
+%\submissionversionfalse
 \submissionversiontrue
@@
 including generating population phylogenies \cite{CavalliSforza1975, treemix},
-dimensionality reduction approaches \cite{meirmans2009genodive}
+dimensionality-reduction approaches, 
 such as principal components analysis 
-\cite{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
+\cite{meirmans2009genodive,menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
 and model-based clustering approaches 
@@
 to estimate the posterior distribution on the parameters.
-Our R package, \texttt{conStruct},
+Our R package, \texttt{conStruct} (for ``\emph{con}tinuous \emph{struct}ure"),
 functions as a wrapper around this inference machinery.
@@
 show that subsequent layers beyond the number used for simulation are unused.
-When data are simulated with $K=1$ but analyzed with $K>1$ 
-additional layers are present, but contribute very little to any population.
+When data are simulated with $K=1$ but analyzed with $K>1$, 
+the layers are present, but contribute very little to any population.
 Even when the spatial model is run with $K=7$, 
@@
 
-To demonstrate the utility of this method, we applied \texttt{conStruct} to empirical population genomic data from two systems:
+To further demonstrate the utility of this method, 
+we also applied \texttt{conStruct} to empirical population genomic data from two systems:
 a contact zone between two poplar species in northwestern North America,
@@
 The details of each were somewhat different 
-(e.g. Fig \ref{populus_nsp_pies} vs. Fig \ref{populus_fastStr}), 
-and also differed across independent runs of each analysis.
-% suggesting a problem of multimodality.
+(Fig \ref{populus_nsp_pies} vs. Fig \ref{populus_fastStr}), 
+and also differed across the replicate analyses.
 
@@
 
-\begin{figure}
+\begin{figure}[h]
 	\centering

### 2017-09-15 writeup/conStruct.tex
ultimate tweaks

@@
 \paragraph{Implications for management and conservation}
-Since isolation by distance is common, 
+Because isolation by distance is common, 
 a likely result of applying \texttt{conStruct} to existing data is that some
@@
 The nonspatial model therefore has $3K$ fewer parameters than the spatial model,
-since three $\alpha$ parameters that describe the continuous differentiation effect of distance in each layer.
+because there are three $\alpha$ parameters that describe the continuous differentiation effect of distance in each layer.
 

### 2017-09-20 writeup/cover_letter.tex
submission to plosgen

@@
 Submission Cover Letter \hfill	\today\\
-\emph{PLoS Biology}
+\emph{PLoS Genetics}
 \\\\\\
@@
 ``Inferring Continuous and Discrete Population Genetic Structure Across Space,"
-for consideration for publication in PLoS Biology.
+for consideration for publication in PLoS Genetics.
 

### 2017-09-20 writeup/cover_letter.tex
tweaked cover letter

@@
 \usepackage[green]{msuletterhead}
+\pagenumbering{gobble}
 \begin{document}
@@
 and Matthew Stephens (mstephens@uchicago.edu) \\
-would be suitable reviewers.
+would be suitable reviewers or external editors.
+(as Coop is an associate editor).
 

### 2017-09-20 writeup/conStruct.tex
submission to plosgen
yaniv caught a typo

@@
 \includesupplementtrue
+%\includesupplementfalse
 \newif\iffloatsatend
 \floatsatendfalse
+%\floatsatendtrue
 
@@
 between populations.
-% This is partly a result of the difficulty of sampling uniformly and continuously 
-% across the range of a population or species, but more, it 
 This reflects a shortcoming of current methods for inferring and 
@@
 modeling population genetic variation.
-% by jointly accommodating both continuous and discrete patterns of differentiation. 
 The method produces useful descriptions of structure in genetic relatedness
@@
 \end{enumerate}
-Distinct popualations likely represent different glacial refugia,
+Distinct populations likely represent different glacial refugia,
 and differing strengths of isolation by distance might indicate 
@@
 
+\begin{figure}[!ht]
+	\centering
+		{\includegraphics[width=\textwidth]{figs/sims/sim_setup.png}}
+		\caption{Schematic of how we simulate datasets with continuous and discrete differentiation, using $K=2$ as an example.  
+			    Going forward in time, the $K$ populations split from a common ancestor at time $\tau_{\text{s}}$,
+			    then expand to each  colonize a lattice of demes with nearest-neighbor symmetric migration at time $\tau_{\text{e}}$,
+			    then finally at time $\tau_{\text{a}}$ collapse into a single lattice consisting of demes 
+			    with ancestry entirely in one or the other of the populations,
+			    or admixed between them.
+			    }\label{sim_setup}
+\end{figure}
+
 Using the program \texttt{ms} \cite{Hudson2002}, 
@@
 
-\begin{figure}
-	\centering
-		{\includegraphics[width=\textwidth]{figs/sims/sim_setup.png}}
-		\caption{Schematic of how we simulate datasets with continuous and discrete differentiation, using $K=2$ as an example.  
-			    Going forward in time, the $K$ populations split from a common ancestor at time $\tau_{\text{s}}$,
-			    then expand to each  colonize a lattice of demes with nearest-neighbor symmetric migration at time $\tau_{\text{e}}$,
-			    then finally at time $\tau_{\text{a}}$ collapse into a single lattice consisting of demes 
-			    with ancestry entirely in one or the other of the populations,
-			    or admixed between them.
-			    }\label{sim_setup}
-\end{figure}
-
 \section*{Acknowledgements}
@@
 
-\begin{figure}[h]
+\clearpage
+
+\begin{figure}
 	\centering
@@
 \end{figure}
+\clearpage
 
@@
 \end{figure}
-
-\newpage
+\clearpage
+
 \begin{figure}

### 2017-09-20 writeup/conStruct.tex
added script for plosgen submission

@@
 % % % % % % % % % % % % % % % % % % % % % % % %
-
 \documentclass[10pt,letterpaper]{article}
 \usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
-
 %% added PLR
@@
 %\floatsatendtrue
-
+\newif\ifincludefigs
+%\includefigstrue
+\includefigsfalse
 \iffloatsatend
-\usepackage{endfloat}
+\usepackage[nomarkers]{endfloat}
 \fi
 %% end added PLR
-
 % amsmath and amssymb packages, useful for mathematical formulas and symbols
 \usepackage{amsmath,amssymb}
-
 % Use adjustwidth environment to exceed column width (see example table in text)
 \usepackage{changepage}
-
 % Use Unicode characters when possible
 \usepackage[utf8x]{inputenc}
-
 % textcomp package and marvosym package for additional characters
 \usepackage{textcomp,marvosym}
-
 % cite package, to clean up citations in the main text. Do not remove.
 \usepackage{cite}
-
 % Use nameref to cite supporting information files (see Supporting Information section for more info)
@@
 \usepackage[hidelinks]{hyperref}
-
 % line numbers
 \usepackage[right]{lineno}
-
 % ligatures disabled
@@
 \DisableLigatures[f]{encoding = *, family = * }
-
 % color can be used to apply background shading to table cells only
 \usepackage[table]{xcolor}
-
 % array package and thick rules for tables
 \usepackage{array}
-
 % create "+" rule type for thick vertical lines
 \newcolumntype{+}{!{\vrule width 2pt}}
-
 % create \thickcline for thick horizontal lines of variable length
@@
 }
-
 % \thickhline command for thick horizontal lines that span the table
@@
 \noalign{\global\arrayrulewidth\savedwidth}}
-
 % Remove comment for double spacing
@@
 %\doublespacing
-
 % Text layout
@@
 \fi
-
 % Bold the 'Figure #' in the caption and separate it from the title/caption with a period
@@
 \renewcommand{\figurename}{Fig}
-
 % Use the PLoS provided BiBTeX style
 \bibliographystyle{latex/plos2015}
-
 % Remove brackets from numbering in List of References
@@
 \makeatother
-
 % Leave date blank
 \date{}
-
 % Header and Footer with logo
@@
 \fi
-
 %% Include all macros below
-
 \newcommand{\e}[1]{{\mathbb E}\left[ #1 \right]}
@@
 \newcommand{\bals}{\textit{balsamifera}}
-
 \newcommand{\gb}[1]{{\it\color{magenta}{(#1)}}}
@@
 \newcommand{\gc}[1]{{\it\color{blue}{(#1)}}}
-
 \newcommand{\R}{\mathbb{R}}
@@
 \newcommand{\bone}{\mathbf{1}}
-
 %% END MACROS SECTION
-
 \begin{document}
 \vspace*{0.2in}
-
 % Title must be 250 characters or less.
@@
 \textbf{c} plr@uoregon.edu
-
 \bigskip
-
 %
@@
 \Yinyang These authors contributed equally to this work.
-
 \end{flushleft}
@@
 \section*{Abstract}
-
 A classic problem in population genetics is the characterization 
@@
 and by applying it to empirical datasets of poplars and black bears in North America.
-
 \ifsubmissionversion
@@
 \fi
-
 % Please keep the Author Summary between 150 and 200 words
@@
 where they can be described by continuous patterns of variation.
-
 %Using simulated data, we show how current approaches can be misleading, 
@@
 %We predict that our method will be an important tool for the analysis of genetic data.
-
 \ifsubmissionversion
@@
 \fi
-
 \newpage
@@
 are required to accurately describe most species' patterns of genetic relatedness.
-
 From a practical standpoint, even while acknowledging continuous
@@
 requiring methods that can disentangle the two.
-
 There have been many methods proposed to characterize population
@@
 are constant in a cluster across the species' range.  
-
 \emph{Isolation by distance} (IBD) refers to a pattern of increasing genetic differentiation
@@
 it is unsurprising that isolation by distance is very widespread in nature \cite{meirmans2012,Sexton_etal_2014}.
-
 The ubiquity of isolation by distance presents a challenge for models of discrete population structure,
@@
 (e.g., \cite{SerrePaabo2004,rosenberg2005clines}).
-
 Existing model-based clustering approaches can only describe continuous patterns of variation using
@@
 the number of discrete clusters present. 
-
 To address this, we set out to develop
@@
 \href{https://github.com/gbradburd/conStruct}{https://github.com/gbradburd/conStruct}.
-
 % \section*{Methods}
 \section*{Results}
-
 \paragraph{Data}
@@
 and reflects the covariance between individual alleles drawn from within the population.
-
 This definition of covariance differs from the usual ``genetic covariance'' \cite{mcvean_genealogical_2009}
@@
 if they are likely to contain a large percentage of errors \cite{linck_battey2017}.
-
 \paragraph{Continuous and discrete differentiation}
@@
 across the entire geographic range.
-
 %they connote a lack of geographical separation within each.
-
 Within each of these layers,
@@
 do so as well. 
-
 To allow flexibility in the form of the decay of allelic covariance with geographic distance within each layer, 
@@
 simple model of allele frequencies underlying this covariance model.
-
 We then allow samples to draw their ancestry from more than one layer.
@@
 A visual representation of the method is shown in Fig \ref{schematic}.
-
-\begin{figure}
-	\centering
+\begin{figure}
+	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/schematic/method_schematic.png}}
+	\fi	
 		\caption{Schematic of our method, using $K=3$ as an example.
@@
 \end{figure}
-
 We can then describe the covariance between samples $i$ and $j$ across all $K$ layers, $\Omega_{i,j}$,
@@
 sample $i$ and sample $j$ have inherited from layer $k$.
-
 In addition to the admixture-weighted sum of the within-layer spatial covariances,
@@
 such as inbreeding and the sampling process.
-
 \paragraph{Likelihood and inference}
@@
 and thereby informally model linkage between loci (e.g., \cite{EEMS}).
-
 We estimate the values of the parameters of the model using a Bayesian approach.
@@
 functions as a wrapper around this inference machinery.
-
 \paragraph{Relationship of this model to nonspatial structure models}
-
 A nice feature of our approach is that the model described in Eq.\ \eqref{cross_layer_covariance} 
@@
 are a close match to those of STRUCTURE-based approaches.
-
 \paragraph{Choice of layer number and cross-validation}
@@
 and between the spatial and nonspatial models.
-
 Several approaches have been used as model choice criteria 
@@
 between several methods.
-
 We use cross-validation (similar in spirit to \cite{ADMIXTURE_xval}) to attack this problem.
@@
 means that isolation by distance is likely a feature of the data.
-
 Cross-validation provides a valuable summary of how much explanatory power
@@
 should probably not be relied on too strongly.
-
 Another way to describe the practical significance of additional layers
@@
 so that they will have a low contribution to overall covariance.
-
 To calculate layer contributions, we use the following alternative description of our covariance model:
@@
 so we take that into account when calculating each layer's contribution to the whole.
-
 % \section*{Results}
-
 \subsection*{Simulations}
-
 To test the method, we first generated data using the coalescent simulator \texttt{ms} \cite{Hudson2002}.
@@
 % inferring admixture under models with $K=2$ through 4 (Figs \ref{fastStr_simK1}--\ref{fastStr_simK3}).
-
 % In Fig \ref{simK1_pies}, we show the results of the dataset simulated with $K=1$, 
@@
 % inferring admixture under models with $K=2$ through 4 (Figs \ref{fastStr_simK1}--\ref{fastStr_simK3}).
-
 With these simulations,
@@
 (Figs \ref{simK2_adprop_fit} and \ref{simK3_adprop_fit}).
-
 In contrast, the nonspatial model describes geographic variation
@@
 because the data were simulated without them.
-
 The spatial model's better fit is reflected by increased predictive accuracy:
@@
 while predictive accuracy of the nonspatial model increases as subsequent clusters are added.
-
 % In contrast, the predictive accuracy of the spatial \texttt{conStruct} model
@@
 % (Figs \ref{sim_xvals}, \ref{simK1_xval}, \ref{simK2_xval}, and \ref{simK3_xval}).
-
 The unimportance of spurious layers can be seen in plots of layer contributions
@@
 in contrast, additional clusters in the nonspatial analyses continue to contribute substantially.
-
-\begin{figure}
-	\centering
+\begin{figure}
+	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/sims/Fig2_simK1_sp_vs_nsp.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-
-\begin{figure}
-	\centering
+\begin{figure}
+	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/sims/sim_xvals.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-
-\begin{figure}
-	\centering
+\begin{figure}
+	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/sims/simK1_laycon_barplots.pdf}}
+	\fi
 		\caption{
@@
 \end{figure}
-
 \subsection*{Empirical Applications}
-
 To further demonstrate the utility of this method, 
@@
 and a large North American sample of black bears.
-
 \subsubsection*{Poplars}
-
 Trees in the genus \textit{Populus} (poplars, aspens, and cottonwoods) 
@@
 where they do, they frequently hybridize \cite{eckenwalder1984, Cronk2005}.
-
 \textit{Populus trichocarpa}, the black cottonwood,
@@
 \begin{enumerate}
-
     \item To what degree has hybridization blurred the boundaries between 
@@
     % rather than a single cline of ancestry across their overlap zone?
-
     \item Does the only significant boundary of population structure
@@
         or is there sub-structuring within species?
-
     \item Does the strength of isolation by distance differ between inferred layers?
@@
         % does it differ between different layers?
-
 \end{enumerate}
-
 We use data from \cite{geraldes_etal_2014}, 
@@
 as well as in Figs \ref{populus_sp_pies} - \ref{populus_fastStr} in the Supplementary Materials.
-
-\begin{figure}
-	\centering
+\begin{figure}
+	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/populus/pop_sp_results.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-
-\begin{figure}
-	\centering
+\begin{figure}
+	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/populus/populus_std_xval.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-
 All models with $K>1$ assigned the majority of each of the two species
@@
 probably as the result of a recent postglacial expansion.
-
 A consistent split between layers within \tri{} fell along the ``no-cottonwood belt," 
@@
 \textit{P. angustifolia} or \textit{fremontii} \cite{Zhou2012,geraldes_etal_2014}
-
 Both nonspatial \texttt{conStruct} and fastSTRUCTURE displayed 
@@
 and also differed across the replicate analyses.
-
 \subsubsection*{Black bears}
-
 The American black bear, \textit{Ursus americanus}, is endemic to North America
@@
 or different postglacial histories.
-
 We use data from \cite{Puckett2015}, consisting of 95 individuals 
@@
 as well as in Figs \ref{bear_sp_pies} - \ref{bear_fastStr} in the Supplementary Materials.
-
-\begin{figure}
-	\centering
+\begin{figure}
+	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/bears/bears_sp_vs_nsp.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-
-\begin{figure}
-	\centering
+\begin{figure}
+	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/bears/bear_std_xval.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-
-\begin{figure}
-	\centering
+\begin{figure}
+	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/bears/bears_laycon_barplots.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-
 The results partition the sampled bears into two main groups:
@@
 with the blue layer presumably then stretching up into British Columbia.
-
 Additional layers in the spatial model have strong statistical support 
@@
 %mitochondrial) DNA. WHY DO WE CARE about BEAR MTDNS
-
 Results from the nonspatial model clearly exhibit the tendency 
@@
 and at $K=4$, it splits them into their own cluster entirely (Fig \ref{bear_fastStr}).
-
 Across all values of $K$ for which we ran \texttt{conStruct},
@@
 (e.g., Fig \ref{bear_sp_pies}d vs \ref{bear_nsp_pies}d).
-
 %%%%%%%%%
 \section*{Discussion}
-
 % The difficulty of dispersing across geography makes
@@
 % including conservation management decisions. 
-
 In this paper, we have presented a statistical framework, \texttt{conStruct}, for simultaneously 
@@
 (but caution against overly strict interpretation of either).
-
 The method performs well on simulated data:
@@
 real datasets can be better described using fewer layers.
-
 The proposed method combines the utility of model-based clustering algorithms 
@@
 especially when the populations exhibit spatial patterns of relatedness.
-
 \paragraph{Comparison to nonspatial model-based clustering}
@@
 Hamiltonian Monte Carlo sampling of our parameters.
-
 For a relatively small number of samples, 
@@
 fastSTRUCTURE (or ADMIXTURE) would be much faster.
-
 \paragraph{Choosing the ``best" number of layers}
@@
 may be the number of samples included in the dataset \cite{Patterson2006}.
-
 From a biological perspective, 
@@
 as it can reveal layers with statistical support but small biological import.
-
 Although we believe our model adds spatial realism to the groups used by clustering methods,
@@
 (see \cite{Falush:16} for an in-depth discussion of these points).
-
 \paragraph{Implications for management and conservation}
@@
 identified using genetic data \cite{Moritz1994,Waples_1998,Moritz_etal_2002}.
-
 It is therefore important to stress that individuals sampled from the same \texttt{conStruct} layer 
@@
 % population assignment is not a replacement for absolute measures of genetic divergence.
-
 \paragraph{Caveats and considerations}
@@
 is not representative of the path traversed in the coalescent of a pair of alleles sampled at those locations.
-
 A second caveat is that, in some instances, 
@@
 and to examine the spatial covariance functions within layers when interpreting results.
-
 \paragraph{Extensions and future directions}
@@
 (e.g, \cite{treemix}).
-
 In addition, here we have assumed that samples have known geographic coordinates, 
@@
 a particular geographic location embedded in a larger pattern of isolation by distance.
-
 A final direction would be to model relatedness within a layer as a spatiotemporal process, 
@@
 in which case a change in discrete population structure across space is comparable to population replacement across time.
-
 \section*{Materials and Methods}
-
 % \section*{Appendix}
@@
 % \setcounter{section}{0}
-
 \section*{Model rationale} \label{rationale}
@@
 We will assume that these four deviations are all uncorrelated with each other.
-
 If we have two samples $i$ and $j$ drawn from layer $k$, 
@@
 as in Eq.\ \eqref{cross_layer_covariance}.
-
 \paragraph{Admixture} 
@@
 \end{equation}
-
 \paragraph{Space}
@@
 and not because it matches any explicit population genetic model of isolation by distance. 
-
 \section*{Allelic covariance}\label{allelic_cov}
@@
 which is $f_i^2 n/(n-1)$.
-
 \paragraph{Likelihood}
@@
 where $\mathcal{W}$ is the Wishart likelihood function.
-
 A benefit of directly modeling the sample allelic covariance is that, 
@@
 so inference can theoretically be done using whole genome data.
-
 \section*{Models, parameters, and priors}\label{model_app}
@@
 because there are three $\alpha$ parameters that describe the continuous differentiation effect of distance in each layer.
-
 \paragraph{Single layer}
@@
 \end{equation}
-
 \paragraph{Priors}
@@
 A table of all parameters, their descriptions, and their priors is given in Table \ref{tab:param_prior_tab}.
-
 \begin{centering}
 \begin{table}
+\ifincludefigs
 \begin{tabular}{| >{\centering\arraybackslash}m{2.1cm} | m{5.2cm} | >{\centering\arraybackslash}m{5.1cm} |}
@@
 \end{tabular}
+\fi
 \caption{
@@
 \end{centering}
-
 \section*{Cross validation}\label{Xvalidation}
-
 We employ Monte Carlo cross-validation approach for model comparison \cite{picard1984}. 
@@
 \end{enumerate}
-
 \section*{Calculating layer contributions}\label{layer_contribution}
-
 Let $A$ and $B$ be randomly chosen alleles from samples $i$ and $j$ respectively,
@@
 which is the genetic divergence.
-
 Now, here is a generative model that gives us the form of the covariance we have postulated.
@@
 \end{aligned}$$
-
 One way to summarize the contribution of each layer is to partition the probability of agreement
@@
 coefficient between $i$ and $j$.
-
 This suggests defining the overall contribution of layer $k$ to agreement, $\xi^{(k)}$, 
@@
 This is the quantity that is plotted in, e.g., Figures \ref{simK1_laycon} and \ref{bear_laycon}.
-
 \section*{Simulation details}\label{sim_details}
@@
 using a cartoon schematic with $K=2$ as an example (Fig \ref{sim_setup}).
-
 \begin{figure}[!ht]
 	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/sims/sim_setup.png}}
+	\fi
 		\caption{Schematic of how we simulate datasets with continuous and discrete differentiation, using $K=2$ as an example.  
@@
 \end{figure}
-
 Using the program \texttt{ms} \cite{Hudson2002}, 
@@
 to nearest neighbors (eight neighbors, including diagonals).
-
 Finally, at time $\tau_{\text{a}}$ in the past 
@@
 as might be expected if a spatial process were mediating admixture between diverged populations.
-
 \section*{Acknowledgements}
-
 We thank Marjorie Weber, Yaniv Brandvain, William Wetzel, 
@@
 \bibliography{reference.library.tex.bib}
-
 %%% BEGIN IF INCLUDE SUPPLEMENT
 \ifincludesupplement
-
 \clearpage
@@
 \setcounter{figure}{0}
-
 % \linespread{1}
-\iffloatsatend
-\processdelayedfloats
-\fi
-
-\clearpage
-
+\clearpage
 \begin{figure}
@@
 \clearpage
-
 \begin{figure}
@@
 \end{figure}
-
 %%% END IF INCLUDE SUPPLEMENT
 \fi
-
+\iffloatsatend
+\renewcommand{\thetable}{\arabic{table}}
+\setcounter{table}{0}
+\renewcommand{\thefigure}{\arabic{figure}}
+\setcounter{figure}{0}
+\processdelayedfloats
+\fi
 \end{document}

### 2017-09-20 writeup/plosgen_prep.R

@@ -1,0 +1,87 @@
+################################################################
+################################################################
+#	ridiculous script that prepares 
+#	paper for plos genetics submission
+################################################################
+################################################################
+
+# Call this script from terminal
+
+setwd("~/Dropbox/conStruct/writeup")
+
+# make sure flags are:
+#	submissionversiontrue
+#	includesupplementtrue
+#	floatsatendfalse
+#	includefigstrue
+call <- c("pdflatex conStruct.tex")
+system(call)
+system(call)
+
+# split out suppmat figs
+call <- c("pdfseparate -f 31 -l 58 conStruct.pdf S%d.pdf")
+system(call)
+
+# rename suppmat figs
+supp.files <- list.files(pattern="^S")
+for(i in 31:58){
+	file.rename(paste0("S",i,".pdf"),paste0("S",i-30,"_fig.pdf"))
+}
+
+# read in text, switch flags, and recompile
+text <- scan("conStruct.tex",what=character(),sep="\n")
+new.text <- text
+new.text[grepl("\\includesupplementtrue",text)] <- "%\\includesupplementtrue"
+new.text[grepl("\\includesupplementfalse",text)] <- "\\includesupplementfalse"
+new.text[grepl("\\\\floatsatendfalse",text)] <- "%\\floatsatendfalse"
+new.text[grepl("\\\\floatsatendtrue",text)] <- "\\floatsatendtrue"
+	cat(new.text,file="conStruct.tex",sep="\n")
+
+call <- c("pdflatex conStruct.tex")
+system(call)
+system(call)
+system(call)
+
+# split out main text figs and tables
+call <- c("pdfseparate -f 26 -l 35 conStruct.pdf Fig%d.pdf")
+system(call)
+
+call <- c("pdfseparate -f 37 -l 37 conStruct.pdf Table%d.pdf")
+system(call)
+
+# rename main text figs and tables
+fig.files <- list.files(pattern="^Fig")
+for(i in 1:length(fig.files)){
+	file.rename(fig.files[i],paste0("Fig",i,".pdf"))
+}
+
+table.files <- list.files(pattern="^Table")
+for(i in 1:length(table.files)){
+	file.rename(table.files[i],paste0("Table",i,".pdf"))
+}
+
+# prepare file to make submission pdf
+#	no floats at end, no figs (but still captions),
[…]

### 2017-09-25 code/conStruct/R/model.comparison.R
fixed bug in x.validation

@@ -60,10 +60,10 @@ x.validation <- function(train.prop=0.9,n.reps,K,freqs,geoDist,coords,prefix,n.i
 										 prefix,
 										 n.iter,
 										 make.figs,
-										 save.files)
+										 save.files) ; 
 					})
 	names(x.val) <- paste0("rep_",1:n.reps)
-	x.val <- standardize.xvals(x.val)
+	x.val <- lapply(x.val,standardize.xvals)
 	return(x.val)
 }
 
@@ -247,7 +247,9 @@ x.validation.rep <- function(rep.no,train.prop,K,freqs,geoDist,coords,prefix,n.i
 	names(test.lnl.nsp) <- K
 	test.lnl <- list("sp" = test.lnl.sp,
 					 "nsp" = test.lnl.nsp)
-	save(test.lnl,file=paste0(prefix,"rep",rep.no,"_test.lnl.Robj"))
+	if(save.files){
+		save(test.lnl,file=paste0(prefix,"rep",rep.no,"_test.lnl.Robj"))
+	}
 	return(test.lnl)
 }
 

### 2017-10-02 code/conStruct/R/data.R
added format-data vignette

@@ -1,0 +1,19 @@
+#' Example dataset used in a \code{conStruct} analysis
+#' 
+#' A simulated dataset containing the allele frequency 
+#' and sampling coordinate data necessary to run a 
+#' \code{conStruct} analysis.
+#' 
+#' @format A list with two elements:
+#' \describe{
+#'		\item{allele.frequencies}{a matrix with one row for each of 
+#'			the 36 samples and one column for each of 10,000 loci, 
+#'			giving the frequency of the counted allele at each locus 
+#'			in each sample}
+#'		\item{coords}{a matrix with one row for each of the 36 samples, 
+#'			in the same order as that of the allele frequency matrix, 
+#'			and two columns, the first giving the x-coordinate 
+#'			(or longitude), the second giving the y-coordinate (or latitude)}
+#' }
+#
+"conStruct.data"

### 2017-10-02 code/conStruct/R/format.data.R

@@ -1,0 +1,89 @@
+#' Convert a dataset from STRUCTURE to conStruct format
+#'
+#' \code{structure2conStruct} converts a STRUCTURE dataset 
+#' to conStruct format
+#' 
+#' This function takes a population genetics dataset in 
+#' STRUCTURE format and converts it to conStruct format. 
+#' The STRUCTURE file must have one row per individual 
+#' and two columns per locus, and can only contain 
+#' bi-allelic SNPs.
+#' 
+#' @param infile The name and path of the file in STRUCTURE format 
+#' 			to be converted to \code{conStruct} format. 
+#' @param start.loci The index of the first column in the dataset 
+#'			that contains genotype data.
+#' @param missing.datum The character or value used to denote 
+#' 			missing data in the STRUCTURE dataset (often -9).
+#' @param outfile The name and path of the file containing the 
+#'			\code{conStruct} formatted dataset to be generated 
+#' 			by this function.
+#'
+#' @details This function takes a STRUCTURE format data file and 
+#'		converts it to a \code{conStruct} format data file. 
+#' 		The STRUCTURE dataset should be in the ONEROWPERIND 
+#' 		file format, with one row per individual and two columns 
+#' 		per locus (this function therefore can only be applied to 
+#' 		diploid organisms). The first column of the STRUCTURE dataset 
+#' 		should be individual names. There may be any number of other 
+#' 		columns that contain non-genotype information before the first
+#'		column that contains genotype data, but there can 
+#' 		be no extraneous columns at the end of the dataset, after the 
+#' 		genotype data.  The genotype data should be bi-allelic 
+#'		single nucleotide polymorphisms (SNPs).
+#'		
+#'	@return This function returns an allele frequency data matrix 
+#'		that can be used as the \code{freqs} argument in a conStruct 
+#'		analysis run using \code{\link{conStruct}}.  It also saves 
+#'		this object as an .RData file so that it can be used in 
+#'		future analyses.
+#'		
+#' @export
+structure2conStruct <- function(infile,start.loci,missing.datum,outfile){
+	structure.data <- utils::read.table(infile,header=FALSE,stringsAsFactors=FALSE)
[…]

### 2017-10-09 code/conStruct/R/plot.output.R
updated run-conStruct vignette and fixed minor bugs

@@ -399,10 +399,12 @@ make.all.chain.plots <- function(conStruct.results,chain.no,data.block,prefix,la
 			plot.admix.props(data.block,conStruct.results,layer.colors)
 		}
 	grDevices::dev.off()
-	grDevices::pdf(file=paste0(prefix,"_model.fit.CIs.chain_",chain.no,".pdf"))
-		plot.model.fit.CIs(data.block,conStruct.results)
-	grDevices::dev.off()
-	if(data.block$spatial | data.block$K > 1){
+	if(!is.null(data.block$geoDist)){
+		grDevices::pdf(file=paste0(prefix,"_model.fit.CIs.chain_",chain.no,".pdf"))
+			plot.model.fit.CIs(data.block,conStruct.results)
+		grDevices::dev.off()
+	}
+	if(data.block$spatial | (data.block$K > 1 & !is.null(data.block$geoDist))){
 		grDevices::pdf(file=paste0(prefix,"_layer.cov.curves.chain_",chain.no,".pdf"),width=5,height=5)
 			plot.layer.covariances(data.block,conStruct.results,layer.colors)
 		grDevices::dev.off()

### 2017-10-09 code/conStruct/R/run.conStruct.R
updated run-conStruct vignette and fixed minor bugs

@@ -279,6 +279,9 @@ process.freq.data <- function(freqs){
 	freqs <- drop.missing(freqs)
 	n.loci <- ncol(freqs)
 	obsCov <- calc.covariance(freqs)
+	if(any(is.na(obsCov))){
+		stop("\n\nAfter dropping invariant loci, one or more pairs of samples have no genotyped loci in common, so relatedness between them cannot be assessed.\n\n")
+	}
 	freq.data <- list("freqs" = freqs,
 					  "obsCov" = obsCov,
 					  "n.loci" = n.loci)
@@ -329,8 +332,10 @@ check.call <- function(args){
 	if(class(args[["freqs"]]) != "matrix" | any(args[["freqs"]] > 1,na.rm=TRUE) | any(args[["freqs"]] < 0,na.rm=TRUE)){	
 		stop("\nyou have specified an invalid value for the \"freqs\" argument \n")
 	}
-	if(class(args[["geoDist"]]) != "matrix" | length(unique(dim(args[["geoDist"]]))) > 1 | any(args[["D"]] < 0)){	
-		stop("\nyou have specified an invalid value for the \"geoDist\" argument \n")
+	if(args[["spatial"]]){
+		if(class(args[["geoDist"]]) != "matrix" | length(unique(dim(args[["geoDist"]]))) > 1 | any(args[["geoDist"]] < 0)){	
+			stop("\nyou have specified an invalid value for the \"geoDist\" argument \n")
+		}
 	}
 	if(class(args[["coords"]]) != "matrix" | ncol(args[["coords"]]) > 2){	
 		stop("\nyou have specified an invalid value for the \"coords\" argument \n")

### 2017-10-16 code/conStruct/R/run.conStruct.R
more specifics on geoDist attributes

@@ -14,7 +14,7 @@
 #' @param freqs A \code{matrix} of allele frequencies with one column per 
 #'				locus and one row per sample.
 #' 				Missing data should be indicated with \code{NA}.
-#' @param geoDist A \code{matrix} of geographic distance between samples. 
+#' @param geoDist A full \code{matrix} of geographic distance between samples. 
 #'					If \code{NULL}, user can only run the nonspatial model.
 #' @param coords A \code{matrix} giving the longitude and latitude 
 #'					(or X and Y coordinates) of the samples.
@@ -334,7 +334,7 @@ check.call <- function(args){
 		stop("\nyou have specified an invalid value for the \"freqs\" argument \n")
 	}
 	if(args[["spatial"]]){
-		if(class(args[["geoDist"]]) != "matrix" | length(unique(dim(args[["geoDist"]]))) > 1 | any(args[["geoDist"]] < 0)){	
+		if(class(args[["geoDist"]]) != "matrix" | length(unique(dim(args[["geoDist"]]))) > 1 | any(args[["geoDist"]] < 0) | !isSymmetric(args[["geoDist"]])){	
 			stop("\nyou have specified an invalid value for the \"geoDist\" argument \n")
 		}
 	}

### 2017-11-29 writeup/response2reviews.tex
response 2 reviewers

\documentclass[11pt]{letter}
\usepackage[hidelinks]{hyperref}
\usepackage[green]{msuletterhead}
\usepackage[usenames,dvipsnames]{xcolor}
\pagenumbering{gobble}

\newcommand{\gb}[1]{{\bf\color{black}{#1}}}
\newcommand{\plr}[1]{{\it\color{purple}{(#1)}}}
\newcommand{\gc}[1]{{\it\color{blue}{(#1)}}}

\begin{document}

Re-Submission Cover Letter \hfill	\today\\
\emph{PLoS Genetics}
\\\\\\
To the Editor(s),

Please find enclosed the re-submission of our manuscript, 
``Inferring Continuous and Discrete Population Genetic Structure Across Space,"
for re-consideration for publication in PLoS Genetics.

We appreciate the opportunity to submit a revised version of the manuscript.  
The comments provided by the Guest Editor 
and two anonymous reviewers are very helpful, 
and we believe the manuscript is substantially improved as a result 
of their critiques.

Our detailed responses to those comments are inset below (in bold), 
but, in brief, WE HAVE DONE SOME STUFF.

We have also edited the text for 
greater clarity.\\\\
%
We hope that the paper is now acceptable for publication.\\\\
%
Thank you for your consideration,\\
Gideon Bradburd (for all of the authors).

\newpage

\textbf{Guest Editor}\\
We believe the concerns of reviewer 1 are addressable, and would like to see some effort put towards addressing them, specifically benchmarking against non-spatial structure identification methods.

\textbf{Reviewer 1}\\
The paper introduces a model-based method called conStruct to account for isolation-by-distance (IBD) when clustering individuals into populations. IBD is known to generate spurious clusters and the proposed software package can avoid spurious cluster detection. Accounting for space in the proposed model based framework amounts at introducing a regularization penalty when fitting admixture model. Regularization explains why overfitting is avoided and correct admixture proportions are found even when performing model-based clustering with a value of K that is too large.

This being said, I have serious concerns about the submitted manuscript.

\begin{itemize}
\item First, many results concern comparisons between clustering based on the spatial and non-spatial models of construct (Figs 2, 3, 4, 6, 7, 9). However the authors should rather compare their spatial approach to existing software for model-based clustering, which do not account for space (Structure, fastSTRUCTURE, ADMIXTURE,...). Results obtained with one of these software should not be available in supplementary material but should rather be visible in the main text.\\\\
\gb{}\\

\item Second, the simulation scenario should correspond to a real biological scenario. Here this layer-based simulation scenario is very difficult to understand. I would be more interesting to stick to more classic spatially explicit simulations such as range expansion, or secondary contact scenarios.\\\\
\gb{}\\

\item Third, I think that the main point of the paper should be more explicit. Non-spatial model based approaches are prone to overfitting when K is too large and this can be avoided when using spatial-based approaches. \\\\
\gb{We disagree that this is the main point of the paper.}\\

\item Last, there are other software that account for space when performing model-based clustering and they should be included in the comparisons (BAPS, Geneland, TESS,...). Right now, they are only mentioned in the introduction and we would like to see results comparing these different model-based approaches.\\\\
\gb{
The methods mentioned here --- BAPS, Geneland, and TESS --- 
place spatial priors on admixture proportions
and/or cluster membership, 
but like STRUCTURE, they assume that allele frequencies 
within a cluster are constant across space.
Therefore, although spatially explicit, this parameterization does not 
account for space in a way that avoids the problem we describe in the paper: 
the characterization of spatially continuous differentiation 
as clines of admixture across spurious clusters.
STRUCTURE is the canonical model-based clustering algorithm
(having been cited $>$ 20,000 times), but is too slow to be run 
on the modern-scale SNP datasets we simulate, 
so fastSTRUCTURE, which implements an extremely similar model 
within a faster inference framework, 
provides the best comparison to our method.}\\

\item In the introduction, the authors write ``placing spatial priors on cluster memberships'' when referring to other methods that deal with ancestry proportions instead of cluster memberships.\\\\
\gb{We have edited our wording for clarity.}\\

\item Formula (1) about the covariance is difficult to understand. It does not correspond to a classic definition of a covariance. Does this correspond to the covariance of some random variables? The discussion following the definition of the covariance looks like a discussion and should be moved to the discussion.\\\\
\gb{}\\

\end{itemize}

\textbf{Reviewer 2}\\
In this paper, Bradburd et al. present a novel analysis method for the joint characterization of discrete and continuous structure. Their model uses a mixed membership Gaussian process, adding a within layer (cluster) Isolation-by-distance-process. I do think that this paper is very clearly written, and presents the model and method in great detail. Besides the presentation of the method, a simulation and two application studies are performed, that lead to a comprehensive overview of the method. I would consider this model to be very valuable for a large audience in ecology and evolution, including biogeographers and conservation biologists, with the only drawback being the very high computational requirements. Overall, however I recommend the paper for acceptance with only very minor changes.

\begin{itemize}
\item One issue I am wondering about is whether there are some issues with ``outlier individuals'', that e.g. represent poorly sampled clusters, which might be an issue with the Alaskan/NW-Canadian bears. In particular, consider the case where the two layers were only fitted from the samples from the 48-states. Presumably, the resulting covariance curve would be similar to Fig S26b (assuming the two layer case). If we then added those NW-pacific samples, we might expect them to be most likely added to the cluster with the flatter right tail, even if they were essentially unrelated. This issue can also arise in kriging (Figure 6 in the first version of the preprint of Caye et al (https://www.biorxiv.org/content/biorxiv/early/2016/10/12/080291.full.pdf is the example I am thinking of). Could this interpretation be excluded somehow?
\\\\
\gb{}\\

\item Another thing that could be added are some additional simulations in more complex demographic models, such as expansions, temporally/spatially heterogeneous migration, or sparser sampling schemes. However, as the paper is already rather long (at almost 20 single-spaced pages), and the space of scenarios for which the model is relevant is enormous, I hope the authors consider further exploring their model in a follow-up study.\\\\
\gb{}\\

\item 100: Is there a reason that allelic covariance is used instead of genetic covariance?\\\\
\gb{}\\

\item Could supp. figures be sorted according to first mentioning in ms?\\\\
\gb{}\\

\item Fig3: increase label size\\\\
\gb{}\\

\item Last line in paragraph on p. 18: as the derivations are otherwise very detailed and easy to follow, it might be worth adding here that this follows from E[UV] = 4 Cov[A, B]\\\\
\gb{}\\

\item 634: Using ms,
\\\\
\gb{Fixed.}\\

\item Fig S4: Could the colors be made to correspond each other (e.g. panel e vs 3). In general, is the exchangeability of layers a practical concern, or similar to the original structure, does the MCMC never ``switch'' two layer labels?\\\\
\gb{}\\

\item Fig S26: maybe keep Y-axis scale consistent between panels?\\\\
\gb{}\\

\item Cross-Validation: As we all know SNP are not uncorrelated. I am wondering if that could influence the CV, as spurious patterns in the testing data might be partially explained by nearby SNP in the training data. Is that a concern?\\\\
\gb{}\\

\end{itemize}

\end{document}

### 2017-12-04 writeup/conStruct.tex
fixed erratum on populus dataset origins
added back in some linebreaks to the latex source (presentation unchanged)

--- 2017-09-20 writeup/conStruct.tex
@@
 \Yinyang These authors contributed equally to this work.
-
 \end{flushleft}
+
 % Please keep the abstract below 300 words
 \section*{Abstract}
-
 A classic problem in population genetics is the characterization 
@@
 \fi
-
 \newpage
+
 \section*{Introduction}
@@
 \paragraph{Relationship of this model to nonspatial structure models}
-
 A nice feature of our approach is that the model described in Eq.\ \eqref{cross_layer_covariance} 
@@
 \subsection*{Simulations}
-
 To test the method, we first generated data using the coalescent simulator \texttt{ms} \cite{Hudson2002}.
@@
 	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/sims/Fig2_simK1_sp_vs_nsp.pdf}}
+	\fi
 	\caption{
@@
 	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/sims/sim_xvals.pdf}}
+	\fi
 	\caption{
@@
 	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/sims/simK1_laycon_barplots.pdf}}
+	\fi
 		\caption{
@@
 \subsection*{Empirical Applications}
-
 To further demonstrate the utility of this method, 
@@
 \subsubsection*{Poplars}
-
 Trees in the genus \textit{Populus} (poplars, aspens, and cottonwoods) 
@@
 with most sampling concentrated on \tri{} drainages.
-The data were generated using RAD-seq, 
+The data were generated using an Infinium 34K array designed for \tri{} \cite{geraldes_2013}, 
 and showed a strong pattern of bias in allelic dropout 
@@
 	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/populus/pop_sp_results.pdf}}
+	\fi
 	\caption{
@@
 	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/populus/populus_std_xval.pdf}}
+	\fi
 	\caption{
@@
 \subsubsection*{Black bears}
-
 The American black bear, \textit{Ursus americanus}, is endemic to North America
@@
 	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/bears/bears_sp_vs_nsp.pdf}}
+	\fi
 	\caption{
@@
 	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/bears/bear_std_xval.pdf}}
+	\fi
 	\caption{
@@
 	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/bears/bears_laycon_barplots.pdf}}
+	\fi
 	\caption{
@@
 \section*{Discussion}
-
 % The difficulty of dispersing across geography makes
@@
 \section*{Materials and Methods}
-
 % \section*{Appendix}
@@
 \section*{Model rationale} \label{rationale}
+
 \subsection*{Drift, admixture, and space}
@@
 covariances, to justify the form given in the main text.
+
 \paragraph{Drift} 
@@
 \section*{Models, parameters, and priors}\label{model_app}
+
 \paragraph{Spatial versus nonspatial}
@@
 
+\begin{table}
 \begin{centering}
-\begin{table}
+\ifincludefigs
 \begin{tabular}{| >{\centering\arraybackslash}m{2.1cm} | m{5.2cm} | >{\centering\arraybackslash}m{5.1cm} |}
@@
 }\label{tab:param_prior_tab}
+\end{centering}
 \end{table}
-\end{centering}
 
 \section*{Cross validation}\label{Xvalidation}
-
 We employ Monte Carlo cross-validation approach for model comparison \cite{picard1984}. 
@@
 \section*{Calculating layer contributions}\label{layer_contribution}
-
 Let $A$ and $B$ be randomly chosen alleles from samples $i$ and $j$ respectively,
@@
 	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/sims/sim_setup.png}}
+	\fi
 		\caption{Schematic of how we simulate datasets with continuous and discrete differentiation, using $K=2$ as an example.  
@@
 \section*{Acknowledgements}
-
 We thank Marjorie Weber, Yaniv Brandvain, William Wetzel, 
@@
 \clearpage
+
 \bibliography{reference.library.tex.bib}
@@
 \ifincludesupplement
-
 \clearpage
@@
 \setcounter{figure}{0}
-
 % \linespread{1}
-\iffloatsatend
-\processdelayedfloats
-\fi
-
 \clearpage
@@
 \fi
+\iffloatsatend
+\renewcommand{\thetable}{\arabic{table}}
+\setcounter{table}{0}
+\renewcommand{\thefigure}{\arabic{figure}}
+\setcounter{figure}{0}
+\processdelayedfloats
+\fi
 

### 2017-12-04 writeup/conStruct.tex
merged back in old linebreaks

--- 2017-09-20 writeup/conStruct.tex
@@
 %\floatsatendtrue
-
+\newif\ifincludefigs
+%\includefigstrue
+\includefigsfalse
 \iffloatsatend
-\usepackage{endfloat}
+\usepackage[nomarkers]{endfloat}
 \fi
@@
 	\centering
+	\ifincludefigs
 		{\includegraphics[width=\textwidth]{figs/schematic/method_schematic.png}}
+	\fi	
 		\caption{Schematic of our method, using $K=3$ as an example.
@@
 }\label{tab:param_prior_tab}
+\end{centering}
 \end{table}
-\end{centering}
 
@@
 % \linespread{1}
-\iffloatsatend
-\processdelayedfloats
-\fi
+% \iffloatsatend
+% \processdelayedfloats
+% \fi
 

### 2017-12-04 writeup/conStruct.tex
edited covariance description

@@
 or frequencies estimated from pooled sequencing.
-We write $N_i$ for the number of chromosomes sequenced in the $i^\text{th}$ sample.
+From these we compute the \emph{allelic covariance} 
+between samples $i$ and $j$, denoted $\widehat{\Omega}_{i,j}$,
+as the expected covariance of distinct individual alleles 
+chosen from each of the two samples at a random locus.
+More precisely, suppose that we pick a random locus uniformly,
+pick a random ``reference'' allelic state from the two seen at that locus,
+and in each sample draw a two random alleles without replacement,
+recording $X_i = 1$ if the first allele drawn in sample $i$ matches the random reference,
+and $X_i = 0$ otherwise;
+and $\widetilde X_i = 1$ or $0$ for the second allele.
+Then, 
+\begin{equation} \label{eqn:first_allelic_cov}
+    \widehat{\Omega}_{i,j} = \cov[X_i, \widetilde X_j] .
+\end{equation}
+
+Although we describe this as a covariance between individually drawn alleles,
+$\widehat{\Omega}_{i,j}$ is in fact also the covariance between the allele frequencies
+of a randomly chosen allele in samples $i$ and $j$, as long as $i \neq j$.
+The diagonal (where $i=j$) does not have this interpretation, 
+and reflects the covariance between individual alleles drawn from within the population.
+To compute it from allele frequencies, first
+write $N_i$ for the number of chromosomes sequenced in the $i^\text{th}$ sample.
 The \emph{sample frequency} at locus $\ell$ in sample $n$, denoted $f_{n,\ell}$, 
@@
 The choice of allele does not affect subsequent calculations, and so may be arbitrary.
-We then calculate the \emph{allelic covariance} between samples $i$ and $j$, denoted $\widehat{\Omega}_{i,j}$,
-as the expected covariance of distinct individual alleles 
-chosen from each of the two samples at a random locus,
-coding the counted allele as `1' (and the other as `0').
-Concretely, this is
+Then $\widehat \Omega$ can be calculated as a covariance
+(derived in \secref{allelic_cov}):
 \begin{equation}
@@
 where $\delta_{i,j}=1$ if $i=j$ and is 0 otherwise.
-Although we describe this as a covariance between individually drawn alleles,
-$\widehat{\Omega}_{i,j}$ is in fact also the covariance between the allele frequencies
-of a randomly chosen allele in samples $i$ and $j$, as long as $i \neq j$.
-The diagonal (where $i=j$) does not have this interpretation, 
-and reflects the covariance between individual alleles drawn from within the population.
-
 This definition of covariance differs from the usual ``genetic covariance'' \cite{mcvean_genealogical_2009}
 in that (a) we do not subtract locus means,
-and so to make the statistic invariant under choice of reference allele, 
-(b) we randomly choose an allele to count at each locus.
-We discuss the derivation of Eq.\ \eqref{allelic_covariance} further in \secref{allelic_cov}.
+and so to make the statistic insensitive to choice of reference allele, 
+(b) we randomly choose a reference allele at each locus.
 As noted in \cite{EEMS},
-this covariance has a close relationship to the pairwise genetic distance:
-if $\pi_{i,j}$ is the mean density of sites at which random samples from $i$ and $j$ differ at a randomly chosen locus,
-then $\Omega_{i,j} = (1 - 2 \pi_{i,j})/4$.
-Therefore, 
-this allelic covariance is more affected in shape by singleton sites
-than the standard genetic covariance,
-so it may be advisable to filter these prior to analysis
-if they are likely to contain a large percentage of errors \cite{linck_battey2017}.
+this can also be calculated as $\Omega_{i,j} = (1 - 2 \pi_{i,j})/4$,
+where $\pi_{i,j}$ is the genetic distance,
+i.e., mean density of sites at which random samples from $i$ and $j$ differ.
 
@@
 
+\paragraph{Allelic or genetic covariance?}
+The choice of allelic covariance, rather than genetic covariance,
+was motivated by two considerations.
+First, it captures within-population variance in a manner more suited to the modeling approach here,
+which led to better performance on test data.
+Second, it is less affected by sample configuration --
+the genetic covariance is calculated after subtracting the mean from the entire sample,
+which is more strongly affected by densely sampled locations.
+However, note that allelic covariance is more affected by singleton sites
+than the standard genetic covariance,
+so it may be advisable to filter these prior to analysis
+if they are likely to contain a large percentage of errors \cite{linck_battey2017}.
+
 \paragraph{Caveats and considerations}
@@
 \section*{Allelic covariance}\label{allelic_cov}
-To see how to arrive at our empirical variance-covariance expression (Eq.\ \eqref{allelic_covariance})
+To see why equations
+\eqref{eqn:first_allelic_cov} and \eqref{allelic_covariance}
+for the allelic covariance are equivalent,
 pick a random locus and
@@
 but otherwise we let $X=1-A$ and $Y=1-B$.
-We then let $\widehat{\Omega}_{i,j} = \cov[X,Y]$. 
+These are $X_i$ and $\widetilde X_j$ in equation \eqref{eqn:first_allelic_cov},
+so that $\widehat{\Omega}_{i,j} = \cov[X,Y]$. 
 Sampling without replacement makes the statistic insensitive to sample size,

### 2017-12-13 sims/structure/datasets/simK1/exe.fastStr.R
more fastStr runs

@@ -2,18 +2,18 @@ setwd("/Applications/STRUCTURE/fastStructure")
 
 K <- 2
 infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK1/simK1"
-outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK1/simK1_estK",K)
+outfile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK1/simK1_estK"
 call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
 system(call)
 
 K <- 3
 infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK1/simK1"
-outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK1/simK1_estK",K)
+outfile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK1/simK1_estK"
 call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
 system(call)
 
 K <- 4
 infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK1/simK1"
-outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK1/simK1_estK",K)
+outfile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK1/simK1_estK"
 call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
 system(call)

### 2017-12-13 sims/structure/datasets/simK2/exe.fastStr.R

@@ -2,18 +2,18 @@ setwd("/Applications/STRUCTURE/fastStructure")
 
 K <- 2
 infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK2/simK2"
-outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK2/simK2_est",K)
+outfile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK2/simK2_estK"
 call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
 system(call)
 
 K <- 3
 infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK2/simK2"
-outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK2/simK2_estK",K)
+outfile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK2/simK2_estK"
 call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
 system(call)
 
 K <- 4
 infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK2/simK2"
-outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK2/simK2_estK",K)
+outfile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK2/simK2_estK"
 call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
 system(call)

### 2017-12-13 sims/structure/datasets/simK3/exe.fastStr.R

@@ -2,18 +2,18 @@ setwd("/Applications/STRUCTURE/fastStructure")
 
 K <- 2
 infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK3/simK3"
-outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK3/simK3_estK",K)
+outfile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK3/simK3_estK"
 call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
 system(call)
 
 K <- 3
 infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK3/simK3"
-outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK3/simK3_estK",K)
+outfile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK3/simK3_estK"
 call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
 system(call)
 
 K <- 4
 infile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK3/simK3"
-outfile <- paste0("/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK3/simK3_estK",K)
+outfile <- "/Users/bradburd/Dropbox/conStruct/sims/structure/datasets/simK3/simK3_estK"
 call <- paste0("python structure.py -K ",K," --input=",infile," --output=",outfile," --prior=logistic --format=str")
 system(call)

### 2017-12-13 writeup/response2reviews.tex
edited covariance description
minor tweaks

@@
 Results obtained with one of these software should not be available in supplementary material but should rather be visible in the main text.\\\\
-\gb{}\\
+\gb{double axis bringing in marginal likelihood of fastSTRUCTURE for Fig 3, or throw in suppmat}\\
 
@@
 I would be more interesting to stick to more classic spatially explicit simulations such as range expansion, or secondary contact scenarios.\\\\
-\gb{}\\
+\gb{trade off w/ space w/ empirical applications, moving the adprop fit figs into the main text}\\
 
@@
 \\\\
-\gb{We disagree that this is the main point of the paper.}\\
+\gb{}\\
 
@@
 Therefore, although spatially explicit, this parameterization does not 
-account for space in a way that avoids the problem we describe in the paper: 
+account for space in a way that avoids the problem we describe in our manuscript: 
 the characterization of spatially continuous differentiation 
@@
 \item In the introduction, the authors write ``placing spatial priors on cluster memberships'' when referring to other methods that deal with ancestry proportions instead of cluster memberships.\\\\
-\gb{We have edited our wording for clarity.}\\
+\gb{We have edited our wording for clarity (see Lines XX-XX) in the revised manuscript.}\\
 
@@
 The discussion following the definition of the covariance looks like a discussion and should be moved to the discussion.\\\\
-\gb{}\\
+\gb{nice property of being invariant w/r/t who's in the sample and invariant w/r/t choice of major or minor allele
+also we've updated the eqns}\\
 
@@
 \\\\
-\gb{}\\
+\gb{make sure it's in the disco}\\
 
@@
 In general, is the exchangeability of layers a practical concern, or similar to the original structure, does the MCMC never ``switch'' two layer labels?\\\\
-\gb{}\\
+\gb{shoutout label-switching }\\
 
@@
 Is that a concern?\\\\
-\gb{}\\
+\gb{shoutout about df of the Wish for CV, should sample SNPs in blocks}\\
 

### 2017-12-13 writeup/conStruct.tex
minor tweaks

--- 2017-12-04 writeup/conStruct.tex
@@
 \end{flushleft}
-
 % Please keep the abstract below 300 words
@@
 \newpage
-
 \section*{Introduction}
@@
 Spatial information has been incorporated into some of these methods, 
-by, for example, placing spatial priors on cluster memberships \cite{geneland,TESS}, 
+by, for example, placing spatial priors on ancestry proportions \cite{geneland,TESS}, 
 but this does not address the underlying issue that 
@@
 functions as a wrapper around this inference machinery.
-
 \paragraph{Relationship of this model to nonspatial structure models}
@@
 so we take that into account when calculating each layer's contribution to the whole.
-
 % \section*{Results}
-
 \subsection*{Simulations}
@@
 in contrast, additional clusters in the nonspatial analyses continue to contribute substantially.
-
 \begin{figure}
@@
 \end{figure}
-
 \begin{figure}
@@
 \end{figure}
-
 \subsection*{Empirical Applications}
@@
 and a large North American sample of black bears.
-
 \subsubsection*{Poplars}
@@
 as well as in Figs \ref{populus_sp_pies} - \ref{populus_fastStr} in the Supplementary Materials.
-
 \begin{figure}
@@
 and also differed across the replicate analyses.
-
 \subsubsection*{Black bears}
@@
 as well as in Figs \ref{bear_sp_pies} - \ref{bear_fastStr} in the Supplementary Materials.
-
 \begin{figure}
@@
 (e.g., Fig \ref{bear_sp_pies}d vs \ref{bear_nsp_pies}d).
-
 %%%%%%%%%
@@
 in which case a change in discrete population structure across space is comparable to population replacement across time.
-
 \section*{Materials and Methods}
@@
 % \setcounter{section}{0}
-
 \section*{Model rationale} \label{rationale}
-
 \subsection*{Drift, admixture, and space}
@@
 covariances, to justify the form given in the main text.
-
 \paragraph{Drift} 
@@
 so inference can theoretically be done using whole genome data.
-
 \section*{Models, parameters, and priors}\label{model_app}
-
 \paragraph{Spatial versus nonspatial}
@@
 A table of all parameters, their descriptions, and their priors is given in Table \ref{tab:param_prior_tab}.
-
+\begin{centering}
 \begin{table}
-\begin{centering}
 \ifincludefigs
@@
 }\label{tab:param_prior_tab}
+\end{table}
 \end{centering}
-\end{table}
-
 \section*{Cross validation}\label{Xvalidation}
@@
 \end{enumerate}
-
 \section*{Calculating layer contributions}\label{layer_contribution}
@@
 as might be expected if a spatial process were mediating admixture between diverged populations.
-
 \section*{Acknowledgements}
@@
 \fi
-
 \newpage
 \clearpage
-
 \bibliography{reference.library.tex.bib}
-
 %%% BEGIN IF INCLUDE SUPPLEMENT
@@
 \end{figure}
-
 %%% END IF INCLUDE SUPPLEMENT
@@
 \fi
-
 \end{document}

### 2017-12-13 writeup/response2reviews.tex

@@
 The discussion following the definition of the covariance looks like a discussion and should be moved to the discussion.\\\\
-\gb{nice property of being invariant w/r/t who's in the sample and invariant w/r/t choice of major or minor allele
-also we've updated the eqns}\\
+
+\gb{
+    The definition is a covariance of some random variables; we have made this fact more clear.
+    We have also moved some of this discussion to the Discussion (heading ``Allelic or genetic covariance?'').
+}\\
 
@@
 \item 100: Is there a reason that allelic covariance is used instead of genetic covariance?\\\\
-\gb{}\\
+\gb{
+    Good question; we've added a paragraph on this to the Discussion.
+}\\
 

### 2017-12-14 code/conStruct/R/run.conStruct.R
added sample size correction to diag

@@ -14,6 +14,10 @@
 #' @param freqs A \code{matrix} of allele frequencies with one column per 
 #'				locus and one row per sample.
 #' 				Missing data should be indicated with \code{NA}.
+#' @param sample.sizes A \code{vector} with one entry per sample giving the 
+#'				number of chromosomes genotyped in the sample.
+#'				E.g., for a diploid sample, the sample size would be 2; 
+#'				for a population of 10 diploids the sample size would be 20.
 #' @param geoDist A full \code{matrix} of geographic distance between samples. 
 #'					If \code{NULL}, user can only run the nonspatial model.
 #' @param coords A \code{matrix} giving the longitude and latitude 
@@ -105,9 +109,9 @@
 #'
 #' @import rstan
 #' @export
-conStruct <- function(spatial=TRUE,K,freqs,geoDist=NULL,coords,prefix="",n.chains=1,n.iter=1e3,make.figs=TRUE,save.files=TRUE){
+conStruct <- function(spatial=TRUE,K,freqs,sample.sizes,geoDist=NULL,coords,prefix="",n.chains=1,n.iter=1e3,make.figs=TRUE,save.files=TRUE){
 	call.check <- check.call(args <- as.list(environment()))
-	freq.data <- process.freq.data(freqs)
+	freq.data <- process.freq.data(freqs,sample.sizes)
 	data.block <- make.data.block(K,freq.data,coords,spatial,geoDist,temp=NULL)
 		if(save.files){
 			save(data.block,file=paste0(prefix,"_data.block.Robj"))
@@ -269,20 +273,20 @@ drop.missing <- function(freqs){
 	return(freqs)
 }
 
-calc.covariance <- function(freqs){
+calc.covariance <- function(freqs,sample.sizes){
 	x <- t(freqs)
 	allelic.covariance <- stats::cov(x,use="pairwise.complete.obs") - 
 									(1/2) * outer( colMeans(x,na.rm=TRUE), 1-colMeans(x,na.rm=TRUE), "*" ) -
 									(1/2) * outer(1-colMeans(x,na.rm=TRUE), colMeans(x,na.rm=TRUE), "*") + 1/4
-	diag(allelic.covariance) <- 1/4
+	diag(allelic.covariance) <- diag(allelic.covariance) - colMeans(x^2) / (sample.sizes - 1)
 	return(allelic.covariance)
 }
 
-process.freq.data <- function(freqs){
[…]

### 2017-12-14 sims/methods/allelic_covariance.Rmd
allelic covariance method

# Allelic covariance

## Math

Suppose that we draw two alleles from a random biallelic site,
and code these as numeric by picking a random reference allele.
We want to compute the covariance between these numbers.
The alleles are chosen without replacement, if drawn from the same population.

First we describe how to compute the covariance of these quantities
from the means and covariances *without* random choice of reference allele.
Let $A$ and $B$ be the two alleles, and
let $U = 1$ or 0 with probability 1/2 each.
Then define
$$ \begin{aligned}
X &= U A + (1-U) (1-A) \\
Y &= U B + (1-U) (1-B) .
\end{aligned} $$
Notice $\E[X] = \E[Y] = 1/2$, thanks to symmetry.
We want to find
$$\begin{aligned}
\cov[X,Y]  
&= \E[(X-1/2)(Y-1/2)] \\
&= \E[XY] - \E[X]/2 - \E[Y]/2 + 1/4
&= \E[XY] - 1/4 .
\end{aligned}$$
Now, averaging over U,
$$\begin{aligned}
\E[XY] 
&= (1/2) \E[AB] + (1/2) \E[(1-A)(1-B)] \\
&= \E[AB - (A+B)/2 + 1/2],
\end{aligned}
and so $\cov[X,Y] = \E[AB] - (\E[A] + \E[B])/2 + 1/4$
and since $\E[AB] = \cov[A,B] + \E[A] \E[B]$,
and hence $\cov[X,Y] = \cov[A,B] + \E[A] \E[B] - (1/2)(\E[A] + \E[B]) + 1/4$,
we have
$$\begin{aligned}
\cov[X,Y] 
&= \cov[A,B] - (1/2){ \E[A] (1 - \E[B]) + (1-\E[A]) \E[B] } + 1/4 ,
\end{aligned}$$
since 
$\E[A] - \E[A] \E[B] = \E[A] (1 - \E[B])$.

Now, how to compute this?
If $A$ is drawn from a population with allele frequency $F_i$ at this randomly chosen site,
then $\E[A] = F_i$.
If $A$ and $B$ are chosen (independently) from different populations,
and the frequency in $B$'s population is $F_j$,
then we can replace $A$ and $B$ above by $F_i$ and $F_j$
(since that is their conditional expectation), so that if $i \neq j$, then
$$\begin{aligned}
\hat \Omega_{ij}
&= \cov[X,Y] 
&= \cov[F_i,F_j] - (1/2){ \E[F_i] (1 - \E[F_j]) + (1-\E[F_i]) \E[F_j] } + 1/4 .
\end{aligned}$$
What if they are from the same population?
For $AB$ to equal 1, both draws must be reference;
if there are $k$ reference and $n-k$ nonreference, this happens with probability
$(k/n)((k-1)/(n-1)) = f (f-1/n) (1 + 1/(n-1)) = f^2 - (1/(n-1)) f (1-f)$ with $f=k/n$.
Therefore,
$$\begin{aligned}
\cov[A,B] 
    &= \E[AB] - \E[A]\E[B] \\
    &= \E[F_i (F_i - 1/N_i)] - \E[F_i]^2 \\
    &= \var[F_i] - \frac{1}{N_i-1} \E[F_i (1-F_i)] .
\end{aligned}$$
This implies that we can extend the formula above to the diagonal, by including
$$\begin{aligned}
\hat \Omega_{ij}
&= \cov[F_i,F_j] - (1/2){ \E[F_i] (1 - \E[F_j]) + (1-\E[F_i]) \E[F_j] } + 1/4 
    - \delta_{ij} \frac{\E[F_i(1-F_i)]}{N_i-1} .
\end{aligned}$$

## Code

Let's check, shall we?
First, simulate data:
```{r sim_data}
nloci <- 1000
npops <- 3
sample_sizes <- c(4,7,11)
global_freqs <- rbeta(nloci, 0.5, 0.5)
freqs <- matrix(rep(global_freqs, npops), ncol=npops)
freqs[,2] <- pmin(1, pmax(0, freqs + rnorm(nloci, sd=1e-3)))
freqs[,npops] <- pmin(1, pmax(0, freqs + rnorm(nloci, sd=0.5)))
counts <- matrix( rbinom(nloci*npops, size=rep(sample_sizes, each=nloci),
                         prob=rep(global_freqs, npops)), ncol=npops)
freqs <- sweep(counts, 2, sample_sizes, "/")
```

Here's what we think the allelic covariance should be:
```{r analytic}
allelic_cov <- function (x, sample_sizes) {
    allelic_covariance <- ( stats::cov(x,use="pairwise.complete.obs") - 
            (1/2) * outer( colMeans(x,na.rm=TRUE), 1-colMeans(x,na.rm=TRUE), "*" ) -
            (1/2) * outer(1-colMeans(x,na.rm=TRUE), colMeans(x,na.rm=TRUE), "*") + 1/4 )
    diag(allelic_covariance) <- diag(allelic_covariance) - colMeans(x*(1-x))/(sample_sizes-1)
    return(allelic_covariance)
}
analytic <- allelic_cov(freqs, sample_sizes)
```

Here's a function to sample loci:
```{r sample_loci}
sample_loci <- function (n, i, j) {
    # sample n loci from pops i and j without replacement
    loci <- sample(1:nloci, n, replace=TRUE)
    f <- counts[loci,i]/sample_sizes[i]
    A <- rbinom(n, size=rep(1,n), prob=f)
    if (i==j) { 
        f <- (counts[loci,j]-A)/(sample_sizes[j]-1)
    } else {
        f <- counts[loci,j]/sample_sizes[j]
    }
    B <- rbinom(n, size=rep(1,n), prob=f)
    return(cbind(A,B))
}

sample_cov <- function (n) {
    # get the allelic covariance by actually drawing random alleles
    out <- matrix(0, nrow=npops, ncol=npops)
    for (i in 1:npops) {
        for (j in i:npops) {
            AB <- sample_loci(n, i, j)
            out[i,j] <- out[j,i] <- cov(AB[,1], AB[,2])
        }
    }
    return(out)
}

sampling <- lapply(1:10, function (i) sample_cov(100000))
mean_sampling <- sampling[[1]]
for (k in 2:length(sampling)) { mean_sampling <- mean_sampling + sampling[[k]] }
mean_sampling <- mean_sampling/length(sampling)
```

Compare these:
```{r comparison}
analytic
mean_sampling
plot(as.vector(analytic), as.vector(mean_sampling), pch=20, col='red',
        ylim=range(unlist(sampling)), xlab='analytic', ylab='sampling')
for (k in 1:length(sampling)) {
    points(as.vector(analytic), as.vector(sampling[[k]]))
}
abline(0,1)
```

### 2017-12-18 code/conStruct/R/run.conStruct.R
removed sample size effec on diag

@@ -14,10 +14,6 @@
 #' @param freqs A \code{matrix} of allele frequencies with one column per 
 #'				locus and one row per sample.
 #' 				Missing data should be indicated with \code{NA}.
-#' @param sample.sizes A \code{vector} with one entry per sample giving the 
-#'				number of chromosomes genotyped in the sample.
-#'				E.g., for a diploid sample, the sample size would be 2; 
-#'				for a population of 10 diploids the sample size would be 20.
 #' @param geoDist A full \code{matrix} of geographic distance between samples. 
 #'					If \code{NULL}, user can only run the nonspatial model.
 #' @param coords A \code{matrix} giving the longitude and latitude 
@@ -109,9 +105,9 @@
 #'
 #' @import rstan
 #' @export
-conStruct <- function(spatial=TRUE,K,freqs,sample.sizes,geoDist=NULL,coords,prefix="",n.chains=1,n.iter=1e3,make.figs=TRUE,save.files=TRUE){
+conStruct <- function(spatial=TRUE,K,freqs,geoDist=NULL,coords,prefix="",n.chains=1,n.iter=1e3,make.figs=TRUE,save.files=TRUE){
 	call.check <- check.call(args <- as.list(environment()))
-	freq.data <- process.freq.data(freqs,sample.sizes)
+	freq.data <- process.freq.data(freqs)
 	data.block <- make.data.block(K,freq.data,coords,spatial,geoDist,temp=NULL)
 		if(save.files){
 			save(data.block,file=paste0(prefix,"_data.block.Robj"))
@@ -273,20 +269,20 @@ drop.missing <- function(freqs){
 	return(freqs)
 }
 
-calc.covariance <- function(freqs,sample.sizes){
+calc.covariance <- function(freqs){
 	x <- t(freqs)
 	allelic.covariance <- stats::cov(x,use="pairwise.complete.obs") - 
 									(1/2) * outer( colMeans(x,na.rm=TRUE), 1-colMeans(x,na.rm=TRUE), "*" ) -
 									(1/2) * outer(1-colMeans(x,na.rm=TRUE), colMeans(x,na.rm=TRUE), "*") + 1/4
-	diag(allelic.covariance) <- diag(allelic.covariance) - colMeans(x^2) / (sample.sizes - 1)
+	diag(allelic.covariance) <- 0.25
 	return(allelic.covariance)
 }
 
-process.freq.data <- function(freqs,sample.sizes){
[…]

### 2017-12-21 writeup/conStruct.tex
accurate cov diag description

@@
 \\
-Gideon S.\  Bradburd\textsuperscript{1,a},
-Graham M.\  Coop\textsuperscript{2,b,\Yinyang},
-Peter L.\  Ralph\textsuperscript{3,c,\Yinyang},
+Gideon S.\  Bradburd\textsuperscript{1,*},
+Graham M.\  Coop\textsuperscript{2,\Yinyang},
+Peter L.\  Ralph\textsuperscript{3,\Yinyang},
 \\
@@
 \\
-\textbf{a} bradburd@msu.edu; 
-\textbf{b} gmcoop@ucdavis.edu;
-\textbf{c} plr@uoregon.edu
+\bigskip
+\textbf{*}bradburd@msu.edu
 
@@
 by continuous migration and drift, as well as by more discrete events, 
-such as geographic barriers, rapid expansions, bottlenecks, and rare
-long-distance migration. 
+such as rapid expansions, bottlenecks, rare long-distance migration,
+and separation by geographic barriers.
 Both discrete and continuous components
@@
 chosen from each of the two samples at a random locus.
-More precisely, suppose that we pick a random locus uniformly,
+More precisely, suppose that we pick a random bi-allelic locus uniformly,
 pick a random ``reference'' allelic state from the two seen at that locus,
-and in each sample draw a two random alleles without replacement,
-recording $X_i = 1$ if the first allele drawn in sample $i$ matches the random reference,
-and $X_i = 0$ otherwise;
-and $\widetilde X_i = 1$ or $0$ for the second allele.
+and, in each sample, draw one random allele,
+recording $X_i = 1$ if the allele drawn in sample $i$ matches the random reference,
+and $X_i = 0$ otherwise.
 Then, 
 \begin{equation} \label{eqn:first_allelic_cov}
-    \widehat{\Omega}_{i,j} = \cov[X_i, \widetilde X_j] .
+    \widehat{\Omega}_{i,j} = \cov[X_i, X_j] .
 \end{equation}
@@
 $\widehat{\Omega}_{i,j}$ is in fact also the covariance between the allele frequencies
-of a randomly chosen allele in samples $i$ and $j$, as long as $i \neq j$.
-The diagonal (where $i=j$) does not have this interpretation, 
-and reflects the covariance between individual alleles drawn from within the population.
-To compute it from allele frequencies, first
-write $N_i$ for the number of chromosomes sequenced in the $i^\text{th}$ sample.
-The \emph{sample frequency} at locus $\ell$ in sample $n$, denoted $f_{n,\ell}$, 
-is calculated by first arbitrarily choosing one of the observed alleles at locus $\ell$ to count, 
-then dividing the number of observations of that counted allele by the total number of chromosomes genotyped at that locus
-in sample $n$.
+of a randomly chosen allele in samples $i$ and $j$, as long as $i \neq j$
+(we discuss the diagonal of $\widehat{\Omega}_{i,j}$
+further in \secref{allelic_cov}).
+
 The choice of allele does not affect subsequent calculations, and so may be arbitrary.
-Then $\widehat \Omega$ can be calculated as a covariance
+The allelic covariance, $\widehat \Omega$, can be calculated as a covariance
 (derived in \secref{allelic_cov}):
@@
 \widehat{\Omega}_{i,j} = 
-    \frac{1}{L} \sum_{\ell=1}^L \left( (f_{i,\ell}-1/2) (f_{j,\ell}-1/2) - \delta_{i,j} f_{i,\ell}^2 / (N_i-1) \right),
+    \frac{1}{L} \sum_{\ell=1}^L \left( (f_{i,\ell}-1/2) (f_{j,\ell}-1/2) \right).
 \label{allelic_covariance}
 \end{equation}
-where $\delta_{i,j}=1$ if $i=j$ and is 0 otherwise.
 This definition of covariance differs from the usual ``genetic covariance'' \cite{mcvean_genealogical_2009}
@@
 let $A$ and $B$ be randomly drawn alleles at that locus from populations $i$ and $j$ respectively.
-If $i=j$, then these are chosen \emph{without} replacement.
+%If $i=j$, then these are chosen \emph{without} replacement.
 Suppose these are each coded as `0' or `1` (where `0' denotes a reference allele),
@@
 but otherwise we let $X=1-A$ and $Y=1-B$.
-These are $X_i$ and $\widetilde X_j$ in equation \eqref{eqn:first_allelic_cov},
+These are $X_i$ and $X_j$ in equation \eqref{eqn:first_allelic_cov},
 so that $\widehat{\Omega}_{i,j} = \cov[X,Y]$. 
-Sampling without replacement makes the statistic insensitive to sample size,
-and random allele flipping makes it independent of the choice of reference allele.
+%Sampling without replacement makes the statistic insensitive to sample size,
+The random allele flipping makes the value of $\widehat{\Omega}$ 
+independent of the choice of reference allele.
 By conditioning on the flip,
@@
 \end{align}
-The factor of $\delta_{i,j}/(N_i-1)$ comes from the fact that if $i=j$,
-then $\E[AB]$ is the probability that both alleles drawn without replacement are reference,
-which is $f_i^2 n/(n-1)$.
+
+The discussion above and Equations \eqref{eqn:first_allelic_cov} and \eqref{allelic_covariance} 
+give the form of the allelic covariance we model for the off-diagonal elements:
+$\widehat{\Omega}$ for $i \neq j$.
+The diagonal elements of $\widehat{\Omega}$ 
+describe the within-population allelic variance:
+$\var{X_i}$, where, as above, 
+$X_i$ gives the randomly `flipped' coding of the genotypes in sample $i$,
+to average over choice of reference allele.
+The within-population allelic variance in sample $i$ is equivalent 
+to the variance of a series of Bernoulli draws across loci, 
+with a probability of `success' at locus $\ell$ given by 
+the allele frequency in $i$ at $\ell$ ($F_{i,\ell}$), with probability 1/2, 
+and $1-F_{i,\ell}$ with probability 1/2.
+With sufficient data, $\widehat{\Omega} \approx 0.25$ for all samples.
+
+Averaging over choice of reference allele therefore removes some information 
+about factors acting within populations that might otherwise leave signatures in, 
+e.g., the genetic covariance, 
+such as population size, extent of inbreeding, and history of bottlenecks.  
+However, as our model is focused on modeling 
+\emph{co}variances between samples as the outcome of some spatial process, 
+we count this a minor loss.
+
+%The factor of $\delta_{i,j}/(N_i-1)$ comes from the fact that if $i=j$,
+%then $\E[AB]$ is the probability that both alleles drawn without replacement are reference,
+%which is $f_i^2 n/(n-1)$.
+
+%and reflects the covariance between individual alleles drawn from within the population.
+%To compute it from allele frequencies, first
+%write $N_i$ for the number of chromosomes sequenced in the $i^\text{th}$ sample.
+%The \emph{sample frequency} at locus $\ell$ in sample $n$, denoted $f_{n,\ell}$, 
+%is calculated by first arbitrarily choosing one of the observed alleles at locus $\ell$ to count, 
+%then dividing the number of observations of that counted allele by the total number of chromosomes genotyped at that locus
+%in sample $n$.
 

### 2018-01-19 writeup/conStruct.tex
ADMIXTURE results

@@
 \newif\ifincludefigs
-%\includefigstrue
-\includefigsfalse
+\includefigstrue
+%\includefigsfalse
 \iffloatsatend
@@
 For comparison,
-we also analyzed each simulated dataset using fastSTRUCTURE \cite{fastStructure}, using $K=2-4$.
-% inferring admixture under models with $K=2$ through 4 (Figs \ref{fastStr_simK1}--\ref{fastStr_simK3}).
-
-% In Fig \ref{simK1_pies}, we show the results of the dataset simulated with $K=1$, 
-% contrasting the results of the nonspatial model runs with 2--4 layers 
-% with those of spatial model with 2--4 layers.
-% 
-% In Fig \ref{sim_xvals} we show the results of the cross-validation analysis for each of the three simulated datasets.
-% The results of all analyses run on datasets simulated using $K=1$ through 3 are shown in 
-% Figs \ref{simK1_nsp_pies}-\ref{simK3_adprop_fit}.
-% 
-% To confirm that results from our nonspatial model were qualitatively similar to those of 
-% other nonspatial model-based clustering algorithms, 
-% we also analyzed each simulated dataset using fastSTRUCTURE \cite{fastStructure}, 
-% inferring admixture under models with $K=2$ through 4 (Figs \ref{fastStr_simK1}--\ref{fastStr_simK3}).
+we also analyzed each simulated dataset using ADMIXTURE \cite{ADMIXTURE}, using $K=1-7$, 
+and compared models using ADMIXTURE's cross-validation procedure with 50 data ``fold" subsets.
 
@@
 % at $K=4$, all four corners, and so on.
-The fastSTRUCTURE results are qualitatively similar, as shown in Figs \ref{fastStr_simK1}-\ref{fastStr_simK3}.
-Each nonspatial cluster is in truth genetically more similar within itself
+The ADMIXTURE results are qualitatively similar, as shown in Figs \ref{admix_simK1}-\ref{admix_simK3}.
+Each nonspatial cluster is genetically more similar within itself
 than it is to other clusters,
@@
 while predictive accuracy of the nonspatial model increases as subsequent clusters are added.
-
+The same holds true for the ADMIXTURE cross-validation results,
+in which models that have the largest number of clusters are preferred over all other models,
+as shown in Fig \ref{sim_admix_CVerrors}.
 % In contrast, the predictive accuracy of the spatial \texttt{conStruct} model
@@
 resulting in just over 20,200 loci retained for analysis.  
-We then analyzed these data, grouped by drainage, using both the spatial and nonspatial \texttt{conStruct} models, 
-using $K = 1$ through 7,
+We then analyzed these data, grouped by drainage, 
+using both the spatial and nonspatial \texttt{conStruct} models with $K = 1$ through 7,
 and compared these models using cross-validation.
-For comparison, we also ran fastSTRUCTURE \cite{fastStructure} 
-using $K = 1$ through 7.
 The results of all these analyses are shown in 
 Figs \ref{populus_pies} and \ref{populus_xvals}, 
-as well as in Figs \ref{populus_sp_pies} - \ref{populus_fastStr} in the Supplementary Materials.
+as well as Figs \ref{populus_sp_pies} - \ref{populus_laycon} in the Supplementary Materials.
+For comparison, we also ran ADMIXTURE \cite{ADMIXTURE} 
+with $K = 1$ through 7, using 50-fold cross-validation to compare model performance
+(Figs \ref{populus_admixture} and \ref{poplar_admix_CVerrors}).
 
@@
 
-Both nonspatial \texttt{conStruct} and fastSTRUCTURE displayed 
+Both nonspatial \texttt{conStruct} and ADMIXTURE displayed 
 the successive partitioning of space and the clines of admixture seen in the simulation results.
 The details of each were somewhat different 
-(Fig \ref{populus_nsp_pies} vs. Fig \ref{populus_fastStr}), 
-and also differed across the replicate analyses.
+(Fig \ref{populus_nsp_pies} vs. Fig \ref{populus_admixture}), 
+and also differed across the replicate analyses, 
+but each strongly favored the most cluster-rich model ran 
+(Figs \ref{populus_xvals} and \ref{poplar_admix_CVerrors}).
 
@@
 and compared these models using cross-validation.
-We also ran fastSTRUCTURE \cite{fastStructure} on the same dataset, 
-using $K = 2$ through 4.
+We also ran ADMIXTURE \cite{ADMIXTURE} on the same dataset, 
+using $K = 1$ through 7, and comparing models using 
+ADMIXTURE's cross-validation procedure with 50 data ``fold" subsets.
 The results of these analyses are shown in Figs \ref{bear_K3} - \ref{bear_laycon}, 
-as well as in Figs \ref{bear_sp_pies} - \ref{bear_fastStr} in the Supplementary Materials.
+as well as in Figs \ref{bear_sp_pies} - \ref{bear_admix_CVerrors} in the Supplementary Materials.
 
@@
 
-Additional layers in the spatial model have strong statistical support 
+The spatial models have strong statistical support 
 up until around $K=5$ or 6 (Fig \ref{bear_xvals}),
@@
 
-Results from the nonspatial model clearly exhibit the tendency 
-of nonspatial clustering algorithms to describe continuous spatial patterns of divergence 
+Results from the nonspatial model and from the ADMIXTURE analyses 
+clearly exhibit the tendency of nonspatial clustering algorithms 
+to describe continuous spatial patterns of divergence 
 using gradients of admixture between clusters.
@@
 overlays the discrete structure between the Southwest cluster and the Southeast.
-The results from fastSTRUCTURE are are not identical to 
+The results from ADMIXTURE are are not identical to 
 those obtained using the nonspatial model, 
 but they do show the same tendency: 
-e.g., at $K=3$, fastSTRUCTURE splits the westernmost Alaskan samples 
+e.g., at $K=3$ 
+-- the preferred model from the cross-validation 
+analysis shown in Fig \ref{bear_admix_CVerrors} -- 
+ADMIXTURE splits the westernmost Alaskan samples 
 out of the cluster with the eastern samples, 
-and at $K=4$, it splits them into their own cluster entirely (Fig \ref{bear_fastStr}).
+and at $K=4$, it subdivides the eastern cluster 
+into two geographically partitioned groups (Fig \ref{bear_admixture}).
+Interestingly, for the nonspatial model implemented in ADMIXTURE, 
+the preferred model has a smaller $K$ ($K=3$) than that of the 
+spatial model with best cross-validation performance in \texttt{conStruct} ($K=5$ or 6). 
+However, comparing results of the two methods, it is clear that even at $K=3$, 
+ADMIXTURE is invoking clusters to describe what seems to be, 
+and is well described by \texttt{conStruct} as, 
+a continuous spatial pattern of genetic variation. 
+The third cluster (shown in gold in Fig \ref{bear_admixture}b), 
+shows  strong spatial autocorrelation in admixture proportions, 
+as would be expected if it is describing continuous spatial differentiation.
 
@@
 For a relatively small number of samples, 
-\texttt{conStruct} can be much faster than other nonspatial clustering methods.
+\texttt{conStruct} can be much faster than existing nonspatial Bayesian clustering methods.
 On a desktop machine, using a single 4.2 GHz Intel Core i7 processor, 
@@
 On the same machine and dataset, 
-a fastSTRUCTURE run with $K=4$ took 90.2 hours.
+a fastSTRUCTURE \cite{fastStructure} run with $K=4$ took 90.2 hours.
 However, for a large number of samples, 
-fastSTRUCTURE (or ADMIXTURE) would be much faster.
+fastSTRUCTURE might be faster.
+And, for almost any dataset of any size, 
+the maximum likelihood algorithm implemented in ADMIXTURE 
+is also quite a bit faster than \texttt{conStruct}; 
+Running ADMIXTURE on the bear dataset over all values of $K$ from 1 to 7, 
+including 50-fold cross-validation for each value of $K$, took only 6.6 minutes.
 
@@
 describe the within-population allelic variance:
-$\var{X_i}$, where, as above, 
+$\var{\left(X_i\right)}$, where, as above, 
 $X_i$ gives the randomly `flipped' coding of the genotypes in sample $i$,
@@
 
+\gb{If the spatial coordinates of the loci in the genome are known, 
+the training/testing partitioning should be designed to accommodate linkage disequilibrium (LD).
+Loci in strong LD do not represent independent instantiations of the coalescent process, 
+so if loci from a single linkage block are randomly split into both training and testing partitions, 
+the independence of the test in the testing partition will be compromised 
+(e.g., the parameters estimated from the training partition 
+might be describing process heterogeneity or noise 
+in a region of the genome that also has loci included in the testing partition).
+If genomic position or LD between SNPs is known, 
+the best practice for cross-validation is to make sure that 
+no loci in the testing dataset are in strong LD with, 
+or spatially proximate to, loci in the testing dataset.}
+
 \section*{Calculating layer contributions}\label{layer_contribution}
@@
 	\centering
-		{\includegraphics[width=\textwidth]{figs/fastStr/fastStr_simK1_pies.pdf}}
-	\caption{
-	Map of admixture proportions estimated using fastSTRUCTURE \cite{fastStructure} for $K=2$ through 4.
+		{\includegraphics[width=\textwidth]{figs/admixture/admixture_simK1_pies.pdf}}
+	\caption{
+	Map of admixture proportions estimated using ADMIXTURE \cite{ADMIXTURE} for $K=2$ through 7.
 	The data were simulated using one layer with nearest-neighbor symmetric migration between demes.
-    }\label{fastStr_simK1}
-\end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=\textwidth]{figs/fastStr/fastStr_simK2_pies.pdf}}
-	\caption{
-	Map of admixture proportions estimated using fastSTRUCTURE \cite{fastStructure} for $K=2$ through 4.
+	The model with the lowest cross-validation error (i.e., the preferred model) is indicated with an asterisk.
+    }\label{admix_simK1}
+\end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=\textwidth]{figs/admixture/admixture_simK2_pies.pdf}}
+	\caption{
+	Map of admixture proportions estimated using ADMIXTURE \cite{ADMIXTURE} for $K=2$ through 7.
 	The data were simulated using two layers with nearest-neighbor symmetric migration between demes.
-    }\label{fastStr_simK2}
-\end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=\textwidth]{figs/fastStr/fastStr_simK3_pies.pdf}}
-	\caption{
-	Map of admixture proportions estimated using fastSTRUCTURE \cite{fastStructure} for $K=2$ through 4.
+	The model with the lowest cross-validation error (i.e., the preferred model) is indicated with an asterisk.
+    }\label{admix_simK2}
+\end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=\textwidth]{figs/admixture/admixture_simK3_pies.pdf}}
+	\caption{
+	Map of admixture proportions estimated using ADMIXTURE \cite{ADMIXTURE} for $K=2$ through 7.
 	The data were simulated using three layers with nearest-neighbor symmetric migration between demes.
-    }\label{fastStr_simK3}
+	The model with the lowest cross-validation error (i.e., the preferred model) is indicated with an asterisk.
+    }\label{admix_simK3}
+\end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=\textwidth]{figs/admixture/sims_CVerror.pdf}}
+	\caption{
+	ADMIXTURE cross-validation results for data simulated under $K=1$, $K=2$, and $K=3$, 
+	run with $K=1$ through 7 using 50 data ``folds" (\texttt{--cv=50}).  
+	The inset plots zoom in on cross-validation results outlined in the dotted boxes.
+	The preferred model (with the lowest cross-validation error) is highlighted with an asterisk.
+    }\label{sim_admix_CVerrors}
 \end{figure}
@@
 	\centering
-		{\includegraphics[width=\textwidth]{figs/populus/poplar_fastStr_results.pdf}}
+		{\includegraphics[width=\textwidth]{figs/populus/poplar_admixture_results.pdf}}
 	\caption{
 	Maps of admixture proportions estimated for the \textit{Populus} dataset 
-	using fastSTRUCTURE \cite{fastStructure} for $K=2$ through 7.
-    }\label{populus_fastStr}
+	using ADMIXTURE \cite{ADMIXTURE} for $K=2$ through 7.
+    }\label{populus_admixture}
+\end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=\textwidth]{figs/populus/poplars_admixture_CVerror.pdf}}
+	\caption{
+	ADMIXTURE cross-validation results for poplar data, 
+	run with $K=1$ through 7 using 50 data ``folds" (\texttt{--cv=50}).  
+	The preferred model (with the lowest cross-validation error) is highlighted with an asterisk.
+    }\label{poplar_admix_CVerrors}
 \end{figure}
@@
 	\centering
-		{\includegraphics[width=\textwidth]{figs/bears/bear_fastStr_results.pdf}}
-	\caption{
-	Maps of admixture proportions estimated for the black bear dataset 
-	using fastSTRUCTURE \cite{fastStructure} for $K=2$ through 4.
-    }\label{bear_fastStr}
-\end{figure}
+		{\includegraphics[width=\textwidth]{figs/bears/bear_admixture_results.pdf}}
+	\caption{
+	Maps of admixture proportions estimated for the bear dataset 
+	using ADMIXTURE \cite{ADMIXTURE} for $K=2$ through 7.
+    }\label{bear_admixture}
+\end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=\textwidth]{figs/bears/bear_admixture_CVerror.pdf}}
+	\caption{
+	ADMIXTURE cross-validation results for bear data, 
+	run with $K=1$ through 7 using 50 data ``folds" (\texttt{--cv=50}).  
+	The preferred model (with the lowest cross-validation error) is highlighted with an asterisk.
+    }\label{bear_admix_CVerrors}
+\end{figure}
+\clearpage
 

### 2018-01-19 data/bears/admixture/exe.admixture.R

@@ -1,0 +1,5 @@
+for(k in 1:7){
+	infile <- "bears.bed"
+	call <- paste0("admixture --cv=50 ",infile ," ",k)
+	system(call)
+}

### 2018-01-19 data/bears/admixture/format.beardata.for.admixture.R

@@ -1,0 +1,43 @@
+convert.str.to.ped <- function(str.file){
+	str <- read.table(str.file,header=FALSE,stringsAsFactors=FALSE,row.names=1)
+	n.loci <- ncol(str)-1
+	n.haps <- nrow(str)
+	hap.IDs <- row.names(str)
+	pop.IDs <- as.numeric(unlist(lapply(lapply(strsplit(hap.IDs,"_"),"[[",2),function(x){strsplit(x,"\\.")[[1]][[1]]}))[seq(1,n.haps,by=2)])
+	mand.cols <- matrix(0,nrow=n.haps/2,ncol=6)
+	#family
+	mand.cols[,1] <- pop.IDs
+	#ind ID
+	mand.cols[,2] <- unlist(lapply(unique(pop.IDs),function(x){1:length(which(pop.IDs==x))}))
+	genos <- matrix(NA,nrow=n.haps/2,ncol=n.loci*2)
+	genos[,seq(1,n.loci*2,by=2)] <- as.matrix(str[seq(1,n.haps,by=2),2:ncol(str)])
+	genos[,seq(2,n.loci*2,by=2)] <- as.matrix(str[seq(2,n.haps,by=2),2:ncol(str)])
+	genos <- genos + 1
+	genos[which(genos < 0)] <- 0
+	ped <- cbind(mand.cols, genos)
+	return(ped)
+}
+
+make.map.file <- function(ped){
+	#recover()
+	n.snps <- (ncol(ped)-6)/2
+	map <- matrix(NA,nrow=n.snps,ncol=3)
+	#CHR
+	map[,1] <- rep(1,n.snps)
+	#RS
+	map[,2] <- paste0("rs",1:n.snps)
+	#BP
+	map[,3] <- seq(1,n.snps*1e3,length.out=n.snps)
+	return(map)
+}
+
+ped_bears <- convert.str.to.ped(str.file="~/Dropbox/conStruct/data/bears/fastStructure/bears.str")
+map_bears <- make.map.file(ped_bears)
+	write.table(ped_bears,file="~/Dropbox/conStruct/data/bears/admixture/bears.ped",sep="\t",quote=FALSE,row.names=FALSE,col.names=FALSE)
+	write.table(map_bears,file="~/Dropbox/conStruct/data/bears/admixture/bears.map",sep="\t",quote=FALSE,row.names=FALSE,col.names=FALSE)
+call <- "plink --noweb --file bears --make-bed --out bears --map3 --allow-no-sex"
[…]

### 2018-01-19 data/poplars/admixture/exe.admixture.R

@@ -1,0 +1,5 @@
+for(k in 1:7){
+	infile <- "poplar.bed"
+	call <- paste0("admixture --cv=50 ",infile ," ",k)
+	system(call)
+}

### 2018-01-19 data/poplars/admixture/format.poplardata.for.admixture.R

@@ -1,0 +1,42 @@
+convert.str.to.ped <- function(str.file){
+	str <- read.table(str.file,header=FALSE,stringsAsFactors=FALSE,row.names=1)
+	n.loci <- ncol(str)-1
+	n.haps <- nrow(str)
+	hap.IDs <- row.names(str)
+	pop.IDs <- as.numeric(unlist(lapply(lapply(strsplit(hap.IDs,"_"),"[[",2),function(x){strsplit(x,"\\.")[[1]][[1]]}))[seq(1,n.haps,by=2)])
+	mand.cols <- matrix(0,nrow=n.haps/2,ncol=6)
+	#family
+	mand.cols[,1] <- pop.IDs
+	#ind ID
+	mand.cols[,2] <- unlist(lapply(unique(pop.IDs),function(x){1:length(which(pop.IDs==x))}))
+	genos <- matrix(NA,nrow=n.haps/2,ncol=n.loci*2)
+	genos[,seq(1,n.loci*2,by=2)] <- as.matrix(str[seq(1,n.haps,by=2),2:ncol(str)])
+	genos[,seq(2,n.loci*2,by=2)] <- as.matrix(str[seq(2,n.haps,by=2),2:ncol(str)])
+	genos <- genos + 1
+	ped <- cbind(mand.cols, genos)
+	return(ped)
+}
+
+make.map.file <- function(ped){
+	#recover()
+	n.snps <- (ncol(ped)-6)/2
+	map <- matrix(NA,nrow=n.snps,ncol=3)
+	#CHR
+	map[,1] <- rep(1,n.snps)
+	#RS
+	map[,2] <- paste0("rs",1:n.snps)
+	#BP
+	map[,3] <- seq(1,n.snps*1e3,length.out=n.snps)
+	return(map)
+}
+
+ped_poplar <- convert.str.to.ped(str.file="~/Dropbox/conStruct/data/poplars/fastStructure/poplars.str")
+map_poplar <- make.map.file(ped_poplar)
+	write.table(ped_poplar,file="~/Dropbox/conStruct/data/poplars/admixture/poplar.ped",sep="\t",quote=FALSE,row.names=FALSE,col.names=FALSE)
+	write.table(map_poplar,file="~/Dropbox/conStruct/data/poplars/admixture/poplar.map",sep="\t",quote=FALSE,row.names=FALSE,col.names=FALSE)
+call <- "plink --noweb --file poplar --make-bed --out poplar --map3 --allow-no-sex"
+system(call)
[…]

### 2018-01-19 sims/admixture/admixture.aux.funcs.R

@@ -1,0 +1,13 @@
+get.CV.error <- function(Rout.file){
+	log <- scan(Rout.file,what="character",sep="\n")
+	CV.error <- as.numeric(
+					unlist(
+						lapply(
+							strsplit(
+								log[grepl("CV error",log)],
+								": "),
+						"[[",2)
+					)
+				)
+	return(CV.error)
+}

### 2018-01-19 sims/admixture/datasets/simK2/exe.admixture.R

@@ -1,0 +1,5 @@
+for(k in 1:7){
+	infile <- "simK2.bed"
+	call <- paste0("admixture --cv=50 ",infile ," ",k)
+	system(call)
+}

### 2018-01-19 sims/admixture/datasets/simK3/exe.admixture.R

@@ -1,0 +1,5 @@
+for(k in 1:7){
+	infile <- "simK3.bed"
+	call <- paste0("admixture --cv=50 ",infile ," ",k)
+	system(call)
+}

### 2018-01-19 sims/admixture/datasets/str.2.ped.R

@@ -1,0 +1,63 @@
+convert.str.to.ped <- function(str.file){
+	str <- read.table(str.file,header=FALSE,stringsAsFactors=FALSE,row.names=1)
+	n.loci <- ncol(str)-1
+	n.haps <- nrow(str)
+	hap.IDs <- row.names(str)
+	pop.IDs <- as.numeric(unlist(lapply(lapply(strsplit(hap.IDs,"_"),"[[",2),function(x){strsplit(x,"\\.")[[1]][[1]]}))[seq(1,n.haps,by=2)])
+	mand.cols <- matrix(0,nrow=n.haps/2,ncol=6)
+	#family
+	mand.cols[,1] <- pop.IDs
+	#ind ID
+	mand.cols[,2] <- unlist(lapply(unique(pop.IDs),function(x){1:length(which(pop.IDs==x))}))
+	genos <- matrix(NA,nrow=n.haps/2,ncol=n.loci*2)
+	genos[,seq(1,n.loci*2,by=2)] <- as.matrix(str[seq(1,n.haps,by=2),2:ncol(str)])
+	genos[,seq(2,n.loci*2,by=2)] <- as.matrix(str[seq(2,n.haps,by=2),2:ncol(str)])
+	genos <- genos + 1
+	ped <- cbind(mand.cols, genos)
+	return(ped)
+}
+
+make.map.file <- function(ped){
+	n.snps <- (ncol(ped)-6)/2
+	map <- matrix(NA,nrow=n.snps,ncol=3)
+	#CHR
+	map[,1] <- rep(1,n.snps)
+	#RS
+	map[,2] <- paste0("rs",1:n.snps)
+	#BP
+	map[,3] <- seq(1,n.snps*1e3,length.out=1000)
+	return(map)
+}
+
+ped_K1 <- convert.str.to.ped(str.file="~/Dropbox/conStruct/sims/admixture/datasets/simK1.str")
+map_K1 <- make.map.file(ped_K1)
+	write.table(ped_K1,file="~/Dropbox/conStruct/sims/admixture/datasets/simK1/simK1.ped",sep="\t",quote=FALSE,row.names=FALSE,col.names=FALSE)
+	write.table(map_K1,file="~/Dropbox/conStruct/sims/admixture/datasets/simK1/simK1.map",sep="\t",quote=FALSE,row.names=FALSE,col.names=FALSE)
+setwd("simK1")
+call <- "plink --file simK1 --out simK1"
[…]

### 2018-01-19 writeup/figs/admixture_figs.R

@@ -1,0 +1,179 @@
+get.CV.error <- function(Rout.file){
+	log <- scan(Rout.file,what="character",sep="\n")
+	CV.error <- as.numeric(
+					unlist(
+						lapply(
+							strsplit(
+								log[grepl("CV error",log)],
+								": "),
+						"[[",2)
+					)
+				)
+	return(CV.error)
+}
+
+collapse.ind.Q <- function(Q,pop.vec){
+	n.pops <- length(unique(pop.vec))
+	admix.props <- matrix(NA,n.pops,ncol(Q))
+	for(i in 1:n.pops){
+		admix.props[i,] <- colMeans(Q[which(pop.vec==i),,drop=FALSE])
+	}
+	return(admix.props)
+}
+
+pie.plot <- function(Q.file,coords,pop.vec,layer.colors){
+	Q <- read.table(Q.file,stringsAsFactors=FALSE)
+	admix.props <- collapse.ind.Q(Q,pop.vec)
+	K <- ncol(admix.props)
+	N <- length(unique(pop.vec))
+	layer.names <- paste0("layer_",1:K)
+	sample.names <- paste0("sample_", 1:N)
+	color.tab <- caroline::nv(c(layer.colors[1:K]), layer.names)
+	pie.list <- lapply(1:N, function(i) {
+	    caroline::nv(admix.props[i, ], layer.names)
+	})
+	names(pie.list) <- sample.names
+	caroline::pies(pie.list, x0 = coords[, 1], 
+	    y0 = coords[, 2], color.table = color.tab, 
+	    border = "black", radii = 3, xlab = "", ylab = "", 
+	    lty = 1, density = NULL,
+	    xlim = range(coords[,1]) + c(-0.5,0.5),
+	    ylim = range(coords[,2]) + c(-0.5,0.5))
+	box(lwd = 2)
+}
+
+get.adprops <- function(Q.file,pop.vec){
+	Q <- read.table(Q.file,stringsAsFactors=FALSE)
+	admix.props <- collapse.ind.Q(Q,pop.vec)
+	return(admix.props)
+}
+
+get.laycols <- function(K,q.file,N){
+	layer.colors <- c("blue", "red", "goldenrod1", "forestgreen", 
+    		          "darkorchid1", "deepskyblue", "darkorange1", "seagreen2", 
[…]

### 2018-01-19 writeup/figs/bear_figs.R
resubmission figs

@@ -27,7 +27,7 @@ load("~/Dropbox/conStruct/data/bears/bear.dataset.Robj")
 pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bear_sampling_map.pdf",width=6,height=4,pointsize=13)
 	#quartz(width=6,height=4,pointsize=13)
 	par(mar=c(1,1.5,0,0),oma=c(0,0,0,0))
-	map(xlim = range(bear.dataset$sample.coords[,1]) + c(-5,5), ylim = range(bear.dataset$sample.coords[,2])+c(-2,2), col="gray")
+	map(xlim = range(bear.dataset$sample.coords[,1]) + c(-5,5), ylim = range(bear.dataset$sample.coords[,2])+c(-2,2), col="gray",lforce="e")
 	points(bear.dataset$sample.coords,pch=19,cex=1.2)
 	map.axes()
 dev.off()
@@ -56,54 +56,6 @@ for(k in 1:7){
 		names(conStruct.results[[1]]$MAP$layer.params[[j]])[4] <- "phi"
 	}
 	output.list.nsp[[k]] <- conStruct.results
-}
-
-if(FALSE){
-csr1.order <- NULL
-for(k in 2:7){
-	data.block$K <- k
-	csr <- output.list.sp[[k]][[1]]
-	if(k > 2){
-		tmp.csr <- output.list.sp[[k-1]][[1]]
-		csr1.order <- match.layers.x.runs(tmp.csr$MAP$admix.proportions,csr$MAP$admix.proportions,csr1.order)
-	}
-	if(is.null(csr1.order)){
-		csr1.order <- 1:k
-	}
-	pdf(file=paste0("~/Dropbox/conStruct/writeup/figs/bears/bear_sp",k,".pdf"),width=7,height=7)
-		make.bear.redux.result.plot(admix.proportions = csr$MAP$admix.proportions,
-									coords = bear.dataset$sample.coords,
-									lump.dist = 200,
-									layer.colors = layer.colors[order(csr1.order)],
-									layer.order=csr1.order,
-									layout = matrix(c(rep(1,10),rep(2,15)),nrow=5,ncol=5,byrow=TRUE))
-	dev.off()
-}
-
-csr1.order <- NULL
-for(k in 2:7){
-	data.block$K <- k
-	csr <- output.list.nsp[[k]][[1]]
-	if(k==2){
[…]

### 2018-01-19 writeup/figs/fig_funcs.R

@@ -108,7 +108,7 @@ plot.layer.curves <- function(data.block, conStruct.results, layer.cols=NULL,add
     return(invisible("layer covs"))
 }
 
-plot.K.layer.curves <- function(Ks,data.block,output.list,col.mat1=NULL,col.mat2=NULL,output.list.sp=NULL){
+plot.K.layer.curves <- function(Ks,data.block,output.list,col.mat1=NULL,col.mat2=NULL,output.list.sp=NULL,y.range=NULL){
 layer.colors <- c("blue","red","goldenrod1","forestgreen","darkorchid1","deepskyblue","darkorange1","seagreen2","yellow1","black")
 if(is.null(col.mat1)){
 	col.mat1 <- matrix(adjustcolor(1,0.7),data.block$N,data.block$N)
@@ -135,12 +135,14 @@ if(is.null(col.mat2)){
 		}
 				par(mar=c(5,5,1,1))
 			    order.mat <- order(data.block$geoDist)
-			    y.range <- range(c(
-			    			unlist(lapply(1:data.block$K,
-			    							function(k){
-										        csr$MAP$layer.params[[k]]$layer.cov
-							})) + csr$MAP$gamma, 
-							data.block$obsCov))			
+			    if(is.null(y.range)){
+				    y.range <- range(c(
+				    			unlist(lapply(1:data.block$K,
+				    							function(k){
+											        csr$MAP$layer.params[[k]]$layer.cov
+								})) + csr$MAP$gamma, 
+								data.block$obsCov))
+				}
 			    plot(data.block$geoDist[upper.tri(data.block$obsCov, diag = TRUE)], 
 			        data.block$obsCov[upper.tri(data.block$obsCov, diag = TRUE)], 
 			        xlim = range(data.block$geoDist), ylim = y.range,
@@ -156,7 +158,7 @@ if(is.null(col.mat2)){
 }
 
 plot.sim.xvals <- function(dir,n.reps,K,simK,y.lim){
-	recover()
+	#recover()
 	for(n in 1:n.reps){
 		load(sprintf("simK%s_rep%s_test.lnl.Robj",simK,n))
 		assign(paste0("tl",n),test.lnl)
@@ -338,7 +340,7 @@ make.bear.redux.result.plot.multipanel1 <- function(admix.proportions,coords,lum
[…]

### 2018-01-19 writeup/figs/make.all.figs.R

@@ -3,7 +3,7 @@ source("~/Dropbox/conStruct/writeup/figs/sim_figs.R")
 	rm(list=ls()) ; gc()
 source("~/Dropbox/conStruct/writeup/figs/populus_figs.R")
 	rm(list=ls()) ; gc()
-source("~/Dropbox/conStruct/writeup/figs/fastStr_figs.R")
+source("~/Dropbox/conStruct/writeup/figs/admixture_figs.R")
 	rm(list=ls()) ; gc()
 source("~/Dropbox/conStruct/writeup/figs/bear_figs.R")
 	rm(list=ls()) ; gc()

### 2018-01-19 writeup/figs/populus_figs.R

@@ -29,7 +29,7 @@ col.mat1 <- matrix(ifelse(poplar.data$sp.ID=="Populus trichocarpa","forestgreen"
 col.mat2 <- matrix(ifelse(poplar.data$sp.ID=="Populus trichocarpa","forestgreen","black"),byrow=FALSE,length(poplar.data$sp.ID),length(poplar.data$sp.ID))
 pdf(file="~/Dropbox/conStruct/writeup/figs/populus/populus_sampling_map.pdf",width=6,height=5,pointsize=13)
 	par(mar=c(4,4,1,1))
-	map(xlim = range(poplar.data$coords[,1]) + c(-5,5), ylim = range(poplar.data$coords[,2])+c(-2,2), col="gray")
+	map(xlim = range(poplar.data$coords[,1]) + c(-5,5), ylim = range(poplar.data$coords[,2])+c(-2,2), col="gray",lforce="e")
 	points(poplar.data$coords,pch=19,cex=1.2,col=ifelse(poplar.data$sp.ID=="Populus trichocarpa","forestgreen","black"))
 	legend(x="bottomleft",pch=19,col=c("forestgreen","black"),legend=c("P. trichocarpa","P. balsamifera"))
 	map.axes()
@@ -77,7 +77,7 @@ for(k in 2:7){
 	}
 	pdf(file=paste0("~/Dropbox/conStruct/writeup/figs/populus/populus_sp",k,".pdf"),width=5,height=5)
 		par(mar=c(5,5,1,1))
-		map(xlim = range(poplar.data$coords[,1]) + c(-5,5), ylim = range(poplar.data$coords[,2])+c(-2,2), col="gray")
+		map(xlim = range(poplar.data$coords[,1]) + c(-5,5), ylim = range(poplar.data$coords[,2])+c(-2,2), col="gray",lforce="e")
 		map.axes()
 		make.admix.pie.plot(csr$MAP$admix.proportions[,csr1.order],data.block$coords,layer.colors=layer.colors,radii=2.5,add=TRUE)
 		box(lwd=2)
@@ -134,7 +134,7 @@ pdf(file="~/Dropbox/conStruct/writeup/figs/populus/pop_sp_results.pdf",width=12,
 			if(is.null(csr1.order)){
 				csr1.order <- 1:k
 			}
[…]

### 2018-01-19 writeup/figs/sim_figs.R

@@ -42,18 +42,6 @@ for(k in 1:7){
 	output.list.nsp[[k]] <- conStruct.results
 }
 
-if(FALSE){
-plot.sim.pies(data.block = data.block,
-			  K = 7,
-			  output.list = output.list.nsp,
-			  file.name = "~/Dropbox/conStruct/writeup/figs/sims/simK1_nsp_pies_K")
-
-plot.sim.pies(data.block = data.block,
-			  K = 7,
-			  output.list = output.list.sp,
-			  file.name = "~/Dropbox/conStruct/writeup/figs/sims/simK1_sp_pies_K")
-}
-
 pdf(file="~/Dropbox/conStruct/writeup/figs/sims/simK1_nsp_pies.pdf",width=8,height=6.3,pointsize=14)
 	par(mfrow=c(2,3),oma=c(1,0,3.5,0))
 	plot.sim.pies.multipanel(data.block = data.block,
@@ -188,25 +176,47 @@ for(k in 1:7){
 	output.list.nsp[[k]] <- conStruct.results
 }
 
-if(FALSE){
-plot.sim.pies(data.block = data.block,
-			  K = 7,
-			  output.list = output.list.nsp,
-			  file.name = "~/Dropbox/conStruct/writeup/figs/sims/simK2_nsp_pies_K")
-
-plot.sim.pies(data.block = data.block,
-			  K = 7,
-			  output.list = output.list.sp,
-			  file.name = "~/Dropbox/conStruct/writeup/figs/sims/simK2_sp_pies_K")
-}
-
 pdf(file="~/Dropbox/conStruct/writeup/figs/sims/simK2_nsp_pies.pdf",width=8,height=6.3,pointsize=14)
+	K <- 7
+	output.list <- output.list.nsp
+	radii <- 1.7
+	trueK <- 2
 	par(mfrow=c(2,3),oma=c(1,0,3.5,0))
-	plot.sim.pies.multipanel(data.block = data.block,
-							 K = 7,
-							 output.list = output.list.nsp,
-							 radii = 1.7,
-							 mar = c(4,2,1,2),trueK=2)
+	layer.colors <- c("blue","red","goldenrod1","forestgreen","darkorchid1","deepskyblue","darkorange1","seagreen2","yellow1","black")
+	csr1.order <- NULL
[…]

### 2018-01-19 writeup/response2reviews.tex
updated response2reviews

@@
 Our detailed responses to those comments are inset below (in bold), 
-but, in brief, WE HAVE DONE SOME STUFF.
+but, in brief, we have implemented a more explicit comparison between 
+\texttt{conStruct}'s spatial model and existing, mainstream nonspatial model-based clustering methods.
+Specifically, we have analyzed all datasets (simulated and empirical) 
+included in the paper using the method ADMIXTURE (Alexander \textit{et al.} 2009),
+and used ADMIXTURE's $n$-fold cross-validation procedure (Alexander and Lange 2011)
+to compare model fit across values of $K$.
+As with the comparison between the spatial and nonspatial models in \texttt{conStruct}, 
+these results highlight both the tendency of nonspatial models 
+to describe continuous differentiation with multiple, spurious clusters, 
+and also the capacity of the spatial model to avoid this pitfall.
 
-We have also edited the text for 
-greater clarity.\\\\
-%
-We hope that the paper is now acceptable for publication.\\\\
+We have also edited the text for greater clarity, and we hope that the paper is now acceptable for publication.\\\\
 %
@@
 Results obtained with one of these software should not be available in supplementary material but should rather be visible in the main text.\\\\
-\gb{double axis bringing in marginal likelihood of fastSTRUCTURE for Fig 3, or throw in suppmat}\\
+\gb{The utility of the nonspatial model implemented in \texttt{conStruct} 
+is that its performance can be directly compared to that of the spatial model; 
+their cross-validation predictive accuracy values are in the same currency 
+because they use the same Wishart likelihood model.
+So, using it we can say not just that that the nonspatial model prefers 
+$K=X$ while the spatial model prefers $K=Y$, 
+but also, e.g., that the spatial model is preferred over the nonspatial model.
+However, as you note, it's important to compare \texttt{conStruct}'s performance to 
+that of existing nonspatial model-based clustering methods as well, 
+which is why we had also included analyses using fastSTRUCTURE.
+Your point here is well-taken though, because the fastSTRUCTURE analyses, 
+which included no model comparisons, did not provide a direct basis of comparison to \texttt{conStruct}.
+We have therefore analyzed all datasets included in the paper using ADMIXTURE, 
+and used ADMIXTURE's cross-validation procedure to compare model performance across values of $K$.  
+These results are discussed extensively throughout the updated manuscript ``Results" section; 
+however, due to the already extensive length of the manuscript (which Reviewer 2 points out), 
+the results of the ADMIXTURE analyses are given in the Supplementary Materials.}\\
 
@@
 I would be more interesting to stick to more classic spatially explicit simulations such as range expansion, or secondary contact scenarios.\\\\
-\gb{trade off w/ space w/ empirical applications, moving the adprop fit figs into the main text}\\
+\gb{We appreciate that the simulation scenario we ran is somewhat artificial, 
+and that it would be good to test the behavior of the method 
+under a classic ``isolation followed by expansion into a zone of secondary contact" model.
+However, as we discuss in the manuscript in the section ``Simulation details",
+we wanted to simulate datasets that had unambiguous true admixture proportions, 
+so that we could test \texttt{conStruct}'s ability to estimate admixture accurately 
+(shown in SuppMat figures XXX and XXX).
+A more organic simulation, with isolation followed by expansion, 
+secondary contact, and subsequent admixture, 
+would not facilitate that test in a straightforward way. 
+As mentioned by Reviewer 2, there are an enormous number 
+of possible simulations we could run to test \texttt{conStruct}'s performance, 
+but fully exploring them is beyond of the scope of this manuscript, 
+so we intend to pursue those in future work.}\\
 
@@
 \\\\
-\gb{}\\
+\gb{WHAT TO SAY HERE}\\
 
@@
 as clines of admixture across spurious clusters.
-STRUCTURE is the canonical model-based clustering algorithm
+
+\ A comprehensive comparison between \texttt{conStruct} 
+and all other model-based clustering methods is beyond the scope of this paper. 
+As STRUCTURE is the canonical model-based clustering algorithm
 (having been cited $>$ 20,000 times), but is too slow to be run 
 on the modern-scale SNP datasets we simulate, 
-so fastSTRUCTURE, which implements an extremely similar model 
+we feel that ADMIXTURE, which implements an extremely similar model 
 within a faster inference framework, 
@@
 
-\gb{
-    The definition is a covariance of some random variables; we have made this fact more clear.
+\gb{The definition given is a covariance of some random variables; we have made this fact more clear, 
+and we have substantially rewritten the section on the allelic covariance
+and have moved some of the intuition-building discussion to the Materials and Methods.
 We have also moved some of this discussion to the Discussion (heading ``Allelic or genetic covariance?'').
@@
 \\\\
-\gb{make sure it's in the disco}\\
+\gb{This is a great point, and highlights a subtle problem.
+Although we feel that isolation-by-distance is a reasonable model for population differentiation, 
+it is possible that, by implementing a spatial model, 
+we can facilitate ``spurious" layer membership in far-flung samples, 
+who are sampled far enough away from all other samples 
+that their spatial covariance is negligible.  
+This situation is made less likely by the $\phi$ parameter in each layer, 
+which induces some amount of covariance between samples in a layer 
+regardless of their geographic separation.
+However, it is possible that $\phi$ in a layer might be estimated to be $\approx 0$. 
+We've discussed heuristics for highlighting such samples 
+(e.g., highlighting pairs of samples with $>$50\% membership in a given layer,
+ that, given their geographic separation, 
+ the parametric covariance curve in that layer, 
+ and the $\phi$ parameter in that layer, have effectively zero covariance), 
+ but we have not yet implemented or tested these checks.
+We are also currently developing a set of spatial \textit{f}-statistics, 
+analogous to those presented in Reich \emph{et al.} (2009) and discussed further in Peter (2016), 
+which could be useful in highlighting such ``outlier" samples.
+But, as you note, this manuscript is already quite long, 
+so this work will be a part of future projects.
+}\\
 
@@
 However, as the paper is already rather long (at almost 20 single-spaced pages), and the space of scenarios for which the model is relevant is enormous, I hope the authors consider further exploring their model in a follow-up study.\\\\
-\gb{}\\
+\gb{We completely agree, and further explorations of the behavior of the method 
+in different scenarios are being planned now.}\\
 
@@
 \gb{
-    Good question; we've added a paragraph on this to the Discussion.
+    Good question; we've added some more text for this, 
+    both explaining what the allelic covariance is and why we use it in lieu of the genetic covariance,
+    (Lines XXX-XXX).
+    The short answer is that the allelic covariance is insensitive to the choice of reference allele, 
+    and, because it doesn't subtract off means like the genetic covariance, 
+    has a constant value between a pair of samples in the dataset, 
+    regardless of which other samples are included.
 }\\
@@
 \item Could supp. figures be sorted according to first mentioning in ms?\\\\
-\gb{}\\
+\gb{MAKE SURE THIS IS RIGHT AFTER MAKING SUBMISSION MS}\\
 
 \item Fig3: increase label size\\\\
-\gb{}\\
+\gb{Fixed.}\\
 
 \item Last line in paragraph on p. 18: as the derivations are otherwise very detailed and easy to follow, it might be worth adding here that this follows from E[UV] = 4 Cov[A, B]\\\\
-\gb{}\\
+\gb{NOT SURE WHAT FOLLOWS FROM THIS}\\
 
@@
 In general, is the exchangeability of layers a practical concern, or similar to the original structure, does the MCMC never ``switch'' two layer labels?\\\\
-\gb{shoutout label-switching }\\
+\gb{In STAN, we declare the vector of $\phi$ parameters to be an object of class ``positive ordered",
+	meaning that values of phi must all have values greater than zero, 
+	and must be in strictly increasing order.
+	In practice, I don't think that much label-switching would occur within the course of a single MCMC run 
+	(like in STRUCTURE), but inducing a specific order to the $\phi$ parameter makes that less likely still.
+	There can be label switching across independent runs for the same value of $K$, 
+	and of course there's label switching across runs with different numbers of $K$.
+	In the \texttt{conStruct} R package, we've included a function that uses the cluster matching 
+	algorithm described in CLUMPP (Jakobsson and Rosenberg 2007) to match layers across runs.
+	We used that algorithm to keep plotting colors consistent in our figures, including in Fig. XXXX.
+	However, something funny happens in that sequence of analyses between $K=5$ and $K=6$, 
+	where the ranking of cluster matching algorithm gives funny results, 
+	leading to the non-intuitive layer color matching.
+	We've tweaked that one by hand to keep the plotting nice.
+ }\\
 
 \item Fig S26: maybe keep Y-axis scale consistent between panels?\\\\
-\gb{}\\
+\gb{Fixed.}\\
 
@@
 Is that a concern?\\\\
-\gb{shoutout about df of the Wish for CV, should sample SNPs in blocks}\\
+\gb{
+	This is a good point and it definitely could be a concern.
+	We've included a discussion of this in the cross-validation procedure description (Lines XXX-XXX), 
+	as well as a description of the CV procedure that should be used if you know LD between loci.
+	But, we also imagine that, for a while at least, the vast majority of \texttt{conStruct} users 
+	won't have genomic position information for their SNP data.
+}\\
 

### 2018-01-21 writeup/conStruct.tex
edits mostly about the diagonal

@@
 \end{equation}
+Each $X_i$ behaves marginally as a fair coin --
+in particular, $\P\{X_i = 1\} = 1$, so $\widehat{\Omega}_{i,i} = 1/4$ for every $i$ --
+all information enters through \emph{correlations}.
 
@@
 $\widehat{\Omega}_{i,j}$ is in fact also the covariance between the allele frequencies
-of a randomly chosen allele in samples $i$ and $j$, as long as $i \neq j$
-(we discuss the diagonal of $\widehat{\Omega}_{i,j}$
-further in \secref{allelic_cov}).
-
+of a randomly chosen allele in samples $i$ and $j$, as long as $i \neq j$.
 The choice of allele does not affect subsequent calculations, and so may be arbitrary.
-The allelic covariance, $\widehat \Omega$, can be calculated as a covariance
+$\widehat \Omega$, can be calculated as a covariance
 (derived in \secref{allelic_cov}):
@@
 \widehat{\Omega}_{i,j} = 
-    \frac{1}{L} \sum_{\ell=1}^L \left( (f_{i,\ell}-1/2) (f_{j,\ell}-1/2) \right).
+    \frac{1}{L} \sum_{\ell=1}^L (f_{i,\ell}-1/2) (f_{j,\ell}-1/2) \qquad \text{for } i \neq j.
 \label{allelic_covariance}
@@
 (b) we randomly choose a reference allele at each locus.
-As noted in \cite{EEMS},
+As noted in \cite{EEMS}, for $i \neq j$,
 this can also be calculated as $\Omega_{i,j} = (1 - 2 \pi_{i,j})/4$,
@@
 For comparison,
-we also analyzed each simulated dataset using ADMIXTURE \cite{ADMIXTURE}, using $K=1-7$, 
+we also analyzed each simulated dataset using ADMIXTURE \cite{ADMIXTURE} with $K$ between 1 and 7, 
 and compared models using ADMIXTURE's cross-validation procedure with 50 data ``fold" subsets.
@@
 and also differed across the replicate analyses, 
-but each strongly favored the most cluster-rich model ran 
+but each strongly favored the most cluster-rich model
 (Figs \ref{populus_xvals} and \ref{poplar_admix_CVerrors}).
@@
 %If $i=j$, then these are chosen \emph{without} replacement.
-Suppose these are each coded as `0' or `1` (where `0' denotes a reference allele),
+Suppose these are each coded as `0' or `1' (where `0' denotes a reference allele),
 but we randomly ``flip'' this coding, so that we let $X=A$ and $Y=B$ with probability 1/2,
@@
 
-The discussion above and Equations \eqref{eqn:first_allelic_cov} and \eqref{allelic_covariance} 
-give the form of the allelic covariance we model for the off-diagonal elements:
-$\widehat{\Omega}$ for $i \neq j$.
-The diagonal elements of $\widehat{\Omega}$ 
-describe the within-population allelic variance:
-$\var{\left(X_i\right)}$, where, as above, 
-$X_i$ gives the randomly `flipped' coding of the genotypes in sample $i$,
-to average over choice of reference allele.
-The within-population allelic variance in sample $i$ is equivalent 
-to the variance of a series of Bernoulli draws across loci, 
-with a probability of `success' at locus $\ell$ given by 
-the allele frequency in $i$ at $\ell$ ($F_{i,\ell}$), with probability 1/2, 
-and $1-F_{i,\ell}$ with probability 1/2.
-With sufficient data, $\widehat{\Omega} \approx 0.25$ for all samples.
-
+Thanks to averaging over choice of alleles,
+the within-population allelic variance in sample $i$, $\widehat Omega_{i,i}$,
+is the variance of a series of Bernoulli($1/2$) draws across loci,
+and therefore $\widehat{\Omega}_{i,i} = 1/4$ for every sample $i$.
 Averaging over choice of reference allele therefore removes some information 

### 2018-01-21 writeup/response2reviews.tex
tweaks about responses

@@
 \\\\
-\gb{WHAT TO SAY HERE}\\
+Good suggestion; we've added a line to the Abstract to this effect.\\
 
@@
 \item Last line in paragraph on p. 18: as the derivations are otherwise very detailed and easy to follow, it might be worth adding here that this follows from E[UV] = 4 Cov[A, B]\\\\
-\gb{NOT SURE WHAT FOLLOWS FROM THIS}\\
+We have substantially rewritten tihs section.  \\
 

### 2018-01-21 writeup/conStruct.tex
tweaks about responses

@@
 This thereby explicitly addresses the ``clines versus clusters" problem in 
-modeling population genetic variation.
+modeling population genetic variation,
+and avoids problems with overfitting that nonspatial models are prone to.
 The method produces useful descriptions of structure in genetic relatedness

### 2018-01-22 writeup/conStruct.tex
tweaking plr edits

--- 2018-01-21 writeup/conStruct.tex
@@
 This thereby explicitly addresses the ``clines versus clusters" problem in 
-modeling population genetic variation.
+modeling population genetic variation,
+and ameliorates the overfitting that nonspatial models are prone to.
 The method produces useful descriptions of structure in genetic relatedness
@@
 Each $X_i$ behaves marginally as a fair coin --
-in particular, $\P\{X_i = 1\} = 1$, so $\widehat{\Omega}_{i,i} = 1/4$ for every $i$ --
+in particular, $\P\{X_i = 1\} = 1/2$, so $\widehat{\Omega}_{i,i} = 1/4$ for every $i$ --
 all information enters through \emph{correlations}.
@@
 of a randomly chosen allele in samples $i$ and $j$, as long as $i \neq j$.
-The choice of allele does not affect subsequent calculations, and so may be arbitrary.
-$\widehat \Omega$, can be calculated as a covariance
+The choice of allele does not affect subsequent calculations, and so may be arbitrary, 
+and $\widehat \Omega$ can be calculated as 
 (derived in \secref{allelic_cov}):

### 2018-01-22 writeup/response2reviews.tex
tweaking plr edits
ADMIXTURE cv error on simxvalsfig

@@
 
-We have also edited the text for greater clarity, and we hope that the paper is now acceptable for publication.\\\\
+We have also edited the description of the allelic covariance we model,
+both for clarity (in response to helpful suggestions from both reviewers), 
+and to bring the text into agreement with what we had actually implemented in the code.
+
+We hope that the paper is now acceptable for publication.\\\\
 %
+
 Thank you for your consideration,\\
@@
 which is why we had also included analyses using fastSTRUCTURE.
-Your point here is well-taken though, because the fastSTRUCTURE analyses, 
+
+\ Your point here is well-taken though, because the fastSTRUCTURE analyses, 
 which included no model comparisons, did not provide a direct basis of comparison to \texttt{conStruct}.
@@
 These results are discussed extensively throughout the updated manuscript ``Results" section; 
-however, due to the already extensive length of the manuscript (which Reviewer 2 points out), 
-the results of the ADMIXTURE analyses are given in the Supplementary Materials.}\\
+and the model that minimizes the cross-validation error in the ADMIXTURE analyses on simulated data 
+is highlighted in the the simulation cross-validation results figure.
+ONLY KEEP THIS IN IF WE LIKE THE NEW FIG.
+However, due to the already extensive length of the manuscript (which Reviewer 2 points out), 
+the results of the ADMIXTURE analyses are given in the Supplementary Materials.
+}\\
 
@@
 \\\\
-Good suggestion; we've added a line to the Abstract to this effect.\\
+\gb{
+Good suggestion; we've added a line to the Abstract to this effect.
+}\\
 
@@
 If we then added those NW-pacific samples, we might expect them to be most likely added to the cluster with the flatter right tail, even if they were essentially unrelated.
-This issue can also arise in kriging (Figure 6 in the first version of the preprint of Caye et al (https://www.biorxiv.org/content/biorxiv/early/2016/10/12/080291.full.pdf is the example I am thinking of).
+This issue can also arise in kriging (Figure 6 in the first version of the preprint of Caye et al (on bioRxiv).
 Could this interpretation be excluded somehow?

### 2018-01-22 writeup/conStruct.tex
ADMIXTURE cv error on simxvalsfig

@@
 modeling population genetic variation,
-and ameliorates the overfitting that nonspatial models are prone to.
+and remedies the overfitting that nonspatial models are prone to.
 The method produces useful descriptions of structure in genetic relatedness
@@
 in which models that have the largest number of clusters are preferred over all other models,
-as shown in Fig \ref{sim_admix_CVerrors}.
+as shown in Fig \ref{sim_xvals} (orange-red points),
+and in more detail in Fig \ref{sim_admix_CVerrors}.
 % In contrast, the predictive accuracy of the spatial \texttt{conStruct} model

### 2018-01-22 writeup/figs/sim_figs.R

@@ -322,17 +322,6 @@ for(k in 1:7){
 	output.list.nsp[[k]] <- conStruct.results
 }
 
-if(FALSE){
-plot.sim.pies(data.block = data.block,
-			  K = 7,
-			  output.list = output.list.nsp,
-			  file.name = "~/Dropbox/conStruct/writeup/figs/sims/simK3_nsp_pies_K")
-
-plot.sim.pies(data.block = data.block,
-			  K = 7,
-			  output.list = output.list.sp,
-			  file.name = "~/Dropbox/conStruct/writeup/figs/sims/simK3_sp_pies_K")
-}
 
 pdf(file="~/Dropbox/conStruct/writeup/figs/sims/simK3_nsp_pies.pdf",width=8,height=6.3,pointsize=14)
 	par(mfrow=c(2,3),oma=c(1,0,3.5,0))
@@ -425,10 +414,15 @@ get.CV.error <- function(Rout.file){
 	return(CV.error)
 }
 
+CV.error1 <- get.CV.error(Rout.file="~/Dropbox/conStruct/sims/admixture/datasets/simK1/exe.admixture.Rout")
+CV.error2 <- get.CV.error(Rout.file="~/Dropbox/conStruct/sims/admixture/datasets/simK2/exe.admixture.Rout")
+CV.error3 <- get.CV.error(Rout.file="~/Dropbox/conStruct/sims/admixture/datasets/simK3/exe.admixture.Rout")
+
+
 pdf(file="~/Dropbox/conStruct/writeup/figs/sims/sim_xvals.pdf",width=14,height=5,pointsize=20)
 #	quartz(width=14,height=5)
 #K1
-setwd("~/Dropbox/gid_runs/mc_runs/sims/simK1")
+setwd("~/Dropbox/conStruct/sims/cross_validation/K_1")
 n.reps <- 10
 K <- 7
 for(n in 1:n.reps){
@@ -441,11 +435,11 @@ xval.CIs <- conStruct:::get.xval.CIs(x.vals.std,K)
 	par(mfrow=c(1,3),mar=c(4.5,4.5,4,1))
 	plot.xval.CIs(xval.CIs,K,xlim=c(0.8,7.2),jitter=0.15)
 		mtext("Predictive accuracy",side=2,padj=-3)
[…]

### 2018-01-22 writeup/response2reviews.tex
minor edits

@@
 \item Last line in paragraph on p. 18: as the derivations are otherwise very detailed and easy to follow, it might be worth adding here that this follows from E[UV] = 4 Cov[A, B]\\\\
-We have substantially rewritten tihs section.  \\
+We have substantially rewritten this section for greater clarity.\\
 

### 2018-01-22 writeup/conStruct.tex
minor edits

@@
 in which models that have the largest number of clusters are preferred over all other models,
-as shown in Fig \ref{sim_xvals} (orange-red points),
-and in more detail in Fig \ref{sim_admix_CVerrors}.
+as shown in Fig \ref{sim_xvals} (vermilion points),
+and, in more detail, in Fig \ref{sim_admix_CVerrors}.
 % In contrast, the predictive accuracy of the spatial \texttt{conStruct} model
@@
     The spatial model shows better model fit at every value of $K$.
+    The ADMIXTURE model with the lowest cross-validation error 
+    (i.e., the preferred model) is highlighted with an vermilion diamond.
+    In all simulations, the preferred ADMIXTURE model 
+    was that with the largest number of clusters.
     }\label{sim_xvals}
@@
 the preferred model has a smaller $K$ ($K=3$) than that of the 
-spatial model with best cross-validation performance in \texttt{conStruct} ($K=5$ or 6). 
+spatial models with best cross-validation performance in \texttt{conStruct} ($K=5$ or 6). 
 However, comparing results of the two methods, it is clear that even at $K=3$, 
@@
 a continuous spatial pattern of genetic variation. 
-The third cluster (shown in gold in Fig \ref{bear_admixture}b), 
+The third cluster in the ADMIXTURE analysis at $K=3$ 
+(shown in gold in Fig \ref{bear_admixture}b), 
 shows  strong spatial autocorrelation in admixture proportions, 
@@
 Thanks to averaging over choice of alleles,
-the within-population allelic variance in sample $i$, $\widehat Omega_{i,i}$,
+the within-population allelic variance in sample $i$, $\widehat{Omega_{i,i}}$,
 is the variance of a series of Bernoulli($1/2$) draws across loci,

### 2018-01-24 writeup/conStruct.tex
minor edits to LD section

@@
 
-\gb{If the spatial coordinates of the loci in the genome are known, 
+If the genomic coordinates of the loci are known, 
 the training/testing partitioning should be designed to accommodate linkage disequilibrium (LD).
-Loci in strong LD do not represent independent instantiations of the coalescent process, 
-so if loci from a single linkage block are randomly split into both training and testing partitions, 
-the independence of the test in the testing partition will be compromised 
-(e.g., the parameters estimated from the training partition 
+Loci in strong LD are not inherited independently,
+so if loci from a single linkage block are included in both training and testing partitions, 
+the independence of the test in the testing partition will be compromised,
+and so the parameters estimated from the training partition 
 might be describing process heterogeneity or noise 
-in a region of the genome that also has loci included in the testing partition).
-If genomic position or LD between SNPs is known, 
-the best practice for cross-validation is to make sure that 
+in a region of the genome that also has loci included in the testing partition.
+The best practice for cross-validation is to make sure that 
 no loci in the testing dataset are in strong LD with, 
-or spatially proximate to, loci in the testing dataset.}
+or near to on the genome, loci in the testing dataset.
 

### 2018-01-24 writeup/conStruct.tex
tweak P edits

@@
 no loci in the testing dataset are in strong LD with, 
-or near to on the genome, loci in the testing dataset.
+-- or near on the genome to -- 
+loci in the testing dataset.
 

### 2018-01-24 writeup/conStruct.tex
minor edits to P edits

--- 2018-01-24 writeup/conStruct.tex
@@
 the independence of the test in the testing partition will be compromised,
-and so the parameters estimated from the training partition 
+because the parameters estimated from the training partition 
 might be describing process heterogeneity or noise 
@@
 The best practice for cross-validation is to make sure that 
-no loci in the testing dataset are in strong LD with, 
-or near to on the genome, loci in the testing dataset.
+no loci in the testing dataset are in strong LD with 
+-- or near on the genome to -- 
+loci in the training dataset.
 

### 2018-01-25 writeup/conStruct.tex
added Graham edits

--- 2018-01-24 writeup/conStruct.tex
@@
 modeling population genetic variation,
-and remedies the overfitting that nonspatial models are prone to.
+and remedies some of the overfitting that nonspatial models are prone to.
 The method produces useful descriptions of structure in genetic relatedness
@@
 Spatial information has been incorporated into some of these methods, 
-by, for example, placing spatial priors on ancestry proportions \cite{geneland,TESS}, 
+by, for example, placing spatial priors on cluster membership \cite{geneland,TESS}, 
 but this does not address the underlying issue that 
@@
 More precisely, suppose that we pick a random bi-allelic locus uniformly,
-pick a random ``reference'' allelic state from the two seen at that locus,
+pick a random ``reference'' allelic state from the two alleles seen at that locus,
 and, in each sample, draw one random allele,
@@
 \end{equation}
-Each $X_i$ behaves marginally as a fair coin --
+Because we randomly choose the reference allele, 
+each $X_i$ behaves marginally as a fair coin --
 in particular, $\P\{X_i = 1\} = 1/2$, so $\widehat{\Omega}_{i,i} = 1/4$ for every $i$ --
@@
 
-If the genomic coordinates of the loci are known, 
+\gb{If the spatial coordinates of the loci in the genome are known, 
 the training/testing partitioning should be designed to accommodate linkage disequilibrium (LD).
-Loci in strong LD are not inherited independently,
-so if loci from a single linkage block are included in both training and testing partitions, 
-the independence of the test in the testing partition will be compromised,
-and so the parameters estimated from the training partition 
+Loci in strong LD do not represent independent instantiations of the coalescent process, 
+so if loci from a single linkage block are randomly split into both training and testing partitions, 
+the independence of the test in the testing partition will be compromised 
+(e.g., the parameters estimated from the training partition 
 might be describing process heterogeneity or noise 
-in a region of the genome that also has loci included in the testing partition.
-The best practice for cross-validation is to make sure that 
+in a region of the genome that also has loci included in the testing partition).
+If genomic position or LD between SNPs is known, 
+the best practice for cross-validation is to make sure that 
 no loci in the testing dataset are in strong LD with, 
-or near to on the genome, loci in the testing dataset.
+or spatially proximate to, loci in the testing dataset.}
 

### 2018-01-25 writeup/response2reviews.tex
incorporated graham edits

--- 2018-01-22 writeup/response2reviews.tex
@@
 which included no model comparisons, did not provide a direct basis of comparison to \texttt{conStruct}.
-We have therefore analyzed all datasets included in the paper using ADMIXTURE, 
+There is a model comparison tool implemented in fastSTRUCTURE, 
+but we were unable to get it to run on our datasets.
+We have contacted the fastSTRUCTURE authors 
+and alerted them to this issue, 
+but have yet to hear back from them.
+
+\ We have therefore analyzed all datasets included in the paper using ADMIXTURE, 
 and used ADMIXTURE's cross-validation procedure to compare model performance across values of $K$.  
@@
 \item Last line in paragraph on p. 18: as the derivations are otherwise very detailed and easy to follow, it might be worth adding here that this follows from E[UV] = 4 Cov[A, B]\\\\
-We have substantially rewritten tihs section.  \\
+\gb{
+We have substantially rewritten this section for greater clarity.\\
+}
 
@@
 	However, something funny happens in that sequence of analyses between $K=5$ and $K=6$, 
-	where the ranking of cluster matching algorithm gives funny results, 
+	where the ranking of cluster matching algorithm gives wonky results, 
 	leading to the non-intuitive layer color matching.

### 2018-01-28 writeup/figs/sim_figs.R
tweaked ADMIX sim CV main fig

@@ -422,7 +422,7 @@ CV.error3 <- get.CV.error(Rout.file="~/Dropbox/conStruct/sims/admixture/datasets
 pdf(file="~/Dropbox/conStruct/writeup/figs/sims/sim_xvals.pdf",width=14,height=5,pointsize=20)
 #	quartz(width=14,height=5)
 #K1
-setwd("~/Dropbox/conStruct/sims/cross_validation/K_1")
+setwd("~/Dropbox/conStruct/sims/cross_validation/K_1/x_validation")
 n.reps <- 10
 K <- 7
 for(n in 1:n.reps){
@@ -435,11 +435,13 @@ xval.CIs <- conStruct:::get.xval.CIs(x.vals.std,K)
 	par(mfrow=c(1,3),mar=c(4.5,4.5,4,1))
 	plot.xval.CIs(xval.CIs,K,xlim=c(0.8,7.2),jitter=0.15)
 		mtext("Predictive accuracy",side=2,padj=-3)
-	legend(x="bottomright",pch=c(19,19,5),col=c("blue","green","orangered"),legend=c("spatial","nonspatial","ADMIXTURE\nmin CV error"),cex=1)
+	legend(x="bottomright",pch=c(19,19,5),
+		col=c("blue","green","orangered"),
+			legend=c("spatial","nonspatial",expression("ADMIXTURE best "*italic("K"))),cex=1)
 	mtext(bquote(paste("True ",italic("K")," = 1")),side=3,adj=0.5,padj=-1.5,font=2,cex=1.2)
 		points(which.min(CV.error1)+0.15,xval.CIs$nsp.means[which.min(CV.error1)],pch=5,col="orangered",cex=1.5)
 #K2
-setwd("~/Dropbox/conStruct/sims/cross_validation/K_2")
+setwd("~/Dropbox/conStruct/sims/cross_validation/K_2/x_validation")
 n.reps <- 10
 K <- 7
 for(n in 1:n.reps){
@@ -466,7 +468,7 @@ xval.CIs <- conStruct:::get.xval.CIs(x.vals.std,K)
 	mtext(bquote(paste("True ",italic("K")," = 2")),side=3,adj=0.5,padj=-1.5,font=2,cex=1.2)
 	mtext("number of layers",side=1,padj=3.4)
 #K3
-setwd("~/Dropbox/conStruct/sims/cross_validation/K_3")
[…]

### 2018-01-28 writeup/figs/populus_figs.R
tidied fig script

@@ -63,28 +63,6 @@ for(k in 1:7){
 	output.list.nsp[[k]] <- conStruct.results
 }
 
-if(FALSE){
-csr1.order <- NULL
-for(k in 2:7){
-	data.block$K <- k
-	csr <- output.list.sp[[k]][[1]]
-	if(k > 2){
-		tmp.csr <- output.list.sp[[k-1]][[1]]
-		csr1.order <- match.layers.x.runs(tmp.csr$MAP$admix.proportions,csr$MAP$admix.proportions,csr1.order)
-	}
-	if(is.null(csr1.order)){
-		csr1.order <- 1:k
-	}
-	pdf(file=paste0("~/Dropbox/conStruct/writeup/figs/populus/populus_sp",k,".pdf"),width=5,height=5)
-		par(mar=c(5,5,1,1))
-		map(xlim = range(poplar.data$coords[,1]) + c(-5,5), ylim = range(poplar.data$coords[,2])+c(-2,2), col="gray",lforce="e")
-		map.axes()
-		make.admix.pie.plot(csr$MAP$admix.proportions[,csr1.order],data.block$coords,layer.colors=layer.colors,radii=2.5,add=TRUE)
-		box(lwd=2)
-	dev.off()
-}
-}
-
 pdf(file=paste0("~/Dropbox/conStruct/writeup/figs/populus/populus_sp_layer_covs.pdf"),width=12,height=8,pointsize=14)
 	#quartz(width=12,height=8)
 	layout(matrix(c(1:6),nrow=2,ncol=3,byrow=TRUE))
@@ -237,136 +215,3 @@ pdf(file="~/Dropbox/conStruct/writeup/figs/populus/poplars_admixture_CVerror.pdf
 	points(best,CV.error[best],pch=8,cex=2)
 	legend(x="topright",pch=8,legend="model with minimum CV error")
 dev.off()
-
-if(FALSE){
-csr1.order <- NULL
-for(k in 2:7){
-	data.block$K <- k
-	data.block$spatial <- FALSE
-	csr <- output.list.nsp[[k]][[1]]
-	if(k==2){
-		tmp.csr <- output.list.sp[[k]][[1]]
-		csr1.order <- match.layers.x.runs(tmp.csr$MAP$admix.proportions,csr$MAP$admix.proportions,csr1.order)
-	}
-	if(k > 2){
-		tmp.csr <- output.list.nsp[[k-1]][[1]]
[…]

### 2018-01-28 writeup/response2reviews.tex
prep for resubmission

@@
 is highlighted in the the simulation cross-validation results figure.
-ONLY KEEP THIS IN IF WE LIKE THE NEW FIG.
 However, due to the already extensive length of the manuscript (which Reviewer 2 points out), 
@@
 \item In the introduction, the authors write ``placing spatial priors on cluster memberships'' when referring to other methods that deal with ancestry proportions instead of cluster memberships.\\\\
-\gb{We have edited our wording for clarity (see Lines XX-XX) in the revised manuscript.}\\
+\gb{
+Guillot \emph{et. al}, in the 2005 paper that introduced the method Geneland, 
+refer to this quantity as ``population membership."
+We kept the wording consistent for clarity, but changed ``population" 
+to ``cluster" to keep a distinction between biological and statistical groupings.}\\
 
@@
     both explaining what the allelic covariance is and why we use it in lieu of the genetic covariance,
-    (Lines XXX-XXX).
+    (in the ``Data" subsection of the ``Results" 
+    and the ``Allelic covariance" subsection of ``Materials and Methods").
     The short answer is that the allelic covariance is insensitive to the choice of reference allele, 
@@
 	This is a good point and it definitely could be a concern.
-	We've included a discussion of this in the cross-validation procedure description (Lines XXX-XXX), 
+	We've included a discussion of this in the cross-validation procedure description, 
 	as well as a description of the CV procedure that should be used if you know LD between loci.

### 2018-01-28 writeup/conStruct.tex
prep for resubmission

@@
 modeling population genetic variation,
-and remedies some of the overfitting that nonspatial models are prone to.
+and remedies some of the overfitting to which nonspatial are prone.
 The method produces useful descriptions of structure in genetic relatedness
@@
 in which models that have the largest number of clusters are preferred over all other models,
-as shown in Fig \ref{sim_xvals} (vermilion points),
+as shown in Fig \ref{sim_xvals} (vermilion diamonds),
 and, in more detail, in Fig \ref{sim_admix_CVerrors}.
@@
     The spatial model shows better model fit at every value of $K$.
-    The ADMIXTURE model with the lowest cross-validation error 
-    (i.e., the preferred model) is highlighted with an vermilion diamond.
+    The vermillion diamond indicates the value of $K$ selected 
+    on the basis of lowest cross-validation error among ADMIXTURE models. 
     In all simulations, the preferred ADMIXTURE model 
@@
 Thanks to averaging over choice of alleles,
-the within-population allelic variance in sample $i$, $\widehat{Omega_{i,i}}$,
+the within-population allelic variance in sample $i$, $\widehat{\Omega}_{i,i}$,
 is the variance of a series of Bernoulli($1/2$) draws across loci,
@@
 
-\gb{If the spatial coordinates of the loci in the genome are known, 
+If the spatial coordinates of the loci in the genome are known, 
 the training/testing partitioning should be designed to accommodate linkage disequilibrium (LD).
@@
 no loci in the testing dataset are in strong LD with, 
-or spatially proximate to, loci in the testing dataset.}
+or spatially proximate to, loci in the testing dataset.
 

### 2018-01-28 writeup/plosgen_prep.R

@@ -9,6 +9,11 @@
 
 setwd("~/Dropbox/conStruct/writeup")
 
+# first, run git-latexdiff to make a diff file 
+#	to the initial submission 55053bbd120f68ed66e199bb39ecd30d3686dc19
+call <- c("git-latexdiff --main conStruct.tex 55053bbd120f68ed66e199bb39ecd30d3686dc19 --bibtex --ignore-makefile -o sub1_vs_sub2_diff.pdf")
+system(call)
+
 # make sure flags are:
 #	submissionversiontrue
 #	includesupplementtrue
@@ -19,12 +24,12 @@ system(call)
 system(call)
 
 # split out suppmat figs
-call <- c("pdfseparate -f 31 -l 58 conStruct.pdf S%d.pdf")
+call <- c("pdfseparate -f 31 -l 62 conStruct.pdf S%d.pdf")
 system(call)
 
 # rename suppmat figs
 supp.files <- list.files(pattern="^S")
-for(i in 31:58){
+for(i in 31:62){
 	file.rename(paste0("S",i,".pdf"),paste0("S",i-30,"_fig.pdf"))
 }
 

### 2018-01-28 writeup/conStruct.tex
further prep for resub

@@
 modeling population genetic variation,
-and remedies some of the overfitting to which nonspatial are prone.
+and remedies some of the overfitting to which nonspatial models are prone.
 The method produces useful descriptions of structure in genetic relatedness
@@
 where they can be described by continuous patterns of variation.
-
-%Using simulated data, we show how current approaches can be misleading, 
-%and also demonstrate that our new method works well.  
-%We predict that our method will be an important tool for the analysis of genetic data.
 
@@
 but substantial work remains to be done \cite{barton2002neutral,barton2013modelling}.
-% Isolation by distance has been modeled in both spatially continuous frameworks 
-% in which an individual's parents are found within some radius of that individual's location,
-% and discrete, ``stepping-stone'' frameworks \cite{kimura1964stepping},
-% in which migration occurs only between neighboring demes.
 Given the generality of the circumstances that generate a pattern of isolation by distance, 
@@
 In this paper, 
-% we present a statistical method for simultaneously 
-% modeling continuous and discrete patterns of population structure.
 we model genetic variation in genotyped individuals as 
@@
 
-%they connote a lack of geographical separation within each.
-
 Within each of these layers,
@@
 and is not chosen to match a particular population genetics model. 
-% (Analytical theory \cite{barton2002neutral} predicts the decay should be proportional to $\exp{-D}/\sqrt{D}$.)
 The $\phi^{(k)}$ is a parameter that describes the background covariance within the layer. 
@@
 allele frequencies at each locus are independently drawn for each 
-cluster \cite{STRUCTURE}, while in conStruct's non-spatial model, 
+cluster \cite{STRUCTURE}, while in conStruct's nonspatial model, 
 it is more natural to envision each cluster's allele frequency 
@@
 (see \secref{sim_details}).
-% so that both the frequencies in layers and the layer memberships are spatially autocorrelated. 
 We simulated datasets using $K=1$, 2, and 3 layers;
@@
 and compared predictive performance of the models using cross-validation.
-% To confirm that results from our nonspatial model were qualitatively similar to those of 
-% other nonspatial model-based clustering algorithms, 
 For comparison,
@@
 (depicted in Figs \ref{simK1_pies} and \ref{simK1_nsp_pies}).
-% as $K$ in the model increases, 
-% the sampled area is increasingly divvied up between clusters 
-% with ``centers of mass" in the $K$ geographical extremes of the simulated space.
-% At $K=2$, the clusters are centered in opposite corners on the map; 
-% at $K=3$, in the lower two corners and the top half; 
-% at $K=4$, all four corners, and so on.
 The ADMIXTURE results are qualitatively similar, as shown in Figs \ref{admix_simK1}-\ref{admix_simK3}.
@@
 and, in more detail, in Fig \ref{sim_admix_CVerrors}.
-% In contrast, the predictive accuracy of the spatial \texttt{conStruct} model
-% either plateaus at or declines from the true value of $K$ used to simulate the data 
-% (Figs \ref{sim_xvals}, \ref{simK1_xval}, \ref{simK2_xval}, and \ref{simK3_xval}).
 
@@
         Results for data simulated using $K=1$,
-	showing maps of admixture proportions estimated using the nonspatial model for $K=2$ through 4 (top row)
-	and the spatial model for $K=2$ through 4 (bottom row).
-	% The data were simulated using a single layer with nearest-neighbor symmetric migration between demes.
-    Since there is only a single layer in the simulation, the spatial model accurately depicts the data
+	showing maps of admixture proportions estimated using the nonspatial \texttt{conStruct} model for $K=2$ through 4 (top row)
+	and the spatial \texttt{conStruct} model for $K=2$ through 4 (bottom row).
+    As there is only a single layer in the simulation, the spatial model accurately depicts the data
     in all cases, while the nonspatial model creates spurious clusters.
@@
 making the system well-suited for application of our method.
-% to test an analysis designed to describe
-% how genetic structure is partitioned both discretely across populations (or species)
-% and continuously with geographic distance.
 We organize the results of our analyses around the following questions: 
@@
         as opposed to a single cline of ancestry?)
-    % Given the known history of hybridization between this pair of poplar species, 
-    % is there support for two discrete populations, 
-    % rather than a single cline of ancestry across their overlap zone?
 
@@
         fall along the species boundary (if any),
-        % If there is evidence for discrete population structuring, 
-        % does it fall solely along species boundaries, 
         or is there sub-structuring within species?
@@
         or primary modes of dispersal.
-        % Is there support for a pattern of isolation by distance within
-        % the inferred layers, and, if so, 
-        % does it differ between different layers?
 
@@
 	Maps of admixture proportions estimated for the \textit{Populus} dataset 
-	using the spatial model for $K=2$ through 4 (a-c), 
+	using the spatial \texttt{conStruct} model for $K=2$ through 4 (a-c), 
 	as well as the corresponding layer-specific covariance curves estimated under each model (d-f).
@@
 but most of this signal seems to derive from the \textit{P. trichocarpa} samples:
-% covariance within predominantly \tri{} layers decays more quickly with distance than within the majority \bals{} layer, 
 as seen in Figs \ref{populus_pies}d-f and \ref{populus_sp_layer_covs},
@@
 \cite{WoodingWard1997,Byun1997,Stone2000,Puckett2015},
-%The retreat of the glaciers was relatively recent in units of 
-%black bear generation times \cite{Onorato2004}, 
-%so we might reasonably expect the divergence accrued between refugia 
-%to be reflected in discrete genetic structure between extant populations.
-%Black bears are strong dispersers \cite{Moore2014}, 
-%but they are distributed across most of the continent, 
-%so we may therefore expect them to experience distance-limited 
-%mating opportunities and exhibit a pattern of isolation by distance.
 likely leading to both continuous and discrete patterns of genetic structure.
@@
 	Maps of admixture proportions estimated for the black bear dataset 
-	using the spatial model (left) and the nonspatial model (right) for $K=3$.
+	using the spatial \texttt{conStruct} model (left) 
+	and the nonspatial \texttt{conStruct} model (right) for $K=3$.
 	Pies show mean admixture results across individuals within their diameter, 
@@
 	for the spatial model (left),
-	and for all clusters using the nonspatial model (right).
+	and for all clusters using the nonspatial \texttt{conStruct} model (right).
 	For each value of $K$ along the x-axis, there are an equal number of contributions plotted.
@@
 but additional spatial layers beyond $K=2$ contribute little to total covariance (Fig \ref{bear_laycon}).
-% However, upon examining layer contributions (Fig \ref{bear_laycon}),
-% it is clear that the additional spatial layers beyond $K=2$ are contributing only a very small amount toward total covariance between individuals.
 The locations of admixed individuals 
-% between the two layers provide information about a possible zone of contact and migration between these groups 
-% along the northwestern coast of North America and near the Cascade Range in the northern U.S.
 are consistent with a scenario of postglacial expansion from two refugia, 
@@
 make more detailed interpretations untrustworthy.
-%In addition, our model is not an explicit test of any demographic scenario, 
-%and also uses a dataset that consists solely of nuclear (rather than
-%mitochondrial) DNA. WHY DO WE CARE about BEAR MTDNS
 
@@
 The results from ADMIXTURE are are not identical to 
-those obtained using the nonspatial model, 
+those obtained using the nonspatial \texttt{conStruct} model, 
 but they do show the same tendency: 
@@
 we see strong support for the spatial model over the nonspatial model (Fig \ref{bear_xvals}).
-% indicating the isolation by distance seems to be a prominent feature of the data.
 This pattern may resolve a discrepancy between our results and 
@@
 \section*{Discussion}
-
-% The difficulty of dispersing across geography makes
-% the pattern of isolation by distance ubiquitous in nature.
-% % this is unsurprising, as mating opportunities are distance-limited in most species.
-% % This common pattern can flummox 
-% Commonly used clustering methods
-% must employ a possibly large number of discrete clusters
-% to describe this simple pattern.
-% The resulting spurious clusters can have implications for 
-% downstream analysis and interpretation of genetic data, 
-% including conservation management decisions.
 
@@
 The computationally limiting step is the inversion of the parametric covariance, 
-% actually it is like O(n^2.3) but I don't know which algorithm STAN uses
-% which is an operation of order $O(n^3)$, 
 which scales more than quadratically with the number of samples,
@@
 is also quite a bit faster than \texttt{conStruct}; 
-Running ADMIXTURE on the bear dataset over all values of $K$ from 1 to 7, 
+running ADMIXTURE on the bear dataset over all values of $K$ from 1 to 7, 
 including 50-fold cross-validation for each value of $K$, took only 6.6 minutes.
@@
 we emphasize that 
-% there is no true best value of $K$. 
-% HOW ABOUT: there is no truth.
 the choice of best $K$ is always relative to the data in hand
 and the questions to be answered.
-%This means both that sampling different individuals or loci 
-%from the same populations may yield a different ``best" $K$, 
 From a statistical perspective, unless the data were generated under the model itself,
 the support for larger values of $K$ is likely to increase with increasing amounts of data.
-%(although perhaps only logarithmically).
 In the limit of infinite data, the best value of $K$ 
@@
 may be quite genetically diverged from one another, 
-perhaps especially at loci underlying adaptive traits, % \citep{toews2016plumage}, 
+perhaps especially at loci underlying adaptive traits, 
 and that a \texttt{conStruct} layer may still contain multiple distinct MUs worthy of independent protections.  
@@
 or of implementing an integrated conservation plan across multiple demes within a layer.
-% Conversely, two populations that are easily distinguishable based on abundant genetic data
-% need not be strongly differentiated --
-% population assignment is not a replacement for absolute measures of genetic divergence.
 
@@
 let $A$ and $B$ be randomly drawn alleles at that locus from populations $i$ and $j$ respectively.
-%If $i=j$, then these are chosen \emph{without} replacement.
 Suppose these are each coded as `0' or `1' (where `0' denotes a reference allele),
@@
 so that $\widehat{\Omega}_{i,j} = \cov[X,Y]$. 
-%Sampling without replacement makes the statistic insensitive to sample size,
 The random allele flipping makes the value of $\widehat{\Omega}$ 
@@
 \begin{align} \label{eqn:cov_xy}
-\cov[X,Y] % &= \E[XY] - 1/4 = (1/2)\E[AB + (1-A)(1-B)] - 1/4\\\notag
+\cov[X,Y] 
     &= \E[(A-1/2)(B-1/2)] .
@@
 we count this a minor loss.
-
-%The factor of $\delta_{i,j}/(N_i-1)$ comes from the fact that if $i=j$,
-%then $\E[AB]$ is the probability that both alleles drawn without replacement are reference,
-%which is $f_i^2 n/(n-1)$.
-
-%and reflects the covariance between individual alleles drawn from within the population.
-%To compute it from allele frequencies, first
-%write $N_i$ for the number of chromosomes sequenced in the $i^\text{th}$ sample.
-%The \emph{sample frequency} at locus $\ell$ in sample $n$, denoted $f_{n,\ell}$, 
-%is calculated by first arbitrarily choosing one of the observed alleles at locus $\ell$ to count, 
-%then dividing the number of observations of that counted allele by the total number of chromosomes genotyped at that locus
-%in sample $n$.
 
@@
 is a special case of the spatial model with all $\alpha$ parameters is set to 0.
-% is described in Eq.\ \eqref{admixed_discrete_covariance}.  % that's not the right eq'n
 The nonspatial model therefore has $3K$ fewer parameters than the spatial model,
@@
 
-If the spatial coordinates of the loci in the genome are known, 
+If the genomic coordinates of the loci are known, 
 the training/testing partitioning should be designed to accommodate linkage disequilibrium (LD).
-Loci in strong LD do not represent independent instantiations of the coalescent process, 
-so if loci from a single linkage block are randomly split into both training and testing partitions, 
+Loci in strong LD are not inherited independently,
+so if loci from a single linkage block are included in both training and testing partitions, 
 the independence of the test in the testing partition will be compromised 
-(e.g., the parameters estimated from the training partition 
+because the parameters estimated from the training partition 
 might be describing process heterogeneity or noise 
-in a region of the genome that also has loci included in the testing partition).
-If genomic position or LD between SNPs is known, 
-the best practice for cross-validation is to make sure that 
+in a region of the genome that also has loci included in the testing partition.
+The best practice for cross-validation is to make sure that 
 no loci in the testing dataset are in strong LD with, 
-or spatially proximate to, loci in the testing dataset.
+or near on the genome to, loci in the training dataset.
 
@@
         &= 2 \P\{ U=V \} - 1 \\
-        % &= 1 - 2 \P\{ U \neq V \} .
         &= 2 \P\{ A=B \} - 1 
@@
 	\caption{
-	Map of admixture proportions estimated using a nonspatial model for $K=2$ through 7.
+	Map of admixture proportions estimated using a nonspatial \texttt{conStruct} model for $K=2$ through 7.
 	The data were simulated using one layer with nearest-neighbor symmetric migration between demes.
@@
 	\caption{
-	Map of admixture proportions estimated using a spatial model for $K=2$ through 7.
+	Map of admixture proportions estimated using a spatial \texttt{conStruct} model for $K=2$ through 7.
 	The data were simulated using one layer with nearest-neighbor symmetric migration between demes.
@@
 	\caption{
-	Map of admixture proportions estimated using a nonspatial model for $K=2$ through 7.
+	Map of admixture proportions estimated using a nonspatial \texttt{conStruct} model for $K=2$ through 7.
 	The data were simulated using two layers with nearest-neighbor symmetric migration between demes.
@@
 	\caption{
-	Map of admixture proportions estimated using a spatial model for $K=2$ through 7.
+	Map of admixture proportions estimated using a spatial \texttt{conStruct} model for $K=2$ through 7.
 	The data were simulated using two layers with nearest-neighbor symmetric migration between demes.
@@
 			for the spatial model (left), 
-			and for all clusters using the nonspatial model (right).
+			and for all clusters using the nonspatial \texttt{conStruct} model (right).
 			Data were simulated using $K=2$.
@@
 	\caption{
-	Map of admixture proportions estimated using a nonspatial model for $K=2$ through 7.
+	Map of admixture proportions estimated using a nonspatial \texttt{conStruct} model for $K=2$ through 7.
 	The data were simulated using three layers with nearest-neighbor symmetric migration between demes.
@@
 	\caption{
-	Map of admixture proportions estimated using a spatial model for $K=2$ through 7.
+	Map of admixture proportions estimated using a spatial \texttt{conStruct} model for $K=2$ through 7.
 	The data were simulated using three layers with nearest-neighbor symmetric migration between demes.
@@
 			for the spatial model (left), 
-			and for all clusters using the nonspatial model (right).
+			and for all clusters using the nonspatial \texttt{conStruct} model (right).
 			Data were simulated using $K=3$.
@@
 	The data were simulated using one layer with nearest-neighbor symmetric migration between demes.
-	The model with the lowest cross-validation error (i.e., the preferred model) is indicated with an asterisk.
+	The model with the lowest cross-validation error (i.e., the preferred model) was $K=7$.
     }\label{admix_simK1}
@@
 	The data were simulated using two layers with nearest-neighbor symmetric migration between demes.
-	The model with the lowest cross-validation error (i.e., the preferred model) is indicated with an asterisk.
+	The model with the lowest cross-validation error (i.e., the preferred model) was $K=7$.
     }\label{admix_simK2}
@@
 	The data were simulated using three layers with nearest-neighbor symmetric migration between demes.
-	The model with the lowest cross-validation error (i.e., the preferred model) is indicated with an asterisk.
+	The model with the lowest cross-validation error (i.e., the preferred model) was $K=7$.
     }\label{admix_simK3}
@@
 	Maps of admixture proportions estimated for the \textit{Populus} dataset 
-	using the spatial model for $K=2$ through 7.
+	using the spatial \texttt{conStruct} model for $K=2$ through 7.
     }\label{populus_sp_pies}
@@
 	Maps of admixture proportions estimated for the \textit{Populus} dataset 
-	using the nonspatial model for $K=2$ through 7.
+	using the nonspatial \texttt{conStruct} model for $K=2$ through 7.
     }\label{populus_nsp_pies}
@@
 	for the spatial model (left), 
-	and for all clusters using the nonspatial model (right).
+	and for all clusters using the nonspatial \texttt{conStruct} model (right).
 	For each value of $K$ along the x-axis, there are an equal number of contributions plotted.
@@
 	Map of admixture proportions estimated for the bear dataset 
-	using the spatial model for $K=2$ through 7.
+	using the spatial \texttt{conStruct} model for $K=2$ through 7.
     }\label{bear_sp_pies}
@@
 	Map of admixture proportions estimated for the bear dataset 
-	using the nonspatial model for $K=2$ through 7.
+	using the nonspatial \texttt{conStruct} model for $K=2$ through 7.
     }\label{bear_nsp_pies}

### 2018-01-29 writeup/response2reviews.tex
resubmission

@@
 so that we could test \texttt{conStruct}'s ability to estimate admixture accurately 
-(shown in SuppMat figures XXX and XXX).
+(shown in SuppMat figures 8 and 13).
 A more organic simulation, with isolation followed by expansion, 
@@
 \gb{
-Guillot \emph{et. al}, in the 2005 paper that introduced the method Geneland, 
+Guillot \emph{et al.}, in the 2005 paper that introduced the method Geneland, 
 refer to this quantity as ``population membership."
@@
 \item Could supp. figures be sorted according to first mentioning in ms?\\\\
-\gb{MAKE SURE THIS IS RIGHT AFTER MAKING SUBMISSION MS}\\
+\gb{Fixed.}\\
 
@@
 	algorithm described in CLUMPP (Jakobsson and Rosenberg 2007) to match layers across runs.
-	We used that algorithm to keep plotting colors consistent in our figures, including in Fig. XXXX.
+	We used that algorithm to keep plotting colors consistent in our figures, including in Fig. S4.
 	However, something funny happens in that sequence of analyses between $K=5$ and $K=6$, 

### 2018-01-29 writeup/conStruct.tex
resubmission

@@
 \newif\ifincludefigs
-\includefigstrue
-%\includefigsfalse
+%\includefigstrue
+\includefigsfalse
 \iffloatsatend
@@
 we also analyzed each simulated dataset using ADMIXTURE \cite{ADMIXTURE} with $K$ between 1 and 7, 
-and compared models using ADMIXTURE's cross-validation procedure with 50 data ``fold" subsets.
+and compared models using ADMIXTURE's cross-validation procedure with 50 data fold subsets.
 
@@
 spatial \texttt{conStruct} does not create spurious discrete groupings when there are none:
-Figures \ref{simK1_pies}, \ref{simK2_sp_pies}, and \ref{simK3_sp_pies} 
+Figures \ref{simK1_pies} and \ref{simK1_sp_pies}-\ref{simK3_sp_pies} 
 show that subsequent layers beyond the number used for simulation are unused.
@@
 Moreover, the method infers the true admixture proportions with high accuracy, tight precision, and good coverage 
-(Figs \ref{simK2_adprop_fit} and \ref{simK3_adprop_fit}).
+(Figs \ref{simK2_adprop_fit}-\ref{simK3_adprop_fit}).
 
@@
 to better approximate the continuous, spatial patterns of relatedness
-(depicted in Figs \ref{simK1_pies} and \ref{simK1_nsp_pies}).
+(depicted in Figs \ref{simK1_pies} and \ref{simK1_nsp_pies}-\ref{simK3_nsp_pies}).
 The ADMIXTURE results are qualitatively similar, as shown in Figs \ref{admix_simK1}-\ref{admix_simK3}.
@@
 and then plateaus or declines
-(Figs \ref{sim_xvals}, \ref{simK1_xval}, \ref{simK2_xval}, and \ref{simK3_xval}),
+(Figs \ref{sim_xvals} and \ref{simK1_xval}-\ref{simK3_xval}),
 while predictive accuracy of the nonspatial model increases as subsequent clusters are added.
@@
 The unimportance of spurious layers can be seen in plots of layer contributions
-(Figs \ref{simK1_laycon}, \ref{simK2_laycon}, and \ref{simK3_laycon}).
+(Figs \ref{simK1_laycon}, \ref{simK2_laycon}-\ref{simK3_laycon}).
 In the spatial analyses, once we pass the true $K$, 
@@
 Figs \ref{populus_pies} and \ref{populus_xvals}, 
-as well as Figs \ref{populus_sp_pies} - \ref{populus_laycon} in the Supplementary Materials.
+as well as Figs \ref{populus_sp_pies}-\ref{populus_laycon} in the Supplementary Materials.
 For comparison, we also ran ADMIXTURE \cite{ADMIXTURE} 
 with $K = 1$ through 7, using 50-fold cross-validation to compare model performance
-(Figs \ref{populus_admixture} and \ref{poplar_admix_CVerrors}).
+(Figs \ref{populus_admixture}-\ref{poplar_admix_CVerrors}).
 
@@
 using $K = 1$ through 7, and comparing models using 
-ADMIXTURE's cross-validation procedure with 50 data ``fold" subsets.
-The results of these analyses are shown in Figs \ref{bear_K3} - \ref{bear_laycon}, 
-as well as in Figs \ref{bear_sp_pies} - \ref{bear_admix_CVerrors} in the Supplementary Materials.
+ADMIXTURE's cross-validation procedure with 50 data fold subsets.
+The results of these analyses are shown in Figs \ref{bear_K3}-\ref{bear_laycon}, 
+as well as in Figs \ref{bear_sp_pies}-\ref{bear_admix_CVerrors} in the Supplementary Materials.
 
@@
 	\centering
+		{\includegraphics[width=\textwidth]{figs/sims/simK1_sp_pies.pdf}}
+	\caption{
+	Map of admixture proportions estimated using a spatial \texttt{conStruct} model for $K=2$ through 7.
+	The data were simulated using one layer with nearest-neighbor symmetric migration between demes.
+    }\label{simK1_sp_pies}
+\end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=\textwidth]{figs/sims/simK2_sp_pies.pdf}}
+	\caption{
+	Map of admixture proportions estimated using a spatial \texttt{conStruct} model for $K=2$ through 7.
+	The data were simulated using two layers with nearest-neighbor symmetric migration between demes.
+    }\label{simK2_sp_pies}
+\end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=\textwidth]{figs/sims/simK3_sp_pies.pdf}}
+	\caption{
+	Map of admixture proportions estimated using a spatial \texttt{conStruct} model for $K=2$ through 7.
+	The data were simulated using three layers with nearest-neighbor symmetric migration between demes.
+    }\label{simK3_sp_pies}
+\end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=\textwidth]{figs/sims/simK2_adprop_fit.pdf}}
+		\caption{
+			Plot of \texttt{conStruct} ability to correctly estimate admixture proportions on simulated data.
+			Results are from an analysis with a spatial model using $K=2$.
+		}\label{simK2_adprop_fit}
+\end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=\textwidth]{figs/sims/simK3_adprop_fit.pdf}}
+		\caption{
+			Plot of \texttt{conStruct} ability to correctly estimate admixture proportions on simulated data.
+			Results are from an analysis with a spatial model using $K=3$.
+		}\label{simK3_adprop_fit}
+\end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
 		{\includegraphics[width=\textwidth]{figs/sims/simK1_nsp_pies.pdf}}
@@
 	\centering
-		{\includegraphics[width=\textwidth]{figs/sims/simK1_sp_pies.pdf}}
-	\caption{
-	Map of admixture proportions estimated using a spatial \texttt{conStruct} model for $K=2$ through 7.
+		{\includegraphics[width=\textwidth]{figs/sims/simK2_nsp_pies.pdf}}
+	\caption{
+	Map of admixture proportions estimated using a nonspatial \texttt{conStruct} model for $K=2$ through 7.
+	The data were simulated using two layers with nearest-neighbor symmetric migration between demes.
+    }\label{simK2_nsp_pies}
+\end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=\textwidth]{figs/sims/simK3_nsp_pies.pdf}}
+	\caption{
+	Map of admixture proportions estimated using a nonspatial \texttt{conStruct} model for $K=2$ through 7.
+	The data were simulated using three layers with nearest-neighbor symmetric migration between demes.
+    }\label{simK3_nsp_pies}
+\end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=\textwidth]{figs/admixture/admixture_simK1_pies.pdf}}
+	\caption{
+	Map of admixture proportions estimated using ADMIXTURE \cite{ADMIXTURE} for $K=2$ through 7.
 	The data were simulated using one layer with nearest-neighbor symmetric migration between demes.
-    }\label{simK1_sp_pies}
+	The model with the lowest cross-validation error (i.e., the preferred model) was $K=7$.
+    }\label{admix_simK1}
+\end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=\textwidth]{figs/admixture/admixture_simK2_pies.pdf}}
+	\caption{
+	Map of admixture proportions estimated using ADMIXTURE \cite{ADMIXTURE} for $K=2$ through 7.
+	The data were simulated using two layers with nearest-neighbor symmetric migration between demes.
+	The model with the lowest cross-validation error (i.e., the preferred model) was $K=7$.
+    }\label{admix_simK2}
+\end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=\textwidth]{figs/admixture/admixture_simK3_pies.pdf}}
+	\caption{
+	Map of admixture proportions estimated using ADMIXTURE \cite{ADMIXTURE} for $K=2$ through 7.
+	The data were simulated using three layers with nearest-neighbor symmetric migration between demes.
+	The model with the lowest cross-validation error (i.e., the preferred model) was $K=7$.
+    }\label{admix_simK3}
 \end{figure}
@@
 	\centering
-		{\includegraphics[width=\textwidth]{figs/sims/simK2_nsp_pies.pdf}}
-	\caption{
-	Map of admixture proportions estimated using a nonspatial \texttt{conStruct} model for $K=2$ through 7.
-	The data were simulated using two layers with nearest-neighbor symmetric migration between demes.
-    }\label{simK2_nsp_pies}
-\end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=\textwidth]{figs/sims/simK2_sp_pies.pdf}}
-	\caption{
-	Map of admixture proportions estimated using a spatial \texttt{conStruct} model for $K=2$ through 7.
-	The data were simulated using two layers with nearest-neighbor symmetric migration between demes.
-    }\label{simK2_sp_pies}
+		{\includegraphics[width=\textwidth]{figs/sims/simK2_std_xval.pdf}}
+		\caption{
+			Cross-validation results for data simulated under $K=2$,
+			comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
+			The right panel zooms in on just the spatial cross-validation results.
+		}\label{simK2_xval}
+\end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=\textwidth]{figs/sims/simK3_std_xval.pdf}}
+		\caption{
+			Cross-validation results for data simulated under $K=3$,
+			comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
+			The right panel zooms in on just the spatial cross-validation results.
+		}\label{simK3_xval}
+\end{figure}
+\clearpage
+
+\begin{figure}
+	\centering
+		{\includegraphics[width=\textwidth]{figs/admixture/sims_CVerror.pdf}}
+	\caption{
+	ADMIXTURE cross-validation results for data simulated under $K=1$, $K=2$, and $K=3$, 
+	run with $K=1$ through 7 using 50 data folds (\texttt{--cv=50}).  
+	The inset plots zoom in on cross-validation results outlined in the dotted boxes.
+	The preferred model (with the lowest cross-validation error) is highlighted with an asterisk.
+    }\label{sim_admix_CVerrors}
 \end{figure}
@@
 	\centering
-		{\includegraphics[width=\textwidth]{figs/sims/simK2_std_xval.pdf}}
-		\caption{
-			Cross-validation results for data simulated under $K=2$,
-			comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
-			The right panel zooms in on just the spatial cross-validation results.
-		}\label{simK2_xval}
-\end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=\textwidth]{figs/sims/simK2_adprop_fit.pdf}}
-		\caption{
-			Plot of \texttt{conStruct} ability to correctly estimate admixture proportions on simulated data.
-			Results are from an analysis with a spatial model using $K=2$.
-		}\label{simK2_adprop_fit}
-\end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=\textwidth]{figs/sims/simK3_nsp_pies.pdf}}
-	\caption{
-	Map of admixture proportions estimated using a nonspatial \texttt{conStruct} model for $K=2$ through 7.
-	The data were simulated using three layers with nearest-neighbor symmetric migration between demes.
-    }\label{simK3_nsp_pies}
-\end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=\textwidth]{figs/sims/simK3_sp_pies.pdf}}
-	\caption{
-	Map of admixture proportions estimated using a spatial \texttt{conStruct} model for $K=2$ through 7.
-	The data were simulated using three layers with nearest-neighbor symmetric migration between demes.
-    }\label{simK3_sp_pies}
-\end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
 		{\includegraphics[width=\textwidth]{figs/sims/simK3_laycon_barplots.pdf}}
@@
 		}\label{simK3_laycon}
-\end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=\textwidth]{figs/sims/simK3_std_xval.pdf}}
-		\caption{
-			Cross-validation results for data simulated under $K=3$,
-			comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
-			The right panel zooms in on just the spatial cross-validation results.
-		}\label{simK3_xval}
-\end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=\textwidth]{figs/sims/simK3_adprop_fit.pdf}}
-		\caption{
-			Plot of \texttt{conStruct} ability to correctly estimate admixture proportions on simulated data.
-			Results are from an analysis with a spatial model using $K=3$.
-		}\label{simK3_adprop_fit}
-\end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=\textwidth]{figs/admixture/admixture_simK1_pies.pdf}}
-	\caption{
-	Map of admixture proportions estimated using ADMIXTURE \cite{ADMIXTURE} for $K=2$ through 7.
-	The data were simulated using one layer with nearest-neighbor symmetric migration between demes.
-	The model with the lowest cross-validation error (i.e., the preferred model) was $K=7$.
-    }\label{admix_simK1}
-\end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=\textwidth]{figs/admixture/admixture_simK2_pies.pdf}}
-	\caption{
-	Map of admixture proportions estimated using ADMIXTURE \cite{ADMIXTURE} for $K=2$ through 7.
-	The data were simulated using two layers with nearest-neighbor symmetric migration between demes.
-	The model with the lowest cross-validation error (i.e., the preferred model) was $K=7$.
-    }\label{admix_simK2}
-\end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=\textwidth]{figs/admixture/admixture_simK3_pies.pdf}}
-	\caption{
-	Map of admixture proportions estimated using ADMIXTURE \cite{ADMIXTURE} for $K=2$ through 7.
-	The data were simulated using three layers with nearest-neighbor symmetric migration between demes.
-	The model with the lowest cross-validation error (i.e., the preferred model) was $K=7$.
-    }\label{admix_simK3}
-\end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
-		{\includegraphics[width=\textwidth]{figs/admixture/sims_CVerror.pdf}}
-	\caption{
-	ADMIXTURE cross-validation results for data simulated under $K=1$, $K=2$, and $K=3$, 
-	run with $K=1$ through 7 using 50 data ``folds" (\texttt{--cv=50}).  
-	The inset plots zoom in on cross-validation results outlined in the dotted boxes.
-	The preferred model (with the lowest cross-validation error) is highlighted with an asterisk.
-    }\label{sim_admix_CVerrors}
 \end{figure}
@@
 	ADMIXTURE cross-validation results for poplar data, 
-	run with $K=1$ through 7 using 50 data ``folds" (\texttt{--cv=50}).  
+	run with $K=1$ through 7 using 50 data folds (\texttt{--cv=50}).  
 	The preferred model (with the lowest cross-validation error) is highlighted with an asterisk.
@@
 	ADMIXTURE cross-validation results for bear data, 
-	run with $K=1$ through 7 using 50 data ``folds" (\texttt{--cv=50}).  
+	run with $K=1$ through 7 using 50 data folds (\texttt{--cv=50}).  
 	The preferred model (with the lowest cross-validation error) is highlighted with an asterisk.

### 2018-01-29 writeup/plosgen_prep.R

@@ -9,38 +9,60 @@
 
 setwd("~/Dropbox/conStruct/writeup")
 
-# first, run git-latexdiff to make a diff file 
-#	to the initial submission 55053bbd120f68ed66e199bb39ecd30d3686dc19
-call <- c("git-latexdiff --main conStruct.tex 55053bbd120f68ed66e199bb39ecd30d3686dc19 --bibtex --ignore-makefile -o sub1_vs_sub2_diff.pdf")
-system(call)
-
-# make sure flags are:
+# first, make sure flags are:
 #	submissionversiontrue
 #	includesupplementtrue
 #	floatsatendfalse
 #	includefigstrue
+text <- readLines("conStruct.tex") 
+text[which(text == "\\submissionversiontrue")] <- "\\submissionversiontrue"
+text[which(text == "%\\submissionversiontrue")] <- "\\submissionversiontrue"
+text[which(text == "\\submissionversionfalse")] <- "%\\submissionversionfalse"
+text[which(text == "%\\submissionversionfalse")] <- "%\\submissionversionfalse"
+
+text[which(text == "\\includesupplementtrue")] <- "\\includesupplementtrue"
+text[which(text == "%\\includesupplementtrue")] <- "\\includesupplementtrue"
+text[which(text == "\\includesupplementfalse")] <- "%\\includesupplementfalse"
+text[which(text == "%\\includesupplementfalse")] <- "%\\includesupplementfalse"
+
+text[which(text == "\\floatsatendtrue")] <- "%\\floatsatendtrue"
+text[which(text == "%\\floatsatendtrue")] <- "%\\floatsatendtrue"
+text[which(text == "\\floatsatendfalse")] <- "\\floatsatendfalse"
+text[which(text == "%\\floatsatendfalse")] <- "\\floatsatendfalse"
+
+text[which(text == "\\includefigstrue")] <- "\\includefigstrue"
+text[which(text == "%\\includefigstrue")] <- "\\includefigstrue"
+text[which(text == "\\includefigsfalse")] <- "%\\includefigsfalse"
+text[which(text == "%\\includefigsfalse")] <- "%\\includefigsfalse"
+writeLines(text,"conStruct.tex")
+
+# compile
 call <- c("pdflatex conStruct.tex")
 system(call)
 system(call)
 
+# run git-latexdiff to make a diff file 
[…]

### 2018-01-29 writeup/conStruct.tex
reformatting

@@
 \newif\ifincludefigs
-%\includefigstrue
-\includefigsfalse
+\includefigstrue
+%\includefigsfalse
+\newif\ifsuppmatfigs
+\suppmatfigsfalse
+%\suppmatfigstrue
 \iffloatsatend
@@
 
+\extrafloats{100}
 % Remove comment for double spacing
@@
 \newpage
+
 \section*{Supplementary Information}
@@
 \setcounter{figure}{0}
-
+%
 % \linespread{1}
@@
 % \fi
-
-\clearpage
-
-\begin{figure}
-	\centering
+%
+\ifsuppmatfigs
+	\clearpage
+\fi
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/sims/simK1_sp_pies.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/sims/simK2_sp_pies.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/sims/simK3_sp_pies.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/sims/simK2_adprop_fit.pdf}}
+	\fi
 		\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/sims/simK3_adprop_fit.pdf}}
+	\fi
 		\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/sims/simK1_nsp_pies.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/sims/simK2_nsp_pies.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/sims/simK3_nsp_pies.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/admixture/admixture_simK1_pies.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/admixture/admixture_simK2_pies.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/admixture/admixture_simK3_pies.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/sims/simK1_std_xval.pdf}}
+	\fi
 		\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/sims/simK2_std_xval.pdf}}
+	\fi
 		\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/sims/simK3_std_xval.pdf}}
+	\fi
 		\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/admixture/sims_CVerror.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/sims/simK2_laycon_barplots.pdf}}
+	\fi
 		\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/sims/simK3_laycon_barplots.pdf}}
+	\fi
 		\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/populus/populus_sampling_map.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/populus/populus_sp_pies.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/populus/populus_nsp_pies.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/populus/populus_sp_layer_covs.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/populus/populus_nsp_layer_covs.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/populus/populus_laycon_barplots.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/populus/poplar_admixture_results.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/populus/poplars_admixture_CVerror.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/bears/bear_sp_results.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/bears/bear_nsp_results.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/bears/bear_sp_layer_covs.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/bears/bear_nsp_layer_covs.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/bears/bear_admixture_results.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
-
-\begin{figure}
-	\centering
+\ifsuppmatfigs
+	\clearpage
+\fi
+
+\begin{figure}[!p]
+	\centering
+	\ifsuppmatfigs
 		{\includegraphics[width=\textwidth]{figs/bears/bear_admixture_CVerror.pdf}}
+	\fi
 	\caption{
@@
 \end{figure}
-\clearpage
+\ifsuppmatfigs
+	\clearpage
+\fi
 

### 2018-01-29 writeup/plosgen_prep.R
reformatting for resub

@@ -34,6 +34,12 @@ text[which(text == "\\includefigstrue")] <- "\\includefigstrue"
 text[which(text == "%\\includefigstrue")] <- "\\includefigstrue"
 text[which(text == "\\includefigsfalse")] <- "%\\includefigsfalse"
 text[which(text == "%\\includefigsfalse")] <- "%\\includefigsfalse"
+
+text[which(text == "\\suppmatfigstrue")] <- "\\suppmatfigstrue"
+text[which(text == "%\\suppmatfigstrue")] <- "\\suppmatfigstrue"
+text[which(text == "\\suppmatfigsfalse")] <- "%\\suppmatfigsfalse"
+text[which(text == "%\\suppmatfigsfalse")] <- "%\\suppmatfigsfalse"
+
 writeLines(text,"conStruct.tex")
 
 # compile
@@ -97,18 +103,22 @@ text[which(text == "%\\floatsatendfalse")] <- "\\floatsatendfalse"
 text[which(text == "\\floatsatendtrue")] <- "%\\floatsatendtrue"
 text[which(text == "%\\includefigsfalse")] <- "\\includefigsfalse"
 text[which(text == "\\includefigstrue")] <- "%\\includefigstrue"
+text[which(text == "%\\suppmatfigsfalse")] <- "\\suppmatfigsfalse"
+text[which(text == "\\suppmatfigstrue")] <- "%\\suppmatfigstrue"
 writeLines(text,"conStruct.tex")
 
 call <- c("pdflatex conStruct.tex")
 system(call)
 system(call)
 
-# split out the pre-suppmat pages
-call <- c("pdfseparate -f 1 -l 26 conStruct.pdf Page%d.pdf")
-system(call)
-
-# unite the main text into a single doc for submission
-call <- paste0("pdfunite ",c(paste0("Page",1:26,".pdf",collapse=" "))," conStruct.pdf")
-system(call)
-
-file.remove(paste0("Page",1:26,".pdf"))
+if(FALSE){
+	# split out the pre-suppmat pages
+	call <- c("pdfseparate -f 1 -l 29 conStruct.pdf Page%d.pdf")
+	system(call)
+	
+	# unite the main text into a single doc for submission
+	call <- paste0("pdfunite ",c(paste0("Page",1:29,".pdf",collapse=" "))," conStruct.pdf")
+	system(call)
+	
+	file.remove(paste0("Page",1:29,".pdf"))
+}

### 2018-01-29 writeup/plosgen_prep.R
resubmission for reals

@@ -49,7 +49,7 @@ system(call)
 
 # run git-latexdiff to make a diff file 
 #	to the initial submission 55053bbd120f68ed66e199bb39ecd30d3686dc19
-call <- c("git-latexdiff --main conStruct.tex 55053bbd120f68ed66e199bb39ecd30d3686dc19 --bibtex --ignore-makefile -o sub1_vs_sub2_diff.pdf")
+call <- c("git-latexdiff --main conStruct.tex 55053bbd120f68ed66e199bb39ecd30d3686dc19 --bibtex --ignore-makefile --ignore-latex-errors -o sub1_vs_sub2_diff.pdf")
 system(call)
 
 # split out suppmat figs

### 2018-05-21 writeup/figs/bear_figs.R
bear figs plus ADMIXTURE

@@ -9,17 +9,28 @@ for(n in 1:n.reps){
 x.vals <- lapply(1:n.reps,function(n){get(sprintf("tl%s",n))})
 x.vals.std <- lapply(x.vals,conStruct:::standardize.xvals)
 xval.CIs <- conStruct:::get.xval.CIs(x.vals.std,K)
-pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bear_std_xval.pdf",width=10,height=5,pointsize=14)
-#	quartz(width=10,height=5)
-	par(mfrow=c(1,2),mar=c(4,5,4,2))
+pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bear_std_xval.pdf",width=12,height=4.25,pointsize=14)
+#	quartz(width=12,height=5)
+	par(mfrow=c(1,3),mar=c(3.75,5,4,1))
 	plot.xval.CIs(xval.CIs,K,jitter=0.1,xlim=c(0.75,7.25))
 		legend(x="bottomright",pch=19,col=c("blue","green"),legend=c("spatial","nonspatial"))
-	mtext("Predictive accuracy",side=2,padj=-5,font=2)
+	mtext("Predictive accuracy",side=2,padj=-3,font=2)
+	mtext("conStruct",side=3,font=2)
 	plot.xval.CIs(xval.CIs,K,ylim=c(-200,0),jitter=0.1,xlim=c(3.75,7.25),xaxt='n')
 		axis(1,at=c(4:7))
 		legend(x="bottomright",pch=c(19,NA),lty=c(NA,1),lwd=c(NA,2),col=c(1,adjustcolor(1,0.8)),legend=c("mean","95% CI"))
-	mtext("number of layers",side=1,adj=-0.9,padj=4,font=2)
-	mtext("Cross-validation results (Bears)",side=3,adj=10.5,padj=-2.5,font=2,cex=1.2)
+	mtext("number of layers",side=1,font=2,padj=2.75)
+	mtext("Cross-validation results (Bears)",side=3,padj=-2.2,font=2,cex=1.2)
+	mtext("non-spatial conStruct",side=3,font=2)
+	mtext("Predictive accuracy",side=2,padj=-3,font=2)
+	CV.error <- get.CV.error("~/Dropbox/conStruct/data/bears/admixture/exe.admixture.Rout")
+	plot(CV.error,pch=19,col="orangered",
+			xlab="",ylab="",cex=2,ylim=rev(range(CV.error)))
[…]

### 2018-05-21 writeup/response2reviews.tex
started edits for genetics reviews

@@
 Re-Submission Cover Letter \hfill	\today\\
-\emph{PLoS Genetics}
+\emph{Genetics}
 \\\\\\
-To the Editor(s),
+To the Editor,
 
@@
 ``Inferring Continuous and Discrete Population Genetic Structure Across Space,"
-for re-consideration for publication in PLoS Genetics.
+for re-consideration for publication in Genetics.
 
 We appreciate the opportunity to submit a revised version of the manuscript.  
-The comments provided by the Guest Editor 
-and two anonymous reviewers are very helpful, 
-and we believe the manuscript is substantially improved as a result 
+The comments provided by the Editor 
+and the anonymous reviewer are very helpful, 
+and we believe the manuscript is improved as a result 
 of their critiques.
@@
 Our detailed responses to those comments are inset below (in bold), 
-but, in brief, we have implemented a more explicit comparison between 
-\texttt{conStruct}'s spatial model and existing, mainstream nonspatial model-based clustering methods.
-Specifically, we have analyzed all datasets (simulated and empirical) 
-included in the paper using the method ADMIXTURE (Alexander \textit{et al.} 2009),
-and used ADMIXTURE's $n$-fold cross-validation procedure (Alexander and Lange 2011)
-to compare model fit across values of $K$.
-As with the comparison between the spatial and nonspatial models in \texttt{conStruct}, 
-these results highlight both the tendency of nonspatial models 
-to describe continuous differentiation with multiple, spurious clusters, 
-and also the capacity of the spatial model to avoid this pitfall.
-
-We have also edited the description of the allelic covariance we model,
-both for clarity (in response to helpful suggestions from both reviewers), 
-and to bring the text into agreement with what we had actually implemented in the code.
+but, in brief, we have included more of the results of the ADMIXTURE analyses 
+in the main body of the text.
+Specifically, we have added figure panels to Figures 7 and 8 
+that explicitly compare \texttt{conStruct}'s results with those of ADMIXTURE 
+on the black bear dataset.
+We have also added text throughout the results to increase the depth of 
+our comparison of \texttt{conStruct} and ADMIXTURE.
 
@@
 
-\textbf{Guest Editor}\\
-We believe the concerns of reviewer 1 are addressable, and would like to see some effort put towards addressing them, specifically benchmarking against non-spatial structure identification methods.
+\textbf{Editor}\\
+We received the previous reviews for your paper, and sent also sent it out to a new reviewer, along with your response to the earlier reviews.
+Both the new reviewer and I believe that you have answered the previous criticisms in a reasonable way, and I am pleased to tell you that your paper is potentially suitable for publication in GENETICS.
+However, the new reviewer did have several comments which you should consider in preparing the final version (see the end of this email).
+In particular, they suggest including more of the ADMIXTURE models in the main text.
+I think that this could be helpful, but would not insist on it.
 
-\textbf{Reviewer 1}\\
-The paper introduces a model-based method called conStruct to account for isolation-by-distance (IBD) when clustering individuals into populations.
-IBD is known to generate spurious clusters and the proposed software package can avoid spurious cluster detection.
-Accounting for space in the proposed model based framework amounts at introducing a regularization penalty when fitting admixture model.
-Regularization explains why overfitting is avoided and correct admixture proportions are found even when performing model-based clustering with a value of K that is too large.
+\textbf{Reviewer}\\
+\begin{itemize}
+\item This manuscript presents a valuable model that addresses an long standing chestnut in population genetic analysis, which is how to jointly analyze discrete and spatially continuous variation within a single dataset.
+The method presented does seem to be something of a breakthrough in addressing this problem sensibly.
+I found the authors response to previous reviewer comments to be generally reasonable.
+However, I do think that there is a strong case for putting some more ADMIXTURE model results into the text and possibly for some additional analyses.
 
-This being said, I have serious concerns about the submitted manuscript.
+\item I agree with the authors point in their letter that ADMIXTURE/STRUCTURE is important because it is what almost everyone would do otherwise.
+The key goal for the manuscript is to show that the method that everyone uses gives results that can be misleading/unhelpful and that the layered model is a substantial improvement.
+This is very clear for the simulated data.
+The previous reviewers are right to ask for additional simulations in the sense that a more biological simulation would make a stronger case for practical utility and it need not lengthen the manuscript significantly, since the current simulations could be largely relegated to supplementary material as representing "simulation according to the model".
+But this is definitely an optional, since there are presentational arguments to be made either way.
+
+\item A more biological simulation is not necessary as long as the real data examples are convincing.
+For the Poplar data, the key insight of the model is that there are three well-supported layers in addition to isolation by distance.
+This is not clear according to ADMIXTURE, which gives continuous improvement for K, although the clustering itself is reasonable.
+The differences in patterns of IBD between populations are also interesting.
+
+\item For the bear data, there is a stronger case to be made, namely that as well as estimating a biologically insightful number of layers, individual ADMIXTURE or non-spatial model components are positively misleading for K=3 or more.
+For the Poplar example, the results are not actively misleading , they just represent finer and finer geography.
+The misleadingness of the bear results can be shown by the fact that although the cross validation shows continuous improvement, the admixture components are highly inconsistent between runs.
+For example at K=3, a component is shown that is shared between East and West, which if interpreted literally would represent a biologically important result about connections between populations but the clustering of the yellow populations disappears entirely for K=4.
+Therefore, I think there is a good case to be made for showing the ADMIXTURE model results in the text and to discuss the misleading/ unstable nature of any particular result.
+
+\item It would be interesting to do some kind of plot for ADMIXTURE results, to demonstrate that populations that share affinity according higher components are not particularly similar when allowing for geographical distance.
+This would be a rather direct way of showing that the ADMIXTURE model is not producing insightful results, despite the continuous improvement in model probability for higher K.
+In this context, I always think of the "clines vs clusters" paper by Rosenberg.
+It would be great to have some kind of figure to show that for K=2, it looks like clusters but for K=3, the clusters add much less.
+
+\item The layer specific covariances show a very funny pattern, tending to infinity for low distances for components 3 and above for the bear data.
+I am not sure I understand this.
+Is this discussed anywhere?\\\\
+\gb{I think what's happening is that these layer-specific covariances for layers $>$2 
+are describing local departures from global patterns of IBD, 
+as might be expected due to, e.g., inbreeding.
+We have added some text to that effect in Lines XXX-XXX.}
+
+\item Minor point, in figure 4, legend, should it read in each run of the spatial model, rather than nonspatial model?\\\\
+\gb{Good catch!  Caption has been fixed.}
+\end{itemize}
+
+\end{document}
 

### 2018-05-21 writeup/conStruct.tex
reformatting for resub
resubmission for reals
submission to genetics
started edits for genetics reviews

@@
 \newif\ifsubmissionversion
-%\submissionversionfalse
-\submissionversiontrue
+\submissionversionfalse
+%\submissionversiontrue
 \newif\ifincludesupplement
@@
 \newif\ifsuppmatfigs
-\suppmatfigsfalse
-%\suppmatfigstrue
+%\suppmatfigsfalse
+\suppmatfigstrue
+\newif\iflinenos
+\linenostrue
+%\linenosfalse
+
 \iffloatsatend
@@
 \usepackage[right]{lineno}
+\iflinenos
+\linenumbers
+\fi
 
@@
     \setlength{\headheight}{27.023pt}
-    \lhead{\includegraphics[width=\textwidth]{latex/PLOS-submission.eps}}
+%    \lhead{\includegraphics[width=\textwidth]{latex/PLOS-submission.eps}}
     \rfoot{\thepage/\pageref{LastPage}}
@@
     \fancyfootoffset[L]{2.25in}
-    \lfoot{\sf PLOS}
+%    \lfoot{\sf PLOS}
 \fi
@@
 			for the spatial model (left), and the nonspatial model (right).
-			In each run of the nonspatial model, a single layer explained nearly all the covariance
+			In each run of the spatial model, a single layer explained nearly all the covariance
             (additional bars are present but not visible).
@@
 
+\paragraph{Comparison to ADMIXTURE}
 Both nonspatial \texttt{conStruct} and ADMIXTURE displayed 
@@
 (Fig \ref{populus_nsp_pies} vs. Fig \ref{populus_admixture}), 
-and also differed across the replicate analyses, 
-but each strongly favored the most cluster-rich model
+and also differed across the replicate analyses.
+These differences between runs and methods may be due to 
+noise in the different inference algorithms employed,
+multi-modality in the likelihood surfaces, or 
+to differences in the specifics of model 
+(e.g., the priors used in nonspatial \texttt{conStruct}, 
+or the fact that ADMIXTURE is modeling each allele's frequency in each cluster, 
+rather than the covariance across all alleles).
+However, overall, the behavior of both methods is quite similar: 
+each recovers the \tri{}/\bals{} split with the first two clusters modeled, 
+then, with higher values of $K$, 
+uses subsequent clusters to subdivide the \tri{} samples 
+into geographically restricted foci of cluster membership.
+And, both nonspatial \texttt{conStruct} and ADMIXTURE 
+strongly favored the most cluster-rich model
 (Figs \ref{populus_xvals} and \ref{poplar_admix_CVerrors}).
@@
 make more detailed interpretations untrustworthy.
+The spatial covariance functions estimated in layers beyond the first two 
+take very large values over small spatial lags, 
+but decay sharply after that.
+This feature, combined with the overall amounts 
+and spatial patterns of ancestry in those layers, 
+suggests that these layers are describing processes 
+that shape genetic variation at local scales, such as inbreeding, 
+which affects covariance between individuals within demes, 
+but has limited impact on covariance between demes.
 
@@
 \fi
-\begin{figure}[!p]
+\begin{figure}[!ht]
 	\centering
@@
 
-\begin{figure}[!p]
+\begin{figure}[!ht]
 	\centering

### 2018-05-21 writeup/response2reviews.tex
responses to Rev1

@@
 I found the authors response to previous reviewer comments to be generally reasonable.
-However, I do think that there is a strong case for putting some more ADMIXTURE model results into the text and possibly for some additional analyses.
+However, I do think that there is a strong case for putting some more ADMIXTURE model results into the text and possibly for some additional analyses.\\\\
+\gb{We have now added some of the ADMIXTURE results to the main text, 
+including the admixture pie plot and the cross-validation results from the ADMIXTURE analyses 
+of the bear dataset in Figures 7 and 8, respectively. 
+We have also included more detailed comparisons between ADMIXTURE and \texttt{conStruct} (see more below).}
 
@@
 The previous reviewers are right to ask for additional simulations in the sense that a more biological simulation would make a stronger case for practical utility and it need not lengthen the manuscript significantly, since the current simulations could be largely relegated to supplementary material as representing "simulation according to the model".
-But this is definitely an optional, since there are presentational arguments to be made either way.
+But this is definitely an optional, since there are presentational arguments to be made either way.\\\\
+\gb{WILL FILL IN AFTER WE FIND OUT WHAT NEWS WEDNESDAY BRINGS}
 
@@
 The differences in patterns of IBD between populations are also interesting.
-
-\item For the bear data, there is a stronger case to be made, namely that as well as estimating a biologically insightful number of layers, individual ADMIXTURE or non-spatial model components are positively misleading for K=3 or more.
+For the bear data, there is a stronger case to be made, namely that as well as estimating a biologically insightful number of layers, individual ADMIXTURE or non-spatial model components are positively misleading for K=3 or more.
 For the Poplar example, the results are not actively misleading , they just represent finer and finer geography.
@@
 For example at K=3, a component is shown that is shared between East and West, which if interpreted literally would represent a biologically important result about connections between populations but the clustering of the yellow populations disappears entirely for K=4.
-Therefore, I think there is a good case to be made for showing the ADMIXTURE model results in the text and to discuss the misleading/ unstable nature of any particular result.
+Therefore, I think there is a good case to be made for showing the ADMIXTURE model results in the text and to discuss the misleading/ unstable nature of any particular result.\\\\
+\gb{We have added more extensive discussions of the 
+comparison between \texttt{conStruct} and ADMIXTURE for both empirical datasets.
+These subsections appear on lines XXX-XXX for the poplar dataset, 
+and lines XXX-XXX for the bear dataset.}
 
@@
 In this context, I always think of the "clines vs clusters" paper by Rosenberg.
-It would be great to have some kind of figure to show that for K=2, it looks like clusters but for K=3, the clusters add much less.
+It would be great to have some kind of figure to show that for K=2, it looks like clusters but for K=3, the clusters add much less.\\\\
+\gb{I didn't follow exactly what this figure might look like, 
+but hopefully this point comes across clearly in the contrast between 
+ADMIXTURE and spatial \texttt{conStruct} results.}
 
@@
 as might be expected due to, e.g., inbreeding.
-We have added some text to that effect in Lines XXX-XXX.}
+We have added some text to that effect in lines XXX-XXX.}
 
@@
 \end{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2018-05-21 writeup/conStruct.tex
responses to Rev1

@@
 rather than the covariance across all alleles).
-However, overall, the behavior of both methods is quite similar: 
-each recovers the \tri{}/\bals{} split with the first two clusters modeled, 
+However, overall, the behavior of both methods was quite similar: 
+each recovered the \tri{}/\bals{} split with the first two clusters modeled, 
 then, with higher values of $K$, 
-uses subsequent clusters to subdivide the \tri{} samples 
+used subsequent clusters to subdivide the \tri{} samples 
 into geographically restricted foci of cluster membership.
@@
 (Figs \ref{populus_xvals} and \ref{poplar_admix_CVerrors}).
+In contrast, the spatial \texttt{conStruct} model clearly did not favor 
+the model with the highest value of $K$, 
+and appears to describe patterns of isolation by distance between
+demes within \tri{} quite well.
 
@@
 	Maps of admixture proportions estimated for the black bear dataset 
-	using the spatial \texttt{conStruct} model (left) 
-	and the nonspatial \texttt{conStruct} model (right) for $K=3$.
+	using the spatial \texttt{conStruct} model (left),
+	the nonspatial \texttt{conStruct} model (center), 
+	and ADMIXTURE (right) for $K=3$.
 	Pies show mean admixture results across individuals within their diameter, 
@@
 	Cross-validation results for the black bear dataset,
-	comparing spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
-	The first panel in each row shows all results; 
-	the second panel zooms in on the results from the spatial analyses run with $K = 4$ through 7.
+	comparing spatial and nonspatial \texttt{conStruct} models, 
+	as well as ADMIXTURE, all run with $K=1$ through 7.  
+	The first panel in each row shows results from spatial and nonspatial \texttt{conStruct} models; 
+	the second panel zooms in on the results from the spatial analyses run with $K = 4$ through 7, 
+	and the third panel shows the results for ADMIXTURE.
+	Note that the admixture plot shows cross-validation error 
+	(rather than predictive accuracy), and that the y-axis has therefore been flipped 
+	for ease of comparison to the \texttt{conStruct} results.
     }\label{bear_xvals}
@@
 
+\paragraph{Comparison to ADMIXTURE}
 Results from the nonspatial model and from the ADMIXTURE analyses 
@@
 spatial models with best cross-validation performance in \texttt{conStruct} ($K=5$ or 6). 
-However, comparing results of the two methods, it is clear that even at $K=3$, 
+This discrepancy may be the result of differences in the cross-validation procedures 
+implemented in \texttt{conStruct} and ADMIXTURE, 
+or perhaps in the ways each accommodate missing data.
+
+Comparing results of the two methods, it is clear that even at $K=3$, 
 ADMIXTURE is invoking clusters to describe what seems to be, 

### 2018-05-21 writeup/figs/admixture_figs.R

@@ -90,7 +90,26 @@ pdf(file="~/Dropbox/conStruct/writeup/figs/admixture/admixture_simK1_pies.pdf",w
 			mtext(side=1,font=2,text=bquote(paste("true ",italic("K")," = ",1)),cex=1.2,padj=2.7)
 		}
 	}
-dev.off()			
+dev.off()
+
+ad.Qs <- lapply(2:7,
+			function(k){
+				pop.vec <- unlist(lapply(1:sim.dataset$N,function(n){rep(n,10)}))
+				Q.file <- sprintf("~/Dropbox/conStruct/sims/admixture/datasets/simK1/simK1.%s.Q",k)
+				Q <- read.table(Q.file,stringsAsFactors=FALSE)
+				admix.props <- collapse.ind.Q(Q,pop.vec)
+				return(admix.props)
+			})
+
+for(k in 1:6){
+	pdf(file=sprintf("~/Dropbox/conStruct/writeup/figs/admixture/simK1_%s_struct.pdf",k+1),width=10,height=4.5)
+		make.structure.plot(ad.Qs[[k]],layer.colors=laycols[[k]])
+	dev.off()
+	pdf(file=sprintf("~/Dropbox/conStruct/writeup/figs/admixture/simK1_%s_pies.pdf",k+1),width=6,height=6)
+		make.admix.pie.plot(ad.Qs[[k]],sim.dataset$coords,layer.colors=laycols[[k]],radii=6,x.lim=c(2.5,8.5),y.lim=c(2.5,8.5))
+		mtext(side=3,font=2,text=bquote(paste(italic("K"),"=",.(k+1))),cex=2)
+	dev.off()
+}
 
 load("~/Dropbox/conStruct/sims/cross_validation/K_2/sim.dataset.Robj")
 laycols <- get.laycols(K=7,q.file="~/Dropbox/conStruct/sims/admixture/datasets/simK2/simK2.%s.Q",N=sim.dataset$N)

### 2018-06-05 writeup/conStruct.tex
data availability, typos, formatting

@@
 \usepackage[hidelinks]{hyperref}
+\usepackage{breakurl}
 
@@
 
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLOS ONE authors please skip this step. 
-% Author Summary not valid for PLOS ONE submissions.   
 \section*{Author summary}
@@
 Here, we will focus on model-based clustering, the most widely used class of approaches for population delineation.
+(We note that the problem of identifying population clusters is distinct from, 
+though of course related to, 
+the problem of detecting barriers to gene flow between populations, 
+e.g.,  
+\cite{BEDASSLE,EEMS,barton2008,Ringbauer2018}.)
 Existing model-based clustering methods model each individual's genotypes 
@@
 \cite{malecot1969mathematics,Slatkin1985,epperson2003geographical}, 
-but substantial work remains to be done \cite{barton2002neutral,barton2013modelling}.
+and some theory has been derived for continuous space 
+\cite{nagylaki1978diffusion,nagylaki1988,barton2002neutral},
+but substantial work remains to be done \cite{barton2013modelling}.
 Given the generality of the circumstances that generate a pattern of isolation by distance, 
@@
 The implementation of this method, \texttt{conStruct} (for ``\emph{con}tinuous \emph{struct}ure"), 
-is available as an R package at 
-\href{https://github.com/gbradburd/conStruct}{https://github.com/gbradburd/conStruct}.
-
-% \section*{Methods}
-\section*{Results}
+is documented and available for general use as an R package at 
+\burl{https://github.com/gbradburd/conStruct}.
+
+\section*{Methods}
 
@@
 
-% \section*{Results}
+\section*{Data availability}
+The method \texttt{conStruct} is implemented as an R package, 
+and is available for installation at \burl{https://github.com/gbradburd/conStruct}.
+Scripts for generating and analyzing all simulated and empirical datasets, 
+as well as the datasets themselves, 
+are also available at the same site.
+\section*{Results}
 
@@
 We then ran both spatial and nonspatial \texttt{conStruct} analyses on each simulated dataset with $K$ between 1 and 7,
-and compared predictive performance of the models using cross-validation.
+and compared predictive performance of the models using cross-validation with 10 replicates.
 For comparison,
@@
 	comparing the spatial and nonspatial \texttt{conStruct} models 
-	(in blue and green, respectively) run with $K=1$ through 7.  
+	(in blue and green, respectively) run with $K=1$ through 7, 
+	with 10 cross-validation replicates.
 	The inset plots zoom in on cross-validation results outlined in the dotted boxes.
@@
 
+\paragraph{Study system and questions}
 Trees in the genus \textit{Populus} (poplars, aspens, and cottonwoods) 
@@
 
+\paragraph{Data and analyses}
 We use data from \cite{geraldes_etal_2014}, 
@@
 using both the spatial and nonspatial \texttt{conStruct} models with $K = 1$ through 7,
-and compared these models using cross-validation.
+and compared these models using cross-validation with 10 replicates.
 The results of all these analyses are shown in 
@@
 	Cross-validation results for \textit{Populus} dataset 
-	comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7.  
+	comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7 
+	with 10 cross-validation replicates.
 	The first panel in each row shows all results; 
@@
 
+\paragraph{Results from \texttt{conStruct}}
 All models with $K>1$ assigned the majority of each of the two species
@@
 
+\paragraph{Study system and questions}
 The American black bear, \textit{Ursus americanus}, is endemic to North America
@@
 
+\paragraph{Data and analyses}
 We use data from \cite{Puckett2015}, consisting of 95 individuals 
@@
 using both the spatial and nonspatial \texttt{conStruct} models with a $K$ of between 1 and 7, 
-and compared these models using cross-validation.
+and compared these models using cross-validation with 10 replicates.
 We also ran ADMIXTURE \cite{ADMIXTURE} on the same dataset, 
@@
 	comparing spatial and nonspatial \texttt{conStruct} models, 
-	as well as ADMIXTURE, all run with $K=1$ through 7.  
+	as well as ADMIXTURE, all run with $K=1$ through 7, 
+	with 10 cross-validation replicates for the \texttt{conStruct} analyses 
+	and 50 data-fold subsets for the ADMIXTURE analyses. 
 	The first panel in each row shows results from spatial and nonspatial \texttt{conStruct} models; 
@@
 
+\paragraph{Results from \texttt{conStruct}}
 The results partition the sampled bears into two main groups:
@@
 \text{Var}(\epsilon) +  \text{Var}\left( \Delta^{(k)} \right) +
-\text{Cov}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell}) + \delta_{i=j} \text{Var}\left( \Delta^{(i)} \right),
+\text{Cov}(\Delta^{(k,i)},\Delta^{(k,j)}) + \delta_{i=j} \text{Var}\left( \Delta^{(i)} \right),
 \end{equation}
@@
   \text{Var}\left( \Delta^{(k)} \right) +
-\text{Cov}(\Delta^{(k,i)}_{\ell},\Delta^{(k,j)}_{\ell}) 	\right) +
+\text{Cov}(\Delta^{(k,i)},\Delta^{(k,j)}) 	\right) +
 \delta_{i=j} \text{Var}(\Delta_i) .
@@
 
+The parameters used to simulate the data were as follows:
+a diploid population size of 1000,
+a migration rate between neighboring demes of $0.4$,
+a deep split time between layers of $500$ 
+(corresponding to $\tau_{\text{s}}$ in Fig \ref{sim_setup}),
+an expansion event across layers of $250$
+(corresponding to $\tau_{\text{e}}$ in Fig \ref{sim_setup}), 
+and an admixture event between layers in the immediate past ($1 \times 10^{-4}$).  
+The times and rates reported above have already been scaled by $4N$ 
+(as per \texttt{ms} syntax), 
+and therefore give the values fed directly to \texttt{ms}.
+We used the \texttt{-s} option to sample a single segregating site per coalescent history, 
+and simulated $1 \times 10^4$ independent histories
+- corresponding to the same number of independent loci - 
+in each dataset, 
+with 10 diploid genotypes generated per deme at each locus.
 \section*{Acknowledgements}

### 2018-06-19 writeup/conStruct.tex

@@
 FINESTRUCTURE, fastStructure, huelsenbeck2007inference, 
-Corander2003,TESS,geneland}), among others.
+Corander2003,TESS,geneland}).
 Each of these methods perform best in particular situations,
@@
 chosen from each of the two samples at a random locus.
-More precisely, suppose that we pick a random bi-allelic locus uniformly,
+More precisely, suppose that we pick a random bi-allelic locus uniformly from the genome,
 pick a random ``reference'' allelic state from the two alleles seen at that locus,
@@
 but this covariance is allowed to decay as geographic distance increases.
-This pattern of spatial decay reflects how migration between neighboring demes  
+This pattern of spatial decay reflects how migration between nearby spatial regions
 homogenizes allele frequency changes that arise locally due to drift, 
-but less effectively homogenizes geographic distant demes,
+but less effectively homogenizes geographically distant regions,
 resulting in a continuous pattern of isolation by distance within each layer.
@@
 is added by spatial structure within each layer, and each additional layer.
-However, we remind users that ``statistical significance does not imply real-world significance','
+However, we remind users that ``statistical significance does not imply real-world significance'',
 and so small but statistically significant differences between models
-should probably not be relied on too strongly.
+should not be relied on too strongly.
 
@@
 by computing the average contribution to increased probability of agreement
-as described in \secref{layer_contribution}). 
+as described in \secref{layer_contribution}. 
 This layer contribution is similar to the ``ancestry contribution'' proposed by \cite{fastStructure}. 
@@
 we also analyzed each simulated dataset using ADMIXTURE \cite{ADMIXTURE} with $K$ between 1 and 7, 
-and compared models using ADMIXTURE's cross-validation procedure with 50 data fold subsets.
+and compared models using ADMIXTURE's cross-validation procedure with 50 folds.
 
@@
 and then plateaus or declines
-(Figs \ref{sim_xvals} and \ref{simK1_xval}-\ref{simK3_xval}),
-while predictive accuracy of the nonspatial model increases as subsequent clusters are added.
+(Figs \ref{sim_xvals} and \ref{simK1_xval}-\ref{simK3_xval}).
+Predictive accuracy of the nonspatial model increases as subsequent clusters are added
+up to $K=7$ (the largest number tested),
+although gains are greatest as layers below the true number are added.
 The same holds true for the ADMIXTURE cross-validation results,
@@
 noise in the different inference algorithms employed,
-multi-modality in the likelihood surfaces, or 
-to differences in the specifics of model 
+multi-modality in the likelihood surfaces, or to model details
 (e.g., the priors used in nonspatial \texttt{conStruct}, 
@@
 into geographically restricted foci of cluster membership.
-And, both nonspatial \texttt{conStruct} and ADMIXTURE 
+Both nonspatial \texttt{conStruct} and ADMIXTURE 
 strongly favored the most cluster-rich model
@@
 the model with the highest value of $K$, 
-and appears to describe patterns of isolation by distance between
-demes within \tri{} quite well.
+and appears to describe patterns of isolation by distance 
+across the \tri{} range quite well.
 
@@
 \paragraph{Results from \texttt{conStruct}}
-The results partition the sampled bears into two main groups:
+The results partition the sampled bears into two main groups (shown in Fig \ref{bear_K3}a):
 one (red) to the east of the Rocky Mountains, 
 which also occurs in Alaska,
-the other primarily west of the Rockies (blue)  (Fig \ref{bear_K3}a).
+the other primarily west of the Rockies (blue).
 The disjointed range of the blue layer likely reflects the fact
@@
 that shape genetic variation at local scales, such as inbreeding, 
-which affects covariance between individuals within demes, 
-but has limited impact on covariance between demes.
+which affects covariance between individuals within each location, 
+but has limited impact on covariance between locations.
 
@@
 e.g., at $K=3$ 
--- the preferred model from the cross-validation 
-analysis shown in Fig \ref{bear_admix_CVerrors} -- 
+--- the preferred model from the cross-validation 
+analysis shown in Fig \ref{bear_admix_CVerrors} --- 
 ADMIXTURE splits the westernmost Alaskan samples 
@@
 spatial models with best cross-validation performance in \texttt{conStruct} ($K=5$ or 6).
-This discrepancy may be the result of differences in the cross-validation procedures 
-implemented in \texttt{conStruct} and ADMIXTURE, 
-or perhaps in the ways each accommodate missing data.
+This discrepancy likely stems from the different features introduced in layers beyond $K=2$ in the two models:
+conStruct uses small contributions of new layers to model very local drift,
+while ADMIXTURE moves to geographically finer subdivisions.
 
 Comparing results of the two methods, it is clear that even at $K=3$, 
-ADMIXTURE is invoking clusters to describe what seems to be, 
+ADMIXTURE invokes clusters to describe what seems to be, 
 and is well described by \texttt{conStruct} as, 
@@
 We anticipate that \texttt{conStruct} will be useful for identifying populations 
-and determining samples' ancestry in and across them, 
-especially when the populations exhibit spatial patterns of relatedness.
+and determining ancestral origins of sampled individuals,
+especially when the populations exhibit geographic patterns of relatedness.
 
@@
 (b) \texttt{conStruct} can concisely capture spatial structure, which
-will be common within populations.
+is common within populations.
 Given this, when should methods without spatial capability be used?
@@
 it currently scales poorly with number of samples.
-The computationally limiting step is the inversion of the parametric covariance, 
+The computationally limiting step is the inversion of the covariance matrix, 
 which scales more than quadratically with the number of samples,
@@
 fastSTRUCTURE might be faster.
-And, for almost any dataset of any size, 
+For almost any size dataset,
 the maximum likelihood algorithm implemented in ADMIXTURE 
-is also quite a bit faster than \texttt{conStruct}; 
+is quite a bit faster than \texttt{conStruct}:
 running ADMIXTURE on the bear dataset over all values of $K$ from 1 to 7, 
@@
 are shaped by complex spatial and hierarchical processes.
-All individuals within a species (and indeed, all individuals across all species), 
-are related to one another in some way, 
-and summarizing that relatedness with a single value of $K$ may be reductive or misleading.
+All individuals within a species are related to one another in some way, 
+and summarizing those patterns of relatedness with a single value of $K$ may be reductive or misleading.
 We therefore encourage users to perform analyses across different values of $K$ and 
@@
 nor does a non-zero admixture proportion indicate that admixture 
-(i.e. gene flow) must have occurred. 
+(i.e., gene flow) must have occurred. 
 Both groupings and admixture proportions
@@
 Because isolation by distance is common, 
-a likely result of applying \texttt{conStruct} to existing data is that some
-populations previously identified using nonspatial clustering methods 
-may be collapsed into each other.  
-This ``lumping'' might better reflect biological reality, 
-but may also have implications for management decisions and conservation policy, 
-both of which are often predicated on the identification of discrete ``management units'' (MUs) 
+a likely result of applying \texttt{conStruct} to existing data is that
+populations previously identified as distinct using nonspatial clustering methods 
+may be grouped into the same layer.
+This ``lumping'' might better reflect the demographic history of these populations,
+but may not contradict the genetic distinctness implied by the nonspatial clustering.
+This genetic distinctness rather than shared history may be more relevant for
+management decisions and conservation policy,
+which are often predicated on the identification of discrete ``management units'' (MUs) 
 identified using genetic data \cite{Moritz1994,Waples_1998,Moritz_etal_2002}.
@@
 and that a \texttt{conStruct} layer may still contain multiple distinct MUs worthy of independent protections.  
+For instance, although both the Alaskan and Eastern Black Bears draw most of their ancestry 
+from the same conStruct layer, they are separated by a great distance, 
+and may therefore differ substantially from each other
+(although less than from the Western bears, as measured by average covariance).
 Alternatively, the inclusion of multiple MUs into a single \texttt{conStruct} layer 
@@
 (or were recently) exchanging migrants, 
-and thus might emphasize the importance of maintaining habitat corridors between demes, 
-or of implementing an integrated conservation plan across multiple demes within a layer.
+and thus might emphasize the importance of maintaining habitat corridors,
+or of implementing an integrated conservation plan across a geographic region.
 
@@
 The choice of allelic covariance, rather than genetic covariance,
-was motivated by two considerations.
-First, it captures within-population variance in a manner more suited to the modeling approach here,
-which led to better performance on test data.
-Second, it is less affected by sample configuration --
+was motivated by the fact that
+it is less affected by sample configuration --
 the genetic covariance is calculated after subtracting the mean from the entire sample,
 which is more strongly affected by densely sampled locations.
+This led to better performance on test data.
 However, note that allelic covariance is more affected by singleton sites
@@
 sampled individuals of greenish warblers \cite{Irwin2001} or \textit{Ensatina} salamanders \cite{wake_schneider1998} 
---- two canonical examples of rings species --- 
+--- two canonical examples of ring species --- 
 we might get misleading results.
@@
 ``admixture proportion'' of sample $i$ in layer $k$.
-(Note that $\sum_{k=1}^K w^{(k)}_{i} = 1$ for each $i$.)
 These can be interpreted as the proportion of the genome in the $i$th
 sample that came from the $k$th layer 
-(or the probability that an allele at a locus is drawn from layer $k$).
+(or the probability that an allele at a locus is drawn from layer $k$),
+so that $\sum_{k=1}^K w^{(k)}_{i} = 1$ for each $i$.
 The allele frequency in the $i$th sample at the $\ell$th locus can therefore be written as:
@@
 \end{equation}
-We note that this form is chosen for its flexibility, 
+We note that this form is chosen for flexibility and convenience,
 and not because it matches any explicit population genetic model of isolation by distance.
@@
 Averaging over choice of reference allele therefore removes some information 
-about factors acting within populations that might otherwise leave signatures in, 
-e.g., the genetic covariance, 
+about factors acting within populations that might otherwise leave signatures in
+the genetic covariance, 
 such as population size, extent of inbreeding, and history of bottlenecks.  
@@
 the computation time of the likelihood is not a function of the number of loci,
-so inference can theoretically be done using whole genome data.
+so inference can be done using whole genome data.
 
@@
 	\begin{enumerate}
-		\item partition the allele frequency data into a 90\% ``training'' partition ($F^x_1$) and a 10 \%``testing" partition ($F^x_2$) \label{partition}
-		\item run our inference procedure on the training partition to estimate model parameters $\theta_{mk}$ for: \label{inference}
+		\item partition the allele frequency data into a 90\% ``training'' partition ($F^x_1$) and a 10\%``testing" partition ($F^x_2$) \label{partition}
+
+		\item run our inference procedure using the training partition to estimate model parameters $\theta_{mk}$ for $2K$ models: \label{inference}
 			\begin{enumerate}
@@
 			\end{enumerate}
+
         \item calculate the mean log likelihood of the testing data partition  
@@
             ($\bar{\mathcal{L}}(F^x_2 \mid \theta_{mk})$, henceforth $\bar{\mathcal{L}}_{xmk}$) \label{lnL}
-	\item generate standardized mean log likelihoods, $\mathcal{Z}_{xmk}$, across models: \label{standardize}
+
+        \item generate standardized mean log likelihoods, $\mathcal{Z}_{xmk}$, across all models run on this data partition: \label{standardize}
             \begin{enumerate}
-			\item identify the highest mean log likelihood, $\bar{\mathcal{L}}^\text{max}_{xmk}$
-			\item subtract $\bar{\mathcal{L}}^\text{max}_{xmk}$ from $\bar{\mathcal{L}}_{xmk}$ across all models,
+                \item identify the highest mean log likelihood, $\bar{\mathcal{L}}^\text{max}_{x}$ across \emph{all} $2K$ models
+                \item subtract $\bar{\mathcal{L}}^\text{max}_{x}$ from $\bar{\mathcal{L}}_{xmk}$ for each model,
                     such that the standardized log likelihood, $\mathcal{Z}_{xmk}$, of the best model is 0,
@@
 	\end{enumerate}
-\item For each model (i.e., each combination of $m$ and $k$) calculate 
-	the mean ($	\bar{\mathcal{Z}}_{mk} $) standardized log likelihood of the testing data partition across $X$ replicates, 
-	as well its standard error ($SE_{\bar{\mathcal{Z}}_{mk}}$) and
+
+\item For each \emph{model} (i.e., each combination of $m$ and $k$) calculate 
+	the mean ($\bar{\mathcal{Z}}_{mk}$) standardized log likelihood of the testing data partition across $X$ replicates, 
+	as well as its standard error ($SE_{\bar{\mathcal{Z}}_{mk}}$) and
     95\% confidence interval ($\bar{\mathcal{Z}}_{mk} \pm 1.96 \times SE_{\bar{\mathcal{Z}}_{mk}}$).
+
 \end{enumerate}
+
+In other words, the ``predictive accuracy'' shown as conStruct crossvalidation results
+are in units of improvement in log-likelihood of that model relative to the best model for that partitioning of the data,
+averaged over data partitions.
+The standardization is necessary because different data partitions can be systematically 
+more or less difficult to fit, resulting in greater differences in mean training data log likelihood 
+between data partitions
+than between models fit to the same partition.
 
@@
 If they do choose the same layer, 
-then they agree with a probability $p_\gamma + q^{(k)}_{ij}$ that depends on their distance apart.
+then they agree with a probability $1/2 + p_\gamma + q^{(k)}_{ij}$ that depends on their distance apart.
 By the above,
 the probability of agreement is $\P\{A=B\} = 2 \cov[A,B] + 1/2$,
-and so
+and so we can define
 $$\begin{aligned}
-    p_\gamma &= (2 (\gamma + \delta_{ij} \eta_i) + 1/2) \\
-    q^{(k)}_{ij} &= \left(2 \alpha_0^{(k)} \exp\left( - \left(\alpha_D^{(k)} D_{ij}\right)^{\alpha_2^{(k)}} \right) + 2 \phi^{(k)}  + 1/2 \right) .
+    p_\gamma &= 2 (\gamma + \delta_{ij} \eta_i) \\
+    q^{(k)}_{ij} &= 2 \alpha_0^{(k)} \exp\left( - \left(\alpha_D^{(k)} D_{ij}\right)^{\alpha_2^{(k)}} \right) + 2 \phi^{(k)} .
 \end{aligned}$$
@@
 So, the contribution from layer $k$ to agreement between $i$ and $j$ is 
-$w_i^{(k)} w_j^{(k)} q^{(k)}_{ij} / (p_\gamma + \sum_{k=1}^K w_i^{(k)} w_j^{(k)} q^{(k)}_{ij})$,
+$$\begin{aligned}
+    \frac{
+        w_i^{(k)} w_j^{(k)} \left( 1/2 + q^{(k)}_{ij} \right) 
+    }{
+        \left(1 - \sum_{k=1}^K w_i^{(k)} w_j^{(k)}\right) \left( 1/2 + p_\gamma \right) 
+        + \sum_{k=1}^K w_i^{(k)} w_j^{(k)} \left( 1/2 + q^{(k)}_{ij} \right) ,
+    }
+\end{aligned}$$
 which is the probability, given that they agree, that they agree thanks to layer $k$.
-Similarly, the ``background" contribution is $p_\gamma / (p_\gamma + \sum_{k=1}^K w_i^{(k)} w_j^{(k)} q^{(k)}_{ij})$.
 Because our signal comes from \emph{variation} in covariance, we omit
-the $p_\gamma$ terms. 
+the $p_\gamma$ terms 
+(i.e., we condition on agreement not due to ``background'' levels of agreement
+in the interpretations above).
 Stated in this way, this quantity is the
-relative contribution of the $k$th layer to the (model-based) kinship
+relative contribution of the $k^\text{th}$ layer to the (model-based) kinship
 coefficient between $i$ and $j$.
@@
 \end{equation}
-This is the quantity that is plotted in, e.g., Figures \ref{simK1_laycon} and \ref{bear_laycon}.
+This is the quantity that is plotted in Figures \ref{simK1_laycon} and \ref{bear_laycon}.
 

### 2018-06-19 writeup/conStruct.tex
fixed postmerge typos

@@
 informing conservation priorities \cite{Moritz1994,Waples_1998,Moritz_etal_2002}.
-Furthermore, we often need to identify subsets of indivduals resulting from reasonably coherent
+Furthermore, we often need to identify subsets of individuals resulting from reasonably coherent
 evolutionary histories for downstream analyses to learn about population history and adaptation.
@@
 
-We employ Monte Carlo cross-validation approach for model comparison \cite{picard1984}. 
+We employ a Monte Carlo cross-validation approach for model comparison \cite{picard1984}. 
 This procedure generates a mean predictive accuracy for each model and each value of $K$, 
@@
 
-In other words, the ``predictive accuracy'' shown as conStruct crossvalidation results
+In other words, the ``predictive accuracy'' shown as conStruct cross-validation results
 are in units of improvement in log-likelihood of that model relative to the best model for that partitioning of the data,

### 2018-06-26 writeup/response2reviews.tex
started response to Rev2

@@
 The comments provided by the Editor 
-and the anonymous reviewer are very helpful, 
+and the anonymous reviewers are very helpful, 
 and we believe the manuscript is improved as a result 
@@
 I think that this could be helpful, but would not insist on it.
+A second review came in late (attached).
+I am sending it on in case you would like to take account of any of the suggestions in your revision.
+The reviewer has quite a few technical suggestions that may be helpful.
+However, even if it had come in before the decision, I would not have required a wider range of simulations.
 
-\textbf{Reviewer}\\
+\textbf{Reviewer 1}\\
 \begin{itemize}
@@
 This is very clear for the simulated data.
-The previous reviewers are right to ask for additional simulations in the sense that a more biological simulation would make a stronger case for practical utility and it need not lengthen the manuscript significantly, since the current simulations could be largely relegated to supplementary material as representing "simulation according to the model".
+The previous reviewers are right to ask for additional simulations in the sense that a more biological simulation would make a stronger case for practical utility and it need not lengthen the manuscript significantly, since the current simulations could be largely relegated to supplementary material as representing ``simulation according to the model".
 But this is definitely an optional, since there are presentational arguments to be made either way.\\\\
@@
 This would be a rather direct way of showing that the ADMIXTURE model is not producing insightful results, despite the continuous improvement in model probability for higher K.
-In this context, I always think of the "clines vs clusters" paper by Rosenberg.
+In this context, I always think of the ``clines vs clusters" paper by Rosenberg.
 It would be great to have some kind of figure to show that for K=2, it looks like clusters but for K=3, the clusters add much less.\\\\
@@
 
+\textbf{Reviewer 2}\\
+\begin{itemize}
+\item Make-or-Break:
+Ideally one tests a new method on simulated data, to assess its power -
+and importantly whether everything was implemented correctly.
+Here, there are simulations of one scenario, but it is neither strictly under-the-model 
+nor a biologically plausible scenario (layers that coexist independent of each other are hard to imagine.
+There seems to be a good fit for the ground-truth admixture coefficients in the scenario you simulate. 
+However, you never assess how well you can fit isolation by distance parameters, 
+and whether your posterior distribution makes any sense. 
+This is a very important step in demonstrating that you implementation works - 
+and would give me and other users much needed reassurance.
+The authors also never give details such as migration rates of this simulations, 
+so the reader has effectively no idea what the extend of isolation by distance and differentiation is in these simulations. 
+Please put them at least in a supplement table. 
+In general it is really hard to get a sense for the power of the method, 
+as there is no exploration done in that direction. 
+Power will obviously depend on a plethora of factors, 
+but at least give an idea of what marker numbers you envision as typical input.
+Given the high-dimensional parameter space there is no realistic way to thoroughly test your inference scheme in this paper - 
+so I would be satisfied if you pick a set of parameters you deem realistic 
+(for instance similar to the ones you estimate in the applications), 
+and show that the inference method can infer them when simulated under the model. 
+In particular, show that your model infers also the isolation by distance parameters well, 
+and that the posteriors make sense. 
+In case it does not (due to model degeneracies or similar issues , see below), 
+it is not big deal as long as the ``important" admixture parameters are unaffected, 
+but then point these out.
+If there are space issues, you can put these plots in a short supplement.\\\\
+\gb{RESPONSE.}
+
+\item Also, I agree with Reviewer 1 that it would be interesting to see how your method behaves on a biologically plausible, 
+idealized, scenario, such as secondary contact with isolation by distance. 
+While the ground truth will be blurry for your model, 
+if the method infers the ``right clusters" and plausible admixture coefficients 
+this would give much credibility to your method - 
+showing that it also works when the model does not strictly hold, 
+as will always be the case in practice. 
+I don't consider this as an essential make-or-break though as the simulation-under-the-model.\\\\
+\gb{RESPONSE.}
+
+\item Issues that should be addressed:
+ There is a typo or something missing in Eq. 8 and Eq. 10., 
+ there appears an l whereas the covariance is already averaged over markers. 
+ Do you miss a sum sign here? 
+ Please clarify that as this masks an important assumption 
+ that is almost certainly violated in most biological realistic scenarios:
+The covariance matrices are not the same across markers, 
+as drift depends on allele frequency. 
+In the extreme case of fixed markers the covariance is always 0! 
+In Population Genetics one often models this with a p(1-p) 
+pref-actor of the covariance matrix OR $F_{ST}$ summary statistics. 
+Obviously averaging over quite different covariance matrices 
+violates the multivariate normal Wishard approximation with 
+degrees of freedom equal to L. 
+Please elaborate on that issue in the ``technical supplement".
+Ringbauer2018 explored a way around this caveat via Fishers-Transform of allele
+frequencies or directly modeling F-matrices. (This sentence falls into III).
+ Throughout the manuscript it is implied that Structure-like methods constitutes a subset of yours, 
+ corresponding to turning off the spatial covariance component. 
+ I have to disagree and think this is misleading: 
+ Importantly, Structures also fits allele frequencies for each loci. 
+ Thus, a concurrent cline of a few of them can already force the introduction of a new cluster. 
+ In stark contrast, ConStruct only uses a single covariance matrix averaged across markers, 
+ so will be relatively blind to this behavior of a few markers! 
+ It seems important to mention that. 
+ In some observed biological scenarios only few markers cause a hybrid zone.\\\\
+\gb{RESPONSE.}
+
+\item Put your work in context. 
+ While I agree with your assessment that detailed comparisons on simulations is more than enough work for a own paper, 
+ it would be nice to put your work into context of methods. 
+ In particular discuss methods that can be applied on datasets that are interesting for your method. 
+ In particular Geneland and EMS come to mind. 
+ The latter will be better to deal with highly heterogeneous landscapes, with varying IBD patterns.
+Interestingly, you also compete with Bedassle. 
+Since it fits allele frequencies, it likely has higher power to confirm 
+hypothesized barriers that aren't detected by this method. 
+You could mention something along these lines.\\\\
+\gb{RESPONSE.}
+
+\item In line 97 to 111: The analogy to ``random flipping of allele state" is mildly interesting, 
+but does not add much insight. 
+You can cut it and start directly with your summary statistic (2) and only mention this analogy in the Supplement/Appendix.\\\\
+\gb{RESPONSE.}
+
+\item On multiple occasions you argue that the ``allelic covariance" matrix is insensitive to the choice of reference allele. 
+This is also true for most covariance matrices of the style (p1-p, p2-p), or do I miss something??
+The main advantage seems to be that you buy is that you don't have to estimate mean allele frequencies, which is MANY parameters. 
+You probably have taught quite a bit about this step ``down" from Bedasssle, so perhaps share some of your views and initial experiments.\\\\
+\gb{RESPONSE.}
+
+\item In line 72-73, you cite that mainly stepping-stone models gives us some analytical predictions. 
+In fact, good analytic approximations are often derived for continuous limits, such as a diffusion model.
+You could add some reference for continuous models that incorporate discrete discontinuities and try to merge these two. 
+Sometimes, there are analytic approximations, which is obviously relevant: 
+Importantly, the landmark work of Nayglaki1988 is missing. 
+Morerecently Barton2008 and Ringbauer2018 made some contributions 
+for the effects of barriers to gene flow on genetic covariance in a continuous setting. 
+They are not directly relevant for this work, but hey could provide useful context.\\\\
+\gb{RESPONSE.}
+
+\item Your covariance matrix has to be a valid covariance matrix, i.e. positive semidefinite. 
+I guess this is why you choose the power of the exponential to be between 0 and 2 
+(for other values it isn't), but you never mention that. 
+In table 1, some draws from the priors would not produce valid covariance matrices, 
+how do you proceed in these cases? 
+Do you have implementations that enforce this or are these areas never sampled in practice? 
+This issue could mess with the Wishart distribution and matrix inversion if the determinant crosses to negative values.\\\\
+\gb{RESPONSE.}
+
+\item There seem to be some possible degeneracy issues, 
+for instance when $\alpha_d$ goes to 0 (which seems to be the case for some of the inferred layers), 
+$\alpha_0$ and $\phi_k$ will becomes indistinguishable.
+ You should mention these issues, and if you have dealt with them 
+ (or think they are not important, as you are not really interested in estimating these parameters).\\\\
+\gb{RESPONSE.}
+
+\item Genetics states ``All articles published by GENETICS must include a 
+Data Availability Statement at the end of the Materials and Methods section." 
+You don't have such a section at the moment.
+For reproducibility of your results, 
+you should make at least the covariance matrix and the positions of each deme you used in your analysis available. 
+Ideally share all of the genetic data and the scripts used to filter it, including the settings (this is often quite important). 
+But if someone is protective of this data, I personally would be fine with omitting it. But this is the editors call to make.\\\\
+\gb{RESPONSE.}
+
+\item Fig. 4 could be relegated to the supplement, it does not show much more really important information as Fig. 2. 
+ In particular if you need space.\\\\
+\gb{RESPONSE.}
+
+\item It would be really helpful if you plot the ground truth admixture coefficients in your K plots (Fig. 2) for K>2. 
+ It seems you sometimes estimate admixture for populations which have none (also observed in my applications), 
+ but not knowing the ground truth this is hard to gauge. 
+ You randomly simulated them, so please add this ``ground truth" as an additional plot. 
+ I am slightly skeptical about Plot S5 (the admixture coefficients) ? did you do averaging? 
+ Having ground truth plots would help to remove my doubts!\\\\
+\gb{RESPONSE.}
+
+\item  Line 386: I think the claim that your model includes a biologically realistic model for isolation by distance is a bit overzealous. 
+ Several non-interacting layers of IBD are a bit hard to imagine, 
+ or do you have a good example?
+ Importantly, your method will infer different clusters when isolation by distance patterns change. 
+ So again you run into a problem of inferring a discrete break where there is none. 
+ In fact, for most realistic populations one would expect different isolation by distance patterns.
+  Could be worth mentioning that caveat, and that in that case EEMS style methods.\\\\
+\gb{RESPONSE.}
+
+\item The headings are a bit off. 
+For instance you start with Results and then lay out the model? 
+Orient yourself on the excellent structure of the BEDASSLE paper.
+ You cite often quite a few works at once (for instance ten in line 60), 
+ check if this goes well with the citation style of Genetics).\\\\
+\gb{RESPONSE.}
+
+\item III) Suggestions (for future)
+ When I run the inference scheme myself, 
+ I noticed that mostly only one core on my machine is under full load, while the other remain idle. 
+ You could conceive ways to incorporate multi-processing. 
+ In particular, if run on a cluster, this could yield huge speed-ups. 
+ There seems to be some literature out there in how to do so for the inversion of huge covariance matrices, 
+ which seems to be the main bottleneck.\\\\
+\gb{RESPONSE.}
+
+ \item You assume a multivariate normal distribution. 
+ This implicitly assumes that all higher moments are 0, 
+ which is quite the assumption for real data and almost certainly violated. 
+ But In some sense, your method fits the covariances matrix - the pattern that it produces it does not matter. 
+ This brings me directly to the next point:
+ You could explore other ways to fit the covariance matrix. 
+ Since typically the model is violated anyways, 
+ you won't lose too much (the posterior is no ``real" posterior)
+For instance, fitting all pairwise covariance by for instance least square 
+is not necessarily inferior, but could give you significant speed updates 
+(quadratic, and no cumbersome inversion of the covariance matrix).
+Ringbauer2018 tried such approaches and found 
+that simpler fitting methods often have similar power than more complex ones 
+(whose assumptions are violated anyways), but are much faster.\\\\
+\gb{RESPONSE.}
+
+ \item Right now you have to give the pairwise distance Matrix as parameter to your method, 
+ and also the sample positions. 
+ You could easily calculate the former from the latter, 
+ and would save users some quick but unnecessary preprocessing step. 
+ I guess in the future you might want to use other distance matrices though.\\\\
+\gb{RESPONSE.}
+
+ \item In line 569 to line 575: When you describe how to deal with linked markers. 
+ You don't mention it very explicitly, 
+ but I think you hint that one can simply do ``block-bootstrap" of whole blocks of genome 
+ as typically done nowadays for many applications, 
+ if at least linkage groups are known.\\\\
+\gb{RESPONSE.}
+
+ \item The Gamma-Exponential Kernels used here are often considered inferior to Matern-Kernels. 
+ In practice it likely wont matter much though.\\\\
+\gb{RESPONSE.}
+
+\item I personally really fancy your writing style, 
+and it is a golden thread throughout your work.  
+But as much as I enjoy learning new words, 
+keep in mind many others in Science are non-mother tongue speakers 
+(and perhaps less literate native speakers). 
+And sometimes it simply feels like you are under contract of an English dictionary company.
+Additionally, a more concise and straightforward formulation 
+would significantly reduce the length of the manuscript and improve its readability 
+- while keeping the information content unchanged.
+
+One example of many, to give you an idea:
+``Existing model-based clustering approaches can only describe 
+continuous patterns of variation using discrete clusters,
+ and tend to erroneously describe continuously distributed variation 
+ with multiple clusters that show spatially autocorrelated cluster membership" (l.82-84)
+
+to
+
+``Most existing model-based clustering methods only use discrete units. 
+Therefore, they often partition continuous variation into spurious clusters. 
+(with spatially autocorrelated membership.)"
+
+This is merely a suggestion, choice of style is obviously the author's and there is
+no ground truth best way.\\\\
+\gb{RESPONSE.}
+
+\item Stress more that an easy-to-use implementation inclusive a manual exists, that is
+certainly a strong point! I was impressed by the ease of installation and use, and quality of the manual.\\\\
+\gb{RESPONSE.}
+
+\end{itemize}
 \end{document}

### 2018-06-26 writeup/response2reviews.tex
started responses to Rev2

@@
 If there are space issues, you can put these plots in a short supplement.\\\\
-\gb{RESPONSE.}
+\gb{First, we want to thank Reviewer 2 for their very thorough  review
+(even taking the time to check out the R package!).  
+We really appreciate the time and thought the reviewer invested here.}
 
@@
  check if this goes well with the citation style of Genetics).\\\\
-\gb{RESPONSE.}
+\gb{These were holdovers from a previous submission format.  
+We have edited them for a more logical flow and greater clarity.}
 
@@
  which seems to be the main bottleneck.\\\\
-\gb{RESPONSE.}
+\gb{Parallelization would definitely be helpful.  
+Right now, the inference algorithm is implemented in STAN, 
+which doesn't have a mechanism for parallelizing a single MCMC chain.
+But, as you say, there are algorithms for parallelizing matrix inversions calculations, 
+so we might look to porting the inference out of STAN to get greater speedups for future releases.}
 
@@
 (whose assumptions are violated anyways), but are much faster.\\\\
-\gb{RESPONSE.}
+\gb{This is a nice idea, and one that we will definitely look into.
+We've talked about including a ``fast" conStruct option  
+that uses a maximum-likelihood algorithm in a future release, 
+and possibly one that uses the TreeMix trick of modeling each cell in the 
+sample covariance matrix as a draw from a Normal with a mean given by the 
+corresponding cell in the parametric covariance matrix, 
+and a standard deviation calculated from by block-boostrapping the genome.
+Your suggestion to use least squares would probably be even faster.}
 
@@
  I guess in the future you might want to use other distance matrices though.\\\\
-\gb{RESPONSE.}
+\gb{Right, it would save users some time for us to calculate pairwise distances for them,
+but at the expense of flexibility.
+Some users may want to use other distance matrices 
+(e.g., distance within a river system, or along a coastline, or around the Tibetan plateau, etc.), 
+so we leave it up to them to specify one.}
 
@@
  if at least linkage groups are known.\\\\
-\gb{RESPONSE.}
+\gb{Yes, we think that's likely the best way to do it when linkage is known.
+We've added some text about that in the ``Cross-validation" section of the Appendix 
+(Lines XXX-XXX).}
 
@@
  In practice it likely wont matter much though.\\\\
-\gb{RESPONSE.}
+\gb{Yes, one of the next updates we plan is to replace the flexible but 
+poorly motivated powered exponential with the form worked out for 
+probability of identity by state in continuous space from Barton, Depaulis, and Etheridge (2000).}
 
@@
 - while keeping the information content unchanged. 
-
+\\\\
 One example of many, to give you an idea:
@@
  with multiple clusters that show spatially autocorrelated cluster membership" (l.82-84)
-
 to 
-
 ``Most existing model-based clustering methods only use discrete units. 
@@
 no ground truth best way.\\\\
-\gb{RESPONSE.}
+\gb{First, thank you for the compliment (seriously - ``a golden thread"!).
+We appreciate that not all scientists are native English speakers, 
+and have tried to be as clear and specific in our writing as possible, 
+both in the manuscript, and even more in the R package vignettes. 
+And, while some of the words we choose are perhaps rare in scientific writing, 
+we put great care into our choice of syntax, 
+and we think that the nuances of meaning these odd word-birds lend us merit their inclusion.
+As Mark Twain wrote, ``The difference between the \emph{almost} right word and the right word 
+is really a large matter -- 
+'tis the difference between the lightning bug and the lightning."}
 
@@
 certainly a strong point! I was impressed by the ease of installation and use, and quality of the manual.\\\\
-\gb{RESPONSE.}
+\gb{Good suggestion - we have now indicated that the method is available as a documented R package.}
 

### 2018-06-26 writeup/response2reviews.tex
more responses to Rev2

@@
 We have also added text throughout the results to increase the depth of 
-our comparison of \texttt{conStruct} and ADMIXTURE.
+our comparison of \texttt{conStruct} and ADMIXTURE, 
+and generally edited the manuscript for greater clarity.
 
@@
 It would be great to have some kind of figure to show that for K=2, it looks like clusters but for K=3, the clusters add much less.\\\\
-\gb{I didn't follow exactly what this figure might look like, 
+\gb{We didn't follow exactly what this figure might look like, 
 but hopefully this point comes across clearly in the contrast between 
@@
 Is this discussed anywhere?\\\\
-\gb{I think what's happening is that these layer-specific covariances for layers $>$2 
+\gb{We think what's happening is that these layer-specific covariances for layers $>$2 
 are describing local departures from global patterns of IBD, 
@@
 (even taking the time to check out the R package!).  
-We really appreciate the time and thought the reviewer invested here.}
+We really appreciate the time and thought the reviewer invested here.
+RESPONSE.}
 
@@
 Ringbauer2018 explored a way around this caveat via Fishers-Transform of allele
-frequencies or directly modeling F-matrices. (This sentence falls into III).
- Throughout the manuscript it is implied that Structure-like methods constitutes a subset of yours, 
+frequencies or directly modeling F-matrices. (This sentence falls into III).\\\\
+\gb{Good catch!  Yes, there was a typo (that $\ell$ subscript didn't belong), 
+and we have fixed the error. 
+Your larger point is correct that the amount of detectable drift is a function of the initial allele frequency,
+so some transformation or normalization of allele frequencies might be useful.
+But, we have avoided this because they make the sample allelic covariance that we model 
+sensitive to which samples are included in the dataset, 
+rather than a quantity that is invariant across choice of samples.
+}
+
+\item Throughout the manuscript it is implied that Structure-like methods constitutes a subset of yours, 
  corresponding to turning off the spatial covariance component. 
@@
 You could mention something along these lines.\\\\
-\gb{RESPONSE.}
+\gb{We have added more text comparing conStruct with Geneland and EEMS and BEDASSLE.}
 
@@
 You can cut it and start directly with your summary statistic (2) and only mention this analogy in the Supplement/Appendix.\\\\
-\gb{RESPONSE.}
+\gb{We have edited that section for clarity and brevity.}
 
@@
 They are not directly relevant for this work, but hey could provide useful context.\\\\
-\gb{RESPONSE.}
+\gb{Thanks for pointing out these missing and important references.  
+We have added them to the manuscript.}
 
@@
 This issue could mess with the Wishart distribution and matrix inversion if the determinant crosses to negative values.\\\\
-\gb{RESPONSE.}
+\gb{Yes, as you say the covariance must be positive semi-definite for us to invert it and calculate the likelihood.
+The inference algorithm implemented by STAN automatically rejects proposed matrices that aren't invertible.
+While it is possible that this could introduce issues into the mixing of the MCMC, 
+in practice we haven't found that concern to be borne out.}
 
@@
  (or think they are not important, as you are not really interested in estimating these parameters).\\\\
-\gb{RESPONSE.}
+\gb{In practice, because those parameters can't be easily interpreted in a biological context, 
+we treat them more as nuisance parameters and focus on inference of the admixture proportions.
+We have plans for a future release that has a more biologically motivated model of IBD (see more below).}
 
@@
 But if someone is protective of this data, I personally would be fine with omitting it. But this is the editors call to make.\\\\
-\gb{RESPONSE.}
+\gb{Thanks for pointing this out.  
+We have added a Data Availability Statement, 
+and all data analyzed and scripts used to perform the analyses have been archived in data dryad.}
 
@@
  In particular if you need space.\\\\
-\gb{RESPONSE.}
+\gb{We like Fig. 4 as a demonstration of the utility of calculating layer 
+contributions as a heuristic of model comparison. 
+And, although the information Fig. 4 contains is a little redundant with that of Fig. 2, 
+that wouldn't have to be the case, because the $phi$ parameters for the extra layers 
+in the spatial model could be so large that even though ancestry in them was negligible, 
+their contributions could still be significant.}
 
-\item It would be really helpful if you plot the ground truth admixture coefficients in your K plots (Fig. 2) for K>2. 
+\item It would be really helpful if you plot the ground truth admixture coefficients in your K plots (Fig. 2) for $K>2$. 
  It seems you sometimes estimate admixture for populations which have none (also observed in my applications), 

### 2018-06-26 writeup/previous_response2reviews.tex
added fancy review commands

--- 2018-06-26 writeup/response2reviews.tex
@@
-\documentclass[11pt]{letter}
-\usepackage[hidelinks]{hyperref}
-\usepackage[green]{msuletterhead}
-\usepackage[usenames,dvipsnames]{xcolor}
-\pagenumbering{gobble}
-
-\newcommand{\gb}[1]{{\bf\color{black}{#1}}}
-\newcommand{\plr}[1]{{\it\color{purple}{(#1)}}}
-\newcommand{\gc}[1]{{\it\color{blue}{(#1)}}}
-
-\begin{document}
-
-Re-Submission Cover Letter \hfill	\today\\
-\emph{Genetics}
-\\\\\\
-To the Editor,
-
-Please find enclosed the re-submission of our manuscript, 
-``Inferring Continuous and Discrete Population Genetic Structure Across Space,"
-for re-consideration for publication in Genetics.
-
-We appreciate the opportunity to submit a revised version of the manuscript.  
-The comments provided by the Editor 
-and the anonymous reviewers are very helpful, 
-and we believe the manuscript is improved as a result 
-of their critiques.
-
-Our detailed responses to those comments are inset below (in bold), 
-but, in brief, we have included more of the results of the ADMIXTURE analyses 
-in the main body of the text.
-Specifically, we have added figure panels to Figures 7 and 8 
-that explicitly compare \texttt{conStruct}'s results with those of ADMIXTURE 
-on the black bear dataset.
-We have also added text throughout the results to increase the depth of 
-our comparison of \texttt{conStruct} and ADMIXTURE.
-
-We hope that the paper is now acceptable for publication.\\\\
-%
-
-Thank you for your consideration,\\
-Gideon Bradburd (for all of the authors).
-
-\newpage
-
-\textbf{Editor}\\
-We received the previous reviews for your paper, and sent also sent it out to a new reviewer, along with your response to the earlier reviews.
-Both the new reviewer and I believe that you have answered the previous criticisms in a reasonable way, and I am pleased to tell you that your paper is potentially suitable for publication in GENETICS.
-However, the new reviewer did have several comments which you should consider in preparing the final version (see the end of this email).
-In particular, they suggest including more of the ADMIXTURE models in the main text.
-I think that this could be helpful, but would not insist on it.
-A second review came in late (attached).
-I am sending it on in case you would like to take account of any of the suggestions in your revision.
-The reviewer has quite a few technical suggestions that may be helpful.
-However, even if it had come in before the decision, I would not have required a wider range of simulations.
-
-\textbf{Reviewer 1}\\
-\begin{itemize}
-\item This manuscript presents a valuable model that addresses an long standing chestnut in population genetic analysis, which is how to jointly analyze discrete and spatially continuous variation within a single dataset.
-The method presented does seem to be something of a breakthrough in addressing this problem sensibly.
-I found the authors response to previous reviewer comments to be generally reasonable.
-However, I do think that there is a strong case for putting some more ADMIXTURE model results into the text and possibly for some additional analyses.\\\\
-\gb{We have now added some of the ADMIXTURE results to the main text, 
-including the admixture pie plot and the cross-validation results from the ADMIXTURE analyses 
-of the bear dataset in Figures 7 and 8, respectively. 
-We have also included more detailed comparisons between ADMIXTURE and \texttt{conStruct} (see more below).}
-
-\item I agree with the authors point in their letter that ADMIXTURE/STRUCTURE is important because it is what almost everyone would do otherwise.
-The key goal for the manuscript is to show that the method that everyone uses gives results that can be misleading/unhelpful and that the layered model is a substantial improvement.
-This is very clear for the simulated data.
-The previous reviewers are right to ask for additional simulations in the sense that a more biological simulation would make a stronger case for practical utility and it need not lengthen the manuscript significantly, since the current simulations could be largely relegated to supplementary material as representing ``simulation according to the model".
-But this is definitely an optional, since there are presentational arguments to be made either way.\\\\
-\gb{WILL FILL IN AFTER WE FIND OUT WHAT NEWS WEDNESDAY BRINGS}
-
-\item A more biological simulation is not necessary as long as the real data examples are convincing.
-For the Poplar data, the key insight of the model is that there are three well-supported layers in addition to isolation by distance.
-This is not clear according to ADMIXTURE, which gives continuous improvement for K, although the clustering itself is reasonable.
-The differences in patterns of IBD between populations are also interesting.
-For the bear data, there is a stronger case to be made, namely that as well as estimating a biologically insightful number of layers, individual ADMIXTURE or non-spatial model components are positively misleading for K=3 or more.
-For the Poplar example, the results are not actively misleading , they just represent finer and finer geography.
-The misleadingness of the bear results can be shown by the fact that although the cross validation shows continuous improvement, the admixture components are highly inconsistent between runs.
-For example at K=3, a component is shown that is shared between East and West, which if interpreted literally would represent a biologically important result about connections between populations but the clustering of the yellow populations disappears entirely for K=4.
-Therefore, I think there is a good case to be made for showing the ADMIXTURE model results in the text and to discuss the misleading/ unstable nature of any particular result.\\\\
-\gb{We have added more extensive discussions of the 
-comparison between \texttt{conStruct} and ADMIXTURE for both empirical datasets.
-These subsections appear on lines XXX-XXX for the poplar dataset, 
-and lines XXX-XXX for the bear dataset.}
-
-\item It would be interesting to do some kind of plot for ADMIXTURE results, to demonstrate that populations that share affinity according higher components are not particularly similar when allowing for geographical distance.
-This would be a rather direct way of showing that the ADMIXTURE model is not producing insightful results, despite the continuous improvement in model probability for higher K.
-In this context, I always think of the ``clines vs clusters" paper by Rosenberg.
-It would be great to have some kind of figure to show that for K=2, it looks like clusters but for K=3, the clusters add much less.\\\\
-\gb{I didn't follow exactly what this figure might look like, 
-but hopefully this point comes across clearly in the contrast between 
-ADMIXTURE and spatial \texttt{conStruct} results.}
-
-\item The layer specific covariances show a very funny pattern, tending to infinity for low distances for components 3 and above for the bear data.
-I am not sure I understand this.
-Is this discussed anywhere?\\\\
-\gb{I think what's happening is that these layer-specific covariances for layers $>$2 
-are describing local departures from global patterns of IBD, 
-as might be expected due to, e.g., inbreeding.
-We have added some text to that effect in lines XXX-XXX.}
-
-\item Minor point, in figure 4, legend, should it read in each run of the spatial model, rather than nonspatial model?\\\\
-\gb{Good catch!  Caption has been fixed.}
-\end{itemize}
-
-\textbf{Reviewer 2}\\
-\begin{itemize}
-\item Make-or-Break:
-Ideally one tests a new method on simulated data, to assess its power -
-and importantly whether everything was implemented correctly.
-Here, there are simulations of one scenario, but it is neither strictly under-the-model 
-nor a biologically plausible scenario (layers that coexist independent of each other are hard to imagine.
-There seems to be a good fit for the ground-truth admixture coefficients in the scenario you simulate. 
-However, you never assess how well you can fit isolation by distance parameters, 
-and whether your posterior distribution makes any sense. 
-This is a very important step in demonstrating that you implementation works - 
-and would give me and other users much needed reassurance.
-The authors also never give details such as migration rates of this simulations, 
-so the reader has effectively no idea what the extend of isolation by distance and differentiation is in these simulations. 
-Please put them at least in a supplement table. 
-In general it is really hard to get a sense for the power of the method, 
-as there is no exploration done in that direction. 
-Power will obviously depend on a plethora of factors, 
-but at least give an idea of what marker numbers you envision as typical input.
-Given the high-dimensional parameter space there is no realistic way to thoroughly test your inference scheme in this paper - 
-so I would be satisfied if you pick a set of parameters you deem realistic 
-(for instance similar to the ones you estimate in the applications), 
-and show that the inference method can infer them when simulated under the model. 
-In particular, show that your model infers also the isolation by distance parameters well, 
-and that the posteriors make sense. 
-In case it does not (due to model degeneracies or similar issues , see below), 
-it is not big deal as long as the ``important" admixture parameters are unaffected, 
-but then point these out.
-If there are space issues, you can put these plots in a short supplement.\\\\
-\gb{RESPONSE.}
-
-\item Also, I agree with Reviewer 1 that it would be interesting to see how your method behaves on a biologically plausible, 
-idealized, scenario, such as secondary contact with isolation by distance. 
-While the ground truth will be blurry for your model, 
-if the method infers the ``right clusters" and plausible admixture coefficients 
-this would give much credibility to your method - 
-showing that it also works when the model does not strictly hold, 
-as will always be the case in practice. 
-I don't consider this as an essential make-or-break though as the simulation-under-the-model.\\\\
-\gb{RESPONSE.}
-
-\item Issues that should be addressed:
- There is a typo or something missing in Eq. 8 and Eq. 10., 
- there appears an l whereas the covariance is already averaged over markers. 
- Do you miss a sum sign here? 
- Please clarify that as this masks an important assumption 
- that is almost certainly violated in most biological realistic scenarios:
-The covariance matrices are not the same across markers, 
-as drift depends on allele frequency. 
-In the extreme case of fixed markers the covariance is always 0! 
-In Population Genetics one often models this with a p(1-p) 
-pref-actor of the covariance matrix OR $F_{ST}$ summary statistics. 
-Obviously averaging over quite different covariance matrices 
-violates the multivariate normal Wishard approximation with 
-degrees of freedom equal to L. 
-Please elaborate on that issue in the ``technical supplement".
-Ringbauer2018 explored a way around this caveat via Fishers-Transform of allele
-frequencies or directly modeling F-matrices. (This sentence falls into III).
- Throughout the manuscript it is implied that Structure-like methods constitutes a subset of yours, 
- corresponding to turning off the spatial covariance component. 
- I have to disagree and think this is misleading: 
- Importantly, Structures also fits allele frequencies for each loci. 
- Thus, a concurrent cline of a few of them can already force the introduction of a new cluster. 
- In stark contrast, ConStruct only uses a single covariance matrix averaged across markers, 
- so will be relatively blind to this behavior of a few markers! 
- It seems important to mention that. 
- In some observed biological scenarios only few markers cause a hybrid zone.\\\\
-\gb{RESPONSE.}
-
-\item Put your work in context. 
- While I agree with your assessment that detailed comparisons on simulations is more than enough work for a own paper, 
- it would be nice to put your work into context of methods. 
- In particular discuss methods that can be applied on datasets that are interesting for your method. 
- In particular Geneland and EMS come to mind. 
- The latter will be better to deal with highly heterogeneous landscapes, with varying IBD patterns.
-Interestingly, you also compete with Bedassle. 
-Since it fits allele frequencies, it likely has higher power to confirm 
-hypothesized barriers that aren't detected by this method. 
-You could mention something along these lines.\\\\
-\gb{RESPONSE.}
-
-\item In line 97 to 111: The analogy to ``random flipping of allele state" is mildly interesting, 
-but does not add much insight. 
-You can cut it and start directly with your summary statistic (2) and only mention this analogy in the Supplement/Appendix.\\\\
-\gb{RESPONSE.}
-
-\item On multiple occasions you argue that the ``allelic covariance" matrix is insensitive to the choice of reference allele. 
-This is also true for most covariance matrices of the style (p1-p, p2-p), or do I miss something??
-The main advantage seems to be that you buy is that you don't have to estimate mean allele frequencies, which is MANY parameters. 
-You probably have taught quite a bit about this step ``down" from Bedasssle, so perhaps share some of your views and initial experiments.\\\\
-\gb{RESPONSE.}
-
-\item In line 72-73, you cite that mainly stepping-stone models gives us some analytical predictions. 
-In fact, good analytic approximations are often derived for continuous limits, such as a diffusion model.
-You could add some reference for continuous models that incorporate discrete discontinuities and try to merge these two. 
-Sometimes, there are analytic approximations, which is obviously relevant: 
-Importantly, the landmark work of Nayglaki1988 is missing. 
-Morerecently Barton2008 and Ringbauer2018 made some contributions 
-for the effects of barriers to gene flow on genetic covariance in a continuous setting. 
-They are not directly relevant for this work, but hey could provide useful context.\\\\
-\gb{RESPONSE.}
-
-\item Your covariance matrix has to be a valid covariance matrix, i.e. positive semidefinite. 
-I guess this is why you choose the power of the exponential to be between 0 and 2 
-(for other values it isn't), but you never mention that. 
-In table 1, some draws from the priors would not produce valid covariance matrices, 
-how do you proceed in these cases? 
-Do you have implementations that enforce this or are these areas never sampled in practice? 
-This issue could mess with the Wishart distribution and matrix inversion if the determinant crosses to negative values.\\\\
-\gb{RESPONSE.}
-
-\item There seem to be some possible degeneracy issues, 
-for instance when $\alpha_d$ goes to 0 (which seems to be the case for some of the inferred layers), 
-$\alpha_0$ and $\phi_k$ will becomes indistinguishable.
- You should mention these issues, and if you have dealt with them 
- (or think they are not important, as you are not really interested in estimating these parameters).\\\\
-\gb{RESPONSE.}
-
-\item Genetics states ``All articles published by GENETICS must include a 
-Data Availability Statement at the end of the Materials and Methods section." 
-You don't have such a section at the moment.
-For reproducibility of your results, 
-you should make at least the covariance matrix and the positions of each deme you used in your analysis available. 
-Ideally share all of the genetic data and the scripts used to filter it, including the settings (this is often quite important). 
-But if someone is protective of this data, I personally would be fine with omitting it. But this is the editors call to make.\\\\
-\gb{RESPONSE.}
-
-\item Fig. 4 could be relegated to the supplement, it does not show much more really important information as Fig. 2. 
- In particular if you need space.\\\\
-\gb{RESPONSE.}
-
-\item It would be really helpful if you plot the ground truth admixture coefficients in your K plots (Fig. 2) for K>2. 
- It seems you sometimes estimate admixture for populations which have none (also observed in my applications), 
- but not knowing the ground truth this is hard to gauge. 
- You randomly simulated them, so please add this ``ground truth" as an additional plot. 
- I am slightly skeptical about Plot S5 (the admixture coefficients) ? did you do averaging? 
- Having ground truth plots would help to remove my doubts!\\\\
-\gb{RESPONSE.}
-
-\item  Line 386: I think the claim that your model includes a biologically realistic model for isolation by distance is a bit overzealous. 
- Several non-interacting layers of IBD are a bit hard to imagine, 
- or do you have a good example?
- Importantly, your method will infer different clusters when isolation by distance patterns change. 
- So again you run into a problem of inferring a discrete break where there is none. 
- In fact, for most realistic populations one would expect different isolation by distance patterns.
-  Could be worth mentioning that caveat, and that in that case EEMS style methods.\\\\
-\gb{RESPONSE.}
-
-\item The headings are a bit off. 
-For instance you start with Results and then lay out the model? 
-Orient yourself on the excellent structure of the BEDASSLE paper.
- You cite often quite a few works at once (for instance ten in line 60), 
- check if this goes well with the citation style of Genetics).\\\\
-\gb{RESPONSE.}
-
-\item III) Suggestions (for future)
- When I run the inference scheme myself, 
- I noticed that mostly only one core on my machine is under full load, while the other remain idle. 
- You could conceive ways to incorporate multi-processing. 
- In particular, if run on a cluster, this could yield huge speed-ups. 
- There seems to be some literature out there in how to do so for the inversion of huge covariance matrices, 
- which seems to be the main bottleneck.\\\\
-\gb{RESPONSE.}
-
- \item You assume a multivariate normal distribution. 
- This implicitly assumes that all higher moments are 0, 
- which is quite the assumption for real data and almost certainly violated. 
- But In some sense, your method fits the covariances matrix - the pattern that it produces it does not matter. 
- This brings me directly to the next point:
- You could explore other ways to fit the covariance matrix. 
- Since typically the model is violated anyways, 
- you won't lose too much (the posterior is no ``real" posterior)
-For instance, fitting all pairwise covariance by for instance least square 
-is not necessarily inferior, but could give you significant speed updates 
-(quadratic, and no cumbersome inversion of the covariance matrix).
-Ringbauer2018 tried such approaches and found 
-that simpler fitting methods often have similar power than more complex ones 
-(whose assumptions are violated anyways), but are much faster.\\\\
-\gb{RESPONSE.}
-
- \item Right now you have to give the pairwise distance Matrix as parameter to your method, 
- and also the sample positions. 
- You could easily calculate the former from the latter, 
- and would save users some quick but unnecessary preprocessing step. 
- I guess in the future you might want to use other distance matrices though.\\\\
-\gb{RESPONSE.}
-
- \item In line 569 to line 575: When you describe how to deal with linked markers. 
- You don't mention it very explicitly, 
- but I think you hint that one can simply do ``block-bootstrap" of whole blocks of genome 
- as typically done nowadays for many applications, 
- if at least linkage groups are known.\\\\
-\gb{RESPONSE.}
-
- \item The Gamma-Exponential Kernels used here are often considered inferior to Matern-Kernels. 
- In practice it likely wont matter much though.\\\\
-\gb{RESPONSE.}
-
-\item I personally really fancy your writing style, 
-and it is a golden thread throughout your work.  
-But as much as I enjoy learning new words, 
-keep in mind many others in Science are non-mother tongue speakers 
-(and perhaps less literate native speakers). 
-And sometimes it simply feels like you are under contract of an English dictionary company.
-Additionally, a more concise and straightforward formulation 
-would significantly reduce the length of the manuscript and improve its readability 
-- while keeping the information content unchanged.
-
-One example of many, to give you an idea:
-``Existing model-based clustering approaches can only describe 
-continuous patterns of variation using discrete clusters,
- and tend to erroneously describe continuously distributed variation 
- with multiple clusters that show spatially autocorrelated cluster membership" (l.82-84)
-
-to
-
-``Most existing model-based clustering methods only use discrete units. 
-Therefore, they often partition continuous variation into spurious clusters. 
-(with spatially autocorrelated membership.)"
-
-This is merely a suggestion, choice of style is obviously the author's and there is
-no ground truth best way.\\\\
-\gb{RESPONSE.}
-
-\item Stress more that an easy-to-use implementation inclusive a manual exists, that is
-certainly a strong point! I was impressed by the ease of installation and use, and quality of the manual.\\\\
-\gb{RESPONSE.}
-
-\end{itemize}
-\end{document}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{itemize}
-
-\end{document}

### 2018-06-26 writeup/response2reviews.tex
added fancy review commands

@@
-\documentclass[11pt]{letter}
-\usepackage[hidelinks]{hyperref}
-\usepackage[green]{msuletterhead}
-\usepackage[usenames,dvipsnames]{xcolor}
-\pagenumbering{gobble}
+%%%%%%
+%%
+%%  Don't reorder the reviewer points; that'll mess up the automatic referencing!
+%%
+%%%%%
 
-\newcommand{\gb}[1]{{\bf\color{black}{#1}}}
-\newcommand{\plr}[1]{{\it\color{purple}{(#1)}}}
-\newcommand{\gc}[1]{{\it\color{blue}{(#1)}}}
+% to-do: get letterhead back in somehow? or not.
+% \usepackage[green]{msuletterhead}
 
-\begin{document}
+\begin{minipage}[b]{2.5in}
+  Resubmission Cover Letter \\
+    {\it Genetics} 
+\end{minipage}
+\hfill
+\begin{minipage}[b]{2.5in}
+  \today
+\end{minipage}
 
-Re-Submission Cover Letter \hfill	\today\\
-\emph{Genetics}
-\\\\\\
+\vskip 2em
+
+\noindent
+
 To the Editor,
@@
 
-We hope that the paper is now acceptable for publication.\\\\
-%
+We hope that the paper is now acceptable for publication.
 
-Thank you for your consideration,\\
-Gideon Bradburd (for all of the authors).
+\noindent \hspace{4em}
+\begin{minipage}{3in}
+\noindent
+    {\bf Thank you for your consideration,}
+
+\vskip 2em
+
+{\bf 
+    Gideon Bradburd (for all of the authors)
+}\\
+\end{minipage}
+
+\vskip 4em
+
+\pagebreak
 
@@
 
-\textbf{Editor}\\
-We received the previous reviews for your paper, and sent also sent it out to a new reviewer, along with your response to the earlier reviews.
-Both the new reviewer and I believe that you have answered the previous criticisms in a reasonable way, and I am pleased to tell you that your paper is potentially suitable for publication in GENETICS.
-However, the new reviewer did have several comments which you should consider in preparing the final version (see the end of this email).
-In particular, they suggest including more of the ADMIXTURE models in the main text.
-I think that this could be helpful, but would not insist on it.
-A second review came in late (attached).
-I am sending it on in case you would like to take account of any of the suggestions in your revision.
-The reviewer has quite a few technical suggestions that may be helpful.
-However, even if it had come in before the decision, I would not have required a wider range of simulations.
+%%%%%%%%%%%%%%
+\reviewersection{Editor}
 
-\textbf{Reviewer 1}\\
-\begin{itemize}
-\item This manuscript presents a valuable model that addresses an long standing chestnut in population genetic analysis, which is how to jointly analyze discrete and spatially continuous variation within a single dataset.
+\begin{quote}
+    We received the previous reviews for your paper, and sent also sent it out
+    to a new reviewer, along with your response to the earlier reviews. Both the
+    new reviewer and I believe that you have answered the previous criticisms in a
+    reasonable way, and I am pleased to tell you that your paper is potentially
+    suitable for publication in GENETICS. However, the new reviewer did have
+    several comments which you should consider in preparing the final version (see
+    the end of this email). In particular, they suggest including more of the
+    ADMIXTURE models in the main text. I think that this could be helpful, but
+    would not insist on it.  A second review came in late (attached).  I am sending
+    it on in case you would like to take account of any of the suggestions in your
+    revision. The reviewer has quite a few technical suggestions that may be
+    helpful.  However, even if it had come in before the decision, I would not have
+    required a wider range of simulations.  
+\end{quote}
+
+%%%%%%%%%%%%%%
+\reviewersection{1}
+
+\begin{point}{}
+    This manuscript presents a valuable model that addresses an long standing chestnut in population genetic analysis, which is how to jointly analyze discrete and spatially continuous variation within a single dataset.
 The method presented does seem to be something of a breakthrough in addressing this problem sensibly.
 I found the authors response to previous reviewer comments to be generally reasonable.
-However, I do think that there is a strong case for putting some more ADMIXTURE model results into the text and possibly for some additional analyses.\\\\
-\gb{We have now added some of the ADMIXTURE results to the main text, 
+However, I do think that there is a strong case for putting some more ADMIXTURE model results into the text and possibly for some additional analyses.
+\end{point}
+
+\reply{We have now added some of the ADMIXTURE results to the main text, 
 including the admixture pie plot and the cross-validation results from the ADMIXTURE analyses 
 of the bear dataset in Figures 7 and 8, respectively. 
-We have also included more detailed comparisons between ADMIXTURE and \texttt{conStruct} (see more below).}
+We have also included more detailed comparisons between ADMIXTURE and \texttt{conStruct},
+for instance at \llname{ll:admixture_sim_results} and \llname{ll:admixture_sim_xval};
+see below for more.}
 
-\item I agree with the authors point in their letter that ADMIXTURE/STRUCTURE is important because it is what almost everyone would do otherwise.
+\begin{point}{}
+    I agree with the authors point in their letter that ADMIXTURE/STRUCTURE is important because it is what almost everyone would do otherwise.
 The key goal for the manuscript is to show that the method that everyone uses gives results that can be misleading/unhelpful and that the layered model is a substantial improvement.
@@
 The previous reviewers are right to ask for additional simulations in the sense that a more biological simulation would make a stronger case for practical utility and it need not lengthen the manuscript significantly, since the current simulations could be largely relegated to supplementary material as representing ``simulation according to the model".
-But this is definitely an optional, since there are presentational arguments to be made either way.\\\\
-\gb{WILL FILL IN AFTER WE FIND OUT WHAT NEWS WEDNESDAY BRINGS}
+But this is definitely an optional, since there are presentational arguments to be made either way.
+\end{point}
 
-\item A more biological simulation is not necessary as long as the real data examples are convincing.
+\reply{
+    We agree that more simulations would be nice,
+    but after serious consideration, have decided it would be too far outside the scope
+    of the already lengthy manuscript.
+    The existing simulations are not quite ``according to the model'' --
+    that would be simulations where we draw sample frequencies directly from a multivariate Gaussian
+    (which is something we have done as a sanity check, but have not included).
+    We feel that the existing simulations draw a nice compromise between realism
+    and being close enough to the model that we still have a good idea of what the correct answer is.
+    We do have plans to follow up with a more detailed simulation study.
+}
+
+\begin{point}{}
+    A more biological simulation is not necessary as long as the real data examples are convincing.
 For the Poplar data, the key insight of the model is that there are three well-supported layers in addition to isolation by distance.
@@
 For example at K=3, a component is shown that is shared between East and West, which if interpreted literally would represent a biologically important result about connections between populations but the clustering of the yellow populations disappears entirely for K=4.
-Therefore, I think there is a good case to be made for showing the ADMIXTURE model results in the text and to discuss the misleading/ unstable nature of any particular result.\\\\
-\gb{We have added more extensive discussions of the 
-comparison between \texttt{conStruct} and ADMIXTURE for both empirical datasets.
-These subsections appear on lines XXX-XXX for the poplar dataset, 
-and lines XXX-XXX for the bear dataset.}
+Therefore, I think there is a good case to be made for showing the ADMIXTURE model results in the text and to discuss the misleading/ unstable nature of any particular result.
+\end{point}
 
-\item It would be interesting to do some kind of plot for ADMIXTURE results, to demonstrate that populations that share affinity according higher components are not particularly similar when allowing for geographical distance.
+\reply{We have added more extensive discussions of the 
+comparison between \texttt{conStruct} and ADMIXTURE for both empirical datasets,
+following \llname{ll:admixture_populus} and \llname{ll:admixture_bears}.
+\plr{We could do this by talking about or including figure S28?}
+}
+
+\begin{point}{}
+    It would be interesting to do some kind of plot for ADMIXTURE results, to demonstrate that populations that share affinity according higher components are not particularly similar when allowing for geographical distance.
 This would be a rather direct way of showing that the ADMIXTURE model is not producing insightful results, despite the continuous improvement in model probability for higher K.
 In this context, I always think of the ``clines vs clusters" paper by Rosenberg.
-It would be great to have some kind of figure to show that for K=2, it looks like clusters but for K=3, the clusters add much less.\\\\
-\gb{We didn't follow exactly what this figure might look like, 
+It would be great to have some kind of figure to show that for K=2, it looks like clusters but for K=3, the clusters add much less.
+\end{point}
+
+\reply{We didn't follow exactly what this figure might look like, 
 but hopefully this point comes across clearly in the contrast between 
@@
 
-\item The layer specific covariances show a very funny pattern, tending to infinity for low distances for components 3 and above for the bear data.
+\begin{point}{}
+    The layer specific covariances show a very funny pattern, tending to infinity for low distances for components 3 and above for the bear data.
 I am not sure I understand this.
-Is this discussed anywhere?\\\\
-\gb{We think what's happening is that these layer-specific covariances for layers $>$2 
+Is this discussed anywhere?
+\end{point}
+
+\reply{We think what's happening is that these layer-specific covariances for layers $>$2 
 are describing local departures from global patterns of IBD, 
@@
 
-\item Minor point, in figure 4, legend, should it read in each run of the spatial model, rather than nonspatial model?\\\\
-\gb{Good catch!  Caption has been fixed.}
-\end{itemize}
+\begin{point}{}
+    Minor point, in figure 4, legend, should it read in each run of the spatial model, rather than nonspatial model?
+\end{point}
 
-\textbf{Reviewer 2}\\
-\begin{itemize}
-\item Make-or-Break:
+\reply{Good catch!  Caption has been fixed.}
+
+%%%%%%%%%%%%%%
+\reviewersection{2}
+
+First, we want to thank Reviewer 2 for their very thorough  review
+(even taking the time to check out the R package!).  
+We really appreciate the time and thought the reviewer invested here.
+
+\begin{point}{}
+    Make-or-Break:
 Ideally one tests a new method on simulated data, to assess its power -
@@
 but then point these out.
-If there are space issues, you can put these plots in a short supplement.\\\\
-\gb{First, we want to thank Reviewer 2 for their very thorough  review
-(even taking the time to check out the R package!).  
-We really appreciate the time and thought the reviewer invested here.
-RESPONSE.}
+If there are space issues, you can put these plots in a short supplement.
+\end{point}
 
-\item Also, I agree with Reviewer 1 that it would be interesting to see how your method behaves on a biologically plausible, 
+\reply{
+As we discussed in reponse to the other reviewer,
+we intentionally chose simulations in this in-between spot,
+that are somewhat biologically realistic 
+while still retaining a clear way of obtaining the ``right'' answer.
+Additional sets of simulations will, in our opinion, make the paper too lengthy and complex;
+the applications to real data we hope demonstrate the real-world utility of the method.
+We have done ``sanity check'' simulations under the model itself,
+but as these are far enough from biological reality (e.g., the ``frequencies'' are not between 0 and 1),
+we think they'd only make the paper more confusing.
+As for the migration rates -- these are given in the text already \llname{ll:sim_params},
+and we have added an explicit reference to this \llname{ll:sim_params_ref}.
+}
+
+\begin{point}{}
+    Also, I agree with Reviewer 1 that it would be interesting to see how your method behaves on a biologically plausible, 
 idealized, scenario, such as secondary contact with isolation by distance. 
@@
 as will always be the case in practice. 
-I don't consider this as an essential make-or-break though as the simulation-under-the-model.\\\\
-\gb{RESPONSE.}
+I don't consider this as an essential make-or-break though as the simulation-under-the-model.
+\end{point}
 
-\item Issues that should be addressed:
+\reply{
+We realize the coalescent simulations we provide may not satisfy the reviewer's curiosity
+regarding isolation by distance,
+but note that the tools required to do genome-scale continuous-space simulations 
+have only recently become available (in SLiM), long after submission of this paper;
+and have not been officially released yet.
+}
+
+\begin{point}{}
+    Issues that should be addressed:
  There is a typo or something missing in Eq. 8 and Eq. 10., 
@@
 Ringbauer2018 explored a way around this caveat via Fishers-Transform of allele
-frequencies or directly modeling F-matrices. (This sentence falls into III).\\\\
-\gb{Good catch!  Yes, there was a typo (that $\ell$ subscript didn't belong), 
+frequencies or directly modeling F-matrices. (This sentence falls into III).
+\end{point}
+
+\reply{Good catch!  Yes, there was a typo (that $\ell$ subscript didn't belong), 
 and we have fixed the error. 
@@
 
-\item Throughout the manuscript it is implied that Structure-like methods constitutes a subset of yours, 
+\begin{point}{}
+    Throughout the manuscript it is implied that Structure-like methods constitutes a subset of yours, 
  corresponding to turning off the spatial covariance component. 
@@
  It seems important to mention that. 
- In some observed biological scenarios only few markers cause a hybrid zone.\\\\
-\gb{RESPONSE.}
+ In some observed biological scenarios only few markers cause a hybrid zone.
+\end{point}
 
-\item Put your work in context. 
+\reply{RESPONSE.}
+
+\begin{point}{}
+    Put your work in context. 
  While I agree with your assessment that detailed comparisons on simulations is more than enough work for a own paper, 
@@
 hypothesized barriers that aren't detected by this method. 
-You could mention something along these lines.\\\\
-\gb{We have added more text comparing conStruct with Geneland and EEMS and BEDASSLE.}
+You could mention something along these lines.
+\end{point}
 
-\item In line 97 to 111: The analogy to ``random flipping of allele state" is mildly interesting, 
+\reply{We have added more text comparing conStruct with Geneland and EEMS and BEDASSLE.}
+
+\begin{point}{}
+    In line 97 to 111: The analogy to ``random flipping of allele state" is mildly interesting, 
 but does not add much insight. 
-You can cut it and start directly with your summary statistic (2) and only mention this analogy in the Supplement/Appendix.\\\\
-\gb{We have edited that section for clarity and brevity.}
+You can cut it and start directly with your summary statistic (2) and only mention this analogy in the Supplement/Appendix.
+\end{point}
 
-\item On multiple occasions you argue that the ``allelic covariance" matrix is insensitive to the choice of reference allele. 
+\reply{We have edited that section for clarity and brevity.}
+
+\begin{point}{}
+    On multiple occasions you argue that the ``allelic covariance" matrix is insensitive to the choice of reference allele. 
 This is also true for most covariance matrices of the style (p1-p, p2-p), or do I miss something??
 The main advantage seems to be that you buy is that you don't have to estimate mean allele frequencies, which is MANY parameters. 
-You probably have taught quite a bit about this step ``down" from Bedasssle, so perhaps share some of your views and initial experiments.\\\\
-\gb{RESPONSE.}
+You probably have taught quite a bit about this step ``down" from Bedasssle, so perhaps share some of your views and initial experiments.
+\end{point}
 
-\item In line 72-73, you cite that mainly stepping-stone models gives us some analytical predictions. 
+\reply{RESPONSE.}
+
+\begin{point}{}
+    In line 72-73, you cite that mainly stepping-stone models gives us some analytical predictions. 
 In fact, good analytic approximations are often derived for continuous limits, such as a diffusion model.
@@
 for the effects of barriers to gene flow on genetic covariance in a continuous setting. 
-They are not directly relevant for this work, but hey could provide useful context.\\\\
-\gb{Thanks for pointing out these missing and important references.  
+They are not directly relevant for this work, but hey could provide useful context.
+\end{point}
+
+\reply{Thanks for pointing out these missing and important references.  
 We have added them to the manuscript.}
 
-\item Your covariance matrix has to be a valid covariance matrix, i.e. positive semidefinite. 
+\begin{point}{}
+    Your covariance matrix has to be a valid covariance matrix, i.e. positive semidefinite. 
 I guess this is why you choose the power of the exponential to be between 0 and 2 
@@
 Do you have implementations that enforce this or are these areas never sampled in practice? 
-This issue could mess with the Wishart distribution and matrix inversion if the determinant crosses to negative values.\\\\
-\gb{Yes, as you say the covariance must be positive semi-definite for us to invert it and calculate the likelihood.
+This issue could mess with the Wishart distribution and matrix inversion if the determinant crosses to negative values.
+\end{point}
+
+\reply{Yes, as you say the covariance must be positive semi-definite for us to invert it and calculate the likelihood.
 The inference algorithm implemented by STAN automatically rejects proposed matrices that aren't invertible.
@@
 
-\item There seem to be some possible degeneracy issues, 
+\begin{point}{}
+    There seem to be some possible degeneracy issues, 
 for instance when $\alpha_d$ goes to 0 (which seems to be the case for some of the inferred layers), 
@@
  You should mention these issues, and if you have dealt with them 
- (or think they are not important, as you are not really interested in estimating these parameters).\\\\
-\gb{In practice, because those parameters can't be easily interpreted in a biological context, 
+ (or think they are not important, as you are not really interested in estimating these parameters).
+\end{point}
+
+\reply{In practice, because those parameters can't be easily interpreted in a biological context, 
 we treat them more as nuisance parameters and focus on inference of the admixture proportions.
@@
 
-\item Genetics states ``All articles published by GENETICS must include a 
+\begin{point}{}
+    Genetics states ``All articles published by GENETICS must include a 
 Data Availability Statement at the end of the Materials and Methods section." 
@@
 Ideally share all of the genetic data and the scripts used to filter it, including the settings (this is often quite important). 
-But if someone is protective of this data, I personally would be fine with omitting it. But this is the editors call to make.\\\\
-\gb{Thanks for pointing this out.  
+But if someone is protective of this data, I personally would be fine with omitting it. But this is the editors call to make.
+\end{point}
+
+\reply{Thanks for pointing this out.  
 We have added a Data Availability Statement, 
@@
 
-\item Fig. 4 could be relegated to the supplement, it does not show much more really important information as Fig. 2. 
- In particular if you need space.\\\\
-\gb{We like Fig. 4 as a demonstration of the utility of calculating layer 
+\begin{point}{}
+    Fig. 4 could be relegated to the supplement, it does not show much more really important information as Fig. 2. 
+ In particular if you need space.
+\end{point}
+
+\reply{We like Fig. 4 as a demonstration of the utility of calculating layer 
 contributions as a heuristic of model comparison. 
@@
 
-\item It would be really helpful if you plot the ground truth admixture coefficients in your K plots (Fig. 2) for $K>2$. 
+\begin{point}{}
+    It would be really helpful if you plot the ground truth admixture coefficients in your K plots (Fig. 2) for $K>2$. 
  It seems you sometimes estimate admixture for populations which have none (also observed in my applications), 
@@
  I am slightly skeptical about Plot S5 (the admixture coefficients) ? did you do averaging? 
- Having ground truth plots would help to remove my doubts!\\\\
-\gb{RESPONSE.}
+ Having ground truth plots would help to remove my doubts!
+\end{point}
 
-\item  Line 386: I think the claim that your model includes a biologically realistic model for isolation by distance is a bit overzealous. 
+\reply{RESPONSE.}
+
+\begin{point}{}
+    Line 386: I think the claim that your model includes a biologically realistic model for isolation by distance is a bit overzealous. 
  Several non-interacting layers of IBD are a bit hard to imagine, 
@@
  In fact, for most realistic populations one would expect different isolation by distance patterns.
-  Could be worth mentioning that caveat, and that in that case EEMS style methods.\\\\
-\gb{RESPONSE.}
+  Could be worth mentioning that caveat, and that in that case EEMS style methods.
+\end{point}
 
-\item The headings are a bit off. 
+\reply{RESPONSE.}
+
+\begin{point}{}
+    The headings are a bit off. 
 For instance you start with Results and then lay out the model? 
@@
  You cite often quite a few works at once (for instance ten in line 60), 
- check if this goes well with the citation style of Genetics).\\\\
-\gb{These were holdovers from a previous submission format.  
+ check if this goes well with the citation style of Genetics).
+\end{point}
+
+\reply{These were holdovers from a previous submission format.  
 We have edited them for a more logical flow and greater clarity.}
 
-\item III) Suggestions (for future)
+\begin{point}{}
+    III) Suggestions (for future)
  When I run the inference scheme myself, 
@@
  There seems to be some literature out there in how to do so for the inversion of huge covariance matrices, 
- which seems to be the main bottleneck.\\\\
-\gb{Parallelization would definitely be helpful.  
+ which seems to be the main bottleneck.
+\end{point}
+
+\reply{Parallelization would definitely be helpful.  
 Right now, the inference algorithm is implemented in STAN, 
@@
 
- \item You assume a multivariate normal distribution. 
+\begin{point}{}
+     You assume a multivariate normal distribution. 
  This implicitly assumes that all higher moments are 0, 
@@
 that simpler fitting methods often have similar power than more complex ones 
-(whose assumptions are violated anyways), but are much faster.\\\\
-\gb{This is a nice idea, and one that we will definitely look into.
+(whose assumptions are violated anyways), but are much faster.
+\end{point}
+
+\reply{This is a nice idea, and one that we will definitely look into.
 We've talked about including a ``fast" conStruct option  
@@
 
- \item Right now you have to give the pairwise distance Matrix as parameter to your method, 
+\begin{point}{}
+    Right now you have to give the pairwise distance Matrix as parameter to your method, 
 and also the sample positions. 
@@
 and would save users some quick but unnecessary preprocessing step. 
- I guess in the future you might want to use other distance matrices though.\\\\
-\gb{Right, it would save users some time for us to calculate pairwise distances for them,
+I guess in the future you might want to use other distance matrices though.
+\end{point}
+
+\reply{Right, it would save users some time for us to calculate pairwise distances for them,
 but at the expense of flexibility.
@@
 
- \item In line 569 to line 575: When you describe how to deal with linked markers. 
+\begin{point}{}
+    In line 569 to line 575: When you describe how to deal with linked markers. 
 You don't mention it very explicitly, 
@@
 as typically done nowadays for many applications, 
- if at least linkage groups are known.\\\\
-\gb{Yes, we think that's likely the best way to do it when linkage is known.
+if at least linkage groups are known.
+\end{point}
+
+\reply{Yes, we think that's likely the best way to do it when linkage is known.
 We've added some text about that in the ``Cross-validation" section of the Appendix 
@@
 
- \item The Gamma-Exponential Kernels used here are often considered inferior to Matern-Kernels. 
- In practice it likely wont matter much though.\\\\
+\begin{point}{}
+    The Gamma-Exponential Kernels used here are often considered inferior to Matern-Kernels. 
+In practice it likely wont matter much though.
+\end{point}
+
 \gb{Yes, one of the next updates we plan is to replace the flexible but 
@@
 
-\item I personally really fancy your writing style, 
+\begin{point}{}
+    I personally really fancy your writing style, 
 and it is a golden thread throughout your work.  
@@
 (with spatially autocorrelated membership.)"
+\\\\
+This is merely a suggestion, choice of style is obviously the author's and there is
+no ground truth best way.
+\end{point}
 
-This is merely a suggestion, choice of style is obviously the author's and there is
-no ground truth best way.\\\\
-\gb{First, thank you for the compliment (seriously - ``a golden thread"!).
+\reply{First, thank you for the compliment (seriously - ``a golden thread"!).
 We appreciate that not all scientists are native English speakers, 
@@
 
-\item Stress more that an easy-to-use implementation inclusive a manual exists, that is
-certainly a strong point! I was impressed by the ease of installation and use, and quality of the manual.\\\\
-\gb{Good suggestion - we have now indicated that the method is available as a documented R package.}
+\begin{point}{}
+    Stress more that an easy-to-use implementation inclusive a manual exists, that is
+certainly a strong point! I was impressed by the ease of installation and use, and quality of the manual.
+\end{point}
 
-\end{itemize}
-\end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{itemize}
-\item First, many results concern comparisons between clustering based on the spatial and non-spatial models of construct (Figs 2, 3, 4, 6, 7, 9).
-However the authors should rather compare their spatial approach to existing software for model-based clustering, which do not account for space (Structure, fastSTRUCTURE, ADMIXTURE,...).
-Results obtained with one of these software should not be available in supplementary material but should rather be visible in the main text.\\\\
-\gb{The utility of the nonspatial model implemented in \texttt{conStruct} 
-is that its performance can be directly compared to that of the spatial model; 
-their cross-validation predictive accuracy values are in the same currency 
-because they use the same Wishart likelihood model.
-So, using it we can say not just that that the nonspatial model prefers 
-$K=X$ while the spatial model prefers $K=Y$, 
-but also, e.g., that the spatial model is preferred over the nonspatial model.
-However, as you note, it's important to compare \texttt{conStruct}'s performance to 
-that of existing nonspatial model-based clustering methods as well, 
-which is why we had also included analyses using fastSTRUCTURE.
-
-\ Your point here is well-taken though, because the fastSTRUCTURE analyses, 
-which included no model comparisons, did not provide a direct basis of comparison to \texttt{conStruct}.
-There is a model comparison tool implemented in fastSTRUCTURE, 
-but we were unable to get it to run on our datasets.
-We have contacted the fastSTRUCTURE authors 
-and alerted them to this issue, 
-but have yet to hear back from them.
-
-\ We have therefore analyzed all datasets included in the paper using ADMIXTURE, 
-and used ADMIXTURE's cross-validation procedure to compare model performance across values of $K$.  
-These results are discussed extensively throughout the updated manuscript ``Results" section; 
-and the model that minimizes the cross-validation error in the ADMIXTURE analyses on simulated data 
-is highlighted in the the simulation cross-validation results figure.
-However, due to the already extensive length of the manuscript (which Reviewer 2 points out), 
-the results of the ADMIXTURE analyses are given in the Supplementary Materials.
-}\\
-
-\item Second, the simulation scenario should correspond to a real biological scenario.
-Here this layer-based simulation scenario is very difficult to understand.
-I would be more interesting to stick to more classic spatially explicit simulations such as range expansion, or secondary contact scenarios.\\\\
-\gb{We appreciate that the simulation scenario we ran is somewhat artificial, 
-and that it would be good to test the behavior of the method 
-under a classic ``isolation followed by expansion into a zone of secondary contact" model.
-However, as we discuss in the manuscript in the section ``Simulation details",
-we wanted to simulate datasets that had unambiguous true admixture proportions, 
-so that we could test \texttt{conStruct}'s ability to estimate admixture accurately 
-(shown in SuppMat figures 8 and 13).
-A more organic simulation, with isolation followed by expansion, 
-secondary contact, and subsequent admixture, 
-would not facilitate that test in a straightforward way. 
-As mentioned by Reviewer 2, there are an enormous number 
-of possible simulations we could run to test \texttt{conStruct}'s performance, 
-but fully exploring them is beyond of the scope of this manuscript, 
-so we intend to pursue those in future work.}\\
-
-\item Third, I think that the main point of the paper should be more explicit.
-Non-spatial model based approaches are prone to overfitting when K is too large and this can be avoided when using spatial-based approaches.
-\\\\
-\gb{
-Good suggestion; we've added a line to the Abstract to this effect.
-}\\
-
-\item Last, there are other software that account for space when performing model-based clustering and they should be included in the comparisons (BAPS, Geneland, TESS,...).
-Right now, they are only mentioned in the introduction and we would like to see results comparing these different model-based approaches.\\\\
-\gb{
-The methods mentioned here --- BAPS, Geneland, and TESS --- 
-place spatial priors on admixture proportions
-and/or cluster membership, 
-but like STRUCTURE, they assume that allele frequencies 
-within a cluster are constant across space.
-Therefore, although spatially explicit, this parameterization does not 
-account for space in a way that avoids the problem we describe in our manuscript: 
-the characterization of spatially continuous differentiation 
-as clines of admixture across spurious clusters.
-
-\ A comprehensive comparison between \texttt{conStruct} 
-and all other model-based clustering methods is beyond the scope of this paper. 
-As STRUCTURE is the canonical model-based clustering algorithm
-(having been cited $>$ 20,000 times), but is too slow to be run 
-on the modern-scale SNP datasets we simulate, 
-we feel that ADMIXTURE, which implements an extremely similar model 
-within a faster inference framework, 
-provides the best comparison to our method.}\\
-
-\item In the introduction, the authors write ``placing spatial priors on cluster memberships'' when referring to other methods that deal with ancestry proportions instead of cluster memberships.\\\\
-\gb{
-Guillot \emph{et al.}, in the 2005 paper that introduced the method Geneland, 
-refer to this quantity as ``population membership."
-We kept the wording consistent for clarity, but changed ``population" 
-to ``cluster" to keep a distinction between biological and statistical groupings.}\\
-
-\item Formula (1) about the covariance is difficult to understand.
-It does not correspond to a classic definition of a covariance.
-Does this correspond to the covariance of some random variables?
-The discussion following the definition of the covariance looks like a discussion and should be moved to the discussion.\\\\
-
-\gb{The definition given is a covariance of some random variables; we have made this fact more clear, 
-and we have substantially rewritten the section on the allelic covariance
-and have moved some of the intuition-building discussion to the Materials and Methods.
-We have also moved some of this discussion to the Discussion (heading ``Allelic or genetic covariance?'').
-}\\
-
-\end{itemize}
-
-\textbf{Reviewer 2}\\
-In this paper, Bradburd et al.
-present a novel analysis method for the joint characterization of discrete and continuous structure.
-Their model uses a mixed membership Gaussian process, adding a within layer (cluster) Isolation-by-distance-process.
-I do think that this paper is very clearly written, and presents the model and method in great detail.
-Besides the presentation of the method, a simulation and two application studies are performed, that lead to a comprehensive overview of the method.
-I would consider this model to be very valuable for a large audience in ecology and evolution, including biogeographers and conservation biologists, with the only drawback being the very high computational requirements.
-Overall, however I recommend the paper for acceptance with only very minor changes.
-
-\begin{itemize}
-\item One issue I am wondering about is whether there are some issues with ``outlier individuals'', that e.g.
-represent poorly sampled clusters, which might be an issue with the Alaskan/NW-Canadian bears.
-In particular, consider the case where the two layers were only fitted from the samples from the 48-states.
-Presumably, the resulting covariance curve would be similar to Fig S26b (assuming the two layer case).
-If we then added those NW-pacific samples, we might expect them to be most likely added to the cluster with the flatter right tail, even if they were essentially unrelated.
-This issue can also arise in kriging (Figure 6 in the first version of the preprint of Caye et al (on bioRxiv).
-Could this interpretation be excluded somehow?
-\\\\
-\gb{This is a great point, and highlights a subtle problem.
-Although we feel that isolation-by-distance is a reasonable model for population differentiation, 
-it is possible that, by implementing a spatial model, 
-we can facilitate ``spurious" layer membership in far-flung samples, 
-who are sampled far enough away from all other samples 
-that their spatial covariance is negligible.  
-This situation is made less likely by the $\phi$ parameter in each layer, 
-which induces some amount of covariance between samples in a layer 
-regardless of their geographic separation.
-However, it is possible that $\phi$ in a layer might be estimated to be $\approx 0$. 
-We've discussed heuristics for highlighting such samples 
-(e.g., highlighting pairs of samples with $>$50\% membership in a given layer,
- that, given their geographic separation, 
- the parametric covariance curve in that layer, 
- and the $\phi$ parameter in that layer, have effectively zero covariance), 
- but we have not yet implemented or tested these checks.
-We are also currently developing a set of spatial \textit{f}-statistics, 
-analogous to those presented in Reich \emph{et al.} (2009) and discussed further in Peter (2016), 
-which could be useful in highlighting such ``outlier" samples.
-But, as you note, this manuscript is already quite long, 
-so this work will be a part of future projects.
-}\\
-
-\item Another thing that could be added are some additional simulations in more complex demographic models, such as expansions, temporally/spatially heterogeneous migration, or sparser sampling schemes.
-However, as the paper is already rather long (at almost 20 single-spaced pages), and the space of scenarios for which the model is relevant is enormous, I hope the authors consider further exploring their model in a follow-up study.\\\\
-\gb{We completely agree, and further explorations of the behavior of the method 
-in different scenarios are being planned now.}\\
-
-\item 100: Is there a reason that allelic covariance is used instead of genetic covariance?\\\\
-\gb{
-    Good question; we've added some more text for this, 
-    both explaining what the allelic covariance is and why we use it in lieu of the genetic covariance,
-    (in the ``Data" subsection of the ``Results" 
-    and the ``Allelic covariance" subsection of ``Materials and Methods").
-    The short answer is that the allelic covariance is insensitive to the choice of reference allele, 
-    and, because it doesn't subtract off means like the genetic covariance, 
-    has a constant value between a pair of samples in the dataset, 
-    regardless of which other samples are included.
-}\\
-
-\item Could supp. figures be sorted according to first mentioning in ms?\\\\
-\gb{Fixed.}\\
-
-\item Fig3: increase label size\\\\
-\gb{Fixed.}\\
-
-\item Last line in paragraph on p. 18: as the derivations are otherwise very detailed and easy to follow, it might be worth adding here that this follows from E[UV] = 4 Cov[A, B]\\\\
-\gb{
-We have substantially rewritten this section for greater clarity.\\
-}
-
-\item 634: Using ms,
-\\\\
-\gb{Fixed.}\\
-
-\item Fig S4: Could the colors be made to correspond each other (e.g.
-panel e vs 3).
-In general, is the exchangeability of layers a practical concern, or similar to the original structure, does the MCMC never ``switch'' two layer labels?\\\\
-\gb{In STAN, we declare the vector of $\phi$ parameters to be an object of class ``positive ordered",
-	meaning that values of phi must all have values greater than zero, 
-	and must be in strictly increasing order.
-	In practice, I don't think that much label-switching would occur within the course of a single MCMC run 
-	(like in STRUCTURE), but inducing a specific order to the $\phi$ parameter makes that less likely still.
-	There can be label switching across independent runs for the same value of $K$, 
-	and of course there's label switching across runs with different numbers of $K$.
-	In the \texttt{conStruct} R package, we've included a function that uses the cluster matching 
-	algorithm described in CLUMPP (Jakobsson and Rosenberg 2007) to match layers across runs.
-	We used that algorithm to keep plotting colors consistent in our figures, including in Fig. S4.
-	However, something funny happens in that sequence of analyses between $K=5$ and $K=6$, 
-	where the ranking of cluster matching algorithm gives wonky results, 
-	leading to the non-intuitive layer color matching.
-	We've tweaked that one by hand to keep the plotting nice.
- }\\
-
-\item Fig S26: maybe keep Y-axis scale consistent between panels?\\\\
-\gb{Fixed.}\\
-
-\item Cross-Validation: As we all know SNP are not uncorrelated.
-I am wondering if that could influence the CV, as spurious patterns in the testing data might be partially explained by nearby SNP in the training data.
-Is that a concern?\\\\
-\gb{
-	This is a good point and it definitely could be a concern.
-	We've included a discussion of this in the cross-validation procedure description, 
-	as well as a description of the CV procedure that should be used if you know LD between loci.
-	But, we also imagine that, for a while at least, the vast majority of \texttt{conStruct} users 
-	won't have genomic position information for their SNP data.
-}\\
-
-\end{itemize}
-
-\end{document}
+\reply{Good suggestion - we have now indicated that the method is available as a documented R package.}

### 2018-06-26 writeup/conStruct.tex
added fancy review commands

@@
 
+% 
+\bibliographystyle{latex/plos2015}
+
 % cite package, to clean up citations in the main text. Do not remove.
@@
 \renewcommand{\figurename}{Fig}
-
-% Use the PLoS provided BiBTeX style
-\bibliographystyle{latex/plos2015}
 
@@
 \newcommand{\bone}{\mathbf{1}}
+
+%% for submitting with responses to reviews
+\newif\ifsubmission
+% \submissiontrue
+\submissionfalse
+\ifsubmission
+    % % stuff for submission
+    \usepackage[nofiglist,nomarkers,figuresonly]{endfloat}
+    \usepackage{setspace}
+    % \doublespacing
+    % \renewcommand\includegraphics{\relax}
+    % \renewcommand{\includegraphics}[2][]{\relax}
+    \renewcommand{\figurename}{Fig.}
+    \newcommand{\Figure}{Fig }
+    \newcommand{\Figures}{Fig }
+\else
+    % for the nice version
+    % \usepackage[square,sort,comma,numbers]{natbib} \bibpunct{(}{)}{;}{author-year}{}{,} 
+    % \bibliographystyle{plainnat}
+    \newcommand{\Figure}{{Figure }}
+    \newcommand{\Figures}{{Figures }}
+\fi
+
+% Responses to reviews
+\input{review-response-commands}
+% set this to show line numbers and include responses to reviews or not
+\newif\ifreviewresponses
+\reviewresponsestrue  % include them
+% \reviewresponsesfalse  % don't include them
+\newcommand{\responsefile}{response2reviews}  % name of the review responses file
 
@@
 choosing various amounts of admixture from these different layers
-(see Fig \ref{sim_setup}). 
-We collapsed the layers together using random, spatially autocorrelated admixture proportions 
-(see \secref{sim_details}).
+(see Fig \ref{sim_setup}),
+and collapsed the layers together using random, spatially autocorrelated admixture proportions.
+See \secref{sim_details} for more details, including parameter values used. \llname{ll:sim_params_ref}
 We simulated datasets using $K=1$, 2, and 3 layers;
@@
 The ADMIXTURE results are qualitatively similar, as shown in Figs \ref{admix_simK1}-\ref{admix_simK3}.
+\linelabel{ll:admixture_sim_results}
 Each nonspatial cluster is genetically more similar within itself
@@
 and, in more detail, in Fig \ref{sim_admix_CVerrors}.
+\linelabel{ll:admixture_sim_xval}
 
@@
 
-\paragraph{Comparison to ADMIXTURE}
+\paragraph{Comparison to ADMIXTURE} \linelabel{ll:admixture_populus}
 Both nonspatial \texttt{conStruct} and ADMIXTURE displayed 
@@
 
-\paragraph{Comparison to ADMIXTURE}
+\paragraph{Comparison to ADMIXTURE} \linelabel{ll:admixture_bears}.
 Results from the nonspatial model and from the ADMIXTURE analyses 
@@
 
-The parameters used to simulate the data were as follows:
+The parameters used to simulate the data were as follows: \linelabel{ll:sim_params}
 a diploid population size of 1000,
@@
 
+% reviews here if toggled above
+\includereviews
+
 \end{document}

### 2018-07-02 data/bears/exe.xval.R
updated dir structure

@@ -1,38 +1,0 @@
-library(conStruct)
-library(doParallel)
-
-x.validation <- function(test.pct,n.reps,K,freqs,geoDist,coords,prefix,n.iter){
-	x.val <- mclapply(1:n.reps,
-					function(i){
-						x.validation.rep(rep.no = i,
-										 test.pct,
-										 K,
-										 freqs,
-										 geoDist,
-										 coords,
-										 prefix,
-										 n.iter)
-					},mc.cores=6,mc.allow.recursive=FALSE,mc.preschedule=FALSE)
-	names(x.val) <- paste0("rep_",1:n.reps)
-	return(x.val)
-}
-
-
-load("bear.dataset.Robj")
-
-missing.data <- apply(bear.dataset$sample.freqs,1,function(x){length(which(is.na(x)))})
-too.much.missing.data <- which(missing.data/ncol(bear.dataset$sample.freqs) > 0.04)
-
-bear.dataset <- list("sample.freqs" = bear.dataset$sample.freqs[-too.much.missing.data,],
-					 "sample.sizes" = bear.dataset$sample.sizes[-too.much.missing.data],
-					 "sample.coords" = bear.dataset$sample.coords[-too.much.missing.data,])
-
-save(bear.dataset,file="bear.dataset.Robj")
-x.validation(test.pct = 0.10,
-			 n.reps = 10,
-			 K = 1:7,
-			 freqs = bear.dataset$sample.freqs,
-			 geoDist = fields::rdist.earth(bear.dataset$sample.coords),
-			 coords = bear.dataset$sample.coords,
-			 prefix = "bear_",
-			 n.iter = 5e3)

### 2018-07-02 writeup/figs/admixture_figs.R
removed all absolute path references

@@ -73,15 +73,14 @@ get.laycols <- function(K,q.file,N){
 	return(laycols)
 }
 
+load("../../sims/sim_data/K_1/sim.dataset.Robj")
+laycols <- get.laycols(K=7,q.file="../../sims/analyses/admixture/datasets/simK1/simK1.%s.Q",N=sim.dataset$N)
 
-load("~/Dropbox/conStruct/sims/cross_validation/K_1/sim.dataset.Robj")
-laycols <- get.laycols(K=7,q.file="~/Dropbox/conStruct/sims/admixture/datasets/simK1/simK1.%s.Q",N=sim.dataset$N)
 
-
-pdf(file="~/Dropbox/conStruct/writeup/figs/admixture/admixture_simK1_pies.pdf",width=12,height=7.75,pointsize=13)
+pdf(file="admixture/admixture_simK1_pies.pdf",width=12,height=7.75,pointsize=13)
 	par(mfrow=c(2,3),mar=c(4.5,4,2.5,4))
 	for(k in 2:7){
-		pie.plot(Q.file = sprintf("~/Dropbox/conStruct/sims/admixture/datasets/simK1/simK1.%s.Q",k),
+		pie.plot(Q.file = sprintf("../../sims/analyses/admixture/datasets/simK1/simK1.%s.Q",k),
 			 coords = sim.dataset$coords,
 			 pop.vec = unlist(lapply(1:sim.dataset$N,function(n){rep(n,10)})),
 			 layer.colors = laycols[[k-1]])
@@ -95,29 +94,29 @@ dev.off()
 ad.Qs <- lapply(2:7,
 			function(k){
 				pop.vec <- unlist(lapply(1:sim.dataset$N,function(n){rep(n,10)}))
-				Q.file <- sprintf("~/Dropbox/conStruct/sims/admixture/datasets/simK1/simK1.%s.Q",k)
+				Q.file <- sprintf("../../sims/analyses/admixture/datasets/simK1/simK1.%s.Q",k)
 				Q <- read.table(Q.file,stringsAsFactors=FALSE)
 				admix.props <- collapse.ind.Q(Q,pop.vec)
 				return(admix.props)
 			})
 
 for(k in 1:6){
-	pdf(file=sprintf("~/Dropbox/conStruct/writeup/figs/admixture/simK1_%s_struct.pdf",k+1),width=10,height=4.5)
[…]

### 2018-07-02 writeup/figs/bear_figs.R

@@ -1,15 +1,15 @@
-source("~/Dropbox/conStruct/writeup/figs/fig_funcs.R")
+source("fig_funcs.R")
 
 n.reps <- 10
 K <- 7
 for(n in 1:n.reps){
-	load(sprintf("~/Dropbox/conStruct/data/bears/bear_rep%s_test.lnl.Robj",n,n))
+	load(sprintf("../../data/bears/conStruct/xvalidation/bear_rep%s_test.lnl.Robj",n,n))
 	assign(paste0("tl",n),test.lnl)
 }
 x.vals <- lapply(1:n.reps,function(n){get(sprintf("tl%s",n))})
 x.vals.std <- lapply(x.vals,conStruct:::standardize.xvals)
 xval.CIs <- conStruct:::get.xval.CIs(x.vals.std,K)
-pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bear_std_xval.pdf",width=12,height=4.25,pointsize=14)
+pdf(file="bears/bear_std_xval.pdf",width=12,height=4.25,pointsize=14)
 #	quartz(width=12,height=5)
 	par(mfrow=c(1,3),mar=c(3.75,5,4,1))
 	plot.xval.CIs(xval.CIs,K,jitter=0.1,xlim=c(0.75,7.25))
@@ -23,7 +23,7 @@ pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bear_std_xval.pdf",width=12,hei
 	mtext("Cross-validation results (Bears)",side=3,padj=-2.2,font=2,cex=1.2)
 	mtext("non-spatial conStruct",side=3,font=2)
 	mtext("Predictive accuracy",side=2,padj=-3,font=2)
-	CV.error <- get.CV.error("~/Dropbox/conStruct/data/bears/admixture/exe.admixture.Rout")
+	CV.error <- get.CV.error("../../data/bears/admixture/exe.admixture.Rout")
 	plot(CV.error,pch=19,col="orangered",
 			xlab="",ylab="",cex=2,ylim=rev(range(CV.error)))
 	best <- which.min(CV.error)
@@ -34,8 +34,8 @@ pdf(file="~/Dropbox/conStruct/writeup/figs/bears/bear_std_xval.pdf",width=12,hei
 dev.off()
 
 library(maps)
-load("~/Dropbox/conStruct/data/bears/bear.dataset.Robj")
[…]

### 2018-07-02 writeup/figs/fig_funcs.R

@@ -173,7 +173,7 @@ make.col.mats <- function(cols){
 plot.sim.xvals <- function(dir,n.reps,K,simK,y.lim){
 	#recover()
 	for(n in 1:n.reps){
-		load(sprintf("simK%s_rep%s_test.lnl.Robj",simK,n))
+		load(paste0(dir,sprintf("/simK%s_rep%s_test.lnl.Robj",simK,n)))
 		assign(paste0("tl",n),test.lnl)
 	}
 	x.vals <- lapply(1:n.reps,function(n){get(sprintf("tl%s",n))})

### 2018-07-02 writeup/figs/make.all.figs.R

@@ -1,9 +1,9 @@
 	rm(list=ls()) ; gc()
-source("~/Dropbox/conStruct/writeup/figs/sim_figs.R")
+source("sim_figs.R")
 	rm(list=ls()) ; gc()
-source("~/Dropbox/conStruct/writeup/figs/populus_figs.R")
+source("populus_figs.R")
 	rm(list=ls()) ; gc()
-source("~/Dropbox/conStruct/writeup/figs/admixture_figs.R")
+source("admixture_figs.R")
 	rm(list=ls()) ; gc()
-source("~/Dropbox/conStruct/writeup/figs/bear_figs.R")
+source("bear_figs.R")
 	rm(list=ls()) ; gc()

### 2018-07-02 writeup/figs/populus_figs.R

@@ -1,16 +1,16 @@
-source("~/Dropbox/conStruct/writeup/figs/fig_funcs.R")
+source("fig_funcs.R")
 
-setwd("~/Dropbox/conStruct/data/poplars")
+
 n.reps <- 10
 K <- 7
 for(n in 1:n.reps){
-	load(sprintf("poplar_rep%s_test.lnl.Robj",n,n))
+	load(sprintf("../../data/poplars/conStruct/xvalidation/poplar_rep%s_test.lnl.Robj",n,n))
 	assign(paste0("tl",n),test.lnl)
 }
 x.vals <- lapply(1:n.reps,function(n){get(sprintf("tl%s",n))})
 x.vals.std <- lapply(x.vals,conStruct:::standardize.xvals)
 xval.CIs <- conStruct:::get.xval.CIs(x.vals.std,K)
-pdf(file="~/Dropbox/conStruct/writeup/figs/populus/populus_std_xval.pdf",width=10,height=5,pointsize=14)
+pdf(file="populus/populus_std_xval.pdf",width=10,height=5,pointsize=14)
 #	quartz(width=10,height=5)
 	par(mfrow=c(1,2),mar=c(4,5,4,2))
 	plot.xval.CIs(xval.CIs,K,jitter=0.1,xlim=c(0.75,7.25))
@@ -24,10 +24,10 @@ dev.off()
 
 
 library(maps)
-load("~/Dropbox/conStruct/data/poplars/poplar.data.Robj")
+load("../../data/poplars/data/poplar.data.Robj")
 col.mat1 <- matrix(ifelse(poplar.data$sp.ID=="Populus trichocarpa","forestgreen","black"),byrow=TRUE,length(poplar.data$sp.ID),length(poplar.data$sp.ID))
 col.mat2 <- matrix(ifelse(poplar.data$sp.ID=="Populus trichocarpa","forestgreen","black"),byrow=FALSE,length(poplar.data$sp.ID),length(poplar.data$sp.ID))
-pdf(file="~/Dropbox/conStruct/writeup/figs/populus/populus_sampling_map.pdf",width=6,height=5,pointsize=13)
+pdf(file="populus/populus_sampling_map.pdf",width=6,height=5,pointsize=13)
 	par(mar=c(4,4,1,1))
 	map(xlim = range(poplar.data$coords[,1]) + c(-5,5), ylim = range(poplar.data$coords[,2])+c(-2,2), col="gray",lforce="e")
[…]

### 2018-07-02 writeup/figs/sim_figs.R

@@ -3,18 +3,20 @@
 #	Sim Figures for conStruct paper
 ################################################################
 ################################################################
-source("~/Dropbox/conStruct/writeup/figs/fig_funcs.R")
 
+source("fig_funcs.R")
+
+################################
 #K1
-setwd("~/Dropbox/conStruct/sims/cross_validation/K_1/x_validation")
-pdf(file="~/Dropbox/conStruct/writeup/figs/sims/simK1_std_xval.pdf",width=10,height=5,pointsize=14)
+################################
+pdf(file="sims/simK1_std_xval.pdf",width=10,height=5,pointsize=14)
 #	quartz(width=10,height=5)
 	par(mfrow=c(1,2),mar=c(4,5,4,2))
-	plot.sim.xvals(dir="~/Dropbox/conStruct/sims/cross_validation/K_1/x_validation",n.reps=10,K=7,simK=1,y.lim=c(-10,0))
+	plot.sim.xvals(dir="../../sims/analyses/conStruct/cross_validation/K_1",n.reps=10,K=7,simK=1,y.lim=c(-10,0))
 dev.off()
 
-setwd("~/Dropbox/conStruct/sims/cross_validation/K_1")
-load("sim.dataset.Robj")
+
+load("../../sims/sim_data/K_1/sim.dataset.Robj")
 freq.data <- conStruct:::process.freq.data(sim.dataset$freq.data$freqs)
 data.block <- conStruct:::make.data.block(K = 1,
 										  freq.data = freq.data,
@@ -24,7 +26,7 @@ data.block <- conStruct:::make.data.block(K = 1,
 
 output.list.sp <- vector("list",7)
 for(k in 1:7){
-	load(sprintf("~/Dropbox/conStruct/sims/cross_validation/K_1/runs/simK1_K%s_sp_conStruct.results.Robj",k))
+	load(sprintf("../../sims/analyses/conStruct/runs/K_1/simK1_K%s_sp_conStruct.results.Robj",k))
 	conStruct.results <- cluster.2.layer(conStruct.results)
 	for(j in 1:k){
 		names(conStruct.results[[1]]$MAP$layer.params[[j]])[4] <- "phi"
@@ -32,17 +34,17 @@ for(k in 1:7){
[…]

### 2018-07-02 data/bears/data/process.bear.data.R
added plink binary files

@@ -43,3 +43,97 @@ bear.dataset <- list("sample.freqs" = bear.dataset$sample.freqs[-too.much.missin
 					 "sample.coords" = bear.dataset$sample.coords[-too.much.missing.data,])
 
 save(bear.dataset,file="bear.dataset.Robj")
+
+################################
+# make STRUCTURE and ADMIXTURE input data files
+################################
+
+################################
+#	structure
+################################
+make.ind.genos <- function(x,N){
+		geno <- numeric(N)
+	if(is.na(x)){
+		geno <- rep(-9,N)
+	} else {
+		geno[sample(1:N,x)] <- 1	
+	}
+	return(geno)
+}
+
+split.pop.into.inds <- function(pop.counts,pop.nChromo){
+	ind.geno <- apply(pop.counts,2,function(x){make.ind.genos(x,pop.nChromo)})
+	return(ind.geno)
+}
+
+make.structure.data.file <- function(counts,sample.sizes){
+	geno.matrix <- matrix(NA,sum(sample.sizes),ncol(freqs))
+	index <- 0
+	for(n in 1:nrow(freqs)){
+		geno.matrix[index+c(1:sample.sizes[n]),] <- split.pop.into.inds(counts[n,,drop=FALSE],sample.sizes[n])
+		index <- index + sample.sizes[n]
+	}
+	PopData <- unlist(lapply(1:length(sample.sizes),
+					function(n){
+						lapply(1:sample.sizes[n],
+							function(i){
+								sprintf("ind_%s.%s",n,i)
+							})
+					}))
+	LocData <- unlist(lapply(1:length(sample.sizes),function(n){rep(n,sample.sizes[n])}))
+	return(cbind(PopData,LocData,geno.matrix))
+}
+
+freqs <- bear.dataset$sample.freqs
+sample.sizes <- matrix(2,nrow(freqs),ncol(freqs))
+sample.sizes[which(is.na(freqs),arr.ind=TRUE)] <- 0
+counts <- freqs*sample.sizes
+str.data <- make.structure.data.file(counts,rep(2,nrow(freqs)))
+write.table(str.data,file="bears.str",quote=FALSE,row.names=FALSE,col.names=FALSE)
+
+################################
+#	admixture
+################################
[…]

### 2018-07-02 writeup/response2reviews.tex
responses to reviews drafted

@@
 following \llname{ll:admixture_populus} and \llname{ll:admixture_bears}.
-\plr{We could do this by talking about or including figure S28?}
 }
@@
 
-\reply{We didn't follow exactly what this figure might look like, 
-but hopefully this point comes across clearly in the contrast between 
-ADMIXTURE and spatial \texttt{conStruct} results.}
+\reply{
+    This is a good suggestion;
+    we think we can make this point (and a related one) by looking at the covariance-versus distance
+    plots for the bear data.
+    The additional discussion of this point is at \revref,
+    and a new figure showing covariances colored by population is at \ref{bear_cov_colorized};
+    we decided not to insert yet another figure in the main text, however,
+    and tried to describe the pattern verbally.
+}
 
@@
  There is a typo or something missing in Eq. 8 and Eq. 10., 
- there appears an l whereas the covariance is already averaged over markers. 
+ there appears an $\ell$ whereas the covariance is already averaged over markers. 
  Do you miss a sum sign here? 
@@
 In the extreme case of fixed markers the covariance is always 0! 
-In Population Genetics one often models this with a p(1-p) 
-pref-actor of the covariance matrix OR $F_{ST}$ summary statistics. 
+In Population Genetics one often models this with a $p(1-p)$ 
+prefactor of the covariance matrix or $F_{ST}$ summary statistics. 
 Obviously averaging over quite different covariance matrices 
-violates the multivariate normal Wishard approximation with 
+violates the multivariate normal Wishart approximation with 
 degrees of freedom equal to L. 
@@
 and we have fixed the error. 
-Your larger point is correct that the amount of detectable drift is a function of the initial allele frequency,
-so some transformation or normalization of allele frequencies might be useful.
-But, we have avoided this because they make the sample allelic covariance that we model 
-sensitive to which samples are included in the dataset, 
-rather than a quantity that is invariant across choice of samples.
+Your larger point is correct that the variance of allele frequency differences
+is a function of the allele frequency,
+so some transformation or normalization of allele frequencies
+can help to standardize variances across loci.
+We've chosen not to do this for a few reasons:
+(a) normalizing by $p(1-p)$ dramatically upweights low-frequency alleles, which can be problematic
+(amplifying the \emph{noise} coming from these loci as well as the signal);
+(b) the Wishart does not hold anyhow, since loci are not independent (due to linkage)
+and have different distributions (depending on frequency and sample size);
+and so (c) we are fitting a phenomenological model rather than a mechanistic one
+(despite our hopefully compelling justifications).
+We have added discussion of this point. \revref
 }
@@
 
-\reply{RESPONSE.}
+\reply{
+    This is a great point.
+    However, we're not sure this is how it would work in practice:
+    the conStruct model does follow from a model with separate ancestral allele frequencies
+    at each locus, in each layer.
+    Our model differs in that it combines information across loci into the covariance matrix
+    before taking likelihoods,
+    while STRUCTURE takes likelihoods first;
+    however, it's not obvious how these differ:
+    as we note at \llname{ll:wishart},
+    if the data were homoskedastic Gaussian, these would be equivalent.
+    We're also not sure how well STRUCTURE could identify the effects of a few loci
+    above the background of a large number of non-differentiated loci.
+    We've reworded the comparison at \revref, and added a caveat to the discussion
+    at \llname{ll:admixture_few_loci}.
+    \plr{perhaps that last caveat is too speculative?}
+}
 
@@
 
-\reply{We have added more text comparing conStruct with Geneland and EEMS and BEDASSLE.}
+\reply{We have added more text comparing conStruct with Geneland and EEMS and BEDASSLE. \revref}
 
@@
 
-\reply{We have edited that section for clarity and brevity.}
+\reply{We have edited that section for clarity and brevity. \revref}
 
 \begin{point}{}
-    On multiple occasions you argue that the ``allelic covariance" matrix is insensitive to the choice of reference allele. 
-This is also true for most covariance matrices of the style (p1-p, p2-p), or do I miss something??
-The main advantage seems to be that you buy is that you don't have to estimate mean allele frequencies, which is MANY parameters. 
-You probably have taught quite a bit about this step ``down" from Bedasssle, so perhaps share some of your views and initial experiments.
+    On multiple occasions you argue that the ``allelic covariance" matrix is
+    insensitive to the choice of reference allele.  This is also true for most
+    covariance matrices of the style (p1-p, p2-p), or do I miss something?
+    The main advantage seems to be that you don't have to estimate mean allele
+    frequencies, which is MANY parameters.  You probably have thought quite a
+    bit about this step ``down" from Bedasssle, so perhaps share some of your
+    views and initial experiments.
 \end{point}
 
-\reply{RESPONSE.}
+\reply{
+    True, insensitivity to reference allele is not the motivation for using allelic covariance;
+    rather, the motivation comes from insensitivity to the distribution of samples.
+    We've tried to make this more clear at \revref{} and at \llname{ll:cov_config}.
+}
 
@@
 \reply{Thanks for pointing out these missing and important references.  
-We have added them to the manuscript.}
+We have added them to the manuscript. \revref}
 
@@
 While it is possible that this could introduce issues into the mixing of the MCMC, 
-in practice we haven't found that concern to be borne out.}
+in practice we haven't found that concern to be borne out.
+We chose not to discuss the bounds on the exponent because the situation is not quite so clear:
+although only those exponents will give consistent covariance functions
+in continuous, flat, Euclidean space,
+the data come from a set of discrete points in a situation where the best notion of distance
+may differ substantially from Euclidean.
+}
 
@@
 
-\reply{In practice, because those parameters can't be easily interpreted in a biological context, 
+\reply{We agree: in practice,
+because those parameters can't be easily interpreted in a biological context, 
 we treat them more as nuisance parameters and focus on inference of the admixture proportions.
-We have plans for a future release that has a more biologically motivated model of IBD (see more below).}
+We have plans for a future release that has a more biologically motivated model of IBD (see more below).
+We've made a note about this. \revref}
 
@@
 We have added a Data Availability Statement, 
-and all data analyzed and scripts used to perform the analyses have been archived in data dryad.}
+and all data analyzed and scripts used to perform the analyses have been archived in data dryad.
+\revref
+}
 
@@
 And, although the information Fig. 4 contains is a little redundant with that of Fig. 2, 
-that wouldn't have to be the case, because the $phi$ parameters for the extra layers 
+that wouldn't have to be the case, because the $\phi$ parameters for the extra layers 
 in the spatial model could be so large that even though ancestry in them was negligible, 
@@
 
-\reply{RESPONSE.}
+\reply{
+    We agree, it would be nice to have the plot that is analogous 
+    to Figures \ref{simK2_adprop_fit} and \ref{simK3_adprop_fit} in the main text, 
+    but we decided that would be too much.
+    We don't think we need a ``ground truth'' plot for figure \ref{simK1pies}, however --
+    since the true $K=1$, all true admxiture coefficients are equal to 1.
+    We've updated the caption of this figure to better point this out.
+    As for Figures \ref{simK2_adprop_fit} and \ref{simK3_adprop_fit},
+    we reassure you that there's no funny business:
+    those figures show the 99\% credible interval for each population's admixture proportions
+    from the spatial conStruct analysis with the appropriate $K$,
+    and plotted them against the values used to simulate the data.
+    We've added more detail to those captions as well.
+}
 
@@
 
-\reply{RESPONSE.}
+\reply{
+    Fair enough! We've removed the phrase ``biologically realistic''. \revref{}
+    We think these caveats are discussed sufficiently later on,
+    at \llname{ll:caveats}.
+}
 
@@
 \begin{point}{}
-    III) Suggestions (for future)
+    Suggestions (for future)
  When I run the inference scheme myself, 
@@
     which is quite the assumption for real data and almost certainly violated. 
- But In some sense, your method fits the covariances matrix - the pattern that it produces it does not matter. 
+    But in some sense, your method fits the covariance matrix 
+    - the pattern that it produces it does not matter. 
     This brings me directly to the next point:
@@
 
-\reply{This is a nice idea, and one that we will definitely look into.
+\reply{This is a nice idea, and one that we have also considered.
 We've talked about including a ``fast" conStruct option  
@@
 
-\reply{Yes, we think that's likely the best way to do it when linkage is known.
-We've added some text about that in the ``Cross-validation" section of the Appendix 
-(Lines XXX-XXX).}
+\reply{
+    We agree -- that's likely the best way to do it when linkage is known.
+    We've added some text about that in the ``Cross-validation" section of the Appendix.
+    \revref
+}
 
@@
 
-\gb{Yes, one of the next updates we plan is to replace the flexible but 
+\reply{
+    Again, agreed! One of the next updates we plan is to replace the flexible but 
     poorly motivated powered exponential with the form worked out for 
-probability of identity by state in continuous space from Barton, Depaulis, and Etheridge (2000).}
+    probability of identity by state in continuous space from Barton, Depaulis, and Etheridge (2000).
+}
 
@@
 \reply{First, thank you for the compliment (seriously - ``a golden thread"!).
-We appreciate that not all scientists are native English speakers, 
-and have tried to be as clear and specific in our writing as possible, 
+And, you've pulled out a good example of an overly convoluted sentence;
+we have edited it following your excellent suggestion \revref{}.
+In this revision, we've tried to be as clear and specific in our writing as possible, 
 both in the manuscript, and even more in the R package vignettes. 
-And, while some of the words we choose are perhaps rare in scientific writing, 
-we put great care into our choice of syntax, 
-and we think that the nuances of meaning these odd word-birds lend us merit their inclusion.
-As Mark Twain wrote, ``The difference between the \emph{almost} right word and the right word 
-is really a large matter -- 
-'tis the difference between the lightning bug and the lightning."}
+We're sure it could still be much better,
+but hope it is improved.
+% And, while some of the words we choose are perhaps rare in scientific writing, 
+% we put great care into our choice of syntax, 
+% and we think that the nuances of meaning these odd word-birds lend us merit their inclusion.
+% As Mark Twain wrote, ``The difference between the \emph{almost} right word and the right word 
+% is really a large matter -- 
+% 'tis the difference between the lightning bug and the lightning."
+}
 
@@
 
-\reply{Good suggestion - we have now indicated that the method is available as a documented R package.}
+\reply{Good suggestion - we have now indicated that the method is available as a documented R package. \revref}

### 2018-07-02 writeup/conStruct.tex
responses to reviews drafted

@@
 \emph{regardless} of their spatial location. 
-Spatial information has been incorporated into some of these methods, 
+Spatial information has been incorporated into some of these methods, \revpoint{2}{5}
 by, for example, placing spatial priors on cluster membership \cite{geneland,TESS}, 
@@
 \cite{nagylaki1978diffusion,nagylaki1988,barton2002neutral},
-but substantial work remains to be done \cite{barton2013modelling}.
+but substantial work remains to be done \cite{barton2013modelling}. \revpoint{2}{8}
 Given the generality of the circumstances that generate a pattern of isolation by distance, 
@@
 
-Existing model-based clustering approaches can only describe continuous patterns of variation using
-discrete clusters, and so tend to erroneously describe continuously distributed variation with multiple clusters that 
-show spatially autocorrelated cluster membership \cite{Frantz2009,meirmans2012}.
+Most existing model-based clustering methods are based on a discrete set of clusters,
+and so tend to partition continuous variation into spurious clusters
+with spatially autocorrelated cluster membership \cite{Frantz2009,meirmans2012}. \revpoint{2}{21}
 In analyses of empirical datasets, which often show strong isolation by distance,
@@
 is documented and available for general use as an R package at 
-\burl{https://github.com/gbradburd/conStruct}.
+\burl{https://github.com/gbradburd/conStruct}. \revpoint{2}{22}
 
@@
 recording $X_i = 1$ if the allele drawn in sample $i$ matches the random reference,
-and $X_i = 0$ otherwise.
+and $X_i = 0$ otherwise. \revpoint{2}{6}
 Then, 
@@
 \end{equation}
-This definition of covariance differs from the usual ``genetic covariance'' \cite{mcvean_genealogical_2009}
-in that (a) we do not subtract locus means,
-and so to make the statistic insensitive to choice of reference allele, 
-(b) we randomly choose a reference allele at each locus.
+This definition of covariance differs from the usual ``genetic covariance'' 
+\cite{mcvean_genealogical_2009}
+in that (a) we do not subtract locus means (to make the statistic insensitive to sample configuration),
+and (b) we randomly choose a reference allele at each locus.
+(to retain insensitivity to choice of reference allele). \revpoint{2}{7}
 As noted in \cite{EEMS}, for $i \neq j$,
@@
 their allelic covariance $\widehat{\Omega}$ 
-would be Wishart distributed with degrees of freedom equal to $L$, 
+would be Wishart distributed with degrees of freedom equal to $L$, \linelabel{ll:wishart}
 the number of loci genotyped.
@@
 while the standard assignment methods assume that the error is binomially distributed \cite{Engelhardt2012}.  
-(We make this approximation for the substantial advantages in
-computational speed.)
-The second small difference is that, in the original STRUCTURE model, 
+(We make this approximation for the substantial advantages in computational speed.)
+The second difference is that, in the original STRUCTURE model, 
 allele frequencies at each locus are independently drawn for each 
@@
 closely related to the ``$F$-model'' prior for allele frequencies of \cite{falush2003}.
-As these differences are relatively small, we can compare the fit of the different models 
+These differences in the underlying model could in principle result in different behavior,
+\revpoint{2}{4}
+but below we show that the nonspatial model indeed produces similar results to ADMIXTURE,
+and use this fact to compare the fit of the different models
 --- spatial vs. nonspatial, across different values of $K$ --- 
 by comparing their performance in a common framework.
-We also validate our claim that the results of our nonspatial method 
-are a close match to those of STRUCTURE-based approaches.
 
@@
 as well as the datasets themselves, 
-are also available at the same site.
+are also available at the same site. \revpoint{2}{11}
+
 \section*{Results}
@@
 and collapsed the layers together using random, spatially autocorrelated admixture proportions.
-See \secref{sim_details} for more details, including parameter values used. \llname{ll:sim_params_ref}
+See \secref{sim_details} for more details, including parameter values used. \linelabel{ll:sim_params_ref}
 We simulated datasets using $K=1$, 2, and 3 layers;
@@
         Results for data simulated using $K=1$,
-	showing maps of admixture proportions estimated using the nonspatial \texttt{conStruct} model for $K=2$ through 4 (top row)
+	showing maps of admixture proportions estimated using the nonspatial \texttt{conStruct} model 
+    for $K=2$ through 4 (top row)
 	and the spatial \texttt{conStruct} model for $K=2$ through 4 (bottom row).
-    As there is only a single layer in the simulation, the spatial model accurately depicts the data
-    in all cases, while the nonspatial model creates spurious clusters.
+    As there is only a single layer in the simulation, no populations should be admixed,
+    which is accurately depicted by the spatial model (second row),
+    while the nonspatial model creates spurious clusters (first row).
     }\label{simK1_pies}
@@
 perhaps due to introgression from the southern neighbors
-\textit{P. angustifolia} or \textit{fremontii} \cite{Zhou2012,geraldes_etal_2014}
-
-\paragraph{Comparison to ADMIXTURE} \linelabel{ll:admixture_populus}
+\textit{P. angustifolia} or \textit{fremontii} \cite{Zhou2012,geraldes_etal_2014}.
+
+\paragraph{Comparison to ADMIXTURE}
 Both nonspatial \texttt{conStruct} and ADMIXTURE displayed 
 the successive partitioning of space and the clines of admixture seen in the simulation results.
+\linelabel{ll:admixture_populus}
 The details of each were somewhat different 
@@
 
-\paragraph{Comparison to ADMIXTURE} \linelabel{ll:admixture_bears}.
+\paragraph{Comparison to ADMIXTURE} \linelabel{ll:admixture_bears}
 Results from the nonspatial model and from the ADMIXTURE analyses 
@@
 as would be expected if it is describing continuous spatial differentiation.
+The allelic covariances plotted against distance (see Fig \ref{bear_cov_colorized})
+provide more information on ADMIXTURE's lack of fit. \revpoint{1}{4}
+First, covariance between different sampling locations of Eastern bears
+falls off with distance,
+indicating a residual pattern best explained by isolation by distance within laters.
+Second, the covariance between bears assigned to ADMIXTURE's gold and red layers
+(the furthest Northwestern (Alaskan) bears and the Eastern bears, respectively)
+appears to be a natural extension of the decay of covariance with distance,
+falling only slightly lower values than covariances 
+between other widely separated pairs of Eastern sampling locations.
 
@@
 The proposed method combines the utility of model-based clustering algorithms 
-with a biologically realistic model of isolation by distance.
+with a model of isolation by distance. \revpoint{2}{14}
 We anticipate that \texttt{conStruct} will be useful for identifying populations 
@@
 including 50-fold cross-validation for each value of $K$, took only 6.6 minutes.
+It should also be noted that there may be situations when the
+binomial-based model underlying ADMXITURE performs better than
+our Gaussian-based model,
+e.g., when clusters differ at only a few, strongly differentiated loci
+\linelabel{ll:admixture_few_loci}.
 
@@
 the genetic covariance is calculated after subtracting the mean from the entire sample,
-which is more strongly affected by densely sampled locations.
-This led to better performance on test data.
+which is more strongly affected by densely sampled locations. \linelabel{ll:cov_config}
+Genetic covariance is also often computed after first dividing each frequency by $\sqrt{p (1-p)}$,
+where $p$ is the global allele frequency,
+with the aim of equalizing variances across loci. \revpoint{2}{3}
+Our definition does not do this, and so is less affected by low-frequency alleles.
+Both of these changes led to better performance on test data.
 However, note that allelic covariance is more affected by singleton sites
@@
 
-\paragraph{Caveats and considerations}
+\paragraph{Caveats and considerations} \linelabel{ll:caveats}
 There are a few important caveats to consider in the interpretation of \texttt{conStruct} results. 
@@
 and at $K=5$, those southernmost samples split out and become their own layer.
+Furthermore, note that in this case $\alpha_0^{(k)}$ and $\phi^{(k)}$ are confounded,
+so differences in $\phi$ between layers should not be overinterpreted. \revpoint{2}{10}
 Again, we encourage users to run analyses across multiple values of $K$, 
-and to examine the spatial covariance functions within layers when interpreting results.
+to examine the spatial covariance functions within layers when interpreting results.
 
@@
 might be describing process heterogeneity or noise 
-in a region of the genome that also has loci included in the testing partition.
+in a region of the genome that also has loci included in the testing partition. \revpoint{2}{19}
 The best practice for cross-validation is to make sure that 
@@
 		\caption{
-			Plot of \texttt{conStruct} ability to correctly estimate admixture proportions on simulated data.
-			Results are from an analysis with a spatial model using $K=2$.
+			Plot of \texttt{conStruct}'s ability to correctly estimate admixture proportions 
+            on simulated data, from an analysis with a spatial model using $K=2$.
+            The horizontal axis shows the admixture proportions used to simulate the data,
+            and the vertical axis shows the 99\% credible intervals for those proportions
+            as reported by conStruct.
 		}\label{simK2_adprop_fit}
@@
 		\caption{
-			Plot of \texttt{conStruct} ability to correctly estimate admixture proportions on simulated data.
-			Results are from an analysis with a spatial model using $K=3$.
+			Plot of \texttt{conStruct} ability to correctly estimate admixture proportions 
+            on simulated data, from an analysis with a spatial model using $K=3$.
+            The horizontal axis shows the admixture proportions used to simulate the data,
+            and the vertical axis shows the 99\% credible intervals for those proportions
+            as reported by conStruct.
 		}\label{simK3_adprop_fit}
@@
 	The data were simulated using one layer with nearest-neighbor symmetric migration between demes.
-	The model with the lowest cross-validation error (i.e., the preferred model) was $K=7$.
+    The true value was $K=1$, but
+	the model with the lowest cross-validation error (i.e., the preferred model) was $K=7$.
     }\label{admix_simK1}
@@
 	The data were simulated using two layers with nearest-neighbor symmetric migration between demes.
-	The model with the lowest cross-validation error (i.e., the preferred model) was $K=7$.
+    The true value was $K=2$, but
+	the model with the lowest cross-validation error (i.e., the preferred model) was $K=7$.
     }\label{admix_simK2}
@@
 	The data were simulated using three layers with nearest-neighbor symmetric migration between demes.
-	The model with the lowest cross-validation error (i.e., the preferred model) was $K=7$.
+    The true value was $K=3$, but
+	the model with the lowest cross-validation error (i.e., the preferred model) was $K=7$.
     }\label{admix_simK3}
@@
 
+\begin{figure}[!ht]
+	\centering
+	\ifsuppmatfigs
+        {\includegraphics[width=\textwidth]{figs/bears/bear_sp_colorized_cov.pdf}}
+	\fi
+	\caption{
+        Allele frequency covariance between each pair of sampling locations for bear data,
+        with each point colored to show which two locations it correponds to
+        (see inset for colors).
+        For instance, set of points showing covariance around 0.18
+        and above 2000 miles distant
+        have dark green centers with white-to-orange borders,
+        indicating they are comparisons between the Northwest-most (Alaskan) bears
+        and the bears on the Eastern half of the map.
+        The two curves show the decay of covariance within each layer
+        for the spatial \texttt{conStruct} model with $K=2$,
+        as in figure \ref{bear_sp_layer_covs}.
+    }\label{bear_cov_colorized}
+\end{figure}
+\ifsuppmatfigs
+	\clearpage
+\fi
+
 %%% END IF INCLUDE SUPPLEMENT

### 2018-07-03 writeup/response2reviews.tex
minor edits

@@
 As we discussed in reponse to the other reviewer,
-we intentionally chose simulations in this in-between spot,
-that are somewhat biologically realistic 
-while still retaining a clear way of obtaining the ``right'' answer.
+we intentionally chose simulations in this in-between spot;
+they are somewhat biologically realistic 
+while still retaining a clear way of obtaining the ``right'' answer 
+(with regard to admixture proportions and number of discrete populations).
 Additional sets of simulations will, in our opinion, make the paper too lengthy and complex;
@@
 We have added a Data Availability Statement, 
-and all data analyzed and scripts used to perform the analyses have been archived in data dryad.
+and all data analyzed and scripts used to perform the analyses will be archived in Data Dryad.
 \revref

### 2018-07-03 writeup/conStruct.tex
minor edits

@@
 as well as the datasets themselves, 
-are also available at the same site. \revpoint{2}{11}
+are also available at the same site, 
+and additionally have been archived at Data Dryad (doi:XXXXX)  \revpoint{2}{11}
 
@@
 (see Fig \ref{sim_setup}),
-and collapsed the layers together using random, spatially autocorrelated admixture proportions.
+with randomly distributed but spatially autocorrelated admixture proportions.
 See \secref{sim_details} for more details, including parameter values used. \linelabel{ll:sim_params_ref}
@@
 our Gaussian-based model,
-e.g., when clusters differ at only a few, strongly differentiated loci
+e.g., when clusters differ at only a few strongly differentiated loci
 \linelabel{ll:admixture_few_loci}.
@@
 but may not contradict the genetic distinctness implied by the nonspatial clustering.
-This genetic distinctness rather than shared history may be more relevant for
+This genetic distinctness -- rather than shared history -- may be more relevant for
 management decisions and conservation policy,
@@
 so differences in $\phi$ between layers should not be overinterpreted. \revpoint{2}{10}
-Again, we encourage users to run analyses across multiple values of $K$, 
-to examine the spatial covariance functions within layers when interpreting results.
+Again, we encourage users to run analyses across multiple values of $K$ 
+and to examine the spatial covariance functions within layers when interpreting results.
 

### 2018-07-03 writeup/response2reviews.tex
finalish edits

@@
 
-Our detailed responses to those comments are inset below (in bold), 
+Our detailed responses to those comments are inset below,
 but, in brief, we have included more of the results of the ADMIXTURE analyses 
@@
 \end{quote}
+
+Thanks for the thoughtful consideration.
+We agree in principle with the reviewer that more detailed tests of a method are always useful,
+but, as discussed below, decided that the paper would be better served without additional simulations.
+We hope that you agree.
 
@@
 \reply{We have added more extensive discussions of the 
-comparison between \texttt{conStruct} and ADMIXTURE for both empirical datasets,
+comparison between \texttt{conStruct} and ADMIXTURE for both empirical datasets along these lines,
 following \llname{ll:admixture_populus} and \llname{ll:admixture_bears}.
@@
 as might be expected due to, e.g., inbreeding.
-We have added some text to that effect in lines XXX-XXX.}
+We have added some text to that effect. \revref}
 
@@
 we think they'd only make the paper more confusing.
-As for the migration rates -- these are given in the text already \llname{ll:sim_params},
+As for the migration rates -- these are given in the text \llname{ll:sim_params},
 and we have added an explicit reference to this \llname{ll:sim_params_ref}.
@@
 and have not been officially released yet.
+We look forward to investigating this in the future.
 }
@@
     at \llname{ll:admixture_few_loci}.
-    \plr{perhaps that last caveat is too speculative?}
 }
@@
     but we decided that would be too much.
-    We don't think we need a ``ground truth'' plot for figure \ref{simK1pies}, however --
+    We don't think we need a ``ground truth'' plot for figure \ref{simK1_pies}, however --
     since the true $K=1$, all true admxiture coefficients are equal to 1.

### 2018-07-03 writeup/conStruct.tex
finalish edits

@@
 process of speciation is usually gradual, and so there is no set point
-in the continuous divergence of populations when they unambiguously become discrete species.
+in the continuous divergence of populations when they unambiguously become distinct species.
 The issue of identifying meaningful biological subunits 
@@
 
-From a practical standpoint, even while acknowledging continuous
-processes, we often need to identify somewhat separable populations 
-from which individuals are sampled \cite{wright1949genetical}.
+From a practical standpoint, 
+we often need to identify somewhat separable populations 
+from which individuals are sampled \cite{wright1949genetical},
+even while acknowledging continuous processes.
 Delineating populations is useful for systematics and for
@@
 including generating population phylogenies \cite{CavalliSforza1975, treemix},
-dimensionality-reduction approaches, 
-such as principal components analysis 
+dimensionality-reduction approaches such as principal components analysis 
 \cite{meirmans2009genodive,menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
@@
 for all individuals assigned to a cluster, 
-\emph{regardless} of their spatial location. 
+regardless of their spatial location. 
 Spatial information has been incorporated into some of these methods, \revpoint{2}{5}
@@
 
-\emph{Isolation by distance} (IBD) refers to a pattern of increasing genetic differentiation
+\emph{Isolation by distance} refers to a pattern of increasing genetic differentiation
 with geographic separation,
@@
 Because we randomly choose the reference allele, 
-each $X_i$ behaves marginally as a fair coin --
-in particular, $\P\{X_i = 1\} = 1/2$, so $\widehat{\Omega}_{i,i} = 1/4$ for every $i$ --
+each $X_i$ behaves marginally as a fair coin ---
+in particular, $\P\{X_i = 1\} = 1/2$, so $\widehat{\Omega}_{i,i} = 1/4$ for every $i$ ---
 all information enters through \emph{correlations}.
@@
 \end{equation}
+Here $f_{i,\ell}$ is the allele frequency in the $i^\text{th}$ sample at locus $\ell$.
 This definition of covariance differs from the usual ``genetic covariance'' 
@@
 in that (a) we do not subtract locus means (to make the statistic insensitive to sample configuration),
-and (b) we randomly choose a reference allele at each locus.
+and (b) we randomly choose a reference allele at each locus
 (to retain insensitivity to choice of reference allele). \revpoint{2}{7}
@@
 where $\pi_{i,j}$ is the genetic distance,
-i.e., mean density of sites at which random samples from $i$ and $j$ differ.
+i.e., the mean density of sites at which random samples from $i$ and $j$ differ.
 
@@
 or merely explain spurious variation introduced by noise
--- in other words, whether additional model complexity
+--- in other words, whether additional model complexity
 provides significant explanatory power.
@@
 or with information theoretic penalizations such as AIC or BIC (e.g., \cite{ADMIXTURE});
-or comparisons of the marginal likelihood, 
+comparisons of the marginal likelihood, 
 generated either via various approximations (e.g., \cite{STRUCTURE})
-or via a more rigorous and computationally intensive approach such as thermodynamic integration \cite{verity_nichols2016}
+or via thermodynamic integration \cite{verity_nichols2016};
 or inference using a Dirichlet process prior \cite{huelsenbeck2007inference}.
@@
 averaged over the posterior.
-Prediction accuracy of a particular model (e.g., value of $K$)
+Prediction accuracy of a particular value of $K$
 is then measured using this log-likelihood,
@@
 When data are simulated with $K=1$ but analyzed with $K>1$, 
-the layers are present, but contribute very little to any population.
+the additional layers contribute very little to any population.
 Even when the spatial model is run with $K=7$, 
@@
 In contrast, the nonspatial model describes geographic variation
-using gradients of admixture between more and more discrete clusters
+using gradients of admixture between increasingly many discrete clusters
 to better approximate the continuous, spatial patterns of relatedness
@@
 the other primarily west of the Rockies (blue).
-The disjointed range of the blue layer likely reflects the fact
+The disjointed range of the red layer likely reflects the fact
 that Canada was not sampled, 
@@
 which affects covariance between individuals within each location, 
-but has limited impact on covariance between locations.
+but has limited impact on covariance between locations. \revpoint{1}{5}
 
@@
 
-Comparing results of the two methods, it is clear that even at $K=3$, 
-ADMIXTURE invokes clusters to describe what seems to be, 
-and is well described by \texttt{conStruct} as, 
-a continuous spatial pattern of genetic variation. 
+Even at $K=3$, ADMIXTURE invokes clusters to describe what seems to be
+a continuous spatial pattern of genetic variation,
+which \texttt{conStruct} describes using only two spatial layers.
 The third cluster in the ADMIXTURE analysis at $K=3$ 
@@
 The allelic covariances plotted against distance (see Fig \ref{bear_cov_colorized})
-provide more information on ADMIXTURE's lack of fit. \revpoint{1}{4}
-First, covariance between different sampling locations of Eastern bears
-falls off with distance,
-indicating a residual pattern best explained by isolation by distance within laters.
+provide more information on ADMIXTURE's lack of fit: \revpoint{1}{4}
+First, covariance between Eastern bears falls off gradually rather than abruptly with distance,
+indicating a residual pattern best explained by isolation by distance within layers.
 Second, the covariance between bears assigned to ADMIXTURE's gold and red layers
-(the furthest Northwestern (Alaskan) bears and the Eastern bears, respectively)
+(the furthest Northwestern and Eastern bears, respectively)
 appears to be a natural extension of the decay of covariance with distance,
@@
 scales linearly with number of samples.
-Our speed, on datasets with large sample sizes, 
-could be improved by adopting rank-one updates 
-to the inverse of the covariance matrix 
-(e.g., \cite{woodbury1950,sherman_morrison1950})
-when updating a sample's admixture proportions, 
-which alters only a single row/column of the covariance matrix. 
-We have not implemented this yet, 
-as it would likely mean losing the ability to do efficient,  
-Hamiltonian Monte Carlo sampling of our parameters.
+% Our speed, on datasets with large sample sizes, 
+% could be improved by adopting rank-one updates 
+% to the inverse of the covariance matrix 
+% (e.g., \cite{woodbury1950,sherman_morrison1950})
+% when updating a sample's admixture proportions, 
+% which alters only a single row/column of the covariance matrix. 
+% We have not implemented this yet, 
+% as it would likely mean losing the ability to do efficient,  
+% Hamiltonian Monte Carlo sampling of our parameters.
 
@@
 our Gaussian-based model,
-e.g., when clusters differ at only a few strongly differentiated loci
-\linelabel{ll:admixture_few_loci}.
+e.g., when clusters differ at only a few strongly differentiated loci,
+although we have not investigated this possibility.
+\linelabel{ll:admixture_few_loci}
 
@@
 management decisions and conservation policy,
-which are often predicated on the identification of discrete ``management units'' (MUs) 
+which are often predicated on the identification of discrete ``management units''
 identified using genetic data \cite{Moritz1994,Waples_1998,Moritz_etal_2002}.
@@
 perhaps especially at loci underlying adaptive traits, 
-and that a \texttt{conStruct} layer may still contain multiple distinct MUs worthy of independent protections.  
+and that a \texttt{conStruct} layer may still contain multiple distinct management units worthy of independent protections.  
 For instance, although both the Alaskan and Eastern Black Bears draw most of their ancestry 
@@
 (although less than from the Western bears, as measured by average covariance).
-Alternatively, the inclusion of multiple MUs into a single \texttt{conStruct} layer 
+Alternatively, the inclusion of multiple management units into a single \texttt{conStruct} layer 
 may occur if these populations are currently 
@@
 Let $A$ and $B$ be randomly chosen alleles from samples $i$ and $j$ respectively,
-at a randomly chosen locus 
-(if the two populations are the same, then choose without replacement).
+at a randomly chosen locus.
 Then, if we let $U=2(A-1/2)$ and $V=2(B-1/2)$,
@@
 and simulated $1 \times 10^4$ independent histories
-- corresponding to the same number of independent loci - 
+--- corresponding to the same number of independent loci ---
 in each dataset, 
 with 10 diploid genotypes generated per deme at each locus.
+
 \section*{Acknowledgements}

### 2018-07-05 writeup/response2reviews.tex
one small change to response

@@
 the applications to real data we hope demonstrate the real-world utility of the method.
-We have done ``sanity check'' simulations under the model itself,
+We have done ``sanity check'' simulations under the
+multivariate-normal model itself,
 but as these are far enough from biological reality (e.g., the ``frequencies'' are not between 0 and 1),

### 2018-07-05 writeup/conStruct.tex
fixed typo

@@
 provide more information on ADMIXTURE's lack of fit: \revpoint{1}{4}
-First, covariance between Eastern bears falls off gradually rather than abruptly with distance,
+covariance between Eastern bears falls off gradually rather than abruptly with distance,
 indicating a residual pattern best explained by isolation by distance within layers.
-Second, the covariance between bears assigned to ADMIXTURE's gold and red layers
+In addition, the covariance between bears assigned to ADMIXTURE's gold and red layers
 (the furthest Northwestern and Eastern bears, respectively)
 appears to be a natural extension of the decay of covariance with distance,
-falling only slightly lower values than covariances 
+falling to only slightly lower values than covariances 
 between other widely separated pairs of Eastern sampling locations.

### 2018-07-05 writeup/conStruct.tex
redef pi

--- 2018-07-03 writeup/conStruct.tex
@@
 this can also be calculated as $\Omega_{i,j} = (1 - 2 \pi_{i,j})/4$,
-where $\pi_{i,j}$ is the genetic distance,
-i.e., the mean density of sites at which random samples from $i$ and $j$ differ.
+where $\pi_{i,j}$ is the genetic distance calculated from those $L$ sites,
+i.e., the proportion of sites at which random samples from $i$ and $j$ differ.
 
@@
 We thank Marjorie Weber, Yaniv Brandvain, William Wetzel, 
-Mariah Meek, Doc Edge, and Matthew Stephens 
+Mariah Meek, Doc Edge, Matthew Stephens, and the anonymous reviewers
 for invaluable comments on the method and manuscript, 

### 2018-07-05 writeup/conStruct.tex
added evan to acknowledgments

@@
 We thank Marjorie Weber, Yaniv Brandvain, William Wetzel, 
-Mariah Meek, Doc Edge, Matthew Stephens, and the anonymous reviewers
+Mariah Meek, Doc Edge, Evan McCartney-Melstad, Matthew Stephens, 
+and the anonymous reviewers
 for invaluable comments on the method and manuscript, 

### 2018-07-18 writeup/conStruct_geneticsformat.tex
prepping for final submission

--- 2018-07-05 writeup/conStruct.tex
@@
-% Template for PLoS
-% Version 3.4 January 2017
-%
-% % % % % % % % % % % % % % % % % % % % % %
-%
-% -- IMPORTANT NOTE
-%
-% This template contains comments intended 
-% to minimize problems and delays during our production 
-% process. Please follow the template instructions
-% whenever possible.
-%
-% % % % % % % % % % % % % % % % % % % % % % % 
-%
-% Once your paper is accepted for publication, 
-% PLEASE REMOVE ALL TRACKED CHANGES in this file 
-% and leave only the final text of your manuscript. 
-% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
-% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
-%
-%
-% There are no restrictions on package use within the LaTeX files except that 
-% no packages listed in the template may be deleted.
-%
-% Please do not include colors or graphics in the text.
-%
-% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
-%
-% % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- FIGURES AND TABLES
-%
-% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
-%
-% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
-% - Figures should be uploaded separately from your manuscript file. 
-% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
-% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
-% For figure citations, please use "Fig" instead of "Figure".
-% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
-%
-% Tables should be cell-based and may not contain:
-% - spacing/line breaks within cells to alter layout or alignment
-% - do not nest tabular environments (no tabular environments within tabular environments)
-% - no graphics or colored text (cell background color/shading OK)
-% See http://journals.plos.org/plosone/s/tables for table guidelines.
-%
-% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
-%
-% IMPORTANT
-% Below are a few tips to help format your equations and other special characters according to our specifications.
-For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
-%
-% For inline equations, please be sure to include all portions of an equation in the math environment.
-For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
-%
-% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
-%
-% Please add line breaks to long display equations when possible in order to fit size of the column. 
-%
-% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
-%
-% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
-%
-% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
-%
-% % % % % % % % % % % % % % % % % % % % % % % % 
-%
-% Please contact latex@plos.org with any questions.
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
-
-\documentclass[10pt,letterpaper]{article}
-\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
-
-%% added PLR
-\newif\ifsubmissionversion
-\submissionversionfalse
-%\submissionversiontrue
-\newif\ifincludesupplement
-\includesupplementtrue
-%\includesupplementfalse
-\newif\iffloatsatend
-\floatsatendfalse
-%\floatsatendtrue
-\newif\ifincludefigs
-\includefigstrue
-%\includefigsfalse
-\newif\ifsuppmatfigs
-%\suppmatfigsfalse
-\suppmatfigstrue
-\newif\iflinenos
-\linenostrue
-%\linenosfalse
-
-\iffloatsatend
-\usepackage[nomarkers]{endfloat}
-\fi
-%% end added PLR
-
-% amsmath and amssymb packages, useful for mathematical formulas and symbols
-\usepackage{amsmath,amssymb}
-
-% Use adjustwidth environment to exceed column width (see example table in text)
-\usepackage{changepage}
-
-% Use Unicode characters when possible
-\usepackage[utf8x]{inputenc}
-
-% textcomp package and marvosym package for additional characters
-\usepackage{textcomp,marvosym}
-
-% 
-\bibliographystyle{latex/plos2015}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-
-% Use nameref to cite supporting information files (see Supporting Information section for more info)
+%TO DO
+%	dryad
+%		split out suppmat, upload as single file?
+%		figshare
+
+\documentclass[9pt,twocolumn,twoside]{gsajnl}
+% Use the documentclass option 'lineno' to view line numbers
+
 \usepackage{nameref}
-\usepackage[hidelinks]{hyperref}
+\usepackage{hyperref}
 \usepackage{breakurl}
-
-% line numbers
-\usepackage[right]{lineno}
-\iflinenos
-\linenumbers
-\fi
-
-% ligatures disabled
-\usepackage{microtype}
-\DisableLigatures[f]{encoding = *, family = * }
-
-% color can be used to apply background shading to table cells only
-\usepackage[table]{xcolor}
-
-% array package and thick rules for tables
-\usepackage{array}
-
-% create "+" rule type for thick vertical lines
-\newcolumntype{+}{!{\vrule width 2pt}}
-
-% create \thickcline for thick horizontal lines of variable length
-\newlength\savedwidth
-\newcommand\thickcline[1]{%
-  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
-  \cline{#1}%
-  \noalign{\vskip\arrayrulewidth}%
-  \noalign{\global\arrayrulewidth\savedwidth}%
-}
-
-% \thickhline command for thick horizontal lines that span the table
-\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
-\hline
-\noalign{\global\arrayrulewidth\savedwidth}}
-
-\extrafloats{100}
-% Remove comment for double spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\ifsubmissionversion
-    \raggedright
-    \setlength{\parindent}{0.5cm}
-    \textwidth 5.25in 
-    \textheight 8.25in
-\else
-    \usepackage{fullpage}
-\fi
-
-% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
-% Captions will be left justified
-\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
-\renewcommand{\figurename}{Fig}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
-% Header and Footer with logo
-\usepackage{graphicx}
-\ifsubmissionversion
-    \usepackage{lastpage,fancyhdr}
-    \usepackage{epstopdf}
-    \pagestyle{myheadings}
-    \pagestyle{fancy}
-    \fancyhf{}
-    \setlength{\headheight}{27.023pt}
-%    \lhead{\includegraphics[width=\textwidth]{latex/PLOS-submission.eps}}
-    \rfoot{\thepage/\pageref{LastPage}}
-    \renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
-    \fancyheadoffset[L]{2.25in}
-    \fancyfootoffset[L]{2.25in}
-%    \lfoot{\sf PLOS}
-\fi
-
-%% Include all macros below
+\usepackage{marvosym}
+\usepackage{marvosym}
 
@@
 \newcommand{\bals}{\textit{balsamifera}}
-
-\newcommand{\gb}[1]{{\it\color{magenta}{(#1)}}}
-\newcommand{\plr}[1]{{\it\color{purple}{(#1)}}}
-\newcommand{\gc}[1]{{\it\color{blue}{(#1)}}}
-
 \newcommand{\R}{\mathbb{R}}
@@
 
-%% for submitting with responses to reviews
-\newif\ifsubmission
-% \submissiontrue
-\submissionfalse
-\ifsubmission
-    % % stuff for submission
-    \usepackage[nofiglist,nomarkers,figuresonly]{endfloat}
-    \usepackage{setspace}
-    % \doublespacing
-    % \renewcommand\includegraphics{\relax}
-    % \renewcommand{\includegraphics}[2][]{\relax}
-    \renewcommand{\figurename}{Fig.}
-    \newcommand{\Figure}{Fig }
-    \newcommand{\Figures}{Fig }
-\else
-    % for the nice version
-    % \usepackage[square,sort,comma,numbers]{natbib} \bibpunct{(}{)}{;}{author-year}{}{,} 
-    % \bibliographystyle{plainnat}
-    \newcommand{\Figure}{{Figure }}
-    \newcommand{\Figures}{{Figures }}
-\fi
-
-% Responses to reviews
-\input{review-response-commands}
-% set this to show line numbers and include responses to reviews or not
-\newif\ifreviewresponses
-\reviewresponsestrue  % include them
-% \reviewresponsesfalse  % don't include them
-\newcommand{\responsefile}{response2reviews}  % name of the review responses file
-
-%% END MACROS SECTION
-
-\begin{document}
-\vspace*{0.2in}
-
-% Title must be 250 characters or less.
-\begin{flushleft}
-{\Large
-\textbf\newline{Inferring continuous and discrete population genetic structure across space}
-}
-\newline
-% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
-\\
-Gideon S.\  Bradburd\textsuperscript{1,*},
-Graham M.\  Coop\textsuperscript{2,\Yinyang},
-Peter L.\  Ralph\textsuperscript{3,\Yinyang},
-\\
-\bigskip
-\textbf{1}
-Department of Integrative Biology, 
-Ecology, Evolutionary Biology, and Behavior Graduate Group,
-Michigan State University, MI 48824
-\\
-\textbf{2}
-Center for Population Biology,
-Department of Evolution and Ecology, 
-University of California, Davis, CA 95616
-\\
-\textbf{3} 
-Institute of Ecology and Evolution,
-Departments of Mathematics and Biology,
-University of Oregon, Eugene, OR 97403
-\\
-\bigskip
-\textbf{*}bradburd@msu.edu
-
-\bigskip
-
-%
-% Primary Equal Contribution Note
-\Yinyang These authors contributed equally to this work.
-
-\end{flushleft}
-% Please keep the abstract below 300 words
-\section*{Abstract}
-
+\articletype{inv} % article type
+% {inv} Investigation 
+% {gs} Genomic Selection
+% {goi} Genetics of Immunity 
+% {gos} Genetics of Sex 
+% {mp} Multiparental Populations
+
+\setlength{\parskip}{0.1em}
+
+\title{Inferring continuous and discrete population genetic structure across space}
+
+\author[$\ast$,1]{Gideon S.\  Bradburd}
+\author[$\dagger$]{Graham M.\  Coop}
+\author[$\ddagger$]{Peter L.\  Ralph}
+
+\affil[$\ast$]{Ecology, Evolutionary Biology, and Behavior Graduate Group, Department of Integrative Biology, Michigan State University, MI 48824}
+\affil[$\dagger$]{Center for Population Biology, Department of Evolution and Ecology, University of California, Davis, CA 95616}
+\affil[$\ddagger$]{Institute of Ecology and Evolution, Departments of Mathematics and Biology, University of Oregon, Eugene, OR 97403}
+
+\keywords{population genetics; isolation by distance; population structure; model-based clustering}
+
+\runningtitle{Modeling spatial population structure} % For use in the footer
+
+%% For the footnote.
+%% Give the last name of the first author if only one author;
+% \runningauthor{FirstAuthorLastname}
+%% last names of both authors if there are two authors;
+% \runningauthor{FirstAuthorLastname and SecondAuthorLastname}
+%% last name of the first author followed by et al, if more than two authors.
+\runningauthor{Bradburd \textit{et al.}}
+
+\begin{abstract}
 A classic problem in population genetics is the characterization 
@@
 and by applying it to empirical datasets of poplars and black bears in North America.
-
-\ifsubmissionversion
-\newpage
-\fi
-
-\section*{Author summary}
-One of the first steps in the analysis of genetic data, 
-and a principal mission of biology, 
-is to describe and categorize natural variation. 
-A continuous pattern of differentiation (isolation by distance), 
-where individuals found closer together in space 
-are, on average, more genetically similar than individuals sampled farther apart, 
-can confound attempts to categorize natural variation into groups.  
-This is because current statistical methods for assigning individuals to discrete clusters 
-cannot accommodate spatial patterns, 
-and so are forced to use clusters to describe what is in fact
-continuous variation. 
-As isolation by distance is common in nature, 
-this is a substantial shortcoming of existing methods.
-In this study, we introduce a new statistical method 
-for categorizing natural genetic variation - 
-one that describes variation as a combination of continuous and discrete patterns.
-We demonstrate that this method works well and can capture patterns in
-population genomic data without resorting to splitting populations
-where they can be described by continuous patterns of variation.
-
-\ifsubmissionversion
-\linenumbers
-\fi
-
-\newpage
-\section*{Introduction}
+\end{abstract}
+
+\begin{document}
+
+\maketitle
+\thispagestyle{firststyle}
+\marginmark
+\firstpagefootnote
+\correspondingauthoraffiliation{Department of Integrative Biology, Michigan State University, 288 Farm Lane, East Lansing, MI 48824.  E-mail: bradburd@msu.edu}
+\vspace{-11pt}%
+
+\section{Introduction}
+
 A fundamental quandary in the description of biological diversity is
@@
 Both discrete and continuous components
-are required to accurately describe most species' patterns of genetic relatedness.
-
+are required to accurately describe most species' patterns of genetic relatedness.\par
 From a practical standpoint, 
 we often need to identify somewhat separable populations 
-from which individuals are sampled \cite{wright1949genetical},
+from which individuals are sampled \citep{wright1949genetical},
 even while acknowledging continuous processes.
 Delineating populations is useful for systematics and for
-informing conservation priorities \cite{Moritz1994,Waples_1998,Moritz_etal_2002}.
+informing conservation priorities \citep{Moritz1994,Waples_1998,Moritz_etal_2002}.
 Furthermore, we often need to identify subsets of individuals resulting from reasonably coherent
@@
 requiring methods that can disentangle the two.
-
+%
 There have been many methods proposed to characterize population
 genetic structure,
-including generating population phylogenies \cite{CavalliSforza1975, treemix},
+including generating population phylogenies \citep{CavalliSforza1975, treemix},
 dimensionality-reduction approaches such as principal components analysis 
-\cite{meirmans2009genodive,menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
+\citep{meirmans2009genodive,menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
 and model-based clustering approaches 
-(e.g., \cite{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, 
+\citep[e.g.,][]{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, 
 FINESTRUCTURE, fastStructure, huelsenbeck2007inference, 
-Corander2003,TESS,geneland}).
+Corander2003,TESS,geneland}.
 Each of these methods perform best in particular situations,
@@
 as that produced by geographic isolation by distance 
-\cite{Wright1943, novembre_interpreting_2008, Frantz2009}.
+\citep{Wright1943, novembre_interpreting_2008, Frantz2009}.
 Here, we will focus on model-based clustering, the most widely used class of approaches for population delineation.
@@
 the problem of detecting barriers to gene flow between populations, 
-e.g.,  
-\cite{BEDASSLE,EEMS,barton2008,Ringbauer2018}.)
+\citep[e.g.,][]{BEDASSLE,EEMS,barton2008,Ringbauer2018}.)
 Existing model-based clustering methods model each individual's genotypes 
@@
 regardless of their spatial location. 
-Spatial information has been incorporated into some of these methods, \revpoint{2}{5}
-by, for example, placing spatial priors on cluster membership \cite{geneland,TESS}, 
+Spatial information has been incorporated into some of these methods,
+by, for example, placing spatial priors on cluster membership \citep{geneland,TESS}, 
 but this does not address the underlying issue that 
@@
 genetic drift to build up differentiation between distant locations
-\cite{Wright1943}. 
+\citep{Wright1943}. 
 Theoretical work,
 mostly derived from ``stepping-stone'' models 
-\cite{kimura1964stepping,sawyer1976stepping,shiga1984stepping},
+\citep{kimura1964stepping,sawyer1976stepping,shiga1984stepping},
 gives us some analytical predictions for isolation by distance
-\cite{malecot1969mathematics,Slatkin1985,epperson2003geographical}, 
+\citep{malecot1969mathematics,Slatkin1985,epperson2003geographical}, 
 and some theory has been derived for continuous space 
-\cite{nagylaki1978diffusion,nagylaki1988,barton2002neutral},
-but substantial work remains to be done \cite{barton2013modelling}. \revpoint{2}{8}
+\citep{nagylaki1978diffusion,nagylaki1988,barton2002neutral},
+but substantial work remains to be done \citep{barton2013modelling}.
 Given the generality of the circumstances that generate a pattern of isolation by distance, 
-it is unsurprising that isolation by distance is very widespread in nature \cite{meirmans2012,Sexton_etal_2014}.
+it is unsurprising that isolation by distance is very widespread in nature \citep{meirmans2012,Sexton_etal_2014}.
 
@@
 about how best to describe sets of individuals using continuous clines and discrete clusters 
-(e.g., \cite{SerrePaabo2004,rosenberg2005clines}).
+\citep[e.g.,][]{SerrePaabo2004,rosenberg2005clines}.
 
@@
 and so tend to partition continuous variation into spurious clusters
-with spatially autocorrelated cluster membership \cite{Frantz2009,meirmans2012}. \revpoint{2}{21}
+with spatially autocorrelated cluster membership \citep{Frantz2009,meirmans2012}.
 In analyses of empirical datasets, which often show strong isolation by distance,
@@
 is documented and available for general use as an R package at 
-\burl{https://github.com/gbradburd/conStruct}. \revpoint{2}{22}
-
-\section*{Methods}
+\burl{https://github.com/gbradburd/conStruct}.
+
+\section{Materials and Methods}
+\label{sec:materials:methods}
 
@@
 recording $X_i = 1$ if the allele drawn in sample $i$ matches the random reference,
-and $X_i = 0$ otherwise. \revpoint{2}{6}
+and $X_i = 0$ otherwise.
 Then, 
@@
 This definition of covariance differs from the usual ``genetic covariance'' 
-\cite{mcvean_genealogical_2009}
+\citep{mcvean_genealogical_2009}
 in that (a) we do not subtract locus means (to make the statistic insensitive to sample configuration),
 and (b) we randomly choose a reference allele at each locus
-(to retain insensitivity to choice of reference allele). \revpoint{2}{7}
+(to retain insensitivity to choice of reference allele).
 As noted in \cite{EEMS}, for $i \neq j$,
@@
 if the proportions of ancestry individuals derive from each layer 
-(the ``admixture'' proportions) 
+(the admixture proportions) 
 do so as well.
@@
 as parameterized by the $\alpha$s, 
-is a flexible and standard choice in spatial statistics \cite{Diggle1998}, 
+is a flexible and standard choice in spatial statistics \citep{Diggle1998}, 
 and is not chosen to match a particular population genetics model. 
@@
 the branch length connecting the $k$th population to the population ancestral to all modeled
-layers (see for example \cite{patterson_ancient_2012,peter_fstats}),
+layers \citep[see, for example,][]{patterson_ancient_2012,peter_fstats},
 although they cannot be directly compared because 
@@
 We then allow samples to draw their ancestry from more than one layer.
-The ``admixture'' proportion of the $i$th sample in the $k$th layer, denoted $w^{(k)}_i$,
+The admixture proportion of the $i$th sample in the $k$th layer, denoted $w^{(k)}_i$,
 gives the genome-wide proportion of alleles from sample $i$ that derive from
@@
 	\centering
-	\ifincludefigs
-		{\includegraphics[width=\textwidth]{figs/schematic/method_schematic.png}}
-	\fi	
+		{\includegraphics[width=\linewidth]{Figure_1.eps}}
 		\caption{Schematic of our method, using $K=3$ as an example.
@@
 their allelic covariance $\widehat{\Omega}$ 
-would be Wishart distributed with degrees of freedom equal to $L$, \linelabel{ll:wishart}
+would be Wishart distributed with degrees of freedom equal to $L$,
 the number of loci genotyped.
@@
 by introducing a parameter to modify the given degrees of freedom 
-and thereby informally model linkage between loci (e.g., \cite{EEMS}).
+and thereby informally model linkage between loci \citep[e.g.,][]{EEMS}.
 
@@
 \begin{equation}
+\begin{split}
 P\left( w,\alpha,\phi,\eta,\gamma \;|\; \widehat{\Omega} \right) 
-    \propto
-    P\left(\widehat{\Omega} \; | \; \Omega(w,\alpha,\phi,\eta,\gamma,D) \right)
-    P(w)P(\alpha)P(\phi)P(\eta)P(\gamma) ,
+    \propto \: &
+    P\left(\widehat{\Omega} \; | \; \Omega(w,\alpha,\phi,\eta,\gamma,D) \right) \\
+    & \times P(w)P(\alpha)P(\phi)P(\eta)P(\gamma) ,
+\end{split}
 \end{equation}
@@
 Parameters are independent between layers.
-We use Hamiltonian Monte Carlo as implemented in STAN \cite{stan, NUTS, stan_lib, rstan} 
+We use Hamiltonian Monte Carlo as implemented in STAN \citep{stan, NUTS, stan_lib, rstan} 
 to estimate the posterior distribution on the parameters.
@@
 and individuals draw a proportion of their ancestry from each cluster. 
-This model is very similar to that of STRUCTURE \cite{STRUCTURE} 
-and related models (e.g., \cite{ADMIXTURE}); 
+This model is very similar to that of STRUCTURE \citep{STRUCTURE} 
+and related models \citep[e.g.,][]{ADMIXTURE}; 
 the main difference is that our likelihood assumes that allele frequencies are normally distributed 
 around their expectations, 
-while the standard assignment methods assume that the error is binomially distributed \cite{Engelhardt2012}.  
+while the standard assignment methods assume that the error is binomially distributed \citep{Engelhardt2012}.  
 (We make this approximation for the substantial advantages in computational speed.)
@@
 allele frequencies at each locus are independently drawn for each 
-cluster \cite{STRUCTURE}, while in conStruct's nonspatial model, 
+cluster \citep{STRUCTURE}, while in conStruct's nonspatial model, 
 it is more natural to envision each cluster's allele frequency 
@@
 This makes our model more
-closely related to the ``$F$-model'' prior for allele frequencies of \cite{falush2003}.
+closely related to the ``$F$-model'' prior for allele frequencies of \citep{falush2003}.
 These differences in the underlying model could in principle result in different behavior,
-\revpoint{2}{4}
 but below we show that the nonspatial model indeed produces similar results to ADMIXTURE,
@@
 comparisons of the likelihood of the data across different values of $K$,
-with various criteria on how to choose a single value (e.g., \cite{Evanno2005}),
-or with information theoretic penalizations such as AIC or BIC (e.g., \cite{ADMIXTURE});
+with various criteria on how to choose a single value \citep[e.g.,][]{Evanno2005},
+or with information theoretic penalizations such as AIC or BIC \citep[e.g.,][]{ADMIXTURE};
 comparisons of the marginal likelihood, 
-generated either via various approximations (e.g., \cite{STRUCTURE})
-or via thermodynamic integration \cite{verity_nichols2016};
-or inference using a Dirichlet process prior \cite{huelsenbeck2007inference}.
+generated either via various approximations \citep[e.g.,][]{STRUCTURE}
+or via thermodynamic integration \citep{verity_nichols2016};
+or inference using a Dirichlet process prior \citep{huelsenbeck2007inference}.
 See \cite{verity_nichols2016} for a discussion of these approaches and comparison
@@
 
-We use cross-validation (similar in spirit to \cite{ADMIXTURE_xval}) to attack this problem.
+We use cross-validation \citep[similar in spirit to][]{ADMIXTURE_xval} to attack this problem.
 To do this,
@@
 
-\section*{Data availability}
+\subsection{Data Availability}
+
 The method \texttt{conStruct} is implemented as an R package, 
@@
 are also available at the same site, 
-and additionally have been archived at Data Dryad (doi:XXXXX)  \revpoint{2}{11}
-
-\section*{Results}
-
-\subsection*{Simulations}
-
-To test the method, we first generated data using the coalescent simulator \texttt{ms} \cite{Hudson2002}.
+and additionally have been archived at Data Dryad (doi:XXXXX).
+
+\section{Results and Discussion}
+
+\subsection{Simulations}
+
+To test the method, we first generated data using the coalescent simulator \texttt{ms} \citep{Hudson2002}.
 In each simulation,
@@
 with randomly distributed but spatially autocorrelated admixture proportions.
-See \secref{sim_details} for more details, including parameter values used. \linelabel{ll:sim_params_ref}
+See \secref{sim_details} for more details, including parameter values used. 
 We simulated datasets using $K=1$, 2, and 3 layers;
@@
 For comparison,
-we also analyzed each simulated dataset using ADMIXTURE \cite{ADMIXTURE} with $K$ between 1 and 7, 
+we also analyzed each simulated dataset using ADMIXTURE \citep{ADMIXTURE} with $K$ between 1 and 7, 
 and compared models using ADMIXTURE's cross-validation procedure with 50 folds.
@@
 
+\begin{figure}
+	\centering
+		{\includegraphics[width=\linewidth]{Figure_2.eps}}
+	\caption{
+        Results for data simulated using $K=1$,
+	showing maps of admixture proportions estimated using the nonspatial \texttt{conStruct} model 
+    for $K=2$ through 4 (top row)
+	and the spatial \texttt{conStruct} model for $K=2$ through 4 (bottom row).
+    As there is only a single layer in the simulation, no populations should be admixed,
+    which is accurately depicted by the spatial model (second row),
+    while the nonspatial model creates spurious clusters (first row).
+    }\label{simK1_pies}
+\end{figure}
+
 In contrast, the nonspatial model describes geographic variation
@@
 The ADMIXTURE results are qualitatively similar, as shown in Figs \ref{admix_simK1}-\ref{admix_simK3}.
-\linelabel{ll:admixture_sim_results}
 Each nonspatial cluster is genetically more similar within itself
@@
 and, in more detail, in Fig \ref{sim_admix_CVerrors}.
-\linelabel{ll:admixture_sim_xval}
-
-The unimportance of spurious layers can be seen in plots of layer contributions
-(Figs \ref{simK1_laycon}, \ref{simK2_laycon}-\ref{simK3_laycon}).
-In the spatial analyses, once we pass the true $K$, 
-subsequent layers add little in terms of (co)variance explained; 
-in contrast, additional clusters in the nonspatial analyses continue to contribute substantially.
-
-\begin{figure}
-	\centering
-	\ifincludefigs
-		{\includegraphics[width=\textwidth]{figs/sims/Fig2_simK1_sp_vs_nsp.pdf}}
-	\fi
-	\caption{
-        Results for data simulated using $K=1$,
-	showing maps of admixture proportions estimated using the nonspatial \texttt{conStruct} model 
-    for $K=2$ through 4 (top row)
-	and the spatial \texttt{conStruct} model for $K=2$ through 4 (bottom row).
-    As there is only a single layer in the simulation, no populations should be admixed,
-    which is accurately depicted by the spatial model (second row),
-    while the nonspatial model creates spurious clusters (first row).
-    }\label{simK1_pies}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\ifincludefigs
-		{\includegraphics[width=\textwidth]{figs/sims/sim_xvals.pdf}}
-	\fi
+
+\begin{figure*}
+	\centering
+		{\includegraphics[width=0.8\linewidth]{Figure_3.eps}}
 	\caption{
@@
     }\label{sim_xvals}
-\end{figure}
+\end{figure*}
+
+The unimportance of spurious layers can be seen in plots of layer contributions
+(Figs \ref{simK1_laycon}, \ref{simK2_laycon}-\ref{simK3_laycon}).
+In the spatial analyses, once we pass the true $K$, 
+subsequent layers add little in terms of (co)variance explained; 
+in contrast, additional clusters in the nonspatial analyses continue to contribute substantially
 
@@
 	\centering
-	\ifincludefigs
-		{\includegraphics[width=\textwidth]{figs/sims/simK1_laycon_barplots.pdf}}
-	\fi
+		{\includegraphics[width=\linewidth]{Figure_4.eps}}
 		\caption{
@@
 
-\subsection*{Empirical Applications}
+\subsection{Empirical Applications}
 
@@
 
-\subsubsection*{Poplars}
+\subsubsection{Poplars}
 
@@
 species in the genus regularly co-occur and, 
-where they do, they frequently hybridize \cite{eckenwalder1984, Cronk2005}.
+where they do, they frequently hybridize \citep{eckenwalder1984, Cronk2005}.
 
@@
 have a broad zone of overlap in the Pacific Northwest,
-where they are hypothesized to hybridize \cite{geraldes_etal_2014, suarezgonzalez_etal_2016}.
+where they are hypothesized to hybridize \citep{geraldes_etal_2014, suarezgonzalez_etal_2016}.
 Both species are sampled over a large geographic region, 
-and show spatial patterns of genetic and phenotypic variation \cite{slavov_etal_2012, mckown_etal_2013},
+and show spatial patterns of genetic and phenotypic variation \citep{slavov_etal_2012, mckown_etal_2013},
 making the system well-suited for application of our method.
@@
 with most sampling concentrated on \tri{} drainages.
-The data were generated using an Infinium 34K array designed for \tri{} \cite{geraldes_2013}, 
+The data were generated using an Infinium 34K array designed for \tri{} \citep{geraldes_2013}, 
 and showed a strong pattern of bias in allelic dropout 
@@
 Figs \ref{populus_pies} and \ref{populus_xvals}, 
-as well as Figs \ref{populus_sp_pies}-\ref{populus_laycon} in the Supplementary Materials.
-For comparison, we also ran ADMIXTURE \cite{ADMIXTURE} 
+as well as Figs \ref{populus_sp_pies}-\ref{populus_laycon} in the \secref{SuppMat}.
+For comparison, we also ran ADMIXTURE \citep{ADMIXTURE} 
 with $K = 1$ through 7, using 50-fold cross-validation to compare model performance
@@
 
-\begin{figure}
-	\centering
-	\ifincludefigs
-		{\includegraphics[width=\textwidth]{figs/populus/pop_sp_results.pdf}}
-	\fi
+\begin{figure*}
+	\centering
+		{\includegraphics[width=0.8\linewidth]{Figure_5.eps}}
 	\caption{
@@
     }\label{populus_pies}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\ifincludefigs
-		{\includegraphics[width=\textwidth]{figs/populus/populus_std_xval.pdf}}
-	\fi
-	\caption{
-	Cross-validation results for \textit{Populus} dataset 
-	comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7 
-	with 10 cross-validation replicates.
-	The first panel in each row shows all results; 
-	the second panel zooms in on the results from analyses run with $K = 2$ through 7.
-    }\label{populus_xvals}
-\end{figure}
+\end{figure*}
 
@@
 Both points are
-in agreement with \cite{keller_etal_2010}, 
-who found low diversity within the region's \bals{},
+in agreement with previous work \citep{keller_etal_2010}, 
+which found low diversity within the region's \bals{},
 probably as the result of a recent postglacial expansion.
@@
 were experimentally shown to display differences in ecologically relevant phenotypes 
-(e.g., pathogen resistance, \cite{xie2009,xie2012}).  
+\citep[e.g., pathogen resistance, ][]{xie2009,xie2012}.  
 At higher values of $K$, drainages at the southern tip of \tri{} sampling 
@@
 perhaps due to introgression from the southern neighbors
-\textit{P. angustifolia} or \textit{fremontii} \cite{Zhou2012,geraldes_etal_2014}.
+\textit{P. angustifolia} or \textit{fremontii} \citep{Zhou2012,geraldes_etal_2014}.
+
+\begin{figure}[!ht]
+	\centering
+		{\includegraphics[width=\linewidth]{Figure_6.eps}}
+	\caption{
+	Cross-validation results for \textit{Populus} dataset 
+	comparing the spatial and nonspatial \texttt{conStruct} models run with $K=1$ through 7 
+	with 10 cross-validation replicates.
+	The first panel in each row shows all results; 
+	the second panel zooms in on the results from analyses run with $K = 2$ through 7.
+    }\label{populus_xvals}
+\end{figure}
 
@@
 the successive partitioning of space and the clines of admixture seen in the simulation results.
-\linelabel{ll:admixture_populus}
 The details of each were somewhat different 
@@
 
-\subsubsection*{Black bears}
+\subsubsection{Black bears}
 
@@
 from which they subsequently expanded to occupy their current range
-\cite{WoodingWard1997,Byun1997,Stone2000,Puckett2015},
+\citep{WoodingWard1997,Byun1997,Stone2000,Puckett2015},
 likely leading to both continuous and discrete patterns of genetic structure.
@@
 
+\begin{figure*}[!t]
+	\centering
+		{\includegraphics[width=0.8\linewidth]{Figure_7.eps}}
+	\caption{
+	Maps of admixture proportions estimated for the black bear dataset 
+	using the spatial \texttt{conStruct} model (left),
+	the nonspatial \texttt{conStruct} model (center), 
+	and ADMIXTURE (right) for $K=3$.
+	Pies show mean admixture results across individuals within their diameter, 
+	and the admixture results for all individuals included within each group are 
+	shown in the plot above.
+    }\label{bear_K3}
+\end{figure*}
+
 \paragraph{Data and analyses}
@@
 and compared these models using cross-validation with 10 replicates.
-We also ran ADMIXTURE \cite{ADMIXTURE} on the same dataset, 
+We also ran ADMIXTURE \citep{ADMIXTURE} on the same dataset, 
 using $K = 1$ through 7, and comparing models using 
@@
 The results of these analyses are shown in Figs \ref{bear_K3}-\ref{bear_laycon}, 
-as well as in Figs \ref{bear_sp_pies}-\ref{bear_admix_CVerrors} in the Supplementary Materials.
-
-\begin{figure}
-	\centering
-	\ifincludefigs
-		{\includegraphics[width=\textwidth]{figs/bears/bears_sp_vs_nsp.pdf}}
-	\fi
-	\caption{
-	Maps of admixture proportions estimated for the black bear dataset 
-	using the spatial \texttt{conStruct} model (left),
-	the nonspatial \texttt{conStruct} model (center), 
-	and ADMIXTURE (right) for $K=3$.
-	Pies show mean admixture results across individuals within their diameter, 
-	and the admixture results for all individuals included within each group are 
-	shown in the plot above.
-    }\label{bear_K3}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\ifincludefigs
-		{\includegraphics[width=\textwidth]{figs/bears/bear_std_xval.pdf}}
-	\fi
-	\caption{
-	Cross-validation results for the black bear dataset,
-	comparing spatial and nonspatial \texttt{conStruct} models, 
-	as well as ADMIXTURE, all run with $K=1$ through 7, 
-	with 10 cross-validation replicates for the \texttt{conStruct} analyses 
-	and 50 data-fold subsets for the ADMIXTURE analyses. 
-	The first panel in each row shows results from spatial and nonspatial \texttt{conStruct} models; 
-	the second panel zooms in on the results from the spatial analyses run with $K = 4$ through 7, 
-	and the third panel shows the results for ADMIXTURE.
-	Note that the admixture plot shows cross-validation error 
-	(rather than predictive accuracy), and that the y-axis has therefore been flipped 
-	for ease of comparison to the \texttt{conStruct} results.
-    }\label{bear_xvals}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\ifincludefigs
-		{\includegraphics[width=\textwidth]{figs/bears/bears_laycon_barplots.pdf}}
-	\fi
-	\caption{
-	Layer/cluster contributions (i.e., how much total covariance is contributed by each layer/cluster), 
-	for all layers estimated in runs using $K = 1$ through 7 
-	for the spatial model (left),
-	and for all clusters using the nonspatial \texttt{conStruct} model (right).
-	For each value of $K$ along the x-axis, there are an equal number of contributions plotted.
-	Colors are consistent with Fig \ref{bear_K3}.
-    }\label{bear_laycon}
-\end{figure}
+as well as in Figs \ref{bear_sp_pies}-\ref{bear_admix_CVerrors} in the \secref{SuppMat}.
 
@@
 which affects covariance between individuals within each location, 
-but has limited impact on covariance between locations. \revpoint{1}{5}
-
-\paragraph{Comparison to ADMIXTURE} \linelabel{ll:admixture_bears}
+but has limited impact on covariance between locations.
+
+\paragraph{Comparison to ADMIXTURE}
 Results from the nonspatial model and from the ADMIXTURE analyses 
@@
 The allelic covariances plotted against distance (see Fig \ref{bear_cov_colorized})
-provide more information on ADMIXTURE's lack of fit: \revpoint{1}{4}
+provide more information on ADMIXTURE's lack of fit:
 covariance between Eastern bears falls off gradually rather than abruptly with distance,
@@
 
+\begin{figure*}[!ht]
+	\centering
+		{\includegraphics[width=0.8\linewidth]{Figure_8.eps}}
+	\caption{
+	Cross-validation results for the black bear dataset,
+	comparing spatial and nonspatial \texttt{conStruct} models, 
+	as well as ADMIXTURE, all run with $K=1$ through 7, 
+	with 10 cross-validation replicates for the \texttt{conStruct} analyses 
+	and 50 data-fold subsets for the ADMIXTURE analyses. 
+	The first panel in each row shows results from spatial and nonspatial \texttt{conStruct} models; 
+	the second panel zooms in on the results from the spatial analyses run with $K = 4$ through 7, 
+	and the third panel shows the results for ADMIXTURE.
+	Note that the admixture plot shows cross-validation error 
+	(rather than predictive accuracy), and that the y-axis has therefore been flipped 
+	for ease of comparison to the \texttt{conStruct} results.
+    }\label{bear_xvals}
+\end{figure*}
+
 Across all values of $K$ for which we ran \texttt{conStruct},
@@
 into their own cluster with an inferred Beringian glacial refugium
-\cite{Byun1997,Stone2000,Puckett2015}.
+\citep{Byun1997,Stone2000,Puckett2015}.
 Our model, which explicitly incorporates a spatial decay of relatedness, 
@@
 
+\begin{figure}[!ht]
+	\centering
+		{\includegraphics[width=\linewidth]{Figure_9.eps}}
+	\caption{
+	Layer/cluster contributions (i.e., how much total covariance is contributed by each layer/cluster), 
+	for all layers estimated in runs using $K = 1$ through 7 
+	for the spatial model (left),
+	and for all clusters using the nonspatial \texttt{conStruct} model (right).
+	For each value of $K$ along the x-axis, there are an equal number of contributions plotted.
+	Colors are consistent with Fig \ref{bear_K3}.
+    }\label{bear_laycon}
+\end{figure}
+
 %%%%%%%%%
-\section*{Discussion}
+\section{Discussion}
 
@@
 The proposed method combines the utility of model-based clustering algorithms 
-with a model of isolation by distance. \revpoint{2}{14}
+with a model of isolation by distance.
 We anticipate that \texttt{conStruct} will be useful for identifying populations 
@@
 scales linearly with number of samples.
-% Our speed, on datasets with large sample sizes, 
-% could be improved by adopting rank-one updates 
-% to the inverse of the covariance matrix 
-% (e.g., \cite{woodbury1950,sherman_morrison1950})
-% when updating a sample's admixture proportions, 
-% which alters only a single row/column of the covariance matrix. 
-% We have not implemented this yet, 
-% as it would likely mean losing the ability to do efficient,  
-% Hamiltonian Monte Carlo sampling of our parameters.
 
@@
 took 2.8 hours.
-On the same machine and dataset, 
-a fastSTRUCTURE \cite{fastStructure} run with $K=4$ took 90.2 hours.
-However, for a large number of samples, 
-fastSTRUCTURE might be faster.
 For almost any size dataset,
@@
 although we have not investigated this possibility.
-\linelabel{ll:admixture_few_loci}
 
@@
 In the limit of infinite data, the best value of $K$ 
-may be the number of samples included in the dataset \cite{Patterson2006}.
+may be the number of samples included in the dataset \citep{Patterson2006}.
 
@@
 should be viewed as hypotheses that should be subject to further testing
-(see \cite{Falush:16} for an in-depth discussion of these points).
+\citep[for an in-depth discussion of these points, see][]{Falush:16}.
 
@@
 which are often predicated on the identification of discrete ``management units''
-identified using genetic data \cite{Moritz1994,Waples_1998,Moritz_etal_2002}.
+identified using genetic data \citep{Moritz1994,Waples_1998,Moritz_etal_2002}.
 
@@
 the genetic covariance is calculated after subtracting the mean from the entire sample,
-which is more strongly affected by densely sampled locations. \linelabel{ll:cov_config}
+which is more strongly affected by densely sampled locations.
 Genetic covariance is also often computed after first dividing each frequency by $\sqrt{p (1-p)}$,
 where $p$ is the global allele frequency,
-with the aim of equalizing variances across loci. \revpoint{2}{3}
+with the aim of equalizing variances across loci.
 Our definition does not do this, and so is less affected by low-frequency alleles.
@@
 so it may be advisable to filter these prior to analysis
-if they are likely to contain a large percentage of errors \cite{linck_battey2017}.
-
-\paragraph{Caveats and considerations} \linelabel{ll:caveats}
+if they are likely to contain a large percentage of errors \citep{linck_battey2017}.
+
+\paragraph{Caveats and considerations}
 There are a few important caveats to consider in the interpretation of \texttt{conStruct} results. 
@@
 For example, if we were to run a \texttt{conStruct} analysis using geographic distances between 
-sampled individuals of greenish warblers \cite{Irwin2001} or \textit{Ensatina} salamanders \cite{wake_schneider1998} 
+sampled individuals of greenish warblers \citep{Irwin2001} or \textit{Ensatina} salamanders \citep{wake_schneider1998} 
 --- two canonical examples of ring species --- 
@@
 Furthermore, note that in this case $\alpha_0^{(k)}$ and $\phi^{(k)}$ are confounded,
-so differences in $\phi$ between layers should not be overinterpreted. \revpoint{2}{10}
+so differences in $\phi$ between layers should not be overinterpreted.
 Again, we encourage users to run analyses across multiple values of $K$ 
@@
 (a star population phylogeny, although the branches can have different lengths thanks to the $\phi^{(k)}$ parameter), 
-similar to the F-model of \cite{falush2003}.
+similar to the F-model of \citep{falush2003}.
 However, we could extend the existing model by implementing 
@@
 estimating a population phylogeny between them
-(e.g, \cite{treemix}).
+\citep[e.g,][]{treemix}.
 
@@
 A natural extension would be to attempt to ``geo-locate'' 
-the ancestry of samples without geographic coordinates \cite{Wasser2004}. 
+the ancestry of samples without geographic coordinates \citep{Wasser2004}. 
 We could also imagine letting samples draw ancestry from other geographic coordinates, 
-as we have done in a previous approach \cite{spacemix} to model long distance dispersal. 
+as we have done in a previous approach \citep{spacemix} to model long distance dispersal. 
 We could even allow entire layers to bud off of a particular location on another layer. 
@@
 it is becoming possible to ask whether there is genetic continuity at a point in space across time, 
-or whether populations are being replaced \cite{lazaridis_ancient_2014, Haak2015, slatkin_racimo2016, Nielsen2017, Schraiber2017}.
+or whether populations are being replaced \citep{lazaridis_ancient_2014, Haak2015, slatkin_racimo2016, Nielsen2017, Schraiber2017}.
 However, we expect allele frequencies to change through time in a population, 
@@
 
-\section*{Materials and Methods}
-
-% \section*{Appendix}
-% \pagenumbering{gobble}
-% \renewcommand{\theequation}{A\arabic{equation}}
-% \setcounter{equation}{0}
-% \renewcommand{\thetable}{A\arabic{table}}
-% \setcounter{table}{0}
-% \renewcommand{\thefigure}{A\arabic{figure}}
-% \setcounter{figure}{0}
-% \renewcommand{\thesection}{A\arabic{section}}
-% \setcounter{section}{0}
-
-\section*{Model rationale} \label{rationale}
-
-\subsection*{Drift, admixture, and space}
+\section{Appendices}
+\subsection{Model rationale: drift, admixture, and space} \label{rationale}
+
 Here we sketch a simple model of allele frequencies and their
@@
 \begin{equation}
-\Omega_{i,j} = \text{Var}(\epsilon) + \sum_{k=1}^K w^{(k)}_iw^{(k)}_j
+\begin{split}
+\Omega_{i,j} = & \text{Var}(\epsilon) + \sum_{k=1}^K w^{(k)}_iw^{(k)}_j
 \left(
   \text{Var}\left( \Delta^{(k)} \right) +
-\text{Cov}(\Delta^{(k,i)},\Delta^{(k,j)}) 	\right) +
-\delta_{i=j} \text{Var}(\Delta_i) .
+\text{Cov}(\Delta^{(k,i)},\Delta^{(k,j)}) 	\right) \\
+& + \delta_{i=j} \text{Var}(\Delta_i) .
 \label{admixed_spatial_cov}
+\end{split}
 \end{equation}
@@
 
-\section*{Allelic covariance}\label{allelic_cov}
+\subsection{Allelic covariance and inference}\label{allelic_cov}
+Here we go into further detail about both 
+the allelic covariance we model 
+and the modeling framework we use.
+
+\paragraph{Allelic covariance}
 To see why equations
@@
 
-\section*{Models, parameters, and priors}\label{model_app}
+\subsection{Models, parameters, and priors}\label{model_app}
+Here we discuss the different models implemented in this paper 
+and give the priors we place on model parameters.
 
@@
 \begin{centering}
-\ifincludefigs
-\begin{tabular}{| >{\centering\arraybackslash}m{2.1cm} | m{5.2cm} | >{\centering\arraybackslash}m{5.1cm} |}
+{\small
+\begin{tabular}{| >{\centering\arraybackslash}m{0.135\linewidth} | m{0.45\linewidth} | >{\centering\arraybackslash}m{0.34\linewidth} |}
 	\hline
@@
 		$w_i \sim \text{Dir}(\alpha_{1} ... \alpha_{K}=0.1)$  \\ \hline
-	\hline
 \end{tabular}
-\fi
+}
 \caption{
@@
 
-\section*{Cross validation}\label{Xvalidation}
-
-We employ a Monte Carlo cross-validation approach for model comparison \cite{picard1984}. 
+\subsection{Cross validation procedure}\label{Xvalidation}
+
+We employ a Monte Carlo cross-validation approach for model comparison \citep{picard1984}. 
 This procedure generates a mean predictive accuracy for each model and each value of $K$, 
@@
 might be describing process heterogeneity or noise 
-in a region of the genome that also has loci included in the testing partition. \revpoint{2}{19}
+in a region of the genome that also has loci included in the testing partition.
 The best practice for cross-validation is to make sure that 
@@
 
-\section*{Calculating layer contributions}\label{layer_contribution}
+\subsection{Calculating layer contributions}\label{layer_contribution}
 
@@
 							\left( -(\alpha^{(k)}_D D_{i,j}) ^ {\alpha^{(k)}_2}\right) 
-						\right) + 2\phi^{(k)} + \frac{1}{2}
+						\right)
+					+ 2\phi^{(k)} + \frac{1}{2}
 				\right) ,
@@
 
-\section*{Simulation details}\label{sim_details}
+\subsection{Simulation details}\label{sim_details}
 We wished to simulate data under a model that had some biological realism, 
@@
 
-\begin{figure}[!ht]
-	\centering
-	\ifincludefigs
-		{\includegraphics[width=\textwidth]{figs/sims/sim_setup.png}}
-	\fi
+\begin{figure}
+	\centering
+		{\includegraphics[width=\linewidth]{Figure_10.eps}}
 		\caption{Schematic of how we simulate datasets with continuous and discrete differentiation, using $K=2$ as an example.  
@@
 
-Using the program \texttt{ms} \cite{Hudson2002}, 
+Using the program \texttt{ms} \citep{Hudson2002}, 
 we generated discrete population structure by simulating $K$ distinct populations,
@@
 
-The parameters used to simulate the data were as follows: \linelabel{ll:sim_params}
+The parameters used to simulate the data were as follows:
 a diploid population size of 1000,
@@
 
-\section*{Acknowledgements}
+\section{Acknowledgments}
 
 We thank Marjorie Weber, Yaniv Brandvain, William Wetzel, 
-Mariah Meek, Doc Edge, Matthew Stephens, and the anonymous reviewers
+Mariah Meek, Doc Edge, Evan McCartney-Melstad, Matthew Stephens, Nick Barton, 
+and the anonymous reviewers
 for invaluable comments on the method and manuscript, 
@@
 whose votes determined the name of the method.
-\ifsubmissionversion
-\else
 This work was supported in part by 
@@
 and the National Science Foundation under award numbers NSF \#1148897 and \#1402725 to GB.
-\fi
-
-\newpage
-\clearpage
-\bibliography{reference.library.tex.bib}
-
-%%% BEGIN IF INCLUDE SUPPLEMENT
-\ifincludesupplement
-
-\clearpage
-\newpage
-
-\section*{Supplementary Information}
+
+\bibliography{reference.library.tex}
+\onecolumn
+\clearpage
+\section{Supplementary Information} \label{SuppMat}
 \renewcommand{\theequation}{S\arabic{equation}}
@@
 \setcounter{figure}{0}
+
+\vspace{3in}
+\begin{center}\Huge
+Supplementary Figures for: 
+Inferring continuous and discrete population genetic structure across space\\
+\bigskip
+\bigskip
+\large
+Gideon S.\  Bradburd\textsuperscript{1,*},
+Graham M.\  Coop\textsuperscript{2,\Yinyang},
+Peter L.\  Ralph\textsuperscript{3,\Yinyang},
+\\
+\bigskip
+\textbf{1}
+Ecology, Evolutionary Biology, and Behavior Graduate Group
+Department of Integrative Biology, 
+Michigan State University, MI 48824
+\\
+\textbf{2}
+Center for Population Biology,
+Department of Evolution and Ecology, 
+University of California, Davis, CA 95616
+\\
+\textbf{3} 
+Institute of Ecology and Evolution,
+Departments of Mathematics and Biology,
+University of Oregon, Eugene, OR 97403
+\\
+\bigskip
+\textbf{*}bradburd@msu.edu
+
+\bigskip
+
 %
-% \linespread{1}
-% \iffloatsatend
-% \processdelayedfloats
-% \fi
-%
-\ifsuppmatfigs
-	\clearpage
-\fi
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+% Primary Equal Contribution Note
+\Yinyang These authors contributed equally to this work.
+
+\clearpage
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/sims/simK1_sp_pies.pdf}}
-	\fi
 	\caption{
@@
     }\label{simK1_sp_pies}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/sims/simK2_sp_pies.pdf}}
-	\fi
 	\caption{
@@
     }\label{simK2_sp_pies}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/sims/simK3_sp_pies.pdf}}
-	\fi
 	\caption{
@@
     }\label{simK3_sp_pies}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/sims/simK2_adprop_fit.pdf}}
-	\fi
 		\caption{
@@
 		}\label{simK2_adprop_fit}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/sims/simK3_adprop_fit.pdf}}
-	\fi
 		\caption{
@@
 		}\label{simK3_adprop_fit}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/sims/simK1_nsp_pies.pdf}}
-	\fi
 	\caption{
@@
     }\label{simK1_nsp_pies}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/sims/simK2_nsp_pies.pdf}}
-	\fi
 	\caption{
@@
     }\label{simK2_nsp_pies}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/sims/simK3_nsp_pies.pdf}}
-	\fi
 	\caption{
@@
     }\label{simK3_nsp_pies}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/admixture/admixture_simK1_pies.pdf}}
-	\fi
 	\caption{
@@
     }\label{admix_simK1}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/admixture/admixture_simK2_pies.pdf}}
-	\fi
 	\caption{
@@
     }\label{admix_simK2}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/admixture/admixture_simK3_pies.pdf}}
-	\fi
 	\caption{
@@
     }\label{admix_simK3}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/sims/simK1_std_xval.pdf}}
-	\fi
 		\caption{
@@
 		}\label{simK1_xval}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/sims/simK2_std_xval.pdf}}
-	\fi
 		\caption{
@@
 		}\label{simK2_xval}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/sims/simK3_std_xval.pdf}}
-	\fi
 		\caption{
@@
 		}\label{simK3_xval}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/admixture/sims_CVerror.pdf}}
-	\fi
 	\caption{
@@
     }\label{sim_admix_CVerrors}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/sims/simK2_laycon_barplots.pdf}}
-	\fi
 		\caption{
@@
 		}\label{simK2_laycon}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/sims/simK3_laycon_barplots.pdf}}
-	\fi
 		\caption{
@@
 		}\label{simK3_laycon}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/populus/populus_sampling_map.pdf}}
-	\fi
 	\caption{
@@
     }\label{populus_map}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/populus/populus_sp_pies.pdf}}
-	\fi
 	\caption{
@@
     }\label{populus_sp_pies}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/populus/populus_nsp_pies.pdf}}
-	\fi
 	\caption{
@@
     }\label{populus_nsp_pies}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/populus/populus_sp_layer_covs.pdf}}
-	\fi
 	\caption{
@@
     }\label{populus_sp_layer_covs}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/populus/populus_nsp_layer_covs.pdf}}
-	\fi
 	\caption{
@@
     }\label{populus_nsp_layer_covs}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/populus/populus_laycon_barplots.pdf}}
-	\fi
 	\caption{
@@
     }\label{populus_laycon}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/populus/poplar_admixture_results.pdf}}
-	\fi
 	\caption{
@@
     }\label{populus_admixture}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/populus/poplars_admixture_CVerror.pdf}}
-	\fi
 	\caption{
@@
     }\label{poplar_admix_CVerrors}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/bears/bear_sp_results.pdf}}
-	\fi
 	\caption{
@@
     }\label{bear_sp_pies}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/bears/bear_nsp_results.pdf}}
-	\fi
 	\caption{
@@
     }\label{bear_nsp_pies}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/bears/bear_sp_layer_covs.pdf}}
-	\fi
 	\caption{
@@
 	 }\label{bear_sp_layer_covs}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/bears/bear_nsp_layer_covs.pdf}}
-	\fi
 	\caption{
@@
     }\label{bear_nsp_layer_covs}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/bears/bear_admixture_results.pdf}}
-	\fi
 	\caption{
@@
     }\label{bear_admixture}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
 		{\includegraphics[width=\textwidth]{figs/bears/bear_admixture_CVerror.pdf}}
-	\fi
 	\caption{
@@
     }\label{bear_admix_CVerrors}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-\begin{figure}[!ht]
-	\centering
-	\ifsuppmatfigs
+\end{figure*}
+	\clearpage
+
+\begin{figure*}[!ht]
+	\centering
         {\includegraphics[width=\textwidth]{figs/bears/bear_sp_colorized_cov.pdf}}
-	\fi
 	\caption{
@@
     }\label{bear_cov_colorized}
-\end{figure}
-\ifsuppmatfigs
-	\clearpage
-\fi
-
-%%% END IF INCLUDE SUPPLEMENT
-\fi
-\iffloatsatend
-\renewcommand{\thetable}{\arabic{table}}
-\setcounter{table}{0}
-\renewcommand{\thefigure}{\arabic{figure}}
-\setcounter{figure}{0}
-\processdelayedfloats
-\fi
-
-% reviews here if toggled above
-\includereviews
+\end{figure*}
+	\clearpage
 

### 2018-07-19 writeup/conStruct_geneticsformat.tex
formatted for Genetics

@@
 %	split out suppmat, upload as single file?
-%		figshare
+%	figshare?
 
@@
 \usepackage{marvosym}
-\usepackage{marvosym}
-
-\newcommand{\e}[1]{{\mathbb E}\left[ #1 \right]}
-\newcommand{\given}{\mid}
+
 \newcommand{\secref}[1]{``\nameref{#1}''}
@@
 \newcommand{\bals}{\textit{balsamifera}}
-\newcommand{\R}{\mathbb{R}}
 \renewcommand{\P}{\mathbb{P}}
 \newcommand{\E}{\mathbb{E}}
-\newcommand{\var}{\mathop{\mbox{Var}}}
 \newcommand{\cov}{\mathop{\mbox{cov}}}
-\newcommand{\supp}{\mathop{\mbox{supp}}}
-\newcommand{\sgn}{\mathop{\mbox{sgn}}}
-\newcommand{\bone}{\mathbf{1}}
 
@@
 \author[$\ast$,1]{Gideon S.\  Bradburd}
-\author[$\dagger$]{Graham M.\  Coop}
-\author[$\ddagger$]{Peter L.\  Ralph}
+\author[$\dagger,2$]{Graham M.\  Coop}
+\author[$\ddagger,2$]{Peter L.\  Ralph}
 
@@
 \correspondingauthoraffiliation{Department of Integrative Biology, Michigan State University, 288 Farm Lane, East Lansing, MI 48824.  E-mail: bradburd@msu.edu}
+\footnotetext[2]{These authors contributed equally to this work.}
 \vspace{-11pt}%
@@
 requiring methods that can disentangle the two.
-%
+
 There have been many methods proposed to characterize population
@@
 are also available at the same site, 
-and additionally have been archived at Data Dryad (doi:XXXXX).
+and additionally have been archived at Data Dryad (doi:10.5061/dryad.5qj7h09).
 
@@
 show that subsequent layers beyond the number used for simulation are unused.
-When data are simulated with $K=1$ but analyzed with $K>1$, 
+When data simulated with $K=1$ are analyzed with $K>1$, 
 the additional layers contribute very little to any population.
@@
 to better approximate the continuous, spatial patterns of relatedness
-(depicted in Figs \ref{simK1_pies} and \ref{simK1_nsp_pies}-\ref{simK3_nsp_pies}).
+(Figs \ref{simK1_pies} and \ref{simK1_nsp_pies}-\ref{simK3_nsp_pies}).
 The ADMIXTURE results are qualitatively similar, as shown in Figs \ref{admix_simK1}-\ref{admix_simK3}.
@@
 it is becoming possible to ask whether there is genetic continuity at a point in space across time, 
-or whether populations are being replaced \citep{lazaridis_ancient_2014, Haak2015, slatkin_racimo2016, Nielsen2017, Schraiber2017}.
+or whether populations are being replaced \citep{lazaridis_ancient_2014, Haak2015, slatkin_racimo2016, Nielsen2017, Schraiber2017,dystruct}.
 However, we expect allele frequencies to change through time in a population, 
@@
 \clearpage
-\section{Supplementary Information} \label{SuppMat}
+\section{Supplementary Materials} \label{SuppMat}
 \renewcommand{\theequation}{S\arabic{equation}}
@@
 \Yinyang These authors contributed equally to this work.
-
-\clearpage
-\begin{figure*}[!ht]
+\end{center}
+
+\clearpage
+\begin{figure}[!ht]
 	\centering
@@
     }\label{simK1_sp_pies}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{simK2_sp_pies}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{simK3_sp_pies}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
 		}\label{simK2_adprop_fit}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
 		}\label{simK3_adprop_fit}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{simK1_nsp_pies}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{simK2_nsp_pies}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{simK3_nsp_pies}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{admix_simK1}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{admix_simK2}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{admix_simK3}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
 		}\label{simK1_xval}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
 		}\label{simK2_xval}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
 		}\label{simK3_xval}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{sim_admix_CVerrors}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
 		}\label{simK2_laycon}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
 		}\label{simK3_laycon}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{populus_map}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{populus_sp_pies}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{populus_nsp_pies}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{populus_sp_layer_covs}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{populus_nsp_layer_covs}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{populus_laycon}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{populus_admixture}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{poplar_admix_CVerrors}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{bear_sp_pies}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{bear_nsp_pies}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
 	 }\label{bear_sp_layer_covs}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{bear_nsp_layer_covs}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{bear_admixture}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{bear_admix_CVerrors}
-\end{figure*}
-	\clearpage
-
-\begin{figure*}[!ht]
+\end{figure}
+	\clearpage
+
+\begin{figure}[!ht]
 	\centering
@@
     }\label{bear_cov_colorized}
-\end{figure*}
+\end{figure}
 	\clearpage

### 2018-08-21 R/model.comparison.R
added model comparison vignette

@@ -204,7 +204,7 @@ match.layers.x.runs <- function(admix.mat1,admix.mat2,admix.mat1.order=NULL){
 #' 			A layer can have a large contribution if many samples draw 
 #'			large amounts of admixture from it, or if it has a very large 
 #'			within-layer covariance parameter (phi), or some combination 
-#'			of the two. layer contribution can be useful for evaluating 
+#'			of the two. Layer contribution can be useful for evaluating 
 #'			an appropriate level of model complexity for the data (e.g., 
 #'			choosing a value of \code{K} or comparing the spatial and 
 #'			nonspatial models).
@@ -230,6 +230,7 @@ make.data.partitions <- function(n.reps,freqs,train.prop){
 							function(i){
 								xval.process.data(freqs,train.prop)								
 							})
+	names(data.partitions) <- paste0("rep",1:n.reps)
 	return(data.partitions)
 }
 

### 2018-08-21 R/conStruct-package.R
updated package description

@@ -1,6 +1,10 @@
 #' The 'conStruct' package.
 #' 
-#' @description A DESCRIPTION OF THE PACKAGE
+#' @description A method for modeling genetic data as a combination of discrete
+#'    layers, within each of which relatedness may decay continuously with geographic
+#'    distance. This package contains code for running analyses (which are implemented
+#'    in the modeling language rstan) and visualizing and interpreting output. See the
+#'    paper (doi: 10.1534/genetics.118.301333) for more details on the model and its utility.
 #' 
 #' @docType package
 #' @name conStruct-package
@@ -12,6 +16,6 @@
 #' @importFrom rstan sampling
 #' 
 #' @references 
+#' G.S. Brdburd, G.M. Coop, and P.L. Ralph (2018). Inferring continuous and discrete population genetic structure across space. Genetics. doi: 10.1534/genetics.118.301333
 #' Stan Development Team (2018). RStan: the R interface to Stan. R package version 2.17.3. http://mc-stan.org
-#' 
 NULL

### 2018-08-28 R/model.comparison.R
end parallelized workers in xval

@@ -75,6 +75,7 @@ x.validation <- function(train.prop = 0.9, n.reps, K, freqs = NULL, data.partiti
 	}
 	check.data.partitions.arg(args <- as.list(environment()))
 	save(data.partitions,file=paste0(prefix, ".xval.data.partitions.Robj"))
+	prespecified <- parallel.prespecify.check(args <- as.list(environment()))
 	`%d%` <- parallelizing(args <- as.list(environment()))
 	i <- 1
     x.val <- foreach::foreach(i=1:n.reps) %d% {
@@ -92,6 +93,7 @@ x.validation <- function(train.prop = 0.9, n.reps, K, freqs = NULL, data.partiti
     x.val <- lapply(x.val, standardize.xvals)
     save(x.val,file=paste0(prefix,".xval.results.Robj"))
 	write.xvals(x.val,prefix)
+	tmp <- end.parallelization(prespecified)
     return(x.val)
 }
 
@@ -491,6 +493,22 @@ write.xvals <- function(xvals,prefix){
 		nsp <- round(nsp,digits=4)
 	utils::write.table(sp,file=paste0(prefix,"_sp_xval_results.txt"),row.names=FALSE,quote=FALSE)
 	utils::write.table(nsp,file=paste0(prefix,"_nsp_xval_results.txt"),row.names=FALSE,quote=FALSE)
+}
+
+parallel.prespecify.check <- function(args){
+	prespecified <- FALSE
+	if(args[["parallel"]] & foreach::getDoParRegistered()){
+		prespecified <- TRUE
+	}
+	return(prespecified)
+}
+
+end.parallelization <- function(prespecified){
+	if(!prespecified){
+		doParallel::stopImplicitCluster()
+		message("\nParallel workers terminated\n\n")
+	}
+	return(invisible("if not prespecified, parallelization ended"))
 }
 
 parallelizing <- function(args){

### 2018-08-28 R/data.R
added example output data objs

@@ -17,3 +17,96 @@
 #' }
 #
 "conStruct.data"
+
+#' Example \code{data.block} generated by a \code{conStruct} analysis
+#' 
+#' An example \code{data.block} object generated in a \code{conStruct} 
+#' analysis from the raw data supplied by the user. This object is 
+#' automatically saved and is used in several subsequent plotting functions.
+#' 
+#' @format A list with 7 elements:
+#' \describe{
+#'		\item{\code{N}}{the number of samples included in the analysis}
+#'		\item{\code{K}}{the number of clusters/layers included in the model}
+#'		\item{\code{spatial}}{a boolean indicating whether the spatial
+#'			model has been specified}
+#'		\item{\code{L}}{the number of loci included in the analysis}
+#'		\item{\code{coords}}{a matrix with one row for each of the \code{N} samples, 
+#'			in the same order as that of the \code{obsCov} matrix, 
+#'			and two columns, the first giving the x-coordinate 
+#'			(or longitude), the second giving the y-coordinate (or latitude)}
+#'		\item{\code{obsCov}}{the sample allelic covariance matrix, 
+#'			in the same order as that of the \code{coords} matrix, 
+#'			with \code{N} rows and columns}
+#'		\item{\code{geoDist}}{a matrix of pairwise geographic distance between , 
+#'			samples in the same order as that of the \code{obsCov}, 
+#'			with \code{N} rows and columns, normalized by the standard 
+#'			deviation of the raw geographic distance matrix specified 
+#'			by the user}
+#'		\item{\code{varMeanFreqs}}{the variance of the mean allele frequencies, 
+#'			averaged over choice of counted allele (passed to the model 
+#'			as a prior on the global covaraince parameter)}
+#' }
+#
+"data.block"
+
+#' Example model output from a \code{conStruct} analysis
+#' 
+#' An example \code{conStruct.results} object generated by 
+#' a \code{conStruct} analysis.
+#' 
+#' @format a list with one entry for each chain run 
+#'		(specified with \code{n.chains}). The entry for each chain is named 
+#'		"chain_X" for the Xth chain.  The components of the entries for each 
+#'		are detailed below: 
+#'		\itemize{
+#'			\item \code{posterior} gives parameter estimates over the posterior 
+#'					distribution of the MCMC
+#'				\itemize{
+#'					\item \code{n.iter} number of MCMC iterations retained for 
+#'							analysis (half of the \code{n.iter} argument 
+#'							specified in the function call)
[…]

### 2018-08-28 R/plot.output.R
suppress pkg caroline warnings

@@ -225,10 +225,12 @@ make.admix.pie.plot <- function(admix.proportions,coords,layer.colors=NULL,radii
 	if(is.null(y.lim)){
 		y.lim <- c(min(coords[,2]) - 1, max(coords[,2]) + 1)
 	}
-	caroline::pies(pie.list,x0=coords[,1],y0=coords[,2],
-				color.table=color.tab,border="black",radii=radii,
-				xlab="",ylab="",main="",lty=1,density=NULL,
-				xlim = x.lim, ylim = y.lim)
+	suppressWarnings(
+		caroline::pies(pie.list,x0=coords[,1],y0=coords[,2],
+					   color.table=color.tab,border="black",radii=radii,
+					   xlab="",ylab="",main="",lty=1,density=NULL,
+					   xlim = x.lim, ylim = y.lim)
+	)
 	return(invisible(0))
 }
 

### 2018-08-31 R/plot.output.R
added compare.two.runs function

@@ -23,7 +23,7 @@
 #'  The plots made are by no means an exhaustive, and users are 
 #' 	encouraged to make further plots, or customize these plots as they 
 #'	see fit.  For each plot, one file is generated for each MCMC chain 
-#'	(specified withe the \code{n.chains} argument in the function 
+#'	(specified with the \code{n.chains} argument in the function 
 #'	\code{conStruct}. The plots generated (as .pdf files) are:
 #'	\itemize{
 #'		\item Structure plot - STRUCTURE-style plot, where each sample 
@@ -234,6 +234,101 @@ make.admix.pie.plot <- function(admix.proportions,coords,layer.colors=NULL,radii
 	return(invisible(0))
 }
 
+#' Compare two conStruct runs
+#'
+#' \code{compare.two.runs} makes figures comparing the output 
+#' 	from two conStruct analyses.
+#'
+#' This function takes the outputs from two conStruct analyses and 
+#' generates a number of plots for comparing results and 
+#' diagnosing MCMC performance.
+#'
+#' @param conStruct.results1 The list output by a 
+#'			\code{conStruct} run.
+#' @param data.block1 A \code{data.block} list saved during a 
+#'			\code{conStruct} run.
+#' @param conStruct.results2 The list output by a second
+#'			\code{conStruct} run.
+#' @param data.block2 A \code{data.block} list saved during a 
+#'			second \code{conStruct} run.
+#' @param prefix A character vector to be prepended to all figures.
+#' @param layer.colors A \code{vector} of colors to be used in 
+#'			plotting results for different layers. Users must 
+#'			specify one color per layer.  If \code{NULL}, plots 
+#'			will use a pre-specified vector of colors.
+#' @return This function has only invisible return values.
+#'
+#'	@details This function produces a variety of plots that can be 
+#'	useful for comparing results from two \code{conStruct} analyses.
+#'  The runs must have the same number of independent MCMC chains, 
+#' 	but may have different values of \code{K}. The spatial and 
+#'	nonspatial models can be compared. If the runs were executed 
+#'	with different values of \code{K}, the run with the smaller 
[…]

### 2018-09-12 R/data.R
slimmed down for cran

@@ -50,64 +50,3 @@
 #' }
 #
 "data.block"
-
-#' Example model output from a \code{conStruct} analysis
-#' 
-#' An example \code{conStruct.results} object generated by 
-#' a \code{conStruct} analysis.
-#' 
-#' @format a list with one entry for each chain run 
-#'		(specified with \code{n.chains}). The entry for each chain is named 
-#'		"chain_X" for the Xth chain.  The components of the entries for each 
-#'		are detailed below: 
-#'		\itemize{
-#'			\item \code{posterior} gives parameter estimates over the posterior 
-#'					distribution of the MCMC
-#'				\itemize{
-#'					\item \code{n.iter} number of MCMC iterations retained for 
-#'							analysis (half of the \code{n.iter} argument 
-#'							specified in the function call)
-#'					\item \code{lpd} vector of log posterior density over the retained 
-#'							MCMC iterations.
-#'					\item \code{nuggets} matrix of estimated nugget parameters with 
-#'							one row per MCMC iteration and one column per sample.
-#'					\item \code{par.cov} array of estimated parametric covariance matrices, 
-#'							for which the first dimension is the number of MCMC iterations.
-#'					\item \code{gamma} vector of estimated gamma parameter.
-#'					\item \code{layer.params} list summarizing estimates of layer-specific 
-#'							parameters. There is one entry for each layer specified, and the 
-#'							entry for the kth layer is named "Layer_k".
-#'						\itemize{
-#'							\item \code{alpha0} vector of estimated alpha0 parameter in the 
-#'									kth layer.
-#'							\item \code{alphaD} vector of estimated alphaD parameter in the 
-#'									kth layer.
-#'							\item \code{alpha2} vector of estimated alpha2 parameter in the 
-#'									kth layer.
-#'							\item \code{mu} vector of estimated mu parameter in the 
-#'									kth layer.
-#'							\item \code{layer.cov} vector of estimated layer-specific 
-#'									covariance parameter in the kth layer.
-#'						}
-#'					\item \code{admix.proportions} array of estimated admixture proportions.
-#'							The first dimension is the number of MCMC iterations, 
-#'							the second is the number of samples, 
-#'								and the third is the number of layers.
-#'				}
-#'		\item \code{MAP} gives point estimates of the parameters listed in the \code{posterior}
-#'							list described above. Values are indexed at the MCMC iteration 
-#'							with the greatest posterior probability.
-#'				\itemize{
[…]

### 2018-10-03 R/format.data.R
more flexible structure2conStruct

@@ -5,33 +5,49 @@
 #' 
 #' This function takes a population genetics dataset in 
 #' STRUCTURE format and converts it to conStruct format. 
-#' The STRUCTURE file must have one row per individual 
-#' and two columns per locus, and can only contain 
-#' bi-allelic SNPs.
+#' The STRUCTURE file can have one row per individual 
+#' and two columns per locus, or one column and two rows 
+#' per individual. It can only contain bi-allelic SNPs.
+#' Missing data is acceptable, but must be indicated with 
+#' a single value throughout the dataset.
 #' 
 #' @param infile The name and path of the file in STRUCTURE format 
 #' 			to be converted to \code{conStruct} format. 
+#' @param onerowperind Indicates whether the file format has 
+#'		one row per individual (\code{TRUE}) or two rows per 
+#'		individual (\code{FALSE}).
 #' @param start.loci The index of the first column in the dataset 
 #'			that contains genotype data.
 #' @param missing.datum The character or value used to denote 
-#' 			missing data in the STRUCTURE dataset (often -9).
+#' 			missing data in the STRUCTURE dataset (often 0 or -9).
 #' @param outfile The name and path of the file containing the 
 #'			\code{conStruct} formatted dataset to be generated 
 #' 			by this function.
 #'
 #' @details This function takes a STRUCTURE format data file and 
-#'		converts it to a \code{conStruct} format data file. 
-#' 		The STRUCTURE dataset should be in the ONEROWPERIND 
-#' 		file format, with one row per individual and two columns 
-#' 		per locus (this function therefore can only be applied to 
-#' 		diploid organisms). The first column of the STRUCTURE dataset 
-#' 		should be individual names. There may be any number of other 
-#' 		columns that contain non-genotype information before the first
-#'		column that contains genotype data, but there can 
-#' 		be no extraneous columns at the end of the dataset, after the 
-#' 		genotype data.  The genotype data should be bi-allelic 
-#'		single nucleotide polymorphisms (SNPs).
+#'		converts it to a \code{conStruct} format data file.
[…]

### 2019-01-02 testing/runs/test.mods.R
v1.0.3

@@ -30,7 +30,7 @@ args <- list("run1" = list("spatial" = FALSE,
 						   "prefix" = "nsp3b")
 		)
 
-cl <- parallel::makeCluster(4,type="FORK")
+cl <- parallel::makeCluster(3,type="FORK")
 doParallel::registerDoParallel(cl)
 
 tmp <- foreach::foreach(i=1:length(args)) %dopar% {

### 2019-07-25 R/format.data.R
skip lines in structure2conStruct

@@ -18,6 +18,9 @@
 #'		individual (\code{FALSE}).
 #' @param start.loci The index of the first column in the dataset 
 #'			that contains genotype data.
+#' @param start.samples The index of the first row in the dataset 
+#'			that contains genotype data (e.g., after any headers). 
+#'			Default value is 1.
 #' @param missing.datum The character or value used to denote 
 #' 			missing data in the STRUCTURE dataset (often 0 or -9).
 #' @param outfile The name and path of the file containing the 
@@ -27,10 +30,10 @@
 #' @details This function takes a STRUCTURE format data file and 
 #'		converts it to a \code{conStruct} format data file.
 #'		This function can only be applied to diploid organisms.
-#'		The STRUCTURE data file must be a plain text file, 
-#'		and the data must start on the first line of the file, 
-#'		with no column headers or extraneous text before the data 
-#'		starts (if present, these extra lines should be deleted by hand).
+#'		The STRUCTURE data file must be a plain text file. 
+#'		If there is extraneous text or column headers before the data 
+#'		starts, those extra lines should be deleted by hand or 
+#'		taken into account via the \code{start.samples} argument.
 #'		
 #' 		The STRUCTURE dataset can either be in the ONEROWPERIND=1 
 #' 		file format, with one row per individual and two columns 
@@ -55,12 +58,12 @@
 #'		future analyses.
 #'		
 #' @export
-structure2conStruct <- function(infile,onerowperind,start.loci,missing.datum,outfile){
+structure2conStruct <- function(infile,onerowperind,start.loci,start.samples=1,missing.datum,outfile){
 	outfile <- paste0(outfile,".RData")
 	if(file.exists(outfile)){
 		stop("\noutfile already exists\n\n")
 	}
-	structure.data <- utils::read.table(infile,header=FALSE,stringsAsFactors=FALSE)
+	structure.data <- utils::read.table(infile,header=FALSE,skip=start.samples-1,stringsAsFactors=FALSE)
 	sample.names <- get.sample.names(structure.data,onerowperind)
 	genos <- structure.data[,start.loci:ncol(structure.data)]
 	rm(structure.data)
@@ -87,6 +90,9 @@ get.sample.names <- function(structure.data,onerowperind){
 
 get.counted.allele <- function(genos,missing.datum){
 	alleles <- unique(genos)
+	if(identical(alleles, missing.datum)){
[…]

### 2019-07-25 R/plot.output.R
typo

@@ -20,7 +20,7 @@
 #'
 #'	@details This function produces a variety of plots that can be 
 #'	useful for visualizing results or diagnosing MCMC performance. 
-#'  The plots made are by no means an exhaustive, and users are 
+#'  The plots made are by no means exhaustive, and users are 
 #' 	encouraged to make further plots, or customize these plots as they 
 #'	see fit.  For each plot, one file is generated for each MCMC chain 
 #'	(specified with the \code{n.chains} argument in the function 
